### 2021-11-04 Rmd/00-b-abstract.Rmd
Rmd folder update

# Abstract {-}

This work is driven by the need for a conceptual framework to tackle temporal analyses in a data-centric workflow. Most research on temporal data focuses on modeling. Corresponding software requires very stringently formatted data, but does little in providing guidelines or tools for wrangling raw data into the required format. This has led to ad-hoc, and once-off solutions, which this research repairs.

There are three original contributions for the temporal data analysis in this research. They are grounded in the spirit of exploratory data analysis for time-indexed data. The first contribution (Chapter \@ref(ch:gravitas)) is a new technique for visualizing data using a calendar layout. It is most useful when the data relates to daily human activity, and patterns can be explored in relation to people's schedules. The second contribution (Chapter \@ref(ch:hakear)) is a new data abstraction which streamlines transformation, visualization, and modeling in temporal contexts. This "tsibble" object is a data infrastructure forming the foundation of temporal data pipelines. The third contribution (Chapter \@ref(ch:gracsr)) is to equip analysts with exploratory and explanatory tools for discovering missing patterns in time, and thus facilitating decisions on appropriate imputation methods for further downstream analysis.

\newpage

### 2021-11-04 Rmd/00-e-preface.Rmd
Rmd folder update

# Preface {-}

Chapter \@ref(ch:gravitas) has been accepted by the *Journal of Computational and Graphical Statistics*. It has won the ACEMS Business Analytics Prize in 2020. The accompanying R package **gravitas** is on CRAN. Chapter \@ref(ch:hakear) and Chapter \@ref(ch:gracsr) are yet to be submitted.

### Open and reproducible research {-}

This thesis is written in R Markdown [@rmarkdown] with **bookdown** [@bookdown], using **renv** [@renv] to create reproducible environments. The online version of this thesis is hosted at <https://sayani.netlify.app/>, powered by [Netlify](http://netlify.com). All materials (including the data sets and source files) required to reproduce this document can be found at the Github repository <https://github.com/Sayani07/thesis>.

### License {-}

<a rel="license" href="http://creativecommons.org/licenses/by-nc-sa/4.0/"><img alt="Creative Commons Licence" style="border-width:0" src="https://i.creativecommons.org/l/by-nc-sa/4.0/88x31.png" /></a>

This work is licensed under a [Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International License](http://creativecommons.org/licenses/by-nc-sa/4.0/).

The code used in this document is available under the [MIT license](https://opensource.org/licenses/MIT).

\clearpage\pagenumbering{arabic}\setcounter{page}{0}

### 2021-11-04 Rmd/01-intro.Rmd
Rmd folder update

# Introduction {#ch:intro}

<!-- Introduction -->
<!-- Motivation & Background -->
<!-- Outline of the thesis -->

To tame the complexity of time, breaking it into years, months, weeks, days and so on in a hierarchical manner is a common way to relate data to time. Such dicrete human made abstractions of time can be thought of as time granularities.(Aigner et al., 2008).

<!-- introducing the overall field-->
With the availability of data at more and more finer time scales, exploration of time series data may be required to be carried out across both finer and coarser scales to draw useful inferences about the underlying process. For example, data collected at an hourly scale could be analyzed using coarser temporal scales such as days, months or quarters. This approach requires deconstructing time in various possible ways. Moreover, often it might be interesting to capture calendar or periodic effects like month-of-year, day-of-week or hour-of-day. They help us in answering questions like if certain levels of those time deconstructions are characterized by unusual/routine values of the observed variable. For example, certain days of the week or months of the year are likely to be characterized by higher values. It is important to be able to navigate through all the temporal deconstructions that accommodate for periodicities to have multiple perspectives of the observed data. This idea aligns with the notion of EDA (Tukey 1977) which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.

<!-- outline the intro chapter structure -->

This chapter will provide an introduction to the study by first discussing the background and context, followed by the research aims, objectives and questions.

<!-- Background -->

The motivation for this work comes from the desire to provide methods to better understand large quantities of measurements on energy usage reported by smart meters
in household across Australia, and indeed many parts of the world. Smart meters currently
provide half-hourly use in kwh for each household, from the time that they were
installed, some as early as 2012. Households are distributed geographically, and have
different demographic properties such as the existence of solar panels, central heating or air conditioning. The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks. It is common to see aggregates of usage across households, total kwh used each half hour
by state, for example, because energy companies need to understand maximum loads
that they will have to plan ahead to accommodate. But studying overall energy use hides
the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.

<!-- introduce the specific problem  -->

<!-- ## change in data structure as a result of transitioning from linear to cyclic deconstructions -->
<!-- + Existing approaches, why we need to drill down, or why we want to -- probability distributions -->
However, restructuring time in this manner leads to restructured data where each level of the time deconstructions correspond to multiple values of the observed variable. It is common to see aggregation or summarization of these multiple observations with a unique value to study calendar effects. For example, using aggregates of usage across each hour/half-hour has been common in the literature because energy companies need to plan for maximum loads on the network. But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency. Summarizing the probability distribution of these multiple observations to capture both the shape and uncertainty could be a potential way to understand the underlying distribution of these observations. Studying probability distributions is likely to focus on features of the data which are not transparent through raw data or a unique summary statistic.

<!-- Following research objectives would facilitate the achievement of this aim -->
<!-- Research objectives -->

<!-- ## Research aims -->

Hence, the overarching research goal is to study the periodic behavior of temporal data in a structured way by studying the probability distributions by best exploiting the characteristics of time. Slicing and dicing the data in all possible temporal scales as suggested by EDA can be a daunting task as it leads to a myriad of possibilities. This inspires the research presented in this thesis, which aims to provide a platform to systematically explore periodicities in temporal data and support finding regular patterns or anomalies, explore clusters of behaviors or summarize the behavior. The first part of the work discusses computation of all possible combinations of cyclic time granularities and a graphical mapping such that distributions of a numeric response variable is displayed across combinations of two cyclic granularities. Even analyzing the distribution of the measured variable across two cyclic granularities at once could amount to displaying many plots in search of potential  patterns. Thus, the first part of the research (@Gupta2020-vo) also introduces "harmony" to denote pairs of granularities that could be analyzed together and reduces the search from all possible options. But this approach is still overwhelming for human consumption because there would still be huge number of harmonies. Hence, the second part of the research extends this work and narrows the search further by finding pair of cyclic granularities which are informative enough and rank them according to their importance. However, to explore periodic patterns of many households, we have to resort to clustering which has been addressed in the third part of the research. Although the motivation came through the smart meter example, this is a problem that is relevant to any temporal data observed more than once per year.

Temporal data analysis has assumed that the entry point to data analysis is at a model-ready data format, which provides little organization or conceptual oversight on how one should get the wild data into a tamed state. This mind-set is related to a long-held belief that exploratory data analysis is a highly ad hoc statistical area, impossible to teach or formalize. However, the **tidyverse** framework implemented in the statistical software R [@R-base], as originating in @wickham2014tidy, fundamentally overturns this thinking. Data plots and data wrangling, which the "tidy data" conceptualization supports, can be formally described using an abstract grammar. The grammar of graphics and data manipulation, as implemented in the **ggplot2** [@R-ggplot2] and **dplyr** [@R-dplyr] R packages respectively, form the core of the **tidyverse** suite of tools. My contributions extend the **tidyverse** way of thinking to the temporal domain, by providing tidy tools, built as R packages, for supporting fluent workflow in temporal data analysis.

## Visualizing probability distributions across bivariate cyclic temporal granularities {#sec:gravitas}

Deconstructing a time index into time granularities can assist in exploration and automated analysis of large temporal data sets. This paper describes classes of time deconstructions using linear and cyclic time granularities. Linear granularities respect the linear progression of time such as hours, days, weeks and months. Cyclic granularities can be circular such as hour-of-the-day, quasi-circular such as day-of-the-month, and aperiodic such as public holidays. The hierarchical structure of granularities creates a nested ordering: hour-of-the-day and second-of-the-minute are single-order-up. Hour-of-the-week is multiple-order-up, because it passes over day-of-the-week. Methods are provided for creating all possible granularities for a time index. A recommendation algorithm provides an indication whether a pair of granularities can be meaningfully examined together (a "harmony"), or when they cannot (a "clash").

 Time granularities can be used to create data visualizations to explore for periodicities, associations and anomalies. The granularities form categorical variables (ordered or unordered) which induce groupings of the observations. Assuming a numeric response variable, the resulting graphics are then displays of distributions compared across combinations of categorical variables.

 The methods implemented in the open source R package `gravitas` are consistent with a tidy workflow, with probability distributions examined using the range of graphics available in `ggplot2`.

## Detecting distributional differences between temporal granularities for exploratory time series analysis {#sec:hakear}

Cyclic temporal granularities, which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can be useful for measuring repetitive patterns in large univariate time series data. The granularities feed new approaches to exploring time series data. One use is to take pairs of granularities, and make plots of response values across the categories induced by the temporal deconstruction. However, when there are many granularities that can be constructed for a time period, there will also be too many possible displays to decide which might be the more interesting to display. This work proposes a new distance metric to screen and rank the possible granularities, and hence choose the most interesting ones to plot. The distance measure is computed for a single or pairs of cyclic granularities can can be compared across different cyclic granularities and also on a collection of time series. The methods are implemented in the open-source R package `hakear`.

## Clustering time series based on probability distributions across temporal granularities {#sec:gracsr}

 With more and more time series data being collected at much finer temporal resolution, for a longer length of time, and for a larger number of individuals/entities, time series clustering research is getting a lot of traction. The sort of noisy, patchy, uneven, and asynchronous time series that is typical in many disciplines limits similarity searches among these lengthy time series. In this work, we suggest a method for overcoming these constraints by grouping time series based on probability distributions over cyclic temporal granularities. Cyclic granularities are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, and so on, and can be helpful for detecting repeating patterns. Looking at probability distributions across cyclic granularities results in an approach that is robust to missing or noisy data, aids in dimension reduction, and ensures small pockets of similar repeated behaviours. The proposed method was tested using a collection of residential electricity customers. The simulated and empirical evidence demonstrates that our method is capable of producing meaningful clusters.

<!-- Time series clustering research is gaining a lot of importance with more and more time series data recorded at much finer temporal resolution, for a longer period of time and for a larger number of individuals/entities.Similarity searches amongst these long time series are often limited by the type of noisy, patchy, unequal and asynchronous time series common in many fields. In this paper we propose a strategy to alleviate these limitations by  clustering time series based on probability distributions across cyclic temporal granularities. Cyclic granularities are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, which can be useful for measuring repetitive patterns in large univariate time series data. Thus, looking at the probability distributions across cyclic granularity leads to a method which is robust to missing or noisy data, helps in dimension reduction while ensuring small pockets of similar repetitive behaviors. The method is applied to electricity smart meter dat -->
<!-- The suggested approach was evaluated on a set of benchmark time series on residential electricity customers. The empirical results show that our approach is able to yield meaningful clusters. -->

## Summary

The thesis is structured as follows. Chapter \@ref(ch:calendar-vis) provides details of the calendar plot, algorithm and applications. This is implemented in the R package **sugrrants**. Chapter \@ref(ch:tsibble) explains the new data abstraction--tsibble--and illustrates how it can be used to for the basis of exploratory methods, visualization and modeling of temporal data. This is available as the R package **tsibble**. Chapter \@ref(ch:mists) describes new procedures for exploring temporal missing patterns, and assisting in handling missing values prior to modeling. This is in the developing R package **mists**.

Chapter \@ref(ch:conclusion) summarizes the software tools developed for the work and their impact, and discusses some future plans.

### 2021-11-04 Rmd/03-hakear.Rmd
Rmd folder update

# Detecting distributional differences between temporal granularities for exploratory time series analysis  {#ch:hakear}

 Cyclic temporal granularities, which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can be useful for measuring repetitive patterns in large univariate time series data. The granularities feed new approaches to exploring time series data. One use is to take pairs of granularities, and make plots of response values across the categories induced by the temporal deconstruction. However, when there are many granularities that can be constructed for a time period, there will also be too many possible displays to decide which might be the more interesting to display. This work proposes a new distance metric to screen and rank the possible granularities, and hence choose the most interesting ones to plot. The distance measure is computed for a single or pairs of cyclic granularities can can be compared across different cyclic granularities and also on a collection of time series. The methods are implemented in the open-source R package `hakear`.

 ```{r external-hakear, include = FALSE, cache = FALSE}
read_chunk("scripts/hakear.R")
```

```{r load}
```

```{r}
set.seed(3215)
```

## Introduction

<!-- <introducing the problem> -->
<!-- background of the problem -->

<!--Exploratory data analysis, as coined by John W. Tukey [@Tukey1977-jx] involves many iterations of finding structures and patterns that allow the data to be informative.-->

Cyclic temporal granularities [@Bettini1998-ed, @Gupta2021-hd] are <!--theoretical--> temporal deconstructions that define cyclic repetitions in time, e.g. hour-of-day, day-of-month, or regularly scheduled public holidays. These granularities form ordered or unordered categorical variables. An example of an ordered granularity is day-of-week, where Tuesday is always followed by Wednesday, and so on. An unordered granularity example is different week types in a semester, orientation, break, exam or regular classes in an academic calendar. Using granularities to  explore patterns in univariate time series can be considered to be examining the distribution of the measured variable across different categories of the cyclic granularities.

Figure \ref{fig:onegran-new} electricity smart meter data plotted against two granularities (hour-of-day, month-of-year). The data was collected on a single household in Melbourne, Australia, over a six month period, as was used in @wang2020calendar. The categorical variable (granularity) is mapped to the x-axis, and the distribution of response variable is displayed using both side-by-side jittered dotplots and boxplots.  From plot (a) it can be seen that energy consumption is higher during the morning hours (5-8), when members in the household wake up, and again in the evening hours (17-20) possibly when members get back from work. In addition, the largest variation in energy use is in the afternoon hours (12-16), as perceived from sizes of the boxes. From plot (b) the variability in energy usage is higher in Jan and Feb, possibly due to the usage of air conditioners on some days. The median usage is highest in January, dips in February-April and rises again in May-June, although not to the height of January usage. This might imply that this household does not use as much energy for heating as it does for air conditioning. A lot of households in Victoria use gas heating and hence the heater use might not be reflected in the electricity data. Many, many different displays could be constructed using different granularities,  day-of-week, day-of-month, weekday/weekend, etc. However, only a few might be interesting, that is, reveal important patterns in energy usage. Determining which displays which have "significant" distributional differences between categories of the cyclic granularity, and plotting only these, would make for efficient exploration.

```{r calendar-elec, fig.height = 8, fig.cap = "Calendar displays faceted by two households are shown. The calendar displays are rich in information and has several components. For example, it can be observed easily that the level of energy consumption by household 2 is higher than household 4. Also, it gives an overview on when these households are away for holidays and which months consumptions are highest (Jan, Feb). However, when it comes to discerning periodic patterns, analyzing all possible periodic patterns through this display is overwhelming due to combination of linear and cyclic representation of time. For example, there is more going on in terms of peaks and troughs in id 2 compared to id 4, but it is not clear which all periodic patterns are dominant and if it differs between households."}

```

```{r onegran-new, fig.cap="A cyclic granularity can be considered to be a categorical variable, and used to break the data into subsets. Here, side-by-side boxplots overlaid on jittered dotplots explore the distribution of of energy use by a household for two different cyclic granularities: (a) hour-of-day and (b) and month-of-year. Daily peaks occur in morning and evening hours, indicating a working household, where members leave for and return from work.  More volatility of usage in summer months (Jan, Feb) is probably due air conditioner use on just some days ."}
```

<!-- comparing one and two grans for one households -->
\noindent Exploring the distribution of the measured variable across two cyclic granularities tends to provide more detailed information on its structure. For example, Figure \ref{fig:id2-new2}(a) slice down further by showing the usage distribution across hour-of-day conditional on month-of-year across two households (id 2 and 4). It shows the hourly usage over a day does not remain the same across months. Unlike other months, the 75th and 90th percentile for all hours of the day in January are    high, pretty close, and are not characterized by a morning and evening peak. The household in Figure \ref{fig:id2-new2}(b) has 90th percentile consumption higher in summer months relative to autumn or winter, but the 75th and 90th percentile are far apart in all months, implying that the second household resorts to air conditioning much less regularly than the first one. The differences seem to be more prominent across month-of-year (facets) than hour-of-day (x-axis) for this household, whereas they are prominent for both cyclic granularities for the first household.

<!-- The problem and its dimension -->
\noindent Are all of these four displays in Figures \ref{fig:onegran-new} and \ref{fig:id2-new2} useful in understanding the distributional difference in energy usage? Which ones are more useful than others? If $N_C$ is the total number of cyclic granularities of interest, the number of displays that could be potentially informative is $N_C$ when considering displays of the form in Figure \ref{fig:onegran-new}. The dimension of the problem, however, increases when considering more than one cyclic granularity. When considering displays of the form in Figure \ref{fig:id2-new2}, 
there are $^{N_C}P_2$ possible pairwise plots exhaustively, with one of the two cyclic granularities acting as the conditioning variable. 
This is huge and overwhelming for human consumption even for moderately large $N_C$. It could be immensely useful to make the transition from all potential displays to the ones that are informative across atleast one cyclic granularity.

<!-- Moreover, the volatility in consumption are almost equal and high for all hours in Jan, whereas variability is most in the evening hours for other months. The distribution of consumption across different hours of the day look similar for Apr, May and Jun, characterized by a distinct morning, afternoon and evening peaks and higher variability in evening hours -->

<!-- months Jan and Feb not only have a high median consumption, but also more variability, possibly due to the usage of air conditioning. The other months shows more consistent behavior with the energy consumption rising in May and June. -->

<!-- Scagnostics literature, other literature and why we need more/different -->

\noindent This problem is similar to Scagnostics (Scatterplot Diagnostics) by @tukey1988computer, which is used to identify meaningful patterns in large collections of scatterplots. Given a set of $v$ variables, there are $v(v-1)/2$ pairs of variables, and thus the same number of possible pairwise scatterplots. Therefore, even for small $v$, the number of scatterplots can be large, and scatterplot matrices (SPLOMs) could easily run out of pixels when presenting high-dimensional data. @Dang2014-tw and @wilkinson2005graph provide potential solutions to this, where few characterizations can be used to locate anomalies in density, shape, trend, and other features in the 2D point scatters. In this paper, we provide a solution to narrowing down the search from $^{N_C}P_2$ plots by introducing a new distance measure that can be used to detect significant distributional differences across cyclic granularities.<!-- harmonies and why it is not enough --> This work is a natural extension of our previous work [@Gupta2021-hd] that narrows down the search from $^{N_C}P_2$ plots by identifying pairs of granularities that can be meaningfully examined together (a "harmony"), or when they cannot (a "clash"). However, even after excluding clashes, the list of harmonies left could be enormous for exhaustive exploration. Hence, there is a need to reduce the search even further by including only those harmonies which are informative enough. @inference and @Majumder2013-hb present methods for statistical significance testing of visual findings using human cognition as the statistical tests. In this paper, the visual discovery of distributional differences is facilitated by choosing a threshold for the proposed numerical distance measure, eventually selecting only those cyclic granularities for which the distributional differences are sufficient to make it an interesting display.

<!-- \ref{fig:id2-new2} shows the distribution of energy consumption across hour-of-day conditional on month-of-year. \ref{fig:id2-new2}(a) shows that energy consumption is higher during the morning hours when members in the household wake up and then again higher in the evening hours possibly when members are back from work with them showing maximum variable behavior in the afternoon hours. \ref{fig:id2-new2}(b) shows the months Jan and Feb not only have a high median consumption, but also more variability, possibly due to the usage of air conditioning. The other months shows more consistent behavior with the energy consumption rising in May and June. \ref{fig:id2-new2} adds to this information, by showing that hours in Jan do not necessarily follow the same pattern across the day (due to Holidays, people visiting). Moreover, the volatility in consumption are almost equal and high for all hours in Jan, whereas variability is most in the evening hours for other months. The distribution of consumption across different hours of the day look similar for Apr, May and Jun, characterized by a distinct morning, afternoon and evening peaks and higher variability in evening hours. -->

<!-- # interaction for two households, importance could be across one -->

<!-- Even when it is known which all interactions to look at, all of them would not be interesting and that too will vary across different households.  -->

<!-- calendar plots -->
<!-- The authors show how hour-of-the-day interact with weekdays and weekends and then move on to use calendar display to show daily schedules. The calendar display has several components in it, which helps us to look at energy consumption across hour-of-day, day-of-week, week-of-month, and month-of-year at once. Some interaction of these cyclic granularities, for example, how  day-of-week relates to month-of-year, could also be interpreted from this display. This is a great way for having an overview of energy consumption. However, if one wants to understand the repetitiveness in energy behavior and how they interact in greater detail, it is not easy to comprehend the interactions of all cyclic granularities from this display, due to the combination of linear and cyclic representation of time. For example, this display might not be the best to understand how hour-of-the-day or month-of-year varies across week-of-the-month as well as with each other. Furthermore, it is not clear what all interactions of cyclic granularities should be read from this display as there could be many combinations that one can look at. Moreover, there could be non-conventional cyclic granularities, which could potentially become useful depending on the context.  -->

<!-- context when it could be useful is monitoring heart rates. There are devices that can detect each heartbeat and transfer the data to a receiver such as a watch or phone. These data could be available for a temporal scale as fine as a minute and it could be of interest to see regular patterns across any deconstruction of time coarser than a minute. -->

```{r intro_all}

```

```{r id2-new2, fig.cap = "Distribution of energy consumption displayed through area quantile plots across two cyclic granularities month-of-year and hour-of-day and two households. The black line is the median, whereas the orange band covers the 25th to 75th percentile and the green band covers the 10th to 90th percentile. Difference between the 90th and 75th quantiles is less for (Jan, Feb) for the first household (a), suggesting that it is a more frequent user of air conditioner than the second household (b). Energy consumption for in (a) changes across both granularities, whereas for (b) daily pattern stays same irrespective of the months."}
```

```{r id4, fig.cap = "something2", eval = FALSE}

```

<!-- Again, look at \ref{fig:} c and d, where energy consumption for these two households are plotted against (weekend/weekday, week-of-month). Here, for both households, the pattern of energy consumption vary  across different weeks of the month irrespective of the fact it is a weekday or weekend. In that respect, the harmony pair (month-of-year, hour-of-day) seems to be more informative than (weekend/weekday, week-of-month) for the first household.  -->

<!-- Take an example of a data set which are observed at fine temporal scales, like that of NYC bike usage available at https://www.citibikenyc.com/system-data. We use the `nyc_bikes` data set from the R package `tsibbledata` which takes a sample of 10 bikes for the year 2018. The `start_time` and the `stop_time` are recorded to a fineness of seconds. We can look at pair of cyclic granularities (hour_day, wknd_wday) or (week_month, day_week) to see how these periodicities interact. But there could be other pairs that are important too. How to understand which pairs are sufficient to explore given the data set without losing much information about the data. -->

<!-- When we need to understand the interplay of different periodicities in a high frequency temporal datasets, we have many choices to consider. In [@wang2020tsibble] and [@wang2020calendar], periodicities are explored across hour of the day and day of the week or months. But calendar effects are not restricted to conventional day-of-week or month-of-year deconstructions. -->

<!-- If we have $n$ periodic linear granularities in the hierarchy table, then $n(n-1)/2$ circular or quasi-circular cyclic granularities could be constructed.  -->

<!-- Our contributions in this paper are: -->

 <!-- * We introduce a new distance measure for detecting periodic interactions. This induces data reduction which allows for identification of patterns, if any, in the time series data. -->

 <!-- * We show that the distance metric could be used to rank the periodic patterns based on how well they capture the variation in the measured variable as they have been normalized for different number of comparisons. -->

 <!-- * We device a framework for choosing a threshold, which will result in detection of only significantly interesting periodic patterns in the time series data. -->

 <!-- *  -->

 <!-- * ; -->

 <!-- * show that the proposed distance metric could be used to rank the interesting patterns across different datasets and temporal granularities since they have been normalized for relevant parameters. -->

\noindent The article is organized as follows. Section \ref{sec:computation-wpd} introduces a distance measure for detecting distributional difference in temporal granularities, which enables identification of patterns in the time series data;
Section \ref{sec:rank-wpd} devises a selection criterion by choosing a threshold, which results in detection of only significantly interesting patterns. Section \ref{sec:simulations} provides some simulation study on the proposed methodology. Section \ref{sec:application-wpd} presents an application to residential smart meter data in Melbourne to show how the proposed methodology can be used to automatically detect temporal granularities along which distributional differences are significant.

<!-- acts as a way to automatically detect periodic patterns in time series. -->

<!-- Also, ranking the remaining harmony pairs based on how well they capture the variation in the measured variable could be potentially useful.   -->

<!-- Talk about Scagnostics: Tukey -->

## Proposed distance measure {#sec:computation-wpd}

We propose a measure called Weighted Pairwise Distances ($wpd$) to detect distributional differences in the measured variable across cyclic granularities.

### Principle

The principle behind the construction of $wpd$ is explained through a simple example explained in Figure \ref{fig:null4by2}. Each of these figures describes a panel with $2$ x-axis categories and $3$ facet levels, but with different designs. Figure \ref{fig:null4by2}a has all categories drawn from N(0, 1) distribution for each facet. It is not an interesting display particularly, as distributions do not vary across x-axis or facet categories. Figure \ref{fig:null4by2}b has x categories drawn from the same distribution, but across facets the distributions are $3$ standard deviations apart. Figure \ref{fig:null4by2}c exhibits an exact opposite situation where distribution between the x-axis categories
are $3$ standard deviations apart, but they do not change across facets. Figure \ref{fig:null4by2}d takes a step further by varying the distribution across both facet and x-axis categories by  $3$ standard deviations. If the panels are to be ranked in order of capturing maximum variation in the measured variable from minimum to maximum, then an obvious choice would be placing (a) followed by (b), (c) and then (d). It might be argued that it is not clear if (b) should precede or succeed (c) in the ranking. Gestalt theory suggests items placed at close proximity can be compared more easily, because people assume that they are in the same group and apart from other groups. With this principle in mind, display (b) is considered less informative as compared to display (c) in emphasizing the distributional differences. Considering one cyclic granularity, we would have only two design choices similar to (a) and (c), corresponding to no difference and significant differences between categories of that cyclic granularity only. The proposed measure $wpd$ is constructed in a way so that it could be used to rank panels of different designs as well as test if a design is interesting. This measure is aimed to be an estimate of the maximum variation in the measured variable explained by the panel. A higher value of  $wpd$ would indicate that the panel is interesting to look at, whereas a lower value would indicate otherwise.

<!-- To compute the maximum variation of the measured variable within each panel, the distributional differences need to be measured for both within-group and between-group categories. Moreover, a tuning parameter specifying the weightage given to within-facet or between-facet categories can help to choose between designs like \ref{fig:null4by2}b and c. The maximum of all these weighted differences serves as an estimate of the maximum variation in the panel and is taken as the value of $wpd$. -->

<!-- Larger differences imply stronger patterns, whereas small difference would imply that the underlying structure is not changing within or between group. -->

<!-- A distance measure aimed to capture the structure of the measured variable in designs like \ref{fig:null4by2}, should ideally estimate these within-group and between-group variations. -->

<!-- \noindent Intuitively, while finding a structure or measuring the strength of patterns in Figure \ref{fig:null4by2}, it makes sense to look for within-group and between-group variations. Larger variation would imply stronger patterns, whereas small variation would imply the underlying structure is not changing within or between group. -->
<!-- Thus, a distance measure aimed to capture this structure should ideally estimate these within-group and between-group variations. One of the potential ways to do this is to measure the distances between distributions of the continuous random variable measured within and between groups, weigh them basis if they are within or between groups and then take the maximum of those distances as an estimate of the maximum variation in the structure. This section starts with possible ways of characterizing distributions and computing distances between them and then describe in details how the measure $wpd$ is defined. -->

<!-- This is similar to @ieee-irish where the authors compute the Jensen Shannon distance between two density estimates by computing percentiles and stresses the advantages to working with percentiles rather than the raw data directly in case of missing observations. Working with quantiles also ensure that unsynchronized time series could be handled. -->

<!-- Hence, with reference to the graphical design in @Gupta2021-hd, therefore the idea would be to rate a harmony pair higher if the variation between different levels of the x-axis variable is higher on an average across all levels of the facet variables.  -->

<!-- To elaborate further, look at the examples in Figure \ref{}, where Figure \ref{}a represents the panel design with distribution of each x categories drawn from N(5, 10) distribution. It could be observed that the graph is not particularly interesting, as there is no significant change in distribution between x-axis levels or facets. Figure \ref{}b represents the same panel design with no difference in distribution of x-axis categories within a facet, but different distribution of x-axis categories for different facets. For example, if there are 4 facet levels and 2 x-axis levels, data is generated in the way as described in Table \ref{}. Figure \ref{}b exhibits an exact opposite situation where the x-axis within facets are different but not across facets. -->
<!-- Figure \ref{}d takes it further by varying the distribution across both facet and x-axis categories. -->

```{r null4by2,fig.cap = "An example illustrating the principle of the proposed distance measure, displaying the distribution of a normally distributed variable in four panels each with $2$ x-axis categories and $3$ facet levels, but with different designs. Display (a) is not interesting as the distribution of the variable does not depend on x or facet categories. Display (b) and (c) are more interesting than (a) since there is a change in distribution either across facets (b) or x-axis (c). Display (d) is most interesting in terms of capturing structure in the variable as the distribution of the variable changes across both facet and x-axis variable. The value of our proposed distance measure is presented for each panel, the relative differences between which will be explained later in Section 3.2."}
```

<!-- A "significant" cyclic granularity have at least two categories which are significantly different from each other. It does not tell you which categories are statistically significantly different from each other -->

<!-- in the same group would be important to bring out different patterns of the data. -->

### Notations

\noindent Let the number of cyclic granularities considered in the display be $m$. The notations and methodology are described in detail for $m=2$. But it can be easily extended to $m>2$. Consider two cyclic granularities $A$ and $B$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$ with $A$ placed across x-axis and $B$ across facets. Let $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ be a continuous variable observed across $T$ time points. This  data structure with $J$ x-axis levels and $K$ facet levels is referred to as a $(J, K)$ panel. For example, a $(2, 3)$ panel will have cyclic granularities with 2 x-axis levels and 3 facet levels. Let the four elementary designs as described in Figure \ref{fig:null4by2} be $D_{null}$ (referred to as "null distribution") where there is no difference in distribution of $v$ for $A$ or $B$, $D_{var_f}$ denotes the set of designs where there is difference in distribution of $v$ for $B$ and not for $A$. Similarly, $D_{var_x}$ denotes the set of designs where difference is observed only across A. Finally, $D_{var_{all}}$ denotes those designs for which difference is observed across both $A$ and $B$. $m = 1$ is a special case of $m = 2$ with $J = 1$.

```{r notations}

```

  <!-- Thus, by CLT, ${\chi^2}_{m} \tilde{} N(m, 2m)$, which would depend on the number of discretization used to approximate the continuous distribution.  -->
  <!-- Then $b_n = 1-1/n$ quantile of the normal distribution and $a_n = 1/[n*\phi(b_n)]$ where $\phi$ is the normal density function. $n$ is the number of pairwise comparisons being made. -->

<!-- The Jensen-Shanon distance between two probability distribution $p_1$ and $p_2$ is given by $$d = [D(p_1, r) + D(p_2, r)]/2 \quad where \quad r = (p_1 + p_2)/2$$ where, -->
<!-- $$D(p_1,p_2) = \int^{\infty}_{-\infty}p_1(x)log\frac{p_1(x)}{p_2(x)}\,dx$$ is the Kullback-Leibler divergence between $p_1$ and $p_2$.   -->

<!-- We call this measure of variation as  Median Maximum Pairwise Distances (MMPD). -->

<!-- #### Distribution of Jensen-Shannon distances -->

<!-- Jensen-Shannon distances (JSD) are distributed as chi-squared with $m$ df where we discretize the continuous distribution with $m$ discrete values. Taking sample percentiles to approximate the integral would mean taking $m = 99$. -->
<!-- With large $m$, chi-squared is asymptotically normal by the CLT. Thus, by CLT, ${\chi^2}_{m} \tilde{} N(m, 2m)$, which would depend on the number of discretization used to approximate the continuous distribution. Then $b_n = 1-1/n$ quantile of the normal distribution and $a_n = 1/[n*\phi(b_n)]$ where $\phi$ is the normal density function. $n$ is the number of pairwise comparisons being made. -->

### Computation

The computation of the distance measure $wpd$ for a panel involves characterizing distributions, computing distances between distributions, choosing a tuning parameter to specify the weightage for different group of distances and summarizing those weighted distances appropriately to estimate maximum variation. Furthermore, the data needs to be appropriately transformed to ensure that the value of $wpd$ emphasizes detection of distributional differences across categories and not across different data generating processes.

<!-- but not when underlying distributions are different -->

<!-- irrespective of the distribution from which the data is generated. -->

<!-- is aimed to capture structure and patterns by estimating the maximum variation of the measured variable within a panel.  $wpd$ -->

<!-- the distributional differences need to be measured for both within-group and between-group categories. Moreover, a tuning parameter specifying the weightage given to within-facet or between-facet categories can help to choose between designs like \ref{fig:null4by2}b and c. The maximum of all these weighted differences serves as an estimate of the maximum variation in the panel and is taken as the value of $wpd$.  -->

#### Data transformation

The intended aim of $wpd$ is to capture differences in categories irrespective of the distribution from which the data is generated. Hence, as a pre-processing step, the raw data is normal-quantile transformed (NQT) (@Krzysztofowicz1997-bv), so that the quantiles of the transformed data follows a standard normal distribution. This sort of transformation is common in the fields of geo-statistics to make most asymmetrical distributed real world measured variables more treatable and normal-like (@Bogner2012-az).The empirical NQT involves the following steps:  
 1. The sample of measured variable $v$ is sorted from the smallest to the largest observation $v_{(1)},\dots, v_{(i)}, .., v_{(n)}$.  
 2. The cumulative probabilities $p_{(1)},\dots, p_{(i)}, .., p_{(n)}$ are estimated using a plotting position like $i/(n + 1)$ such that $p_{(i)} = P(v \leq v_{(i)})$.  
 3. Each observation $v_{(i)}$ of $v$ is transformed into observation $v^*(i) = Q^{-1}(p(i))$ of the standard normal variate $v^*$, with $Q$ denoting the standard normal distribution and $Q^{-1}$ its inverse.

#### Characterising distributions

Multiple observations of $v$ correspond to the subset  $v_{jk} = \{s: A(s) = j, B(s) = k\}$. The number of observations might vary widely across subsets due to the structure of the calendar, missing observations or uneven locations of events in the time domain. in this paper, quantiles of $v_{jk}$'s are chosen as a way to characterize distributions for the category $(a_j, b_k)$, $\forall j\in \{1, 2, \dots, J\}, k\in \{1, 2, \dots, K\}\}$. The quantile of a distribution with probability $p$ is defined as $Q(p)=F^{-1}(p) = inf\{x: F(x) >p\}$, $0<p< 1$ where $F(x)$ is the distribution function. There are two broad approaches to quantile estimation, viz, parametric and non-parametric. Sample quantiles (@Hyndman1996-ty) are used for estimating population quantiles in a non-parametric setup, which is desirable because of less rigid assumptions made about the nature of the underlying distribution of the data. The default quantile chosen in this paper is percentiles computed for $p = {0.01, 0.02, \dots, 0.99}$, where for example, the $99^{th}$ percentile would be the value corresponding to $p=0.99$ and hence 99% of the observations would lie below that.

<!-- The benefit of using a non-parametric estimator is that there are less rigid assumptions made about the nature of the underlying distribution of the data. Sample quantiles could be used for estimating population quantiles in a non-parametric setup. describes the many ways of defining sample quantiles and recommends the use of median-unbiased estimator because of _desirable properties of a quantile estimator and can be defined independently of the underlying distribution._.  -->

<!-- in this paper. -->

<!-- Each $v_{jk}$'s $\forall j\in \{1, 2, \dots, J\}, k\in \{1, 2, \dots, K\}\}$ are assumed to be drawn from a continuous probability distribution and have certain characteristics. Shape, central tendency, and variability are the common characteristics used to describe a distribution. -->

<!-- Mean, median or mode are generally used to describe the center of the distribution, while range, standard deviation, quantiles, standard errors and confidence intervals are used to describe variability. -->

<!-- The quantile of a distribution with probability $p$ is defined as $Q(p)=F^{-1}(p) = inf\{x: F(x) >p\}$, $0<p< 1$ where $F(x)$ is the distribution function. There are two broad approaches to quantile estimation, viz, parametric and non-parametric. The benefit of using a non-parametric estimator is that there are less rigid assumptions made about the nature of the underlying distribution of the data. Sample quantiles could be used for estimating population quantiles in a non-parametric setup. @Hyndman1996-ty describes the many ways of defining sample quantiles and recommends the use of median-unbiased estimator because of _desirable properties of a quantile estimator and can be defined independently of the underlying distribution._. The `stats::quantile()` function in @R-language could be used for practical implementation where type = 8 refers to the algorithm corresponding to the median-unbiased estimator. The default quantile chosen in this paper is percentiles computed for $p = {0.01, 0.02, \dots, 0.99}$, where for example, the $99^{th}$ percentile would be the value corresponding to $p=0.99$ and hence 99% of the observations would lie below that. -->

#### Distance between distributions

<!-- One of the most important class of divergence is the f-divergence and includes measures like Kullback-Leibler divergence, Hellinger distance etc. The continuous version of f -divergence is given by -->
<!-- $$D_f(P||Q) := \int q(x)f(\frac{p(x)}{q(x)})$$, where -->
<!-- $f : [0,\infty) \rightarrow R \cup \{\infty\}$ is a continuous convex function, and $f(1) = 0$.  -->

One of the most common ways to measure divergence between distributions is the Kullback-Leibler (KL) divergence [@Kullback1951-jy]. The KL divergence denoted by $D(q_1||q_2)$ is a non-symmetric measure of the difference between two probability distributions $q_1$ and $q_2$ and is interpreted as the amount of information lost when $q_2$ is used to approximate $q_1$. The KL divergence, however, is not symmetric and hence can not be considered as a "distance" measure. The Jensen-Shannon divergence [@Menendez1997-in] based on the Kullback-Leibler divergence is symmetric and has a finite value. Hence, in this paper, the pairwise distances between the distributions of the measured variable are obtained through the square root of the Jensen-Shannon divergence, called Jensen-Shannon distance (JSD) and is defined by,

$$JSD(q_1||q_2) = \frac{1}{2}D(q_1||M) + \frac{1}{2}D(q_2||M)$$
where $M = \frac{q_1+q_2}{2}$ and 
$D(q_1||q_2) := \int^\infty_{-\infty} q_1(x)f(\frac{q_1(x)}{q_2(x)})$ is the KL divergence between distributions $q_1$ and $q_2$. Other common measures of distance between distributions are Hellinger distance, total variation distance and Fisher information metric.

<!-- Furthermore, these distances are distributed as chi-squared with $m$ degrees of freedom (@Menendez1997-in), if the continuous distribution is being discretized with $m$ discrete values. Taking sample percentiles to approximate the integral would mean taking $m = 99$. As the degrees of freedom $m$ get larger, the chi-square distribution approaches the normal distribution. -->

#### Within-facet and between-facet distances

\noindent Pairwise distances could be within-facets or between-facets for $m \geq 2$. Figure \ref{fig:distance-explain} illustrates how they are defined. Pairwise distances are within-facets when $b_{k} = b_{k'}$, that is, between pairs of the form $(a_{j}b_{k}, a_{j'}b_{k})$ as shown in panel (3) of Figure \ref{fig:distance-explain}. If categories are ordered (like all temporal cyclic granularities), then only distances between pairs where $a_{j'} = (a_{j+1})$ are considered (panel (4)). Pairwise distances are between-facets when they are considered between pairs of the form $(a_{j}b_{k}, a_{j}b_{k'})$. Number of between-facet distances would be $^{K}C_2*J$ and number of within-facet distances are $K*(J-1)$ (ordered) and $^{J}C_2*K$ (un-ordered).

#### Tuning parameter

A tuning parameter specifying the weightage given to the within-facet or between-facet categories can help to balance weightage between designs like \ref{fig:null4by2}(b) and (c). The tuning parameters should be chosen such that $\sum_{i=1}^m \lambda_i = 1$. When $m=2$, following the principle of Gestalt theory, $\lambda = \frac{2}{3} = 0.67$ is chosen to put a relative weightage of $2:1$ for within-facet and between-facet distances. No human experiment is conducted to justify this ratio, however, typically a tuning parameter $\lambda>0.5$ would tend to upweigh the within-facet distances and that with $<0.5$ would upweigh the between-facet distances (refer to the Supplementary paper for more details). For $m = 1$, there are no conditioning variables or groups, and hence $\lambda = 1$.

#### Raw distance measure

The raw distance measure, denoted by $wpd_{raw}$, is computed after combining all the weighted distance measures appropriately. First, NQT is performed on the measured variable $v_t$ to obtain $v^*_t$ (_data transformation_). Then, for a fixed harmony pair $(A, B)$, percentiles of $v^*_{jk}$ are computed and stored in $q_{jk}$ (_distribution characterization_). This is repeated for all pairs of categories of the form $(a_{j} b_{k}, a_{j'}b_{k'}): \{a_j: j = 1, 2, \dots, J\}, B = \{ b_k: k = 1, 2, \dots, K\}$. The pairwise distances between pairs $(a_{j} b_{k}, a_{j'}b_{k'})$  denoted by $d_{(jk, j'k')} = JSD(q_{jk}, q_{j'k'})$ is computed (_distance between distributions_). The pairwise distances (_Within-facet and between-facet_) are transformed using a suitable _tuning parameter_ ($0<\lambda<1$) depending on if they are within-facet($d_w$) or between-facets($d_b$) as follows:
\begin{equation}\label{def:raw-wpd}
d*_{(j,k), (j'k')} =
\begin{cases}
\lambda d_{(jk), (j'k')},& \text{if } d = d_w\\
(1-\lambda)  d_{(jk), (j'k')},              & \text{if } d = d_b\\
\end{cases}
\end{equation}

\noindent The $wpd_{raw}$ is then computed as $$wpd = max_{j, j', k, k'}(d*_{(jk), (j'k')}) \forall j, j' \in \{1, 2, \dots, J\}, k, k' \in \{1, 2, \dots, K\}$$

\noindent The statistic "maximum" is chosen to combine the weighted pairwise distances since the distance measure is aimed at capturing the maximum variation of the measured variable within a panel. The statistic "maximum" is, however, affected by the number of comparisons (resulting pairwise distances). For example, for a $(2, 3)$ panel, there are $6$ possible subsets of obervations corresponding to the combinations $(a_1, b_1), (a_1, b_2), (a_1, b_3), (a_2 ,b_1), (a_2 ,b_2), (a_2, b_3)$, whereas for a $(2, 2)$ panel, there are only $4$ possible subsets $(a_1, b_1), (a_1, b_2), (a_2 ,b_1), (a_2 ,b_2)$. Consequently, the measure would have higher values for the panel $(2, 3)$ as compared to $(2, 2)$, since maximum is taken over higher number of pairwise distances.

#### Adjusting for the number of comparisons {#sec:norm-wpd}

Ideally, it is desired that the proposed distance measure takes a higher value only if there is a significant difference between distributions across categories, and not because the number of categories $J$ or $K$ is high. That is, under designs like $D_{null}$, their distribution should not differ for a different number of categories. Only then the distance measure could be compared across panels with different levels. This calls for an adjusted measure, which normalizes for the different number of comparisons. We denote it by $wpd$. Two approaches for adjusting the number of comparisons are discussed, both of which are substantiated using simulations. The first one defines an adjusted measure $wpd_{perm}$ based on the permutation method to remove the effect of different comparisons. The second approach fits a model to represent the relationship between $wpd_{raw}$ and the number of comparisons and defines the adjusted measure ($wpd_{glm}$) as the residual from the model.

_Permutation approach_

This method is somewhat similar in spirit to bootstrap or permutation tests, where the goal is to test the hypothesis that the groups under study have identical distributions. This method accomplishes a different goal of finding the null distribution for different groups (panels in our case) and standardizing the raw values using that distribution. The values of $wpd_{raw}$ is computed on many ($nperm$) permuted data sets and stored in $wpd_{perm-data}$. Then $wpd_{perm}$ is computed as follows:

\begin{align*}
wpd_{perm} = \frac{(wpd_{raw} - \bar{wpd_{perm-data}})}{sd(wpd_{perm-data})}
\end{align*}

\noindent where $\bar{wpd_{perm-data}}$ and $sd(wpd_{perm-data})$ are the mean and standard deviation of $wpd_{perm-data}$ respectively. Standardizing  $wpd$ in the permutation approach ensures that the distribution of $wpd_{perm}$ under $D_{null}$ has the same $mean = 0$ and $\sigma^2_{perm} = 1$ across all comparisons. While this works successfully to make the location and scale similar across different $nx$ and $nfacet$, it is computationally heavy and time consuming, and hence less user friendly when being actually used in practice. Hence, another approach to adjustment, with potentially less computational time, is proposed.

<!-- for the $l^{th}$ panel and -->

_Modeling approach_

In this approach, a Gamma generalized linear model (GLM) for $wpd_{raw}$ is fitted with number of comparisons as the explanatory variable. Since, $wpd_{raw}$ is a Jensen-Shannon distance, it follows a Chi-square distribution (@Menendez1997-in), which is a special case of Gamma distribution. Furthermore, the mean response is bounded, since any JSD is bounded by $1$ given that base $2$ logarithm is used (@Lin1991-pt). Hence, by @Faraway2016-uk, an inverse link is used for the model, which is of the  form $y = a+b*log(z) + e$, where $y = wpd_{raw}$, $z = (nx*nfacet)$ is the number of groups and $e$ are idiosyncratic errors. Let $E(y) = \mu$ and $a + b*log(z) = g(\mu)$ where $g$ is the link function. Then $g(\mu)=  1/\mu$ and $\hat \mu  = 1/(\hat a + \hat b log(z))$. The residuals from this model $(y-\hat y) = (y-1/(\hat a + \hat b log(z)))$ would be expected to have no dependency on $z$. Thus, $wpd_{glm}$ is chosen as the residuals from this model and is defined as:
$$wpd_{glm} = wpd_{raw} - 1/(\hat a + \hat b*log(nx*nfacet))$$
The distribution of $wpd_{glm}$ under $D_{null}$ will have $mean = 0$, since it is the residuals from the model, and a constant variance $\sigma^2_{glm}$, which might not equal 1.

_Combination approach_

The simulation results (in Section \ref{sec:simulations}) show that the distribution of $wpd_{glm}$ under null is similar for high $nx$ and $nfacet$ (levels higher than $5$) and not so much for lower $nx$ and $nfacet$. Hence, a combination approach is proposed which chooses permutation approach for categories with smaller levels and modeling approach for categories with higher levels. This ensures that the computational load of the permutation approach is alleviated while maintaining similar null distribution across different categories. This approach, however, requires that the adjusted variables from the two approaches are brought to the same scale. We define $wpd_{glm-scaled} = wpd_{glm}*\sigma^2_{perm}/\sigma^2_{glm}$ as the transformed $wpd_{glm}$ with a similar scale as $wpd_{perm}$. The adjusted measure from the combination approach, denoted by $wpd$ is then defined as follows:

\begin{equation}
wpd =
\begin{cases}
wpd_{perm},& \text{if } J, K <=5\\
wpd_{glm-scaled}             & \text{otherwise}\\
\end{cases}
\end{equation}

<!-- ### Algorithm -->

<!-- The distance measure $wpd$ between two cyclic granularities $A$ and $B$ is aimed to capture the strength of the structure by estimating the maximum within-group and between-group variations. Furthermore,  -->

<!-- The steps employed for computing $wpd$ is summarized as follows: -->

<!-- The values of $wpd$ indeed depend upon the underlying distribution. This is also not desirable as the intended aim of $wpd$ is to capture differences in categories irrespective of the distribution from which the data is generated. The steps employed for computing the distance measure is summarized as follows: -->

<!-- Hence, differences in distributions of the measured variable are computed between all pairs of categories $(a_{j} b_{k}, a_{j'}b_{k'}): j = 1, 2, \dots, J\}, B = \{ b_k: k = 1, 2, \dots, K\}$. Percentiles of $v_{jk}$ and $v_{j'k'}$ are computed and stored in $q_{jk}$ and $q_{j'k'}$ respectively. Then the pairwise distances between pairs $(a_{j} b_{k}, a_{j'}b_{k'})$ be denoted as $d_{(jk, j'k')} = JSD(q_{jk}, q_{j'k'})$ is computed. Pairwise distances could be within-facets or between-facets. Figure \ref{fig:distance-explain} illustrates how the within-facet or between-facet distances are defined. Pairwise distances are within-facets ($d_{w}$) when $b_{k} = b_{k'}$, that is, between pairs of the form $(a_{j}b_{k}, a_{j'}b_{k})$ as shown in panel (3) of Figure \ref{fig:distance-explain}. If categories are ordered (like all temporal cyclic granularities), then only distances between pairs where $a_{j'} = (a_{j+1})$ are considered (panel (4)). Pairwise distances are between-facets ($d_{b}$) when they are considered between pairs of the form $(a_{j}b_{k}, a_{j}b_{k'})$. Number of between-facet distances would be $^{K}C_2*J$ and number of within-facet distances are $K*(J-1)$ (ordered) and $^{J}C_2*K$ (un-ordered). The pairwise distances $d_{(jk, j'k')}$ are transformed using a suitable tuning parameter ($0<\lambda<1$) depending on if they are $d_{b}$ or $d_{w}$ as follows: -->

```{r distance-explain, fig.cap = "Within and between-facet distances shown for two cyclic granularities A and B, where A is mapped to x-axis and B is mapped to facets. The dotted lines represent the distances between different categories. Panel 1) and 2) show the between-facet distances. Panel 3) and 4) are used to illustrate within-facet distances when categories are un-ordered or ordered respectively. When categories are ordered, distances should only be considered for consecutive x-axis categories. Between-facet distances are distances between different facet levels for the same x-axis category, for example, distances between {($a_1$,$b_1$) and ($a_1$, $b_2$)} or {($a_1$,$b_1$) and ($a_1$, $b_3$)}."}

```

<!-- 5. Compute distribution of maximum distance ($M$) by shuffling the data $200$ times and finishing Steps1-4 in each case. Scale $M$ using the mean and sd of the distribution.

5. Use Steps 1-4 to compute maximum distance for $\forall k \in  \{1, 2, \ldots, K\}$.

6. Compute the distance measure MMPD_raw = median $(M_1, M_2, \dots, M_K)$.

<!-- All of these distances are then aggregated by taking the maximum from these distances to obtain $wpd$. -->

<!-- _j$ maps index set to a set $\{B_\ell \mid \ell =1,\dots,L\}$. Here, $A_k$ and $B_\ell$ are the levels/categories corresponding to $C_i$ and $C_j$ respectively. Let $S_{k\ell}$ be a subset of the index set such that for all $s \in S_{k\ell}$, $C_i(s) = A_k$ and $C_j(s) = B_\ell$. There are $KL$ such data subsets, one for each combination of levels ($A_k$, $B_\ell$). Moreover, consider that in the graphical space, $C_i$ is mapped to facets and $C_j$ is mapped to x-axis. -->

 <!-- old -->
<!-- Consider two cyclic granularities $C_i$ and $C_j$, such that $C_i$ maps index set to a set $\{A_k \mid k=1,\dots,K\}$ and $C_j$ maps index set to a set $\{B_\ell \mid \ell =1,\dots,L\}$. Here, $A_k$ and $B_\ell$ are the levels/categories corresponding to $C_i$ and $C_j$ respectively. Let $S_{k\ell}$ be a subset of the index set such that for all $s \in S_{k\ell}$, $C_i(s) = A_k$ and $C_j(s) = B_\ell$. There are $KL$ such data subsets, one for each combination of levels ($A_k$, $B_\ell$). Moreover, consider that in the graphical space, $C_i$ is mapped to facets and $C_j$ is mapped to x-axis.  -->

<!-- The algorithm employed for computing the distance measure is summarized as follows: -->

<!-- 1. Fix harmony pair $(C_i, C_j)$. -->

<!-- 2. Fix $k$. Then there are $L$ groups corresponding to level $A_k$ of $C_i$. -->

<!-- 3. Compute  $m = \binom{L}{2}$ pairwise distances between distributions of $L$ unordered levels and $m = L-1$ pairwise distances for $L$ ordered categories. -->

<!-- 4. Identify maximum within the $m$ computed distances. -->

<!-- <!-- 5. Compute distribution of maximum distance ($M$) by shuffling the data $200$ times and finishing Steps1-4 in each case. Scale $M$ using the mean and sd of the distribution.  -->

<!-- 5. Use Steps 1-4 to compute maximum distance for $\forall k \in  \{1, 2, \ldots, K\}$. -->

<!-- 6. Compute the distance measure MMPD_raw = median $(M_1, M_2, \dots, M_K)$. -->

<!-- \begin{algorithm}[!thb] -->
<!-- 	\caption{Calculation for a raw distance measure between two cyclic granularities $A = \{ a_j: j = 1, 2, \dots, J\}$, $B = \{ b_k: k = 1, 2, \dots, K\}$ with $A$ placed across x-axis and $B$ across facets.}\label{alg:scoreopt}. -->
<!-- 	\begin{algorithmic}[1] -->
<!-- 		\Procedure{RawMMPD}{$A = \{ a_j: j = 1, 2, \dots, J\}$, $B = \{ b_k: k = 1, 2, \dots, K\}$, $v = \{ v_t: t = 1, 2, \dots, T\}$}. -->
<!-- 		\For{$k=1:K$, $j=1:J$} -->
<!-- 		\State Find distances between pairs of all possible combinations of categories $(a_jb_k,a_j'b_k')$ by computing JSD between quantiles of the measured variable $q(v)$ across these combinations. -->
<!-- 		\State $d \gets JSD(\tilde{q(v)_{a_{j}b_{k}}},\tilde{q(v)_{a_{j}'b_{k}'}})$ -->
<!-- 		\If {$b_k = b_k'$} -->
<!-- 		\State $d* \gets \lambda d$  \Comment{upweight within-facet distances} -->
<!-- 		\Else -->
<!-- 		\State $d* \gets 1/\lambda d $ \Comment{downweight across-facet distances} -->
<!-- 		\EndFor -->
<!-- 		\State Set the raw distance measure as $max(d*)$ where max is taken over all $j, j', k, k'$. -->

<!-- 		\EndProcedure -->
<!-- 	\end{algorithmic} -->
<!-- \end{algorithm} -->

<!-- A stronger measure "max" is chosen for aggregating x-axis categories compared to "median" for aggregating facet categories as the measure is intended to put more importance in pointing towards distributional differences between x-axis categories than for facet categories. -->

<!-- ## Properties of $wpd$ -->

<!-- Simulations were carried out to explore the behavior of $wpd$ under the following factors that could potentially impact the values of $wpd$: $nx$, $nfacet$, $\lambda$, $\omega$, $dist$ (normal/non-normal distributions with different location and scale), $ntimes$, and $designs$ and results are presented in two parts. The dependence of $wpd$ on $nx$ and $nfacet$ under $D_{null}$ is presented here, which lays the foundation for the next section. The rest of the results that discuss the relationship of the $wpd$ with other factors is presented in details in the Supplementary section of the paper. They show that the designs $D_{var_f}$ and $D_{var_x}$ intersect at $\lambda = 0.5$ and hence for up-weighing designs of the form $D_{var_x}$, $\lambda = 0.67$ has been considered for computation of $wpd$ in the rest of the paper.  -->

<!-- - $nx$ (number of levels of x-axis) -->
<!-- - $nfacet$ (number of levels of facets) -->
<!-- - $\lambda$ (tuning parameter) -->
<!-- - $\omega$ (increment in each panel design) -->
<!-- - $dist$ (normal/non-normal distributions with different location and scale) -->
<!-- - $n$ (sample size for each combination of categories) -->
<!-- - $designs$ ($D_{null}$, $D_{var_f}$, $D_{var_x}$ and $D_{var_{all}}$) -->

<!-- - $nsim$ (number of simulations) -->
<!-- - $nperm$ (number of permutations of data)   -->

<!-- ### Simulation design{#sec:sim-study} -->

<!-- Observations are generated from a Gamma(2,1) distribution for each combination of $nx$ and $nfacet$ from the following sets: $nx = nfacet = \{2, 3, 5, 7, 14, 20, 31, 50\}$ to cover a wide range of levels from very low to moderately high. Each combination is being referred to as a _panel_. That is, data is being generated for each of the panels $\{nx = 2, nfacet = 2\}, \{nx = 2, nfacet = 3\}, \{nx = 2, nfacet = 5\},  \dots, \{nx = 50, nfacet = 31\}, \{nx = 50, nfacet = 50\}$. For each of the $64$ panels, $ntimes = 500$ observations are drawn for each combination of the categories. That is, if we consider the panel $\{nx = 2, nfacet = 2\}$, $500$ observations are generated for each of the combination of categories from the panel, namely, $\{(1, 1), (1, 2), (2, 1), (2, 2)\}$. The values -->
<!-- of $wpd$ is obtained for each of the panels. This design corresponds to $D_{null}$ as each combination of categories in a panel are drawn from the same distribution. Furthermore, the data is simulated for each of the panels $nsim=200$ times, so that the distribution oaf $wpd$ under $D_{null}$ could be observed. $wpd_{l, s}$ denotes the value of $wpd$ obtained for the $l^{th}$ panel and $s^{th}$ simulation. -->

<!-- ### Results -->

<!-- Figure \ref{fig:raw} shows the distribution of $wpd$ plotted across different $nx$ and $nfacet$ categories. Since under $D_{null}$, there is no difference in distributions across different categories, we expect the distance measure $wpd$ to reflect that as well and have the same distribution across categories. But Figure \ref{fig:raw} shows that both the location and scale of the distributions change across panels. This is not desirable under $D_{null}$ as it would mean comparisons of $wpd$ values is not appropriate across different $nx$ and $nfacet$. Figure \ref{fig:quadratic} shows how the median of $wpd$ varies with the total number of distances $nx*nfacet$ for each panel. The median increases abruptly for lower values of $nx*nfacet$ and slowly for higher $nx*nfacet$.  -->

<!-- Furthermore, simulation results also show that the values of $wpd$ indeed depend upon the underlying distribution. This is also not desirable as the intended aim of $wpd$ is to capture differences in categories irrespective of the distribution from which the data is generated. -->

<!-- last writeup -->
<!-- In the linear model approach, $wpd\in R$ was assumed, whereas, $wpd$ is a Jensen-Shannon Distance (JSD) and lies between 0 and 1 (@JSD). Furthermore, JSD follows a Chi-square distribution, which is a special case of Gamma distribution. Therefore, a generalized linear model could be fitted instead of a linear model to allow for the response variable to follow a Gamma distribution. The inverse link is used when we know that the mean response is bounded, which is applicable in our case since $0 \leq wpd\leq 1$. -->

<!-- We fit a Gamma generalized linear model with the inverse link which is of the form:   -->
<!-- $$y_l = a+b*log(z_l) + e_l$$, where $y_l = median_m(wpd_{l, m})$, $z_l$ is the $l^{th}$ panel and $e_l$ are idiosyncratic errors. Let $E(y) = \mu$ and $a + b*log(z) = g(\mu)$ where $g$ is the link function. Then $g(\mu)=  1/\mu$ and $\hat \mu  = 1/(\hat a + \hat b log(z))$. The residuals from this model $(y-\hat y) = (y-1/(\hat a + \hat b log(z)))$ would be expected to have no dependency on $z$. Thus, $wpd_{glm}$ is chosen as the residuals from this model and is defined as: -->
<!-- $wpd_{glm} = wpd - 1/(\hat a + \hat b*log(nx*nfacet))$. -->

<!-- which is more approximate than exact but still has the similar accuracy when compared to the permutation approach. -->

<!-- use later -->
<!-- The measure $wpd$ could potentially lead to comparison of the measure across different panels and also help distinguishing the interesting panels from a data set. We discuss two approaches for normalization, both of which are substantiated using simulations. -->

<!-- ## Methodology -->

<!-- The transformed ${wpd}$ which is normalized for the values of $nx$ and $nfacet$ is denoted by $wpd$. Two approaches have been employed - the first one involves a permutation method to make the distribution of the transformed $wpd$ similar for different comparisons and the second one fits a model to represent the relationship between the two variables and defines $wpd$ as the residual of the model. -->

<!-- ### Notations -->

<!-- Let $\{nx_{i}, i = 1, 2, \dots, nx\}$, $\{nfacet_{j}, j = 1, 2, \dots, nfacet\}$ be the set of x-axis and facet categories respectively. Each combination of $nx_{i}$ and $nfacet_{j}$is being referred to as a _panel_. Then the total number of panel is $nx*nfacet$. Let the total number of pairwise distances that could result in each panel be $\{z_k, k = 1, 2 , \dots, nx*nfacet\}$. Here, $\{z_1 = nx_1*nfacet_1\}$, $\{z_2 = nx_2*nfacet_2\}$ and $\{z_k = nx*nfacet\}$. Now, let $\{x_{k,l}, k = 1, 2, \dots, nx*nfacet, l = 1, 2, \dots, nsim\}$ denote the values of $wpd_{raw}$ obtained from the simulation study for $k^{th}$ panel in the $i^{th}$ simulation. Hence, for each of those $k$ panel, we have $nsim$ values of $wpd_{raw}$. -->

<!-- use soon -->
<!-- ## Properties -->

<!-- This section reports the results of a simulation study that was carried out to evaluate the behavior of $wpd$ under different designs and other potential factors. The behavior of $wpd$ is explored in designs where there is in fact difference in distribution between facet categories ($D_{var_f}$) or across x-categories ($D_{var_x}$) or both ($D_{var_{all}}$). Using $\omega =  \{1, 2, \dots, 10\}$ and $\lambda = seq(from  = 0.1, to = 0.9, by = 0.05)$, observations are drawn from a N(0,1) distribution for each combination of $nx$ and $nfacet$ from the following sets: $nx = nfacet =  \{2, 3, 5, 7, 14, 20, 31, 50\}$. $ntimes = 500$ is assumed for this setup as well. Furthermore, to generate different distributions across different combination of facet and x levels, the following method is deployed - suppose the distribution of the combination of first levels of $x$ and $facet$ category is $N(\mu,\sigma)$ and $\mu_{jk}$ denotes the mean of the combination $(a_jb_k)$, then $\mu_{j.} = \mu + j\omega$ (for design $D_{var_x}$) and $\mu_{.k} = \mu + k\omega$ (for design $D_{var_f}$).  -->

<!-- The tabulated values and graphical representations of the simulation results are provided in the Supplementary paper. The learning from the simulations are as follows: The values of $wpd$ is least for $D_{null}$, followed by -->
<!-- $D_{var_f}$, $D_{var_x}$ and $D_{var_{all}}$. This is a desirable result since the measure $wpd$ was designed such that this relationship holds. Furthermore, the distribution of the measure $wpd$ does not change for different facet and x categories. The distribution of $wpd$ looks similar with at least the mean and standard of the distributions being uniform across panels. This means $wpd$ could be used to measure differences in distribution across panels. Also, note that since the data is processed using normal-quantile-transform, this measure is independent of the initial distribution of the underlying data and hence is also comparable across different data sets. This is valid for the case when sample size $ntimes$ for each combination of categories is at least 30 and  $nperm$ used for computing $wpd$ is at least 100. More detailed results about the properties of $wpd$ could be found in the Supplementary paper. -->

## Ranking and selection of cyclic granularities {#sec:rank-wpd}

A cyclic granularity is referred to as "significant" if there is a significant distributional difference of the measured variable between different categories of the harmony. In this section, a selection criterion to choose significant harmonies is provided, thereby eliminating all harmonies that exhibit complete randomness in the measured variable. The distance measure $wpd$ is used as a test statistic to test the null hypothesis that no harmony/cyclic granularity is significant. We select only those harmonies/cyclic granularities for which the test fails. They are then ranked basis how well they capture variation in the measured variable.

<!-- . Randomization tests (permutation tests) generates a random distribution by re-ordering the observed data and allows us to test if the observed data is significantly different from any random distribution.  -->
<!-- the harmonies are not significant under null hypothesis and then test if our hypotheses are true: -->

<!--   $H_0$: harmonies are not significant -->
<!--   $H_1$: at least one harmony is significant -->

<!-- we provide a method to select important harmonies by eliminating all harmonies for which patterns are not significant by employing randomization test. Randomization tests (permutation tests) generates a random distribution by re-ordering our observed data and allow to test if the observed data is significantly different from any random distribution. Complete randomness in the measured variable indicates that the process follows a homogeneous underlying distribution over the whole time series, which essentially implies there is no interesting distinction across any different categories of the cyclic granularities.  -->

### Selection

<!-- Under the null hypothesis, the harmonies are not significant. -->

A threshold and consequently a selection criterion is chosen using the notion of Randomization tests (@edgington2007randomization). The data is permuted several times and $wpd$ is computed for each of the permuted data sets to obtain the sampling distribution of $wpd$ under the null hypothesis. If the null hypothesis is true, then $wpd$ obtained from the original data set would be a likely value in the sampling distribution. But in case the null hypothesis is not true, then it is less probable that $wpd$ obtained for the original data will be from the same distribution. This idea is utilized to come up with a threshold for selection, denoted by $wpd_{threshold}$, defined as the $99^{th}$ percentile of the sampling distribution. A harmony is selected if the value of $wpd$ for that harmony is greater than the chosen threshold. The detailed algorithm for choosing a threshold and selection procedure (for two cyclic granularities) is listed as follows:

<!-- The selection criterion is stated as follows: for $(i \in {1, 2, \dots, H_{N_C}})$, select $H_i$ if $wpd_i > wpd_{threshold}$, else reject. -->

- **Input:** All harmonies of the form $\{(A, B), A = \{ a_j: j = 1, 2, \dots, J\}, B = \{ b_k: k = 1, 2, \dots, K\}\}$, $\forall (A, B) \in H_{N_C}$.

- **Output:** Harmony pairs $(A, B)$ for which $wpd$ is significant.

1. Fix harmony pair $(A, B)$.

2. Given the measured variable; $\{v_t: t=0, 1, 2, \dots, T-1\}$, $wpd$ is computed and is represented by $wpd^{A, B}_{obs}$.

3. From the original sequence a random permutation is obtained: $\{v_t^{1}: t=0, 1, 2, \dots, T-1\}$.

4. $wpd$ is computed for the permuted sequence of the data and is represented by $wpd^{A, B}_{1}$.

5.  Steps (3) and (4) are repeated a large number of times $M$ (e.g. $M = 200$).

6. For each permutation, one $wpd^{A, B}_{i}$ is obtained. Define $wpd_{sample} = \{wpd^{A, B}_{1} wpd^{A, B}_{2}, \dots, wpd^{A, B}_{M}\}$.

7. Repeat Steps (1-6) for all harmony pairs $(A, B) \in H_{N_C}$ and stored $wpd^{all}_{sample}$.

8. $99^{th}$ percentiles of $wpd^{all}_{sample}$ is computed and stored in $wpd_{threshold99}$

7. If $wpd^{A, B}_{obs} > wpd_{threshold99}$, harmony pair $(A, B)$ is selected, otherwise rejected.

Similarly, a harmony pair $(A, B)$ could be selected if $wpd^{A, B}_{obs} > wpd_{threshold95}$ and $wpd^{A, B}_{obs} > wpd_{threshold90}$, where  $wpd_{threshold95}$ and $wpd_{threshold90}$ denote the $95^{th}$ and $90^{th}$ percentile of $wpd^{all}_{sample}$ respectively. A harmony selected using $99^{th}$,  $95^{th}$  and  $90^{th}$ threshold are tagged as \*\*\*, \*\*, \* respectively.

### Ranking

The distribution of $wpd$ is expected to be similar for all harmonies under the null hypothesis, since they have been adjusted for different number of categories for the harmonies or underlying distribution of the measured variable. Hence, the values of $wpd$ for different harmonies are comparable and can be used to rank the significant harmonies. A higher value of $wpd$ for a harmony indicates that higher maximum variation in the measured variable is captured through that harmony. Figure \ref{fig:null4by2} presents the results of $wpd$ from the illustrative designs in Section \ref{sec:computation-wpd}. The value of $wpd$ under null design (a) is the least, followed by (b), (c) and (d). This aligns with the principle of $wpd$, which is expected to have lowest value for null designs and highest for designs of the form $D_{var_{all}}$ (d). Moreover, note the relative differences in $wpd$  values between (b) and (c). The value of the tuning parameter $\lambda$ is set to $0.67$, which has resulted in giving more emphasis to differences in x-axis categories. Again consider \ref{fig:onegran-new}(a) and \ref{fig:onegran-new}(b) with a $wpd$ value of 20.5 and 145 respectively. This is because there is more gradual increase across hours of the day than months of the year. If order of the categories is not considered, they result in a $wpd$ value of $97.8$ and $161$ respectively, which follows from the fact that if we consider difference between any hours of the day, the magnitude will be much larger than if we consider difference between consecutive categories. Similarly, \ref{fig:onegran-new}(a) and (b) have $wpd$ values of   $110.79$ and $125.82$ respectively. The ranking implies that the distributional differences are more prominent for the second household, as is also seen from the bigger fluctuations in the $90^{th}$ percentile than the first household.

<!-- ## Choosing a threshold -->

<!-- A randomization test involves calculating a test statistic, randomly shuffling the data and calculating the test statistic several times to obtain a distribution of the test statistic. But we will use this procedure to obtain a threshold such that harmony pairs with a $wpd$ value higher than this threshold will only be considered significant. The process of choosing a threshold is described as follows: -->

<!-- to test if there is any interesting pattern captured by the harmonies, which essentially implies if $wpd$ is significantly different from zero. The percentages of times the $wpd$ obtained from the permuted data is greater than or equal to the observed $wpd$ is the p-value. The randomization test is described as follows: -->

<!-- <!-- We can remove the harmonies for which no interesting patterns are observed through a randomization permutation method. --> 
<!-- Essentially, the assumption is that under the null hypothesis, there is no difference in categories between the pair of cyclic granularities in the chosen harmony. This method is based on the generation of randomly chosen reassignments (permutations) of the data across different cyclic granularities and the computation of  -->
<!-- $wpd$ for each of these reassignments. -->

<!-- **Assumption:** random permutation of the data keeping the categories of the cyclic granularities constant. -->

<!-- Complete randomness in the measured variable indicates that the process follows a homogeneous underlying distribution over the whole time series, which essentially implies there is no interesting distinction across any different categories of the cyclic granularities. We can remove the harmonies for which no interesting patterns are observed through a randomization permutation method. Essentially, the assumption is that under the null hypothesis, there is no difference in categories between the pair of cyclic granularities in the chosen harmony. This method is based on the generation of randomly chosen reassignments (permutations) of the data across different cyclic granularities and the computation of  -->
<!-- $wpd$ for each of these reassignments. The percentages of times the theoretical distribution greater than or equal to the respective observed $wpd$ values are calculated and are used to obtain the P value. The procedure for the permutation test is: -->

<!-- **Assumption:** random permutation of the data keeping the categories of the cyclic granularities constant. -->

<!-- 14, 20, 31, 50\}$. For each of the $64$ panels, $ntimes = 500$ observations are drawn for each combination of the categories. The values of $wpd$ is obtained for each of the panels for the designs $D_{null}$, $D_{var_x}$, $D_{var_f}$ and $D_{var_{all}}$. $wpd_{threshold}$ is computed from all of these panels together and the number of times a harmony pair $(A, B) \in H_{N_C}$ is selected when in fact it was of the design $D_{null}$ is noted. This entire process is repeated for several null data sets to see the number of times any harmony pair $(A, B) \in H_{N_C}$ is selected under null. -->

<!-- ### Results -->

```{r varall-new}

```

```{r varx-new}

```

```{r varf-new}
```

```{r null-new}

```

```{r plot-all-new3, fig.cap = "$wpd$ values for the four illustrative designs are presented here. As expected, the value of $wpd$ under null design is the least (a), followed by (b, c and d). Moreover, the relative difference is interesting as c is closer to d than b, since the differences between x categories are up-weighed by design.", dependson= "null4by2"}

```

## Simulations {#sec:simulations}

### Behavior of raw and adjusted distance measures

\noindent _Simulation design_

$m=1$

\noindent Observations are generated from a N(0,1) distribution for a wide range of levels from very low to moderately high. $nx = \{2, 3, 5, 7, 9, 14, 17, 20, 24, 31, 42, 50\}$ is considered. $ntimes = 500$ observations are drawn for each combination of the categories, that is, for a panel with $nx=3$, $500$ observations are simulated for each of the categories. This design corresponds to $D_{null}$ as each combination of categories in a panel are drawn from the same distribution. Furthermore, the data is simulated for each of the categories $nsim=200$ times, so that the distribution of $wpd$ under $D_{null}$ could be observed. The values of $wpd$ is obtained for each of the panels. $wpd_{l, s}$ denotes the value of $wpd$ obtained for the $l^{th}$ panel and $s^{th}$ simulation.

$m=2$  
\noindent Similarly, observations are generated from a N(0,1) distribution for each combination of $nx$ and $nfacet$ from the following sets: $nx = nfacet = \{2, 3, 5, 7, 14, 20, 31, 50\}$. That is, data is being generated for each of the panels $(2, 2), (2, 3), (2, 5) \dots, (50, 31), (50, 50)$. For each of the $64$ panels, $ntimes = 500$ observations are drawn for each combination of the categories. That is, if we consider a $(2, 2)$ panel, $500$ observations are generated for each of the possible subsets, namely, $\{(1, 1), (1, 2), (2, 1), (2, 2)\}$.

\noindent _Results_

Figure \ref{fig:raw} shows that both the location and scale of the distributions change across panels. This is not desirable under $D_{null}$ as it would mean comparisons of $wpd$ values is not appropriate across different $nx$ and $nfacet$. Table \ref{tab:glm-tab} gives the summary of the generalized linear model to capture the relationship between $wpd_{raw}$ and number of comparisons. For example, the model considered for $m = 2$ is $wpd_{l,s} = 23.4 - 0.96*log(nx * nfacet)) + e$. The intercepts and slopes are similar independent of the starting distributions (see supplementary paper for details) and hence the coefficients are shown for the case when observations are drawn from a N(0, 1) distribution. Figure \ref{fig:dist-new-same-scale-link} shows the distribution of $wpd_{perm}$ and $wpd_{glm-scaled}$ in the same scale to show that a combination approach could be used for higher values of levels to alleviate the computational time of permutation approach.

```{r raw, fig.cap = "Distribution of $wpd_{raw}$ is plotted across different $nx$ and $nfacet$ categories under $D_{null}$ through density and rug plots. Both location (blue line) and scale (orange marks) of the distribution shifts for different panels. This is not desirable since under null design, the distribution  is not expected to capture any differences."}
```

```{r glm-tab}

```

```{r wpd-glm-dist}

```

```{r hist-qq-new, eval = FALSE}

```

```{r quadratic, eval=FALSE}
```

```{r norm, eval}
```

```{r dist-new-same-scale-link, fig.cap = "The distribution of $wpd_{perm}$ and $wpd_{glm-scaled}$ are overlaid to compare the location and scale across different $nx$ and $nfacet$. $wpd_{norm}$ takes the value of $wpd_{perm}$ for lower levels, and $wpd_{glm-scaled}$ for higher levels to to alleviate the problem of computational time in permutation approaches. This is possible as the distribution of the adjusted measure looks similar for both appraches for higher levels."}
```

### Choosing threshold

\noindent _Simulation design_

Observations are generated from a N(0,1) distribution for each combination of $nx$ and $nfacet$ from the following sets: $nx = \{3, 7, 14\}$ and $nfacet = \{2, 9, 10\}$. This would result in 9 panels, viz, $(3, 2), (3, 9), (3, 10), \dots, (14,9), (14, 10)$. Few experiments were conducted. In the first scenario, data for all panels are simulated using the null design  $D_{null}$. In other scenarios, data simulated from the panel $(14, 2)$ and $(3, 10)$ are considered under $D_{vary_{all}}$. Moreover, $\omega = \{0.5, 2, 5\}$ are considered to examine if the proposed test is able to capture subtle differences and non-subtle differences when we shift from the null design.

<!-- In the last scenario, we consider the panel $(3, 2), (7,9), (14, 10)$ to be under $D_{null}$, the panels $(7, 2), (14, 9)$ to be under $D_{var_f}$. $(14, 2), (3, 10)$ under $D_{var_x}$ and the rest under $D_{var_{null}}$. This is done to check if the consequent ranking procedure leads to designs like $D_{vary_{all}}$ to be chosen first followed by $D_{vary_{all}}$. We generate only one data set each for which these scenarios were simulated and consider this as the original data set. We generate $1000$ repetitions of this experiment with different seeds. -->

\noindent _Results_

For the first scenario, size of the test is obtained as $0.1$ with $wpd_{threshold99}$ as the threshold. This implies that the proportion of times a panel is rejected when it is under $D_{null}$ is $0.1$. The level of significance for each test is 1% (as a result of choosing $99^{th}$ percentile as the threshold) and we have $9$ multiple tests. Hence, it is reasonable
that the level of significance of the composite tests would be larger than the individual tests. We also compute the proportion of times a panel is rejected when it actually belongs to a non-null design. These proportions constitute to be the estimated size and power of the test.
Power can depend on many things like sample size, number of designs that deviate from the null and extent to which from the null. It is found that as we increase from low to high changes from the null distribution, the power increased. <!---The power also increases with increased sample size, as is expected.---> The results and graphics are included in details in the Supplementary paper.

```{r, eval = FALSE}
all_data <- read_rds("simulations/supplementary/test-hpc/data-agg/all_data.rds")

# computes size for 99 percentile threshold
data_summary <- all_data %>%
  mutate(sig = if_else(significance!=99, 0, 99)) %>% 
  group_by(seed_id) %>% 
  summarize(sumsig = sum(sig)) %>% 
  count(sumsig==0) %>% 
  mutate(p_value = n/sum(n)) %>% 
  slice(1)

# plots
ggplot() + 
  geom_histogram(data = all_data, aes(x = wpd))  + 
  geom_vline(data = data_summary, aes(xintercept = p_value),colour = "red") +
  facet_grid(x_levels~facet_levels)

```

### Environment

Simulation studies were carried out to study the behavior of $wpd$, build the normalization method as well as compare and evaluate different normalization approaches. R version 4.0.1 (2020-06-06) is used with the platform: x86_64-apple-darwin17.0 (64-bit) running under: macOS Mojave 10.14.6 and MonaRCH, which is a next-generation High Power Computing (HPC) Cluster, addressing the needs of the Monash HPC community.

## Application to residential smart meter dataset {#sec:application-wpd}

The smart meter data set for eight households in Melbourne has been utilized to see the use of $wpd$ proposed in the paper. The data has been cleaned to be a `tsibble` (@wang2020tsibble) containing half-hourly electricity consumption from Jul-2019 to Dec-2019 for each of the households, which is procured by them by downloading their data from the energy supplier/retailer. 
No behavioral pattern is likely to be discerned from the line graph of energy usage over the entire period, since the plot will have too many measurements squeezed in a linear representation. When we zoom into the linear representation of this series in Figure \ref{fig:dotplot-8} (b) for September, some patterns are visible in terms of peaks and troughs, but we do not know if they are regular or what is their period.
Electricity demand, in general, has a daily and weekly periodic pattern. However, it is not apparent from this view if all of these households have those patterns and in case they have if they are significant enough. Also, it is not clear if any other periodic patterns are present in any household which might have been hidden with this view. We start the analysis by choosing few harmonies, ranking them for each of these households, compare households to get more insights into what these rankings imply. Furthermore, the ranking and selection of significant harmonies is validated by analyzing the distribution of energy usage across significant harmonies.

<!-- if we are benefiting by moving from a linear to cyclic representation of time to . -->

```{r linear-scale-8, fig.cap = "Electricity demand for eight households is shown in different facets from Jul-19 to Dec-19 in Fig a and it has been zoomed in for Sep-19 in Fig b, where a week in Sep-19 has been highlighted. From the scales of Fig a, it is apparent that these households vary in consumption levels, but all their periodic patterns have been squeezed with this representation. In fig b, we can see some weekly patterns for the entire period and daily patterns for the highlighted week and there could be some other patterns visible in terms of peaks and troughs, but it is not clear if they are regular or significant.", fig.height=8, eval = FALSE}

```

_Choosing cyclic granularities of interest and removing clashes_

Let $v_{i, t}$ denote the electricity demand for $i^{th}$ household for time period $t$. The series $v_{i, t}$ is the linear granularity corresponding to half-hour since the interval of the tsibble is 30 minutes. We consider coarser linear granularities like hour, day, week and month from the commonly used Gregorian calendar. Considering $4$ linear granularities hour, day, week, month, the number of cyclic granularities is $N_C = (4 * 3/2) = 6$. We obtain cyclic granularities namely "hour_day", "hour_week", "hour_month", "day_ week", "day_month" and "week_month", read as "hour of the day", etc. Further, we add cyclic granularity day-type ("wknd wday") to capture weekend and weekday behavior. Thus, $7$ cyclic granularities are considered to be of interest. The set consisting of pairs of cyclic granularities ($C_{N_C}$) will have $7_{P_2}=42$ elements which could be analyzed for detecting possible periodicities. The set of possible harmonies $H_{N_C}$ from $C_{N_C}$ are chosen by removing clashes using procedures described in [@Gupta2021-hd]. Table \ref{tab:tab-rank-8}  shows $14$ harmony pairs that belong to $H_{N_C}$.

 <!-- The paper also shows how the number of harmonies could be identified among  by removing clashes. This section shows how we could refine the search further by only looking at significant harmonies. -->

_Selecting and Ranking harmonies for all households_

$wpd_{i}$ is computed on $v_{i, t}$ for all harmony pairs $\in H_{N_C}$ and for each households $i \in \{1, 2, \dots, 8\}$. The harmony pairs are then arranged in descending order and highlighted with `***`, `**` and `*` corresponding to the  $99^{th}$,  $95^{th}$ and $90^{th}$ percentile threshold. Table \ref{tab:tab-rank-8} shows the rank of the harmonies for different households. The rankings are different for different households, which is a reflection of their varied behaviors. Most importantly, there are at most $3$ harmonies that are significant for any household. This is a huge reduction in the number of potential harmonies to explore closely, starting from $42$.

_Detecting patterns not apparent from linear display_

\noindent Figure \ref{fig:dotplot-8} helps to compare households through the heatmap (a) across harmony pairs with the cyclic granularity mapped to x-axis and facet being plotted on the x-axis and y-axis of the heatmap. Here $dom$, $dow$, $wdwnd$ are abbreviations for day-of-month, day-of-week, weekday/weekend and so on. The colors represent the value of $wpd$. Darker cells correspond to more significant harmony pairs. Also, the ones with `*` corresponds to the ones above $wpd_{threshold95}$. 
Few observations that emphasizes patterns not discernible through (b) includes - (1) id 7 and 8 have the same significant harmonies despite having very different total energy usage, (2) id 6 and 7 differ in the sense that for id 6, the difference in patterns only during weekday/weekends, whereas for id 7, all or few other days of the week are also important. This might be due to their flexible work routines or different day-off, (3) there are no significant periodic patterns for id 5 when we fix the threshold to $wpd_{threshold95}$. Note that the $wpd$ values are computed over the entire range, but the linear display in (b) is zoomed only for September, with the major and minor x-axis corresponding to weeks and days respectively.

<!-- Households id 2 and 3 are similar, in terms of linear display and heatmap, which implies that similar periodic behavior can stem from households with different demographics. -->

```{r rank-household-mclapply-8}
```

```{r elec_select_harmony-8}
```

```{r dotplot-8, fig.cap = "An ensemble plot with a heatmap (a), line plot (b), parallel coordinate plot (c) to demonstrate energy behavior of the households in different ways. (b) is the zoomed-in raw demand series for September to highlight the repetitive patterns of energy demand. (a) shows $wpd$ values across harmonies with the x variable of harmony placed across x-axis and facet variable placed across y-axis. The darker the colour in (a), the higher the harmony is ranked. Visualizing harmonies through (a) allows to view the significant cyclic granularities along the x-axis, facet or both for each household. For eg, ids 7 and 8 have significant patterns across (hod, dow) and (dow, hod), which was not apparent through (a). (c) is useful for comparing households across harmonies, for eg, for the harmony (dow-hod), ids 1 and 7 have the least and highest wpd respectively.", fig.height = 7}
```

```{r tab-rank-8}
```

_Comparing households and validating rank of harmonies_

According to Figure \ref{fig:dotplot-8}(c), for the harmony pair (dow-hod), household id 7 has the greatest value of $wpd$, while id 1 has the least. Also, from table \ref{tab:tab-rank-8} it could be seen that the harmony pair $(dow, hod)$ is important for id 7, however it has been labeled as an inconsequential pair for id 1. The distribution of energy demand for both of these households, with $dow$ as the facet and $hod$ on the x-axis, may help explain the choice. Figure \ref{fig:gravitas-plot-8} demonstrates that for id 7, the median (black) and quartile deviation (orange) of energy consumption fluctuates throughout for most hours of the day and day of the week, while for id 1, daily patterns are consistent within weekdays and weekends. As a result, for id 1, it is more appropriate to examine the distributional difference solely across (dow, wdwnd), which has been rated higher in Table \ref{tab:tab-rank-8}.

<!-- Although the differences might seem significant at first, with closer inspection it could be seen that the scale of the demand is lower in this case and hence the differences are not large enough to cross the threshold for significance. -->

<!-- Figure \ref{fig:gravitas-plot-8} is used to show -->
<!-- if this selection and ranking of harmony pairs makes sense for this household. Panel a) of  Figure \ref{fig:gravitas-plot-8} shows the distribution of energy demand with weekday/weekend as the x-axis and hour-of-day as the facets and helps to compare the weekend/weekday patterns for different hours of the day. It could be observed that the difference between weekend and weekday is the highest from 15 to 19 hours of the day. Panel b) shows the distribution of energy demand with the variables swapped and helps to compare the daily patterns within weekday and weekend. It could be observed that the daily pattern is similar for weekdays and weekends with a morning and evening peak. However, the difference between morning and evening peaks are higher for weekends. Since $wpd$ is designed to put more weightage on within-facet differences for $\lambda>0.5$, it makes sense that the pair (hod, wdwnd) has been ranked higher than (wdwnd, hod). Panel c) shows the distribution of energy demand with weekday/weekend as the x-axis and week-of-month as the facets. Although the differences might seem significant at first, with closer inspection it could be seen that the scale of the demand is lower in this case and hence the differences are not large enough to cross the threshold for significance. -->

```{r gravitas-plot-8, message = FALSE, warning=FALSE,fig.cap="Comparing distribution of energy demand shown for household id 1 (a) and 7 (b) across hod in x-axis and dow in facets through quantile area plots. The value of $wpd$ in Table 3 suggests that the harmony pair (dow, hod) is significant for household id 7, but not for 1. This implies that distributional differences are captured more by this harmony for id 7, which is apparent from this display with more fluctuations across median and 75th percentile for different hours of the day and day of week. For id 1, patterns look similar within weekdays and weekends. Here, the median is represented by the black line, the orange area corresponds to quartile deviation and the green area corresponds to area between $10^{th}$ and $90^{th}$ quantile."}

```

<!-- From Figure \ref{fig:gravitas-plot-8-id5}, it could further be observed that id5 has only one significant harmony $(hod, dow)$. Apparently, $(hod, wnd/wday)$ which is an important harmony for most households is not important for this one. -->

## Discussion

Exploratory data analysis involves many iterations of finding and summarizing patterns. With temporal data available at finer scales, exploring time series has become overwhelming with so many possible granularities to explore. A common solution is to aggregate and look at the patterns across usual granularities like hour-of-day or day-of-week, but there is no way to know the “interesting” granularities a priori. A huge number of displays need to be analyzed or we might end up missing informative granularities. This work refines the search of informative granularities by identifying those for which the differences between the displayed distributions are greatest and rating them in order of importance of capturing maximum variation.

The significant granularities across different datasets (individuals/subjects) do not imply similar patterns across different datasets. They simply mean that maximum distributional differences are being captured across those granularities. A future direction of work is to be able to explore and compare many individuals/subjects together for similar patterns across significant granularities.

<!-- patterns are interesting and others are not. The subjects for which behaviors across a group of cyclic granularities are similar, could be grouped together as their periodic behavior is similar.  to group subjects  -->

## Acknowledgments {-}

The Australian authors thank the ARC Centre of Excellence for Mathematical and Statistical Frontiers [(ACEMS)](https://acems.org.au/home) for supporting this research. Sayani Gupta was partially funded by [Data61 CSIRO](https://data61.csiro.au/) during her PhD. The Github repository,  [github.com/Sayani07/paper-hakear](https://github.com/Sayani07/paper-gravitas), contains all materials required to reproduce this article and the code is also available online in the supplemental materials. This article was created with R [@R-language], `knitr` [@knitr2015, @R-knitr] and `rmarkdown` [@rmarkdown2018, @R-rmarkdown]. Graphics are produced with @Wickham2009pk.

## Supplementary Materials

**Data and scripts:** Data sets and R code to reproduce all figures in this article (main.R).

\noindent
**Simulation results and scripts:** All simulation table, graphics and and R code to reproduce it (paper-supplementary.pdf, paper-supplementary.Rmd)

\noindent
**R-package:** The open-source R [@R-language] package `hakear` is available on Github (https://github.com/Sayani07/hakear) to implement ideas presented in this paper.

### 2021-11-04 Rmd/04-gracsr.Rmd
Rmd folder update

# Data representation, visual and analytical techniques for demystifying temporal missing data {#ch:mists}

### 2021-11-04 data/hakear/simulations/norm/null_design_quantrans/R-ind/norm_pairwise_Gamma01.R
all data delete

@@ -1,0 +1,63 @@
+#This script calculates raw mmpd for each simulation scenario
+##Read Simulation Table
+
+.libPaths(c("~/R/libs", .libPaths()))
+library(distributional)
+library(readr)
+library(drake)
+library(tidyverse)
+library(hakear)
+
+set.seed(9999)
+
+nsim = 200
+# change path while running it on HPC
+# simtable<-read_csv(here::here('simulations/null/sim_table.csv'))
+simtable<-read_csv('../../../sim_table/sim_table.csv')
+
+### Extract flags from simulation scenario
+
+#scen<- 2 #If running within R uncomment this.  This will only run first scenario
+scen<-as.numeric(commandArgs()[[6]]) # If running batch job uncomment this
+
+simj<-simtable[scen,] #Extract row of table
+nfacetj<-simj$nfacet # Which nfacet level
+nxj<-simj$nx #Which nx level
+
+#create data for each row for null normal
+
+sim_null_normal = function(nxj, nfacetj){
+  rep(distributional::dist_gamma(0.5, 1), 
+      times = nxj*nfacetj)
+}
+
+sim_panel_data = 
+  hakear::sim_panel(nx = nxj,
+                    nfacet = nfacetj, 
+                    ntimes = 500, 
+                    sim_dist = sim_null_normal) %>% 
+  unnest(c(data)) %>% ungroup()
+
+set.seed(1111)
+
+raw_dist <- map(seq_len(nsim), function(i)
+{
+  new_sim_data = sample(sim_panel_data$sim_data, 
+                        size = nrow(sim_panel_data))
+  new_data = sim_panel_data %>% 
+    select(-sim_data) %>% 
+    mutate(sim_data = new_sim_data)
+  
+  # for creating one raw mmpd
+  raw_mmpd = compute_wpd_norm(new_data, 
+                                  gran_x = "id_x",
+                                  gran_facet = "id_facet",
+                                  response = sim_data) %>% 
+    as_tibble() %>% mutate(perm_id = i)
+  
+}) %>% bind_rows()
+
+saveRDS(raw_dist,
+        paste0('../data-ind/wpd_Gamma01/',
+               nxj,'_',
+               nfacetj,'_wpd.rds'))

### 2021-11-04 data/hakear/simulations/norm/null_design_quantrans/job/job-norm-N05.sh

@@ -1,0 +1,12 @@
+#! /bin/bash
+#SBATCH --job-name=norm_pairwise_N05
+#SBATCH --output=trace_%a.out
+#SBATCH --error=error_%a.err
+#SBATCH --mem-per-cpu=4096
+#SBATCH --time=168:00:00
+#SBATCH --array=1-64
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --export=NONE
+module load R/4.0.0-openblas
+Rscript ../R-ind/norm_pairwise_N05.R $SLURM_ARRAY_TASK_ID 

### 2021-11-04 data/hakear/simulations/norm/null_design_quantrans/job/job-norm-N51.sh

@@ -1,0 +1,12 @@
+#! /bin/bash
+#SBATCH --job-name=norm_pairwise_N51
+#SBATCH --output=trace_%a.out
+#SBATCH --error=error_%a.err
+#SBATCH --mem-per-cpu=4096
+#SBATCH --time=168:00:00
+#SBATCH --array=1-64
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --export=NONE
+module load R/4.0.0-openblas
+Rscript ../R-ind/norm_pairwise_N51.R $SLURM_ARRAY_TASK_ID 

### 2021-11-04 data/hakear/simulations/norm/null_design_quantrans_nperm/R-agg/aggregate01.R

@@ -1,0 +1,56 @@
+# contains function to aggregate data each for from N(0,1), N(0,5), N(5,1), N(5,5), Gamma(0.5, 1), Gamma(2, 1)
+#  aggregated data from the file simulations/results/norm/folder_name and tidy them up
+library(tidyverse)
+library(here)
+library(readr)
+library(rlang)
+
+aggregate01 <-  function(folder_name){
+  
+all_files = list.files(path = paste0("simulations/norm/null_design_quantrans_nperm/data-ind/", folder_name), 
+                       pattern = ".rds")
+
+names_levels <- map_dfr(all_files, 
+                        function(x){
+                          z = str_split(str_remove(x, "_dist.rds"), "_") %>% 
+                            unlist()
+                          bind_cols(nx = as.numeric(z[1]),
+                                    nfacet = as.numeric(z[2]))
+                        })
+
+# len_file = read_rds(("simulations/results/norm/tuning_param/2_2_tuning_param.rds"))
+# 
+# names_rep <- names_levels %>% slice(rep(1:n(), each = nrow(len_file)))
+
+  
+all_files_path <- paste0("simulations/norm/null_design_quantrans_nperm/data-ind/",folder_name,"/",
+         all_files)  
+  
+
+all_data <- lapply(1:length(all_files_path), function(x){
+    
+  data = all_files_path %>% magrittr::extract2(x) %>% 
+    readRDS()
+  
+  names = names_levels %>% magrittr::extract(x,)
+  names_rep =   names %>% slice(rep(1:n(), each = nrow(data)))
+  bind_cols(names_rep, data)
+  }) %>% bind_rows() %>% 
+    arrange(nfacet, nx)
+    
+write_rds(all_data, paste0("simulations/norm/null_design_quantrans_nperm/data-agg/all_data_", folder_name, ".rds"))
+}
+
+
+# use functions for all folders
+
+aggregate01(folder_name = "wpd_N01")
+aggregate01(folder_name = "wpd_N05")
+aggregate01(folder_name = "wpd_N51")
+aggregate01(folder_name = "wpd_N55")
+
+
+aggregate01(folder_name = "wpd_Gamma01")
+aggregate01(folder_name = "wpd_Gamma21")
+
+

### 2021-11-04 data/hakear/simulations/norm/null_design_quantrans_nperm/R-agg/makegraph02.R

@@ -1,0 +1,119 @@
+library(here)
+library(readr)
+library(tidyverse)
+# run for  norm MMPD files aggregation
+#all_data <- read_rds("simulations/result_report/all_data_ norm.rds")
+# run for  norm max pairwise distances files aggregation
+
+
+makegraph02 <- function(folder_name){
+  
+all_data <- read_rds(paste0("simulations/norm/null_design_quantrans_nperm/data-agg/all_data_", folder_name, ".rds"))
+  
+
+nxbyfacet_density <- all_data %>% 
+  ggplot(aes(x = value)) + 
+  geom_density(fill = "blue") +
+  facet_grid(nx~nfacet,
+             labeller = "label_both",
+             scales = "free_y") + 
+  xlab("wpd")
+
+ggsave(nxbyfacet_density, filename = paste0("simulations/norm/null_design_quantrans_nperm/figs/", "nxbyfacet_density_", folder_name,".png"))
+
+
+nxbyfacet_ridge <- all_data %>% 
+  ggplot(aes(x = value, y = as.factor(nx))) +
+  ggridges::geom_density_ridges() +
+  facet_wrap(~nfacet, labeller = "label_both", nrow = 2) + 
+  xlab("mmpd") +
+  ylab("nx")
+
+ggsave(nxbyfacet_ridge, filename = paste0("simulations/norm/null_design_quantrans_nperm/figs/", "nxbyfacet_ridge_", folder_name,".png"))
+
+
+nfacetbynx_ridge <- all_data %>% 
+  ggplot(aes(x = value, y = as.factor(nfacet))) +
+  ggridges::geom_density_ridges() +
+  facet_wrap(~nx, labeller = "label_both", nrow = 2) + 
+  xlab("mmpd") +
+  ylab("nfacet")
+
+ggsave(nfacetbynx_ridge, filename = paste0("simulations/norm/null_design_quantrans_nperm/figs/", "nfacetbynx_ridge_", folder_name,".png"))
+}
+
+makegraph02(folder_name = "wpd_N01")
+makegraph02(folder_name = "wpd_N05")
+makegraph02(folder_name = "wpd_N51")
+makegraph02(folder_name = "wpd_N55")
+makegraph02(folder_name = "wpd_Gamma01")
+makegraph02(folder_name = "wpd_Gamma21")
+
+
+
+# N(0,1) and N(5,1) together as they will go in the paper
+
+normal_0_1 <- aggregate01(folder_name = "wpd_N01")
[…]

### 2021-11-04 data/hakear/simulations/norm/null_design_quantrans_nperm/R-ind/norm_pairwise_N51.R

@@ -1,0 +1,63 @@
+#This script calculates raw mmpd for each simulation scenario
+##Read Simulation Table
+
+.libPaths(c("~/R/libs", .libPaths()))
+library(distributional)
+library(readr)
+library(drake)
+library(tidyverse)
+library(hakear)
+library(parallel)
+set.seed(9999)
+
+nsim = 200
+# change path while running it on HPC
+# simtable<-read_csv(here::here('simulations/null/sim_table.csv'))
+simtable<-read_csv('../../../sim_table/sim_table.csv')
+
+### Extract flags from simulation scenario
+
+#scen<- 2 #If running within R uncomment this.  This will only run first scenario
+scen<-as.numeric(commandArgs()[[6]]) # If running batch job uncomment this
+
+simj<-simtable[scen,] #Extract row of table
+nfacetj<-simj$nfacet # Which nfacet level
+nxj<-simj$nx #Which nx level
+
+#create data for each row for null normal
+
+sim_null_normal = function(nxj, nfacetj){
+  rep(distributional::dist_normal(5, 1), 
+      times = nxj*nfacetj)
+}
+
+sim_panel_data = 
+  hakear::sim_panel(nx = nxj,
+                    nfacet = nfacetj, 
+                    ntimes = 500, 
+                    sim_dist = sim_null_normal) %>% 
+  unnest(c(data)) %>% ungroup()
+
+set.seed(1111)
+
+raw_dist <- map(seq_len(nsim), function(i)
+{
+  new_sim_data = sample(sim_panel_data$sim_data, 
+                        size = nrow(sim_panel_data))
+  new_data = sim_panel_data %>% 
+    select(-sim_data) %>% 
+    mutate(sim_data = new_sim_data)
+  
+  # for creating one raw mmpd
+  raw_mmpd = compute_pairwise_norm(new_data, 
+                                  gran_x = "id_x",
+                                  gran_facet = "id_facet",
+                                  response = sim_data) %>% 
+    as_tibble() %>% mutate(perm_id = i)
+  
+}) %>% bind_rows()
+
+saveRDS(raw_dist,
+        paste0('../data-ind/wpd_N51/',
+               nxj,'_',
+               nfacetj,'_wpd.rds'))

### 2021-11-04 data/hakear/simulations/norm/null_design_quantrans_nperm/job/job-norm-Gamma21.sh

@@ -1,0 +1,12 @@
+#! /bin/bash
+#SBATCH --job-name=norm_pairwise_Gamma21
+#SBATCH --output=trace_%a.out
+#SBATCH --error=error_%a.err
+#SBATCH --mem-per-cpu=4096
+#SBATCH --time=168:00:00
+#SBATCH --array=1-64
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --export=NONE
+module load R/4.0.0-openblas
+Rscript ../R-ind/norm_pairwise_Gamma21.R $SLURM_ARRAY_TASK_ID 

### 2021-11-04 data/hakear/simulations/norm/null_design_quantrans_nperm/job/job-norm-N01.sh

@@ -1,0 +1,12 @@
+#! /bin/bash
+#SBATCH --job-name=norm_pairwise_N01
+#SBATCH --output=trace_%a.out
+#SBATCH --error=error_%a.err
+#SBATCH --mem-per-cpu=4096
+#SBATCH --time=168:00:00
+#SBATCH --array=1-64
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --export=NONE
+module load R/4.0.0-openblas
+Rscript ../R-ind/norm_pairwise_N01.R $SLURM_ARRAY_TASK_ID 

### 2021-11-04 data/hakear/simulations/raw/null_design_quantrans/R-agg/makegraph02.R

@@ -1,0 +1,169 @@
+library(here)
+library(readr)
+library(tidyverse)
+# run for  norm MMPD files aggregation
+#all_data <- read_rds("simulations/result_report/all_data_ norm.rds")
+# run for  norm max pairwise distances files aggregation
+
+
+makegraph02 <- function(folder_name){
+  
+all_data <- read_rds(paste0("simulations/raw/null_design_quantrans/data-agg/all_data_wpd_", folder_name, ".rds"))
+  
+summary_data <- all_data %>% 
+  group_by(nx, nfacet) %>% 
+  summarise(mean = mean(value))
+
+nxbyfacet_density <- all_data %>% 
+  ggplot(aes(x = value)) + 
+  geom_density(fill = "#999999") +
+  facet_grid(nx~nfacet,
+             labeller = "label_both") + 
+  xlab("raw wpd") +
+  geom_vline(data = summary_data, aes(xintercept  = mean), color = "#0072B2") +
+  geom_rug(sides = "b", colour = "#D55E00") + 
+ scale_x_continuous(breaks = scales::breaks_extended(3.5)) +
+  theme_bw() +
+  theme(panel.grid.major.x = element_blank()) 
+
+ggsave(nxbyfacet_density, filename = paste0("simulations/raw/null_design_quantrans/figs/", "nxbyfacet_density_wpd_", folder_name,".png"))
+
+
+nxbyfacet_ridge <- all_data %>% 
+  ggplot(aes(x = value, y = as.factor(nx))) +
+  ggridges::geom_density_ridges() +
+  facet_wrap(~nfacet, labeller = "label_both", nrow = 2) + 
+  xlab("raw wpd") +
+  ylab("nx") +
+  scale_x_continuous(breaks = scales::breaks_extended(4)) +
+  theme_bw()
+
+ggsave(nxbyfacet_ridge, filename = paste0("simulations/raw/null_design_quantrans/figs/", "nxbyfacet_ridge_wpd_", folder_name,".png"))
+
+
+nfacetbynx_ridge <- all_data %>% 
+  ggplot(aes(x = value, y = as.factor(nfacet))) +
+  ggridges::geom_density_ridges() +
+  facet_wrap(~nx, labeller = "label_both", nrow = 2) + 
+  xlab("raw wpd") +
+  ylab("nfacet") +
+  scale_x_continuous(breaks = scales::breaks_extended(4)) +
+  theme_bw()
+
+
[…]

### 2021-11-04 data/hakear/simulations/raw/null_design_quantrans/R-ind/raw_pairwise_Gamma21.R

@@ -1,0 +1,63 @@
+#This script calculates raw mmpd for each simulation scenario
+##Read Simulation Table
+
+.libPaths(c("~/R/libs", .libPaths()))
+library(distributional)
+library(readr)
+library(drake)
+library(tidyverse)
+library(hakear)
+
+set.seed(9999)
+
+nsim = 200
+# change path while running it on HPC
+# simtable<-read_csv(here::here('simulations/null/sim_table.csv'))
+simtable<-read_csv('../../../sim_table/sim_table.csv')
+
+### Extract flags from simulation scenario
+
+#scen<- 2 #If running within R uncomment this.  This will only run first scenario
+scen<-as.numeric(commandArgs()[[6]]) # If running batch job uncomment this
+
+simj<-simtable[scen,] #Extract row of table
+nfacetj<-simj$nfacet # Which nfacet level
+nxj<-simj$nx #Which nx level
+
+#create data for each row for null normal
+
+sim_null_normal = function(nxj, nfacetj){
+  rep(distributional::dist_gamma(2, 1), 
+      times = nxj*nfacetj)
+}
+
+sim_panel_data = 
+  hakear::sim_panel(nx = nxj,
+                    nfacet = nfacetj, 
+                    ntimes = 500, 
+                    sim_dist = sim_null_normal) %>% 
+  unnest(c(data)) %>% ungroup()
+
+set.seed(1111)
+
+raw_dist <- map(seq_len(nsim), function(i)
+{
+  new_sim_data = sample(sim_panel_data$sim_data, 
+                        size = nrow(sim_panel_data))
+  new_data = sim_panel_data %>% 
+    select(-sim_data) %>% 
+    mutate(sim_data = new_sim_data)
+  
+  # for creating one raw mmpd
+  raw_mmpd = compute_pairwise_max(new_data, 
+                                  gran_x = "id_x",
+                                  gran_facet = "id_facet",
+                                  response = sim_data) %>% 
+    as_tibble() %>% mutate(perm_id = i)
+  
+}) %>% bind_rows()
+
+saveRDS(raw_dist,
+        paste0('../data-ind/wpd_Gamma21/',
+               nxj,'_',
+               nfacetj,'_wpd.rds')) 

### 2021-11-04 data/hakear/simulations/raw/null_design_quantrans/R-ind/raw_pairwise_N01.R

@@ -1,0 +1,63 @@
+#This script calculates raw mmpd for each simulation scenario
+##Read Simulation Table
+
+.libPaths(c("~/R/libs", .libPaths()))
+library(distributional)
+library(readr)
+library(drake)
+library(tidyverse)
+library(hakear)
+
+set.seed(9999)
+
+nsim = 200
+# change path while running it on HPC
+# simtable<-read_csv(here::here('simulations/null/sim_table.csv'))
+simtable<-read_csv('../../../sim_table/sim_table.csv')
+
+### Extract flags from simulation scenario
+
+#scen<- 2 #If running within R uncomment this.  This will only run first scenario
+scen<-as.numeric(commandArgs()[[6]]) # If running batch job uncomment this
+
+simj<-simtable[scen,] #Extract row of table
+nfacetj<-simj$nfacet # Which nfacet level
+nxj<-simj$nx #Which nx level
+
+#create data for each row for null normal
+
+sim_null_normal = function(nxj, nfacetj){
+  rep(distributional::dist_normal(0, 1), 
+      times = nxj*nfacetj)
+}
+
+sim_panel_data = 
+  hakear::sim_panel(nx = nxj,
+                    nfacet = nfacetj, 
+                    ntimes = 500, 
+                    sim_dist = sim_null_normal) %>% 
+  unnest(c(data)) %>% ungroup()
+
+set.seed(1111)
+
+  raw_dist <- map(seq_len(nsim), function(i)
+{
+  new_sim_data = sample(sim_panel_data$sim_data, 
+                        size = nrow(sim_panel_data))
+  new_data = sim_panel_data %>% 
+    select(-sim_data) %>% 
+    mutate(sim_data = new_sim_data)
+  
+  # for creating one raw mmpd
+  raw_mmpd = compute_pairwise_max(new_data, 
+                                 gran_x = "id_x",
+                                 gran_facet = "id_facet",
+                                 response = sim_data) %>% 
+    as_tibble() %>% mutate(perm_id = i)
+  
+}) %>% bind_rows()
+
+saveRDS(raw_dist,
+          paste0('../data-ind/wpd_N01/',
+                 nxj,'_',
+                 nfacetj,'_wpd.rds'))

### 2021-11-04 data/hakear/simulations/raw/null_design_quantrans/R-ind/raw_pairwise_N05.R

@@ -1,0 +1,65 @@
+#This script calculates raw mmpd for each simulation scenario
+##Read Simulation Table
+
+.libPaths(c("~/R/libs", .libPaths()))
+library(distributional)
+library(readr)
+library(drake)
+library(tidyverse)
+library(hakear)
+
+set.seed(9999)
+
+nsim = 200
+# change path while running it on HPC
+# simtable<-read_csv(here::here('simulations/null/sim_table.csv'))
+simtable<-read_csv('../../../sim_table/sim_table.csv')
+
+### Extract flags from simulation scenario
+
+#scen<- 2 #If running within R uncomment this.  This will only run first scenario
+scen<-as.numeric(commandArgs()[[6]]) # If running batch job uncomment this
+
+simj<-simtable[scen,] #Extract row of table
+nfacetj<-simj$nfacet # Which nfacet level
+nxj<-simj$nx #Which nx level
+
+#create data for each row for null normal
+
+sim_null_normal = function(nxj, nfacetj){
+  rep(distributional::dist_normal(0, 5), 
+      times = nxj*nfacetj)
+}
+
+sim_panel_data = 
+  hakear::sim_panel(nx = nxj,
+                    nfacet = nfacetj, 
+                    ntimes = 500, 
+                    sim_dist = sim_null_normal) %>% 
+  unnest(c(data)) %>% ungroup()
+
+set.seed(1111)
+
+raw_dist <- map(seq_len(nsim), function(i)
+{
+  new_sim_data = sample(sim_panel_data$sim_data, 
+                        size = nrow(sim_panel_data))
+  new_data = sim_panel_data %>% 
+    select(-sim_data) %>% 
+    mutate(sim_data = new_sim_data)
+  
+  # for creating one raw mmpd
+  raw_mmpd = compute_pairwise_max(new_data, 
+                                  gran_x = "id_x",
+                                  gran_facet = "id_facet",
+                                  response = sim_data) %>% 
+    as_tibble() %>% mutate(perm_id = i)
+  
+}) %>% bind_rows()
+
+saveRDS(raw_dist,
+        paste0('../data-ind/wpd_N05/',
+               nxj,'_',
+               nfacetj,'_wpd.rds'))
+
+

### 2021-11-04 data/hakear/simulations/raw/null_design_scale/R-ind/raw_pairwise_N01.R

@@ -1,0 +1,64 @@
+#This script calculates raw mmpd for each simulation scenario
+##Read Simulation Table
+
+.libPaths(c("~/R/libs", .libPaths()))
+library(distributional)
+library(readr)
+library(drake)
+library(tidyverse)
+library(hakear)
+
+set.seed(9999)
+
+nsim = 200
+# change path while running it on HPC
+# simtable<-read_csv(here::here('simulations/null/sim_table.csv'))
+simtable<-read_csv('sim_table.csv')
+
+### Extract flags from simulation scenario
+
+#scen<- 2 #If running within R uncomment this.  This will only run first scenario
+scen<-as.numeric(commandArgs()[[6]]) # If running batch job uncomment this
+
+simj<-simtable[scen,] #Extract row of table
+nfacetj<-simj$nfacet # Which nfacet level
+nxj<-simj$nx #Which nx level
+
+#create data for each row for null normal
+
+sim_null_normal = function(nxj, nfacetj){
+  rep(distributional::dist_normal(0, 1), 
+      times = nxj*nfacetj)
+}
+
+sim_panel_data = 
+  hakear::sim_panel(nx = nxj,
+                    nfacet = nfacetj, 
+                    ntimes = 500, 
+                    sim_dist = sim_null_normal) %>% 
+  unnest(c(data)) %>% ungroup()
+
+set.seed(1111)
+
+  raw_dist <- map(seq_len(nsim), function(i)
+{
+  new_sim_data = sample(sim_panel_data$sim_data, 
+                        size = nrow(sim_panel_data))
+  new_data = sim_panel_data %>% 
+    select(-sim_data) %>% 
+    mutate(sim_data = new_sim_data)
+  
+  # for creating one raw mmpd
+  raw_mmpd = compute_pairwise_max(new_data, 
+                                 gran_x = "id_x",
+                                 gran_facet = "id_facet",
+                                 response = sim_data) %>% 
+    as_tibble() %>% mutate(perm_id = i)
+  
+}) %>% bind_rows()
+
+saveRDS(raw_dist, paste0('../results/raw/wpd_N01/',
+                         nxj,'_',
+                         nfacetj,'_wpd_N01.rds'))
+
+

### 2021-11-04 data/hakear/simulations/raw/null_design_scale/job/job-raw-N51.sh

@@ -1,0 +1,12 @@
+#! /bin/bash
+#SBATCH --job-name=raw_pairwise_N51
+#SBATCH --output=trace_%a.out
+#SBATCH --error=error_%a.err
+#SBATCH --mem-per-cpu=4096
+#SBATCH --time=168:00:00
+#SBATCH --array=1-64
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --export=NONE
+module load R/4.0.0-openblas
+Rscript raw_pairwise_N51.R $SLURM_ARRAY_TASK_ID 

### 2021-11-04 data/hakear/simulations/supplementary/R-ind/wpd_pairwise_N01.R

@@ -1,0 +1,156 @@
+#This script calculates raw mmpd for each simulation scenario
+##Read Simulation Table
+
+.libPaths(c("~/R/libs", .libPaths()))
+library(distributional)
+library(readr)
+library(drake)
+library(tidyverse)
+library(hakear)
+
+# seed has been set later in the code when sampling part comes - to be included as a parameter later at least for two more seeds results hold true
+
+nsim = 200
+# change path while running it on HPC
+simtable<-read_csv(here::here('simulations/supplementary/sim_table/sim_table.csv'))
+
+### Extract flags from simulation scenario
+
+#scen<- 28800 #If running within R uncomment this.  This will only run first scenario
+scen<-as.numeric(commandArgs()[[6]]) # If running batch job uncomment this
+
+simj<-simtable[scen,] #Extract row of table
+nfacetj<-simj$nfacet # Which nfacet level
+nxj<-simj$nx #Which nx level
+wj <- simj$w
+designj <- simj$design
+npermj <- simj$nperm
+ntimesj <- simj$ntimes
+distj <- simj$dist
+seedj <- simj$seed
+typej <- simj$type
+
+
+if(designj=="null")
+{
+  sim_function <-  function(nxj, nfacetj, mean, sd, w = 0){
+    rep(distributional::dist_normal(mu = mean, sigma = sd), 
+        times = nxj*nfacetj)
+  }
+}
+if(designj=="vary_x")
+{
+  sim_function <- function(nx, nfacet, mean, sd, w)
+  {
+    rep(dist_normal((mean + seq(0, nx-1, by  = 1)*w), sd), nfacet)
+  }
+  
+}
+if(designj=="vary_f")
+{
+  sim_function <- function(nx, nfacet, mean, sd, w)
+  {
+    rep(dist_normal((mean + seq(0, nfacet-1, by  = 1)*w), sd), each = nx)
+  }
+  
+}
+if(designj=="vary_all"){
+  
+  sim_function <- function(nx, nfacet, mean, sd, w)
+  {
+    dist_normal((mean + seq(0, (nx*nfacet - 1), by  = 1)*w), sd)
+  }
+  
+  
+}
+
+#create data for each row for null normal
+
+set.seed(seedj)
+
+
+sim_panel_data = 
+  hakear::sim_panel(nx = nxj,
+                    nfacet = nfacetj, 
+                    ntimes = ntimesj, 
[…]

### 2021-11-04 data/hakear/simulations/supplementary/job/job-raw-N01.sh

@@ -1,0 +1,12 @@
+#! /bin/bash
+#SBATCH --job-name=raw_pairwise_N01
+#SBATCH --output=trace_%a.out
+#SBATCH --error=error_%a.err
+#SBATCH --mem-per-cpu=4096
+#SBATCH --time=168:00:00
+#SBATCH --array=1-14400
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --export=NONE
+module load R/4.0.0-openblas
+Rscript ../R-ind/wpd_pairwise_N01.R $SLURM_ARRAY_TASK_ID 

### 2021-11-04 data/hakear/simulations/supplementary/job/job_1-1000.sh

@@ -1,0 +1,12 @@
+#! /bin/bash
+#SBATCH --job-name=raw_pairwise_N01
+#SBATCH --output=trace_%a.out
+#SBATCH --error=error_%a.err
+#SBATCH --mem-per-cpu=4096
+#SBATCH --time=168:00:00
+#SBATCH --array=1-1000
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --export=NONE
+module load R/4.0.0-openblas
+Rscript ../R-ind/wpd_pairwise_N01.R $SLURM_ARRAY_TASK_ID 

### 2021-11-04 data/hakear/simulations/supplementary/one-gran/job-power-one.sh

@@ -1,0 +1,12 @@
+#! /bin/bash
+#SBATCH --job-name=power-one
+#SBATCH --output=trace_%a.out
+#SBATCH --error=error_%a.err
+#SBATCH --mem-per-cpu=4096
+#SBATCH --time=24:00:00
+#SBATCH --array=1-12000
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --export=NONE
+module load R/4.0.5
+Rscript ../R-ind/power-one.R $SLURM_ARRAY_TASK_ID 

### 2021-11-04 data/hakear/simulations/supplementary/one-gran/norm/null_design_quantrans_nperm/R-ind/norm_pairwise_N01.R

@@ -1,0 +1,65 @@
+
+#This script calculates normalised wpd for each simulation scenario
+##Read Simulation Table
+
+.libPaths(c("~/R/libs", .libPaths()))
+library(distributional)
+library(readr)
+library(drake)
+library(tidyverse)
+library(hakear)
+
+set.seed(9999)
+
+nsim = 200
+simtable<-read_csv('../../../sim_table.csv')
+# change path while running locally
+#simtable<-read_csv('simulations/supplementary/one-gran/sim_table.csv')
+
+### Extract flags from simulation scenario
+
+#scen<- 2 #If running within R uncomment this.  This will only run first scenario
+scen<-as.numeric(commandArgs()[[6]]) # If running batch job uncomment this
+
+simj<-simtable[scen,] #Extract row of table
+nfacetj<-simj$nfacet # Which nfacet level
+nxj<-simj$nx #Which nx level
+
+#create data for each row for null normal
+
+sim_null_normal = function(nxj, nfacetj){
+  rep(distributional::dist_normal(0, 1), 
+      times = nxj*nfacetj)
+}
+
+sim_panel_data = 
+  hakear::sim_panel(nx = nxj,
+                    nfacet = nfacetj, 
+                    ntimes = 500, 
+                    sim_dist = sim_null_normal(nxj, nfacetj)) %>% 
+  unnest(c(data)) %>% ungroup()
+
+set.seed(1111)
+
+  norm_dist <- map(seq_len(nsim), function(i)
+{
+  new_sim_data = sample(sim_panel_data$sim_data, 
+                        size = nrow(sim_panel_data))
+  new_data = sim_panel_data %>% 
+    select(-sim_data) %>% 
+    mutate(sim_data = new_sim_data)
+  
+  # for creating one raw mmpd
+  raw_mmpd = compute_pairwise_norm(new_data, 
+                                 gran_x = "id_x",
+                                 gran_facet = "id_facet",
+                                 response = sim_data) %>% 
+    as_tibble() %>% mutate(perm_id = i)
+  
+}) %>% bind_rows()
+
+saveRDS(norm_dist,
+          paste0('../data-ind/wpd_N01/',
+                 nxj,'_',
+                 nfacetj,'_wpd.rds'))
+  

### 2021-11-04 data/hakear/simulations/supplementary/one-gran/norm/null_design_quantrans_nperm/job/job-norm-N01.sh

@@ -1,0 +1,12 @@
+#! /bin/bash
+#SBATCH --job-name=norm_pairwise_N01
+#SBATCH --output=trace_%a.out
+#SBATCH --error=error_%a.err
+#SBATCH --mem-per-cpu=4096
+#SBATCH --time=168:00:00
+#! /bin/bash#SBATCH --array=1-12
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --export=NONE
+module load R/4.0.5
+Rscript ../R-ind/norm_pairwise_N01.R $SLURM_ARRAY_TASK_ID 

### 2021-11-04 data/hakear/simulations/supplementary/one-gran/raw/null_design_quantrans/R-agg/aggregate01.R

@@ -1,0 +1,57 @@
+# contains function to aggregate data each for from N(0,1), N(0,5), N(5,1), N(5,5), Gamma(0.5, 1), Gamma(2, 1)
+#  aggregated data from the file simulations/results/norm/folder_name and tidy them up
+library(tidyverse)
+library(here)
+library(readr)
+library(rlang)
+
+
+aggregate01 <-  function(folder_name){
+  
+all_files = list.files(path = paste0("simulations/supplementary/one-gran/raw/null_design_quantrans/data-ind/", folder_name), 
+                       pattern = ".rds")
+
+# all_files = list.files(path = paste0("simulations/vary_all/raw/data-ind/", folder_name), 
+#                        pattern = ".rds")
+# 
+# 
+
+names_levels <- map_dfr(all_files, 
+                        function(x){
+                          z = str_split(str_remove(x, "_wpd.rds"), "_") %>% 
+                            unlist()
+                          bind_cols(nx = as.numeric(z[1]),
+                                    nfacet = as.numeric(z[2]))
+                        })
+  
+all_files_path <- paste0("simulations/supplementary/one-gran/raw/null_design_quantrans/data-ind/",folder_name,"/",
+         all_files)  
+  
+
+all_data <- lapply(1:length(all_files_path), function(x){
+    
+  data = all_files_path %>% magrittr::extract2(x) %>% 
+    readRDS()
+  
+  names = names_levels %>% magrittr::extract(x,)
+  names_rep =   names %>% slice(rep(1:n(), each = nrow(data)))
+  bind_cols(names_rep, data)
+  }) %>% bind_rows() %>% 
+    arrange(nfacet, nx)
+    
+write_rds(all_data, paste0("simulations/supplementary/one-gran/raw/null_design_quantrans/data-agg/all_data_", folder_name, ".rds"))
+}
+
+
+# use functions for all folders
+
+aggregate01(folder_name = "wpd_N01")
+# aggregate01(folder_name = "wpd_N05")
+# aggregate01(folder_name = "wpd_N51")
+# aggregate01(folder_name = "wpd_N55")
+# 
+# 
+# aggregate01(folder_name = "wpd_Gamma01")
+# aggregate01(folder_name = "wpd_Gamma21")
+# 
+# 

### 2021-11-04 data/hakear/simulations/supplementary/one-gran/raw/null_design_quantrans/R-ind/example.R

@@ -1,0 +1,8 @@
+library(readr)
+read_csv("../../../sim_table/sim_table.csv")
+
+
+
+
+#setwd("../../../")
+#print(getwd())

### 2021-11-04 data/hakear/simulations/supplementary/one-gran/raw/null_design_quantrans/R-ind/raw_pairwise_Gamma21.R

@@ -1,0 +1,63 @@
+#This script calculates raw mmpd for each simulation scenario
+##Read Simulation Table
+
+.libPaths(c("~/R/libs", .libPaths()))
+library(distributional)
+library(readr)
+library(drake)
+library(tidyverse)
+library(hakear)
+
+set.seed(9999)
+
+nsim = 200
+# change path while running it on HPC
+simtable<-read_csv(here::here('simulations/supplementary/one-gran/sim_table.csv'))
+#simtable<-read_csv('../../../sim_table/sim_table.csv')
+
+### Extract flags from simulation scenario
+
+#scen<- 2 #If running within R uncomment this.  This will only run first scenario
+scen<-as.numeric(commandArgs()[[6]]) # If running batch job uncomment this
+
+simj<-simtable[scen,] #Extract row of table
+nfacetj<-simj$nfacet # Which nfacet level
+nxj<-simj$nx #Which nx level
+
+#create data for each row for null normal
+
+sim_null_normal = function(nxj, nfacetj){
+  rep(distributional::dist_gamma(2, 1), 
+      times = nxj*nfacetj)
+}
+
+sim_panel_data = 
+  hakear::sim_panel(nx = nxj,
+                    nfacet = nfacetj, 
+                    ntimes = 500, 
+                    sim_dist = sim_null_normal(nxj, nfacetj)) %>% 
+  unnest(c(data)) %>% ungroup()
+
+set.seed(1111)
+
+raw_dist <- map(seq_len(nsim), function(i)
+{
+  new_sim_data = sample(sim_panel_data$sim_data, 
+                        size = nrow(sim_panel_data))
+  new_data = sim_panel_data %>% 
+    select(-sim_data) %>% 
+    mutate(sim_data = new_sim_data)
+  
+  # for creating one raw mmpd
+  raw_mmpd = compute_pairwise_max(new_data, 
+                                  gran_x = "id_x",
+                                  gran_facet = "id_facet",
+                                  response = sim_data) %>% 
+    as_tibble() %>% mutate(perm_id = i)
+  
+}) %>% bind_rows()
+
+saveRDS(raw_dist,
+        paste0('simulations/supplementary/one-gran/raw/null_design_quantrans/data-ind/wpd_Gamma21/',
+               nxj,'_',
+               nfacetj,'_wpd.rds')) 

### 2021-11-04 data/hakear/simulations/supplementary/one-gran/raw/null_design_quantrans/job/job-raw-Gamma01.sh

@@ -1,0 +1,12 @@
+#! /bin/bash
+#SBATCH --job-name=raw_pairwise_Gamma01
+#SBATCH --output=trace_%a.out
+#SBATCH --error=error_%a.err
+#SBATCH --mem-per-cpu=4096
+#SBATCH --time=168:00:00
+#SBATCH --array=1-64
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --export=NONE
+module load R/4.0.0-openblas
+Rscript ../R-ind/raw_pairwise_Gamma01.R $SLURM_ARRAY_TASK_ID 

### 2021-11-04 data/hakear/simulations/supplementary/one-gran/raw/null_design_quantrans/job/job-raw-N51.sh

@@ -1,0 +1,12 @@
+#! /bin/bash
+#SBATCH --job-name=raw_pairwise_N51
+#SBATCH --output=trace_%a.out
+#SBATCH --error=error_%a.err
+#SBATCH --mem-per-cpu=4096
+#SBATCH --time=168:00:00
+#SBATCH --array=1-64
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --export=NONE
+module load R/4.0.0-openblas
+Rscript ../R-ind/raw_pairwise_N51.R $SLURM_ARRAY_TASK_ID 

### 2021-11-04 data/hakear/simulations/supplementary/one-gran/raw/null_design_quantrans/job/job-raw-N55.sh

@@ -1,0 +1,12 @@
+#! /bin/bash
+#SBATCH --job-name=raw_pairwise_N55
+#SBATCH --output=trace_%a.out
+#SBATCH --error=error_%a.err
+#SBATCH --mem-per-cpu=4096
+#SBATCH --time=168:00:00
+#SBATCH --array=1-64
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --export=NONE
+module load R/4.0.0-openblas
+Rscript ../R-ind/raw_pairwise_N55.R $SLURM_ARRAY_TASK_ID 

### 2021-11-04 data/hakear/simulations/supplementary/one-gran/raw/null_design_scale/R-ind/norm_pairwise_Gamma01.R

@@ -1,0 +1,65 @@
+#This script calculates norm mmpd for each simulation scenario
+##Read Simulation Table
+
+.libPaths(c("~/R/libs", .libPaths()))
+library(distributional)
+library(readr)
+library(drake)
+library(tidyverse)
+library(hakear)
+library(here)
+
+set.seed(9999)
+
+nsim = 200
+# change path while running it on HPC
+# simtable<-read_csv(here::here('simulations/null/sim_table.csv'))
+simtable<-read_csv('../null/sim_table.csv')
+
+### Extract flags from simulation scenario
+
+#scen<- 2 #If running within R uncomment this.  This will only run first scenario
+scen<-as.numeric(commandArgs()[[6]]) # If running batch job uncomment this
+
+simj<-simtable[scen,] #Extract row of table
+nfacetj<-simj$nfacet # Which nfacet level
+nxj<-simj$nx #Which nx level
+
+#create data for each row for null gamma
+
+sim_null_gamma = function(nxj, nfacetj){
+  rep(distributional::dist_gamma(0.5, 1), 
+      times = nxj*nfacetj)
+}
+
+sim_panel_data = 
+  hakear::sim_panel(nx = nxj,
+                    nfacet = nfacetj, 
+                    ntimes = 500, 
+                    sim_dist = sim_null_gamma) %>% 
+  unnest(c(data)) %>% ungroup()
+
+set.seed(1111)
+
+  norm_dist <- map(seq_len(nsim), function(i)
+{
+  new_sim_data = sample(sim_panel_data$sim_data, 
+                        size = nrow(sim_panel_data))
+  new_data = sim_panel_data %>% 
+    select(-sim_data) %>% 
+    mutate(sim_data = new_sim_data)
+  
+  # for creating one norm mmpd
+  norm_mmpd = compute_wpd_norm(new_data, 
+                                 gran_x = "id_x",
+                                 gran_facet = "id_facet",
+                                 response = sim_data) %>% 
+    as_tibble() %>% mutate(perm_id = i)
+  
+}) %>% bind_rows()
+
+  saveRDS(norm_dist, here(paste0('../results/norm/wpd_Gamma01/',
+                           nxj,'_',
+                           nfacetj,'_wpd_Gamma01.rds')))
+
+

### 2021-11-04 data/hakear/simulations/supplementary/one-gran/raw/null_design_scale/job/job-raw-N55.sh

@@ -1,0 +1,12 @@
+#! /bin/bash
+#SBATCH --job-name=raw_pairwise_N55
+#SBATCH --output=trace_%a.out
+#SBATCH --error=error_%a.err
+#SBATCH --mem-per-cpu=4096
+#SBATCH --time=168:00:00
+#SBATCH --array=1-64
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --export=NONE
+module load R/4.0.0-openblas
+Rscript raw_pairwise_N55.R $SLURM_ARRAY_TASK_ID 

### 2021-11-04 data/hakear/simulations/supplementary/sim_table/simtable.R

@@ -1,0 +1,32 @@
+### This code creates a table of data for different simulations scenarios ###
+library(tidyverse)
+# nperm and nsim can run parallelly
+sim_table <- expand.grid(#data = c(99, 9999), # different data with same specifications
+  nperm = c(10, 30, 50, 100, 300),
+  #nperm = c(10, 30, 500),
+  ntimes = c(5, 30, 100, 500),
+  #ntimes = c(5, 500),
+  # different permutations to compute mean and sd
+  #nobs_each_comb  = c(10, 40, 500), # different number of observations for each combinations for each panel
+  #nobs_diff_comb  = c("comb1", "comb2"), # different number of observations for different combinations for each panel
+  #nsim = c(100, 500), # number of simulations of the entire exercise to compute distribution and CI of MMPD 
+  #dist = c("N01", "N05", "N51", "N55", "G01", "G21"), #
+  dist = c("N01"), #distributions from where data is generated
+   nfacet = c(5, 7, 14, 20, 31), # range of facet levels 
+  # 50 removed for now
+   #nfacet = c(2, 3),
+   #nx = c(2, 3)
+   nx = c(5, 7, 14, 20, 31), # range of x levels
+  design = c("null", "vary_x", "vary_f", "vary_all"),
+  #seed = seq(1500, 7500, 1000),
+  seed = c(1500),
+  #w = c(seq(1, 20, 7), 50),
+  w = c(1, 10, 50),
+  type = c("norm-nperm")
+             
+) %>% tibble()
+
+
+#Export as .csv
+
+write_csv(sim_table, here::here('simulations/supplementary/sim_table/sim_table.csv'))

### 2021-11-04 data/hakear/simulations/supplementary/test-hpc/R-agg/aggregate01-power-one-alternate-low-inc.R

@@ -1,0 +1,37 @@
+#  aggregated data from the file simulations/results/norm/folder_name and tidy them up
+library(tidyverse)
+library(here)
+library(readr)
+library(rlang)
+
+folder = "data-ind-power-one-alternate-low-inc/"
+  
+all_files = list.files(path = paste0("simulations/supplementary/test-hpc/", folder), 
+                       pattern = ".csv")
+
+names_levels <- map(all_files, 
+                        function(x){
+                          z = str_remove(str_remove(x, "null-oneseed-id-"), ".csv") 
+                        }) %>% unlist()
+
+  
+all_files_path <- paste0("simulations/supplementary/test-hpc/", folder, 
+         all_files)  
+  
+
+all_data <- lapply(1:length(all_files_path), function(x){
+    
+  data = all_files_path %>% magrittr::extract2(x) %>% 
+    read_csv()
+  
+  names = rep(names_levels[x], each = nrow(data))
+  data = data %>% mutate(seed_id = names, 
+                         select_harmony = as.character(select_harmony)) %>% mutate(significance = if_else(!is.na(str_extract(select_harmony, "\\*\\*\\*")), 99, if_else(!is.na(str_extract(select_harmony, "\\*\\*")), 95, if_else(!is.na(str_extract(select_harmony, "\\*")), 90, 0))))
+  }) %>% bind_rows() 
+
+
+
+ write_rds(all_data, paste0("simulations/supplementary/test-hpc/data-agg/all_data-power-one-alternate-low-inc.rds"))
+
+
+

### 2021-11-04 data/hakear/simulations/supplementary/test-hpc/R-agg/makegraph02.R

@@ -1,0 +1,159 @@
+library(here)
+library(readr)
+library(tidyverse)
+# run for  norm MMPD files aggregation
+#all_data <- read_rds("simulations/result_report/all_data_ norm.rds")
+# run for  norm max pairwise distances files aggregation
+
+
+makegraph02 <- function(folder_name){
+  
+all_data <- read_rds(paste0("simulations/raw/null_design_quantrans/data-agg/all_data_", folder_name, ".rds"))
+  
+
+nxbyfacet_density <- all_data %>% 
+  ggplot(aes(x = value)) + 
+  geom_density(fill = "blue") +
+  facet_grid(nx~nfacet,
+             labeller = "label_both") + 
+  xlab("wpd") +
+  scale_x_continuous(breaks = scales::breaks_extended(3))
+
+ggsave(nxbyfacet_density, filename = paste0("simulations/raw/null_design_quantrans/figs/", "nxbyfacet_density_", folder_name,".png"))
+
+
+nxbyfacet_ridge <- all_data %>% 
+  ggplot(aes(x = value, y = as.factor(nx))) +
+  ggridges::geom_density_ridges() +
+  facet_wrap(~nfacet, labeller = "label_both", nrow = 2) + 
+  xlab("mmpd") +
+  ylab("nx") +
+  scale_x_continuous(breaks = scales::breaks_extended(4))
+
+ggsave(nxbyfacet_ridge, filename = paste0("simulations/raw/null_design_quantrans/figs/", "nxbyfacet_ridge_", folder_name,".png"))
+
+
+nfacetbynx_ridge <- all_data %>% 
+  ggplot(aes(x = value, y = as.factor(nfacet))) +
+  ggridges::geom_density_ridges() +
+  facet_wrap(~nx, labeller = "label_both", nrow = 2) + 
+  xlab("mmpd") +
+  ylab("nfacet") +
+  scale_x_continuous(breaks = scales::breaks_extended(4))
+
+ggsave(nfacetbynx_ridge, filename = paste0("simulations/raw/null_design_quantrans/figs/", "nfacetbynx_ridge_", folder_name,".png"))
+}
+
+makegraph02(folder_name = "wpd_N01")
+makegraph02(folder_name = "wpd_N05")
+makegraph02(folder_name = "wpd_N51")
+makegraph02(folder_name = "wpd_N55")
+
+
+
+
+# N(0,1) and N(5,1) together as they will go in the paper
+
+normal_0_1 <- aggregate01(folder_name = "wpd_N01")
[…]

### 2021-11-04 data/hakear/simulations/supplementary/test-hpc/R-ind/job-power-one.sh

@@ -1,0 +1,12 @@
+#! /bin/bash
+#SBATCH --power-one
+#SBATCH --output=trace_%a.out
+#SBATCH --error=error_%a.err
+#SBATCH --mem-per-cpu=4096
+#SBATCH --time=72:00:00
+#SBATCH --array=1-27000
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --export=NONE
+module load R/4.0.5
+Rscript ../R-ind/power-one.R $SLURM_ARRAY_TASK_ID 

### 2021-11-04 data/hakear/simulations/supplementary/test-hpc/R-ind/power-comp.R

@@ -1,0 +1,18 @@
+all_data <- read_rds("simulations/supplementary/test-hpc/data-agg/all_data-power-one-alternate-low-inc.rds.rds")
+
+
+# computes size for 99 percentile threshold
+data_summary <- all_data %>%
+  mutate(sig = if_else(significance!=99, 0, 99)) %>% 
+  group_by(seed_id) %>% 
+  summarize(sumsig = sum(sig)) %>% 
+  count(sumsig==0) %>% 
+  mutate(p_value = n/sum(n)) %>% 
+  slice(1)
+
+
+# plots
+ggplot() + 
+  geom_histogram(data = all_data, aes(x = wpd))  + 
+  geom_vline(data = data_summary, aes(xintercept = p_value),colour = "red")
+

### 2021-11-04 data/hakear/simulations/supplementary/test-hpc/R-ind/power-one-alternate-high-inc.R

@@ -1,0 +1,120 @@
+# This script computes power of the test in our paper 
+# low increment : 0.5 , medium increment : 2, high increment: 5
+# initial dist : N(0, 1)
+# only one alternate considered 
+# since all the harmonies need to be run at once, you can only run the seeds parallely, not all the harmonies.
+# sim_table should only contain the seeds in that case
+
+library(readr)
+library(dplyr)
+library(distributional)
+library(gravitas)
+library(here)
+library(purrr)
+library(tidyverse)
+library(purrr)
+library(hakear)
+
+# harmony_table<-read_csv(here::here('simulations/sim_table/sim_table.csv'))
+
+sample_seed <- seq(1000, 10999, by = 10) %>% as_tibble()
+simtable <- sample_seed
+scen <- as.numeric(commandArgs()[[6]])
+#scen = 2 # testing if works for one seed
+simj <- simtable[scen, ] # Extract row of table
+seedj <- simj$value # which seed with which simulations should be run
+
+
+harmony_table <- expand.grid(
+  x_levels = c(3, 7, 14),
+  facet_levels = c(2, 9, 10)
+) %>%
+  mutate(design = c("null", "null", "varall", "null", "null", "null", "null", "null", "null"))
+
+# parameters and design used for this simulation
+npermj <- 200
+ntimesj <- 500
+nsampj <- 100
+wj <- 5 # change this to change increment
+
+sim_null <- function(nxj, nfacetj, mean, sd, w = 0) {
+  rep(distributional::dist_normal(mu = mean, sigma = sd),
+      times = nxj * nfacetj
+  )
+}
+sim_varall <- function(nx, nfacet, mean, sd, w)
+{
+  dist_normal((mean + seq(0,
+                          (nx*
+                             nfacet - 1), by  = 1)*w), sd)
+}
+
+set.seed(seedj)
+## code for one harmony table starts
+sim_data_all <- map(
+  seq_len(nrow(harmony_table)),
+  function(x) {
+    y <- harmony_table %>% magrittr::extract(x, )
+    
+    # assigning different rows for different design
+    
+    # setting value of parameters
+    designj <- y$design
+    nfacetj <- y$facet_levels # Which nfacet level
+    nxj <- y$x_levels
+    
+    
+    if (designj == "null") {
+      sim_function <- sim_null
+    } else if (designj == "varx") {
+      sim_function <- sim_varx
+    } else if (designj == "varf") {
+      sim_function <- sim_varf
+    } else {
[…]

### 2021-11-04 data/hakear/simulations/supplementary/test-hpc/R-ind/power-one-alternate-low-inc.R

@@ -1,0 +1,120 @@
+# This script computes power of the test in our paper 
+# low increment : 0.5 , medium increment : 2, high increment: 5
+# initial dist : N(0, 1)
+# only one alternate considered 
+# since all the harmonies need to be run at once, you can only run the seeds parallely, not all the harmonies.
+# sim_table should only contain the seeds in that case
+
+library(readr)
+library(dplyr)
+library(distributional)
+library(gravitas)
+library(here)
+library(purrr)
+library(tidyverse)
+library(purrr)
+library(hakear)
+
+# harmony_table<-read_csv(here::here('simulations/sim_table/sim_table.csv'))
+
+sample_seed <- seq(1000, 10999, by = 10) %>% as_tibble()
+simtable <- sample_seed
+scen <- as.numeric(commandArgs()[[6]])
+#scen = 2 # testing if works for one seed
+simj <- simtable[scen, ] # Extract row of table
+seedj <- simj$value # which seed with which simulations should be run
+
+
+harmony_table <- expand.grid(
+  x_levels = c(3, 7, 14),
+  facet_levels = c(2, 9, 10)
+) %>%
+  mutate(design = c("null", "null", "varall", "null", "null", "null", "null", "null", "null"))
+
+# parameters and design used for this simulation
+npermj <- 200
+ntimesj <- 500
+nsampj <- 100
+wj <- 0.5 # change this to change increment
+
+sim_null <- function(nxj, nfacetj, mean, sd, w = 0) {
+  rep(distributional::dist_normal(mu = mean, sigma = sd),
+      times = nxj * nfacetj
+  )
+}
+sim_varall <- function(nx, nfacet, mean, sd, w)
+{
+  dist_normal((mean + seq(0,
+                          (nx*
+                             nfacet - 1), by  = 1)*w), sd)
+}
+
+set.seed(seedj)
+## code for one harmony table starts
+sim_data_all <- map(
+  seq_len(nrow(harmony_table)),
+  function(x) {
+    y <- harmony_table %>% magrittr::extract(x, )
+    
+    # assigning different rows for different design
+    
+    # setting value of parameters
+    designj <- y$design
+    nfacetj <- y$facet_levels # Which nfacet level
+    nxj <- y$x_levels
+    
+    
+    if (designj == "null") {
+      sim_function <- sim_null
+    } else if (designj == "varx") {
+      sim_function <- sim_varx
+    } else if (designj == "varf") {
+      sim_function <- sim_varf
+    } else {
[…]

### 2021-11-04 data/hakear/simulations/supplementary/test-hpc/R-ind/power-two-alternate-high-inc.R

@@ -1,0 +1,120 @@
+# This script computes power of the test in our paper 
+# low increment : 0.5 , medium increment : 2, high increment: 5
+# initial dist : N(0, 1)
+# only one alternate considered 
+# since all the harmonies need to be run at once, you can only run the seeds parallely, not all the harmonies.
+# sim_table should only contain the seeds in that case
+
+library(readr)
+library(dplyr)
+library(distributional)
+library(gravitas)
+library(here)
+library(purrr)
+library(tidyverse)
+library(purrr)
+library(hakear)
+
+# harmony_table<-read_csv(here::here('simulations/sim_table/sim_table.csv'))
+
+sample_seed <- seq(1000, 10999, by = 10) %>% as_tibble()
+simtable <- sample_seed
+scen <- as.numeric(commandArgs()[[6]])
+#scen = 2 # testing if works for one seed
+simj <- simtable[scen, ] # Extract row of table
+seedj <- simj$value # which seed with which simulations should be run
+
+
+harmony_table <- expand.grid(
+  x_levels = c(3, 7, 14),
+  facet_levels = c(2, 9, 10)
+) %>%
+  mutate(design = c("null", "null", "varall", "null", "null", "null", "varall", "null", "null"))
+
+# parameters and design used for this simulation
+npermj <- 200
+ntimesj <- 500
+nsampj <- 100
+wj <- 5 # change this to change increment
+
+sim_null <- function(nxj, nfacetj, mean, sd, w = 0) {
+  rep(distributional::dist_normal(mu = mean, sigma = sd),
+      times = nxj * nfacetj
+  )
+}
+sim_varall <- function(nx, nfacet, mean, sd, w)
+{
+  dist_normal((mean + seq(0,
+                          (nx*
+                             nfacet - 1), by  = 1)*w), sd)
+}
+
+set.seed(seedj)
+## code for one harmony table starts
+sim_data_all <- map(
+  seq_len(nrow(harmony_table)),
+  function(x) {
+    y <- harmony_table %>% magrittr::extract(x, )
+    
+    # assigning different rows for different design
+    
+    # setting value of parameters
+    designj <- y$design
+    nfacetj <- y$facet_levels # Which nfacet level
+    nxj <- y$x_levels
+    
+    
+    if (designj == "null") {
+      sim_function <- sim_null
+    } else if (designj == "varx") {
+      sim_function <- sim_varx
+    } else if (designj == "varf") {
+      sim_function <- sim_varf
+    } else {
[…]

### 2021-11-04 data/hakear/simulations/supplementary/test-hpc/R-ind/power-two-alternate-low-inc.R

@@ -1,0 +1,120 @@
+# This script computes power of the test in our paper 
+# low increment : 0.5 , medium increment : 2, high increment: 5
+# initial dist : N(0, 1)
+# only one alternate considered 
+# since all the harmonies need to be run at once, you can only run the seeds parallely, not all the harmonies.
+# sim_table should only contain the seeds in that case
+
+library(readr)
+library(dplyr)
+library(distributional)
+library(gravitas)
+library(here)
+library(purrr)
+library(tidyverse)
+library(purrr)
+library(hakear)
+
+# harmony_table<-read_csv(here::here('simulations/sim_table/sim_table.csv'))
+
+sample_seed <- seq(1000, 10999, by = 10) %>% as_tibble()
+simtable <- sample_seed
+scen <- as.numeric(commandArgs()[[6]])
+#scen = 2 # testing if works for one seed
+simj <- simtable[scen, ] # Extract row of table
+seedj <- simj$value # which seed with which simulations should be run
+
+
+harmony_table <- expand.grid(
+  x_levels = c(3, 7, 14),
+  facet_levels = c(2, 9, 10)
+) %>%
+  mutate(design = c("null", "null", "varall", "null", "null", "null", "varall", "null", "null"))
+
+# parameters and design used for this simulation
+npermj <- 200
+ntimesj <- 500
+nsampj <- 100
+wj <- 0.5 # change this to change increment
+
+sim_null <- function(nxj, nfacetj, mean, sd, w = 0) {
+  rep(distributional::dist_normal(mu = mean, sigma = sd),
+      times = nxj * nfacetj
+  )
+}
+sim_varall <- function(nx, nfacet, mean, sd, w)
+{
+  dist_normal((mean + seq(0,
+                          (nx*
+                             nfacet - 1), by  = 1)*w), sd)
+}
+
+set.seed(seedj)
+## code for one harmony table starts
+sim_data_all <- map(
+  seq_len(nrow(harmony_table)),
+  function(x) {
+    y <- harmony_table %>% magrittr::extract(x, )
+    
+    # assigning different rows for different design
+    
+    # setting value of parameters
+    designj <- y$design
+    nfacetj <- y$facet_levels # Which nfacet level
+    nxj <- y$x_levels
+    
+    
+    if (designj == "null") {
+      sim_function <- sim_null
+    } else if (designj == "varx") {
+      sim_function <- sim_varx
+    } else if (designj == "varf") {
+      sim_function <- sim_varf
+    } else {
[…]

### 2021-11-04 data/hakear/simulations/supplementary/test-hpc/R-ind/power-two-alternate-med-inc.R

@@ -1,0 +1,120 @@
+# This script computes power of the test in our paper 
+# low increment : 0.5 , medium increment : 2, high increment: 5
+# initial dist : N(0, 1)
+# only one alternate considered 
+# since all the harmonies need to be run at once, you can only run the seeds parallely, not all the harmonies.
+# sim_table should only contain the seeds in that case
+
+library(readr)
+library(dplyr)
+library(distributional)
+library(gravitas)
+library(here)
+library(purrr)
+library(tidyverse)
+library(purrr)
+library(hakear)
+
+# harmony_table<-read_csv(here::here('simulations/sim_table/sim_table.csv'))
+
+sample_seed <- seq(1000, 10999, by = 10) %>% as_tibble()
+simtable <- sample_seed
+scen <- as.numeric(commandArgs()[[6]])
+#scen = 2 # testing if works for one seed
+simj <- simtable[scen, ] # Extract row of table
+seedj <- simj$value # which seed with which simulations should be run
+
+
+harmony_table <- expand.grid(
+  x_levels = c(3, 7, 14),
+  facet_levels = c(2, 9, 10)
+) %>%
+  mutate(design = c("null", "null", "varall", "null", "null", "null", "varall", "null", "null"))
+
+# parameters and design used for this simulation
+npermj <- 200
+ntimesj <- 500
+nsampj <- 100
+wj <- 2 # change this to change increment
+
+sim_null <- function(nxj, nfacetj, mean, sd, w = 0) {
+  rep(distributional::dist_normal(mu = mean, sigma = sd),
+      times = nxj * nfacetj
+  )
+}
+sim_varall <- function(nx, nfacet, mean, sd, w)
+{
+  dist_normal((mean + seq(0,
+                          (nx*
+                             nfacet - 1), by  = 1)*w), sd)
+}
+
+set.seed(seedj)
+## code for one harmony table starts
+sim_data_all <- map(
+  seq_len(nrow(harmony_table)),
+  function(x) {
+    y <- harmony_table %>% magrittr::extract(x, )
+    
+    # assigning different rows for different design
+    
+    # setting value of parameters
+    designj <- y$design
+    nfacetj <- y$facet_levels # Which nfacet level
+    nxj <- y$x_levels
+    
+    
+    if (designj == "null") {
+      sim_function <- sim_null
+    } else if (designj == "varx") {
+      sim_function <- sim_varx
+    } else if (designj == "varf") {
+      sim_function <- sim_varf
+    } else {
[…]

### 2021-11-04 data/hakear/simulations/supplementary/test-hpc/R-ind/size-supplementary.R

@@ -1,0 +1,110 @@
+# This script computes size of the test in our paper and shows it is close to 95%
+
+library(readr)
+library(dplyr)
+library(distributional)
+library(gravitas)
+library(here)
+library(purrr)
+library(tidyverse)
+library(purrr)
+library(hakear)
+
+# make choices for x levels and seed for which size needs to be computed
+simtable <-  expand.grid(x_levels = c(3, 7, 14),
+                         facet_levels = c(2, 9, 10)) %>%
+  mutate(design = c("null", "varf","varx", "varall", "null", "varf", "varx", "varall", "null"))
+
+sample_seed = seq(1000, 10999, by = 10)
+npermj <- 200
+ntimesj <- 500
+nsampj <- 100
+wj <- 2
+
+# different designs
+
+sim_null <- function(nxj, nfacetj, mean, sd, w = 0){
+  rep(distributional::dist_normal(mu = mean, sigma = sd),
+      times = nxj*nfacetj)}
+
+sim_varx <- function(nx, nfacet, mean, sd, w)
+{
+  rep(dist_normal((mean + seq(0, nx-1, by  = 1)*w), sd), nfacet)
+}
+
+sim_varf <- function(nx, nfacet, mean, sd, w)
+{
+  rep(dist_normal((mean + seq(0, nfacet-1, by  = 1)*w), sd), each = nx)
+}
+
+sim_varall <- function(nx, nfacet, mean, sd, w)
+{
+  dist_normal((mean + seq(0,
+                          (nx*
+                             nfacet - 1), by  = 1)*w), sd)
+}
+
+
+
+sim_harmony_sim <-   map(seq_len(length(sample_seed)),
+                         function(seedj){
+                           set.seed(seedj) 
+  ## code for one harmony table starts
+  sim_data_all <- map(seq_len(nrow(simtable)),
+                      function(x){
+                        y = simtable %>% magrittr::extract(x,)
+                        
+                        # assigning different rows for different design
+                        
+                        # setting value of parameters
+                        designj = y$design
+                        nfacetj<- y$facet_levels # Which nfacet level
+                        nxj<- y$x_levels
+                    
+                        
+                        if(designj=="null")
+                          sim_function = sim_null
+                        else if(designj=="varx")
+                          sim_function =  sim_varx
+                        else if(designj=="varf")
+                          sim_function =  sim_varf
+                        else
+                          sim_function =  sim_varall
+                        
+                        # making data for each row using customized sim_function
+                     
[…]

### 2021-11-04 data/hakear/simulations/supplementary/test-hpc/R-ind/size-test.R

@@ -1,0 +1,90 @@
+library(readr)
+library(dplyr)
+library(hakear)
+library(distributional)
+library(gravitas)
+library(here)
+sample_seed = seq(1000,2999, by = 1000)
+nperm = 2
+
+#scen <-as.numeric(commandArgs()[[6]])
+sample_seedj <- sample_seed[scen]
+#sample_seedj = 1000
+
+simtable<-read_csv(here::here('simulations/sim_table/sim_table.csv'))
+
+set.seed(sample_seedj)
+
+library(distributional)
+sim_dist <- distributional::dist_normal(0,1)
+sim_orig <-  distributional::generate(sim_dist, 50000)
+sim_orig <- sim_orig[[1]]
+lensim <- length(sim_orig)
+
+# for each row of the sim table
+each_seed <- lapply(seq_len(nrow(simtable)), 
+                    function(scen){
+                      
+                      simj<-simtable[scen,] 
+                      nfacetj<-simj$nfacet
+                      nxj<-simj$nx
+                      
+                      id_x <- rep_len(seq_len(nxj), length.out = lensim) 
+                      id_facet <- rep_len(rep(seq_len(nfacetj), each = nxj), length.out = lensim) 
+                      
+                      sim_panel_orig <- bind_cols(id_x = id_x, id_facet = id_facet, sim_data = sim_orig)
+                      
+                      wpd_orig = compute_pairwise_norm_scalar(sim_panel_orig, 
+                                                              gran_x = "id_x",
+                                                              gran_facet = "id_facet",
+                                                              response = sim_data)
+                      
+                      wpd_sample <- lapply(seq_len(nperm), function(x){
+                        
+                        sample_data <- sample(sim_orig, size =  lensim)
+                        
+                        sim_panel_sample <- bind_cols(id_x = id_x, 
+                                                      id_facet = id_facet, 
+                                                      sim_data = sample_data)
+                        
+                        wpd =  compute_pairwise_norm_scalar(sim_panel_sample, 
+                                                            gran_x = "id_x",
+                                                            gran_facet = "id_facet",
+                                                            response = sim_data) %>% 
+                          as_tibble() %>% mutate(perm_id = x)
+                        
+                      }) %>% bind_rows()
+                      
[…]

### 2021-11-04 data/hakear/simulations/supplementary/test-hpc/job/job-power-1-alternate-high-inc.sh

@@ -1,0 +1,12 @@
+#! /bin/bash
+#SBATCH --job-name=power-one-alternate-high-inc
+#SBATCH --output=trace_%a.out
+#SBATCH --error=error_%a.err
+#SBATCH --mem-per-cpu=4096
+#SBATCH --time=168:00:00
+#SBATCH --array=1-1000
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --export=NONE
+module load R/4.0.5
+Rscript ../R-ind/power-one-alternate-high-inc.R $SLURM_ARRAY_TASK_ID 

### 2021-11-04 data/hakear/simulations/supplementary/test-hpc/job/job-power-2-alternate-med-inc.sh

@@ -1,0 +1,12 @@
+#! /bin/bash
+#SBATCH --job-name=power-two-alternate-med-inc
+#SBATCH --output=trace_%a.out
+#SBATCH --error=error_%a.err
+#SBATCH --mem-per-cpu=4096
+#SBATCH --time=168:00:00
+#SBATCH --array=1-1000
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --export=NONE
+module load R/4.0.5
+Rscript ../R-ind/power-two-alternate-med-inc.R $SLURM_ARRAY_TASK_ID 

### 2021-11-04 data/hakear/simulations/supplementary/test-hpc/job/job-size-nullonly.sh

@@ -1,0 +1,12 @@
+#! /bin/bash
+#SBATCH --job-name=size-nullonly
+#SBATCH --output=trace_%a.out
+#SBATCH --error=error_%a.err
+#SBATCH --mem-per-cpu=4096
+#SBATCH --time=168:00:00
+#SBATCH --array=1-1000
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --export=NONE
+module load R/4.0.0-openblas
+Rscript ../R-ind/size-nullonly.R $SLURM_ARRAY_TASK_ID 

### 2021-11-04 index.Rmd
make preview

---
title: 'Visualization and analysis of probability distributions of large temporal data'
author: 'Sayani Gupta'
degreetype: 'Doctor of Philosophy'
site: bookdown::bookdown_site
degrees: 'M.Stat, B.Sc (Stat Hons)'
bibliography: [bib/gravitas.bib, bib/hakear.bib, bib/gracsr.bib, bib/thesisrefs.bib]
link-citations: yes
---

```{r initial, echo = FALSE, cache = FALSE, results = 'hide'}
options("knitr.graphics.auto_pdf" = TRUE)
library(knitr)
opts_chunk$set(
  warning = FALSE, message = FALSE, echo = FALSE,
  fig.path = 'figure/', fig.align = 'center', fig.show = 'hold',
  cache = TRUE, external = TRUE, comment = "#>", dpi = 300,
  out.width = ifelse(is_html_output(), "100%", "\\textwidth"),
  cache.path = ifelse(is_html_output(), 'cache-gitbook/', 'cache-pdfbook/')
)
hook_output <-  knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```

`r if (knitr::is_latex_output()) '<!--'`

# Welcome {-}

This is the website for my PhD thesis at Monash University (Australia), titled "Visualization and analysis of probability distributions of large temporal data".

`r Sys.Date()`

`r if (knitr::is_latex_output()) '-->'`

### 2021-11-04 thesis.Rmd
make preview

---
title: 'Visualization and analysis of probability distributions of large temporal data'
author: 'Sayani Gupta'
degreetype: 'Doctor of Philosophy'
site: bookdown::bookdown_site
degrees: 'M.Stat, B.Sc (Stat Hons)'
bibliography: [bib/gravitas.bib, bib/hakear.bib, bib/gracsr.bib, bib/thesisrefs.bib]
link-citations: yes
---

```{r initial, echo = FALSE, cache = FALSE, results = 'hide'}
options("knitr.graphics.auto_pdf" = TRUE)
library(knitr)
opts_chunk$set(
  warning = FALSE, message = FALSE, echo = FALSE,
  fig.path = 'figure/', fig.align = 'center', fig.show = 'hold',
  cache = TRUE, external = TRUE, comment = "#>", dpi = 300,
  out.width = ifelse(is_html_output(), "100%", "\\textwidth"),
  cache.path = ifelse(is_html_output(), 'cache-gitbook/', 'cache-pdfbook/')
)
hook_output <-  knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```

`r if (knitr::is_latex_output()) '<!--'`

# Welcome {-}

This is the website for my PhD thesis at Monash University (Australia), titled "Visualization and analysis of probability distributions of large temporal data".

`r Sys.Date()`

`r if (knitr::is_latex_output()) '-->'`

<!--chapter:end:index.Rmd-->

`r if (knitr::is_html_output()) '<!--'`

# Copyright notice {-}

\textcopyright { }Sayani Gupta (\number\the\year).

I certify that I have made all reasonable efforts to secure copyright permissions for third-party content included in this thesis and have not knowingly added copyright content to my work without the owner's permission.

\newpage

`r if (knitr::is_html_output()) '-->'`

<!--chapter:end:Rmd//00-a-copyright.Rmd-->

# Abstract {-}

This work is driven by the need for a conceptual framework to tackle temporal analyses in a data-centric workflow. Most research on temporal data focuses on modeling. Corresponding software requires very stringently formatted data, but does little in providing guidelines or tools for wrangling raw data into the required format. This has led to ad-hoc, and once-off solutions, which this research repairs.

There are three original contributions for the temporal data analysis in this research. They are grounded in the spirit of exploratory data analysis for time-indexed data. The first contribution (Chapter \@ref(ch:gravitas)) is a new technique for visualizing data using a calendar layout. It is most useful when the data relates to daily human activity, and patterns can be explored in relation to people's schedules. The second contribution (Chapter \@ref(ch:hakear)) is a new data abstraction which streamlines transformation, visualization, and modeling in temporal contexts. This "tsibble" object is a data infrastructure forming the foundation of temporal data pipelines. The third contribution (Chapter \@ref(ch:gracsr)) is to equip analysts with exploratory and explanatory tools for discovering missing patterns in time, and thus facilitating decisions on appropriate imputation methods for further downstream analysis.

\newpage

<!--chapter:end:Rmd//00-b-abstract.Rmd-->

# Declaration {-}

Placeholder

<!--chapter:end:Rmd//00-c-declaration.Rmd-->

# Acknowledgements {-}

This thesis would not be possible without many people.

First and foremost, I would like to express my gratitude to my supervisors, Dianne Cook and Rob J Hyndman. They are outstanding supervisors, providing excellent insights on statistical computing and graphics during our weekly supervisory meetings. I'm deeply impressed by their dynamics, enthusiasm, and visions about open source software and data science, all of which have formed the support and encouragement for me to continue with this non-traditional research path.

I thank Heike Hofmann for hosting me at Iowa State University during the summer of 2018. I enjoyed many discussions with Heike, where I received brilliant questions and insightful advice on my research. I thank Stuart Lee and Mitchell O'Hara-Wild for countless conversations on software development, that helped to shape this thesis. I thank David Frazier for being the best pingpong partner, truly boosting the creative thinking process.

I thank Emi Tanaka for sending me PRs for proofreading my thesis on Github, unlocking the power of an open-sourced research compendium. I thank Wenjing Wang and Carson Sievert for being so helpful in the early days of my PhD. I thank the R community, the tidyverse team, and all NUMBATs for ongoing inspiration.

Last but not least, I'd like to thank my mother for her unceasing love.

<!--chapter:end:Rmd//00-d-acknowlegements.Rmd-->

# Preface {-}

Chapter \@ref(ch:gravitas) has been accepted by the *Journal of Computational and Graphical Statistics*. It has won the ACEMS Business Analytics Prize in 2020. The accompanying R package **gravitas** is on CRAN. Chapter \@ref(ch:hakear) and Chapter \@ref(ch:gracsr) are yet to be submitted.

### Open and reproducible research {-}

This thesis is written in R Markdown [@rmarkdown] with **bookdown** [@bookdown], using **renv** [@renv] to create reproducible environments. The online version of this thesis is hosted at <https://sayani.netlify.app/>, powered by [Netlify](http://netlify.com). All materials (including the data sets and source files) required to reproduce this document can be found at the Github repository <https://github.com/Sayani07/thesis>.

### License {-}

<a rel="license" href="http://creativecommons.org/licenses/by-nc-sa/4.0/"><img alt="Creative Commons Licence" style="border-width:0" src="https://i.creativecommons.org/l/by-nc-sa/4.0/88x31.png" /></a>

This work is licensed under a [Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International License](http://creativecommons.org/licenses/by-nc-sa/4.0/).

The code used in this document is available under the [MIT license](https://opensource.org/licenses/MIT).

\clearpage\pagenumbering{arabic}\setcounter{page}{0}

<!--chapter:end:Rmd//00-e-preface.Rmd-->

# Introduction {#ch:intro}

Placeholder

## Visualizing probability distributions across bivariate cyclic temporal granularities {#sec:gravitas}
## Detecting distributional differences between temporal granularities for exploratory time series analysis {#sec:hakear}
## Clustering time series based on probability distributions across temporal granularities {#sec:gracsr}
## Summary

<!--chapter:end:Rmd//01-intro.Rmd-->

# Visualizing probability distributions across bivariate cyclic temporal granularities {#ch:gravitas}

Placeholder

## Introduction
## Linear time granularities {#sec:linear-time}
### Definitions
### Relativities
## Cyclic time granularities {#sec:cyclic-gran}
### Circular granularities {#sec:circular-gran-def}
### Quasi-circular granularities {#sec:quasi-circular-gran-def}
### Aperiodic granularities {#sec:aperiodic-gran-def}
### Relativities {#sec:cyclic-calendar}
### Computation
#### Single-to-multiple order-up {-}
#### Multiple-to-single order-up {-}
#### Multiple order-up quasi-circular granularities {-}
## Data structure {#sec:data-structure}
### Harmonies and clashes {#sec:synergy}
### Near-clashes {#sec:near-clashes}
## Visualization {#sec:visualization}
### Data summarization
### Display choices for univariate distributions
### Comparison across sub-groups induced by conditioning {#sec:aesthetics}
#### Levels {-}
#### Synergy of cyclic granularities {-}
#### Conditioning variable {-}
## Applications {#sec:application}
### Smart meter data of Australia {#sec:smartmeter}
#### Cyclic granularities search and computation {-}
#### Screening and visualizing harmonies {-}
### T20 cricket data of Indian Premier League {#sec:cricket}
## Discussion {#sec:discussion}
## Acknowledgments {-}
## Supplementary Materials

<!--chapter:end:Rmd//02-gravitas.Rmd-->

# Detecting distributional differences between temporal granularities for exploratory time series analysis  {#ch:hakear}

Placeholder

## Introduction
## Proposed distance measure {#sec:computation-wpd}
### Principle 
### Notations
### Computation
#### Data transformation
#### Characterising distributions
#### Distance between distributions
#### Within-facet and between-facet distances
#### Tuning parameter
#### Raw distance measure
#### Adjusting for the number of comparisons {#sec:norm-wpd}
## Ranking and selection of cyclic granularities {#sec:rank-wpd}
### Selection
### Ranking
## Simulations {#sec:simulations}
### Behavior of raw and adjusted distance measures
### Choosing threshold
### Environment
## Application to residential smart meter dataset {#sec:application-wpd}
## Discussion
## Acknowledgments {-}
## Supplementary Materials

<!--chapter:end:Rmd//03-hakear.Rmd-->

# Data representation, visual and analytical techniques for demystifying temporal missing data {#ch:mists}

<!--chapter:end:Rmd//04-gracsr.Rmd-->

# Conclusion and future plans {#ch:conclusion}

Placeholder

## Software development
### sugrrants
### tsibble
### mists
## Future work
### Process for generating missing data in time
### Visual methods for temporal data of nesting and crossing interactions
## Final words

<!--chapter:end:Rmd//05-conclusion.Rmd-->

# (APPENDIX) Appendix {-}

# Data dictionary

```{r wdi-data-dict}
```

(ref:wdi-data-dict) Data dictionary for the World Development Indicators.

<!--chapter:end:Rmd//06-appendix.Rmd-->

`r if (knitr::is_html_output()) '# Bibliography {-}'`

<!--chapter:end:Rmd//99-references.Rmd-->

### 2021-11-04 Rmd/00-c-declaration.Rmd
remove last 2 commits

`r if (knitr::is_html_output()) '<!--'`

\newpage

# Declaration {-}

I have not renumbered sections of submitted or published papers in order to generate a consistent presentation within the thesis.

\textbf{Student name:} Sayani Gupta

\textbf{Student signature:}

```{r signature, out.width = "20%", fig.align = "left", eval = file.exists("img/signature.png")}
knitr::include_graphics("img/signature.png")
```

\textbf{Date:} `r Sys.Date()`

`r if (knitr::is_html_output()) '-->'`

### 2021-11-04 Rmd/00-d-acknowlegements.Rmd
remove last 2 commits

# Acknowledgements {-}

## supervisors
Rob and Di are phenomenal leaders in their respective fields, who lead by example. I am inspired by their creativity, wisdom, discipline, and dedication to really contribute to the society through their research. Thank you for constantly pushing me to improve as a programmer and researcher, and for regularly sharing best practices for conducting research. Looking back, I am grateful for how my thoughts and work in statistical computing, graphics, and data analytics in general have evolved over the years. I still have a lot to learn, but I am a much more self-reliant and independent researcher than I was when I started. Di has been instrumental in exposing me to the potential benefits of effective data visualization. As a female researcher, I am encouraged by her willingness to pursue unconventional avenues and have frequently noticed her make a conscious decision to question existing  stereotypes and biases. Thank you Di for being a fantastic female role model.

## department and PhD colleagues

## R community

## friends

PhD in Covid has been harder and I would not have been able to make it through this if not for the emotional support of my friends and family. Thank you Puwasala, Sium for being an epitome of kindness and acting as a pillar of support. Thank you Tushar, Samarpita and Nairita for always motivating me and believing in me when I didn't believe in myself. Thank you Ian for always putting things in perspective when I lost sight. Thank you each one of you for being available, listening to me, and not giving up on me at my lowest. Thanks to my housemates Anjali, Surbhi and Dulaji for the fun company and conversations that kept me sane in the covid lockdowns.

## family
Finally, a big thanks to my family for always being supportive of my choices. Thanks to my mum (Nupur Gupta) and dad (Arun Prasad Gupta), from whom I learnt that no matter where you start from, if you persevere and are sincere in your efforts, you can sail through any difficult situation. They are my constant cheerleaders for all little and big endeavours. Thanks to my brother (Avijit) for always having my back and inspiring me to dream bigger. Thank you, Juhi (sister-in-law) for instilling in me the value of organisation in all aspects of life and the importance of prioritizing my physical health from time to time. Thank you all for your part in supporting me in this journey. I love you.

### 2021-11-04 Rmd/02-gravitas.Rmd
remove last 2 commits

# Visualizing probability distributions across bivariate cyclic temporal granularities {#ch:gravitas}

Deconstructing a time index into time granularities can assist in exploration and automated analysis of large temporal data sets. This paper describes classes of time deconstructions using linear and cyclic time granularities. Linear granularities respect the linear progression of time such as hours, days, weeks and months. Cyclic granularities can be circular such as hour-of-the-day, quasi-circular such as day-of-the-month, and aperiodic such as public holidays. The hierarchical structure of granularities creates a nested ordering: hour-of-the-day and second-of-the-minute are single-order-up. Hour-of-the-week is multiple-order-up, because it passes over day-of-the-week. Methods are provided for creating all possible granularities for a time index. A recommendation algorithm provides an indication whether a pair of granularities can be meaningfully examined together (a "harmony"), or when they cannot (a "clash").

 Time granularities can be used to create data visualizations to explore for periodicities, associations and anomalies. The granularities form categorical variables (ordered or unordered) which induce groupings of the observations. Assuming a numeric response variable, the resulting graphics are then displays of distributions compared across combinations of categorical variables.

 The methods implemented in the open source R package `gravitas` are consistent with a tidy workflow, with probability distributions examined using the range of graphics available in `ggplot2`.

```{r external, include = FALSE, cache = FALSE}
read_chunk('scripts/gravitas.R')
```

```{r load}
```

## Introduction

Temporal data are available at various resolutions depending on the context. Social and economic data are often collected and reported at coarse temporal scales such as monthly, quarterly or annually. With recent advancements in technology, more and more data are recorded at much finer temporal scales. Energy consumption may be collected every half an hour, energy supply may be collected every minute, and web search data might be recorded every second. As the frequency of data increases, the number of questions about the periodicity of the observed variable also increases. For example, data collected at an hourly scale can be analyzed using coarser temporal scales such as days, months or quarters. This approach requires deconstructing time in various possible ways called time granularities [@aigner2011visualization].

It is important to be able to navigate through all of these time granularities to have multiple perspectives on the periodicity of the observed data. This aligns with the notion of exploratory data analysis (EDA) [@Tukey1977-jx] which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing. Visualizing probability distributions conditional on one or more granularities is an indispensable tool for exploration. Analysts are expected to comprehensively explore the many ways to view and consider temporal data. However, the plethora of choices and the lack of a systematic approach to do so quickly can make the task overwhelming.

Calendar-based graphics [@wang2020calendar] are useful in visualizing patterns in the weekly and monthly structure and are helpful when checking for the effects of weekends or special days. Any temporal data at sub-daily resolution can also be displayed using this type of faceting [@Wickham2009pk] with days of the week, month of the year, or another sub-daily deconstruction of time. But calendar effects are not restricted to conventional day-of-week or month-of-year deconstructions. There can be many different time deconstructions, based on the calendar or on categorizations of time granularities.

Linear time granularities (such as hours, days, weeks and months) respect the linear progression of time and are non-repeating. One of the first attempts to characterize these granularities is due to @Bettini1998-ed. However, the definitions and rules defined are inadequate for describing non-linear granularities. Hence, there is a need to define some new time granularities, that can be useful in visualizations. Cyclic time granularities can be circular, quasi-circular or aperiodic. Examples of circular granularities are hour of the day and day of the week; an example of a quasi-circular granularity is day of the month; examples of aperiodic granularities are public holidays and school holidays.

Time deconstructions can also be based on the hierarchical structure of time. For example, hours are nested within days, days within weeks, weeks within months, and so on. Hence, it is possible to construct single-order-up granularities such as second of the minute, or multiple-order-up granularities such as second of the hour. The lubridate package [@Grolemund2011-vm] provides tools to access and manipulate common date-time objects. However, most of its accessor functions are limited to single-order-up granularities.

The motivation for this work stems from the desire to provide methods to better understand large quantities of measurements on energy usage reported by smart meters in households across Australia, and indeed many parts of the world. Smart meters currently provide half-hourly use in kWh for each household, from the time they were installed, some as early as 2012. Households are distributed geographically and have different demographic properties as well as physical properties such as the existence of solar panels, central heating or air conditioning. The behavioral patterns in households vary substantially; for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks. It is common to see aggregates [see @Goodwin_2012] of usage across households, such as half-hourly total usage by state, because energy companies need to plan for maximum loads on the network. But studying overall energy use hides the distribution of usage at finer scales, and makes it more difficult to find solutions to improve energy efficiency. We propose that the analysis of smart meter data will benefit from systematically exploring energy consumption by visualizing the probability distributions across different deconstructions of time to find regular patterns and anomalies. Although we were motivated by the smart meter example, the problem and the solutions we propose are practically relevant to any temporal data observed more than once per year. In a broader sense, it could be even suitable for data observed by years, decades, and centuries as might be in weather or astronomical data.

This work provides tools for systematically exploring bivariate granularities within the tidy workflow (@wickham2016r). In particular, we

  * provide a formal characterization of cyclic granularities;
  * facilitate manipulation of single- and multiple-order-up time granularities through cyclic calendar algebra;
  * develop an approach to check the feasibility of creating plots or drawing inferences for any two cyclic granularities.

The remainder of the paper is organized as follows: Section&nbsp;\ref{sec:linear-time} provides some background material on linear granularities and calendar algebra for computing different linear granularities. Section&nbsp;\ref{sec:cyclic-gran} formally characterizes different cyclic time granularities by extending the framework of linear time granularities, and introducing cyclic calendar algebra for computing cyclic time granularities. The data structure for exploring the conditional distributions of the associated time series across pairs of cyclic time granularities is discussed in Section&nbsp;\ref{sec:data-structure}. Section&nbsp;\ref{sec:visualization} discusses the role of different factors in constructing an informative and trustworthy visualization. Section&nbsp;\ref{sec:application} examines how systematic exploration can be carried out for a temporal and non-temporal application. Finally, we summarize our results and discuss possible future directions in Section&nbsp;\ref{sec:discussion}.

## Linear time granularities {#sec:linear-time}

Discrete abstractions of time such as weeks, months or holidays can be thought of as "time granularities". Time granularities are **linear** if they respect the linear progression of time. There have been several attempts to provide a framework for formally characterizing time granularities, including @Bettini1998-ed which forms the basis of the work described here.

### Definitions

\begin{defn}\label{def:definition}
A \textbf{time domain} is a pair $(T; \le)$ where $T$ is a non-empty set of time instants (equivalently, moments or points) and $\le$ is a total order on $T$.
\end{defn}

\noindent The time domain is assumed to be *discrete*, and there is unique predecessor and successor for every element in the time domain except for the first and last.

\begin{defn}\label{def:index set}
The \textbf{index set}, $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$, uniquely maps the time instants to the set of non-negative integers.
\end{defn}

\begin{defn}\label{def:linear}
A \textbf{linear granularity} is a mapping $G$ from the index set, $Z$, to subsets of the time domain such that:
  (1) if $i < j$ and $G(i)$ and $G(j)$ are non-empty, then each element of $G(i)$ is less than all elements of $G(j)$; and
  (2) if $i < k < j$ and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.
Each non-empty subset $G(i)$ is called a \textbf{granule}.
\end{defn}

\noindent This implies that the granules in a linear granularity are non-overlapping, continuous and ordered. The indexing for each granule can also be associated with a textual representation, called the label. A discrete time model often uses a fixed smallest linear granularity named by @Bettini1998-ed **bottom granularity**. \autoref{fig:linear-time} illustrates some common linear time granularities. Here, "hour" is the bottom granularity and "day", "week", "month" and "year" are linear granularities formed by mapping the index set to subsets of the hourly time domain. If we have "hour" running from $\{0, 1, \dots,t\}$, we will have "day" running from $\{0, 1, \dots, \lfloor t/24\rfloor\}$. These linear granularities are uni-directional and non-repeating.

```{r linear-time, echo=FALSE, fig.cap="(ref:linear-time)", fig.pos="!htbp"}
```

(ref:linear-time) Illustration of time domain, linear granularities and index set. Hour, day, week, month and year are linear granularities and can also be considered to be time domains. These are ordered with ordering guided by integers and hence is unidirectional and non-repeating. Hours could also be considered the index set, and a bottom granularity.

### Relativities

Properties of pairs of granularities fall into various categories.

\begin{defn}\label{def:finerthan}
A linear granularity $G$ is \textbf{finer than} a linear granularity $H$, denoted $G \preceq H$, if for each index $i$, there exists an index $j$ such that
$G(i) \subset H(j).$
\end{defn}

\begin{defn}\label{def:groupsinto}
A linear granularity $G$ \textbf{groups into} a linear granularity $H$, denoted
$G \trianglelefteq H$, if for each index $j$ there exists a (possibly infinite) subset $S$ of the integers such that $H(j) = \bigcup_{i \in S}G(i).$
\end{defn}

\noindent For example, both $day \trianglelefteq week$ and $day \preceq week$ hold, since every granule of $week$ is the union of some set of granules of day and each day is a subset of a $week$. These definitions are not equivalent. Consider another example, where $G_1$ denotes "weekend" and $H_1$ denotes "week". Then, $G_1 \preceq H_1$, but $G_1 \ntrianglelefteq H_1$. Further, with $G_2$ denoting "days" and $H_2$ denoting "business-week", $G_2 \npreceq H_2$, but $G_2 \trianglelefteq H_2$, since each business-week can be expressed as an union of some days, but Saturdays and Sundays are not subset of any business-week.
Moreover, with  $H_3$ denoting "public holidays", $G_1 \npreceq H_3$ and $G_1 \ntrianglelefteq H_3$.

<!-- # talk about period after definition 6 after you have explained period -->
<!-- The relationship has period 7. -->
<!-- The relationship $day \trianglelefteq month$ has a more complicated period. If leap years are ignored, each month is a grouping of the same number of days over years, hence the period of the grouping $(day, month)$ is one year. With the inclusion of leap years, the grouping period is 400 years. -->

\begin{defn}\label{def:periodic}
A granularity $G$ is \textbf{periodic} with respect to a finite granularity $H$ if:
(1) $G \trianglelefteq H$; and
(2) there exist $R$, $P \in \mathbb{Z}_+$, where $R$ is less than the number of granules of $H$, such that for all $i \in \mathbb{Z}_{\ge {0}}$ , if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then $H (i + R) = \bigcup_{j \in S} G(j + P)$.
\end{defn}

If $G$ groups into $H$, it would imply that any granule $H(i)$ is the union of some granules of $G$, for example, $G(a_1), G(a_2), \dots, G(a_k)$. Condition (2) in Definition \ref{def:periodic} implies that if $H(i + R) \neq \emptyset$, then $H(i + R) = \bigcup (G(a_1 + P), G(a_2 + P), \dots, G(a_k + P))$, resulting in a "periodic" pattern of the composition of $H$ using granules of $G$. In this pattern, each granule of $H$ is shifted by $P$ granules of $G$. $P$ is called the \textbf{Period} (@Bettini2000-qk).

For example, day is periodic with respect to week with $R=1$ and $P=7$, while (if we ignore leap years) day is periodic with respect to month with $R=12$ and $P=365$ as any month would consist of the same number of days across years.
Since the idea of period involves a pair of granularities, we say that the pair $(day, week)$ has period 7, while the pair $(day, month)$ has a period 365 (ignoring leap years).

Granularities can also be periodic with respect to other granularities, _"except for a finite number of spans of time where they behave in an anomalous way"_; these are called **quasi-periodic** relationships [@Bettini2000-vy]. In a Gregorian calendar with leap years, day groups quasi-periodically into month with the exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.

\begin{defn}\label{def:order}
The \textbf{order} of a linear granularity is the level of coarseness associated with a linear granularity. A linear granularity G will have lower order than H if each granule of G is composed of lower number of granules of bottom granularity than each granule of H.
\end{defn}

With two linear granularities $G$ and $H$, if $G$ *groups into* or *finer than* $H$ then $G$ is of lower order than $H$. For example, if the bottom granularity is hour, then granularity $day$ will have lower order than $week$ since each day consist of fewer hours than each week.

Granules in any granularity may be aggregated to form a coarser granularity. A system of multiple granularities in lattice structures is referred to as a **calendar** by @Dyreson_2000. Linear time granularities are computed through "calendar algebra" operations [@Ning_2002] designed to generate new granularities recursively from the bottom granularity. For example, due to the constant length of day and week, we can derive them from hour using
$$
  D(j) = \lfloor H(i)/24\rfloor, \qquad W(k) = \lfloor H(i)/(24*7)\rfloor,
$$
where $H$, $D$ and $W$ denote hours, days and weeks respectively.

## Cyclic time granularities {#sec:cyclic-gran}

Cyclic granularities represent cyclical repetitions in time. They can be thought of as additional categorizations of time that are not linear. Cyclic granularities can be constructed from two linear granularities, that relate periodically; the resulting cycles can be either _regular_ (**circular**), or _irregular_ (**quasi-circular**).

### Circular granularities {#sec:circular-gran-def}

\begin{defn}\label{def:circular}
A \textbf{circular granularity} $C_{B, G}$ relates linear granularity $G$ to bottom granularity $B$ if
\begin{equation} \label{eq:circular-gran}
\begin{split}
C_{B, G}(z) & = z\mod P(B, G) \quad \forall z \in \mathbb{Z}_{\geq 0} \\
\end{split}
\end{equation}
where
$z$ denotes the index set,
$B$ groups periodically into $G$ with regular mapping and period $P(B, G)$.
\end{defn}

```{r circular-dow, echo=FALSE, fig.cap= "(ref:circular-dow)",fig.pos="!htb"}
```

(ref:circular-dow) Index sets for some linear and circular granularities (a). Circular granularities can be constructed by slicing the linear granularity into pieces and stacking them (b).

\autoref{fig:circular-dow} illustrates some linear and cyclical granularities. Cyclical granularities are constructed by cutting the linear granularity into pieces, and stacking them to match the cycles (as shown in b). $B, G, H$ (day, week, fortnight, respectively) are linear granularities. The circular granularity $C_{B, G}$ (day-of-week) is constructed from $B$ and $G$, while  circular granularity $C_{B, H}$ (day-of-fortnight) is constructed from $B$ and $H$. These overlapping cyclical granularities share elements from the linear granularity. Each of $C_{B , G}$ and $C_{B , H}$ consist of repeated patterns $\{0, 1, \dots, 6\}$ and $\{0, 1, \dots, 13\}$ with $P=7$ and $P=14$ respectively.

Suppose ${L}$ is a label mapping that defines a unique label for each index $\ell \in \{ 0,1,\dots, (P-1)\}$. For example, the label mapping $L$ for $C_{B, G}$ can be defined as
$$
  L: \{0,1, \dots, 6\} \longmapsto\ \{\text{Sunday}, \text{Monday}, \dots, \text{Saturday}\}.
$$

In general, any circular granularity relating two linear granularities can be expressed as
$$
  C_{G, H}(z) = \lfloor z/P(B,G) \rfloor\mod P(G,H),
$$
where $H$ is periodic with respect to $G$ with regular mapping and period $P(G,H)$. Table&nbsp;\ref{tab:definitions} shows several circular granularities constructed using minutes as the bottom granularity.

\begin{table}[ht]
\begin{center}
\begin{tabular}{lll}
\toprule
Circular granularity & Expression & Period \\
\midrule
minute-of-hour                               &
  $C_1 = z \mod 60$                     &
  $P_1 = \phantom{99}60$ \\
minute-of-day                                &
  $C_2 = z \mod 60*24$                  &
  $P_2= `r 60*24`$\\
hour-of-day                                  &
  $C_3 = \lfloor z/60\rfloor\mod 24$    &
  $P_3 = \phantom{99}24$ \\
hour-of-week                                 &
  $C_4 = \lfloor z/60\rfloor\mod 24*7$  &
  $P_4= \phantom{9}`r 24*7`$\\
day-of-week                                  &
  $C_5 = \lfloor z/24*60\rfloor \mod 7$ &
  $P_5= \phantom{999}7$\\
\bottomrule
\end{tabular}
\end{center}
\caption{Examples of circular granularities with bottom granularity minutes. Circular granularity $C_i$ relates two linear granularities one of which groups periodically into the other with regular mapping and period $P_i$. Circular granularities can be expressed using modular arithmetic due to their regular mapping.}
\label{tab:definitions}
\end{table}

### Quasi-circular granularities {#sec:quasi-circular-gran-def}

A **quasi-circular** granularity cannot be defined using modular arithmetic because of the irregular mapping. However, they are still formed with linear granularities, one of which groups periodically into the other. \autoref{tab:quasi} shows some examples of quasi-circular granularities.

\begin{table}[ht]
\centering
\begin{tabular}{lr@{~}lr@{~}r}
\toprule
Quasi-circular granularity && Possible period lengths\\
\midrule
$Q_1 =$ day-of-month && $P_1 = 31, 30, 29, 28$\\
$Q_2 =$ hour-of-month && $P_2 = 24\times 31, 24\times 30, 24\times 29, 24\times 28$\\
$Q_3 =$ day-of-year && $P_3 = 366, 365$\\
\bottomrule
\end{tabular}
\caption{Examples of quasi-circular granularities relating two linear granularities with irregular mapping leading to several possible period lengths.}
\label{tab:quasi}
\end{table}

\begin{defn}\label{def:quasicircular}
A \textbf{quasi-circular granularity} $Q_{B, G'}$ is formed when bottom granularity $B$ groups periodically into linear granularity $G'$ with irregular mapping such that the granularities are given by
\begin{equation}\label{eq:quasi}
Q_{B, G'}(z) =
z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert, \quad \text{for}\quad z \in T_{k},
\end{equation}
where
$z$ denotes the index set,
$w$ denotes the index of $G'$,
$R'$ is the number of granules of $G'$ in each period of $(B, G')$,
$T_w$ are the sets of indices of $B$ such that $G'(w) = \bigcup_{z \in T_w}B(z)$,
and $\vert T_w \vert$ is the cardinality of set $T_w$.
\end{defn}

For example, day-of-year is quasi-periodic with either 365 or 366 granules of $B$ (days) within each granule of $G'$ (years). The pattern repeats every $4$ years (ignoring leap seconds). Hence $R'= 4$. $Q_{B, G'}$ is a repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not the same across different granules of $G'$.

### Aperiodic granularities {#sec:aperiodic-gran-def}

Aperiodic linear granularities are those that cannot be specified as a periodic repetition of a pattern of granules as described in Definition \ref{def:periodic}. Aperiodic cyclic granularities capture repetitions of these aperiodic linear granularities.
Examples include public holidays which repeat every year, but there is no reasonably small span of time within which their behavior remains constant. A classic example is Easter (in the western tradition) whose dates repeat only after 5.7 million years [@Reingold2001-kf]. In Australia, if a standard public holiday falls on a weekend, a substitute public holiday will sometimes be observed on the first non-weekend day (usually Monday) after the weekend. Examples of aperiodic granularity may also include school holidays or a scheduled event. All of these are recurring events, but with non-periodic patterns. Consequently, $P_i$ (as given in \autoref{tab:quasi}) are essentially infinite for aperiodic granularities.

\begin{defn}\label{def:aperiodic}
An \textbf{aperiodic cyclic granularity} is formed when bottom granularity $B$ groups into an aperiodic linear granularity $M$ such that the granularities are given by
\begin{equation}\label{eq:aperiodic}
A_{B, M}(z) = \begin{cases}
                  i, & \text{for}\quad z \in T_{i_j} \\
                  0  & \text{otherwise},
                \end{cases}
\end{equation}
where
$z$ denotes the index set,
$T_{i_j}$ are the sets of indices of $B$ describing aperiodic linear granularities $M_{i}$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$, and $M = \bigcup_{i=1}^{n}M_{i}$, $n$ being the number of  aperiodic linear granularities in consideration.
\end{defn}

```{r aperiodic-example, echo=FALSE, fig.cap= "(ref:aperiodic-example)", fig.pos="!htb"}
```

(ref:aperiodic-example) Quasi-circular and aperiodic cyclic granularities illustrated by linear (a) and 
stacked-displays (b) progression of time. The linear display shows the granularities days ($B$), weeks ($G$), semester days ($B'$), and stages of a semester ($M$) indexed over a linear representation of time. The granules of $B'$ is only defined for days when the semester is running. Here a semester spans 18 weeks and 2 days, and consists of 6 stages. It starts with a week of orientation, followed by an in-session period (6 weeks), a break (1 week), the second half of semester (7 weeks), a 1-week study break before final exams, which spans the next 16 days. This distribution of semester days remains relatively similar for every semester. $Q_{B',M}$ with P = 128 is a quasi-circular granularity with repeating patterns, while $A_{B,M}$ is an aperiodic cyclic granularity as the placement of the semester within a year varies from year to year with no fixed start and end dates.

For example, consider the school semester shown in \autoref{fig:aperiodic-example}. Let the linear granularities $M_1$ and $M_2$ denote the teaching and non-teaching stages of the semester respectively. Both $M_1$, $M_2$ and $M = M_{1}\bigcup M_{2}$ denoting the "stages" of the semester are aperiodic with respect to days ($B$) or weeks ($G$). Hence $A_{B, M}$ denoting day-of-the-stage would be an aperiodic cyclic granularity because the placement of the semester within a year would vary across years. Here, $Q_{B', M}$ denoting semester-day-of-the-stage would be a quasi-circular granularity
since the distribution of semester days within a semester is assumed to remain constant over years. Here semester-day is denoted by "sem day" ($B'$) and its granules are only defined for the span of the semesters.

### Relativities {#sec:cyclic-calendar}

The hierarchical structure of time creates a natural nested ordering which can be used in the computation of relative pairs of granularities.

\begin{defn}\label{def:hierarchy}
The nested ordering of linear granularities can be organized into a \textbf{hierarchy table}, denoted as $H_n: (G, C, K)$, which arranges them from lowest to highest in order. It shows how the $n$ granularities relate through $K$, and how the cyclic granularities, $C$, can be defined relative to the linear granularities. Let $G_{\ell}$ and $G_{m}$ represent the linear granularity of order $\ell$ and $m$ respectively with $\ell<m$. Then $K \equiv P(\ell,m)$ represents the period length of the grouping $(G_{\ell}, G_{m})$, if $C_{G_{\ell}, G_{m}}$ is a circular granularity and $K \equiv k(\ell,m)$ represents the operation to obtain $G_{m}$ from $G_{\ell}$, if $C_{G_{\ell}, G_{m}}$ is quasi-circular.
\end{defn}

For example, \autoref{tab:tab-mayan} shows the hierarchy table for the Mayan calendar. In the Mayan calendar, one day was referred to as a kin and the calendar was structured such that 1 kin = 1 day; 1 uinal = 20 kin; 1 tun  = 18 uinal (about a year); 1 katun = 20 tun (20 years) and 1 baktun = 20 katun.

```{r tab-mayan}
```

Like most calendars, the Mayan calendar used the day as the basic unit of time [@Reingold2001-kf]. The structuring of larger units, weeks, months, years and cycle of years, though, varies substantially between calendars. For example,  the French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds", the duration of which is 0.864 common seconds. Nevertheless, for any calendar, a hierarchy table can be defined. Note that it is not always possible to organize an aperiodic linear granularity in a hierarchy table. Hence, we assume that the hierarchy table consists of periodic linear granularities only, and that the cyclic granularity $C_{G(\ell),G(m)}$ is either circular or quasi-circular.

\begin{defn}\label{def:norderup}
The hierarchy table contains \textbf{multiple-order-up} granularities which are cyclic granularities that are nested within multiple levels.
A \textbf{single-order-up} is a cyclic granularity which is nested within a single level. It is a special case of multiple-order-up granularity.
\end{defn}

\noindent In the Mayan calendar (Table \@ref(tab:tab-mayan)), kin-of-tun or kin-of-baktun are examples of multiple-order-up granularities and single-order-up granularities are kin-of-uinal, uinal-of-tun etc.

### Computation

Following the calendar algebra of @Ning_2002 for linear granularities, we can define cyclic calendar algebra to compute cyclic granularities. Cyclic calendar algebra comprises two kinds of operations:
(1) **single-to-multiple** (the calculation of _multiple-order-up_ cyclic granularities from _single-order-up_ cyclic granularities) and (2) **multiple-to-single** (the reverse).

#### Single-to-multiple order-up {-}

Methods to obtain multiple-order-up granularity will depend on whether the hierarchy consists of all circular single-order-up granularities or a mix of circular and quasi-circular single-order-up granularities. Circular single-order-up granularities can be used recursively to obtain a multiple-order-up circular granularity using
\begin{equation} \label{eq:eq7}
C_{G_\ell,G_m}(z)
  = \sum_{i=0}^{m - \ell - 1} P(\ell, \ell+i)C_{G_{\ell+i},G_{\ell+i+1}}(z),
\end{equation}
where $\ell < m - 1$ and $P(i, i) = 1$ for $i=0,1,\dots,m-\ell-1$, and
$C_{B, G}(z)  = z\mod P(B, G)$ as per Equation \eqref{eq:circular-gran}.

For example, the multiple-order-up granularity $C_{\text{uinal},\text{katun}}$ for the Mayan calendar could be obtained using
\begin{align*}
C_{\text{uinal}, \text{baktun}}(z)  &=  C_{\text{uinal}, \text{tun}}(z) + P(\text{uinal}, \text{tun})C_{\text{tun},\text{katun}}(z) + P(\text{uinal},\text{katun})C_{\text{katun}, \text{baktun}}(z) \\
         &=  C_{\text{uinal}, \text{tun}}(z) + 18 \times{ C_{\text{tun},\text{katun}}(z)} + 18 \times 20 \times{C_{\text{katun}, \text{baktun}}(z)}
\end{align*}

, where $z$ is the index of the bottom granularity $kin$.

<!--                &=  \lfloor \frac{z}{20}\rfloor \mod 18 + 18\lfloor  \frac{z}{20 \times18} \rfloor \mod 20 -->
<!--                 + 18 \times20 \lfloor \frac{z}{(20\times18\times20} \rfloor \mod 20 -->

Now consider the case where there is one quasi-circular single order-up granularity in the hierarchy table while computing a multiple-order-up quasi-circular granularity. Any multiple-order-up quasi-circular granularity $C_{\ell, m}(z)$ could then be obtained as a discrete combination of circular and quasi-circular granularities.

Depending on the order of the combination, two different approaches need to be employed leading to the following cases:

  * $C_{G_{\ell},G_{m'}}$ is circular and $C_{G_{m'},G_{m}}$ is quasi-circular
\begin{equation} \label{eq:multifromsingle-quasi1}
C_{G_\ell,G_{m}}(z) = C_{G_{\ell},G_{m'}}(z) + P(\ell, m')C_{G_{m'},G_{m}}(z)
\end{equation}

  * $C_{G_{\ell},G_{m'}}$ is quasi-circular and $C_{G_{m'},G_{m}}$ is circular
\begin{equation} \label{eq:multifromsingle-quasi2}
C_{G_\ell,G_{m}}(z)  = C_{G_{\ell},G_{m'}}(z) + \sum_{w=0}^{C_{G_{m'},G_{m}}(z) -1}(\vert T_{w} \vert)
\end{equation}
where, $T_w$ is such that $G_{m'}(w) = \bigcup_{z \in T_w}G_{\ell}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.

```{r tab-gregorian}
```

For example, the Gregorian calendar (\autoref{tab:tab-gregorian}) has day-of-month as a single-order-up quasi-circular granularity, with the other granularities being circular. Using Equations \eqref{eq:multifromsingle-quasi1} and \eqref{eq:multifromsingle-quasi2}, we then have:
$$
  C_{hour, month}(z) = C_{hour, day}(z) + P(hour, day)*C_{day, month}(z)
$$
$$
  C_{day, year}(z) = C_{day,month}(z) + \sum_{w=0}^{C_{month, year}(z)-1}(\vert T_{w} \vert),
$$
where $T_w$ is such that $month(w) = \bigcup_{z \in T_w}day(z)$.

#### Multiple-to-single order-up {-}

Similar to single-to-multiple operations, multiple-to-single operations involve different approaches for all circular single-order-up granularities and a mix of circular and quasi-circular single-order-up granularities in the hierarchy. For a hierarchy table $H_n: (G, C, K)$ with only circular single-order-up granularities and $\ell_1, \ell_2, m_1, m_2 \in {1, 2, \dots, n}$ and $\ell_2<\ell_1$ and $m_2>m_1$, multiple-order-up granularities can be obtained using Equation \eqref{eq:all-circular-multiple}.
\begin{equation} \label{eq:all-circular-multiple}
C_{G_{\ell_1}, G_{m_1}}(z) = \lfloor C_{G_{\ell_2}, G_{m_2}}(z)/P(\ell_2,\ell_1) \rfloor \mod P(\ell_1, m_1)
\end{equation}
For example, in the Mayan Calendar, it is possible to compute the single-order-up granularity tun-of-katun from uinal-of-baktun, since $C_{tun, katun}(z) = \lfloor C_{uinal, baktun}(z)/18\rfloor \mod 20$.

#### Multiple order-up quasi-circular granularities {-}

Single-order-up quasi-circular granularity can be obtained from multiple-order-up quasi-circular granularity and single/multiple-order-up circular granularity using Equations \eqref{eq:multifromsingle-quasi1} and \eqref{eq:multifromsingle-quasi2}.

## Data structure {#sec:data-structure}

Effective exploration and visualization benefit from well-organized data structures. @wang2020tsibble introduced the tidy "tsibble" data structure to support exploration and modeling of temporal data. This forms the basis of the structure for cyclic granularities. A tsibble comprises an index, optional key(s), and measured variables. An index is a variable with inherent ordering from past to present and a key is a set of variables that define observational units over time. A linear granularity is a mapping of the index set to subsets of the time domain. For example, if the index of a tsibble is days, then a linear granularity might be weeks, months or years. A bottom granularity is represented by the index of the tsibble.

\novspacing

```{r data-structure, fig.pos="!htb"}
library(dplyr)
as.data.frame(matrix(" ", ncol=7, nrow=1)) %>%
  mutate(
    V1 = cell_spec(V1, "latex", background="#dbe3f1"),
    V2 = cell_spec(V2, "latex", background="#dbe3f1"),
    V3 = cell_spec(V3, "latex", background="#dbe3f1"),
    V4 = cell_spec(V4, "latex", background="#fdf2d0"),
    V5 = cell_spec(V5, "latex", background="#fdf2d0"),
    V6 = cell_spec(V6, "latex", background="#fdf2d0"),
    V7 = cell_spec(V7, "latex", background="#fdf2d0"),
  ) %>%
  knitr::kable(format='latex',
               booktabs=TRUE,
               escape=FALSE,
               caption = "The data structure for exploring periodicities in data by including cyclic granularities in the tsibble structure with index, key and measured variables.",
               col.names = c("\\cellcolor[HTML]{dbe3f1}{index}","\\cellcolor[HTML]{dbe3f1}{key}","\\cellcolor[HTML]{dbe3f1}{measurements}","$C_1$","$C_2$","$\\cdots$","$C_{N_C}$")
  ) %>%
  row_spec(0, bold=TRUE,background="#fdf2d0") %>%
  kable_styling(position = "center")
```

\vspacing

All cyclic granularities can be expressed in terms of the index set. \autoref{tab:data-structure} shows the tsibble structure (index, key, measurements) augmented by columns of cyclic granularities. The total number of cyclic granularities depends on the number of linear granularities considered in the hierarchy table and the presence of any aperiodic cyclic granularities. For example, if we have $n$ periodic linear granularities in the hierarchy table, then $n(n-1)/2$ circular or quasi-circular cyclic granularities can be constructed. Let $N_C$ be the total number of contextual circular, quasi-circular and aperiodic cyclic granularities that can originate from the underlying periodic and aperiodic linear granularities. Simultaneously encoding more than a few of these cyclic granularities when visualizing the data overwhelms human comprehension. Instead, we focus on visualizing the data split by pairs of cyclic granularities ($C_i$, $C_j$). Data sets of the form <$C_i$, $C_j$, $v$> then allow exploration and analysis of the measured variable $v$.

### Harmonies and clashes {#sec:synergy}

The way granularities are related is important when we consider data visualizations. Consider two cyclic granularities $C_i$ and $C_j$, such that $C_i$ maps index set to a set $\{A_k \mid k=1,\dots,K\}$ and $C_j$ maps index set to a set $\{B_\ell \mid \ell =1,\dots,L\}$. Here, $A_k$ and $B_\ell$ are the levels/categories corresponding to $C_i$ and $C_j$ respectively. Let $S_{k\ell}$ be a subset of the index set such that for all $s \in S_{k\ell}$, $C_i(s) = A_k$ and $C_j(s) = B_\ell$. There are $KL$ such data subsets, one for each combination of levels ($A_k$, $B_\ell$). Some of these sets may be empty due to the structure of the calendar, or because of the duration and location of events in a calendar.

\begin{defn}\label{def:clash}
A \textbf{clash} is a pair of cyclic granularities that contains empty combinations of categories.
\end{defn}

\begin{defn}\label{def:harmony}
A \textbf{harmony} is a pair of cyclic granularities that does not contain any empty combinations of its categories.
\end{defn}

Structurally empty combinations can arise due to the structure of the calendar or hierarchy. For example, let $C_i$ be day-of-month with 31 levels and $C_j$ be day-of-year with 365 levels. There will be $31\times 365=11315$ sets $S_{k\ell}$ corresponding to possible combinations of $C_i$ and $C_j$. Many of these are empty. For example, $S_{1,5}$ is empty because the first day of the month can never correspond to the fifth day of the year. Hence the pair (day-of-month, day-of-year) is a clash.

Event-driven empty combinations arise due to differences in event location or duration in a calendar. For example, let $C_i$ be day-of-week with 7 levels and $C_j$ be working-day/non-working-day with 2 levels. While potentially all of these 14 sets $S_{k\ell}$ can be non-empty (it is possible to have a public holiday on any day-of-week), in practice many of these will probably have very few observations. For example, there are few (if any) public holidays on Wednesdays or Thursdays in any given year in Melbourne, Australia.

An example of harmony is where $C_i$ and $C_j$ denote day-of-week and month-of-year respectively. So $C_i$ will have 7 levels while $C_j$ will have 12 levels, giving $12\times7=84$ sets $S_{k\ell}$. All of these are non-empty because every day-of-week can occur in every month. Hence, the pair (day-of-week, month-of-year) is a harmony.

### Near-clashes {#sec:near-clashes}

Suppose $C_i$ denotes day-of-year and $C_j$ denotes day-of-week. While any day of the week can occur on any day of the year, some combinations will be very rare. For example, the 366th day of the year will only coincide with a Wednesday approximately every 28 years on average. We refer to these as "near-clashes".

## Visualization {#sec:visualization}

The purpose is to visualize the distribution of the continuous variable ($v$) conditional on the values of two granularities, $C_i$ and $C_j$. Since $C_i$ and $C_j$ are factors or categorical variables, data subsets corresponding to each combination of their levels form a subgroup and the visualization amounts to having displays of distributions for different subgroups. The response variable $(v)$ is plotted on the y-axis and the levels of $C_i (C_j)$ on the x-axis, conditional on the levels of $C_j (C_i)$. This means, carrying out the same plot corresponding to each level of the conditioning variable. This is consistent with the widely used grammar of graphics
which is a framework to construct statistical graphics by relating the data space to the graphic space [@Wilkinson1999-nk, @Wickham2009pk].

<!-- When a data is made up of different groups, it is often informative to compare the distribution of the variables across the groups. -->

### Data summarization

There are several ways to summarize the distribution of a data set such as estimating the empirical distribution or density of the data, or computing a few quantiles or other statistics. This estimation or summarization could be potentially misleading if it is performed on rarely occurring categories (Section&nbsp;\ref{sec:near-clashes}). Even when there are no rarely occurring events, the number of observations may vary greatly within or across each facet, due to missing observations or uneven locations of events in the time domain. In such cases, data summarization should be used with caution as sample sizes will directly affect the accuracy of the estimated quantities being displayed.

<!-- Each plot option has parameters that need to be estimated or summarized from the data. -->
<!-- For example, for histogram, empirical distribution is estimated, for boxplot or its variation, number summaries are computed, for violin or any density-based plots, the density of the variable's distribution needs to be estimated. -->

### Display choices for univariate distributions

<!-- Plotting each point for each combination through dot plots or rug plots is not useful for our data structure since that leads to a blob of points for each subgroup. Further, they are anyway a poor choice for large datasets. Hence, -->

The basic plot choice for our data structure is one that can display distributions. For displaying the distribution of a continuous univariate variable, many options are available. Displays based on descriptive statistics include box plots [@Tukey1977-jx] and its variants such as notched box plots [@McGill1978-hg] or other variations as mentioned in @boxplots. They also include line or area quantile plots which can display any quantiles and not only quartiles like in a boxplot. Plots based on kernel density estimates include violin plots [@Hintze1998-zi], summary plot [@Potter2010-qc], ridge line plots [@R-ggridges], and highest density region (HDR) plots [@Hyndman1996-ft]. The less commonly used Letter-Value plots [@Hofmann2017-sg] is midway between boxplots and density plots. Letter values are order statistics with specific depths, for example, the median ($M$) is a letter value that divides the data set into halves. Each of the next letter values splits the remaining parts into two separate regions so that the fourths ($F$), eighths ($E$), sixteenths ($D$), etc. are obtained. They are useful for displaying the distributions beyond the quartiles especially for large data,
where boxplots mislabel data points as outliers.
One of the best approaches in exploratory data analysis is to draw a variety of plots to reveal information while keeping in mind the drawbacks and benefits of each of the plot choices. For example, boxplots obscure multimodality, and interpretation of density estimates and histograms may change depending on the bandwidth and binwidths respectively. In R package `gravitas` [@R-gravitas], boxplots, violin, ridge, letter-value, line and area quantile plots are implemented, but it is potentially possible to use any plots which can display the distribution of the data.

### Comparison across sub-groups induced by conditioning {#sec:aesthetics}

<!-- While it is often beneficial to compare distributions in this way, there are some caveats that we should be kept in mind while exploring these displays, subject to our data structure as follows: -->

#### Levels {-}

The levels of cyclic granularities affect plotting choices since space and resolution may be problematic with too many levels. A potential approach could be to categorize the number of levels as low/medium/high/very high for each cyclic granularity and define some criteria based on human cognitive power, available display size and the aesthetic mappings. Default values for these categorizations could be chosen based on levels of common temporal granularities like days of the month, days of the fortnight, or days of the week.

#### Synergy of cyclic granularities {-}

The synergy of the two cyclic granularities will affect plotting choices for exploratory analysis. Cyclic granularities that form clashes (Section \@ref(sec:synergy)) or near-clashes lead to potentially ineffective graphs. Harmonies tend to be more useful for exploring patterns. \autoref{fig:allFig}a shows the distribution of half-hourly electricity consumption through letter value plots across months of the year conditional on quarters of the year. This plot does not work because quarter-of-year clashes with month-of-year, leading to empty subsets. For example, the first quarter never corresponds to December.

```{r allFig, fig.pos="!p", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap= "(ref:allFig)", fig.show = 'hold'}
```

(ref:allFig) Distribution of energy consumption displayed as letter value plots, illustrating harmonies and clashes, and how mappings change emphasis: **a** weekday/weekend faceted by quarter-of-year produces a harmony, **b** quarter-of-year faceted by weekday/weekend produces a harmony, **c** month-of-year faceted by quarter-of-year produces a clash, as indicated by the empty sets and white space. Placement within a facet should be done for primary comparisons. For example, arrangement in **a** makes it easier to compare across weekday type (x-axis) within a quarter (facet). It can be seen that in quarter 2, there is more mass occupied the lower tail on the weekends (letter value E corresponding to tail area 1/8) relative to that of the weekdays (letter value D 1/16), which corresponds to more days with lower energy use in this period.

<!-- ; **b** across weekday/weekend faceted by quarter-of-year; **c** across quarter-of-year faceted by weekday/weekend. **a** shows a clash since there are empty combinations. **b** and **c** show harmonies since each quarter include both weekdays and weekends. It can be seen in **b** that for every quarter (mapped to facet), weekend and weekday consumption are fairly similar except for the second quarter. -->
<!-- For example, the letter value D for weekdays mostly corresponds to letter value E for weekends, implying there are lower values of energy consumption during weekdays in the second quarter. This is probably because of lower temperatures in the second quarter compared to the first quarter (summer in Australia). **c** switches the granularities mapped to the x-axis and the facets and helps to compare quarters within weekdays and weekends. For example, the lower D and E letter values are most different across different quarters for weekdays and weekends respectively. -->
<!-- For example, for weekdays the interquartile range of consumption reduces over the year, whereas this pattern is not true for weekends. -->

#### Conditioning variable {-}

When $C_i$ is mapped to the $x$ position and $C_j$ to facets, then the $A_k$ levels are juxtaposed and each $B_\ell$ represents a group/facet. Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups. Hence, in this case, the $A_k$'s are compared against each other within each group. With the mapping of $C_i$ and $C_j$ reversed, the emphasis will shift to comparing $B_\ell$ levels rather than $A_k$ levels. For example, \autoref{fig:allFig}b shows the letter value plot across weekday/weekend partitioned by quarters of the year and \autoref{fig:allFig}c shows the same two cyclic granularities with their mapping reversed. \autoref{fig:allFig}b helps us to compare weekday and weekend within each quarter and \autoref{fig:allFig}c helps to compare quarters within weekend and weekday.

## Applications {#sec:application}

### Smart meter data of Australia {#sec:smartmeter}

Smart meters provide large quantities of measurements on energy usage for households across Australia. One of the customer trials [@smart-meter] conducted as part of the Smart Grid Smart City project in Newcastle and parts of Sydney provides customer level data on energy consumption for every half hour from February 2012 to March 2014. We can use this data set to visualize the distribution of energy consumption across different cyclic granularities in a systematic way to identify different behavioral patterns.

#### Cyclic granularities search and computation {-}

The tsibble object `smart_meter10` from R package `gravitas` [@R-gravitas] includes the variables `reading_datetime`, `customer_id` and `general_supply_kwh` denoting the index, key and measured variable respectively. The interval of this tsibble is 30 minutes.

To identify the available cyclic time granularities, consider the conventional time deconstructions for a Gregorian calendar that can be formed from the 30-minute time index: half-hour, hour, day, week, month, quarter, half-year, year. In this example, we will consider the granularities hour, day, week and month giving six cyclic granularities "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month", read as "hour of the day", etc. To these, we add day-type ("wknd_wday") to capture weekend and weekday behavior. Now that we have a list of cyclic granularities to look at, we can compute them using the results in Section&nbsp;\ref{sec:cyclic-calendar}.

#### Screening and visualizing harmonies {-}

Using these seven cyclic granularities, we want to explore patterns of energy behavior. Each of these seven cyclic granularities can either be mapped to the x-axis or to facets. Choosing $2$ of the possible $7$ granularities, gives $^{7}P_2 = 42$ candidates for visualization. Harmonies can be identified among those $42$ possibilities to narrow the search. \autoref{tab:harmony-tab} shows $16$ harmony pairs after removing clashes and any cyclic granularities with more than $31$ levels, as effective exploration becomes difficult with many levels (Section&nbsp;\ref{sec:aesthetics}).

```{r harmony-tab, echo=FALSE}
```

```{r bothcust, fig.cap = "(ref:bothcust)", fig.pos="!p"}
```

(ref:bothcust) Energy consumption of a single customer shown with different distribution displays, and granularity arrangements: hour of the day; and weekday/weekend. **a** The side-by-side boxplots make the comparison between day types easier, and suggest that there is generally lower energy use on the weekend. Interestingly, this is the opposite to what might be expected. Plots **b**, **c** examine the temporal trend of consumption over the course of a day, separately for the type of day. The area quantile emphasizes time, and indicates that median consumption shows prolonged high usage in the morning on weekdays. The violin plot emphasizes subtler distributional differences across hours: morning use is bimodal.

A few harmony pairs are displayed in \autoref{fig:bothcust} to illustrate the impact of different distribution plots and reverse mapping. For each of \autoref{fig:bothcust}b and c, $C_i$ denotes day-type (weekday/weekend) and $C_j$ is hour-of-day. The geometry used for displaying the distribution is chosen as area-quantiles and violins in \autoref{fig:bothcust}b and c respectively. \autoref{fig:bothcust}a shows the reverse mapping of $C_i$ and $C_j$ with $C_i$ denoting hour-of-day and $C_j$ denoting day-type with distribution geometrically displayed as boxplots.

In \autoref{fig:bothcust}b, the black line is the median, whereas the purple (narrow) band covers the 25th to 75th percentile, the orange (middle) band covers the 10th to 90th percentile, and the green (broad) band covers the 1st to 99th percentile. The first facet represents the weekday behavior while the second facet displays the weekend behavior; energy consumption across each hour of the day is shown inside each facet. The energy consumption is extremely skewed with the 1st, 10th and 25th percentile lying relatively close whereas 75th, 90th and 99th lying further away from each other. This is common across both weekdays and weekends. For the first few hours on weekdays, median energy consumption starts and continues to be higher for longer compared to weekends.

The same data is shown using violin plots instead of quantile plots in \autoref{fig:bothcust}c. There is bimodality in the early hours of the day for weekdays and weekends. If we visualize the same data with reverse mapping of the cyclic granularities (\autoref{fig:bothcust}a), then the natural tendency would be to compare weekend and weekday behavior within each hour and not across hours. Then it can be seen that median energy consumption for the early morning hours is higher for weekdays than weekends. Also, outliers are more prominent in the later hours of the day. All of these indicate that looking at different distribution geometry or changing the mapping can shed light on different aspects of energy behavior for the same sample.

### T20 cricket data of Indian Premier League {#sec:cricket}

Our proposed approach can be generalized to other hierarchical granularities where there is an underlying ordered index. We illustrate this with data from the sport cricket. Cricket is played with two teams of 11 players each, with each team taking turns batting and fielding. This is similar to baseball, wherein the _batsman_ and _bowler_ in cricket are analogous to a batter and pitcher in baseball. A _wicket_ is a structure with three sticks, stuck into the ground at the end of the cricket pitch behind the batsman. One player from the fielding team acts as the bowler, while another takes up the role of the _wicket-keeper_ (similar to a catcher in baseball). The bowler tries to hit the wicket with a _ball_, and the batsman defends the wicket using a _bat_. At any one time, two of the batting team and all of the fielding team are on the field. The batting team aims to score as many _runs_ as possible, while the fielding team aims to successively _dismiss_ 10 players from the batting team. The team with the highest number of runs wins the match.

Cricket is played in various formats and Twenty20 cricket (T20) is a shortened format, where the two teams have a single _innings_ each, which is restricted to a maximum of 20 _overs_. An over will consist of 6 balls (with some exceptions). A single _match_ will consist of 2 innings and a _season_ consists of several matches. Although there is no conventional time component in cricket, each ball can be thought to represent an ordering over the course of the game. Then, we can conceive a hierarchy where the ball is nested within overs, overs nested within innings, innings within matches, and matches within seasons. Cyclic granularities can be constructed using this hierarchy. Example granularities include ball of the over, over of the innings, and ball of the innings. The hierarchy table is given in \autoref{tab:hierarchy-cric}. Although most of these cyclic granularities are circular by the design of the hierarchy, in practice some granularities are aperiodic. For example, most overs will consist of 6 balls, but there are exceptions due to wide balls, no-balls, or when an innings finishes before the over finishes. Thus, the cyclic granularity ball-of-over may be aperiodic.

```{r hierarchy-cric}
```

The Indian Premier League (IPL) is a professional T20 cricket league in India contested by eight teams representing eight different cities in India. The IPL ball-by-ball data is provided in the `cricket` data set in the `gravitas` package for a sample of 214 matches spanning 9 seasons (2008 to 2016).

Many interesting questions could be addressed with the `cricket` data set. For example, does the distribution of total runs depend on whether a team bats in the first or second innings? The Mumbai Indians (MI) and Chennai Super Kings (CSK) appeared in the final playoffs from 2010 to 2015. Using data from these two teams, it can be observed (\autoref{fig:cricex}a) that for the team batting in the first innings there is an upward trend of runs per over, while there is no clear upward trend in the median and quartile deviation of runs for the team batting in the second innings after the first few overs. This suggests that players feel mounting pressure to score more runs as they approach the end of the first innings, while teams batting second have a set target in mind and are not subjected to such mounting pressure and therefore may adopt a more conservative run-scoring strategy.

Another question that can be addressed is if good fielding or bowling (defending) in the previous over affects the scoring rate in the subsequent over? To measure the defending quality, we use an indicator function on dismissals (1 if there was at least one wicket in the previous over, 0 otherwise). The scoring rate is measured by runs per over. \autoref{fig:cricex}b shows that no dismissals in the previous over leads to a higher median and quartile spread of runs per over compared to the case when there has been at least one dismissal in the previous over. This seems to be unaffected by the over of the innings (the faceting variable). This might be because the new batsman needs to "play himself in" or the dismissals lead the (not-dismissed) batsman to adopt a more defensive playstyle. Run rates will also vary depending on which player is facing the next over and when the wicket falls in the previous over.

Here, wickets per over is an aperiodic cyclic granularity, so it does not appear in the hierarchy table. These are similar to holidays or special events in temporal data.

```{r cricex, fig.cap= "(ref:cricex)", warning = FALSE, message = FALSE, fig.height=4, fig.width=8}
```

(ref:cricex) Examining distribution of runs per by innings, over of the innings and number of wickets in previous innings. Plot **a** displays distribution using letter value plots. A gradual upward trend in runs per over can be seen in innings 1, which is not present in innings 2. Plot **b** shows quantile plots of runs per over across an indicator of wickets in the previous over, faceted by current over. When a wicket occurred in the previous over, the runs per over tends to be lower throughout the innings.

## Discussion {#sec:discussion}

Exploratory data analysis involves many iterations of finding and summarizing patterns. With temporal data available at ever finer scales, exploring periodicity can become overwhelming with so many possible granularities to explore. This work provides tools to classify and compute possible cyclic granularities from an ordered (usually temporal) index. We also provide a framework to systematically explore the distribution of a univariate variable conditional on two cyclic time granularities using visualizations based on the synergy and levels of the cyclic granularities.

The `gravitas` package provides very general tools to compute and manipulate cyclic granularities, and to generate plots displaying distributions conditional on those granularities.

A missing piece in the package `gravitas` is the computation of cyclic aperiodic granularities which would require computing aperiodic linear granularities first. A few R packages including `almanac`[@R-almanac] and `gs`[@R-gs] provide the tools to create recurring aperiodic events. These functions can be used with the `gravitas` package to accommodate aperiodic cyclic granularities.

We propose producing plots based on pairs of cyclic granularities that form harmonies rather than clashes or near-clashes. A future direction of work could be to further refine the selection of appropriate pairs of granularities by identifying those for which the differences between the displayed distributions is greatest and rating these selected harmony pairs in order of importance for exploration.

## Acknowledgments {-}

The Australian authors thank the ARC Centre of Excellence for Mathematical and Statistical Frontiers [(ACEMS)](https://acems.org.au/home) for supporting this research. Thanks to [Data61 CSIRO](https://data61.csiro.au/) for partially funding Sayani's research and Dr Peter Toscas for providing useful inputs on improving the analysis of the smart meter application. We would also like to thank Nicholas Spyrison for many useful discussions, sketching figures and feedback on the manuscript. The package `gravitas` was built during the [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/). More details about the package can be found at [sayani07.github.io/gravitas](https://sayani07.github.io/gravitas/). The Github repository,  [github.com/Sayani07/paper-gravitas](https://github.com/Sayani07/paper-gravitas), contains all materials required to reproduce this article and the code is also available online in the supplemental materials. This article was created with `knitr` [@knitr, @R-knitr] and `rmarkdown` [@rmarkdown, @R-rmarkdown].

## Supplementary Materials

**Data and scripts:** Data sets and R code to reproduce all figures in this article (main.R).

\noindent
**R-package:** The ideas presented in this article have been implemented in the open-source R [@R-language] package `gravitas` [@R-gravitas], available from CRAN. The R-package facilitates manipulation of single and multiple-order-up time granularities through cyclic calendar algebra, checks feasibility of creating plots or drawing inferences for any two cyclic granularities by providing list of harmonies and recommends possible visual summaries through factors described in the article. Version 0.1.3 of the package was used for the results presented in the article and is available on Github (https://github.com/Sayani07/gravitas).

### 2021-11-04 Rmd/05-conclusion.Rmd
remove last 2 commits

# Conclusion and future plans {#ch:conclusion}

In this thesis, I present a systematic approach for visualizing and analyzing large temporal data distributions. The building blocks of this framework are presented in this thesis. Chapter \@ref(ch:gravitas) includes tools for computing all cyclic granularities as well as a recommendation system for selecting pairs of granularities that may be effectively investigated together. These temporal granularities may be used to generate data visualizations to search for patterns, associations, and anomalies. However, when there are many granularities that can be constructed for a time period, there will also be too many possible displays to decide which might be the more interesting to display. Chapter \@ref(ch:hakear) presents a framework for selecting displays that are interesting, with the greatest differences between the displayed distributions, and ranking them in order of priority for capturing the most variation. Both of these are used for studying patterns in individual time series or comparing a few time series together. In Chapter \@ref(ch:gracsr), it is extended to allow for the exploration of distributions for many time series at the same time by clustering them based on probability distributions across informative cyclic granularities. Through the use of probability distribution, this technique is more comprehensive in recognizing clusters with recurring patterns over many important granularities and more resilient to noisy, patchy, and uneven length time series.

```{r software-initial, echo = FALSE, cache = FALSE, include = FALSE}
read_chunk('scripts/software.R')
```

## Software development

```{r software-impact}
```

This thesis focuses on integrating research approaches into open source R packages such as **gravitas**, **hakear**, and **gracsr**.

<!-- Figure \@ref(fig:software-ghcommits) gives an overview of my Git commits to these repositories, and Figure \@ref(fig:software-downloads) shows the daily downloads of the packages from the RStudio mirror (one of 90 CRAN mirrors) since they were available on CRAN. -->

<!-- ```{r software-ghcommits, fig.cap = "(ref:software-ghcommits-cap)"} -->
<!-- include_graphics("img/pkg-commits.png") -->
<!-- ``` -->

<!-- (ref:software-ghcommits-cap) Patterns of my package development effort during my PhD years based on Git commits to three repositories, sugrrants, tsibble, and mists. Scatter plots of weekly totals are overlaid with a loess smoother. The **sugrrants** package was the first project with much initial energy, followed by small constant maintenance. The **tsibble** package has been a major project with ongoing constant development and bursts of effort in response to users' feedback. The **mists** package has been a recent intense project. -->

<!-- # ```{r software-downloads, fig.height = 3, fig.cap = "(ref:software-downloads-cap)"} -->
<!-- # ``` -->
<!-- #  -->
<!-- # (ref:software-downloads-cap) Impact of these works (sugrrants and tsibble) as measured by daily downloads (on square root scale) from the RStudio mirror since they landed on CRAN. The **tsibble** package has an increasing trend, suggesting the steady adoption of the new data structure. -->

### gravitas

The **gravitas** package provides very general tools to compute and manipulate cyclic granularities, and to generate plots displaying distributions conditional on those granularities. It can be utilized in non-temporal cases for which a hierarchical structure can be construed similar to time. It is on CRAN. The website (<https://sayani07.github.io/gravitas>) includes full documentation and two vignettes about the package usage.  There has been a grand total of 12K downloads from the RStudio mirror dating from 2020-11-01 to 2021-11-01.

### hakear

The open-source R package `hakear` is available on Github (https://github.com/Sayani07/hakear) to implement ideas presented in Chapter \@ref(ch:hakear). Given a `tsibble` and context granularities of interest, the function `wpd()` provides support for computing the wpd for each cyclic granularities or pair of granularities and `select_harmonies()` chooses the ones with significant patterns and ranks them from highest to lowest wpd.

### gracsr

The open-source R package `gracsr` is available on Github (https://github.com/Sayani07/gracsr) to implement ideas presented in Chapter \@ref(ch:gracsr). The package provides functions to carry out the entire clustering methodology discussed in the paper. It is still a work in progress and has won the ACEMS Business Analytics Prize 2021 with a prize money of AUD 3000, which would be utilized in polishing this package and preparing it for CRAN.

## Future work

### Putting all functionalities on CRAN

I plan to integrate `hakear` with `gravitas` as one R package to systematically exploring few time series, whereas `gracsr` would provide the clustering framework for exploring many time series together. I plan to run it through https://ropensci.org/software-review/ to develop them further for more visibility and efficient usage.

### Scaling up the clustering method to incorporate large uncertainty and improved computational efficiency that comes with large data

With the volume of data projected to grow in the future, potentially leading to increased variability in patterns, research is needed to understand the issues that may arise with the existing methodology and how to adapt our current algorithm. Computational efficiency is also critical when scaling up for analysis of huge data sets, let alone when adding features to existing highly dimensional data structures.

### Comparing our clustering method with other benchmark methods

Testing needs to be carried out with non-hierarchy based clustering methods to see whether these distances perform better with other algorithms. Also, to evaluate how much information is lost when aggregating individual level demand versus distributions.

### Check generalizability of our methods

We provide solutions that are realistically applicable to any temporal data observed more than once per year. We could just verify its value in the electricity smart meter context. With numerous open-source benchmark data sets accessible, it is necessary to test how well the approaches operate in various disciplines.

## Final words

### 2021-11-04 thesis.Rmd
remove last 2 commits

@@
 ---
-title: 'Visualization and analysis of probability distributions of large temporal data'
-author: 'Sayani Gupta'
-degreetype: 'Doctor of Philosophy'
+title: "Visualization and analysis of probability distributions of large temporal data"
+author: "Sayani Gupta"
+degreetype: Doctor of Philosophy
 site: bookdown::bookdown_site
-degrees: 'M.Stat, B.Sc (Stat Hons)'
-bibliography: [bib/gravitas.bib, bib/hakear.bib, bib/gracsr.bib, bib/thesisrefs.bib]
+degrees: M.Stat, B.Sc (Stat Hons)
+bibliography:
+- bib/gravitas.bib
+- bib/hakear.bib
+- bib/gracsr.bib
+- bib/thesisrefs.bib
 link-citations: yes
@@
 
-This work is driven by the need for a conceptual framework to tackle temporal analyses in a data-centric workflow.
-Most research on temporal data focuses on modeling.
-Corresponding software requires very stringently formatted data, but does little in providing guidelines or tools for wrangling raw data into the required format.
-This has led to ad-hoc, and once-off solutions, which this research repairs.
+The research was motivated by the desire to understand the Australian smart meters data, which was collected half-hourly for two years at the household level.
+With temporal data available at ever finer scales, exploring periodicity can become overwhelming with so many possible temporal deconstructions to explore.
+Analysts are expected to comprehensively explore the many ways to view and consider temporal data.
+However, the plethora of choices and the lack of a systematic approach to do so quickly can make the task overwhelming.
 
-There are three original contributions for the temporal data analysis in this research.
-They are grounded in the spirit of exploratory data analysis for time-indexed data.
-The first contribution (Chapter \@ref(ch:gravitas)) is a new technique for visualizing data using a calendar layout.
-It is most useful when the data relates to daily human activity, and patterns can be explored in relation to people's schedules.
-The second contribution (Chapter \@ref(ch:hakear)) is a new data abstraction which streamlines transformation, visualization, and modeling in temporal contexts.
-This "tsibble" object is a data infrastructure forming the foundation of temporal data pipelines.
-The third contribution (Chapter \@ref(ch:gracsr)) is to equip analysts with exploratory and explanatory tools for discovering missing patterns in time, and thus facilitating decisions on appropriate imputation methods for further downstream analysis.
+<!-- exploration may be required to be carried out across both finer and coarser scales to draw useful inferences about the underlying process.
+It is important to be able to navigate through all the temporal deconstructions that accommodate for periodicities to have multiple perspectives of the observed data.
+This idea aligns with the notion of EDA (Tukey 1977) which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
+Slicing and dicing the data in all possible temporal scales as suggested by EDA can be a daunting task as it leads to a -->
+<!-- myriad of possibilities. -->
 
+<!-- Analysts are expected to comprehensively explore the many ways to view and consider temporal data.
+However, the plethora of choices and the lack of a systematic approach to do so quickly can make the task overwhelming -->
+
+<!-- exploring periodicity can become overwhelming with so many possible temporal deconstructions to explore.
+Analysts are expected to comprehensively explore the many ways to view and consider temporal data.
+However, the plethora of choices and the lack of a systematic approach to do so quickly can make the task overwhelming.
+-->
+
+<!-- <WHY> <WHAT> <HOW> -->
+
+This work investigates how time may be dissected, resulting in alternative data segmentation and, as a result, different visualizations that can aid in the identification of underlying patterns.
+The first contribution (Chapter \@ref(ch:gravitas)) describes classes of time deconstructions using linear and cyclic time granularities.
+It provides tools to compute possible cyclic granularities from an ordered (usually temporal) index and also a framework to systematically explore the distribution of a univariate variable conditional on two cyclic time granularities by defining "harmony".
+"harmony" denotes pairs of granularities that could be analyzed together and reduces the search from all possible options.
+This approach is still overwhelming for human consumption because there would still be a huge number of harmonies.
+Hence, the second contribution (Chapter \@ref(ch:hakear)) refines the search of informative granularities by identifying those for which the differences between the displayed distributions are greatest and rating them in order of importance of capturing maximum variation.
+The third contribution (Chapter \@ref(ch:gracsr)) builds upon the first two to provide methods for exploring heterogeneities in repetitive behavior for many households and over multiple granularities.
+It accomplishes this by providing a way to cluster time series based on probability distributions across informative cyclic granularities.
+Although we were motivated by the smart meter example, the problem and the solutions we propose are practically relevant to any temporal data observed more than once per year.
+
+<!-- there is a need to resort to clustering, which has been addressed in  It approaches the problem by computing pairwise distances between time series based on probability distributions over multiple cyclic granularities at once.
+-->
 \newpage
@@
 
+`r if (knitr::is_html_output()) '<!--'`
+
+\newpage
+
 # Declaration {-}
 
-Placeholder
+I have not renumbered sections of submitted or published papers in order to generate a consistent presentation within the thesis.
+
+\textbf{Student name:} Sayani Gupta
+
+\textbf{Student signature:}
+
+```{r signature, out.width = "20%", fig.align = "left", eval = file.exists("img/signature.png")}
+knitr::include_graphics("img/signature.png")
+```
+
+\textbf{Date:} `r Sys.Date()`
+
+`r if (knitr::is_html_output()) '-->'`
 
@@
 
-This thesis would not be possible without many people.
+## supervisors
+Rob and Di are phenomenal leaders in their respective fields, who lead by example.
+I am inspired by their creativity, wisdom, discipline, and dedication to really contribute to the society through their research.
+Thank you for constantly pushing me to improve as a programmer and researcher, and for regularly sharing best practices for conducting research.
+Looking back, I am grateful for how my thoughts and work in statistical computing, graphics, and data analytics in general have evolved over the years.
+I still have a lot to learn, but I am a much more self-reliant and independent researcher than I was when I started.
+Di has been instrumental in exposing me to the potential benefits of effective data visualization.
+As a female researcher, I am encouraged by her willingness to pursue unconventional avenues and have frequently noticed her make a conscious decision to question existing  stereotypes and biases.
+Thank you Di for being a fantastic female role model.
 
-First and foremost, I would like to express my gratitude to my supervisors, Dianne Cook and Rob J Hyndman.
-They are outstanding supervisors, providing excellent insights on statistical computing and graphics during our weekly supervisory meetings.
-I'm deeply impressed by their dynamics, enthusiasm, and visions about open source software and data science, all of which have formed the support and encouragement for me to continue with this non-traditional research path.
+## department and PhD colleagues
 
-I thank Heike Hofmann for hosting me at Iowa State University during the summer of 2018.
-I enjoyed many discussions with Heike, where I received brilliant questions and insightful advice on my research.
-I thank Stuart Lee and Mitchell O'Hara-Wild for countless conversations on software development, that helped to shape this thesis.
-I thank David Frazier for being the best pingpong partner, truly boosting the creative thinking process.
+## R community
 
-I thank Emi Tanaka for sending me PRs for proofreading my thesis on Github, unlocking the power of an open-sourced research compendium.
-I thank Wenjing Wang and Carson Sievert for being so helpful in the early days of my PhD.
-I thank the R community, the tidyverse team, and all NUMBATs for ongoing inspiration.
+## friends
 
-Last but not least, I'd like to thank my mother for her unceasing love.
+PhD in Covid has been harder and I would not have been able to make it through this if not for the emotional support of my friends and family.
+Thank you Puwasala, Sium for being an epitome of kindness and acting as a pillar of support.
+Thank you Tushar, Samarpita and Nairita for always motivating me and believing in me when I didn't believe in myself.
+Thank you Ian for always putting things in perspective when I lost sight.
+Thank you each one of you for being available, listening to me, and not giving up on me at my lowest.
+Thanks to my housemates Anjali, Surbhi and Dulaji for the fun company and conversations that kept me sane in the covid lockdowns.
+
+## family
+Finally, a big thanks to my family for always being supportive of my choices.
+Thanks to my mum (Nupur Gupta) and dad (Arun Prasad Gupta), from whom I learnt that no matter where you start from, if you persevere and are sincere in your efforts, you can sail through any difficult situation.
+They are my constant cheerleaders for all little and big endeavours.
+Thanks to my brother (Avijit) for always having my back and inspiring me to dream bigger.
+Thank you, Juhi (sister-in-law) for instilling in me the value of organisation in all aspects of life and the importance of prioritizing my physical health from time to time.
+Thank you all for your part in supporting me in this journey.
+I love you.
 
@@
 
-# Data representation, visual and analytical techniques for demystifying temporal missing data {#ch:mists}
+# Clustering time series based on probability distributions across temporal granularities {#ch:gracsr}
+
+Placeholder
+
+## Introduction
+## Clustering methodology {#sec:methodology}
+## Validation {#sec:validation}
+### Data generating processes
+### Data designs  
+#### Individual granularities
+#### Interaction of granularities
+### Visual exploration of findings
+## Case study {#sec:cases}
+### Clustering
+#### JS-based distances
+#### wpd-based distances
+## Conclusion and future work {#sec:conclusion}
+## Acknowledgments {-}
+## Supplementary Materials
 
@@
 ## Software development
-### sugrrants
-### tsibble
-### mists
+### gravitas
+### hakear 
+### gracsr
 ## Future work
-### Process for generating missing data in time
-### Visual methods for temporal data of nesting and crossing interactions
+### Putting all functionalities on CRAN
+### Scaling up the clustering method to incorporate large uncertainty and improved computational efficiency that comes with large data
+### Comparing our clustering method with other benchmark methods
+### Check generalizability of our methods
 ## Final words

### 2021-11-04 scripts/software.R

@@ -1,0 +1,10 @@
+## ---- software-impact
+library(cranlogs)
+library(tidyverse)
+library(lubridate)
+
+my_pkgs <- c("sugrrants", "tsibble", "mists")
+
+## ---- software-downloads
+gravitas_dl <- cran_downloads(my_pkgs[1], from = "2020-01-01", to = "2021-11-01")
+tsibble_dl <- cran_downloads(my_pkgs[2], from = "2018-01-09", to = "2019-09-20")

### 2021-11-04 Rmd/00-c-declaration.Rmd
index and gitignore files update

@@
 
+I hereby declare that this thesis contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.
+
+This thesis includes 3 publications, two of which have been revised and resubmitted and one not yet submitted.
+The core theme of the thesis is "tidy tools for temporal data".
+The ideas, development and writing up of all the papers in the thesis were the principal responsibility of myself, the student, working within the Department of Econometrics and Business Statistics under the supervision of Professor Dianne Cook and Professor Rob J Hyndman.
+
+The inclusion of co-authors reflects the fact that the work came from active collaboration between researchers and acknowledges input into team-based research.
+
+In the case of Chapter 2, 3 and 4 my contribution to the work involved the following:
+
+\begin{table}
+\centering\footnotesize\tabcolsep=0.12cm
+\begin{tabular}{|p{1cm}|p{2cm}|p{1.5cm}|p{3.5cm}|p{3.5cm}|p{1.5cm}|}
+\hline
+\RaggedRight\textbf{Thesis Chapter}  & \RaggedRight\textbf{Publication Title}  & \RaggedRight\textbf{Status (published, in press, accepted or returned for revision)}  & \RaggedRight\textbf{Nature and} {\%} \RaggedRight\textbf{of student contribution} & \RaggedRight\textbf{Co-author name(s) Nature and} {\%} \RaggedRight\textbf{of Co-author’s contribution} & \RaggedRight\textbf{Co-author(s), Monash student Y/N} \\ \hline
+2 & Calendar-based graphics for visualizing people's daily schedules & Returned for revision & 70\%.
+Concept and developing software and writing first draft & (1) Dianne Cook, Concept and input into manuscript 20\% (2) Rob J Hyndman, input into manuscript 10\% & \multicolumn{1}{c|}{N} \\
+\hline
+3 & A new tidy data structure to support exploration and modeling of temporal data & Returned for revision & 80\%.
+Concept and developing software and writing first draft & (1) Dianne Cook, input into manuscript 15\% (2) Rob J Hyndman, input into manuscript 5\% & \multicolumn{1}{c|}{N} \\
+\hline
+4 & Data representation, visual and analytical techniques for demystifying temporal missing data & To be submitted & 80\%.
+Concept and developing software and writing first draft & (1) Dianne Cook, Concept and input into manuscript 15\% (2) Rob J Hyndman, input into manuscript 5\% & \multicolumn{1}{c|}{N} \\
+\hline
+\end{tabular}
+\end{table}
+
 I have not renumbered sections of submitted or published papers in order to generate a consistent presentation within the thesis.
 
-\textbf{Student name:} Sayani Gupta
+\textbf{Student name:} Yiru (Earo) Wang
 

### 2021-11-04 index.Rmd
index and gitignore files update

@@
 degrees: 'M.Stat, B.Sc (Stat Hons)'
-bibliography: [bib/gravitas.bib, bib/hakear.bib, bib/gracsr.bib, bib/thesisrefs.bib]
+bibliography: [bib/gravitas.bib, bib/tsibble.bib, bib/mists.bib, bib/thesisrefs.bib]
 link-citations: yes

### 2021-11-04 Rmd/00-d-acknowlegements.Rmd
running without references

@@
 
-## supervisors
+# supervisors
 Rob and Di are phenomenal leaders in their respective fields, who lead by example.
@@
 
-## department and PhD colleagues
+# department and PhD colleagues
 
-## R community
+# R community
 
-## friends
+# friends
 
@@
 
-## family
+# family
 Finally, a big thanks to my family for always being supportive of my choices.

### 2021-11-04 Rmd/01-intro.Rmd
running without references

@@
 
-To tame the complexity of time, breaking it into years, months, weeks, days and so on in a hierarchical manner is a common way to relate data to time.
+<!-- To tame the complexity of time, breaking it into years, months, weeks, days and so on in a hierarchical manner is a common way to relate data to time.
 Such dicrete human made abstractions of time can be thought of as time granularities.(Aigner et al., 2008).
+-->
 
@@
 
-The motivation for this work comes from the desire to provide methods to better understand large quantities of measurements on energy usage reported by smart meters
-in household across Australia, and indeed many parts of the world. Smart meters currently
-provide half-hourly use in kwh for each household, from the time that they were
-installed, some as early as 2012. Households are distributed geographically, and have
-different demographic properties such as the existence of solar panels, central heating or air conditioning.
+The motivation for this work comes from the desire to provide methods to better understand large quantities of measurements on energy usage reported by smart meters in household across Australia, and indeed many parts of the world.
+Smart meters currently provide half-hourly use in kwh for each household, from the time that they were installed, some as early as 2012.
+Households are distributed geographically, and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
 The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
-It is common to see aggregates of usage across households, total kwh used each half hour
-by state, for example, because energy companies need to understand maximum loads
-that they will have to plan ahead to accommodate. But studying overall energy use hides
-the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
+It is common to see aggregates of usage across households, total kwh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
+But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
 
@@
 Although the motivation came through the smart meter example, this is a problem that is relevant to any temporal data observed more than once per year.
-
-Temporal data analysis has assumed that the entry point to data analysis is at a model-ready data format, which provides little organization or conceptual oversight on how one should get the wild data into a tamed state.
-This mind-set is related to a long-held belief that exploratory data analysis is a highly ad hoc statistical area, impossible to teach or formalize.
-However, the **tidyverse** framework implemented in the statistical software R [@R-base], as originating in @wickham2014tidy, fundamentally overturns this thinking.
-Data plots and data wrangling, which the "tidy data" conceptualization supports, can be formally described using an abstract grammar.
-The grammar of graphics and data manipulation, as implemented in the **ggplot2** [@R-ggplot2] and **dplyr** [@R-dplyr] R packages respectively, form the core of the **tidyverse** suite of tools.
-My contributions extend the **tidyverse** way of thinking to the temporal domain, by providing tidy tools, built as R packages, for supporting fluent workflow in temporal data analysis.
 
@@
 The thesis is structured as follows.
-Chapter \@ref(ch:calendar-vis) provides details of the calendar plot, algorithm and applications.
-This is implemented in the R package **sugrrants**.
-Chapter \@ref(ch:tsibble) explains the new data abstraction--tsibble--and illustrates how it can be used to for the basis of exploratory methods, visualization and modeling of temporal data.
-This is available as the R package **tsibble**.
-Chapter \@ref(ch:mists) describes new procedures for exploring temporal missing patterns, and assisting in handling missing values prior to modeling.
-This is in the developing R package **mists**.
-
-Chapter \@ref(ch:conclusion) summarizes the software tools developed for the work and their impact, and discusses some future plans.
+Chapter \@ref(ch:gravitas) provides details of the cyclic granularities, different classes, and computation, and also its usage in exploratory time series analysis through applications.
+This is implemented in the R package **gravitas**.
+Chapter \@ref(ch:hakear) provides guidance on how to choose significant cyclic granularities, which are likely to have interesting patterns across its categories.
+This is available as the R package **hakear**.
+The chapter @ref (ch: gracsr) provides similarity measures for comparing multiple time series.This is in the developing R package **gracsr**.
+Chapter \@ref(ch:conclusion) summarizes the software tools developed for the work, and discusses some future plans.

### 2021-11-04 Rmd/04-gracsr.Rmd
running without references

@@
-# Data representation, visual and analytical techniques for demystifying temporal missing data {#ch:mists}
+# Clustering time series based on probability distributions across temporal granularities {#ch:gracsr}
+
+With more and more time series data being collected at much finer temporal resolution, for a longer length of time, and for a larger number of individuals/entities, time series clustering research is getting a lot of traction.
+Long, noisy, patchy, uneven, and asynchronous time series are common in many fields, limiting similarity searches or lowering method efficiency when clustering is based on a distance metric.
+In this work, we suggest two approaches for obtaining similarity between time series based on probability distributions over cyclic temporal granularities for distance-based clustering approaches.
+Cyclic granularities like hour-of-the-day, work-day/weekend, month-of-the-year and so on, are useful for finding repeated patterns in the data.
+Looking at probability distributions across cyclic granularities serves two purposes: (a) "Probability distributions" characterise the inherent temporal data structure of these large unequal-length time series and are robust to missing or noisy data.
+(b) Using probability distributions over "cyclic granularities" ensures small pockets of similar "repeated" behaviors.
+Our method is capable of producing useful clusters, as demonstrated by testing on validation data designs and a sample of residential smart meter consumers.
+
+```{r external3, include = FALSE, cache = FALSE}
+knitr::read_chunk('scripts/gracsr.R')
+```
+
+```{r load-lib}
+
+```
+
+```{r mytheme-application}
+```
+
+```{r theme_characterisation}
+```
+
+```{r theme-validation}
+
+```
+
+## Introduction
+
+<!-- time series clustering and its challenges -->
+Time-series clustering is the process of unsupervised partitioning of $n$ time-series data into $k$ ($k<n$) meaningful groups such that homogeneous time-series are grouped together based on a certain similarity measure.
+The time-series features, length of time-series, representation technique, and, of course, the purpose of clustering time-series all influence the suitable similarity measure or distance metric to a meaningful level.
+The three primary methods to time series clustering (@liao2005clustering) are algorithms that operate directly with distances or raw data points in the time or frequency domain (distance-based), with features derived from raw data (feature-based), or indirectly with models constructed from raw data (model-based).
+The efficacy of distance-based techniques is highly dependent on the distance measure utilized.
+Defining an appropriate distance measure for the raw time series may be a difficult task since it must take into account noise, variable lengths of time series, asynchronous time series, different scales, and missing data.
+Commonly used Distance-based similarity measures as suggested by a decade review of time series clustering approaches (@Aghabozorgi2015-ct) are Euclidean, Pearson's correlation coefficient and related distances, Dynamic Time Warping, Autocorrelation, Short time series distance, Piecewise regularization, cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances (@liao2007clustering) but on a vector time series data.
+Among these alternatives, Euclidean distances have high performance but need the same length of data over the same period, resulting in information loss regardless of whether it is on raw data or a smaller collection of features.
+DTW works well with time series of different lengths (@corradini2001dynamic), but it is incapable of handling missing observations.
+Surprisingly, probability distributions, which may reflect the inherent temporal structure of a time series have not been considered in determining time series similarity.
+
+\noindent We consider the problem of clustering a large number of univariate time series of continuous values which are available at fine temporal scales, being motivated by the residential smart meter data.
+These time series data are long (with more and more data collected at finer resolutions), are asynchronous, with varying time lengths for different houses and missing observations and characterized by noisy and patchy behavior that can quickly become overwhelming and hard to interpret, requiring summarizing the large number of customers into pockets of similar energy behavior.
+Choosing probability distributions seem to be a natural way to analyze these types of data sets since they are robust to uneven length, missing data, or noise.
+This paper proposes two approaches for obtaining pairwise similarities based on Jensen-Shannon distances between probability distributions across significant cyclic granularities.
+Cyclic temporal granularities [@Gupta2021-hakear], which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can be useful for measuring repetitive patterns in large univariate time series data.
+The resulting clusters are expected to group customers that have similar repetitive behaviors across each of the interesting cyclic granularities.
+Below are some of the benefits of our method:
+
+- When using probability distributions, data does not have to be the same length or observed during the exact same time period (unless there is a structural pattern);
+
+<!-- (for example, some customers observed only for holidays, others observed over non holidays or year of observation is different which can lead to technological advancement.) -->
+
+- Jensen-Shannon distances evaluate the distance between two distributions rather than raw data, which is less sensitive to missing observations and outliers than other conventional distance methods;
+
+- While most clustering algorithms produce clusters similar across just one temporal granularity, this technique takes a broader approach to the problem, attempting to group observations with similar distributions across all interesting cyclic granularities;
+
+- It is reasonable to define a time series based on its degree of trend and seasonality, and to take these characteristics into account while clustering it.
+The modification of the data structure by taking into account probability distributions across cyclic granularities assures that there is no trend and that seasonal variations are handled independently.
+As a result, there is no need to de-trend or de-seasonalize the data before applying the clustering method.
+For similar reasons, there is no need to exclude holiday or weekend routines.
+
+_Background and motivation_
+
+Large spatio-temporal data sets, both from open and administrative sources, offer up a world of possibilities for research.
+One such data sets for Australia is the Smart Grid, Smart City (SGSC) project (2010–2014) available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) The project provides half-hourly data of over 13,000 household electricity smart meters distributed unevenly from October 2011 to March 2014.
+Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.
+Households vary in size, location, and amenities such as solar panels, central heating, and air conditioning.
+The behavioral patterns differ amongst customers due to many temporal dependencies.
+Some households use a dryer, while others dry their clothes on a line.
+Their weekly profile may reflect this.
+They may vary monthly, with some customers using more air conditioners or heaters than others, while having equivalent electrical equipment and weather circumstances.
+Some customers are night owls, while others are morning larks.
+Day-off energy use varies depending on whether customers stay home or go outside.
+Age, lifestyle, family composition, building attributes, weather, availability of diverse electrical equipment, among other factors, make the task of properly segmenting customers into comparable energy behavior a fascinating one.
+This challenge is worsened when all we know about our consumers is their energy use history (@Ushakova2020-rl).
+To safeguard the customers' privacy, it is probable that such information is not accessible.
+Also, energy suppliers may not always update client information, such as property features, in a timely manner.
+Thus, there is a growing need to have research that examines how much energy usage heterogeneity can be found in smart meter data and what are some of the most common power consumption patterns, rather than explaining why consumption differs.
+
+_Related work_
+
+A multitude of papers have emerged around smart meter time series clustering for deepening our knowledge of consumption patterns.
+@Tureczek2017-pb conducted a systematic study of over $2100$ peer-reviewed papers on smart meter data analytics.
+None of the $34$ articles chosen for their emphasis use Australian smart meter data.
+The most often used algorithm is K-Means.
+Using K-Means without considering time series structure or correlation results in inefficient clusters.
+Principal Component Analysis (PCA) or Self-Organizing Maps (SOM) eliminate correlation patterns and decrease feature space, but lose interpretability.
+To reduce dimensionality, several studies use principal component analysis or factor analysis to pre-process smart-meter data before clustering (@Ndiaye2011-pf).
+Other algorithms utilized in the literature include k-means variants, hierarchical approaches, and greedy k-medoids.
+Time series data, such as smart meter data, are not well-suited to any of the techniques mentioned in @Tureczek2017-pb.
+Only one study [@ozawa2016determining] identified time series characteristics using Fourier transformation, which converts data from time to frequency and then uses K-Means to cluster by greatest frequency.
+@Motlagh2019-yj suggests that the time feature extraction is limited by the type of noisy, patchy, and unequal time-series common in residential and addresses model-based clustering by transforming the series into other other objects such as structure or set of parameters which can be more easily characterized and clustered.
+[@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
+@Melnykov2013-sp discusses how outliers, noisy observations and scattered observations can complicate estimating mixture model parameters and hence the partitions.
+To our knowledge, none of the methods focus on exploring heterogeneity in repetitive behaviors based on the dynamics of multiple temporal dependencies using probability distributions.
+
+<!-- Given none of the methods use probability distributions across cyclic granularity to explore heterogeneity in different repetitive behaviors, we present a way to do that by presenting similarity through probability distributions across cyclic granularities.
+-->
+
+The remainder of the paper is organized as follows: Section&nbsp;\ref{sec:methodology} provides the clustering methodology.
+Section&nbsp;\ref{sec:validation} shows data designs to validate our methods.
+Section&nbsp;\ref{sec:cases} discusses the application of the method to a subset of the real data.
+Finally, we summarize our results and discuss possible future directions in Section&nbsp;\ref{sec:conclusion}.
+
+<!-- A typical clustering technique includes the following steps: (a) establishing distance (dissimilarity) and similarity through feature or model extraction and selection; and (b) selecting the clustering algorithm design.
+Distance measures could be time-domain based or frequency-domain (fast Fourier transform).
+-->
+
+<!-- A model-based clustering works by transforming the series into other other objects such as structure or set of parameters which can be more easily characterized and clustered (@Motlagh2019-yj).
+[@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
+The essential temporal characteristics of the curves are defined or extracted using feature-based clustering.
+-->
+
+<!-- This work -->
+<!-- This is similar to a stochastic approach (@Motlagh2019-yj) to clustering, which proposes interpreting electricity demand as a random process and extracting time-series characteristics, or a model of the series, to enable unsupervised clustering.
+Unsupervised clustering is only as good as the features that are extracted/selected or the distance metrics that were utilized.
+Well-designed additional features may collect characteristics that default features cannot.
+Based on the underlying structure of the temporal data, this article offers new distance metric and features for clustering and applies them to actual smart-meter data.
+Firstly, the distance metric is based on probability distribution, which in our knowledge is the first attempt to cluster smart meter data using probability distributions.
+These recorded time series are asynchronous, with varying time lengths for different houses and missing observations.
+Taking probability distributions helps to deal with such data, while helping with dimension reduction in one hand but not losing too much information due to aggregation.
+Secondly, we recognise that most clustering algorithms only provide hourly energy profiles during the day, but this approach provides a wider approach to the issue, seeking to group consumers with similar shapes over all important cyclic granularities.
+Since cyclic granularities are considered instead of linear granularities, clustering would group customers that have similar repetitive behavior across more than one cyclic granularities across which patterns are expected to be significant.
+-->
+
+<!-- common similarity measures -->
+
+<!-- electricity data structure -->
+
+<!-- common similarity measures used there -->
+
+<!-- lit review -->
+
+## Clustering methodology {#sec:methodology}
+
+<!-- In contrast to models, a feature-based strategy is used to explicitly define or automatically extract the curves’ key time features, for instance by application of PCA on the daily curves [ --
+
+<!-- Most papers discussed in Tureczek2017-pb fail to accept smart meter readings as time series data, a data type which contains a temporal component.
+The omission of the essential time series features in the analysis leads to the application of methods that are not designed for handling temporal components.
+K-Means ignores autocorrelation, unless the input data is pre-processed.
+The clusters identified in the papers are validated by a variety of indices, with the most prevalent -->
+<!-- being the cluster dispersion index (CDI) [22–24], the Davies–Bouldin index (DBI) [25,26] and the mean index adequacy (MIA) [8,13]. -->
+
+<!-- The data set solely contains readings from smart meters and no information about the consumers' specific physical, geographical, or behavioral attributes.
+As a result, no attempt is made to explain why consumption varies.
+Instead, this work investigates how much energy usage heterogeneity can be found in smart meter data and what some of the most common electricity use patterns are.
+-->
+
+<!-- Because of this importance, countless approaches to estimate time series similarity have been proposed.  -->
+
+The foundation of our method is unsupervised clustering algorithms based exclusively on the time-series data.
+The proposed methodology aims to leverage the intrinsic data structure hidden within cyclic temporal granularities.
+The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
+<!--So going back to the smart meter example, $f_i(v)$ is the distribution of customer $i$ and $v$ is electricity demand.--> In this work, instead of considering the probability distributions of the linear time series, we assume it across different categories of any cyclic granularity.
+We can consider categories of an individual cyclic granularity ($A$) or combination of categories for two interacting granularities ($A*B$) to have a distribution, where $A, B$ are defined as $A = \{a_j: j=1, 2, \dots J\}$ and $B = \{b_k: k = 1, 2, \dots K\}$.
+For example, let us consider two cyclic granularities $A$ and $B$ representing hour-of-day and day-of-week.
+Then $A = \{0, 1, 2, \dots, 23\}$ and $B = \{Mon, Tue, Wed, \dots, Sun\}$.
+In case individual granularities are considered, there are $J = 24$ distributions of the form $f_{i,j}(v)$ or $K = 7$ distributions of the form $f_{i,k}(v)$ for each customer $i$.
+In case of interaction, $J*K=168$ distributions of the form $f_{i,j, k}(v)$ could be conceived for each customer $i$.
+Hence clustering these customers is equivalent to clustering these collections of conditional univariate probability distributions.
+Towards this goal, we need to decide how to measure similarities between collections of univariate probability distributions.
+There are multiple ways to measure similarities depending on the aim of the analysis.
+<!--This paper focuses on looking at the (dis) similarity between underlying distributions that may have resulted in different patterns across different cyclic temporal granularities, that eventually have resulted in the (dis) similarity between time series.
+--> This paper considers a two approaches for finding distances between time series.
+Both of these approaches may be useful in a practical context and, depending on the data set, may or may not propose the same customer classification.
+The obtained distances could be fed into a clustering algorithm to break large data sets into subgroups that can then be analyzed separately.
+<!--These clusters may be commonly associated with real-world data segmentation.
+However, since the data is unlabeled a priori, more information is required to corroborate this.--> The methodology is explained in the Figure \ref{fig:flowchart} and each element of the pipeline is discussed.
+
+```{r flowchart, fig.cap = "Flow chart illustrating the pipeline for methodology", out.width="100%"}
+knitr::include_graphics("img/flowchart.png")
+
+```
+
+- _Find significant granularities or harmonies_
+
+[@Gupta2021-hakear] proposes a method for choosing significant cyclic granularities and harmonies (interacting granularities that can be studied together) [@Gupta2021-gravitas], which is used in this work.
+We define "significant" granularities as those with significant distributional differences across categories.
+It is preferable to consider those since it is more probable that there will be some intriguing repeated behavior worth investigating.
+It should be noted that all of the observations in the study may not have the same set of important granularities.
+The following is a method for generating a list ($S_c$) of significant granularities for all observations:
+
+ (a) Remove granularities from the comprehensive list that are insignificant for all observations.
+
+ (b) select only the granularities that are significant for the majority of observations.
+
+There will be observations in both cases where one or a few selected granularities are boring.
+Even in that case, having this group of observations with no interesting patterns at a granularity that regularly discovers patterns may be valuable.
+If, on the other hand, the granularities under examination are truly important for a group of data, unique patterns may be identified while clustering them.
+
+- _Data transformation_
+
+Observations often have a somewhat skewed time series distribution and their ranges might vary greatly.
+Statistical transformations are employed to bring all of them to the same range or normalize each observation.
+This is important because we are not interested in trivial clusters that vary in magnitudes, but rather in uncovering comparable patterns of distributional differences between categories.
+For the JS-based approaches, two data transformation techniques are utilized viz, Normal-Quantile Transform (NQT) and Robust scaling.
+NQT is a built-in transformation for computing $wpd$, which is the foundation of wpd-based distances.
+
+_Robust scaling_ The normalized $i^{th}$ observation is denoted by $v_{norm} = \frac{v_t - q_{0.5}}{q_{0.75}-q_{0.25}}$, where $v_t$ is the actual value at the $t^{th}$ time point and $q_{0.25}$, $q_{0.5}$ and $q_{0.75}$ are the $25^{th}$, $50^{th}$ and $75^{th}$ percentile of the time series for the $i^{th}$ observation.
+$v_{norm}$ has zero mean and median, as well as a standard deviation of one, with the outliers having same relative connections to other values.
+
+_Normal-Quantile transform_ The raw data for all observations is individually normal-quantile transformed (NQT) [@Krzysztofowicz1997-bv], so that the transformed data follows a standard normal distribution.
+NQT will make the skewed distributions bell-shaped.
+As a result, determining which raw distribution was used is difficult using the modified distribution.
+Multimodality is also concealed or reversed.
+As a result, while displaying transformed data using NQT, one must exercise caution.
+But NQT is a useful transformation that often improves clustering performance.
+
+<!-- NQT, however, is an useful transformation which often improves the clustering performance.
+Since our variables are cyclic granularities which are derived from linear granularities, NQT ensures that the conditional distribution across each variable is also normalised.
+-->
+
+<!-- is not a problem for implementation of this methodology as distributions are characterized by quantiles and the order of the quantiles is reserved under NQT. -->
+
+<!-- First, the original data is ranked in ascending order and the probabilities $P(Y<=y(i)) = i/(n+1)$ are attached to $y(i)$, in terms of their ranking order.
+A NQT based transformation is applied by computing from a standard normal distribution a variable $\eta(i)$, which corresponds to the same probability $P(\eta< \eta(i)) = i/n+1$.By doing this, the new variables $\eta(i)$ will be marginally distributed according to standard Normal, N(0,1).
+-->
+
+<!-- It is worth noting that when studying these similarities, a variety of objectives may be pursued.
+One objective could be to group time series with similar shapes over all relevant cyclic granularities.
+In this scenario, the variation in customers within each group is in magnitude rather than shape, while the variation between groups is only in shape.
+There -->
+<!-- are distance measures are used for shape-based clustering [Ding et al. 2008; Wang -->
+<!-- et al.
+2013] and many more but none of them look at the probability distributions while computing similarity.
+Moreover, most distance measures offer similar shape across just one dimension.
+For example, we often see "similar" daily energy profiles across hours of the day, but we suggest a broader approach to the problem, aiming to group consumers with similar distributional shape across all significant cyclic granularities.
+Another purpose of clustering could be to group customers that have similar differences in patterns across all major cyclic granularities, capturing similar jumps across categories regardless of the overall shape.
+For example, in the first goal, similar shapes across hours of the day will be grouped together, resulting in customers with similar behavior across all hours of the day, whereas in the second goal, any similar big-enough jumps across hours of the day will be clubbed together, regardless of which hour of the day it is.
+Both of these objectives may be useful in a practical context and, depending on the data set, may or may not propose the same customer classification.
+Depending on the goal of clustering, the distance metric for defining similarity would be different.
+These distance metrics could be fed into a clustering algorithm to break large data sets into subgroups that can then be analyzed separately.
+These clusters may be commonly associated with real-world data segmentation.
+However, since the data is unlabeled a priori, more information is required to corroborate this.
+This section presents the work flow of the methodology:  -->
+
+- _Data pre-preprocessing_
+
+We start with a "tsibble" (@wang2020tsibble) data structure with an index variable representing inherent ordering from past to present and a key variable that defines observational units over time.
+The measured variable for an observation is a time-indexed sequence of values.
+This sequence, however, could be shown in several ways.
+A shuffle of the raw sequence may represent hourly consumption throughout a day, a week, or a year.
+Cyclic granularities can be expressed in terms of the index set in the "tsibble" data structure.
+But the data structure changes while transporting from linear to cyclic scale of time as multiple observations now correspond to each category and induce a probability distribution.
+Directly computing Jensen-Shannon distances between the entire probability distributions can be very costly.
+Hence, in this paper, quantiles are chosen to characterize the probability distributions.
+So, in the final data structure, each category of a cyclic granularity corresponds to a list of numbers which is essentially a few chosen quantiles.
+
+- _Distance metrics_
+
+<!-- Considering each individual or combined categories of cyclic granularities to have an underlying distribution lead to a collection of conditional distributions for each customer $i$. -->
+
+The total (dis) similarity between each pair of customers is obtained by combining the distances between the collections of conditional distributions.
+This needs to be done in a way such that the resulting metric is a distance metric, and could be fed into the clustering algorithm.
+Two types of distance metric is considered:
+
+<!-- The choice of distance measures is a critical step in clustering. It defines how the similarity of two elements (x, y) is calculated. -->
+
+**JS-based distances**
+
+This distance metric considers two time series to be similar if the distributions of each category of an individual cyclic granularity or combination of categories for interacting cyclic granularities are similar.
+In this study, the distribution for each category is characterized using deciles (can potentially consider any list of quantiles), and the distances between distributions are calculated using the Jensen-Shannon distances (@Menendez1997-in), which are symmetric and thus could be used as a distance measure.
+
+\noindent The sum of the distances between two observations $x$ and $y$ in terms of cyclic granularity $A$ is defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity $A$) or $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$.
+After determining the distance between two series in terms of one granularity, we must combine them to produce a distance based on all significant granularities.
+When combining distances from individual $L$ cyclic granularities $C_l$ with $n_l$ levels, $$S_{x, y} = \sum_lS^{C_l}_{x,y}/n_l$$ is employed, which is also a distance metric since it is the sum of JS distances.
+This approach is expected to yield groups, such that the variation in observations within each group is in magnitude rather than distributional pattern, while the variation between groups is only in distributional pattern across categories.
+
+**wpd-based distances**
+
+Compute weighted pairwise distances $wpd$ (@Gupta2021-hakear) for all considered granularities for all observations.
+$wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction and is estimated by the maximum pairwise distances between consecutive categories normalized by appropriate parameters.
+A higher value of $wpd$ indicates that some interesting pattern is expected, whereas a lower value would indicate otherwise.
+
+\noindent Once we have chosen $wpd$ as a relevant feature for characterizing the distributions across one cyclic granularity, we have to decide how we combine differences between the multiple features (corresponding to multiple granularities) into a single number.
+The Euclidean distance between them is chosen, with the granularities acting as variables and $wpd$ representing the value under each variable.
+With this approach, we should expect the observations with similar $wpd$ values to be clustered together.
+Thus, this approach is useful for grouping observations that have a similar significance of patterns across different granularities.
+Similar significance does not imply a similar pattern, which is where this technique varies from JS-based distances, which detect differences in patterns across categories.
+
+<!-- Consider a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$.
+Each household consists of a $J*K$ distributions one harmony.
+We compute the distributional difference -->
+<!-- between $(A, B)$ for the $s^{th}$ household using $wpd_{{s}}(A,B)$. -->
+<!-- $wpd_{{s}}(A,B)$ denotes the normalized weighted-pairwise distributional distances between $(A, B)$ and is a feature which measures distributional difference between harmonies.
+If we have $H_{N_C}$ harmonies in the harmony table, then for each household we have a vector of $wpd_{{s}}$ of $H_{N_C}$ elements with each element corresponding to one harmony.
+We aim to have pockets of households showing similar periodic behavior by considering $wpd$ values for different harmonies and some time series features.
+The features should also characterize probability distributions of different household.
+-->
+
+<!-- grouping probability distributions across a harmony.
+This clustering algorithm is adopted to remove or appropriately adjust for auto correlation and unequal length in the data.
+The method could be further extended by clustering probability distributions conditional on one or more cyclic granularities.
+The following are some of the advantages of our proposed method.
+-->
+
+- _Clustering algorithm_
+
+With a way to obtain pairwise distances, any clustering algorithm can be employed that supports the given distance metric as input.
+A good comprehensive list of algorithms can be found in @Xu2015-ja based on traditional ways like partition, hierarchy, or more recent approaches like distribution, density, and others.
+We employ agglomerative hierarchical clustering in conjunction with Ward's linkage.
+Hierarchical cluster techniques fuse neighboring points sequentially to form bigger clusters, beginning with a full pairwise distance matrix.
+The distance between clusters is described using a "linkage technique".
+This agglomerative approach successively merges the pair of clusters with the shortest between-cluster distance using Ward's linkage method.
+
+<!--Hierarchical algorithms are one of the most widely used, can operate with data of any shape, has reasonable scalability, and the number of clusters is not needed as a parameter.-->
+
+- _Characterization of clusters_
+
+\noindent Cluster characterization is an important element of cluster analysis.
+@Cook2007-qe provides several methods for characterizing clusters.
+_Parallel coordinate plots_ (@wegman1990hyperdimensional), _Scatterplot matrix_, _Displaying cluster statistics_ (@dasu2005grouping), _MDS_ (@borg2005modern), PCA, _t-SNE_(@R-tsne), _Tour_ (@wickham2011tourr) are some of the graphical approaches used in this study.
+A Parallel Coordinates Plot features parallel axes for each variable and each axis is linked by lines.
+Changing the axes may reveal patterns or relationships between variables for categorical variables.
+However, for categories with cyclic temporal granularities, preserving the underlying ordering of time is more desirable.
+Displaying cluster statistics is useful when we have larger problems and it is difficult to read the parallel coordinate plots due to congestion.
+All of MDS, PCA and t-SNE use a distance or dissimilarity matrix to construct a reduced-dimension space representation, their goals are diverse.
+Multidimensional scaling (@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space.
+The t-SNE embedding will compress data points that are close in high-dimensional space.
+Tour is a collection of interpolated linear projections of multivariate data into lower-dimensional space.
+The cluster characterization approach varies depending on the distance metric used.
+Parallel coordinate plots, scatter plot matrices, MDS or PCA are potentially useful ways to characterize clusters using wpd-based distances.
+For JS-based distances, plotting cluster statistics is beneficial for characterization and variable importance could be displayed through parallel coordinate plots.
+
+<!-- CUT IT SHORT -->
+
+<!-- (a) _Parallel coordinate plots_ (@wegman1990hyperdimensional) are often used to visualize high-dimensional and multivariate data, allowing visual grouping and pattern detection.
+A Parallel Coordinates Plot features parallel axes for each variable.
+Each axis is linked by lines.
+Changing the axes may reveal patterns or relationships between variables for categorical variables.
+However, for categories with cyclic temporal granularities, preserving the underlying ordering of time is more desirable.
+-->
+
+<!-- (b) _Scatterplot matrix_ contains pairwise scatter plots of the p variables.
+Pairwise scatter plots are useful for figuring out how variables relate to each other and how factors determine the clustering.
+-->
+
+<!-- (c) _Displaying cluster statistics_ are useful when we have larger problems and it is difficult to read the Parallel coordinate plots due to congestion. (@dasu2005grouping) -->
+
+<!-- (d) _MDS, PCA and t-SNE_ While all of them use a distance or dissimilarity matrix to construct a reduced-dimension space representation, their goals are diverse.
+PCA seeks to retain data variance.
+Multidimensional scaling (@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space.
+t-SNE, on the other hand, is concerned with preserving neighborhood data points.
+The t-SNE embedding will compress data points which are close in high-dimensional space.
+-->
+
+<!-- (e) _Tour_ is a collection of interpolated linear projections of multivariate data into lower-dimensional space.
+As a result, the viewer may observe the high-dimensional data's shadows from a low-dimensional perspective.
+-->
+
+<!-- The cluster characterization approach varies depending on the distance metric used.
+Parallel coordinate plots, scatter plot matrices, MDS or PCA are potentially useful ways to characterize clusters using wpd-based distances.
+For JS-based distances, plotting cluster statistics is beneficial for characterization and variable importance could be displayed through parallel coordinate plots.
+-->
+
+<!-- This part of the work uses R packages `GGally` (@R-GGally), `Rtsne` (@R-tsne), `ggplot2` (Wickham2009pk), `tour` (@wickham2011tourr), `stats` (@R-language). -->
+
+<!-- - A random sample of the original data is taken for clustering analysis and includes missing and noisy observations (detailed description in Appendix) -->
+
+<!-- - All harmonies are computed for each customer in the sample. -->
+<!-- Cyclic granularities which are clashes for all customers in the sample are removed. -->
+
+<!-- - It is worth noting that a number of other solutions may be considered at the pre-processing stage of the method.
+We have considered a) Normal-Quantile Transform and b) Robust transformation.
+-->
+
+<!-- - Two methods are considered for computing dissimilarity between two customers.
+The first one involves computing according to one granularity is computed as the sum of the JS distances between distribution of all the categories of the granularity.
+When we consider more than one granularity, we consider the sum of the average distances for all the granularity so that the combined metric is also a distance.
+-->
+
+<!-- - Given the scale of dissimilarity among the energy readings, the model chooses optimal number of clusters -->
+
+<!-- - Once clusters have been allocated, the groups are explored visually. -->
+
+<!-- - Results are reported and compared. -->
+
+<!-- Two methods are used for computing distances between subjects and then hierarchical clustering algorithm is used. -->
+
+<!-- The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the probability distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
+In our work, we are using $i$ as denoting a customer and the underlying variable as the electricity demand.
+So $f_i(v)$ is the distribution of household $i$ and $v$ is electricity demand.
+-->
+
+<!-- We want to cluster distributions of the form $f_{i,j,k}(v)$, where $i$ and $j$ denote -->
+
+<!-- Consider a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$.
+Each household consists of a $J*K$ distributions one harmony.
+We compute the distributional difference -->
+<!-- between $(A, B)$ for the $s^{th}$ household using $wpd_{{s}}(A,B)$. -->
+<!-- $wpd_{{s}}(A,B)$ denotes the normalized weighted-pairwise distributional distances between $(A, B)$ and is a feature which measures distributional difference between harmonies.
+If we have $H_{N_C}$ harmonies in the harmony table, then for each household we have a vector of $wpd_{{s}}$ of $H_{N_C}$ elements with each element corresponding to one harmony.
+We aim to have pockets of households showing similar periodic behavior by considering $wpd$ values for different harmonies and some time series features.
+The features should also characterize probability distributions of different household.
+-->
+
+<!-- ### Notations -->
+
+<!-- Consider an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the probability distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
+In our work, we are using $i$ as denoting a household and the underlying variable as the electricity demand.
+Further consider a cyclic granularity of the form $B = \{ b_k: k = 1, 2, \dots, K\}$.
+Each customer consists of collection of probability distributions.
+-->
+
+<!-- So $f_i(v)$ is the distribution of household $i$ and $v$ is electricity demand.
+We want to cluster distributions of the form $f_{i,j,k}(v)$, where $i$ and $j$ denote $i^{th}$ and $j^{th}$ customer respectively.
+-->
+
+<!-- a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$. -->
+
+<!-- ### A single or pair of granularities together (change names) -->
+
+<!-- The methodology can be summarized in the following steps: -->
+
+<!-- - _Pre-processing step_ -->
+
+<!-- Robust scaling method or NQT used for each customer. -->
+
+<!-- - _NQT_ -->
+
+<!-- - _Treatment to outliers_ -->
+
+<!-- ### Many granularities together (change names) -->
+
+<!-- The methodology can be summarized in the following steps: -->
+
+<!-- 1. Compute quantiles of distributions across each category of the cyclic granularity -->
+<!-- 2. Compute JS distance between households for each each category of the cyclic granularity -->
+<!-- 3. Total distance between households computed as sum of JS distances for all hours -->
+<!-- 4. Cluster using this distance with hierarchical clustering algorithm (method "Ward.D") -->
+
+<!-- _Pro:_  -->
+<!-- - distance metric makes sense to group different shapes together  -->
+<!-- - simulation results look great on typical designs  -->
+<!-- _Cons:_   -->
+<!-- - Can only take one granularity at once   -->
+<!-- - Clustering a big blob of points together whereas the aim is to groups these big blob into smaller ones   -->
+
+<!-- ### Multiple-granularities -->
+
+<!-- _Description:_   -->
+
+<!-- Choose all significant granularities and compute wpd for all these granularities for all customers.
+Distance between customers is taken as the euclidean distances between them with the granularities being the variables and wpd being the value under each variable for which Euclidean distance needs to be measured.
+-->
+<!-- _Pro:_   -->
+<!-- - Can only take many granularities at once -->
+<!-- - can apply variable selection PCP and other interesting clustering techniques -->
+<!-- - simulation results look great on typical designs -->
+<!-- - splitting the data into similar sized groups   -->
+<!-- _Cons:_   -->
+<!-- - distance metric does not make sense to split the data into similar shaped clusters  -->
+
+## Validation {#sec:validation}
+
+To validate our clustering methods, we spiked many attributes in the data to create different data designs.
+<!--to see where one method works better than the other and where they might give us the same outcome --or the effect of missing data <!---and trends on the proposed methods.--> Three circular granularities $g1$, $g2$ and $g3$ are considered with categories denoted by $\{g10,g11\}$, $\{g20, g21, g22\}$ and $\{g30, g31, g32, g33, g34\}$ and levels $n_{g_1}=2$, $n_{g_2}=3$ and $n_{g_3}=5$.
+These categories could be integers or some more meaningful labels.
+For example, the granularity "day-of-week" could be either represented by $\{0, 1, 2, \dots, 6\}$ or $\{Mon, Tue, \dots, Sun\}$.
+Here categories of $g1$, $g2$ and $g3$ are represented by $\{0, 1\}$, $\{0, 1, 2\}$ and $\{0, 1, 2, 3, 4\}$ respectively.
+A continuous measured variable $v$ of length $T$ indexed by $\{0, 1, \dots T-1\}$ is simulated such that it follows the structure across $g1$, $g2$ and $g3$.
+We constructed independent replications of all data designs $R = \{25, 250, 500\}$ to investigate if our proposed clustering method can discover distinct designs in small, medium, and big number of series.
+All designs employ $T=\{300, 1000, 5000\}$ sample sizes to evaluate small, medium, and large-sized series.
+Variations in method performance may be due to different jumps between categories.
+So a mean difference of $\mu = \{1, 2, 5\}$ between categories is considered.
+The performance of the approaches varies with the number of granularities which has interesting patterns across its categories.
+So three scenarios are considered to accommodate that.
+
+<!-- The results for $T=300$ and $R=25$ is shown, that means we have $25$ time series each with length $300$. The rest of the results could be found in the supplementary paper. -->
+
+### Data generating processes
+
+<!-- An ARMA (p,q) process is used to generate series, where $p$ and $q$ are selected at random such that the series is stationary.
+The various designs on $g1$, $g2$, and $g3$ are introduced by adding matching designs to this series' innovations.
+The innovations are considered to have a normal distribution, although they follow the same pattern as the designs.
+To eliminate the effect of starting values, the first 500 observations in each series are discarded.
+-->
+
+Each category or combination of categories from $g1$, $g2$ and $g3$ are assumed to come from the same distribution, a subset of them from the same distribution, a subset of them from separate distributions, or all from different distributions, resulting in various data designs.
+As the methods ignore the linear progression of time, there is little value in adding time dependency to the data generating process.
+The data type is set to be "continuous," and the setup is assumed to be Gaussian.
+When the distribution of a granularity is "fixed", it means distributions across categories do not vary and are considered to be from N (0,1).
+$\mu$ alters in the "varying" designs, leading to varying distributions across categories.
+
+<!-- An ARMA (p,q) process is used to generate series, where $p$ and $q$ are selected at random such that the series is stationary.
+The various designs on $g1$, $g2$, and $g3$ are introduced by adding matching designs to this series' innovations.
+The innovations are considered to have a normal distribution, although they follow the same pattern as the designs.
+To eliminate the effect of starting values, the first 500 observations in each series are discarded.
+-->
+
+<!-- It is often reasonable to construct a time series using properties such as trend, seasonality, and auto-correlation.
+However, when examining distributions across categories of cyclic granularities, these time series features are lost or addressed independently by considering seasonal fluctuations through cyclic granularities.
+Because the time span during which an entity is observed in order to ascertain its behavior is not very long, -->
+<!-- the behavior of the entity will not change drastically and hence the time series can be assumed to remain stationary throughout the observation period.
+If the observation period is very long (for e.g more than 3 years), property, physical or geographical attributes might change leading to a non-stationary time series.
+But such a scenario is not considered here and the resulting clusters are assumed to be time invariant in the observation period.
+-->
+
+### Data designs
+
+#### Individual granularities
+
+_Scenario (a): All significant granularities_
+
+<!-- Consider a case where all the three granularities $g1$, $g2$ and $g3$ would be responsible for making the designs distinct.
+That would mean, the pattern for each of $g1$, $g2$ and $g3$ will change for at least one design.
+We consider a situation with all the null cases corresponding to no difference in distribution across categories, that is, all categories follow the same distribution N(0,1).
+-->
+
+\noindent Consider the instance where $g1$, $g2$, and $g3$ all contribute to design distinction.
+This means that each granularity will have significantly different patterns at least across one of the designs to be clustered.
+In Table \ref{tab:tab-dist-design} various distributions across categories are considered (top) which lead to different designs (bottom).
+Figure \ref{fig:plot-3gran-new} shows the simulated variable's linear (left) and cyclic (right) representations for each of these five designs.
+The structural difference in the time series variable is impossible to discern from the linear view, with all of them looking very similar.
+The shift in structure may be seen clearly in the distribution of cyclic granularities.
+The following scenarios use solely graphical displays across cyclic granularities to highlight distributional differences in categories.
+
+_Scenario (b): Few significant granularities_
+
+\noindent This is the case where one granularity will remain the same across all designs.
+We consider the case where the distribution of $v$ varies across $g2$ levels for all designs, across $g3$ levels for a few designs, and $g1$ does not vary across designs.
+The proposed design is shown in Figure \ref{fig:gran2and1-clubbed}(b).
+
+_Scenario (c): One significant granularity_
+
+\noindent Only one granularity is responsible for identifying the designs in this case.
+This is depicted in Figure \ref{fig:gran2and1-clubbed} (right) where only $g3$ affects the designs significantly.
+
+```{r tab-distribution}
+```
+
+```{r tab-design}
+```
+
+```{r tab-dist-design}
+```
+
+```{r generate-D3change}
+
+```
+
+```{r generate-design-3change}
+
+```
+
+```{r plot-3gran-new, out.width="100%", fig.cap="The linear (left) and cyclic (right) representation of the simulated variable is shown.
+Each row represents a design in Scenario (a).
+In this scenario, all of $g1$, $g2$ and $g3$ changes across at least one design.
+Also, it is not possible to comprehend these differences in patterns just by looking at or considering the linear representation."}
+
+```
+
+```{r generate-design-new}
+
+```
+
+```{r generate-design-2gran-data}
+
+```
+
+```{r generate-design-2gran-plot}
+
+```
+
+```{r generate-design-1gran-data}
+
+```
+
+```{r generate-design-1gran-plot}
+
+```
+
+```{r gran2and1-clubbed, fig.cap=" Plots (b) and (c) correspond to Design (b) and (c) respectively.
+In (b) $g2$, $g3$ changes across atleast one design but $g1$ remains constant.
+Only $g3$ changes across different designs in (c)."}
+#gran2_change  + gran1_change + plot_layout(widths = c(1,1))
+#ggpubr::ggarrange(gran1_change, gran2_change, labels = c("a", "b"))
+```
+
+#### Interaction of granularities
+
+The proposed methods could be extended when two granularities of interest interact and we want to group subjects based on the interaction of the two granularities.
+Consider a group that has a different weekday and weekend behavior in the summer but not in the winter.
+This type of combined behavior across granularities can be discovered by evaluating the distribution across combinations of categories for different interacting granularities (Weekend/Weekday and month-of-year in this example).
+As a result, in this scenario, we analyze a combination of categories generated from different distributions.
+Consider a case in which there are only two interacting granularities of interest, $g1$ and $g2$.
+In contrast to the previous situation, when we could study distributions across $n_{g_1} + n_{g_2} = 5$ separate categories, with interaction, we must evaluate the distribution of the $n_{g_1}*n_{g_2}=6$ combination of categories.
+Consider the $4$ designs in Figure \ref{fig:interaction-gran}, where various distributions are assumed for different combinations of categories, resulting in different designs.
+Design $D1$ exhibits no change in distributions across $g1$ or $g2$, whereas Designs $D2$ and $D3$ alter across only $g1$ and $g2$, respectively.
+D4 varies across both $g1$ and $g2$ categories.
+D3 and D4 appear similar based on their relative differences across consecutive categories, but $D4$ also changes across facets, unlike $D3$, which has all facets look the same.
+
+```{r interaction-gran, fig.cap = "Distribution of the simulated variable across $g1$ conditional on $g2$ is shown through boxplots for 4 designs.
+D1 has no change in distributions across different categories of $g1$ or $g2$, while D2 and D3 change across only $g1$ and $g2$ respectively.
+D4 changes across categories of both $g1$ and $g2$."}
+
+```
+
+<!-- When two granularities of interest interact, the connection between a granularity and the measured variable is determined by the value of the other interacting granularity.
+This happens when the effects of the two granularities on the measured variable are not additive.
+For simplicity, consider a case with just two interacting granularities $g1$ and $g2$ of interest.
+As opposed to the last case, where we could play with the distribution of $5$ individual categories, with interaction we can play with the distribution of $6$ combination of categories.
+Consider $4$ designs in Figure \ref{fig:} where different distributions are assumed for different designs to get some distinction across designs.
+For example, in application, think about the scenario when customers need to grouped basis their joint behavior across hour-of-day and month-of-year.
+-->
+
+<!-- | Granularity type                                                   	| # Significant 	| # Replications 	| -->
+<!-- |--------------------------------------------------------------------	|---------------	|----------------	| -->
+<!-- | **Individual**  <br><br># obs: 300, 500, 2000  <br># clusters: 4/5 	| 1/2/3         	| 25, 100, 200   	| -->
+<!-- | **Interaction**  <br><br># obs: 500, 2000  <br># clusters: 4       	| 2           	| 25, 100, 200   	| -->
+
+### Visual exploration of findings
+
+All of the approaches were fitted to each data design and for each combination of the considered parameters.
+The formed clusters have to match the design, be well separated, and have minimal intra-cluster variation.
+It is possible to study these desired clustering traits visually in a more comprehensive way than just looking at index values.
+So we use MDS and parallel coordinate graphs to demonstrate the findings:
+
+- In Figure \ref{fig:mds-plot-validation}, we tried to see how separated our clusters are.
+We observe that in all scenarios and for different mean differences, clusters are separated.
+However, the separation increases with an increase in mean differences across scenarios.
+This is intuitive because, as the difference between categories increases, it gets easier for the methods to correctly distinguish the designs.
+
+- Figure \ref{fig:parcoord-sim} depicts a parallel coordinate plot with the vertical bar showing total inter-cluster distances with regard to granularities $g1$, $g2$, and $g3$ for all simulation settings and scenarios.
+So one line in the figure shows the inter-cluster distances for one simulation setting and scenarios vary across facets.
+The lines are not colored by group since the purpose is to highlight the contribution of the factors to categorization rather than class separation.
+The first plot shows that no variable stands out in the clustering, but the following two designs show that {g1} and {g1, g2} have very low inter cluster distances, meaning that they did not contribute to the clustering.
+It is worth noting that these facts correspond to our original assumptions when developing the scenarios, which incorporate distributional differences over three (a), two (b), and one (c) significant granularities.
+Hence, Figure \ref{fig:parcoord-sim} (a), (b), and (c) validate the construction of scenarios (a), (b), and (c) respectively.
+
+ - The js-robust and wpd methods perform worse for $nT=300$, then improve for higher $nT$ evaluated in the study.
+Although, a complete year of data is the minimum requirement to capture distributional differences in winter and summer profiles, for example.
+Even if the data is only available for a month, $nT$ with half-hourly data is expected to be at least $1000$.
+As a result, as long as the performance is promising for higher $nT=300$, this is not a challenge.
+
+- In our study sample, the method js-nqt outperforms the method js-robust for smaller differences between categories. More testing, however, is required to corroborate this.
+
+For more detailed results, please refer to the supplementary paper.
+The code for creating these data designs and running the methodologies is available at (https://github.com/Sayani07/paper-gracsR/Validation).
+
+```{r parcoord-sim, fig.cap = "The parallel coordinate plot illustrates the total inter-cluster distances for granularities $g1$, $g2$, and $g3$.
+One line in the figure depicts the inter-cluster distances for a single simulation scenario.
+While the first plot indicates that no variable stands out during clustering, the next two designs demonstrate that {g1} and {g1, g2} have extremely low inter-cluster distances, indicating that they did not contribute to clustering.
+It is worth emphasising that these facts are consistent with our initial assumptions when designing the scenarios and (a), (b), and (c) correspond to Scenario (a), (b) and (c) respectively."}
+
+```
+
+```{r mds-plot-validation, fig.cap = "Relative positions of clusters corresponding to different scenarios (columns) for different values of mean differences between categories (rows) are shown using the first two dimensions of MDS.
+It can be observed that clusters become more compact and separated for higher mean differences between categories across all designs.
+Between designs, separation is least prominent corresponding to scenario (c) where only granularity is responsible for the clusters."}
+
+```
+
+<!-- "Parallel coordinate plots are used to identify key variables for classification.
+Each line in this figure represents a formed group for a certain set of simulated parameters.
+(a), (b) and (c) corresponds to Scenarios (a), (b) and (c).
+The cyclic granularities $g1$, $g2$, and $g3$ are plotted on the x-axis, while total inter-cluster distances are plotted on the y-axis.
+Each line in this figure represents a group for a certain configuration of the various parameters ($nT$, $diff$, $R$).
+(c) demonstrates that intercluster distances are only significant for the variable $g3$, meaning that the clusters formed are primarily due to $g3$.
+This corresponds to our design in Scenario (c), in which distributional differences were implemented solely for g3.
+(b) demonstrates the role of g2 and g3 in clustering.
+(a) depicts a heterogeneous pattern of inter-cluster distances across variables, indicating that no single variable is to account for the clusters." -->
+
+<!-- A confusion table can come alive with linked brushing, so that mismatches and agreements between methods can be explored. -->
+
+```{r, eval=FALSE}
+# code in append_3gran_change.R
+table <- read_rds("data/append_3gran_change.rds")
+table %>% kable()
+```
+
+<!-- starts getting better with increasing difference and get worse with increasing number of replications.
+Length of series do not show to have any effect on the performance of the methods.
+It does not depend on if time series is ar or arma.
+-->
+
+<!-- - confusion matrix could be used for showing results if proper labeling is used -->
+
+<!-- - write about features that we have spiked into the data set -->
+<!-- - write about you incorporated noise -->
+<!-- - What is the additional structure you can incorporate that will lead to failing of method1 and method2? -->
+<!-- - And both gives the same result? Basically say when method 1 works better than method 2 and vice versa! -->
+
+<!-- - -->
+
+## Case study {#sec:cases}
+
+The use of our methodology is illustrated on smart meter energy usage for a sample of customers from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) which was available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) and Data61 CSIRO.
+It contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.
+In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual.
+We do not learn about these repetitive behaviors from the linear view because too many measurements all squeezed in that representation.
+Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior.
+The raw data for these consumers is of unequal length, with varying start and finish dates.
+Because our proposed methods evaluate probability distributions rather than raw data, neither of these data features would pose any threat to our methodology unless they contained any structure or systematic patterns.
+Additionally, there were missing values in the database but further investigation revealed that there is no structure in the missingness (see Supplementary paper for raw data features and missingness).
+The study begins by subsetting a data set along all dimensions of interest using data filtering and prototyping.
+By grouping the prototypes using our methods and assessing their meaning, the study hopes to unravel some of the heterogeneities observed in energy usage data.
+Because our application does not employ additional customer data, we cannot explain why consumption varies, but rather try to identify how it varies.
+
+_Data filtering and variable selection_
+
+- Choose a smaller subset of randomly selected $600$ customers with no implicit missing values for 2013.
+
+- Obtain $wpd$ for all cyclic granularities considered for these customers.
+It was found that `hod` (hour-of-day), `moy` (month-of-year) and `wkndwday` (weekend/weekday) are coming out to be significant for most customers.
+We use these three granularities while clustering.
+
+- Remove customers whose data for an entire category of `hod`, `moy` or `wnwd` is empty.
+For example, a customer who does not have data for an entire month is excluded because their monthly behavior cannot be analyzed.
+
+- Remove customers whose energy consumption is 0 in all deciles.
+These are the clients whose consumption is likely to remain essentially flat and with no intriguing repeated patterns that we are interested in studying.
+
+_Prototype selection_
+
+<!-- Why instance selection -->
+Supervised learning uses a training set of known information to categorize new events through instance selection.
+Instance selection (@olvera2010review) is a method of rejecting instances that are not helpful for classification.
+This is analogous to subsampling the population along all dimensions of interest such that the sampled data represents the primary features of the underlying distribution.
+Instance selection in unsupervised learning has received little attention in the literature, yet it could be a useful tool for evaluating model or method performance.
+There are several ways to approach prototype selection.
+Following @Fan2021-bq's idea of picking related examples (neighbors) for each instance (anchor), we can first use any dimensionality reduction techniques like MDS or PCA to project the data into a 2D space.
+Then pick a few "anchor" customers who are far apart in 2D space and pick a few neighbors for each.
+Unfortunately, this does not ensure that consumers with significant patterns across all variables are chosen.
+Tours can reveal variable separation that was hidden in a single variable display better than static projections.
+Hence we perform a linked tour with a t-SNE layout using the R package `liminal` (@R-liminal) to identify customers who are more likely to have distinct patterns across the variables studied.
+(Refer to Supplementary article for further details).
+Figure \ref{fig:hod-ind-group} shows the distribution across `hod` (a), `moy`(b) and `wnwd` (c) for the set of chosen $24$ customers that were chosen.
+Few of these customers have similar distribution across `moy` and some are similar in their `hod` distribution.
+
+```{r prototype-data-pick}
+
+```
+
+```{r assemble}
+
+```
+
+```{r data-pick}
+```
+
+<!-- - can be done in several ways -->
+
+<!-- - show raw data of moy, hod, wkndwday patterns for them -->
+
+```{r hod-moy-wkndwday plots, fig.cap="The distribution across moy, hod and wkndwday for the selected designs.
+Few are similar in their hod pattern, while others are similar in moy behavior.
+Some customers have distinct behavior as compared to all other customers.For example, although patterns across wkndwday do not look distinctly different for most households, there is one household for whom weekend behavior is standing out from the rest."}
+```
+
+### Clustering
+
+The $24$ prototypes <!--and the data subset consisting of $353$ customers--> are clustered using the methodology described in Section \ref{sec:methodology} and results are reported below.
+In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$.
+All customers with the same color represent the same clustered groups.
+Groups by JS-based distances and wpd-based distances are colored differently as they represent different groupings.
+The plotting scales are not displayed since we want to emphasize comparable shapes rather than scales.
+The idea is that a customer in a cluster may have low total energy usage, but their behavior may be quite similar to a customer with high usage with respect to distributional pattern or significance across cyclic granularities.
+
+#### JS-based distances
+
+For clustering based on JS-based distances, we chose the optimal number of clusters using (@Hennig2014-ah) as $5$.
+The groupings are shown in Figure \ref{fig:hod-ind-group}.
+Customers with the same color represent the same clustered groups.
+Our methodology is useful for grouping similar distributions over `hod` and `moy` and they are placed closely for easy comparison.
+Few groups have mixed patterns across `hod` and `moy`, but few have all customers in the group having a similar profile.
+Figure \ref{fig:combined-groups-js} shows the summarized distributions across $5$ groups and assists us in characterizing each cluster.
+It shows Groups $2$ and $4$ have `hod` pattern with a typical morning and evening peak, whereas groups $1$, $3$, and $5$ show a `moy` pattern with higher usage in winter months.
+Differences in Weekend/Weekday between groups are not discernible, implying that it may not be a relevant variable in distinguishing various clusters unless maybe conditioned by `moy` or `hod`.
+It may be interesting to compare these two plots to verify if the summarized distributions across groups correctly characterised the groupings.
+If it has, then the majority of the group's members should share a similar profile.
+
+<!--  It appears that the aim of grouping comparable distributions across categories of multiple temporal dependencies has been accomplished to some extent. -->
+
+```{r all-data}
+```
+
+```{r clustering}
+
+```
+
+```{r opt-clusters}
+```
+
+```{r groups-24}
+```
+
+```{r hod-ind-group, fig.cap = "The distribution of selected customers over hod (a), moy (b), and wnwd (c) for the $24$ selected customers is shown.
+In each case, the same colour denotes the same group using js-based clustering and are placed together to facilitate comparison.
+Individuals with similar distributional differences across categories of `hod` or `moy` are grouped together, with some exceptions.
+" }
+```
+
+<!-- It appears that the aim of grouping comparable distributions over considered variables has been accomplished to some extent. -->
+
+```{r data-heatmap-hod-group}
+```
+
+```{r data-heatmap-moy-group}
+
+```
+
+```{r data-heatmap-wkndwday-group}
+
+```
+
+```{r combined-groups-js, fig.cap = "The distribution of electricity demand over hod (a), moy (b), and wnwd*hod (c).
+Groups 2 and 5 appear to have a hod pattern among its members, whereas groups 1, 3, and 5 appear to have a moy pattern.
+Weekend/weekday differences across groups are not discernible, showing that it is not a critical variable for clustering on its own.
+It is probable that the distribution over Weekend/Weekday will differ conditional on the categories of hod or moy.
+(c) demonstrates that the distribution is not different, except for the tails.
+It is useful to compare the summarised distributions of groups to those of individuals in order to establish that the majority of individuals in the group share the same characteristics."}
+```
+
+#### wpd-based distances
+
+We chose the optimal number of clusters using (@Hennig2014-ah) as $3$.
+A parallel coordinate plot with the three significant cyclic granularities is used to characterise the groups here.
+The variables are sorted according to their separation across classes (rather than their overall variation between classes).
+This means that $moy$ is the most important variable in distinguishing the groups followed by $hod$ and $wnwd$.
+There is only one customer who has significant $wpd$ across $wnwd$ and stands out from the rest of the customers.
+Group 3 has a higher wpd for hod than moy or wkndwday.
+Group 2 has the most distinct pattern across moy.
+Group 1 is a mixed group that has strong patterns on at least one of the three variables.
+The findings vary from js-based clustering, yet it is a helpful grouping.
+
+```{r data-pick-wpd}
+```
+
+```{r parcoord, fig.cap="Each of the 24 customers is represented by a parallel coordinate plot (a) with three wpd-based groupings.
+The plot shows that moy is the most important variable in identifying clusters, whereas wkdn-wday is the least significant and has the least fluctuation.
+One particular customer with high wpd across wknwday stands out in this display.
+Group 3 has a higher wpd for hod than moy or wkndwday.
+Group 2 has most discernible pattern across moy.
+Group 1 is a mixed group with strong patterns on atleast one of the three variables.
+All of these could be observed from the plot or the table (b) which shows median wpd values for each group.", message=FALSE, warning=FALSE}
+
+```
+
+Things become far more complicated when we consider a larger data set with more uncertainty, as they do with any clustering problem.
+Summarizing distributions across clusters with varied or outlying customers can result in a shape that does not represent the group.
+Furthermore, combining heterogeneous customers may result in similar-looking final clusters that are not effective for visually differentiating them.
+It is also worth noting that the Weekend/Weekday behavior in the given case does not characterize any cluster.
+This, however, will not be true for all of the customers in the data set.
+If more extensive prototype selection is used, resulting in more comprehensive prototypes in the data set, this method might be used to classify the entire data set into these prototype behaviors.
+However, the goal of this section was to have a few customers that have significant patterns over one or more cyclic granularities, apply our methodology to cluster them, and demonstrate that the method produces useful clusters.
+
+## Conclusion and future work {#sec:conclusion}
+
+We offer two approaches for calculating pairwise distances between time series based on probability distributions over multiple cyclic granularities at once.
+Depending on the goal of the clustering, these distance metrics, when fed into a hierarchical clustering algorithm using Ward's linkage, yield meaningful clusters.
+Probability distributions provide an intuitive method to characterise noisy, patchy, long, and unequal-length time series data.
+Distributions over cyclic granularities help to characterise the formed clusters in terms of their repeating behavior over these cyclic granularities.
+Furthermore, unlike earlier efforts that group customers based on behavior across only one cyclic granularity (such as hour-of-day), our method is more comprehensive in detecting clusters with repeated patterns at all relevant granularities.
+<!--The JS-based approaches are faster than wpd-based approaches.-->
+
+There are few areas to extend this research.
+First, larger data sets with more uncertainty complicate matters, as is true for any clustering task.
+Characterizing clusters with varied or outlying customers can result in a shape that does not represent the group.
+Moreover, integrating heterogeneous consumers may result in visually identical end clusters, which are potentially not useful.
+Hence, a way of appropriately scaling it up to many customers such that anomalies are removed before clustering would be useful for bringing forth meaningful, compact and separated clusters.
+Secondly, we have assumed the time series to be stationary, and hence the distributions are assumed to remain constant for the observation period.
+In reality, however, it might change.
+For the smart meter example, the distribution for a customer moving to a different house or changing electrical equipment can change drastically.
+Our current approach can not detect these dynamic changes.
+Thirdly, it is possible that for a few customers, data for some categories from the list of considered significant granularities are missing.
+In our application, we have removed those customers and done the analysis but the metrics used should be able to incorporate those customers in the clustering by handling their missing categories.
+Finally, $wpd$ is computationally heavy even under parallel computation.
+Future work can make the computations more efficient so that they are easily scalable to a large number of customers.
+Moreover, experiments can also be run with non-hierarchy based clustering algorithms to verify if these distances work better with other algorithms.
+
+## Acknowledgments {-}
+
+The authors thank the ARC Centre of Excellence for Mathematical and Statistical Frontiers [(ACEMS)](https://acems.org.au/home) for supporting this research.
+Sayani Gupta was partially funded by [Data61 CSIRO](https://data61.csiro.au/) during her PhD.
+The Monash eResearch Centre and eSolutions-Study Support Services supported this research in part through the resource usage of the MonARCH HPC Cluster.
+The Github repository, [github.com/Sayani07/paper-gracsr](https://github.com/Sayani07/paper-gracsr), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
+This article was created with R [@R-language], `knitr` [@knitr, @R-knitr] and `rmarkdown` [@rmarkdown, @R-rmarkdown].
+Graphics are produced with `ggplot2` [@Wickham2009pk] and GGally [@R-GGally].
+
+## Supplementary Materials
+
+**Data and scripts:** Data sets and R code to reproduce all figures in this article (main.R).
+
+\noindent
+**Supplementary paper:** Additional tables, graphics and and R code to reproduce it (paper-supplementary.pdf, paper-supplementary.Rmd)
+
+\noindent
+**R-package:** To implement the ideas provided in this research, the open-source R package 'gracsr' is available on Github (https://github.com/Sayani07/gracsr).

### 2021-11-04 index.Rmd
running without references

@@
 ---
-title: 'Visualization and analysis of probability distributions of large temporal data'
-author: 'Sayani Gupta'
-degreetype: 'Doctor of Philosophy'
+title: "Visualization and analysis of probability distributions of large temporal data"
+author: "Sayani Gupta"
+degreetype: Doctor of Philosophy
 site: bookdown::bookdown_site
-degrees: 'M.Stat, B.Sc (Stat Hons)'
-bibliography: [bib/gravitas.bib, bib/tsibble.bib, bib/mists.bib, bib/thesisrefs.bib]
+degrees: M.Stat, B.Sc (Stat Hons)
+bibliography:
+- bib/gravitas.bib
+- bib/hakear.bib
+- bib/gracsr.bib
+- bib/thesisrefs.bib
 link-citations: yes

### 2021-11-04 thesis.Rmd
running without references

@@
 
-Placeholder
+<!-- Introduction -->
+<!-- Motivation & Background -->
+<!-- Outline of the thesis -->
+
+<!-- To tame the complexity of time, breaking it into years, months, weeks, days and so on in a hierarchical manner is a common way to relate data to time.
+Such dicrete human made abstractions of time can be thought of as time granularities.(Aigner et al., 2008).
+-->
+
+<!-- introducing the overall field-->
+With the availability of data at more and more finer time scales, exploration of time series data may be required to be carried out across both finer and coarser scales to draw useful inferences about the underlying process.
+For example, data collected at an hourly scale could be analyzed using coarser temporal scales such as days, months or quarters.
+This approach requires deconstructing time in various possible ways.
+Moreover, often it might be interesting to capture calendar or periodic effects like month-of-year, day-of-week or hour-of-day.
+They help us in answering questions like if certain levels of those time deconstructions are characterized by unusual/routine values of the observed variable.
+For example, certain days of the week or months of the year are likely to be characterized by higher values.
+It is important to be able to navigate through all the temporal deconstructions that accommodate for periodicities to have multiple perspectives of the observed data.
+This idea aligns with the notion of EDA (Tukey 1977) which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
+
+<!-- outline the intro chapter structure -->
+
+This chapter will provide an introduction to the study by first discussing the background and context, followed by the research aims, objectives and questions.
+
+<!-- Background -->
+
+The motivation for this work comes from the desire to provide methods to better understand large quantities of measurements on energy usage reported by smart meters in household across Australia, and indeed many parts of the world.
+Smart meters currently provide half-hourly use in kwh for each household, from the time that they were installed, some as early as 2012.
+Households are distributed geographically, and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
+The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
+It is common to see aggregates of usage across households, total kwh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
+But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
+
+<!-- introduce the specific problem  -->
+
+<!-- ## change in data structure as a result of transitioning from linear to cyclic deconstructions -->
+<!-- + Existing approaches, why we need to drill down, or why we want to -- probability distributions -->
+However, restructuring time in this manner leads to restructured data where each level of the time deconstructions correspond to multiple values of the observed variable.
+It is common to see aggregation or summarization of these multiple observations with a unique value to study calendar effects.
+For example, using aggregates of usage across each hour/half-hour has been common in the literature because energy companies need to plan for maximum loads on the network.
+But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
+Summarizing the probability distribution of these multiple observations to capture both the shape and uncertainty could be a potential way to understand the underlying distribution of these observations.
+Studying probability distributions is likely to focus on features of the data which are not transparent through raw data or a unique summary statistic.
+
+<!-- Following research objectives would facilitate the achievement of this aim -->
+<!-- Research objectives -->
+
+<!-- ## Research aims -->
+
+Hence, the overarching research goal is to study the periodic behavior of temporal data in a structured way by studying the probability distributions by best exploiting the characteristics of time.
+Slicing and dicing the data in all possible temporal scales as suggested by EDA can be a daunting task as it leads to a myriad of possibilities.
+This inspires the research presented in this thesis, which aims to provide a platform to systematically explore periodicities in temporal data and support finding regular patterns or anomalies, explore clusters of behaviors or summarize the behavior.
+The first part of the work discusses computation of all possible combinations of cyclic time granularities and a graphical mapping such that distributions of a numeric response variable is displayed across combinations of two cyclic granularities.
+Even analyzing the distribution of the measured variable across two cyclic granularities at once could amount to displaying many plots in search of potential  patterns.
+Thus, the first part of the research (@Gupta2020-vo) also introduces "harmony" to denote pairs of granularities that could be analyzed together and reduces the search from all possible options.
+But this approach is still overwhelming for human consumption because there would still be huge number of harmonies.
+Hence, the second part of the research extends this work and narrows the search further by finding pair of cyclic granularities which are informative enough and rank them according to their importance.
+However, to explore periodic patterns of many households, we have to resort to clustering which has been addressed in the third part of the research.
+Although the motivation came through the smart meter example, this is a problem that is relevant to any temporal data observed more than once per year.
 
 ## Visualizing probability distributions across bivariate cyclic temporal granularities {#sec:gravitas}
+
+Deconstructing a time index into time granularities can assist in exploration and automated analysis of large temporal data sets.
+This paper describes classes of time deconstructions using linear and cyclic time granularities.
+Linear granularities respect the linear progression of time such as hours, days, weeks and months.
+Cyclic granularities can be circular such as hour-of-the-day, quasi-circular such as day-of-the-month, and aperiodic such as public holidays.
+The hierarchical structure of granularities creates a nested ordering: hour-of-the-day and second-of-the-minute are single-order-up.
+Hour-of-the-week is multiple-order-up, because it passes over day-of-the-week.
+Methods are provided for creating all possible granularities for a time index.
+A recommendation algorithm provides an indication whether a pair of granularities can be meaningfully examined together (a "harmony"), or when they cannot (a "clash").
+
+ Time granularities can be used to create data visualizations to explore for periodicities, associations and anomalies.
+The granularities form categorical variables (ordered or unordered) which induce groupings of the observations.
+Assuming a numeric response variable, the resulting graphics are then displays of distributions compared across combinations of categorical variables.
+
+ The methods implemented in the open source R package `gravitas` are consistent with a tidy workflow, with probability distributions examined using the range of graphics available in `ggplot2`.
+
 ## Detecting distributional differences between temporal granularities for exploratory time series analysis {#sec:hakear}
+
+Cyclic temporal granularities, which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can be useful for measuring repetitive patterns in large univariate time series data.
+The granularities feed new approaches to exploring time series data.
+One use is to take pairs of granularities, and make plots of response values across the categories induced by the temporal deconstruction.
+However, when there are many granularities that can be constructed for a time period, there will also be too many possible displays to decide which might be the more interesting to display.
+This work proposes a new distance metric to screen and rank the possible granularities, and hence choose the most interesting ones to plot.
+The distance measure is computed for a single or pairs of cyclic granularities can can be compared across different cyclic granularities and also on a collection of time series.
+The methods are implemented in the open-source R package `hakear`.
+
 ## Clustering time series based on probability distributions across temporal granularities {#sec:gracsr}
+
+ With more and more time series data being collected at much finer temporal resolution, for a longer length of time, and for a larger number of individuals/entities, time series clustering research is getting a lot of traction.
+The sort of noisy, patchy, uneven, and asynchronous time series that is typical in many disciplines limits similarity searches among these lengthy time series.
+In this work, we suggest a method for overcoming these constraints by grouping time series based on probability distributions over cyclic temporal granularities.
+Cyclic granularities are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, and so on, and can be helpful for detecting repeating patterns.
+Looking at probability distributions across cyclic granularities results in an approach that is robust to missing or noisy data, aids in dimension reduction, and ensures small pockets of similar repeated behaviours.
+The proposed method was tested using a collection of residential electricity customers.
+The simulated and empirical evidence demonstrates that our method is capable of producing meaningful clusters.
+
+<!-- Time series clustering research is gaining a lot of importance with more and more time series data recorded at much finer temporal resolution, for a longer period of time and for a larger number of individuals/entities.Similarity searches amongst these long time series are often limited by the type of noisy, patchy, unequal and asynchronous time series common in many fields.
+In this paper we propose a strategy to alleviate these limitations by  clustering time series based on probability distributions across cyclic temporal granularities.
+Cyclic granularities are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, which can be useful for measuring repetitive patterns in large univariate time series data.
+Thus, looking at the probability distributions across cyclic granularity leads to a method which is robust to missing or noisy data, helps in dimension reduction while ensuring small pockets of similar repetitive behaviors.
+The method is applied to electricity smart meter dat -->
+<!-- The suggested approach was evaluated on a set of benchmark time series on residential electricity customers. The empirical results show that our approach is able to yield meaningful clusters. -->
+
 ## Summary
+
+The thesis is structured as follows.
+Chapter \@ref(ch:gravitas) provides details of the cyclic granularities, different classes, and computation, and also its usage in exploratory time series analysis through applications.
+This is implemented in the R package **gravitas**.
+Chapter \@ref(ch:hakear) provides guidance on how to choose significant cyclic granularities, which are likely to have interesting patterns across its categories.
+This is available as the R package **hakear**.
+The chapter @ref (ch: gracsr) provides similarity measures for comparing multiple time series.This is in the developing R package **gracsr**.
+Chapter \@ref(ch:conclusion) summarizes the software tools developed for the work, and discusses some future plans.
 
@@
 
-Placeholder
+Deconstructing a time index into time granularities can assist in exploration and automated analysis of large temporal data sets.
+This paper describes classes of time deconstructions using linear and cyclic time granularities.
+Linear granularities respect the linear progression of time such as hours, days, weeks and months.
+Cyclic granularities can be circular such as hour-of-the-day, quasi-circular such as day-of-the-month, and aperiodic such as public holidays.
+The hierarchical structure of granularities creates a nested ordering: hour-of-the-day and second-of-the-minute are single-order-up.
+Hour-of-the-week is multiple-order-up, because it passes over day-of-the-week.
+Methods are provided for creating all possible granularities for a time index.
+A recommendation algorithm provides an indication whether a pair of granularities can be meaningfully examined together (a "harmony"), or when they cannot (a "clash").
+
+ Time granularities can be used to create data visualizations to explore for periodicities, associations and anomalies.
+The granularities form categorical variables (ordered or unordered) which induce groupings of the observations.
+Assuming a numeric response variable, the resulting graphics are then displays of distributions compared across combinations of categorical variables.
+
+ The methods implemented in the open source R package `gravitas` are consistent with a tidy workflow, with probability distributions examined using the range of graphics available in `ggplot2`.
+
+```{r external, include = FALSE, cache = FALSE}
+read_chunk('scripts/gravitas.R')
+```
+
+```{r load}
+```
 
 ## Introduction
+
+Temporal data are available at various resolutions depending on the context.
+Social and economic data are often collected and reported at coarse temporal scales such as monthly, quarterly or annually.
+With recent advancements in technology, more and more data are recorded at much finer temporal scales.
+Energy consumption may be collected every half an hour, energy supply may be collected every minute, and web search data might be recorded every second.
+As the frequency of data increases, the number of questions about the periodicity of the observed variable also increases.
+For example, data collected at an hourly scale can be analyzed using coarser temporal scales such as days, months or quarters.
+This approach requires deconstructing time in various possible ways called time granularities [@aigner2011visualization].
+
+It is important to be able to navigate through all of these time granularities to have multiple perspectives on the periodicity of the observed data.
+This aligns with the notion of exploratory data analysis (EDA) [@Tukey1977-jx] which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
+Visualizing probability distributions conditional on one or more granularities is an indispensable tool for exploration.
+Analysts are expected to comprehensively explore the many ways to view and consider temporal data.
+However, the plethora of choices and the lack of a systematic approach to do so quickly can make the task overwhelming.
+
+Calendar-based graphics [@wang2020calendar] are useful in visualizing patterns in the weekly and monthly structure and are helpful when checking for the effects of weekends or special days.
+Any temporal data at sub-daily resolution can also be displayed using this type of faceting [@Wickham2009pk] with days of the week, month of the year, or another sub-daily deconstruction of time.
+But calendar effects are not restricted to conventional day-of-week or month-of-year deconstructions.
+There can be many different time deconstructions, based on the calendar or on categorizations of time granularities.
+
+Linear time granularities (such as hours, days, weeks and months) respect the linear progression of time and are non-repeating.
+One of the first attempts to characterize these granularities is due to @Bettini1998-ed.
+However, the definitions and rules defined are inadequate for describing non-linear granularities.
+Hence, there is a need to define some new time granularities, that can be useful in visualizations.
+Cyclic time granularities can be circular, quasi-circular or aperiodic.
+Examples of circular granularities are hour of the day and day of the week; an example of a quasi-circular granularity is day of the month; examples of aperiodic granularities are public holidays and school holidays.
+
+Time deconstructions can also be based on the hierarchical structure of time.
+For example, hours are nested within days, days within weeks, weeks within months, and so on.
+Hence, it is possible to construct single-order-up granularities such as second of the minute, or multiple-order-up granularities such as second of the hour.
+The lubridate package [@Grolemund2011-vm] provides tools to access and manipulate common date-time objects.
+However, most of its accessor functions are limited to single-order-up granularities.
+
+The motivation for this work stems from the desire to provide methods to better understand large quantities of measurements on energy usage reported by smart meters in households across Australia, and indeed many parts of the world.
+Smart meters currently provide half-hourly use in kWh for each household, from the time they were installed, some as early as 2012.
+Households are distributed geographically and have different demographic properties as well as physical properties such as the existence of solar panels, central heating or air conditioning.
+The behavioral patterns in households vary substantially; for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
+It is common to see aggregates [see @Goodwin_2012] of usage across households, such as half-hourly total usage by state, because energy companies need to plan for maximum loads on the network.
+But studying overall energy use hides the distribution of usage at finer scales, and makes it more difficult to find solutions to improve energy efficiency.
+We propose that the analysis of smart meter data will benefit from systematically exploring energy consumption by visualizing the probability distributions across different deconstructions of time to find regular patterns and anomalies.
+Although we were motivated by the smart meter example, the problem and the solutions we propose are practically relevant to any temporal data observed more than once per year.
+In a broader sense, it could be even suitable for data observed by years, decades, and centuries as might be in weather or astronomical data.
+
+This work provides tools for systematically exploring bivariate granularities within the tidy workflow (@wickham2016r). In particular, we
+
+  * provide a formal characterization of cyclic granularities;
+  * facilitate manipulation of single- and multiple-order-up time granularities through cyclic calendar algebra;
+  * develop an approach to check the feasibility of creating plots or drawing inferences for any two cyclic granularities.
+
+The remainder of the paper is organized as follows: Section&nbsp;\ref{sec:linear-time} provides some background material on linear granularities and calendar algebra for computing different linear granularities.
+Section&nbsp;\ref{sec:cyclic-gran} formally characterizes different cyclic time granularities by extending the framework of linear time granularities, and introducing cyclic calendar algebra for computing cyclic time granularities.
+The data structure for exploring the conditional distributions of the associated time series across pairs of cyclic time granularities is discussed in Section&nbsp;\ref{sec:data-structure}.
+Section&nbsp;\ref{sec:visualization} discusses the role of different factors in constructing an informative and trustworthy visualization.
+Section&nbsp;\ref{sec:application} examines how systematic exploration can be carried out for a temporal and non-temporal application.
+Finally, we summarize our results and discuss possible future directions in Section&nbsp;\ref{sec:discussion}.
+
 ## Linear time granularities {#sec:linear-time}
+
+Discrete abstractions of time such as weeks, months or holidays can be thought of as "time granularities".
+Time granularities are **linear** if they respect the linear progression of time.
+There have been several attempts to provide a framework for formally characterizing time granularities, including @Bettini1998-ed which forms the basis of the work described here.
+
 ### Definitions
+
+\begin{defn}\label{def:definition}
+A \textbf{time domain} is a pair $(T; \le)$ where $T$ is a non-empty set of time instants (equivalently, moments or points) and $\le$ is a total order on $T$.
+\end{defn}
+
+\noindent The time domain is assumed to be *discrete*, and there is unique predecessor and successor for every element in the time domain except for the first and last.
+
+\begin{defn}\label{def:index set}
+The \textbf{index set}, $Z=\{z: z \in \mathbb{Z}_{\geq 0}\}$, uniquely maps the time instants to the set of non-negative integers.
+\end{defn}
+
+\begin{defn}\label{def:linear}
+A \textbf{linear granularity} is a mapping $G$ from the index set, $Z$, to subsets of the time domain such that:
+  (1) if $i < j$ and $G(i)$ and $G(j)$ are non-empty, then each element of $G(i)$ is less than all elements of $G(j)$; and
+  (2) if $i < k < j$ and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.
+Each non-empty subset $G(i)$ is called a \textbf{granule}.
+\end{defn}
+
+\noindent This implies that the granules in a linear granularity are non-overlapping, continuous and ordered.
+The indexing for each granule can also be associated with a textual representation, called the label.
+A discrete time model often uses a fixed smallest linear granularity named by @Bettini1998-ed **bottom granularity**.
+\autoref{fig:linear-time} illustrates some common linear time granularities.
+Here, "hour" is the bottom granularity and "day", "week", "month" and "year" are linear granularities formed by mapping the index set to subsets of the hourly time domain.
+If we have "hour" running from $\{0, 1, \dots,t\}$, we will have "day" running from $\{0, 1, \dots, \lfloor t/24\rfloor\}$.
+These linear granularities are uni-directional and non-repeating.
+
+```{r linear-time, echo=FALSE, fig.cap="(ref:linear-time)", fig.pos="!htbp"}
+```
+
+(ref:linear-time) Illustration of time domain, linear granularities and index set.
+Hour, day, week, month and year are linear granularities and can also be considered to be time domains.
+These are ordered with ordering guided by integers and hence is unidirectional and non-repeating.
+Hours could also be considered the index set, and a bottom granularity.
+
 ### Relativities
+
+Properties of pairs of granularities fall into various categories.
+
+\begin{defn}\label{def:finerthan}
+A linear granularity $G$ is \textbf{finer than} a linear granularity $H$, denoted $G \preceq H$, if for each index $i$, there exists an index $j$ such that
+$G(i) \subset H(j).$
+\end{defn}
+
+\begin{defn}\label{def:groupsinto}
+A linear granularity $G$ \textbf{groups into} a linear granularity $H$, denoted
+$G \trianglelefteq H$, if for each index $j$ there exists a (possibly infinite) subset $S$ of the integers such that $H(j) = \bigcup_{i \in S}G(i).$
+\end{defn}
+
+\noindent For example, both $day \trianglelefteq week$ and $day \preceq week$ hold, since every granule of $week$ is the union of some set of granules of day and each day is a subset of a $week$.
+These definitions are not equivalent.
+Consider another example, where $G_1$ denotes "weekend" and $H_1$ denotes "week".
+Then, $G_1 \preceq H_1$, but $G_1 \ntrianglelefteq H_1$.
+Further, with $G_2$ denoting "days" and $H_2$ denoting "business-week", $G_2 \npreceq H_2$, but $G_2 \trianglelefteq H_2$, since each business-week can be expressed as an union of some days, but Saturdays and Sundays are not subset of any business-week.
+Moreover, with  $H_3$ denoting "public holidays", $G_1 \npreceq H_3$ and $G_1 \ntrianglelefteq H_3$.
+
+<!-- # talk about period after definition 6 after you have explained period -->
+<!-- The relationship has period 7. -->
+<!-- The relationship $day \trianglelefteq month$ has a more complicated period.
+If leap years are ignored, each month is a grouping of the same number of days over years, hence the period of the grouping $(day, month)$ is one year.
+With the inclusion of leap years, the grouping period is 400 years.
+-->
+
+\begin{defn}\label{def:periodic}
+A granularity $G$ is \textbf{periodic} with respect to a finite granularity $H$ if:
+(1) $G \trianglelefteq H$; and
+(2) there exist $R$, $P \in \mathbb{Z}_+$, where $R$ is less than the number of granules of $H$, such that for all $i \in \mathbb{Z}_{\ge {0}}$ , if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then $H (i + R) = \bigcup_{j \in S} G(j + P)$.
+\end{defn}
+
+If $G$ groups into $H$, it would imply that any granule $H(i)$ is the union of some granules of $G$, for example, $G(a_1), G(a_2), \dots, G(a_k)$.
+Condition (2) in Definition \ref{def:periodic} implies that if $H(i + R) \neq \emptyset$, then $H(i + R) = \bigcup (G(a_1 + P), G(a_2 + P), \dots, G(a_k + P))$, resulting in a "periodic" pattern of the composition of $H$ using granules of $G$.
+In this pattern, each granule of $H$ is shifted by $P$ granules of $G$.
+$P$ is called the \textbf{Period} (@Bettini2000-qk).
+
+For example, day is periodic with respect to week with $R=1$ and $P=7$, while (if we ignore leap years) day is periodic with respect to month with $R=12$ and $P=365$ as any month would consist of the same number of days across years.
+Since the idea of period involves a pair of granularities, we say that the pair $(day, week)$ has period 7, while the pair $(day, month)$ has a period 365 (ignoring leap years).
+
+Granularities can also be periodic with respect to other granularities, _"except for a finite number of spans of time where they behave in an anomalous way"_; these are called **quasi-periodic** relationships [@Bettini2000-vy].
+In a Gregorian calendar with leap years, day groups quasi-periodically into month with the exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.
+
+\begin{defn}\label{def:order}
+The \textbf{order} of a linear granularity is the level of coarseness associated with a linear granularity.
+A linear granularity G will have lower order than H if each granule of G is composed of lower number of granules of bottom granularity than each granule of H.
+\end{defn}
+
+With two linear granularities $G$ and $H$, if $G$ *groups into* or *finer than* $H$ then $G$ is of lower order than $H$.
+For example, if the bottom granularity is hour, then granularity $day$ will have lower order than $week$ since each day consist of fewer hours than each week.
+
+Granules in any granularity may be aggregated to form a coarser granularity.
+A system of multiple granularities in lattice structures is referred to as a **calendar** by @Dyreson_2000.
+Linear time granularities are computed through "calendar algebra" operations [@Ning_2002] designed to generate new granularities recursively from the bottom granularity.
+For example, due to the constant length of day and week, we can derive them from hour using
+$$
+  D(j) = \lfloor H(i)/24\rfloor, \qquad W(k) = \lfloor H(i)/(24*7)\rfloor,
+$$
+where $H$, $D$ and $W$ denote hours, days and weeks respectively.
+
 ## Cyclic time granularities {#sec:cyclic-gran}
+
+Cyclic granularities represent cyclical repetitions in time.
+They can be thought of as additional categorizations of time that are not linear.
+Cyclic granularities can be constructed from two linear granularities, that relate periodically; the resulting cycles can be either _regular_ (**circular**), or _irregular_ (**quasi-circular**).
+
 ### Circular granularities {#sec:circular-gran-def}
+
+\begin{defn}\label{def:circular}
+A \textbf{circular granularity} $C_{B, G}$ relates linear granularity $G$ to bottom granularity $B$ if
+\begin{equation} \label{eq:circular-gran}
+\begin{split}
+C_{B, G}(z) & = z\mod P(B, G) \quad \forall z \in \mathbb{Z}_{\geq 0} \\
+\end{split}
+\end{equation}
+where
+$z$ denotes the index set,
+$B$ groups periodically into $G$ with regular mapping and period $P(B, G)$.
+\end{defn}
+
+```{r circular-dow, echo=FALSE, fig.cap= "(ref:circular-dow)",fig.pos="!htb"}
+```
+
+(ref:circular-dow) Index sets for some linear and circular granularities (a). Circular granularities can be constructed by slicing the linear granularity into pieces and stacking them (b).
+
+\autoref{fig:circular-dow} illustrates some linear and cyclical granularities.
+Cyclical granularities are constructed by cutting the linear granularity into pieces, and stacking them to match the cycles (as shown in b).
+$B, G, H$ (day, week, fortnight, respectively) are linear granularities.
+The circular granularity $C_{B, G}$ (day-of-week) is constructed from $B$ and $G$, while  circular granularity $C_{B, H}$ (day-of-fortnight) is constructed from $B$ and $H$.
+These overlapping cyclical granularities share elements from the linear granularity.
+Each of $C_{B , G}$ and $C_{B , H}$ consist of repeated patterns $\{0, 1, \dots, 6\}$ and $\{0, 1, \dots, 13\}$ with $P=7$ and $P=14$ respectively.
+
+Suppose ${L}$ is a label mapping that defines a unique label for each index $\ell \in \{ 0,1,\dots, (P-1)\}$. For example, the label mapping $L$ for $C_{B, G}$ can be defined as
+$$
+  L: \{0,1, \dots, 6\} \longmapsto\ \{\text{Sunday}, \text{Monday}, \dots, \text{Saturday}\}.
+$$
+
+In general, any circular granularity relating two linear granularities can be expressed as
+$$
+  C_{G, H}(z) = \lfloor z/P(B,G) \rfloor\mod P(G,H),
+$$
+where $H$ is periodic with respect to $G$ with regular mapping and period $P(G,H)$.
+Table&nbsp;\ref{tab:definitions} shows several circular granularities constructed using minutes as the bottom granularity.
+
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{lll}
+\toprule
+Circular granularity & Expression & Period \\
+\midrule
+minute-of-hour                               &
+  $C_1 = z \mod 60$                     &
+  $P_1 = \phantom{99}60$ \\
+minute-of-day                                &
+  $C_2 = z \mod 60*24$                  &
+  $P_2= `r 60*24`$\\
+hour-of-day                                  &
+  $C_3 = \lfloor z/60\rfloor\mod 24$    &
+  $P_3 = \phantom{99}24$ \\
+hour-of-week                                 &
+  $C_4 = \lfloor z/60\rfloor\mod 24*7$  &
+  $P_4= \phantom{9}`r 24*7`$\\
+day-of-week                                  &
+  $C_5 = \lfloor z/24*60\rfloor \mod 7$ &
+  $P_5= \phantom{999}7$\\
+\bottomrule
+\end{tabular}
+\end{center}
+\caption{Examples of circular granularities with bottom granularity minutes.
+Circular granularity $C_i$ relates two linear granularities one of which groups periodically into the other with regular mapping and period $P_i$.
+Circular granularities can be expressed using modular arithmetic due to their regular mapping.}
+\label{tab:definitions}
+\end{table}
+
 ### Quasi-circular granularities {#sec:quasi-circular-gran-def}
+
+A **quasi-circular** granularity cannot be defined using modular arithmetic because of the irregular mapping.
+However, they are still formed with linear granularities, one of which groups periodically into the other.
+\autoref{tab:quasi} shows some examples of quasi-circular granularities.
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{lr@{~}lr@{~}r}
+\toprule
+Quasi-circular granularity && Possible period lengths\\
+\midrule
+$Q_1 =$ day-of-month && $P_1 = 31, 30, 29, 28$\\
+$Q_2 =$ hour-of-month && $P_2 = 24\times 31, 24\times 30, 24\times 29, 24\times 28$\\
+$Q_3 =$ day-of-year && $P_3 = 366, 365$\\
+\bottomrule
+\end{tabular}
+\caption{Examples of quasi-circular granularities relating two linear granularities with irregular mapping leading to several possible period lengths.}
+\label{tab:quasi}
+\end{table}
+
+\begin{defn}\label{def:quasicircular}
+A \textbf{quasi-circular granularity} $Q_{B, G'}$ is formed when bottom granularity $B$ groups periodically into linear granularity $G'$ with irregular mapping such that the granularities are given by
+\begin{equation}\label{eq:quasi}
+Q_{B, G'}(z) =
+z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert, \quad \text{for}\quad z \in T_{k},
+\end{equation}
+where
+$z$ denotes the index set,
+$w$ denotes the index of $G'$,
+$R'$ is the number of granules of $G'$ in each period of $(B, G')$,
+$T_w$ are the sets of indices of $B$ such that $G'(w) = \bigcup_{z \in T_w}B(z)$,
+and $\vert T_w \vert$ is the cardinality of set $T_w$.
+\end{defn}
+
+For example, day-of-year is quasi-periodic with either 365 or 366 granules of $B$ (days) within each granule of $G'$ (years).
+The pattern repeats every $4$ years (ignoring leap seconds).
+Hence $R'= 4$.
+$Q_{B, G'}$ is a repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not the same across different granules of $G'$.
+
 ### Aperiodic granularities {#sec:aperiodic-gran-def}
+
+Aperiodic linear granularities are those that cannot be specified as a periodic repetition of a pattern of granules as described in Definition \ref{def:periodic}.
+Aperiodic cyclic granularities capture repetitions of these aperiodic linear granularities.
+Examples include public holidays which repeat every year, but there is no reasonably small span of time within which their behavior remains constant.
+A classic example is Easter (in the western tradition) whose dates repeat only after 5.7 million years [@Reingold2001-kf].
+In Australia, if a standard public holiday falls on a weekend, a substitute public holiday will sometimes be observed on the first non-weekend day (usually Monday) after the weekend.
+Examples of aperiodic granularity may also include school holidays or a scheduled event.
+All of these are recurring events, but with non-periodic patterns.
+Consequently, $P_i$ (as given in \autoref{tab:quasi}) are essentially infinite for aperiodic granularities.
+
+\begin{defn}\label{def:aperiodic}
+An \textbf{aperiodic cyclic granularity} is formed when bottom granularity $B$ groups into an aperiodic linear granularity $M$ such that the granularities are given by
+\begin{equation}\label{eq:aperiodic}
+A_{B, M}(z) = \begin{cases}
+                  i, & \text{for}\quad z \in T_{i_j} \\
+                  0  & \text{otherwise},
+                \end{cases}
+\end{equation}
+where
+$z$ denotes the index set,
+$T_{i_j}$ are the sets of indices of $B$ describing aperiodic linear granularities $M_{i}$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$, and $M = \bigcup_{i=1}^{n}M_{i}$, $n$ being the number of  aperiodic linear granularities in consideration.
+\end{defn}
+
+```{r aperiodic-example, echo=FALSE, fig.cap= "(ref:aperiodic-example)", fig.pos="!htb"}
+```
+
+(ref:aperiodic-example) Quasi-circular and aperiodic cyclic granularities illustrated by linear (a) and 
+stacked-displays (b) progression of time.
+The linear display shows the granularities days ($B$), weeks ($G$), semester days ($B'$), and stages of a semester ($M$) indexed over a linear representation of time.
+The granules of $B'$ is only defined for days when the semester is running.
+Here a semester spans 18 weeks and 2 days, and consists of 6 stages.
+It starts with a week of orientation, followed by an in-session period (6 weeks), a break (1 week), the second half of semester (7 weeks), a 1-week study break before final exams, which spans the next 16 days.
+This distribution of semester days remains relatively similar for every semester.
+$Q_{B',M}$ with P = 128 is a quasi-circular granularity with repeating patterns, while $A_{B,M}$ is an aperiodic cyclic granularity as the placement of the semester within a year varies from year to year with no fixed start and end dates.
+
+For example, consider the school semester shown in \autoref{fig:aperiodic-example}.
+Let the linear granularities $M_1$ and $M_2$ denote the teaching and non-teaching stages of the semester respectively.
+Both $M_1$, $M_2$ and $M = M_{1}\bigcup M_{2}$ denoting the "stages" of the semester are aperiodic with respect to days ($B$) or weeks ($G$).
+Hence $A_{B, M}$ denoting day-of-the-stage would be an aperiodic cyclic granularity because the placement of the semester within a year would vary across years.
+Here, $Q_{B', M}$ denoting semester-day-of-the-stage would be a quasi-circular granularity
+since the distribution of semester days within a semester is assumed to remain constant over years.
+Here semester-day is denoted by "sem day" ($B'$) and its granules are only defined for the span of the semesters.
+
 ### Relativities {#sec:cyclic-calendar}
+
+The hierarchical structure of time creates a natural nested ordering which can be used in the computation of relative pairs of granularities.
+
+\begin{defn}\label{def:hierarchy}
+The nested ordering of linear granularities can be organized into a \textbf{hierarchy table}, denoted as $H_n: (G, C, K)$, which arranges them from lowest to highest in order.
+It shows how the $n$ granularities relate through $K$, and how the cyclic granularities, $C$, can be defined relative to the linear granularities.
+Let $G_{\ell}$ and $G_{m}$ represent the linear granularity of order $\ell$ and $m$ respectively with $\ell<m$.
+Then $K \equiv P(\ell,m)$ represents the period length of the grouping $(G_{\ell}, G_{m})$, if $C_{G_{\ell}, G_{m}}$ is a circular granularity and $K \equiv k(\ell,m)$ represents the operation to obtain $G_{m}$ from $G_{\ell}$, if $C_{G_{\ell}, G_{m}}$ is quasi-circular.
+\end{defn}
+
+For example, \autoref{tab:tab-mayan} shows the hierarchy table for the Mayan calendar.
+In the Mayan calendar, one day was referred to as a kin and the calendar was structured such that 1 kin = 1 day; 1 uinal = 20 kin; 1 tun  = 18 uinal (about a year); 1 katun = 20 tun (20 years) and 1 baktun = 20 katun.
+
+```{r tab-mayan}
+```
+
+Like most calendars, the Mayan calendar used the day as the basic unit of time [@Reingold2001-kf].
+The structuring of larger units, weeks, months, years and cycle of years, though, varies substantially between calendars.
+For example,  the French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds", the duration of which is 0.864 common seconds.
+Nevertheless, for any calendar, a hierarchy table can be defined.
+Note that it is not always possible to organize an aperiodic linear granularity in a hierarchy table.
+Hence, we assume that the hierarchy table consists of periodic linear granularities only, and that the cyclic granularity $C_{G(\ell),G(m)}$ is either circular or quasi-circular.
+
+\begin{defn}\label{def:norderup}
+The hierarchy table contains \textbf{multiple-order-up} granularities which are cyclic granularities that are nested within multiple levels.
+A \textbf{single-order-up} is a cyclic granularity which is nested within a single level. It is a special case of multiple-order-up granularity.
+\end{defn}
+
+\noindent In the Mayan calendar (Table \@ref(tab:tab-mayan)), kin-of-tun or kin-of-baktun are examples of multiple-order-up granularities and single-order-up granularities are kin-of-uinal, uinal-of-tun etc.
+
 ### Computation
+
+Following the calendar algebra of @Ning_2002 for linear granularities, we can define cyclic calendar algebra to compute cyclic granularities. Cyclic calendar algebra comprises two kinds of operations:
+(1) **single-to-multiple** (the calculation of _multiple-order-up_ cyclic granularities from _single-order-up_ cyclic granularities) and (2) **multiple-to-single** (the reverse).
+
 #### Single-to-multiple order-up {-}
+
+Methods to obtain multiple-order-up granularity will depend on whether the hierarchy consists of all circular single-order-up granularities or a mix of circular and quasi-circular single-order-up granularities.
+Circular single-order-up granularities can be used recursively to obtain a multiple-order-up circular granularity using
+\begin{equation} \label{eq:eq7}
+C_{G_\ell,G_m}(z)
+  = \sum_{i=0}^{m - \ell - 1} P(\ell, \ell+i)C_{G_{\ell+i},G_{\ell+i+1}}(z),
+\end{equation}
+where $\ell < m - 1$ and $P(i, i) = 1$ for $i=0,1,\dots,m-\ell-1$, and
+$C_{B, G}(z)  = z\mod P(B, G)$ as per Equation \eqref{eq:circular-gran}.
+
+For example, the multiple-order-up granularity $C_{\text{uinal},\text{katun}}$ for the Mayan calendar could be obtained using
+\begin{align*}
+C_{\text{uinal}, \text{baktun}}(z)  &=  C_{\text{uinal}, \text{tun}}(z) + P(\text{uinal}, \text{tun})C_{\text{tun},\text{katun}}(z) + P(\text{uinal},\text{katun})C_{\text{katun}, \text{baktun}}(z) \\
+         &=  C_{\text{uinal}, \text{tun}}(z) + 18 \times{ C_{\text{tun},\text{katun}}(z)} + 18 \times 20 \times{C_{\text{katun}, \text{baktun}}(z)}
+\end{align*}
+
+, where $z$ is the index of the bottom granularity $kin$.
+
+<!--                &=  \lfloor \frac{z}{20}\rfloor \mod 18 + 18\lfloor  \frac{z}{20 \times18} \rfloor \mod 20 -->
+<!--                 + 18 \times20 \lfloor \frac{z}{(20\times18\times20} \rfloor \mod 20 -->
+
+Now consider the case where there is one quasi-circular single order-up granularity in the hierarchy table while computing a multiple-order-up quasi-circular granularity.
+Any multiple-order-up quasi-circular granularity $C_{\ell, m}(z)$ could then be obtained as a discrete combination of circular and quasi-circular granularities.
+
+Depending on the order of the combination, two different approaches need to be employed leading to the following cases:
+
+  * $C_{G_{\ell},G_{m'}}$ is circular and $C_{G_{m'},G_{m}}$ is quasi-circular
+\begin{equation} \label{eq:multifromsingle-quasi1}
+C_{G_\ell,G_{m}}(z) = C_{G_{\ell},G_{m'}}(z) + P(\ell, m')C_{G_{m'},G_{m}}(z)
+\end{equation}
+
+  * $C_{G_{\ell},G_{m'}}$ is quasi-circular and $C_{G_{m'},G_{m}}$ is circular
+\begin{equation} \label{eq:multifromsingle-quasi2}
+C_{G_\ell,G_{m}}(z)  = C_{G_{\ell},G_{m'}}(z) + \sum_{w=0}^{C_{G_{m'},G_{m}}(z) -1}(\vert T_{w} \vert)
+\end{equation}
+where, $T_w$ is such that $G_{m'}(w) = \bigcup_{z \in T_w}G_{\ell}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
+
+```{r tab-gregorian}
+```
+
+For example, the Gregorian calendar (\autoref{tab:tab-gregorian}) has day-of-month as a single-order-up quasi-circular granularity, with the other granularities being circular.
+Using Equations \eqref{eq:multifromsingle-quasi1} and \eqref{eq:multifromsingle-quasi2}, we then have:
+$$
+  C_{hour, month}(z) = C_{hour, day}(z) + P(hour, day)*C_{day, month}(z)
+$$
+$$
+  C_{day, year}(z) = C_{day,month}(z) + \sum_{w=0}^{C_{month, year}(z)-1}(\vert T_{w} \vert),
+$$
+where $T_w$ is such that $month(w) = \bigcup_{z \in T_w}day(z)$.
+
 #### Multiple-to-single order-up {-}
+
+Similar to single-to-multiple operations, multiple-to-single operations involve different approaches for all circular single-order-up granularities and a mix of circular and quasi-circular single-order-up granularities in the hierarchy.
+For a hierarchy table $H_n: (G, C, K)$ with only circular single-order-up granularities and $\ell_1, \ell_2, m_1, m_2 \in {1, 2, \dots, n}$ and $\ell_2<\ell_1$ and $m_2>m_1$, multiple-order-up granularities can be obtained using Equation \eqref{eq:all-circular-multiple}.
+\begin{equation} \label{eq:all-circular-multiple}
+C_{G_{\ell_1}, G_{m_1}}(z) = \lfloor C_{G_{\ell_2}, G_{m_2}}(z)/P(\ell_2,\ell_1) \rfloor \mod P(\ell_1, m_1)
+\end{equation}
+For example, in the Mayan Calendar, it is possible to compute the single-order-up granularity tun-of-katun from uinal-of-baktun, since $C_{tun, katun}(z) = \lfloor C_{uinal, baktun}(z)/18\rfloor \mod 20$.
+
 #### Multiple order-up quasi-circular granularities {-}
+
+Single-order-up quasi-circular granularity can be obtained from multiple-order-up quasi-circular granularity and single/multiple-order-up circular granularity using Equations \eqref{eq:multifromsingle-quasi1} and \eqref{eq:multifromsingle-quasi2}.
+
 ## Data structure {#sec:data-structure}
+
+Effective exploration and visualization benefit from well-organized data structures.
+@wang2020tsibble introduced the tidy "tsibble" data structure to support exploration and modeling of temporal data.
+This forms the basis of the structure for cyclic granularities.
+A tsibble comprises an index, optional key(s), and measured variables.
+An index is a variable with inherent ordering from past to present and a key is a set of variables that define observational units over time.
+A linear granularity is a mapping of the index set to subsets of the time domain.
+For example, if the index of a tsibble is days, then a linear granularity might be weeks, months or years.
+A bottom granularity is represented by the index of the tsibble.
+
+\novspacing
+
+```{r data-structure, fig.pos="!htb"}
+library(dplyr)
+as.data.frame(matrix(" ", ncol=7, nrow=1)) %>%
+  mutate(
+    V1 = cell_spec(V1, "latex", background="#dbe3f1"),
+    V2 = cell_spec(V2, "latex", background="#dbe3f1"),
+    V3 = cell_spec(V3, "latex", background="#dbe3f1"),
+    V4 = cell_spec(V4, "latex", background="#fdf2d0"),
+    V5 = cell_spec(V5, "latex", background="#fdf2d0"),
+    V6 = cell_spec(V6, "latex", background="#fdf2d0"),
+    V7 = cell_spec(V7, "latex", background="#fdf2d0"),
+  ) %>%
+  knitr::kable(format='latex',
+               booktabs=TRUE,
+               escape=FALSE,
+               caption = "The data structure for exploring periodicities in data by including cyclic granularities in the tsibble structure with index, key and measured variables.",
+               col.names = c("\\cellcolor[HTML]{dbe3f1}{index}","\\cellcolor[HTML]{dbe3f1}{key}","\\cellcolor[HTML]{dbe3f1}{measurements}","$C_1$","$C_2$","$\\cdots$","$C_{N_C}$")
+  ) %>%
+  row_spec(0, bold=TRUE,background="#fdf2d0") %>%
+  kable_styling(position = "center")
+```
+
+\vspacing
+
+All cyclic granularities can be expressed in terms of the index set.
+\autoref{tab:data-structure} shows the tsibble structure (index, key, measurements) augmented by columns of cyclic granularities.
+The total number of cyclic granularities depends on the number of linear granularities considered in the hierarchy table and the presence of any aperiodic cyclic granularities.
+For example, if we have $n$ periodic linear granularities in the hierarchy table, then $n(n-1)/2$ circular or quasi-circular cyclic granularities can be constructed.
+Let $N_C$ be the total number of contextual circular, quasi-circular and aperiodic cyclic granularities that can originate from the underlying periodic and aperiodic linear granularities.
+Simultaneously encoding more than a few of these cyclic granularities when visualizing the data overwhelms human comprehension.
+Instead, we focus on visualizing the data split by pairs of cyclic granularities ($C_i$, $C_j$).
+Data sets of the form <$C_i$, $C_j$, $v$> then allow exploration and analysis of the measured variable $v$.
+
 ### Harmonies and clashes {#sec:synergy}
+
+The way granularities are related is important when we consider data visualizations.
+Consider two cyclic granularities $C_i$ and $C_j$, such that $C_i$ maps index set to a set $\{A_k \mid k=1,\dots,K\}$ and $C_j$ maps index set to a set $\{B_\ell \mid \ell =1,\dots,L\}$.
+Here, $A_k$ and $B_\ell$ are the levels/categories corresponding to $C_i$ and $C_j$ respectively.
+Let $S_{k\ell}$ be a subset of the index set such that for all $s \in S_{k\ell}$, $C_i(s) = A_k$ and $C_j(s) = B_\ell$.
+There are $KL$ such data subsets, one for each combination of levels ($A_k$, $B_\ell$).
+Some of these sets may be empty due to the structure of the calendar, or because of the duration and location of events in a calendar.
+
+\begin{defn}\label{def:clash}
+A \textbf{clash} is a pair of cyclic granularities that contains empty combinations of categories.
+\end{defn}
+
+\begin{defn}\label{def:harmony}
+A \textbf{harmony} is a pair of cyclic granularities that does not contain any empty combinations of its categories.
+\end{defn}
+
+Structurally empty combinations can arise due to the structure of the calendar or hierarchy.
+For example, let $C_i$ be day-of-month with 31 levels and $C_j$ be day-of-year with 365 levels.
+There will be $31\times 365=11315$ sets $S_{k\ell}$ corresponding to possible combinations of $C_i$ and $C_j$.
+Many of these are empty.
+For example, $S_{1,5}$ is empty because the first day of the month can never correspond to the fifth day of the year.
+Hence the pair (day-of-month, day-of-year) is a clash.
+
+Event-driven empty combinations arise due to differences in event location or duration in a calendar.
+For example, let $C_i$ be day-of-week with 7 levels and $C_j$ be working-day/non-working-day with 2 levels.
+While potentially all of these 14 sets $S_{k\ell}$ can be non-empty (it is possible to have a public holiday on any day-of-week), in practice many of these will probably have very few observations.
+For example, there are few (if any) public holidays on Wednesdays or Thursdays in any given year in Melbourne, Australia.
+
+An example of harmony is where $C_i$ and $C_j$ denote day-of-week and month-of-year respectively.
+So $C_i$ will have 7 levels while $C_j$ will have 12 levels, giving $12\times7=84$ sets $S_{k\ell}$.
+All of these are non-empty because every day-of-week can occur in every month.
+Hence, the pair (day-of-week, month-of-year) is a harmony.
+
 ### Near-clashes {#sec:near-clashes}
+
+Suppose $C_i$ denotes day-of-year and $C_j$ denotes day-of-week.
+While any day of the week can occur on any day of the year, some combinations will be very rare.
+For example, the 366th day of the year will only coincide with a Wednesday approximately every 28 years on average.
+We refer to these as "near-clashes".
+
 ## Visualization {#sec:visualization}
+
+The purpose is to visualize the distribution of the continuous variable ($v$) conditional on the values of two granularities, $C_i$ and $C_j$.
+Since $C_i$ and $C_j$ are factors or categorical variables, data subsets corresponding to each combination of their levels form a subgroup and the visualization amounts to having displays of distributions for different subgroups.
+The response variable $(v)$ is plotted on the y-axis and the levels of $C_i (C_j)$ on the x-axis, conditional on the levels of $C_j (C_i)$.
+This means, carrying out the same plot corresponding to each level of the conditioning variable.
+This is consistent with the widely used grammar of graphics
+which is a framework to construct statistical graphics by relating the data space to the graphic space [@Wilkinson1999-nk, @Wickham2009pk].
+
+<!-- When a data is made up of different groups, it is often informative to compare the distribution of the variables across the groups. -->
+
 ### Data summarization
+
+There are several ways to summarize the distribution of a data set such as estimating the empirical distribution or density of the data, or computing a few quantiles or other statistics.
+This estimation or summarization could be potentially misleading if it is performed on rarely occurring categories (Section&nbsp;\ref{sec:near-clashes}).
+Even when there are no rarely occurring events, the number of observations may vary greatly within or across each facet, due to missing observations or uneven locations of events in the time domain.
+In such cases, data summarization should be used with caution as sample sizes will directly affect the accuracy of the estimated quantities being displayed.
+
+<!-- Each plot option has parameters that need to be estimated or summarized from the data. -->
+<!-- For example, for histogram, empirical distribution is estimated, for boxplot or its variation, number summaries are computed, for violin or any density-based plots, the density of the variable's distribution needs to be estimated.
+-->
+
 ### Display choices for univariate distributions
+
+<!-- Plotting each point for each combination through dot plots or rug plots is not useful for our data structure since that leads to a blob of points for each subgroup.
+Further, they are anyway a poor choice for large datasets.
+Hence, -->
+
+The basic plot choice for our data structure is one that can display distributions.
+For displaying the distribution of a continuous univariate variable, many options are available.
+Displays based on descriptive statistics include box plots [@Tukey1977-jx] and its variants such as notched box plots [@McGill1978-hg] or other variations as mentioned in @boxplots.
+They also include line or area quantile plots which can display any quantiles and not only quartiles like in a boxplot.
+Plots based on kernel density estimates include violin plots [@Hintze1998-zi], summary plot [@Potter2010-qc], ridge line plots [@R-ggridges], and highest density region (HDR) plots [@Hyndman1996-ft].
+The less commonly used Letter-Value plots [@Hofmann2017-sg] is midway between boxplots and density plots.
+Letter values are order statistics with specific depths, for example, the median ($M$) is a letter value that divides the data set into halves.
+Each of the next letter values splits the remaining parts into two separate regions so that the fourths ($F$), eighths ($E$), sixteenths ($D$), etc.
+are obtained.
+They are useful for displaying the distributions beyond the quartiles especially for large data,
+where boxplots mislabel data points as outliers.
+One of the best approaches in exploratory data analysis is to draw a variety of plots to reveal information while keeping in mind the drawbacks and benefits of each of the plot choices.
+For example, boxplots obscure multimodality, and interpretation of density estimates and histograms may change depending on the bandwidth and binwidths respectively.
+In R package `gravitas` [@R-gravitas], boxplots, violin, ridge, letter-value, line and area quantile plots are implemented, but it is potentially possible to use any plots which can display the distribution of the data.
+
 ### Comparison across sub-groups induced by conditioning {#sec:aesthetics}
+
+<!-- While it is often beneficial to compare distributions in this way, there are some caveats that we should be kept in mind while exploring these displays, subject to our data structure as follows: -->
+
 #### Levels {-}
+
+The levels of cyclic granularities affect plotting choices since space and resolution may be problematic with too many levels.
+A potential approach could be to categorize the number of levels as low/medium/high/very high for each cyclic granularity and define some criteria based on human cognitive power, available display size and the aesthetic mappings.
+Default values for these categorizations could be chosen based on levels of common temporal granularities like days of the month, days of the fortnight, or days of the week.
+
 #### Synergy of cyclic granularities {-}
+
+The synergy of the two cyclic granularities will affect plotting choices for exploratory analysis.
+Cyclic granularities that form clashes (Section \@ref(sec:synergy)) or near-clashes lead to potentially ineffective graphs.
+Harmonies tend to be more useful for exploring patterns.
+\autoref{fig:allFig}a shows the distribution of half-hourly electricity consumption through letter value plots across months of the year conditional on quarters of the year.
+This plot does not work because quarter-of-year clashes with month-of-year, leading to empty subsets.
+For example, the first quarter never corresponds to December.
+
+```{r allFig, fig.pos="!p", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap= "(ref:allFig)", fig.show = 'hold'}
+```
+
+(ref:allFig) Distribution of energy consumption displayed as letter value plots, illustrating harmonies and clashes, and how mappings change emphasis: **a** weekday/weekend faceted by quarter-of-year produces a harmony, **b** quarter-of-year faceted by weekday/weekend produces a harmony, **c** month-of-year faceted by quarter-of-year produces a clash, as indicated by the empty sets and white space.
+Placement within a facet should be done for primary comparisons.
+For example, arrangement in **a** makes it easier to compare across weekday type (x-axis) within a quarter (facet).
+It can be seen that in quarter 2, there is more mass occupied the lower tail on the weekends (letter value E corresponding to tail area 1/8) relative to that of the weekdays (letter value D 1/16), which corresponds to more days with lower energy use in this period.
+
+<!-- ; **b** across weekday/weekend faceted by quarter-of-year; **c** across quarter-of-year faceted by weekday/weekend.
+**a** shows a clash since there are empty combinations.
+**b** and **c** show harmonies since each quarter include both weekdays and weekends.
+It can be seen in **b** that for every quarter (mapped to facet), weekend and weekday consumption are fairly similar except for the second quarter.
+-->
+<!-- For example, the letter value D for weekdays mostly corresponds to letter value E for weekends, implying there are lower values of energy consumption during weekdays in the second quarter.
+This is probably because of lower temperatures in the second quarter compared to the first quarter (summer in Australia).
+**c** switches the granularities mapped to the x-axis and the facets and helps to compare quarters within weekdays and weekends.
+For example, the lower D and E letter values are most different across different quarters for weekdays and weekends respectively.
+-->
+<!-- For example, for weekdays the interquartile range of consumption reduces over the year, whereas this pattern is not true for weekends. -->
+
 #### Conditioning variable {-}
+
+When $C_i$ is mapped to the $x$ position and $C_j$ to facets, then the $A_k$ levels are juxtaposed and each $B_\ell$ represents a group/facet.
+Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups.
+Hence, in this case, the $A_k$'s are compared against each other within each group.
+With the mapping of $C_i$ and $C_j$ reversed, the emphasis will shift to comparing $B_\ell$ levels rather than $A_k$ levels.
+For example, \autoref{fig:allFig}b shows the letter value plot across weekday/weekend partitioned by quarters of the year and \autoref{fig:allFig}c shows the same two cyclic granularities with their mapping reversed.
+\autoref{fig:allFig}b helps us to compare weekday and weekend within each quarter and \autoref{fig:allFig}c helps to compare quarters within weekend and weekday.
+
 ## Applications {#sec:application}
+
 ### Smart meter data of Australia {#sec:smartmeter}
+
+Smart meters provide large quantities of measurements on energy usage for households across Australia.
+One of the customer trials [@smart-meter] conducted as part of the Smart Grid Smart City project in Newcastle and parts of Sydney provides customer level data on energy consumption for every half hour from February 2012 to March 2014.
+We can use this data set to visualize the distribution of energy consumption across different cyclic granularities in a systematic way to identify different behavioral patterns.
+
 #### Cyclic granularities search and computation {-}
+
+The tsibble object `smart_meter10` from R package `gravitas` [@R-gravitas] includes the variables `reading_datetime`, `customer_id` and `general_supply_kwh` denoting the index, key and measured variable respectively.
+The interval of this tsibble is 30 minutes.
+
+To identify the available cyclic time granularities, consider the conventional time deconstructions for a Gregorian calendar that can be formed from the 30-minute time index: half-hour, hour, day, week, month, quarter, half-year, year.
+In this example, we will consider the granularities hour, day, week and month giving six cyclic granularities "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month", read as "hour of the day", etc.
+To these, we add day-type ("wknd_wday") to capture weekend and weekday behavior.
+Now that we have a list of cyclic granularities to look at, we can compute them using the results in Section&nbsp;\ref{sec:cyclic-calendar}.
+
 #### Screening and visualizing harmonies {-}
+
+Using these seven cyclic granularities, we want to explore patterns of energy behavior.
+Each of these seven cyclic granularities can either be mapped to the x-axis or to facets.
+Choosing $2$ of the possible $7$ granularities, gives $^{7}P_2 = 42$ candidates for visualization.
+Harmonies can be identified among those $42$ possibilities to narrow the search.
+\autoref{tab:harmony-tab} shows $16$ harmony pairs after removing clashes and any cyclic granularities with more than $31$ levels, as effective exploration becomes difficult with many levels (Section&nbsp;\ref{sec:aesthetics}).
+
+```{r harmony-tab, echo=FALSE}
+```
+
+```{r bothcust, fig.cap = "(ref:bothcust)", fig.pos="!p"}
+```
+
+(ref:bothcust) Energy consumption of a single customer shown with different distribution displays, and granularity arrangements: hour of the day; and weekday/weekend.
+**a** The side-by-side boxplots make the comparison between day types easier, and suggest that there is generally lower energy use on the weekend.
+Interestingly, this is the opposite to what might be expected.
+Plots **b**, **c** examine the temporal trend of consumption over the course of a day, separately for the type of day.
+The area quantile emphasizes time, and indicates that median consumption shows prolonged high usage in the morning on weekdays.
+The violin plot emphasizes subtler distributional differences across hours: morning use is bimodal.
+
+A few harmony pairs are displayed in \autoref{fig:bothcust} to illustrate the impact of different distribution plots and reverse mapping.
+For each of \autoref{fig:bothcust}b and c, $C_i$ denotes day-type (weekday/weekend) and $C_j$ is hour-of-day.
+The geometry used for displaying the distribution is chosen as area-quantiles and violins in \autoref{fig:bothcust}b and c respectively.
+\autoref{fig:bothcust}a shows the reverse mapping of $C_i$ and $C_j$ with $C_i$ denoting hour-of-day and $C_j$ denoting day-type with distribution geometrically displayed as boxplots.
+
+In \autoref{fig:bothcust}b, the black line is the median, whereas the purple (narrow) band covers the 25th to 75th percentile, the orange (middle) band covers the 10th to 90th percentile, and the green (broad) band covers the 1st to 99th percentile.
+The first facet represents the weekday behavior while the second facet displays the weekend behavior; energy consumption across each hour of the day is shown inside each facet.
+The energy consumption is extremely skewed with the 1st, 10th and 25th percentile lying relatively close whereas 75th, 90th and 99th lying further away from each other.
+This is common across both weekdays and weekends.
+For the first few hours on weekdays, median energy consumption starts and continues to be higher for longer compared to weekends.
+
+The same data is shown using violin plots instead of quantile plots in \autoref{fig:bothcust}c.
+There is bimodality in the early hours of the day for weekdays and weekends.
+If we visualize the same data with reverse mapping of the cyclic granularities (\autoref{fig:bothcust}a), then the natural tendency would be to compare weekend and weekday behavior within each hour and not across hours.
+Then it can be seen that median energy consumption for the early morning hours is higher for weekdays than weekends.
+Also, outliers are more prominent in the later hours of the day.
+All of these indicate that looking at different distribution geometry or changing the mapping can shed light on different aspects of energy behavior for the same sample.
+
 ### T20 cricket data of Indian Premier League {#sec:cricket}
+
+Our proposed approach can be generalized to other hierarchical granularities where there is an underlying ordered index.
+We illustrate this with data from the sport cricket.
+Cricket is played with two teams of 11 players each, with each team taking turns batting and fielding.
+This is similar to baseball, wherein the _batsman_ and _bowler_ in cricket are analogous to a batter and pitcher in baseball.
+A _wicket_ is a structure with three sticks, stuck into the ground at the end of the cricket pitch behind the batsman.
+One player from the fielding team acts as the bowler, while another takes up the role of the _wicket-keeper_ (similar to a catcher in baseball).
+The bowler tries to hit the wicket with a _ball_, and the batsman defends the wicket using a _bat_.
+At any one time, two of the batting team and all of the fielding team are on the field.
+The batting team aims to score as many _runs_ as possible, while the fielding team aims to successively _dismiss_ 10 players from the batting team.
+The team with the highest number of runs wins the match.
+
+Cricket is played in various formats and Twenty20 cricket (T20) is a shortened format, where the two teams have a single _innings_ each, which is restricted to a maximum of 20 _overs_.
+An over will consist of 6 balls (with some exceptions).
+A single _match_ will consist of 2 innings and a _season_ consists of several matches.
+Although there is no conventional time component in cricket, each ball can be thought to represent an ordering over the course of the game.
+Then, we can conceive a hierarchy where the ball is nested within overs, overs nested within innings, innings within matches, and matches within seasons.
+Cyclic granularities can be constructed using this hierarchy.
+Example granularities include ball of the over, over of the innings, and ball of the innings.
+The hierarchy table is given in \autoref{tab:hierarchy-cric}.
+Although most of these cyclic granularities are circular by the design of the hierarchy, in practice some granularities are aperiodic.
+For example, most overs will consist of 6 balls, but there are exceptions due to wide balls, no-balls, or when an innings finishes before the over finishes.
+Thus, the cyclic granularity ball-of-over may be aperiodic.
+
+```{r hierarchy-cric}
+```
+
+The Indian Premier League (IPL) is a professional T20 cricket league in India contested by eight teams representing eight different cities in India.
+The IPL ball-by-ball data is provided in the `cricket` data set in the `gravitas` package for a sample of 214 matches spanning 9 seasons (2008 to 2016).
+
+Many interesting questions could be addressed with the `cricket` data set.
+For example, does the distribution of total runs depend on whether a team bats in the first or second innings?
+The Mumbai Indians (MI) and Chennai Super Kings (CSK) appeared in the final playoffs from 2010 to 2015.
+Using data from these two teams, it can be observed (\autoref{fig:cricex}a) that for the team batting in the first innings there is an upward trend of runs per over, while there is no clear upward trend in the median and quartile deviation of runs for the team batting in the second innings after the first few overs.
+This suggests that players feel mounting pressure to score more runs as they approach the end of the first innings, while teams batting second have a set target in mind and are not subjected to such mounting pressure and therefore may adopt a more conservative run-scoring strategy.
+
+Another question that can be addressed is if good fielding or bowling (defending) in the previous over affects the scoring rate in the subsequent over?
+To measure the defending quality, we use an indicator function on dismissals (1 if there was at least one wicket in the previous over, 0 otherwise).
+The scoring rate is measured by runs per over.
+\autoref{fig:cricex}b shows that no dismissals in the previous over leads to a higher median and quartile spread of runs per over compared to the case when there has been at least one dismissal in the previous over.
+This seems to be unaffected by the over of the innings (the faceting variable).
+This might be because the new batsman needs to "play himself in" or the dismissals lead the (not-dismissed) batsman to adopt a more defensive playstyle.
+Run rates will also vary depending on which player is facing the next over and when the wicket falls in the previous over.
+
+Here, wickets per over is an aperiodic cyclic granularity, so it does not appear in the hierarchy table. These are similar to holidays or special events in temporal data.
+
+```{r cricex, fig.cap= "(ref:cricex)", warning = FALSE, message = FALSE, fig.height=4, fig.width=8}
+```
+
+(ref:cricex) Examining distribution of runs per by innings, over of the innings and number of wickets in previous innings.
+Plot **a** displays distribution using letter value plots.
+A gradual upward trend in runs per over can be seen in innings 1, which is not present in innings 2.
+Plot **b** shows quantile plots of runs per over across an indicator of wickets in the previous over, faceted by current over.
+When a wicket occurred in the previous over, the runs per over tends to be lower throughout the innings.
+
 ## Discussion {#sec:discussion}
+
+Exploratory data analysis involves many iterations of finding and summarizing patterns.
+With temporal data available at ever finer scales, exploring periodicity can become overwhelming with so many possible granularities to explore.
+This work provides tools to classify and compute possible cyclic granularities from an ordered (usually temporal) index.
+We also provide a framework to systematically explore the distribution of a univariate variable conditional on two cyclic time granularities using visualizations based on the synergy and levels of the cyclic granularities.
+
+The `gravitas` package provides very general tools to compute and manipulate cyclic granularities, and to generate plots displaying distributions conditional on those granularities.
+
+A missing piece in the package `gravitas` is the computation of cyclic aperiodic granularities which would require computing aperiodic linear granularities first.
+A few R packages including `almanac`[@R-almanac] and `gs`[@R-gs] provide the tools to create recurring aperiodic events.
+These functions can be used with the `gravitas` package to accommodate aperiodic cyclic granularities.
+
+We propose producing plots based on pairs of cyclic granularities that form harmonies rather than clashes or near-clashes.
+A future direction of work could be to further refine the selection of appropriate pairs of granularities by identifying those for which the differences between the displayed distributions is greatest and rating these selected harmony pairs in order of importance for exploration.
+
 ## Acknowledgments {-}
+
+The Australian authors thank the ARC Centre of Excellence for Mathematical and Statistical Frontiers [(ACEMS)](https://acems.org.au/home) for supporting this research.
+Thanks to [Data61 CSIRO](https://data61.csiro.au/) for partially funding Sayani's research and Dr Peter Toscas for providing useful inputs on improving the analysis of the smart meter application.
+We would also like to thank Nicholas Spyrison for many useful discussions, sketching figures and feedback on the manuscript.
+The package `gravitas` was built during the [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/).
+More details about the package can be found at [sayani07.github.io/gravitas](https://sayani07.github.io/gravitas/).
+The Github repository,  [github.com/Sayani07/paper-gravitas](https://github.com/Sayani07/paper-gravitas), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
+This article was created with `knitr` [@knitr, @R-knitr] and `rmarkdown` [@rmarkdown, @R-rmarkdown].
+
 ## Supplementary Materials
+
+**Data and scripts:** Data sets and R code to reproduce all figures in this article (main.R).
+
+\noindent
+**R-package:** The ideas presented in this article have been implemented in the open-source R [@R-language] package `gravitas` [@R-gravitas], available from CRAN.
+The R-package facilitates manipulation of single and multiple-order-up time granularities through cyclic calendar algebra, checks feasibility of creating plots or drawing inferences for any two cyclic granularities by providing list of harmonies and recommends possible visual summaries through factors described in the article.
+Version 0.1.3 of the package was used for the results presented in the article and is available on Github (https://github.com/Sayani07/gravitas).
 
@@
 
-Placeholder
+ Cyclic temporal granularities, which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can be useful for measuring repetitive patterns in large univariate time series data.
+The granularities feed new approaches to exploring time series data.
+One use is to take pairs of granularities, and make plots of response values across the categories induced by the temporal deconstruction.
+However, when there are many granularities that can be constructed for a time period, there will also be too many possible displays to decide which might be the more interesting to display.
+This work proposes a new distance metric to screen and rank the possible granularities, and hence choose the most interesting ones to plot.
+The distance measure is computed for a single or pairs of cyclic granularities can can be compared across different cyclic granularities and also on a collection of time series.
+The methods are implemented in the open-source R package `hakear`.
+
+ ```{r external-hakear, include = FALSE, cache = FALSE}
+read_chunk("scripts/hakear.R")
+```
+
+```{r load}
+```
+
+```{r}
+set.seed(3215)
+```
 
 ## Introduction
+
+<!-- <introducing the problem> -->
+<!-- background of the problem -->
+
+<!--Exploratory data analysis, as coined by John W. Tukey [@Tukey1977-jx] involves many iterations of finding structures and patterns that allow the data to be informative.-->
+
+Cyclic temporal granularities [@Bettini1998-ed, @Gupta2021-hd] are <!--theoretical--> temporal deconstructions that define cyclic repetitions in time, e.g.
+hour-of-day, day-of-month, or regularly scheduled public holidays.
+These granularities form ordered or unordered categorical variables.
+An example of an ordered granularity is day-of-week, where Tuesday is always followed by Wednesday, and so on.
+An unordered granularity example is different week types in a semester, orientation, break, exam or regular classes in an academic calendar.
+Using granularities to  explore patterns in univariate time series can be considered to be examining the distribution of the measured variable across different categories of the cyclic granularities.
+
+Figure \ref{fig:onegran-new} electricity smart meter data plotted against two granularities (hour-of-day, month-of-year).
+The data was collected on a single household in Melbourne, Australia, over a six month period, as was used in @wang2020calendar.
+The categorical variable (granularity) is mapped to the x-axis, and the distribution of response variable is displayed using both side-by-side jittered dotplots and boxplots.
+From plot (a) it can be seen that energy consumption is higher during the morning hours (5-8), when members in the household wake up, and again in the evening hours (17-20) possibly when members get back from work.
+In addition, the largest variation in energy use is in the afternoon hours (12-16), as perceived from sizes of the boxes.
+From plot (b) the variability in energy usage is higher in Jan and Feb, possibly due to the usage of air conditioners on some days.
+The median usage is highest in January, dips in February-April and rises again in May-June, although not to the height of January usage.
+This might imply that this household does not use as much energy for heating as it does for air conditioning.
+A lot of households in Victoria use gas heating and hence the heater use might not be reflected in the electricity data.
+Many, many different displays could be constructed using different granularities,  day-of-week, day-of-month, weekday/weekend, etc.
+However, only a few might be interesting, that is, reveal important patterns in energy usage.
+Determining which displays which have "significant" distributional differences between categories of the cyclic granularity, and plotting only these, would make for efficient exploration.
+
+```{r calendar-elec, fig.height = 8, fig.cap = "Calendar displays faceted by two households are shown.
+The calendar displays are rich in information and has several components.
+For example, it can be observed easily that the level of energy consumption by household 2 is higher than household 4.
+Also, it gives an overview on when these households are away for holidays and which months consumptions are highest (Jan, Feb).
+However, when it comes to discerning periodic patterns, analyzing all possible periodic patterns through this display is overwhelming due to combination of linear and cyclic representation of time.
+For example, there is more going on in terms of peaks and troughs in id 2 compared to id 4, but it is not clear which all periodic patterns are dominant and if it differs between households."}
+
+```
+
+```{r onegran-new, fig.cap="A cyclic granularity can be considered to be a categorical variable, and used to break the data into subsets.
+Here, side-by-side boxplots overlaid on jittered dotplots explore the distribution of of energy use by a household for two different cyclic granularities: (a) hour-of-day and (b) and month-of-year.
+Daily peaks occur in morning and evening hours, indicating a working household, where members leave for and return from work.
+More volatility of usage in summer months (Jan, Feb) is probably due air conditioner use on just some days ."}
+```
+
+<!-- comparing one and two grans for one households -->
+\noindent Exploring the distribution of the measured variable across two cyclic granularities tends to provide more detailed information on its structure.
+For example, Figure \ref{fig:id2-new2}(a) slice down further by showing the usage distribution across hour-of-day conditional on month-of-year across two households (id 2 and 4).
+It shows the hourly usage over a day does not remain the same across months.
+Unlike other months, the 75th and 90th percentile for all hours of the day in January are    high, pretty close, and are not characterized by a morning and evening peak.
+The household in Figure \ref{fig:id2-new2}(b) has 90th percentile consumption higher in summer months relative to autumn or winter, but the 75th and 90th percentile are far apart in all months, implying that the second household resorts to air conditioning much less regularly than the first one.
+The differences seem to be more prominent across month-of-year (facets) than hour-of-day (x-axis) for this household, whereas they are prominent for both cyclic granularities for the first household.
+
+<!-- The problem and its dimension -->
+\noindent Are all of these four displays in Figures \ref{fig:onegran-new} and \ref{fig:id2-new2} useful in understanding the distributional difference in energy usage?
+Which ones are more useful than others?
+If $N_C$ is the total number of cyclic granularities of interest, the number of displays that could be potentially informative is $N_C$ when considering displays of the form in Figure \ref{fig:onegran-new}.
+The dimension of the problem, however, increases when considering more than one cyclic granularity.
+When considering displays of the form in Figure \ref{fig:id2-new2},
+there are $^{N_C}P_2$ possible pairwise plots exhaustively, with one of the two cyclic granularities acting as the conditioning variable. 
+This is huge and overwhelming for human consumption even for moderately large $N_C$.
+It could be immensely useful to make the transition from all potential displays to the ones that are informative across atleast one cyclic granularity.
+
+<!-- Moreover, the volatility in consumption are almost equal and high for all hours in Jan, whereas variability is most in the evening hours for other months.
+The distribution of consumption across different hours of the day look similar for Apr, May and Jun, characterized by a distinct morning, afternoon and evening peaks and higher variability in evening hours -->
+
+<!-- months Jan and Feb not only have a high median consumption, but also more variability, possibly due to the usage of air conditioning.
+The other months shows more consistent behavior with the energy consumption rising in May and June.
+-->
+
+<!-- Scagnostics literature, other literature and why we need more/different -->
+
+\noindent This problem is similar to Scagnostics (Scatterplot Diagnostics) by @tukey1988computer, which is used to identify meaningful patterns in large collections of scatterplots.
+Given a set of $v$ variables, there are $v(v-1)/2$ pairs of variables, and thus the same number of possible pairwise scatterplots.
+Therefore, even for small $v$, the number of scatterplots can be large, and scatterplot matrices (SPLOMs) could easily run out of pixels when presenting high-dimensional data.
+@Dang2014-tw and @wilkinson2005graph provide potential solutions to this, where few characterizations can be used to locate anomalies in density, shape, trend, and other features in the 2D point scatters.
+In this paper, we provide a solution to narrowing down the search from $^{N_C}P_2$ plots by introducing a new distance measure that can be used to detect significant distributional differences across cyclic granularities.<!-- harmonies and why it is not enough --> This work is a natural extension of our previous work [@Gupta2021-hd] that narrows down the search from $^{N_C}P_2$ plots by identifying pairs of granularities that can be meaningfully examined together (a "harmony"), or when they cannot (a "clash").
+However, even after excluding clashes, the list of harmonies left could be enormous for exhaustive exploration.
+Hence, there is a need to reduce the search even further by including only those harmonies which are informative enough.
+@inference and @Majumder2013-hb present methods for statistical significance testing of visual findings using human cognition as the statistical tests.
+In this paper, the visual discovery of distributional differences is facilitated by choosing a threshold for the proposed numerical distance measure, eventually selecting only those cyclic granularities for which the distributional differences are sufficient to make it an interesting display.
+
+<!-- \ref{fig:id2-new2} shows the distribution of energy consumption across hour-of-day conditional on month-of-year.
+\ref{fig:id2-new2}(a) shows that energy consumption is higher during the morning hours when members in the household wake up and then again higher in the evening hours possibly when members are back from work with them showing maximum variable behavior in the afternoon hours.
+\ref{fig:id2-new2}(b) shows the months Jan and Feb not only have a high median consumption, but also more variability, possibly due to the usage of air conditioning.
+The other months shows more consistent behavior with the energy consumption rising in May and June.
+\ref{fig:id2-new2} adds to this information, by showing that hours in Jan do not necessarily follow the same pattern across the day (due to Holidays, people visiting).
+Moreover, the volatility in consumption are almost equal and high for all hours in Jan, whereas variability is most in the evening hours for other months.
+The distribution of consumption across different hours of the day look similar for Apr, May and Jun, characterized by a distinct morning, afternoon and evening peaks and higher variability in evening hours.
+-->
+
+<!-- # interaction for two households, importance could be across one -->
+
+<!-- Even when it is known which all interactions to look at, all of them would not be interesting and that too will vary across different households.  -->
+
+<!-- calendar plots -->
+<!-- The authors show how hour-of-the-day interact with weekdays and weekends and then move on to use calendar display to show daily schedules.
+The calendar display has several components in it, which helps us to look at energy consumption across hour-of-day, day-of-week, week-of-month, and month-of-year at once.
+Some interaction of these cyclic granularities, for example, how  day-of-week relates to month-of-year, could also be interpreted from this display.
+This is a great way for having an overview of energy consumption.
+However, if one wants to understand the repetitiveness in energy behavior and how they interact in greater detail, it is not easy to comprehend the interactions of all cyclic granularities from this display, due to the combination of linear and cyclic representation of time.
+For example, this display might not be the best to understand how hour-of-the-day or month-of-year varies across week-of-the-month as well as with each other.
+Furthermore, it is not clear what all interactions of cyclic granularities should be read from this display as there could be many combinations that one can look at.
+Moreover, there could be non-conventional cyclic granularities, which could potentially become useful depending on the context.
+-->
+
+<!-- context when it could be useful is monitoring heart rates.
+There are devices that can detect each heartbeat and transfer the data to a receiver such as a watch or phone.
+These data could be available for a temporal scale as fine as a minute and it could be of interest to see regular patterns across any deconstruction of time coarser than a minute.
+-->
+
+```{r intro_all}
+
+```
+
+```{r id2-new2, fig.cap = "Distribution of energy consumption displayed through area quantile plots across two cyclic granularities month-of-year and hour-of-day and two households.
+The black line is the median, whereas the orange band covers the 25th to 75th percentile and the green band covers the 10th to 90th percentile.
+Difference between the 90th and 75th quantiles is less for (Jan, Feb) for the first household (a), suggesting that it is a more frequent user of air conditioner than the second household (b).
+Energy consumption for in (a) changes across both granularities, whereas for (b) daily pattern stays same irrespective of the months."}
+```
+
+```{r id4, fig.cap = "something2", eval = FALSE}
+
+```
+
+<!-- Again, look at \ref{fig:} c and d, where energy consumption for these two households are plotted against (weekend/weekday, week-of-month).
+Here, for both households, the pattern of energy consumption vary  across different weeks of the month irrespective of the fact it is a weekday or weekend.
+In that respect, the harmony pair (month-of-year, hour-of-day) seems to be more informative than (weekend/weekday, week-of-month) for the first household.
+-->
+
+<!-- Take an example of a data set which are observed at fine temporal scales, like that of NYC bike usage available at https://www.citibikenyc.com/system-data.
+We use the `nyc_bikes` data set from the R package `tsibbledata` which takes a sample of 10 bikes for the year 2018.
+The `start_time` and the `stop_time` are recorded to a fineness of seconds.
+We can look at pair of cyclic granularities (hour_day, wknd_wday) or (week_month, day_week) to see how these periodicities interact.
+But there could be other pairs that are important too.
+How to understand which pairs are sufficient to explore given the data set without losing much information about the data.
+-->
+
+<!-- When we need to understand the interplay of different periodicities in a high frequency temporal datasets, we have many choices to consider.
+In [@wang2020tsibble] and [@wang2020calendar], periodicities are explored across hour of the day and day of the week or months.
+But calendar effects are not restricted to conventional day-of-week or month-of-year deconstructions.
+-->
+
+<!-- If we have $n$ periodic linear granularities in the hierarchy table, then $n(n-1)/2$ circular or quasi-circular cyclic granularities could be constructed.  -->
+
+<!-- Our contributions in this paper are: -->
+
+ <!-- * We introduce a new distance measure for detecting periodic interactions. This induces data reduction which allows for identification of patterns, if any, in the time series data. -->
+
+ <!-- * We show that the distance metric could be used to rank the periodic patterns based on how well they capture the variation in the measured variable as they have been normalized for different number of comparisons.
+-->
+
+ <!-- * We device a framework for choosing a threshold, which will result in detection of only significantly interesting periodic patterns in the time series data. -->
+
+ <!-- *  -->
+
+ <!-- * ; -->
+
+ <!-- * show that the proposed distance metric could be used to rank the interesting patterns across different datasets and temporal granularities since they have been normalized for relevant parameters.
+-->
+
+\noindent The article is organized as follows.
+Section \ref{sec:computation-wpd} introduces a distance measure for detecting distributional difference in temporal granularities, which enables identification of patterns in the time series data;
+Section \ref{sec:rank-wpd} devises a selection criterion by choosing a threshold, which results in detection of only significantly interesting patterns.
+Section \ref{sec:simulations} provides some simulation study on the proposed methodology.
+Section \ref{sec:application-wpd} presents an application to residential smart meter data in Melbourne to show how the proposed methodology can be used to automatically detect temporal granularities along which distributional differences are significant.
+
+<!-- acts as a way to automatically detect periodic patterns in time series. -->
+
+<!-- Also, ranking the remaining harmony pairs based on how well they capture the variation in the measured variable could be potentially useful.   -->
+
+<!-- Talk about Scagnostics: Tukey -->
+
 ## Proposed distance measure {#sec:computation-wpd}
+
+We propose a measure called Weighted Pairwise Distances ($wpd$) to detect distributional differences in the measured variable across cyclic granularities.
+
 ### Principle
+
+The principle behind the construction of $wpd$ is explained through a simple example explained in Figure \ref{fig:null4by2}.
+Each of these figures describes a panel with $2$ x-axis categories and $3$ facet levels, but with different designs.
+Figure \ref{fig:null4by2}a has all categories drawn from N(0, 1) distribution for each facet.
+It is not an interesting display particularly, as distributions do not vary across x-axis or facet categories.
+Figure \ref{fig:null4by2}b has x categories drawn from the same distribution, but across facets the distributions are $3$ standard deviations apart.
+Figure \ref{fig:null4by2}c exhibits an exact opposite situation where distribution between the x-axis categories
+are $3$ standard deviations apart, but they do not change across facets.
+Figure \ref{fig:null4by2}d takes a step further by varying the distribution across both facet and x-axis categories by  $3$ standard deviations.
+If the panels are to be ranked in order of capturing maximum variation in the measured variable from minimum to maximum, then an obvious choice would be placing (a) followed by (b), (c) and then (d).
+It might be argued that it is not clear if (b) should precede or succeed (c) in the ranking.
+Gestalt theory suggests items placed at close proximity can be compared more easily, because people assume that they are in the same group and apart from other groups.
+With this principle in mind, display (b) is considered less informative as compared to display (c) in emphasizing the distributional differences.
+Considering one cyclic granularity, we would have only two design choices similar to (a) and (c), corresponding to no difference and significant differences between categories of that cyclic granularity only.
+The proposed measure $wpd$ is constructed in a way so that it could be used to rank panels of different designs as well as test if a design is interesting.
+This measure is aimed to be an estimate of the maximum variation in the measured variable explained by the panel.
+A higher value of  $wpd$ would indicate that the panel is interesting to look at, whereas a lower value would indicate otherwise.
+
+<!-- To compute the maximum variation of the measured variable within each panel, the distributional differences need to be measured for both within-group and between-group categories.
+Moreover, a tuning parameter specifying the weightage given to within-facet or between-facet categories can help to choose between designs like \ref{fig:null4by2}b and c.
+The maximum of all these weighted differences serves as an estimate of the maximum variation in the panel and is taken as the value of $wpd$.
+-->
+
+<!-- Larger differences imply stronger patterns, whereas small difference would imply that the underlying structure is not changing within or between group. -->
+
+<!-- A distance measure aimed to capture the structure of the measured variable in designs like \ref{fig:null4by2}, should ideally estimate these within-group and between-group variations. -->
+
+<!-- \noindent Intuitively, while finding a structure or measuring the strength of patterns in Figure \ref{fig:null4by2}, it makes sense to look for within-group and between-group variations.
+Larger variation would imply stronger patterns, whereas small variation would imply the underlying structure is not changing within or between group.
+-->
+<!-- Thus, a distance measure aimed to capture this structure should ideally estimate these within-group and between-group variations.
+One of the potential ways to do this is to measure the distances between distributions of the continuous random variable measured within and between groups, weigh them basis if they are within or between groups and then take the maximum of those distances as an estimate of the maximum variation in the structure.
+This section starts with possible ways of characterizing distributions and computing distances between them and then describe in details how the measure $wpd$ is defined.
+-->
+
+<!-- This is similar to @ieee-irish where the authors compute the Jensen Shannon distance between two density estimates by computing percentiles and stresses the advantages to working with percentiles rather than the raw data directly in case of missing observations.
+Working with quantiles also ensure that unsynchronized time series could be handled.
+-->
+
+<!-- Hence, with reference to the graphical design in @Gupta2021-hd, therefore the idea would be to rate a harmony pair higher if the variation between different levels of the x-axis variable is higher on an average across all levels of the facet variables.
+-->
+
+<!-- To elaborate further, look at the examples in Figure \ref{}, where Figure \ref{}a represents the panel design with distribution of each x categories drawn from N(5, 10) distribution.
+It could be observed that the graph is not particularly interesting, as there is no significant change in distribution between x-axis levels or facets.
+Figure \ref{}b represents the same panel design with no difference in distribution of x-axis categories within a facet, but different distribution of x-axis categories for different facets.
+For example, if there are 4 facet levels and 2 x-axis levels, data is generated in the way as described in Table \ref{}.
+Figure \ref{}b exhibits an exact opposite situation where the x-axis within facets are different but not across facets.
+-->
+<!-- Figure \ref{}d takes it further by varying the distribution across both facet and x-axis categories. -->
+
+```{r null4by2,fig.cap = "An example illustrating the principle of the proposed distance measure, displaying the distribution of a normally distributed variable in four panels each with $2$ x-axis categories and $3$ facet levels, but with different designs.
+Display (a) is not interesting as the distribution of the variable does not depend on x or facet categories.
+Display (b) and (c) are more interesting than (a) since there is a change in distribution either across facets (b) or x-axis (c).
+Display (d) is most interesting in terms of capturing structure in the variable as the distribution of the variable changes across both facet and x-axis variable.
+The value of our proposed distance measure is presented for each panel, the relative differences between which will be explained later in Section 3.2."}
+```
+
+<!-- A "significant" cyclic granularity have at least two categories which are significantly different from each other.
+It does not tell you which categories are statistically significantly different from each other -->
+
+<!-- in the same group would be important to bring out different patterns of the data. -->
+
 ### Notations
+
+\noindent Let the number of cyclic granularities considered in the display be $m$.
+The notations and methodology are described in detail for $m=2$.
+But it can be easily extended to $m>2$.
+Consider two cyclic granularities $A$ and $B$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$ with $A$ placed across x-axis and $B$ across facets.
+Let $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ be a continuous variable observed across $T$ time points.
+This  data structure with $J$ x-axis levels and $K$ facet levels is referred to as a $(J, K)$ panel.
+For example, a $(2, 3)$ panel will have cyclic granularities with 2 x-axis levels and 3 facet levels.
+Let the four elementary designs as described in Figure \ref{fig:null4by2} be $D_{null}$ (referred to as "null distribution") where there is no difference in distribution of $v$ for $A$ or $B$, $D_{var_f}$ denotes the set of designs where there is difference in distribution of $v$ for $B$ and not for $A$.
+Similarly, $D_{var_x}$ denotes the set of designs where difference is observed only across A.
+Finally, $D_{var_{all}}$ denotes those designs for which difference is observed across both $A$ and $B$.
+$m = 1$ is a special case of $m = 2$ with $J = 1$.
+
+```{r notations}
+
+```
+
+  <!-- Thus, by CLT, ${\chi^2}_{m} \tilde{} N(m, 2m)$, which would depend on the number of discretization used to approximate the continuous distribution.  -->
+  <!-- Then $b_n = 1-1/n$ quantile of the normal distribution and $a_n = 1/[n*\phi(b_n)]$ where $\phi$ is the normal density function. $n$ is the number of pairwise comparisons being made. -->
+
+<!-- The Jensen-Shanon distance between two probability distribution $p_1$ and $p_2$ is given by $$d = [D(p_1, r) + D(p_2, r)]/2 \quad where \quad r = (p_1 + p_2)/2$$ where, -->
+<!-- $$D(p_1,p_2) = \int^{\infty}_{-\infty}p_1(x)log\frac{p_1(x)}{p_2(x)}\,dx$$ is the Kullback-Leibler divergence between $p_1$ and $p_2$.   -->
+
+<!-- We call this measure of variation as  Median Maximum Pairwise Distances (MMPD). -->
+
+<!-- #### Distribution of Jensen-Shannon distances -->
+
+<!-- Jensen-Shannon distances (JSD) are distributed as chi-squared with $m$ df where we discretize the continuous distribution with $m$ discrete values.
+Taking sample percentiles to approximate the integral would mean taking $m = 99$.
+-->
+<!-- With large $m$, chi-squared is asymptotically normal by the CLT.
+Thus, by CLT, ${\chi^2}_{m} \tilde{} N(m, 2m)$, which would depend on the number of discretization used to approximate the continuous distribution.
+Then $b_n = 1-1/n$ quantile of the normal distribution and $a_n = 1/[n*\phi(b_n)]$ where $\phi$ is the normal density function.
+$n$ is the number of pairwise comparisons being made.
+-->
+
 ### Computation
+
+The computation of the distance measure $wpd$ for a panel involves characterizing distributions, computing distances between distributions, choosing a tuning parameter to specify the weightage for different group of distances and summarizing those weighted distances appropriately to estimate maximum variation.
+Furthermore, the data needs to be appropriately transformed to ensure that the value of $wpd$ emphasizes detection of distributional differences across categories and not across different data generating processes.
+
+<!-- but not when underlying distributions are different -->
+
+<!-- irrespective of the distribution from which the data is generated. -->
+
+<!-- is aimed to capture structure and patterns by estimating the maximum variation of the measured variable within a panel.  $wpd$ -->
+
+<!-- the distributional differences need to be measured for both within-group and between-group categories.
+Moreover, a tuning parameter specifying the weightage given to within-facet or between-facet categories can help to choose between designs like \ref{fig:null4by2}b and c.
+The maximum of all these weighted differences serves as an estimate of the maximum variation in the panel and is taken as the value of $wpd$.
+-->
+
 #### Data transformation
+
+The intended aim of $wpd$ is to capture differences in categories irrespective of the distribution from which the data is generated.
+Hence, as a pre-processing step, the raw data is normal-quantile transformed (NQT) (@Krzysztofowicz1997-bv), so that the quantiles of the transformed data follows a standard normal distribution.
+This sort of transformation is common in the fields of geo-statistics to make most asymmetrical distributed real world measured variables more treatable and normal-like (@Bogner2012-az).The empirical NQT involves the following steps:
+ 1. The sample of measured variable $v$ is sorted from the smallest to the largest observation $v_{(1)},\dots, v_{(i)}, .., v_{(n)}$.  
+ 2. The cumulative probabilities $p_{(1)},\dots, p_{(i)}, .., p_{(n)}$ are estimated using a plotting position like $i/(n + 1)$ such that $p_{(i)} = P(v \leq v_{(i)})$.  
+ 3.
+Each observation $v_{(i)}$ of $v$ is transformed into observation $v^*(i) = Q^{-1}(p(i))$ of the standard normal variate $v^*$, with $Q$ denoting the standard normal distribution and $Q^{-1}$ its inverse.
+
 #### Characterising distributions
+
+Multiple observations of $v$ correspond to the subset  $v_{jk} = \{s: A(s) = j, B(s) = k\}$.
+The number of observations might vary widely across subsets due to the structure of the calendar, missing observations or uneven locations of events in the time domain.
+in this paper, quantiles of $v_{jk}$'s are chosen as a way to characterize distributions for the category $(a_j, b_k)$, $\forall j\in \{1, 2, \dots, J\}, k\in \{1, 2, \dots, K\}\}$.
+The quantile of a distribution with probability $p$ is defined as $Q(p)=F^{-1}(p) = inf\{x: F(x) >p\}$, $0<p< 1$ where $F(x)$ is the distribution function.
+There are two broad approaches to quantile estimation, viz, parametric and non-parametric.
+Sample quantiles (@Hyndman1996-ty) are used for estimating population quantiles in a non-parametric setup, which is desirable because of less rigid assumptions made about the nature of the underlying distribution of the data.
+The default quantile chosen in this paper is percentiles computed for $p = {0.01, 0.02, \dots, 0.99}$, where for example, the $99^{th}$ percentile would be the value corresponding to $p=0.99$ and hence 99% of the observations would lie below that.
+
+<!-- The benefit of using a non-parametric estimator is that there are less rigid assumptions made about the nature of the underlying distribution of the data.
+Sample quantiles could be used for estimating population quantiles in a non-parametric setup.
+describes the many ways of defining sample quantiles and recommends the use of median-unbiased estimator because of _desirable properties of a quantile estimator and can be defined independently of the underlying distribution._.
+-->
+
+<!-- in this paper. -->
+
+<!-- Each $v_{jk}$'s $\forall j\in \{1, 2, \dots, J\}, k\in \{1, 2, \dots, K\}\}$ are assumed to be drawn from a continuous probability distribution and have certain characteristics.
+Shape, central tendency, and variability are the common characteristics used to describe a distribution.
+-->
+
+<!-- Mean, median or mode are generally used to describe the center of the distribution, while range, standard deviation, quantiles, standard errors and confidence intervals are used to describe variability.
+-->
+
+<!-- The quantile of a distribution with probability $p$ is defined as $Q(p)=F^{-1}(p) = inf\{x: F(x) >p\}$, $0<p< 1$ where $F(x)$ is the distribution function.
+There are two broad approaches to quantile estimation, viz, parametric and non-parametric.
+The benefit of using a non-parametric estimator is that there are less rigid assumptions made about the nature of the underlying distribution of the data.
+Sample quantiles could be used for estimating population quantiles in a non-parametric setup.
+@Hyndman1996-ty describes the many ways of defining sample quantiles and recommends the use of median-unbiased estimator because of _desirable properties of a quantile estimator and can be defined independently of the underlying distribution._.
+The `stats::quantile()` function in @R-language could be used for practical implementation where type = 8 refers to the algorithm corresponding to the median-unbiased estimator.
+The default quantile chosen in this paper is percentiles computed for $p = {0.01, 0.02, \dots, 0.99}$, where for example, the $99^{th}$ percentile would be the value corresponding to $p=0.99$ and hence 99% of the observations would lie below that.
+-->
+
 #### Distance between distributions
+
+<!-- One of the most important class of divergence is the f-divergence and includes measures like Kullback-Leibler divergence, Hellinger distance etc.
+The continuous version of f -divergence is given by -->
+<!-- $$D_f(P||Q) := \int q(x)f(\frac{p(x)}{q(x)})$$, where -->
+<!-- $f : [0,\infty) \rightarrow R \cup \{\infty\}$ is a continuous convex function, and $f(1) = 0$.  -->
+
+One of the most common ways to measure divergence between distributions is the Kullback-Leibler (KL) divergence [@Kullback1951-jy].
+The KL divergence denoted by $D(q_1||q_2)$ is a non-symmetric measure of the difference between two probability distributions $q_1$ and $q_2$ and is interpreted as the amount of information lost when $q_2$ is used to approximate $q_1$.
+The KL divergence, however, is not symmetric and hence can not be considered as a "distance" measure.
+The Jensen-Shannon divergence [@Menendez1997-in] based on the Kullback-Leibler divergence is symmetric and has a finite value.
+Hence, in this paper, the pairwise distances between the distributions of the measured variable are obtained through the square root of the Jensen-Shannon divergence, called Jensen-Shannon distance (JSD) and is defined by,
+
+$$JSD(q_1||q_2) = \frac{1}{2}D(q_1||M) + \frac{1}{2}D(q_2||M)$$
+where $M = \frac{q_1+q_2}{2}$ and 
+$D(q_1||q_2) := \int^\infty_{-\infty} q_1(x)f(\frac{q_1(x)}{q_2(x)})$ is the KL divergence between distributions $q_1$ and $q_2$.
+Other common measures of distance between distributions are Hellinger distance, total variation distance and Fisher information metric.
+
+<!-- Furthermore, these distances are distributed as chi-squared with $m$ degrees of freedom (@Menendez1997-in), if the continuous distribution is being discretized with $m$ discrete values.
+Taking sample percentiles to approximate the integral would mean taking $m = 99$.
+As the degrees of freedom $m$ get larger, the chi-square distribution approaches the normal distribution.
+-->
+
 #### Within-facet and between-facet distances
+
+\noindent Pairwise distances could be within-facets or between-facets for $m \geq 2$.
+Figure \ref{fig:distance-explain} illustrates how they are defined.
+Pairwise distances are within-facets when $b_{k} = b_{k'}$, that is, between pairs of the form $(a_{j}b_{k}, a_{j'}b_{k})$ as shown in panel (3) of Figure \ref{fig:distance-explain}.
+If categories are ordered (like all temporal cyclic granularities), then only distances between pairs where $a_{j'} = (a_{j+1})$ are considered (panel (4)).
+Pairwise distances are between-facets when they are considered between pairs of the form $(a_{j}b_{k}, a_{j}b_{k'})$.
+Number of between-facet distances would be $^{K}C_2*J$ and number of within-facet distances are $K*(J-1)$ (ordered) and $^{J}C_2*K$ (un-ordered).
+
 #### Tuning parameter
+
+A tuning parameter specifying the weightage given to the within-facet or between-facet categories can help to balance weightage between designs like \ref{fig:null4by2}(b) and (c).
+The tuning parameters should be chosen such that $\sum_{i=1}^m \lambda_i = 1$.
+When $m=2$, following the principle of Gestalt theory, $\lambda = \frac{2}{3} = 0.67$ is chosen to put a relative weightage of $2:1$ for within-facet and between-facet distances.
+No human experiment is conducted to justify this ratio, however, typically a tuning parameter $\lambda>0.5$ would tend to upweigh the within-facet distances and that with $<0.5$ would upweigh the between-facet distances (refer to the Supplementary paper for more details).
+For $m = 1$, there are no conditioning variables or groups, and hence $\lambda = 1$.
+
 #### Raw distance measure
+
+The raw distance measure, denoted by $wpd_{raw}$, is computed after combining all the weighted distance measures appropriately.
+First, NQT is performed on the measured variable $v_t$ to obtain $v^*_t$ (_data transformation_).
+Then, for a fixed harmony pair $(A, B)$, percentiles of $v^*_{jk}$ are computed and stored in $q_{jk}$ (_distribution characterization_).
+This is repeated for all pairs of categories of the form $(a_{j} b_{k}, a_{j'}b_{k'}): \{a_j: j = 1, 2, \dots, J\}, B = \{ b_k: k = 1, 2, \dots, K\}$.
+The pairwise distances between pairs $(a_{j} b_{k}, a_{j'}b_{k'})$  denoted by $d_{(jk, j'k')} = JSD(q_{jk}, q_{j'k'})$ is computed (_distance between distributions_).
+The pairwise distances (_Within-facet and between-facet_) are transformed using a suitable _tuning parameter_ ($0<\lambda<1$) depending on if they are within-facet($d_w$) or between-facets($d_b$) as follows:
+\begin{equation}\label{def:raw-wpd}
+d*_{(j,k), (j'k')} =
+\begin{cases}
+\lambda d_{(jk), (j'k')},& \text{if } d = d_w\\
+(1-\lambda)  d_{(jk), (j'k')},              & \text{if } d = d_b\\
+\end{cases}
+\end{equation}
+
+\noindent The $wpd_{raw}$ is then computed as $$wpd = max_{j, j', k, k'}(d*_{(jk), (j'k')}) \forall j, j' \in \{1, 2, \dots, J\}, k, k' \in \{1, 2, \dots, K\}$$
+
+\noindent The statistic "maximum" is chosen to combine the weighted pairwise distances since the distance measure is aimed at capturing the maximum variation of the measured variable within a panel.
+The statistic "maximum" is, however, affected by the number of comparisons (resulting pairwise distances).
+For example, for a $(2, 3)$ panel, there are $6$ possible subsets of obervations corresponding to the combinations $(a_1, b_1), (a_1, b_2), (a_1, b_3), (a_2 ,b_1), (a_2 ,b_2), (a_2, b_3)$, whereas for a $(2, 2)$ panel, there are only $4$ possible subsets $(a_1, b_1), (a_1, b_2), (a_2 ,b_1), (a_2 ,b_2)$.
+Consequently, the measure would have higher values for the panel $(2, 3)$ as compared to $(2, 2)$, since maximum is taken over higher number of pairwise distances.
+
 #### Adjusting for the number of comparisons {#sec:norm-wpd}
+
+Ideally, it is desired that the proposed distance measure takes a higher value only if there is a significant difference between distributions across categories, and not because the number of categories $J$ or $K$ is high.
+That is, under designs like $D_{null}$, their distribution should not differ for a different number of categories.
+Only then the distance measure could be compared across panels with different levels.
+This calls for an adjusted measure, which normalizes for the different number of comparisons.
+We denote it by $wpd$.
+Two approaches for adjusting the number of comparisons are discussed, both of which are substantiated using simulations.
+The first one defines an adjusted measure $wpd_{perm}$ based on the permutation method to remove the effect of different comparisons.
+The second approach fits a model to represent the relationship between $wpd_{raw}$ and the number of comparisons and defines the adjusted measure ($wpd_{glm}$) as the residual from the model.
+
+_Permutation approach_
+
+This method is somewhat similar in spirit to bootstrap or permutation tests, where the goal is to test the hypothesis that the groups under study have identical distributions.
+This method accomplishes a different goal of finding the null distribution for different groups (panels in our case) and standardizing the raw values using that distribution.
+The values of $wpd_{raw}$ is computed on many ($nperm$) permuted data sets and stored in $wpd_{perm-data}$.
+Then $wpd_{perm}$ is computed as follows:
+
+\begin{align*}
+wpd_{perm} = \frac{(wpd_{raw} - \bar{wpd_{perm-data}})}{sd(wpd_{perm-data})}
+\end{align*}
+
+\noindent where $\bar{wpd_{perm-data}}$ and $sd(wpd_{perm-data})$ are the mean and standard deviation of $wpd_{perm-data}$ respectively.
+Standardizing  $wpd$ in the permutation approach ensures that the distribution of $wpd_{perm}$ under $D_{null}$ has the same $mean = 0$ and $\sigma^2_{perm} = 1$ across all comparisons.
+While this works successfully to make the location and scale similar across different $nx$ and $nfacet$, it is computationally heavy and time consuming, and hence less user friendly when being actually used in practice.
+Hence, another approach to adjustment, with potentially less computational time, is proposed.
+
+<!-- for the $l^{th}$ panel and -->
+
+_Modeling approach_
+
+In this approach, a Gamma generalized linear model (GLM) for $wpd_{raw}$ is fitted with number of comparisons as the explanatory variable.
+Since, $wpd_{raw}$ is a Jensen-Shannon distance, it follows a Chi-square distribution (@Menendez1997-in), which is a special case of Gamma distribution.
+Furthermore, the mean response is bounded, since any JSD is bounded by $1$ given that base $2$ logarithm is used (@Lin1991-pt).
+Hence, by @Faraway2016-uk, an inverse link is used for the model, which is of the  form $y = a+b*log(z) + e$, where $y = wpd_{raw}$, $z = (nx*nfacet)$ is the number of groups and $e$ are idiosyncratic errors.
+Let $E(y) = \mu$ and $a + b*log(z) = g(\mu)$ where $g$ is the link function.
+Then $g(\mu)=  1/\mu$ and $\hat \mu  = 1/(\hat a + \hat b log(z))$.
+The residuals from this model $(y-\hat y) = (y-1/(\hat a + \hat b log(z)))$ would be expected to have no dependency on $z$.
+Thus, $wpd_{glm}$ is chosen as the residuals from this model and is defined as:
+$$wpd_{glm} = wpd_{raw} - 1/(\hat a + \hat b*log(nx*nfacet))$$
+The distribution of $wpd_{glm}$ under $D_{null}$ will have $mean = 0$, since it is the residuals from the model, and a constant variance $\sigma^2_{glm}$, which might not equal 1.
+
+_Combination approach_
+
+The simulation results (in Section \ref{sec:simulations}) show that the distribution of $wpd_{glm}$ under null is similar for high $nx$ and $nfacet$ (levels higher than $5$) and not so much for lower $nx$ and $nfacet$.
+Hence, a combination approach is proposed which chooses permutation approach for categories with smaller levels and modeling approach for categories with higher levels.
+This ensures that the computational load of the permutation approach is alleviated while maintaining similar null distribution across different categories.
+This approach, however, requires that the adjusted variables from the two approaches are brought to the same scale.
+We define $wpd_{glm-scaled} = wpd_{glm}*\sigma^2_{perm}/\sigma^2_{glm}$ as the transformed $wpd_{glm}$ with a similar scale as $wpd_{perm}$.
+The adjusted measure from the combination approach, denoted by $wpd$ is then defined as follows:
+
+\begin{equation}
+wpd =
+\begin{cases}
+wpd_{perm},& \text{if } J, K <=5\\
+wpd_{glm-scaled}             & \text{otherwise}\\
+\end{cases}
+\end{equation}
+
+<!-- ### Algorithm -->
+
+<!-- The distance measure $wpd$ between two cyclic granularities $A$ and $B$ is aimed to capture the strength of the structure by estimating the maximum within-group and between-group variations.
+Furthermore,  -->
+
+<!-- The steps employed for computing $wpd$ is summarized as follows: -->
+
+<!-- The values of $wpd$ indeed depend upon the underlying distribution.
+This is also not desirable as the intended aim of $wpd$ is to capture differences in categories irrespective of the distribution from which the data is generated.
+The steps employed for computing the distance measure is summarized as follows: -->
+
+<!-- Hence, differences in distributions of the measured variable are computed between all pairs of categories $(a_{j} b_{k}, a_{j'}b_{k'}): j = 1, 2, \dots, J\}, B = \{ b_k: k = 1, 2, \dots, K\}$.
+Percentiles of $v_{jk}$ and $v_{j'k'}$ are computed and stored in $q_{jk}$ and $q_{j'k'}$ respectively.
+Then the pairwise distances between pairs $(a_{j} b_{k}, a_{j'}b_{k'})$ be denoted as $d_{(jk, j'k')} = JSD(q_{jk}, q_{j'k'})$ is computed.
+Pairwise distances could be within-facets or between-facets.
+Figure \ref{fig:distance-explain} illustrates how the within-facet or between-facet distances are defined.
+Pairwise distances are within-facets ($d_{w}$) when $b_{k} = b_{k'}$, that is, between pairs of the form $(a_{j}b_{k}, a_{j'}b_{k})$ as shown in panel (3) of Figure \ref{fig:distance-explain}.
+If categories are ordered (like all temporal cyclic granularities), then only distances between pairs where $a_{j'} = (a_{j+1})$ are considered (panel (4)).
+Pairwise distances are between-facets ($d_{b}$) when they are considered between pairs of the form $(a_{j}b_{k}, a_{j}b_{k'})$.
+Number of between-facet distances would be $^{K}C_2*J$ and number of within-facet distances are $K*(J-1)$ (ordered) and $^{J}C_2*K$ (un-ordered).
+The pairwise distances $d_{(jk, j'k')}$ are transformed using a suitable tuning parameter ($0<\lambda<1$) depending on if they are $d_{b}$ or $d_{w}$ as follows: -->
+
+```{r distance-explain, fig.cap = "Within and between-facet distances shown for two cyclic granularities A and B, where A is mapped to x-axis and B is mapped to facets.
+The dotted lines represent the distances between different categories.
+Panel 1) and 2) show the between-facet distances.
+Panel 3) and 4) are used to illustrate within-facet distances when categories are un-ordered or ordered respectively.
+When categories are ordered, distances should only be considered for consecutive x-axis categories.
+Between-facet distances are distances between different facet levels for the same x-axis category, for example, distances between {($a_1$,$b_1$) and ($a_1$, $b_2$)} or {($a_1$,$b_1$) and ($a_1$, $b_3$)}."}
+
+```
+
+<!-- 5. Compute distribution of maximum distance ($M$) by shuffling the data $200$ times and finishing Steps1-4 in each case. Scale $M$ using the mean and sd of the distribution.
+
+5. Use Steps 1-4 to compute maximum distance for $\forall k \in  \{1, 2, \ldots, K\}$.
+
+6. Compute the distance measure MMPD_raw = median $(M_1, M_2, \dots, M_K)$.
+
+<!-- All of these distances are then aggregated by taking the maximum from these distances to obtain $wpd$. -->
+
+<!-- _j$ maps index set to a set $\{B_\ell \mid \ell =1,\dots,L\}$.
+Here, $A_k$ and $B_\ell$ are the levels/categories corresponding to $C_i$ and $C_j$ respectively.
+Let $S_{k\ell}$ be a subset of the index set such that for all $s \in S_{k\ell}$, $C_i(s) = A_k$ and $C_j(s) = B_\ell$.
+There are $KL$ such data subsets, one for each combination of levels ($A_k$, $B_\ell$).
+Moreover, consider that in the graphical space, $C_i$ is mapped to facets and $C_j$ is mapped to x-axis.
+-->
+
+ <!-- old -->
+<!-- Consider two cyclic granularities $C_i$ and $C_j$, such that $C_i$ maps index set to a set $\{A_k \mid k=1,\dots,K\}$ and $C_j$ maps index set to a set $\{B_\ell \mid \ell =1,\dots,L\}$.
+Here, $A_k$ and $B_\ell$ are the levels/categories corresponding to $C_i$ and $C_j$ respectively.
+Let $S_{k\ell}$ be a subset of the index set such that for all $s \in S_{k\ell}$, $C_i(s) = A_k$ and $C_j(s) = B_\ell$.
+There are $KL$ such data subsets, one for each combination of levels ($A_k$, $B_\ell$).
+Moreover, consider that in the graphical space, $C_i$ is mapped to facets and $C_j$ is mapped to x-axis.
+-->
+
+<!-- The algorithm employed for computing the distance measure is summarized as follows: -->
+
+<!-- 1. Fix harmony pair $(C_i, C_j)$. -->
+
+<!-- 2. Fix $k$. Then there are $L$ groups corresponding to level $A_k$ of $C_i$. -->
+
+<!-- 3. Compute  $m = \binom{L}{2}$ pairwise distances between distributions of $L$ unordered levels and $m = L-1$ pairwise distances for $L$ ordered categories. -->
+
+<!-- 4. Identify maximum within the $m$ computed distances. -->
+
+<!-- <!-- 5. Compute distribution of maximum distance ($M$) by shuffling the data $200$ times and finishing Steps1-4 in each case. Scale $M$ using the mean and sd of the distribution.  -->
+
+<!-- 5. Use Steps 1-4 to compute maximum distance for $\forall k \in  \{1, 2, \ldots, K\}$. -->
+
+<!-- 6. Compute the distance measure MMPD_raw = median $(M_1, M_2, \dots, M_K)$. -->
+
+<!-- \begin{algorithm}[!thb] -->
+<!-- 	\caption{Calculation for a raw distance measure between two cyclic granularities $A = \{ a_j: j = 1, 2, \dots, J\}$, $B = \{ b_k: k = 1, 2, \dots, K\}$ with $A$ placed across x-axis and $B$ across facets.}\label{alg:scoreopt}.
+-->
+<!-- 	\begin{algorithmic}[1] -->
+<!-- 		\Procedure{RawMMPD}{$A = \{ a_j: j = 1, 2, \dots, J\}$, $B = \{ b_k: k = 1, 2, \dots, K\}$, $v = \{ v_t: t = 1, 2, \dots, T\}$}. -->
+<!-- 		\For{$k=1:K$, $j=1:J$} -->
+<!-- 		\State Find distances between pairs of all possible combinations of categories $(a_jb_k,a_j'b_k')$ by computing JSD between quantiles of the measured variable $q(v)$ across these combinations.
+-->
+<!-- 		\State $d \gets JSD(\tilde{q(v)_{a_{j}b_{k}}},\tilde{q(v)_{a_{j}'b_{k}'}})$ -->
+<!-- 		\If {$b_k = b_k'$} -->
+<!-- 		\State $d* \gets \lambda d$  \Comment{upweight within-facet distances} -->
+<!-- 		\Else -->
+<!-- 		\State $d* \gets 1/\lambda d $ \Comment{downweight across-facet distances} -->
+<!-- 		\EndFor -->
+<!-- 		\State Set the raw distance measure as $max(d*)$ where max is taken over all $j, j', k, k'$. -->
+
+<!-- 		\EndProcedure -->
+<!-- 	\end{algorithmic} -->
+<!-- \end{algorithm} -->
+
+<!-- A stronger measure "max" is chosen for aggregating x-axis categories compared to "median" for aggregating facet categories as the measure is intended to put more importance in pointing towards distributional differences between x-axis categories than for facet categories.
+-->
+
+<!-- ## Properties of $wpd$ -->
+
+<!-- Simulations were carried out to explore the behavior of $wpd$ under the following factors that could potentially impact the values of $wpd$: $nx$, $nfacet$, $\lambda$, $\omega$, $dist$ (normal/non-normal distributions with different location and scale), $ntimes$, and $designs$ and results are presented in two parts.
+The dependence of $wpd$ on $nx$ and $nfacet$ under $D_{null}$ is presented here, which lays the foundation for the next section.
+The rest of the results that discuss the relationship of the $wpd$ with other factors is presented in details in the Supplementary section of the paper.
+They show that the designs $D_{var_f}$ and $D_{var_x}$ intersect at $\lambda = 0.5$ and hence for up-weighing designs of the form $D_{var_x}$, $\lambda = 0.67$ has been considered for computation of $wpd$ in the rest of the paper.
+-->
+
+<!-- - $nx$ (number of levels of x-axis) -->
+<!-- - $nfacet$ (number of levels of facets) -->
+<!-- - $\lambda$ (tuning parameter) -->
+<!-- - $\omega$ (increment in each panel design) -->
+<!-- - $dist$ (normal/non-normal distributions with different location and scale) -->
+<!-- - $n$ (sample size for each combination of categories) -->
+<!-- - $designs$ ($D_{null}$, $D_{var_f}$, $D_{var_x}$ and $D_{var_{all}}$) -->
+
+<!-- - $nsim$ (number of simulations) -->
+<!-- - $nperm$ (number of permutations of data)   -->
+
+<!-- ### Simulation design{#sec:sim-study} -->
+
+<!-- Observations are generated from a Gamma(2,1) distribution for each combination of $nx$ and $nfacet$ from the following sets: $nx = nfacet = \{2, 3, 5, 7, 14, 20, 31, 50\}$ to cover a wide range of levels from very low to moderately high.
+Each combination is being referred to as a _panel_.
+That is, data is being generated for each of the panels $\{nx = 2, nfacet = 2\}, \{nx = 2, nfacet = 3\}, \{nx = 2, nfacet = 5\},  \dots, \{nx = 50, nfacet = 31\}, \{nx = 50, nfacet = 50\}$.
+For each of the $64$ panels, $ntimes = 500$ observations are drawn for each combination of the categories.
+That is, if we consider the panel $\{nx = 2, nfacet = 2\}$, $500$ observations are generated for each of the combination of categories from the panel, namely, $\{(1, 1), (1, 2), (2, 1), (2, 2)\}$.
+The values -->
+<!-- of $wpd$ is obtained for each of the panels.
+This design corresponds to $D_{null}$ as each combination of categories in a panel are drawn from the same distribution.
+Furthermore, the data is simulated for each of the panels $nsim=200$ times, so that the distribution oaf $wpd$ under $D_{null}$ could be observed.
+$wpd_{l, s}$ denotes the value of $wpd$ obtained for the $l^{th}$ panel and $s^{th}$ simulation.
+-->
+
+<!-- ### Results -->
+
+<!-- Figure \ref{fig:raw} shows the distribution of $wpd$ plotted across different $nx$ and $nfacet$ categories.
+Since under $D_{null}$, there is no difference in distributions across different categories, we expect the distance measure $wpd$ to reflect that as well and have the same distribution across categories.
+But Figure \ref{fig:raw} shows that both the location and scale of the distributions change across panels.
+This is not desirable under $D_{null}$ as it would mean comparisons of $wpd$ values is not appropriate across different $nx$ and $nfacet$.
+Figure \ref{fig:quadratic} shows how the median of $wpd$ varies with the total number of distances $nx*nfacet$ for each panel.
+The median increases abruptly for lower values of $nx*nfacet$ and slowly for higher $nx*nfacet$.
+-->
+
+<!-- Furthermore, simulation results also show that the values of $wpd$ indeed depend upon the underlying distribution.
+This is also not desirable as the intended aim of $wpd$ is to capture differences in categories irrespective of the distribution from which the data is generated.
+-->
+
+<!-- last writeup -->
+<!-- In the linear model approach, $wpd\in R$ was assumed, whereas, $wpd$ is a Jensen-Shannon Distance (JSD) and lies between 0 and 1 (@JSD).
+Furthermore, JSD follows a Chi-square distribution, which is a special case of Gamma distribution.
+Therefore, a generalized linear model could be fitted instead of a linear model to allow for the response variable to follow a Gamma distribution.
+The inverse link is used when we know that the mean response is bounded, which is applicable in our case since $0 \leq wpd\leq 1$.
+-->
+
+<!-- We fit a Gamma generalized linear model with the inverse link which is of the form:   -->
+<!-- $$y_l = a+b*log(z_l) + e_l$$, where $y_l = median_m(wpd_{l, m})$, $z_l$ is the $l^{th}$ panel and $e_l$ are idiosyncratic errors.
+Let $E(y) = \mu$ and $a + b*log(z) = g(\mu)$ where $g$ is the link function.
+Then $g(\mu)=  1/\mu$ and $\hat \mu  = 1/(\hat a + \hat b log(z))$.
+The residuals from this model $(y-\hat y) = (y-1/(\hat a + \hat b log(z)))$ would be expected to have no dependency on $z$.
+Thus, $wpd_{glm}$ is chosen as the residuals from this model and is defined as: -->
+<!-- $wpd_{glm} = wpd - 1/(\hat a + \hat b*log(nx*nfacet))$. -->
+
+<!-- which is more approximate than exact but still has the similar accuracy when compared to the permutation approach. -->
+
+<!-- use later -->
+<!-- The measure $wpd$ could potentially lead to comparison of the measure across different panels and also help distinguishing the interesting panels from a data set.
+We discuss two approaches for normalization, both of which are substantiated using simulations.
+-->
+
+<!-- ## Methodology -->
+
+<!-- The transformed ${wpd}$ which is normalized for the values of $nx$ and $nfacet$ is denoted by $wpd$.
+Two approaches have been employed - the first one involves a permutation method to make the distribution of the transformed $wpd$ similar for different comparisons and the second one fits a model to represent the relationship between the two variables and defines $wpd$ as the residual of the model.
+-->
+
+<!-- ### Notations -->
+
+<!-- Let $\{nx_{i}, i = 1, 2, \dots, nx\}$, $\{nfacet_{j}, j = 1, 2, \dots, nfacet\}$ be the set of x-axis and facet categories respectively.
+Each combination of $nx_{i}$ and $nfacet_{j}$is being referred to as a _panel_.
+Then the total number of panel is $nx*nfacet$.
+Let the total number of pairwise distances that could result in each panel be $\{z_k, k = 1, 2 , \dots, nx*nfacet\}$.
+Here, $\{z_1 = nx_1*nfacet_1\}$, $\{z_2 = nx_2*nfacet_2\}$ and $\{z_k = nx*nfacet\}$.
+Now, let $\{x_{k,l}, k = 1, 2, \dots, nx*nfacet, l = 1, 2, \dots, nsim\}$ denote the values of $wpd_{raw}$ obtained from the simulation study for $k^{th}$ panel in the $i^{th}$ simulation.
+Hence, for each of those $k$ panel, we have $nsim$ values of $wpd_{raw}$.
+-->
+
+<!-- use soon -->
+<!-- ## Properties -->
+
+<!-- This section reports the results of a simulation study that was carried out to evaluate the behavior of $wpd$ under different designs and other potential factors.
+The behavior of $wpd$ is explored in designs where there is in fact difference in distribution between facet categories ($D_{var_f}$) or across x-categories ($D_{var_x}$) or both ($D_{var_{all}}$).
+Using $\omega =  \{1, 2, \dots, 10\}$ and $\lambda = seq(from  = 0.1, to = 0.9, by = 0.05)$, observations are drawn from a N(0,1) distribution for each combination of $nx$ and $nfacet$ from the following sets: $nx = nfacet =  \{2, 3, 5, 7, 14, 20, 31, 50\}$.
+$ntimes = 500$ is assumed for this setup as well.
+Furthermore, to generate different distributions across different combination of facet and x levels, the following method is deployed - suppose the distribution of the combination of first levels of $x$ and $facet$ category is $N(\mu,\sigma)$ and $\mu_{jk}$ denotes the mean of the combination $(a_jb_k)$, then $\mu_{j.} = \mu + j\omega$ (for design $D_{var_x}$) and $\mu_{.k} = \mu + k\omega$ (for design $D_{var_f}$).
+-->
+
+<!-- The tabulated values and graphical representations of the simulation results are provided in the Supplementary paper.
+The learning from the simulations are as follows: The values of $wpd$ is least for $D_{null}$, followed by -->
+<!-- $D_{var_f}$, $D_{var_x}$ and $D_{var_{all}}$.
+This is a desirable result since the measure $wpd$ was designed such that this relationship holds.
+Furthermore, the distribution of the measure $wpd$ does not change for different facet and x categories.
+The distribution of $wpd$ looks similar with at least the mean and standard of the distributions being uniform across panels.
+This means $wpd$ could be used to measure differences in distribution across panels.
+Also, note that since the data is processed using normal-quantile-transform, this measure is independent of the initial distribution of the underlying data and hence is also comparable across different data sets.
+This is valid for the case when sample size $ntimes$ for each combination of categories is at least 30 and  $nperm$ used for computing $wpd$ is at least 100.
+More detailed results about the properties of $wpd$ could be found in the Supplementary paper.
+-->
+
 ## Ranking and selection of cyclic granularities {#sec:rank-wpd}
+
+A cyclic granularity is referred to as "significant" if there is a significant distributional difference of the measured variable between different categories of the harmony.
+In this section, a selection criterion to choose significant harmonies is provided, thereby eliminating all harmonies that exhibit complete randomness in the measured variable.
+The distance measure $wpd$ is used as a test statistic to test the null hypothesis that no harmony/cyclic granularity is significant.
+We select only those harmonies/cyclic granularities for which the test fails.
+They are then ranked basis how well they capture variation in the measured variable.
+
+<!-- .
+Randomization tests (permutation tests) generates a random distribution by re-ordering the observed data and allows us to test if the observed data is significantly different from any random distribution.
+-->
+<!-- the harmonies are not significant under null hypothesis and then test if our hypotheses are true: -->
+
+<!--   $H_0$: harmonies are not significant -->
+<!--   $H_1$: at least one harmony is significant -->
+
+<!-- we provide a method to select important harmonies by eliminating all harmonies for which patterns are not significant by employing randomization test.
+Randomization tests (permutation tests) generates a random distribution by re-ordering our observed data and allow to test if the observed data is significantly different from any random distribution.
+Complete randomness in the measured variable indicates that the process follows a homogeneous underlying distribution over the whole time series, which essentially implies there is no interesting distinction across any different categories of the cyclic granularities.
+-->
+
 ### Selection
+
+<!-- Under the null hypothesis, the harmonies are not significant. -->
+
+A threshold and consequently a selection criterion is chosen using the notion of Randomization tests (@edgington2007randomization).
+The data is permuted several times and $wpd$ is computed for each of the permuted data sets to obtain the sampling distribution of $wpd$ under the null hypothesis.
+If the null hypothesis is true, then $wpd$ obtained from the original data set would be a likely value in the sampling distribution.
+But in case the null hypothesis is not true, then it is less probable that $wpd$ obtained for the original data will be from the same distribution.
+This idea is utilized to come up with a threshold for selection, denoted by $wpd_{threshold}$, defined as the $99^{th}$ percentile of the sampling distribution.
+A harmony is selected if the value of $wpd$ for that harmony is greater than the chosen threshold.
+The detailed algorithm for choosing a threshold and selection procedure (for two cyclic granularities) is listed as follows:
+
+<!-- The selection criterion is stated as follows: for $(i \in {1, 2, \dots, H_{N_C}})$, select $H_i$ if $wpd_i > wpd_{threshold}$, else reject. -->
+
+- **Input:** All harmonies of the form $\{(A, B), A = \{ a_j: j = 1, 2, \dots, J\}, B = \{ b_k: k = 1, 2, \dots, K\}\}$, $\forall (A, B) \in H_{N_C}$.
+
+- **Output:** Harmony pairs $(A, B)$ for which $wpd$ is significant.
+
+1. Fix harmony pair $(A, B)$.
+
+2. Given the measured variable; $\{v_t: t=0, 1, 2, \dots, T-1\}$, $wpd$ is computed and is represented by $wpd^{A, B}_{obs}$.
+
+3. From the original sequence a random permutation is obtained: $\{v_t^{1}: t=0, 1, 2, \dots, T-1\}$.
+
+4. $wpd$ is computed for the permuted sequence of the data and is represented by $wpd^{A, B}_{1}$.
+
+5.  Steps (3) and (4) are repeated a large number of times $M$ (e.g. $M = 200$).
+
+6. For each permutation, one $wpd^{A, B}_{i}$ is obtained. Define $wpd_{sample} = \{wpd^{A, B}_{1} wpd^{A, B}_{2}, \dots, wpd^{A, B}_{M}\}$.
+
+7. Repeat Steps (1-6) for all harmony pairs $(A, B) \in H_{N_C}$ and stored $wpd^{all}_{sample}$.
+
+8. $99^{th}$ percentiles of $wpd^{all}_{sample}$ is computed and stored in $wpd_{threshold99}$
+
+7. If $wpd^{A, B}_{obs} > wpd_{threshold99}$, harmony pair $(A, B)$ is selected, otherwise rejected.
+
+Similarly, a harmony pair $(A, B)$ could be selected if $wpd^{A, B}_{obs} > wpd_{threshold95}$ and $wpd^{A, B}_{obs} > wpd_{threshold90}$, where  $wpd_{threshold95}$ and $wpd_{threshold90}$ denote the $95^{th}$ and $90^{th}$ percentile of $wpd^{all}_{sample}$ respectively.
+A harmony selected using $99^{th}$,  $95^{th}$  and  $90^{th}$ threshold are tagged as \*\*\*, \*\*, \* respectively.
+
 ### Ranking
+
+The distribution of $wpd$ is expected to be similar for all harmonies under the null hypothesis, since they have been adjusted for different number of categories for the harmonies or underlying distribution of the measured variable.
+Hence, the values of $wpd$ for different harmonies are comparable and can be used to rank the significant harmonies.
+A higher value of $wpd$ for a harmony indicates that higher maximum variation in the measured variable is captured through that harmony.
+Figure \ref{fig:null4by2} presents the results of $wpd$ from the illustrative designs in Section \ref{sec:computation-wpd}.
+The value of $wpd$ under null design (a) is the least, followed by (b), (c) and (d).
+This aligns with the principle of $wpd$, which is expected to have lowest value for null designs and highest for designs of the form $D_{var_{all}}$ (d).
+Moreover, note the relative differences in $wpd$  values between (b) and (c).
+The value of the tuning parameter $\lambda$ is set to $0.67$, which has resulted in giving more emphasis to differences in x-axis categories.
+Again consider \ref{fig:onegran-new}(a) and \ref{fig:onegran-new}(b) with a $wpd$ value of 20.5 and 145 respectively.
+This is because there is more gradual increase across hours of the day than months of the year.
+If order of the categories is not considered, they result in a $wpd$ value of $97.8$ and $161$ respectively, which follows from the fact that if we consider difference between any hours of the day, the magnitude will be much larger than if we consider difference between consecutive categories.
+Similarly, \ref{fig:onegran-new}(a) and (b) have $wpd$ values of   $110.79$ and $125.82$ respectively.
+The ranking implies that the distributional differences are more prominent for the second household, as is also seen from the bigger fluctuations in the $90^{th}$ percentile than the first household.
+
+<!-- ## Choosing a threshold -->
+
+<!-- A randomization test involves calculating a test statistic, randomly shuffling the data and calculating the test statistic several times to obtain a distribution of the test statistic.
+But we will use this procedure to obtain a threshold such that harmony pairs with a $wpd$ value higher than this threshold will only be considered significant.
+The process of choosing a threshold is described as follows: -->
+
+<!-- to test if there is any interesting pattern captured by the harmonies, which essentially implies if $wpd$ is significantly different from zero.
+The percentages of times the $wpd$ obtained from the permuted data is greater than or equal to the observed $wpd$ is the p-value.
+The randomization test is described as follows: -->
+
+<!-- <!-- We can remove the harmonies for which no interesting patterns are observed through a randomization permutation method. --> 
+<!-- Essentially, the assumption is that under the null hypothesis, there is no difference in categories between the pair of cyclic granularities in the chosen harmony.
+This method is based on the generation of randomly chosen reassignments (permutations) of the data across different cyclic granularities and the computation of  -->
+<!-- $wpd$ for each of these reassignments. -->
+
+<!-- **Assumption:** random permutation of the data keeping the categories of the cyclic granularities constant. -->
+
+<!-- Complete randomness in the measured variable indicates that the process follows a homogeneous underlying distribution over the whole time series, which essentially implies there is no interesting distinction across any different categories of the cyclic granularities.
+We can remove the harmonies for which no interesting patterns are observed through a randomization permutation method.
+Essentially, the assumption is that under the null hypothesis, there is no difference in categories between the pair of cyclic granularities in the chosen harmony.
+This method is based on the generation of randomly chosen reassignments (permutations) of the data across different cyclic granularities and the computation of  -->
+<!-- $wpd$ for each of these reassignments.
+The percentages of times the theoretical distribution greater than or equal to the respective observed $wpd$ values are calculated and are used to obtain the P value.
+The procedure for the permutation test is: -->
+
+<!-- **Assumption:** random permutation of the data keeping the categories of the cyclic granularities constant. -->
+
+<!-- 14, 20, 31, 50\}$.
+For each of the $64$ panels, $ntimes = 500$ observations are drawn for each combination of the categories.
+The values of $wpd$ is obtained for each of the panels for the designs $D_{null}$, $D_{var_x}$, $D_{var_f}$ and $D_{var_{all}}$.
+$wpd_{threshold}$ is computed from all of these panels together and the number of times a harmony pair $(A, B) \in H_{N_C}$ is selected when in fact it was of the design $D_{null}$ is noted.
+This entire process is repeated for several null data sets to see the number of times any harmony pair $(A, B) \in H_{N_C}$ is selected under null.
+-->
+
+<!-- ### Results -->
+
+```{r varall-new}
+
+```
+
+```{r varx-new}
+
+```
+
+```{r varf-new}
+```
+
+```{r null-new}
+
+```
+
+```{r plot-all-new3, fig.cap = "$wpd$ values for the four illustrative designs are presented here.
+As expected, the value of $wpd$ under null design is the least (a), followed by (b, c and d).
+Moreover, the relative difference is interesting as c is closer to d than b, since the differences between x categories are up-weighed by design.", dependson= "null4by2"}
+
+```
+
 ## Simulations {#sec:simulations}
+
 ### Behavior of raw and adjusted distance measures
+
+\noindent _Simulation design_
+
+$m=1$
+
+\noindent Observations are generated from a N(0,1) distribution for a wide range of levels from very low to moderately high.
+$nx = \{2, 3, 5, 7, 9, 14, 17, 20, 24, 31, 42, 50\}$ is considered.
+$ntimes = 500$ observations are drawn for each combination of the categories, that is, for a panel with $nx=3$, $500$ observations are simulated for each of the categories.
+This design corresponds to $D_{null}$ as each combination of categories in a panel are drawn from the same distribution.
+Furthermore, the data is simulated for each of the categories $nsim=200$ times, so that the distribution of $wpd$ under $D_{null}$ could be observed.
+The values of $wpd$ is obtained for each of the panels.
+$wpd_{l, s}$ denotes the value of $wpd$ obtained for the $l^{th}$ panel and $s^{th}$ simulation.
+
+$m=2$  
+\noindent Similarly, observations are generated from a N(0,1) distribution for each combination of $nx$ and $nfacet$ from the following sets: $nx = nfacet = \{2, 3, 5, 7, 14, 20, 31, 50\}$.
+That is, data is being generated for each of the panels $(2, 2), (2, 3), (2, 5) \dots, (50, 31), (50, 50)$.
+For each of the $64$ panels, $ntimes = 500$ observations are drawn for each combination of the categories.
+That is, if we consider a $(2, 2)$ panel, $500$ observations are generated for each of the possible subsets, namely, $\{(1, 1), (1, 2), (2, 1), (2, 2)\}$.
+
+\noindent _Results_
+
+Figure \ref{fig:raw} shows that both the location and scale of the distributions change across panels.
+This is not desirable under $D_{null}$ as it would mean comparisons of $wpd$ values is not appropriate across different $nx$ and $nfacet$.
+Table \ref{tab:glm-tab} gives the summary of the generalized linear model to capture the relationship between $wpd_{raw}$ and number of comparisons.
+For example, the model considered for $m = 2$ is $wpd_{l,s} = 23.4 - 0.96*log(nx * nfacet)) + e$.
+The intercepts and slopes are similar independent of the starting distributions (see supplementary paper for details) and hence the coefficients are shown for the case when observations are drawn from a N(0, 1) distribution.
+Figure \ref{fig:dist-new-same-scale-link} shows the distribution of $wpd_{perm}$ and $wpd_{glm-scaled}$ in the same scale to show that a combination approach could be used for higher values of levels to alleviate the computational time of permutation approach.
+
+```{r raw, fig.cap = "Distribution of $wpd_{raw}$ is plotted across different $nx$ and $nfacet$ categories under $D_{null}$ through density and rug plots.
+Both location (blue line) and scale (orange marks) of the distribution shifts for different panels.
+This is not desirable since under null design, the distribution  is not expected to capture any differences."}
+```
+
+```{r glm-tab}
+
+```
+
+```{r wpd-glm-dist}
+
+```
+
+```{r hist-qq-new, eval = FALSE}
+
+```
+
+```{r quadratic, eval=FALSE}
+```
+
+```{r norm, eval}
+```
+
+```{r dist-new-same-scale-link, fig.cap = "The distribution of $wpd_{perm}$ and $wpd_{glm-scaled}$ are overlaid to compare the location and scale across different $nx$ and $nfacet$.
+$wpd_{norm}$ takes the value of $wpd_{perm}$ for lower levels, and $wpd_{glm-scaled}$ for higher levels to to alleviate the problem of computational time in permutation approaches.
+This is possible as the distribution of the adjusted measure looks similar for both appraches for higher levels."}
+```
+
 ### Choosing threshold
+
+\noindent _Simulation design_
+
+Observations are generated from a N(0,1) distribution for each combination of $nx$ and $nfacet$ from the following sets: $nx = \{3, 7, 14\}$ and $nfacet = \{2, 9, 10\}$.
+This would result in 9 panels, viz, $(3, 2), (3, 9), (3, 10), \dots, (14,9), (14, 10)$.
+Few experiments were conducted.
+In the first scenario, data for all panels are simulated using the null design  $D_{null}$.
+In other scenarios, data simulated from the panel $(14, 2)$ and $(3, 10)$ are considered under $D_{vary_{all}}$.
+Moreover, $\omega = \{0.5, 2, 5\}$ are considered to examine if the proposed test is able to capture subtle differences and non-subtle differences when we shift from the null design.
+
+<!-- In the last scenario, we consider the panel $(3, 2), (7,9), (14, 10)$ to be under $D_{null}$, the panels $(7, 2), (14, 9)$ to be under $D_{var_f}$.
+$(14, 2), (3, 10)$ under $D_{var_x}$ and the rest under $D_{var_{null}}$.
+This is done to check if the consequent ranking procedure leads to designs like $D_{vary_{all}}$ to be chosen first followed by $D_{vary_{all}}$.
+We generate only one data set each for which these scenarios were simulated and consider this as the original data set.
+We generate $1000$ repetitions of this experiment with different seeds.
+-->
+
+\noindent _Results_
+
+For the first scenario, size of the test is obtained as $0.1$ with $wpd_{threshold99}$ as the threshold.
+This implies that the proportion of times a panel is rejected when it is under $D_{null}$ is $0.1$.
+The level of significance for each test is 1% (as a result of choosing $99^{th}$ percentile as the threshold) and we have $9$ multiple tests.
+Hence, it is reasonable
+that the level of significance of the composite tests would be larger than the individual tests.
+We also compute the proportion of times a panel is rejected when it actually belongs to a non-null design.
+These proportions constitute to be the estimated size and power of the test.
+Power can depend on many things like sample size, number of designs that deviate from the null and extent to which from the null.
+It is found that as we increase from low to high changes from the null distribution, the power increased.
+<!---The power also increases with increased sample size, as is expected.---> The results and graphics are included in details in the Supplementary paper.
+
+```{r, eval = FALSE}
+all_data <- read_rds("simulations/supplementary/test-hpc/data-agg/all_data.rds")
+
+# computes size for 99 percentile threshold
+data_summary <- all_data %>%
+  mutate(sig = if_else(significance!=99, 0, 99)) %>% 
+  group_by(seed_id) %>% 
+  summarize(sumsig = sum(sig)) %>% 
+  count(sumsig==0) %>% 
+  mutate(p_value = n/sum(n)) %>% 
+  slice(1)
+
+# plots
+ggplot() + 
+  geom_histogram(data = all_data, aes(x = wpd))  + 
+  geom_vline(data = data_summary, aes(xintercept = p_value),colour = "red") +
+  facet_grid(x_levels~facet_levels)
+
+```
+
 ### Environment
+
+Simulation studies were carried out to study the behavior of $wpd$, build the normalization method as well as compare and evaluate different normalization approaches.
+R version 4.0.1 (2020-06-06) is used with the platform: x86_64-apple-darwin17.0 (64-bit) running under: macOS Mojave 10.14.6 and MonaRCH, which is a next-generation High Power Computing (HPC) Cluster, addressing the needs of the Monash HPC community.
+
 ## Application to residential smart meter dataset {#sec:application-wpd}
+
+The smart meter data set for eight households in Melbourne has been utilized to see the use of $wpd$ proposed in the paper.
+The data has been cleaned to be a `tsibble` (@wang2020tsibble) containing half-hourly electricity consumption from Jul-2019 to Dec-2019 for each of the households, which is procured by them by downloading their data from the energy supplier/retailer.
+No behavioral pattern is likely to be discerned from the line graph of energy usage over the entire period, since the plot will have too many measurements squeezed in a linear representation.
+When we zoom into the linear representation of this series in Figure \ref{fig:dotplot-8} (b) for September, some patterns are visible in terms of peaks and troughs, but we do not know if they are regular or what is their period.
+Electricity demand, in general, has a daily and weekly periodic pattern.
+However, it is not apparent from this view if all of these households have those patterns and in case they have if they are significant enough.
+Also, it is not clear if any other periodic patterns are present in any household which might have been hidden with this view.
+We start the analysis by choosing few harmonies, ranking them for each of these households, compare households to get more insights into what these rankings imply.
+Furthermore, the ranking and selection of significant harmonies is validated by analyzing the distribution of energy usage across significant harmonies.
+
+<!-- if we are benefiting by moving from a linear to cyclic representation of time to . -->
+
+```{r linear-scale-8, fig.cap = "Electricity demand for eight households is shown in different facets from Jul-19 to Dec-19 in Fig a and it has been zoomed in for Sep-19 in Fig b, where a week in Sep-19 has been highlighted.
+From the scales of Fig a, it is apparent that these households vary in consumption levels, but all their periodic patterns have been squeezed with this representation.
+In fig b, we can see some weekly patterns for the entire period and daily patterns for the highlighted week and there could be some other patterns visible in terms of peaks and troughs, but it is not clear if they are regular or significant.", fig.height=8, eval = FALSE}
+
+```
+
+_Choosing cyclic granularities of interest and removing clashes_
+
+Let $v_{i, t}$ denote the electricity demand for $i^{th}$ household for time period $t$.
+The series $v_{i, t}$ is the linear granularity corresponding to half-hour since the interval of the tsibble is 30 minutes.
+We consider coarser linear granularities like hour, day, week and month from the commonly used Gregorian calendar.
+Considering $4$ linear granularities hour, day, week, month, the number of cyclic granularities is $N_C = (4 * 3/2) = 6$.
+We obtain cyclic granularities namely "hour_day", "hour_week", "hour_month", "day_ week", "day_month" and "week_month", read as "hour of the day", etc.
+Further, we add cyclic granularity day-type ("wknd wday") to capture weekend and weekday behavior.
+Thus, $7$ cyclic granularities are considered to be of interest.
+The set consisting of pairs of cyclic granularities ($C_{N_C}$) will have $7_{P_2}=42$ elements which could be analyzed for detecting possible periodicities.
+The set of possible harmonies $H_{N_C}$ from $C_{N_C}$ are chosen by removing clashes using procedures described in [@Gupta2021-hd].
+Table \ref{tab:tab-rank-8}  shows $14$ harmony pairs that belong to $H_{N_C}$.
+
+ <!-- The paper also shows how the number of harmonies could be identified among  by removing clashes.
+This section shows how we could refine the search further by only looking at significant harmonies.
+-->
+
+_Selecting and Ranking harmonies for all households_
+
+$wpd_{i}$ is computed on $v_{i, t}$ for all harmony pairs $\in H_{N_C}$ and for each households $i \in \{1, 2, \dots, 8\}$.
+The harmony pairs are then arranged in descending order and highlighted with `***`, `**` and `*` corresponding to the  $99^{th}$,  $95^{th}$ and $90^{th}$ percentile threshold.
+Table \ref{tab:tab-rank-8} shows the rank of the harmonies for different households.
+The rankings are different for different households, which is a reflection of their varied behaviors.
+Most importantly, there are at most $3$ harmonies that are significant for any household.
+This is a huge reduction in the number of potential harmonies to explore closely, starting from $42$.
+
+_Detecting patterns not apparent from linear display_
+
+\noindent Figure \ref{fig:dotplot-8} helps to compare households through the heatmap (a) across harmony pairs with the cyclic granularity mapped to x-axis and facet being plotted on the x-axis and y-axis of the heatmap.
+Here $dom$, $dow$, $wdwnd$ are abbreviations for day-of-month, day-of-week, weekday/weekend and so on.
+The colors represent the value of $wpd$.
+Darker cells correspond to more significant harmony pairs.
+Also, the ones with `*` corresponds to the ones above $wpd_{threshold95}$.
+Few observations that emphasizes patterns not discernible through (b) includes - (1) id 7 and 8 have the same significant harmonies despite having very different total energy usage, (2) id 6 and 7 differ in the sense that for id 6, the difference in patterns only during weekday/weekends, whereas for id 7, all or few other days of the week are also important.
+This might be due to their flexible work routines or different day-off, (3) there are no significant periodic patterns for id 5 when we fix the threshold to $wpd_{threshold95}$.
+Note that the $wpd$ values are computed over the entire range, but the linear display in (b) is zoomed only for September, with the major and minor x-axis corresponding to weeks and days respectively.
+
+<!-- Households id 2 and 3 are similar, in terms of linear display and heatmap, which implies that similar periodic behavior can stem from households with different demographics. -->
+
+```{r rank-household-mclapply-8}
+```
+
+```{r elec_select_harmony-8}
+```
+
+```{r dotplot-8, fig.cap = "An ensemble plot with a heatmap (a), line plot (b), parallel coordinate plot (c) to demonstrate energy behavior of the households in different ways.
+(b) is the zoomed-in raw demand series for September to highlight the repetitive patterns of energy demand.
+(a) shows $wpd$ values across harmonies with the x variable of harmony placed across x-axis and facet variable placed across y-axis.
+The darker the colour in (a), the higher the harmony is ranked.
+Visualizing harmonies through (a) allows to view the significant cyclic granularities along the x-axis, facet or both for each household.
+For eg, ids 7 and 8 have significant patterns across (hod, dow) and (dow, hod), which was not apparent through (a).
+(c) is useful for comparing households across harmonies, for eg, for the harmony (dow-hod), ids 1 and 7 have the least and highest wpd respectively.", fig.height = 7}
+```
+
+```{r tab-rank-8}
+```
+
+_Comparing households and validating rank of harmonies_
+
+According to Figure \ref{fig:dotplot-8}(c), for the harmony pair (dow-hod), household id 7 has the greatest value of $wpd$, while id 1 has the least.
+Also, from table \ref{tab:tab-rank-8} it could be seen that the harmony pair $(dow, hod)$ is important for id 7, however it has been labeled as an inconsequential pair for id 1.
+The distribution of energy demand for both of these households, with $dow$ as the facet and $hod$ on the x-axis, may help explain the choice.
+Figure \ref{fig:gravitas-plot-8} demonstrates that for id 7, the median (black) and quartile deviation (orange) of energy consumption fluctuates throughout for most hours of the day and day of the week, while for id 1, daily patterns are consistent within weekdays and weekends.
+As a result, for id 1, it is more appropriate to examine the distributional difference solely across (dow, wdwnd), which has been rated higher in Table \ref{tab:tab-rank-8}.
+
+<!-- Although the differences might seem significant at first, with closer inspection it could be seen that the scale of the demand is lower in this case and hence the differences are not large enough to cross the threshold for significance.
+-->
+
+<!-- Figure \ref{fig:gravitas-plot-8} is used to show -->
+<!-- if this selection and ranking of harmony pairs makes sense for this household.
+Panel a) of  Figure \ref{fig:gravitas-plot-8} shows the distribution of energy demand with weekday/weekend as the x-axis and hour-of-day as the facets and helps to compare the weekend/weekday patterns for different hours of the day.
+It could be observed that the difference between weekend and weekday is the highest from 15 to 19 hours of the day.
+Panel b) shows the distribution of energy demand with the variables swapped and helps to compare the daily patterns within weekday and weekend.
+It could be observed that the daily pattern is similar for weekdays and weekends with a morning and evening peak.
+However, the difference between morning and evening peaks are higher for weekends.
+Since $wpd$ is designed to put more weightage on within-facet differences for $\lambda>0.5$, it makes sense that the pair (hod, wdwnd) has been ranked higher than (wdwnd, hod).
+Panel c) shows the distribution of energy demand with weekday/weekend as the x-axis and week-of-month as the facets.
+Although the differences might seem significant at first, with closer inspection it could be seen that the scale of the demand is lower in this case and hence the differences are not large enough to cross the threshold for significance.
+-->
+
+```{r gravitas-plot-8, message = FALSE, warning=FALSE,fig.cap="Comparing distribution of energy demand shown for household id 1 (a) and 7 (b) across hod in x-axis and dow in facets through quantile area plots.
+The value of $wpd$ in Table 3 suggests that the harmony pair (dow, hod) is significant for household id 7, but not for 1.
+This implies that distributional differences are captured more by this harmony for id 7, which is apparent from this display with more fluctuations across median and 75th percentile for different hours of the day and day of week.
+For id 1, patterns look similar within weekdays and weekends.
+Here, the median is represented by the black line, the orange area corresponds to quartile deviation and the green area corresponds to area between $10^{th}$ and $90^{th}$ quantile."}
+
+```
+
+<!-- From Figure \ref{fig:gravitas-plot-8-id5}, it could further be observed that id5 has only one significant harmony $(hod, dow)$.
+Apparently, $(hod, wnd/wday)$ which is an important harmony for most households is not important for this one.
+-->
+
 ## Discussion
+
+Exploratory data analysis involves many iterations of finding and summarizing patterns.
+With temporal data available at finer scales, exploring time series has become overwhelming with so many possible granularities to explore.
+A common solution is to aggregate and look at the patterns across usual granularities like hour-of-day or day-of-week, but there is no way to know the “interesting” granularities a priori.
+A huge number of displays need to be analyzed or we might end up missing informative granularities.
+This work refines the search of informative granularities by identifying those for which the differences between the displayed distributions are greatest and rating them in order of importance of capturing maximum variation.
+
+The significant granularities across different datasets (individuals/subjects) do not imply similar patterns across different datasets.
+They simply mean that maximum distributional differences are being captured across those granularities.
+A future direction of work is to be able to explore and compare many individuals/subjects together for similar patterns across significant granularities.
+
+<!-- patterns are interesting and others are not.
+The subjects for which behaviors across a group of cyclic granularities are similar, could be grouped together as their periodic behavior is similar.
+to group subjects  -->
+
 ## Acknowledgments {-}
+
+The Australian authors thank the ARC Centre of Excellence for Mathematical and Statistical Frontiers [(ACEMS)](https://acems.org.au/home) for supporting this research.
+Sayani Gupta was partially funded by [Data61 CSIRO](https://data61.csiro.au/) during her PhD.
+The Github repository,  [github.com/Sayani07/paper-hakear](https://github.com/Sayani07/paper-gravitas), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
+This article was created with R [@R-language], `knitr` [@knitr2015, @R-knitr] and `rmarkdown` [@rmarkdown2018, @R-rmarkdown].
+Graphics are produced with @Wickham2009pk.
+
 ## Supplementary Materials
+
+**Data and scripts:** Data sets and R code to reproduce all figures in this article (main.R).
+
+\noindent
+**Simulation results and scripts:** All simulation table, graphics and and R code to reproduce it (paper-supplementary.pdf, paper-supplementary.Rmd)
+
+\noindent
+**R-package:** The open-source R [@R-language] package `hakear` is available on Github (https://github.com/Sayani07/hakear) to implement ideas presented in this paper.
 
@@
 
-Placeholder
+With more and more time series data being collected at much finer temporal resolution, for a longer length of time, and for a larger number of individuals/entities, time series clustering research is getting a lot of traction.
+Long, noisy, patchy, uneven, and asynchronous time series are common in many fields, limiting similarity searches or lowering method efficiency when clustering is based on a distance metric.
+In this work, we suggest two approaches for obtaining similarity between time series based on probability distributions over cyclic temporal granularities for distance-based clustering approaches.
+Cyclic granularities like hour-of-the-day, work-day/weekend, month-of-the-year and so on, are useful for finding repeated patterns in the data.
+Looking at probability distributions across cyclic granularities serves two purposes: (a) "Probability distributions" characterise the inherent temporal data structure of these large unequal-length time series and are robust to missing or noisy data.
+(b) Using probability distributions over "cyclic granularities" ensures small pockets of similar "repeated" behaviors.
+Our method is capable of producing useful clusters, as demonstrated by testing on validation data designs and a sample of residential smart meter consumers.
+
+```{r external3, include = FALSE, cache = FALSE}
+knitr::read_chunk('scripts/gracsr.R')
+```
+
+```{r load-lib}
+
+```
+
+```{r mytheme-application}
+```
+
+```{r theme_characterisation}
+```
+
+```{r theme-validation}
+
+```
 
 ## Introduction
+
+<!-- time series clustering and its challenges -->
+Time-series clustering is the process of unsupervised partitioning of $n$ time-series data into $k$ ($k<n$) meaningful groups such that homogeneous time-series are grouped together based on a certain similarity measure.
+The time-series features, length of time-series, representation technique, and, of course, the purpose of clustering time-series all influence the suitable similarity measure or distance metric to a meaningful level.
+The three primary methods to time series clustering (@liao2005clustering) are algorithms that operate directly with distances or raw data points in the time or frequency domain (distance-based), with features derived from raw data (feature-based), or indirectly with models constructed from raw data (model-based).
+The efficacy of distance-based techniques is highly dependent on the distance measure utilized.
+Defining an appropriate distance measure for the raw time series may be a difficult task since it must take into account noise, variable lengths of time series, asynchronous time series, different scales, and missing data.
+Commonly used Distance-based similarity measures as suggested by a decade review of time series clustering approaches (@Aghabozorgi2015-ct) are Euclidean, Pearson's correlation coefficient and related distances, Dynamic Time Warping, Autocorrelation, Short time series distance, Piecewise regularization, cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances (@liao2007clustering) but on a vector time series data.
+Among these alternatives, Euclidean distances have high performance but need the same length of data over the same period, resulting in information loss regardless of whether it is on raw data or a smaller collection of features.
+DTW works well with time series of different lengths (@corradini2001dynamic), but it is incapable of handling missing observations.
+Surprisingly, probability distributions, which may reflect the inherent temporal structure of a time series have not been considered in determining time series similarity.
+
+\noindent We consider the problem of clustering a large number of univariate time series of continuous values which are available at fine temporal scales, being motivated by the residential smart meter data.
+These time series data are long (with more and more data collected at finer resolutions), are asynchronous, with varying time lengths for different houses and missing observations and characterized by noisy and patchy behavior that can quickly become overwhelming and hard to interpret, requiring summarizing the large number of customers into pockets of similar energy behavior.
+Choosing probability distributions seem to be a natural way to analyze these types of data sets since they are robust to uneven length, missing data, or noise.
+This paper proposes two approaches for obtaining pairwise similarities based on Jensen-Shannon distances between probability distributions across significant cyclic granularities.
+Cyclic temporal granularities [@Gupta2021-hakear], which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can be useful for measuring repetitive patterns in large univariate time series data.
+The resulting clusters are expected to group customers that have similar repetitive behaviors across each of the interesting cyclic granularities.
+Below are some of the benefits of our method:
+
+- When using probability distributions, data does not have to be the same length or observed during the exact same time period (unless there is a structural pattern);
+
+<!-- (for example, some customers observed only for holidays, others observed over non holidays or year of observation is different which can lead to technological advancement.) -->
+
+- Jensen-Shannon distances evaluate the distance between two distributions rather than raw data, which is less sensitive to missing observations and outliers than other conventional distance methods;
+
+- While most clustering algorithms produce clusters similar across just one temporal granularity, this technique takes a broader approach to the problem, attempting to group observations with similar distributions across all interesting cyclic granularities;
+
+- It is reasonable to define a time series based on its degree of trend and seasonality, and to take these characteristics into account while clustering it.
+The modification of the data structure by taking into account probability distributions across cyclic granularities assures that there is no trend and that seasonal variations are handled independently.
+As a result, there is no need to de-trend or de-seasonalize the data before applying the clustering method.
+For similar reasons, there is no need to exclude holiday or weekend routines.
+
+_Background and motivation_
+
+Large spatio-temporal data sets, both from open and administrative sources, offer up a world of possibilities for research.
+One such data sets for Australia is the Smart Grid, Smart City (SGSC) project (2010–2014) available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) The project provides half-hourly data of over 13,000 household electricity smart meters distributed unevenly from October 2011 to March 2014.
+Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.
+Households vary in size, location, and amenities such as solar panels, central heating, and air conditioning.
+The behavioral patterns differ amongst customers due to many temporal dependencies.
+Some households use a dryer, while others dry their clothes on a line.
+Their weekly profile may reflect this.
+They may vary monthly, with some customers using more air conditioners or heaters than others, while having equivalent electrical equipment and weather circumstances.
+Some customers are night owls, while others are morning larks.
+Day-off energy use varies depending on whether customers stay home or go outside.
+Age, lifestyle, family composition, building attributes, weather, availability of diverse electrical equipment, among other factors, make the task of properly segmenting customers into comparable energy behavior a fascinating one.
+This challenge is worsened when all we know about our consumers is their energy use history (@Ushakova2020-rl).
+To safeguard the customers' privacy, it is probable that such information is not accessible.
+Also, energy suppliers may not always update client information, such as property features, in a timely manner.
+Thus, there is a growing need to have research that examines how much energy usage heterogeneity can be found in smart meter data and what are some of the most common power consumption patterns, rather than explaining why consumption differs.
+
+_Related work_
+
+A multitude of papers have emerged around smart meter time series clustering for deepening our knowledge of consumption patterns.
+@Tureczek2017-pb conducted a systematic study of over $2100$ peer-reviewed papers on smart meter data analytics.
+None of the $34$ articles chosen for their emphasis use Australian smart meter data.
+The most often used algorithm is K-Means.
+Using K-Means without considering time series structure or correlation results in inefficient clusters.
+Principal Component Analysis (PCA) or Self-Organizing Maps (SOM) eliminate correlation patterns and decrease feature space, but lose interpretability.
+To reduce dimensionality, several studies use principal component analysis or factor analysis to pre-process smart-meter data before clustering (@Ndiaye2011-pf).
+Other algorithms utilized in the literature include k-means variants, hierarchical approaches, and greedy k-medoids.
+Time series data, such as smart meter data, are not well-suited to any of the techniques mentioned in @Tureczek2017-pb.
+Only one study [@ozawa2016determining] identified time series characteristics using Fourier transformation, which converts data from time to frequency and then uses K-Means to cluster by greatest frequency.
+@Motlagh2019-yj suggests that the time feature extraction is limited by the type of noisy, patchy, and unequal time-series common in residential and addresses model-based clustering by transforming the series into other other objects such as structure or set of parameters which can be more easily characterized and clustered.
+[@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
+@Melnykov2013-sp discusses how outliers, noisy observations and scattered observations can complicate estimating mixture model parameters and hence the partitions.
+To our knowledge, none of the methods focus on exploring heterogeneity in repetitive behaviors based on the dynamics of multiple temporal dependencies using probability distributions.
+
+<!-- Given none of the methods use probability distributions across cyclic granularity to explore heterogeneity in different repetitive behaviors, we present a way to do that by presenting similarity through probability distributions across cyclic granularities.
+-->
+
+The remainder of the paper is organized as follows: Section&nbsp;\ref{sec:methodology} provides the clustering methodology.
+Section&nbsp;\ref{sec:validation} shows data designs to validate our methods.
+Section&nbsp;\ref{sec:cases} discusses the application of the method to a subset of the real data.
+Finally, we summarize our results and discuss possible future directions in Section&nbsp;\ref{sec:conclusion}.
+
+<!-- A typical clustering technique includes the following steps: (a) establishing distance (dissimilarity) and similarity through feature or model extraction and selection; and (b) selecting the clustering algorithm design.
+Distance measures could be time-domain based or frequency-domain (fast Fourier transform).
+-->
+
+<!-- A model-based clustering works by transforming the series into other other objects such as structure or set of parameters which can be more easily characterized and clustered (@Motlagh2019-yj).
+[@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
+The essential temporal characteristics of the curves are defined or extracted using feature-based clustering.
+-->
+
+<!-- This work -->
+<!-- This is similar to a stochastic approach (@Motlagh2019-yj) to clustering, which proposes interpreting electricity demand as a random process and extracting time-series characteristics, or a model of the series, to enable unsupervised clustering.
+Unsupervised clustering is only as good as the features that are extracted/selected or the distance metrics that were utilized.
+Well-designed additional features may collect characteristics that default features cannot.
+Based on the underlying structure of the temporal data, this article offers new distance metric and features for clustering and applies them to actual smart-meter data.
+Firstly, the distance metric is based on probability distribution, which in our knowledge is the first attempt to cluster smart meter data using probability distributions.
+These recorded time series are asynchronous, with varying time lengths for different houses and missing observations.
+Taking probability distributions helps to deal with such data, while helping with dimension reduction in one hand but not losing too much information due to aggregation.
+Secondly, we recognise that most clustering algorithms only provide hourly energy profiles during the day, but this approach provides a wider approach to the issue, seeking to group consumers with similar shapes over all important cyclic granularities.
+Since cyclic granularities are considered instead of linear granularities, clustering would group customers that have similar repetitive behavior across more than one cyclic granularities across which patterns are expected to be significant.
+-->
+
+<!-- common similarity measures -->
+
+<!-- electricity data structure -->
+
+<!-- common similarity measures used there -->
+
+<!-- lit review -->
+
 ## Clustering methodology {#sec:methodology}
+
+<!-- In contrast to models, a feature-based strategy is used to explicitly define or automatically extract the curves’ key time features, for instance by application of PCA on the daily curves [ --
+
+<!-- Most papers discussed in Tureczek2017-pb fail to accept smart meter readings as time series data, a data type which contains a temporal component.
+The omission of the essential time series features in the analysis leads to the application of methods that are not designed for handling temporal components.
+K-Means ignores autocorrelation, unless the input data is pre-processed.
+The clusters identified in the papers are validated by a variety of indices, with the most prevalent -->
+<!-- being the cluster dispersion index (CDI) [22–24], the Davies–Bouldin index (DBI) [25,26] and the mean index adequacy (MIA) [8,13]. -->
+
+<!-- The data set solely contains readings from smart meters and no information about the consumers' specific physical, geographical, or behavioral attributes.
+As a result, no attempt is made to explain why consumption varies.
+Instead, this work investigates how much energy usage heterogeneity can be found in smart meter data and what some of the most common electricity use patterns are.
+-->
+
+<!-- Because of this importance, countless approaches to estimate time series similarity have been proposed.  -->
+
+The foundation of our method is unsupervised clustering algorithms based exclusively on the time-series data.
+The proposed methodology aims to leverage the intrinsic data structure hidden within cyclic temporal granularities.
+The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
+<!--So going back to the smart meter example, $f_i(v)$ is the distribution of customer $i$ and $v$ is electricity demand.--> In this work, instead of considering the probability distributions of the linear time series, we assume it across different categories of any cyclic granularity.
+We can consider categories of an individual cyclic granularity ($A$) or combination of categories for two interacting granularities ($A*B$) to have a distribution, where $A, B$ are defined as $A = \{a_j: j=1, 2, \dots J\}$ and $B = \{b_k: k = 1, 2, \dots K\}$.
+For example, let us consider two cyclic granularities $A$ and $B$ representing hour-of-day and day-of-week.
+Then $A = \{0, 1, 2, \dots, 23\}$ and $B = \{Mon, Tue, Wed, \dots, Sun\}$.
+In case individual granularities are considered, there are $J = 24$ distributions of the form $f_{i,j}(v)$ or $K = 7$ distributions of the form $f_{i,k}(v)$ for each customer $i$.
+In case of interaction, $J*K=168$ distributions of the form $f_{i,j, k}(v)$ could be conceived for each customer $i$.
+Hence clustering these customers is equivalent to clustering these collections of conditional univariate probability distributions.
+Towards this goal, we need to decide how to measure similarities between collections of univariate probability distributions.
+There are multiple ways to measure similarities depending on the aim of the analysis.
+<!--This paper focuses on looking at the (dis) similarity between underlying distributions that may have resulted in different patterns across different cyclic temporal granularities, that eventually have resulted in the (dis) similarity between time series.
+--> This paper considers a two approaches for finding distances between time series.
+Both of these approaches may be useful in a practical context and, depending on the data set, may or may not propose the same customer classification.
+The obtained distances could be fed into a clustering algorithm to break large data sets into subgroups that can then be analyzed separately.
+<!--These clusters may be commonly associated with real-world data segmentation.
+However, since the data is unlabeled a priori, more information is required to corroborate this.--> The methodology is explained in the Figure \ref{fig:flowchart} and each element of the pipeline is discussed.
+
+```{r flowchart, fig.cap = "Flow chart illustrating the pipeline for methodology", out.width="100%"}
+knitr::include_graphics("img/flowchart.png")
+
+```
+
+- _Find significant granularities or harmonies_
+
+[@Gupta2021-hakear] proposes a method for choosing significant cyclic granularities and harmonies (interacting granularities that can be studied together) [@Gupta2021-gravitas], which is used in this work.
+We define "significant" granularities as those with significant distributional differences across categories.
+It is preferable to consider those since it is more probable that there will be some intriguing repeated behavior worth investigating.
+It should be noted that all of the observations in the study may not have the same set of important granularities.
+The following is a method for generating a list ($S_c$) of significant granularities for all observations:
+
+ (a) Remove granularities from the comprehensive list that are insignificant for all observations.
+
+ (b) select only the granularities that are significant for the majority of observations.
+
+There will be observations in both cases where one or a few selected granularities are boring.
+Even in that case, having this group of observations with no interesting patterns at a granularity that regularly discovers patterns may be valuable.
+If, on the other hand, the granularities under examination are truly important for a group of data, unique patterns may be identified while clustering them.
+
+- _Data transformation_
+
+Observations often have a somewhat skewed time series distribution and their ranges might vary greatly.
+Statistical transformations are employed to bring all of them to the same range or normalize each observation.
+This is important because we are not interested in trivial clusters that vary in magnitudes, but rather in uncovering comparable patterns of distributional differences between categories.
+For the JS-based approaches, two data transformation techniques are utilized viz, Normal-Quantile Transform (NQT) and Robust scaling.
+NQT is a built-in transformation for computing $wpd$, which is the foundation of wpd-based distances.
+
+_Robust scaling_ The normalized $i^{th}$ observation is denoted by $v_{norm} = \frac{v_t - q_{0.5}}{q_{0.75}-q_{0.25}}$, where $v_t$ is the actual value at the $t^{th}$ time point and $q_{0.25}$, $q_{0.5}$ and $q_{0.75}$ are the $25^{th}$, $50^{th}$ and $75^{th}$ percentile of the time series for the $i^{th}$ observation.
+$v_{norm}$ has zero mean and median, as well as a standard deviation of one, with the outliers having same relative connections to other values.
+
+_Normal-Quantile transform_ The raw data for all observations is individually normal-quantile transformed (NQT) [@Krzysztofowicz1997-bv], so that the transformed data follows a standard normal distribution.
+NQT will make the skewed distributions bell-shaped.
+As a result, determining which raw distribution was used is difficult using the modified distribution.
+Multimodality is also concealed or reversed.
+As a result, while displaying transformed data using NQT, one must exercise caution.
+But NQT is a useful transformation that often improves clustering performance.
+
+<!-- NQT, however, is an useful transformation which often improves the clustering performance.
+Since our variables are cyclic granularities which are derived from linear granularities, NQT ensures that the conditional distribution across each variable is also normalised.
+-->
+
+<!-- is not a problem for implementation of this methodology as distributions are characterized by quantiles and the order of the quantiles is reserved under NQT. -->
+
+<!-- First, the original data is ranked in ascending order and the probabilities $P(Y<=y(i)) = i/(n+1)$ are attached to $y(i)$, in terms of their ranking order.
+A NQT based transformation is applied by computing from a standard normal distribution a variable $\eta(i)$, which corresponds to the same probability $P(\eta< \eta(i)) = i/n+1$.By doing this, the new variables $\eta(i)$ will be marginally distributed according to standard Normal, N(0,1).
+-->
+
+<!-- It is worth noting that when studying these similarities, a variety of objectives may be pursued.
+One objective could be to group time series with similar shapes over all relevant cyclic granularities.
+In this scenario, the variation in customers within each group is in magnitude rather than shape, while the variation between groups is only in shape.
+There -->
+<!-- are distance measures are used for shape-based clustering [Ding et al. 2008; Wang -->
+<!-- et al.
+2013] and many more but none of them look at the probability distributions while computing similarity.
+Moreover, most distance measures offer similar shape across just one dimension.
+For example, we often see "similar" daily energy profiles across hours of the day, but we suggest a broader approach to the problem, aiming to group consumers with similar distributional shape across all significant cyclic granularities.
+Another purpose of clustering could be to group customers that have similar differences in patterns across all major cyclic granularities, capturing similar jumps across categories regardless of the overall shape.
+For example, in the first goal, similar shapes across hours of the day will be grouped together, resulting in customers with similar behavior across all hours of the day, whereas in the second goal, any similar big-enough jumps across hours of the day will be clubbed together, regardless of which hour of the day it is.
+Both of these objectives may be useful in a practical context and, depending on the data set, may or may not propose the same customer classification.
+Depending on the goal of clustering, the distance metric for defining similarity would be different.
+These distance metrics could be fed into a clustering algorithm to break large data sets into subgroups that can then be analyzed separately.
+These clusters may be commonly associated with real-world data segmentation.
+However, since the data is unlabeled a priori, more information is required to corroborate this.
+This section presents the work flow of the methodology:  -->
+
+- _Data pre-preprocessing_
+
+We start with a "tsibble" (@wang2020tsibble) data structure with an index variable representing inherent ordering from past to present and a key variable that defines observational units over time.
+The measured variable for an observation is a time-indexed sequence of values.
+This sequence, however, could be shown in several ways.
+A shuffle of the raw sequence may represent hourly consumption throughout a day, a week, or a year.
+Cyclic granularities can be expressed in terms of the index set in the "tsibble" data structure.
+But the data structure changes while transporting from linear to cyclic scale of time as multiple observations now correspond to each category and induce a probability distribution.
+Directly computing Jensen-Shannon distances between the entire probability distributions can be very costly.
+Hence, in this paper, quantiles are chosen to characterize the probability distributions.
+So, in the final data structure, each category of a cyclic granularity corresponds to a list of numbers which is essentially a few chosen quantiles.
+
+- _Distance metrics_
+
+<!-- Considering each individual or combined categories of cyclic granularities to have an underlying distribution lead to a collection of conditional distributions for each customer $i$. -->
+
+The total (dis) similarity between each pair of customers is obtained by combining the distances between the collections of conditional distributions.
+This needs to be done in a way such that the resulting metric is a distance metric, and could be fed into the clustering algorithm.
+Two types of distance metric is considered:
+
+<!-- The choice of distance measures is a critical step in clustering. It defines how the similarity of two elements (x, y) is calculated. -->
+
+**JS-based distances**
+
+This distance metric considers two time series to be similar if the distributions of each category of an individual cyclic granularity or combination of categories for interacting cyclic granularities are similar.
+In this study, the distribution for each category is characterized using deciles (can potentially consider any list of quantiles), and the distances between distributions are calculated using the Jensen-Shannon distances (@Menendez1997-in), which are symmetric and thus could be used as a distance measure.
+
+\noindent The sum of the distances between two observations $x$ and $y$ in terms of cyclic granularity $A$ is defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity $A$) or $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$.
+After determining the distance between two series in terms of one granularity, we must combine them to produce a distance based on all significant granularities.
+When combining distances from individual $L$ cyclic granularities $C_l$ with $n_l$ levels, $$S_{x, y} = \sum_lS^{C_l}_{x,y}/n_l$$ is employed, which is also a distance metric since it is the sum of JS distances.
+This approach is expected to yield groups, such that the variation in observations within each group is in magnitude rather than distributional pattern, while the variation between groups is only in distributional pattern across categories.
+
+**wpd-based distances**
+
+Compute weighted pairwise distances $wpd$ (@Gupta2021-hakear) for all considered granularities for all observations.
+$wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction and is estimated by the maximum pairwise distances between consecutive categories normalized by appropriate parameters.
+A higher value of $wpd$ indicates that some interesting pattern is expected, whereas a lower value would indicate otherwise.
+
+\noindent Once we have chosen $wpd$ as a relevant feature for characterizing the distributions across one cyclic granularity, we have to decide how we combine differences between the multiple features (corresponding to multiple granularities) into a single number.
+The Euclidean distance between them is chosen, with the granularities acting as variables and $wpd$ representing the value under each variable.
+With this approach, we should expect the observations with similar $wpd$ values to be clustered together.
+Thus, this approach is useful for grouping observations that have a similar significance of patterns across different granularities.
+Similar significance does not imply a similar pattern, which is where this technique varies from JS-based distances, which detect differences in patterns across categories.
+
+<!-- Consider a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$.
+Each household consists of a $J*K$ distributions one harmony.
+We compute the distributional difference -->
+<!-- between $(A, B)$ for the $s^{th}$ household using $wpd_{{s}}(A,B)$. -->
+<!-- $wpd_{{s}}(A,B)$ denotes the normalized weighted-pairwise distributional distances between $(A, B)$ and is a feature which measures distributional difference between harmonies.
+If we have $H_{N_C}$ harmonies in the harmony table, then for each household we have a vector of $wpd_{{s}}$ of $H_{N_C}$ elements with each element corresponding to one harmony.
+We aim to have pockets of households showing similar periodic behavior by considering $wpd$ values for different harmonies and some time series features.
+The features should also characterize probability distributions of different household.
+-->
+
+<!-- grouping probability distributions across a harmony.
+This clustering algorithm is adopted to remove or appropriately adjust for auto correlation and unequal length in the data.
+The method could be further extended by clustering probability distributions conditional on one or more cyclic granularities.
+The following are some of the advantages of our proposed method.
+-->
+
+- _Clustering algorithm_
+
+With a way to obtain pairwise distances, any clustering algorithm can be employed that supports the given distance metric as input.
+A good comprehensive list of algorithms can be found in @Xu2015-ja based on traditional ways like partition, hierarchy, or more recent approaches like distribution, density, and others.
+We employ agglomerative hierarchical clustering in conjunction with Ward's linkage.
+Hierarchical cluster techniques fuse neighboring points sequentially to form bigger clusters, beginning with a full pairwise distance matrix.
+The distance between clusters is described using a "linkage technique".
+This agglomerative approach successively merges the pair of clusters with the shortest between-cluster distance using Ward's linkage method.
+
+<!--Hierarchical algorithms are one of the most widely used, can operate with data of any shape, has reasonable scalability, and the number of clusters is not needed as a parameter.-->
+
+- _Characterization of clusters_
+
+\noindent Cluster characterization is an important element of cluster analysis.
+@Cook2007-qe provides several methods for characterizing clusters.
+_Parallel coordinate plots_ (@wegman1990hyperdimensional), _Scatterplot matrix_, _Displaying cluster statistics_ (@dasu2005grouping), _MDS_ (@borg2005modern), PCA, _t-SNE_(@R-tsne), _Tour_ (@wickham2011tourr) are some of the graphical approaches used in this study.
+A Parallel Coordinates Plot features parallel axes for each variable and each axis is linked by lines.
+Changing the axes may reveal patterns or relationships between variables for categorical variables.
+However, for categories with cyclic temporal granularities, preserving the underlying ordering of time is more desirable.
+Displaying cluster statistics is useful when we have larger problems and it is difficult to read the parallel coordinate plots due to congestion.
+All of MDS, PCA and t-SNE use a distance or dissimilarity matrix to construct a reduced-dimension space representation, their goals are diverse.
+Multidimensional scaling (@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space.
+The t-SNE embedding will compress data points that are close in high-dimensional space.
+Tour is a collection of interpolated linear projections of multivariate data into lower-dimensional space.
+The cluster characterization approach varies depending on the distance metric used.
+Parallel coordinate plots, scatter plot matrices, MDS or PCA are potentially useful ways to characterize clusters using wpd-based distances.
+For JS-based distances, plotting cluster statistics is beneficial for characterization and variable importance could be displayed through parallel coordinate plots.
+
+<!-- CUT IT SHORT -->
+
+<!-- (a) _Parallel coordinate plots_ (@wegman1990hyperdimensional) are often used to visualize high-dimensional and multivariate data, allowing visual grouping and pattern detection.
+A Parallel Coordinates Plot features parallel axes for each variable.
+Each axis is linked by lines.
+Changing the axes may reveal patterns or relationships between variables for categorical variables.
+However, for categories with cyclic temporal granularities, preserving the underlying ordering of time is more desirable.
+-->
+
+<!-- (b) _Scatterplot matrix_ contains pairwise scatter plots of the p variables.
+Pairwise scatter plots are useful for figuring out how variables relate to each other and how factors determine the clustering.
+-->
+
+<!-- (c) _Displaying cluster statistics_ are useful when we have larger problems and it is difficult to read the Parallel coordinate plots due to congestion. (@dasu2005grouping) -->
+
+<!-- (d) _MDS, PCA and t-SNE_ While all of them use a distance or dissimilarity matrix to construct a reduced-dimension space representation, their goals are diverse.
+PCA seeks to retain data variance.
+Multidimensional scaling (@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space.
+t-SNE, on the other hand, is concerned with preserving neighborhood data points.
+The t-SNE embedding will compress data points which are close in high-dimensional space.
+-->
+
+<!-- (e) _Tour_ is a collection of interpolated linear projections of multivariate data into lower-dimensional space.
+As a result, the viewer may observe the high-dimensional data's shadows from a low-dimensional perspective.
+-->
+
+<!-- The cluster characterization approach varies depending on the distance metric used.
+Parallel coordinate plots, scatter plot matrices, MDS or PCA are potentially useful ways to characterize clusters using wpd-based distances.
+For JS-based distances, plotting cluster statistics is beneficial for characterization and variable importance could be displayed through parallel coordinate plots.
+-->
+
+<!-- This part of the work uses R packages `GGally` (@R-GGally), `Rtsne` (@R-tsne), `ggplot2` (Wickham2009pk), `tour` (@wickham2011tourr), `stats` (@R-language). -->
+
+<!-- - A random sample of the original data is taken for clustering analysis and includes missing and noisy observations (detailed description in Appendix) -->
+
+<!-- - All harmonies are computed for each customer in the sample. -->
+<!-- Cyclic granularities which are clashes for all customers in the sample are removed. -->
+
+<!-- - It is worth noting that a number of other solutions may be considered at the pre-processing stage of the method.
+We have considered a) Normal-Quantile Transform and b) Robust transformation.
+-->
+
+<!-- - Two methods are considered for computing dissimilarity between two customers.
+The first one involves computing according to one granularity is computed as the sum of the JS distances between distribution of all the categories of the granularity.
+When we consider more than one granularity, we consider the sum of the average distances for all the granularity so that the combined metric is also a distance.
+-->
+
+<!-- - Given the scale of dissimilarity among the energy readings, the model chooses optimal number of clusters -->
+
+<!-- - Once clusters have been allocated, the groups are explored visually. -->
+
+<!-- - Results are reported and compared. -->
+
+<!-- Two methods are used for computing distances between subjects and then hierarchical clustering algorithm is used. -->
+
+<!-- The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the probability distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
+In our work, we are using $i$ as denoting a customer and the underlying variable as the electricity demand.
+So $f_i(v)$ is the distribution of household $i$ and $v$ is electricity demand.
+-->
+
+<!-- We want to cluster distributions of the form $f_{i,j,k}(v)$, where $i$ and $j$ denote -->
+
+<!-- Consider a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$.
+Each household consists of a $J*K$ distributions one harmony.
+We compute the distributional difference -->
+<!-- between $(A, B)$ for the $s^{th}$ household using $wpd_{{s}}(A,B)$. -->
+<!-- $wpd_{{s}}(A,B)$ denotes the normalized weighted-pairwise distributional distances between $(A, B)$ and is a feature which measures distributional difference between harmonies.
+If we have $H_{N_C}$ harmonies in the harmony table, then for each household we have a vector of $wpd_{{s}}$ of $H_{N_C}$ elements with each element corresponding to one harmony.
+We aim to have pockets of households showing similar periodic behavior by considering $wpd$ values for different harmonies and some time series features.
+The features should also characterize probability distributions of different household.
+-->
+
+<!-- ### Notations -->
+
+<!-- Consider an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the probability distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
+In our work, we are using $i$ as denoting a household and the underlying variable as the electricity demand.
+Further consider a cyclic granularity of the form $B = \{ b_k: k = 1, 2, \dots, K\}$.
+Each customer consists of collection of probability distributions.
+-->
+
+<!-- So $f_i(v)$ is the distribution of household $i$ and $v$ is electricity demand.
+We want to cluster distributions of the form $f_{i,j,k}(v)$, where $i$ and $j$ denote $i^{th}$ and $j^{th}$ customer respectively.
+-->
+
+<!-- a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$. -->
+
+<!-- ### A single or pair of granularities together (change names) -->
+
+<!-- The methodology can be summarized in the following steps: -->
+
+<!-- - _Pre-processing step_ -->
+
+<!-- Robust scaling method or NQT used for each customer. -->
+
+<!-- - _NQT_ -->
+
+<!-- - _Treatment to outliers_ -->
+
+<!-- ### Many granularities together (change names) -->
+
+<!-- The methodology can be summarized in the following steps: -->
+
+<!-- 1. Compute quantiles of distributions across each category of the cyclic granularity -->
+<!-- 2. Compute JS distance between households for each each category of the cyclic granularity -->
+<!-- 3. Total distance between households computed as sum of JS distances for all hours -->
+<!-- 4. Cluster using this distance with hierarchical clustering algorithm (method "Ward.D") -->
+
+<!-- _Pro:_  -->
+<!-- - distance metric makes sense to group different shapes together  -->
+<!-- - simulation results look great on typical designs  -->
+<!-- _Cons:_   -->
+<!-- - Can only take one granularity at once   -->
+<!-- - Clustering a big blob of points together whereas the aim is to groups these big blob into smaller ones   -->
+
+<!-- ### Multiple-granularities -->
+
+<!-- _Description:_   -->
+
+<!-- Choose all significant granularities and compute wpd for all these granularities for all customers.
+Distance between customers is taken as the euclidean distances between them with the granularities being the variables and wpd being the value under each variable for which Euclidean distance needs to be measured.
+-->
+<!-- _Pro:_   -->
+<!-- - Can only take many granularities at once -->
+<!-- - can apply variable selection PCP and other interesting clustering techniques -->
+<!-- - simulation results look great on typical designs -->
+<!-- - splitting the data into similar sized groups   -->
+<!-- _Cons:_   -->
+<!-- - distance metric does not make sense to split the data into similar shaped clusters  -->
+
 ## Validation {#sec:validation}
+
+To validate our clustering methods, we spiked many attributes in the data to create different data designs.
+<!--to see where one method works better than the other and where they might give us the same outcome --or the effect of missing data <!---and trends on the proposed methods.--> Three circular granularities $g1$, $g2$ and $g3$ are considered with categories denoted by $\{g10,g11\}$, $\{g20, g21, g22\}$ and $\{g30, g31, g32, g33, g34\}$ and levels $n_{g_1}=2$, $n_{g_2}=3$ and $n_{g_3}=5$.
+These categories could be integers or some more meaningful labels.
+For example, the granularity "day-of-week" could be either represented by $\{0, 1, 2, \dots, 6\}$ or $\{Mon, Tue, \dots, Sun\}$.
+Here categories of $g1$, $g2$ and $g3$ are represented by $\{0, 1\}$, $\{0, 1, 2\}$ and $\{0, 1, 2, 3, 4\}$ respectively.
+A continuous measured variable $v$ of length $T$ indexed by $\{0, 1, \dots T-1\}$ is simulated such that it follows the structure across $g1$, $g2$ and $g3$.
+We constructed independent replications of all data designs $R = \{25, 250, 500\}$ to investigate if our proposed clustering method can discover distinct designs in small, medium, and big number of series.
+All designs employ $T=\{300, 1000, 5000\}$ sample sizes to evaluate small, medium, and large-sized series.
+Variations in method performance may be due to different jumps between categories.
+So a mean difference of $\mu = \{1, 2, 5\}$ between categories is considered.
+The performance of the approaches varies with the number of granularities which has interesting patterns across its categories.
+So three scenarios are considered to accommodate that.
+
+<!-- The results for $T=300$ and $R=25$ is shown, that means we have $25$ time series each with length $300$. The rest of the results could be found in the supplementary paper. -->
+
 ### Data generating processes
+
+<!-- An ARMA (p,q) process is used to generate series, where $p$ and $q$ are selected at random such that the series is stationary.
+The various designs on $g1$, $g2$, and $g3$ are introduced by adding matching designs to this series' innovations.
+The innovations are considered to have a normal distribution, although they follow the same pattern as the designs.
+To eliminate the effect of starting values, the first 500 observations in each series are discarded.
+-->
+
+Each category or combination of categories from $g1$, $g2$ and $g3$ are assumed to come from the same distribution, a subset of them from the same distribution, a subset of them from separate distributions, or all from different distributions, resulting in various data designs.
+As the methods ignore the linear progression of time, there is little value in adding time dependency to the data generating process.
+The data type is set to be "continuous," and the setup is assumed to be Gaussian.
+When the distribution of a granularity is "fixed", it means distributions across categories do not vary and are considered to be from N (0,1).
+$\mu$ alters in the "varying" designs, leading to varying distributions across categories.
+
+<!-- An ARMA (p,q) process is used to generate series, where $p$ and $q$ are selected at random such that the series is stationary.
+The various designs on $g1$, $g2$, and $g3$ are introduced by adding matching designs to this series' innovations.
+The innovations are considered to have a normal distribution, although they follow the same pattern as the designs.
+To eliminate the effect of starting values, the first 500 observations in each series are discarded.
+-->
+
+<!-- It is often reasonable to construct a time series using properties such as trend, seasonality, and auto-correlation.
+However, when examining distributions across categories of cyclic granularities, these time series features are lost or addressed independently by considering seasonal fluctuations through cyclic granularities.
+Because the time span during which an entity is observed in order to ascertain its behavior is not very long, -->
+<!-- the behavior of the entity will not change drastically and hence the time series can be assumed to remain stationary throughout the observation period.
+If the observation period is very long (for e.g more than 3 years), property, physical or geographical attributes might change leading to a non-stationary time series.
+But such a scenario is not considered here and the resulting clusters are assumed to be time invariant in the observation period.
+-->
+
 ### Data designs
+
 #### Individual granularities
+
+_Scenario (a): All significant granularities_
+
+<!-- Consider a case where all the three granularities $g1$, $g2$ and $g3$ would be responsible for making the designs distinct.
+That would mean, the pattern for each of $g1$, $g2$ and $g3$ will change for at least one design.
+We consider a situation with all the null cases corresponding to no difference in distribution across categories, that is, all categories follow the same distribution N(0,1).
+-->
+
+\noindent Consider the instance where $g1$, $g2$, and $g3$ all contribute to design distinction.
+This means that each granularity will have significantly different patterns at least across one of the designs to be clustered.
+In Table \ref{tab:tab-dist-design} various distributions across categories are considered (top) which lead to different designs (bottom).
+Figure \ref{fig:plot-3gran-new} shows the simulated variable's linear (left) and cyclic (right) representations for each of these five designs.
+The structural difference in the time series variable is impossible to discern from the linear view, with all of them looking very similar.
+The shift in structure may be seen clearly in the distribution of cyclic granularities.
+The following scenarios use solely graphical displays across cyclic granularities to highlight distributional differences in categories.
+
+_Scenario (b): Few significant granularities_
+
+\noindent This is the case where one granularity will remain the same across all designs.
+We consider the case where the distribution of $v$ varies across $g2$ levels for all designs, across $g3$ levels for a few designs, and $g1$ does not vary across designs.
+The proposed design is shown in Figure \ref{fig:gran2and1-clubbed}(b).
+
+_Scenario (c): One significant granularity_
+
+\noindent Only one granularity is responsible for identifying the designs in this case.
+This is depicted in Figure \ref{fig:gran2and1-clubbed} (right) where only $g3$ affects the designs significantly.
+
+```{r tab-distribution}
+```
+
+```{r tab-design}
+```
+
+```{r tab-dist-design}
+```
+
+```{r generate-D3change}
+
+```
+
+```{r generate-design-3change}
+
+```
+
+```{r plot-3gran-new, out.width="100%", fig.cap="The linear (left) and cyclic (right) representation of the simulated variable is shown.
+Each row represents a design in Scenario (a).
+In this scenario, all of $g1$, $g2$ and $g3$ changes across at least one design.
+Also, it is not possible to comprehend these differences in patterns just by looking at or considering the linear representation."}
+
+```
+
+```{r generate-design-new}
+
+```
+
+```{r generate-design-2gran-data}
+
+```
+
+```{r generate-design-2gran-plot}
+
+```
+
+```{r generate-design-1gran-data}
+
+```
+
+```{r generate-design-1gran-plot}
+
+```
+
+```{r gran2and1-clubbed, fig.cap=" Plots (b) and (c) correspond to Design (b) and (c) respectively.
+In (b) $g2$, $g3$ changes across atleast one design but $g1$ remains constant.
+Only $g3$ changes across different designs in (c)."}
+#gran2_change  + gran1_change + plot_layout(widths = c(1,1))
+#ggpubr::ggarrange(gran1_change, gran2_change, labels = c("a", "b"))
+```
+
 #### Interaction of granularities
+
+The proposed methods could be extended when two granularities of interest interact and we want to group subjects based on the interaction of the two granularities.
+Consider a group that has a different weekday and weekend behavior in the summer but not in the winter.
+This type of combined behavior across granularities can be discovered by evaluating the distribution across combinations of categories for different interacting granularities (Weekend/Weekday and month-of-year in this example).
+As a result, in this scenario, we analyze a combination of categories generated from different distributions.
+Consider a case in which there are only two interacting granularities of interest, $g1$ and $g2$.
+In contrast to the previous situation, when we could study distributions across $n_{g_1} + n_{g_2} = 5$ separate categories, with interaction, we must evaluate the distribution of the $n_{g_1}*n_{g_2}=6$ combination of categories.
+Consider the $4$ designs in Figure \ref{fig:interaction-gran}, where various distributions are assumed for different combinations of categories, resulting in different designs.
+Design $D1$ exhibits no change in distributions across $g1$ or $g2$, whereas Designs $D2$ and $D3$ alter across only $g1$ and $g2$, respectively.
+D4 varies across both $g1$ and $g2$ categories.
+D3 and D4 appear similar based on their relative differences across consecutive categories, but $D4$ also changes across facets, unlike $D3$, which has all facets look the same.
+
+```{r interaction-gran, fig.cap = "Distribution of the simulated variable across $g1$ conditional on $g2$ is shown through boxplots for 4 designs.
+D1 has no change in distributions across different categories of $g1$ or $g2$, while D2 and D3 change across only $g1$ and $g2$ respectively.
+D4 changes across categories of both $g1$ and $g2$."}
+
+```
+
+<!-- When two granularities of interest interact, the connection between a granularity and the measured variable is determined by the value of the other interacting granularity.
+This happens when the effects of the two granularities on the measured variable are not additive.
+For simplicity, consider a case with just two interacting granularities $g1$ and $g2$ of interest.
+As opposed to the last case, where we could play with the distribution of $5$ individual categories, with interaction we can play with the distribution of $6$ combination of categories.
+Consider $4$ designs in Figure \ref{fig:} where different distributions are assumed for different designs to get some distinction across designs.
+For example, in application, think about the scenario when customers need to grouped basis their joint behavior across hour-of-day and month-of-year.
+-->
+
+<!-- | Granularity type                                                   	| # Significant 	| # Replications 	| -->
+<!-- |--------------------------------------------------------------------	|---------------	|----------------	| -->
+<!-- | **Individual**  <br><br># obs: 300, 500, 2000  <br># clusters: 4/5 	| 1/2/3         	| 25, 100, 200   	| -->
+<!-- | **Interaction**  <br><br># obs: 500, 2000  <br># clusters: 4       	| 2           	| 25, 100, 200   	| -->
+
 ### Visual exploration of findings
+
+All of the approaches were fitted to each data design and for each combination of the considered parameters.
+The formed clusters have to match the design, be well separated, and have minimal intra-cluster variation.
+It is possible to study these desired clustering traits visually in a more comprehensive way than just looking at index values.
+So we use MDS and parallel coordinate graphs to demonstrate the findings:
+
+- In Figure \ref{fig:mds-plot-validation}, we tried to see how separated our clusters are.
+We observe that in all scenarios and for different mean differences, clusters are separated.
+However, the separation increases with an increase in mean differences across scenarios.
+This is intuitive because, as the difference between categories increases, it gets easier for the methods to correctly distinguish the designs.
+
+- Figure \ref{fig:parcoord-sim} depicts a parallel coordinate plot with the vertical bar showing total inter-cluster distances with regard to granularities $g1$, $g2$, and $g3$ for all simulation settings and scenarios.
+So one line in the figure shows the inter-cluster distances for one simulation setting and scenarios vary across facets.
+The lines are not colored by group since the purpose is to highlight the contribution of the factors to categorization rather than class separation.
+The first plot shows that no variable stands out in the clustering, but the following two designs show that {g1} and {g1, g2} have very low inter cluster distances, meaning that they did not contribute to the clustering.
+It is worth noting that these facts correspond to our original assumptions when developing the scenarios, which incorporate distributional differences over three (a), two (b), and one (c) significant granularities.
+Hence, Figure \ref{fig:parcoord-sim} (a), (b), and (c) validate the construction of scenarios (a), (b), and (c) respectively.
+
+ - The js-robust and wpd methods perform worse for $nT=300$, then improve for higher $nT$ evaluated in the study.
+Although, a complete year of data is the minimum requirement to capture distributional differences in winter and summer profiles, for example.
+Even if the data is only available for a month, $nT$ with half-hourly data is expected to be at least $1000$.
+As a result, as long as the performance is promising for higher $nT=300$, this is not a challenge.
+
+- In our study sample, the method js-nqt outperforms the method js-robust for smaller differences between categories. More testing, however, is required to corroborate this.
+
+For more detailed results, please refer to the supplementary paper.
+The code for creating these data designs and running the methodologies is available at (https://github.com/Sayani07/paper-gracsR/Validation).
+
+```{r parcoord-sim, fig.cap = "The parallel coordinate plot illustrates the total inter-cluster distances for granularities $g1$, $g2$, and $g3$.
+One line in the figure depicts the inter-cluster distances for a single simulation scenario.
+While the first plot indicates that no variable stands out during clustering, the next two designs demonstrate that {g1} and {g1, g2} have extremely low inter-cluster distances, indicating that they did not contribute to clustering.
+It is worth emphasising that these facts are consistent with our initial assumptions when designing the scenarios and (a), (b), and (c) correspond to Scenario (a), (b) and (c) respectively."}
+
+```
+
+```{r mds-plot-validation, fig.cap = "Relative positions of clusters corresponding to different scenarios (columns) for different values of mean differences between categories (rows) are shown using the first two dimensions of MDS.
+It can be observed that clusters become more compact and separated for higher mean differences between categories across all designs.
+Between designs, separation is least prominent corresponding to scenario (c) where only granularity is responsible for the clusters."}
+
+```
+
+<!-- "Parallel coordinate plots are used to identify key variables for classification.
+Each line in this figure represents a formed group for a certain set of simulated parameters.
+(a), (b) and (c) corresponds to Scenarios (a), (b) and (c).
+The cyclic granularities $g1$, $g2$, and $g3$ are plotted on the x-axis, while total inter-cluster distances are plotted on the y-axis.
+Each line in this figure represents a group for a certain configuration of the various parameters ($nT$, $diff$, $R$).
+(c) demonstrates that intercluster distances are only significant for the variable $g3$, meaning that the clusters formed are primarily due to $g3$.
+This corresponds to our design in Scenario (c), in which distributional differences were implemented solely for g3.
+(b) demonstrates the role of g2 and g3 in clustering.
+(a) depicts a heterogeneous pattern of inter-cluster distances across variables, indicating that no single variable is to account for the clusters." -->
+
+<!-- A confusion table can come alive with linked brushing, so that mismatches and agreements between methods can be explored. -->
+
+```{r, eval=FALSE}
+# code in append_3gran_change.R
+table <- read_rds("data/append_3gran_change.rds")
+table %>% kable()
+```
+
+<!-- starts getting better with increasing difference and get worse with increasing number of replications.
+Length of series do not show to have any effect on the performance of the methods.
+It does not depend on if time series is ar or arma.
+-->
+
+<!-- - confusion matrix could be used for showing results if proper labeling is used -->
+
+<!-- - write about features that we have spiked into the data set -->
+<!-- - write about you incorporated noise -->
+<!-- - What is the additional structure you can incorporate that will lead to failing of method1 and method2? -->
+<!-- - And both gives the same result? Basically say when method 1 works better than method 2 and vice versa! -->
+
+<!-- - -->
+
 ## Case study {#sec:cases}
+
+The use of our methodology is illustrated on smart meter energy usage for a sample of customers from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) which was available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) and Data61 CSIRO.
+It contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.
+In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual.
+We do not learn about these repetitive behaviors from the linear view because too many measurements all squeezed in that representation.
+Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior.
+The raw data for these consumers is of unequal length, with varying start and finish dates.
+Because our proposed methods evaluate probability distributions rather than raw data, neither of these data features would pose any threat to our methodology unless they contained any structure or systematic patterns.
+Additionally, there were missing values in the database but further investigation revealed that there is no structure in the missingness (see Supplementary paper for raw data features and missingness).
+The study begins by subsetting a data set along all dimensions of interest using data filtering and prototyping.
+By grouping the prototypes using our methods and assessing their meaning, the study hopes to unravel some of the heterogeneities observed in energy usage data.
+Because our application does not employ additional customer data, we cannot explain why consumption varies, but rather try to identify how it varies.
+
+_Data filtering and variable selection_
+
+- Choose a smaller subset of randomly selected $600$ customers with no implicit missing values for 2013.
+
+- Obtain $wpd$ for all cyclic granularities considered for these customers.
+It was found that `hod` (hour-of-day), `moy` (month-of-year) and `wkndwday` (weekend/weekday) are coming out to be significant for most customers.
+We use these three granularities while clustering.
+
+- Remove customers whose data for an entire category of `hod`, `moy` or `wnwd` is empty.
+For example, a customer who does not have data for an entire month is excluded because their monthly behavior cannot be analyzed.
+
+- Remove customers whose energy consumption is 0 in all deciles.
+These are the clients whose consumption is likely to remain essentially flat and with no intriguing repeated patterns that we are interested in studying.
+
+_Prototype selection_
+
+<!-- Why instance selection -->
+Supervised learning uses a training set of known information to categorize new events through instance selection.
+Instance selection (@olvera2010review) is a method of rejecting instances that are not helpful for classification.
+This is analogous to subsampling the population along all dimensions of interest such that the sampled data represents the primary features of the underlying distribution.
+Instance selection in unsupervised learning has received little attention in the literature, yet it could be a useful tool for evaluating model or method performance.
+There are several ways to approach prototype selection.
+Following @Fan2021-bq's idea of picking related examples (neighbors) for each instance (anchor), we can first use any dimensionality reduction techniques like MDS or PCA to project the data into a 2D space.
+Then pick a few "anchor" customers who are far apart in 2D space and pick a few neighbors for each.
+Unfortunately, this does not ensure that consumers with significant patterns across all variables are chosen.
+Tours can reveal variable separation that was hidden in a single variable display better than static projections.
+Hence we perform a linked tour with a t-SNE layout using the R package `liminal` (@R-liminal) to identify customers who are more likely to have distinct patterns across the variables studied.
+(Refer to Supplementary article for further details).
+Figure \ref{fig:hod-ind-group} shows the distribution across `hod` (a), `moy`(b) and `wnwd` (c) for the set of chosen $24$ customers that were chosen.
+Few of these customers have similar distribution across `moy` and some are similar in their `hod` distribution.
+
+```{r prototype-data-pick}
+
+```
+
+```{r assemble}
+
+```
+
+```{r data-pick}
+```
+
+<!-- - can be done in several ways -->
+
+<!-- - show raw data of moy, hod, wkndwday patterns for them -->
+
+```{r hod-moy-wkndwday plots, fig.cap="The distribution across moy, hod and wkndwday for the selected designs.
+Few are similar in their hod pattern, while others are similar in moy behavior.
+Some customers have distinct behavior as compared to all other customers.For example, although patterns across wkndwday do not look distinctly different for most households, there is one household for whom weekend behavior is standing out from the rest."}
+```
+
 ### Clustering
+
+The $24$ prototypes <!--and the data subset consisting of $353$ customers--> are clustered using the methodology described in Section \ref{sec:methodology} and results are reported below.
+In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$.
+All customers with the same color represent the same clustered groups.
+Groups by JS-based distances and wpd-based distances are colored differently as they represent different groupings.
+The plotting scales are not displayed since we want to emphasize comparable shapes rather than scales.
+The idea is that a customer in a cluster may have low total energy usage, but their behavior may be quite similar to a customer with high usage with respect to distributional pattern or significance across cyclic granularities.
+
 #### JS-based distances
+
+For clustering based on JS-based distances, we chose the optimal number of clusters using (@Hennig2014-ah) as $5$.
+The groupings are shown in Figure \ref{fig:hod-ind-group}.
+Customers with the same color represent the same clustered groups.
+Our methodology is useful for grouping similar distributions over `hod` and `moy` and they are placed closely for easy comparison.
+Few groups have mixed patterns across `hod` and `moy`, but few have all customers in the group having a similar profile.
+Figure \ref{fig:combined-groups-js} shows the summarized distributions across $5$ groups and assists us in characterizing each cluster.
+It shows Groups $2$ and $4$ have `hod` pattern with a typical morning and evening peak, whereas groups $1$, $3$, and $5$ show a `moy` pattern with higher usage in winter months.
+Differences in Weekend/Weekday between groups are not discernible, implying that it may not be a relevant variable in distinguishing various clusters unless maybe conditioned by `moy` or `hod`.
+It may be interesting to compare these two plots to verify if the summarized distributions across groups correctly characterised the groupings.
+If it has, then the majority of the group's members should share a similar profile.
+
+<!--  It appears that the aim of grouping comparable distributions across categories of multiple temporal dependencies has been accomplished to some extent. -->
+
+```{r all-data}
+```
+
+```{r clustering}
+
+```
+
+```{r opt-clusters}
+```
+
+```{r groups-24}
+```
+
+```{r hod-ind-group, fig.cap = "The distribution of selected customers over hod (a), moy (b), and wnwd (c) for the $24$ selected customers is shown.
+In each case, the same colour denotes the same group using js-based clustering and are placed together to facilitate comparison.
+Individuals with similar distributional differences across categories of `hod` or `moy` are grouped together, with some exceptions.
+" }
+```
+
+<!-- It appears that the aim of grouping comparable distributions over considered variables has been accomplished to some extent. -->
+
+```{r data-heatmap-hod-group}
+```
+
+```{r data-heatmap-moy-group}
+
+```
+
+```{r data-heatmap-wkndwday-group}
+
+```
+
+```{r combined-groups-js, fig.cap = "The distribution of electricity demand over hod (a), moy (b), and wnwd*hod (c).
+Groups 2 and 5 appear to have a hod pattern among its members, whereas groups 1, 3, and 5 appear to have a moy pattern.
+Weekend/weekday differences across groups are not discernible, showing that it is not a critical variable for clustering on its own.
+It is probable that the distribution over Weekend/Weekday will differ conditional on the categories of hod or moy.
+(c) demonstrates that the distribution is not different, except for the tails.
+It is useful to compare the summarised distributions of groups to those of individuals in order to establish that the majority of individuals in the group share the same characteristics."}
+```
+
 #### wpd-based distances
+
+We chose the optimal number of clusters using (@Hennig2014-ah) as $3$.
+A parallel coordinate plot with the three significant cyclic granularities is used to characterise the groups here.
+The variables are sorted according to their separation across classes (rather than their overall variation between classes).
+This means that $moy$ is the most important variable in distinguishing the groups followed by $hod$ and $wnwd$.
+There is only one customer who has significant $wpd$ across $wnwd$ and stands out from the rest of the customers.
+Group 3 has a higher wpd for hod than moy or wkndwday.
+Group 2 has the most distinct pattern across moy.
+Group 1 is a mixed group that has strong patterns on at least one of the three variables.
+The findings vary from js-based clustering, yet it is a helpful grouping.
+
+```{r data-pick-wpd}
+```
+
+```{r parcoord, fig.cap="Each of the 24 customers is represented by a parallel coordinate plot (a) with three wpd-based groupings.
+The plot shows that moy is the most important variable in identifying clusters, whereas wkdn-wday is the least significant and has the least fluctuation.
+One particular customer with high wpd across wknwday stands out in this display.
+Group 3 has a higher wpd for hod than moy or wkndwday.
+Group 2 has most discernible pattern across moy.
+Group 1 is a mixed group with strong patterns on atleast one of the three variables.
+All of these could be observed from the plot or the table (b) which shows median wpd values for each group.", message=FALSE, warning=FALSE}
+
+```
+
+Things become far more complicated when we consider a larger data set with more uncertainty, as they do with any clustering problem.
+Summarizing distributions across clusters with varied or outlying customers can result in a shape that does not represent the group.
+Furthermore, combining heterogeneous customers may result in similar-looking final clusters that are not effective for visually differentiating them.
+It is also worth noting that the Weekend/Weekday behavior in the given case does not characterize any cluster.
+This, however, will not be true for all of the customers in the data set.
+If more extensive prototype selection is used, resulting in more comprehensive prototypes in the data set, this method might be used to classify the entire data set into these prototype behaviors.
+However, the goal of this section was to have a few customers that have significant patterns over one or more cyclic granularities, apply our methodology to cluster them, and demonstrate that the method produces useful clusters.
+
 ## Conclusion and future work {#sec:conclusion}
+
+We offer two approaches for calculating pairwise distances between time series based on probability distributions over multiple cyclic granularities at once.
+Depending on the goal of the clustering, these distance metrics, when fed into a hierarchical clustering algorithm using Ward's linkage, yield meaningful clusters.
+Probability distributions provide an intuitive method to characterise noisy, patchy, long, and unequal-length time series data.
+Distributions over cyclic granularities help to characterise the formed clusters in terms of their repeating behavior over these cyclic granularities.
+Furthermore, unlike earlier efforts that group customers based on behavior across only one cyclic granularity (such as hour-of-day), our method is more comprehensive in detecting clusters with repeated patterns at all relevant granularities.
+<!--The JS-based approaches are faster than wpd-based approaches.-->
+
+There are few areas to extend this research.
+First, larger data sets with more uncertainty complicate matters, as is true for any clustering task.
+Characterizing clusters with varied or outlying customers can result in a shape that does not represent the group.
+Moreover, integrating heterogeneous consumers may result in visually identical end clusters, which are potentially not useful.
+Hence, a way of appropriately scaling it up to many customers such that anomalies are removed before clustering would be useful for bringing forth meaningful, compact and separated clusters.
+Secondly, we have assumed the time series to be stationary, and hence the distributions are assumed to remain constant for the observation period.
+In reality, however, it might change.
+For the smart meter example, the distribution for a customer moving to a different house or changing electrical equipment can change drastically.
+Our current approach can not detect these dynamic changes.
+Thirdly, it is possible that for a few customers, data for some categories from the list of considered significant granularities are missing.
+In our application, we have removed those customers and done the analysis but the metrics used should be able to incorporate those customers in the clustering by handling their missing categories.
+Finally, $wpd$ is computationally heavy even under parallel computation.
+Future work can make the computations more efficient so that they are easily scalable to a large number of customers.
+Moreover, experiments can also be run with non-hierarchy based clustering algorithms to verify if these distances work better with other algorithms.
+
 ## Acknowledgments {-}
+
+The authors thank the ARC Centre of Excellence for Mathematical and Statistical Frontiers [(ACEMS)](https://acems.org.au/home) for supporting this research.
+Sayani Gupta was partially funded by [Data61 CSIRO](https://data61.csiro.au/) during her PhD.
+The Monash eResearch Centre and eSolutions-Study Support Services supported this research in part through the resource usage of the MonARCH HPC Cluster.
+The Github repository, [github.com/Sayani07/paper-gracsr](https://github.com/Sayani07/paper-gracsr), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
+This article was created with R [@R-language], `knitr` [@knitr, @R-knitr] and `rmarkdown` [@rmarkdown, @R-rmarkdown].
+Graphics are produced with `ggplot2` [@Wickham2009pk] and GGally [@R-GGally].
+
 ## Supplementary Materials
+
+**Data and scripts:** Data sets and R code to reproduce all figures in this article (main.R).
+
+\noindent
+**Supplementary paper:** Additional tables, graphics and and R code to reproduce it (paper-supplementary.pdf, paper-supplementary.Rmd)
+
+\noindent
+**R-package:** To implement the ideas provided in this research, the open-source R package 'gracsr' is available on Github (https://github.com/Sayani07/gracsr).
 
@@
 
-Placeholder
+In this thesis, I present a systematic approach for visualizing and analyzing large temporal data distributions.
+The building blocks of this framework are presented in this thesis.
+Chapter \@ref(ch:gravitas) includes tools for computing all cyclic granularities as well as a recommendation system for selecting pairs of granularities that may be effectively investigated together.
+These temporal granularities may be used to generate data visualizations to search for patterns, associations, and anomalies.
+However, when there are many granularities that can be constructed for a time period, there will also be too many possible displays to decide which might be the more interesting to display.
+Chapter \@ref(ch:hakear) presents a framework for selecting displays that are interesting, with the greatest differences between the displayed distributions, and ranking them in order of priority for capturing the most variation.
+Both of these are used for studying patterns in individual time series or comparing a few time series together.
+In Chapter \@ref(ch:gracsr), it is extended to allow for the exploration of distributions for many time series at the same time by clustering them based on probability distributions across informative cyclic granularities.
+Through the use of probability distribution, this technique is more comprehensive in recognizing clusters with recurring patterns over many important granularities and more resilient to noisy, patchy, and uneven length time series.
+
+```{r software-initial, echo = FALSE, cache = FALSE, include = FALSE}
+read_chunk('scripts/software.R')
+```
 
 ## Software development
+
+```{r software-impact}
+```
+
+This thesis focuses on integrating research approaches into open source R packages such as **gravitas**, **hakear**, and **gracsr**.
+
+<!-- Figure \@ref(fig:software-ghcommits) gives an overview of my Git commits to these repositories, and Figure \@ref(fig:software-downloads) shows the daily downloads of the packages from the RStudio mirror (one of 90 CRAN mirrors) since they were available on CRAN.
+-->
+
+<!-- ```{r software-ghcommits, fig.cap = "(ref:software-ghcommits-cap)"} -->
+<!-- include_graphics("img/pkg-commits.png") -->
+<!-- ``` -->
+
+<!-- (ref:software-ghcommits-cap) Patterns of my package development effort during my PhD years based on Git commits to three repositories, sugrrants, tsibble, and mists.
+Scatter plots of weekly totals are overlaid with a loess smoother.
+The **sugrrants** package was the first project with much initial energy, followed by small constant maintenance.
+The **tsibble** package has been a major project with ongoing constant development and bursts of effort in response to users' feedback.
+The **mists** package has been a recent intense project.
+-->
+
+<!-- # ```{r software-downloads, fig.height = 3, fig.cap = "(ref:software-downloads-cap)"} -->
+<!-- # ``` -->
+<!-- #  -->
+<!-- # (ref:software-downloads-cap) Impact of these works (sugrrants and tsibble) as measured by daily downloads (on square root scale) from the RStudio mirror since they landed on CRAN.
+The **tsibble** package has an increasing trend, suggesting the steady adoption of the new data structure.
+-->
+
 ### gravitas
+
+The **gravitas** package provides very general tools to compute and manipulate cyclic granularities, and to generate plots displaying distributions conditional on those granularities.
+It can be utilized in non-temporal cases for which a hierarchical structure can be construed similar to time.
+It is on CRAN.
+The website (<https://sayani07.github.io/gravitas>) includes full documentation and two vignettes about the package usage.
+There has been a grand total of 12K downloads from the RStudio mirror dating from 2020-11-01 to 2021-11-01.
+
 ### hakear
+
+The open-source R package `hakear` is available on Github (https://github.com/Sayani07/hakear) to implement ideas presented in Chapter \@ref(ch:hakear).
+Given a `tsibble` and context granularities of interest, the function `wpd()` provides support for computing the wpd for each cyclic granularities or pair of granularities and `select_harmonies()` chooses the ones with significant patterns and ranks them from highest to lowest wpd.
+
 ### gracsr
+
+The open-source R package `gracsr` is available on Github (https://github.com/Sayani07/gracsr) to implement ideas presented in Chapter \@ref(ch:gracsr).
+The package provides functions to carry out the entire clustering methodology discussed in the paper.
+It is still a work in progress and has won the ACEMS Business Analytics Prize 2021 with a prize money of AUD 3000, which would be utilized in polishing this package and preparing it for CRAN.
+
 ## Future work
+
 ### Putting all functionalities on CRAN
+
+I plan to integrate `hakear` with `gravitas` as one R package to systematically exploring few time series, whereas `gracsr` would provide the clustering framework for exploring many time series together.
+I plan to run it through https://ropensci.org/software-review/ to develop them further for more visibility and efficient usage.
+
 ### Scaling up the clustering method to incorporate large uncertainty and improved computational efficiency that comes with large data
+
+With the volume of data projected to grow in the future, potentially leading to increased variability in patterns, research is needed to understand the issues that may arise with the existing methodology and how to adapt our current algorithm.
+Computational efficiency is also critical when scaling up for analysis of huge data sets, let alone when adding features to existing highly dimensional data structures.
+
 ### Comparing our clustering method with other benchmark methods
+
+Testing needs to be carried out with non-hierarchy based clustering methods to see whether these distances perform better with other algorithms.
+Also, to evaluate how much information is lost when aggregating individual level demand versus distributions.
+
 ### Check generalizability of our methods
+
+We provide solutions that are realistically applicable to any temporal data observed more than once per year.
+We could just verify its value in the electricity smart meter context.
+With numerous open-source benchmark data sets accessible, it is necessary to test how well the approaches operate in various disciplines.
+
 ## Final words

### 2021-11-04 scripts/mists-wdi.R

@@ -1,0 +1,164 @@
+## ---- wdi-load
+library(tidyverse)
+library(mists)
+library(imputeTS)
+library(fable)
+
+wdi_ts <- wdi %>%
+  tsibble::as_tsibble(key = country_code, index = year)
+
+cutoff_grids <-
+  tibble(
+    measures = seq(0.4, 0.9, 0.1), key = measures,
+    index = measures, index2 = measures
+  ) %>%
+  expand(measures, key, index, index2) %>%
+  rename_all(~ paste0("na_polish_", .))
+
+## ---- wdi-vis
+wdi %>%
+  select(-year, -country_code) %>%
+  rename_at(vars(ag_lnd_frst_k2:tx_val_tech_mf_zs), ~ str_replace_all(., "_", " ")) %>%
+  visdat::vis_miss() +
+  theme(axis.text.x = element_text(angle = 90))
+
+## ---- wdi-run
+polish_res01 <- polish_res00 <- vector("list", NROW(cutoff_grids))
+for (i in seq_len(NROW(cutoff_grids))) {
+  polish_res01[[i]] <- na_polish_autotrace(wdi_ts, cutoff = c(cutoff_grids[i, ]))
+}
+
+polish_res_df01 <- bind_rows(polish_res01, .id = ".id") %>%
+  mutate(tol = .1)
+
+for (i in seq_len(NROW(cutoff_grids))) {
+  polish_res00[[i]] <- na_polish_autotrace(wdi_ts, cutoff = c(cutoff_grids[i, ]),
+    tol = 0)
+}
+polish_res_df00 <- bind_rows(polish_res00, .id = ".id") %>%
+  mutate(tol = 0)
+polish_res <- bind_rows(polish_res_df01, polish_res_df00)
+write_rds(polish_res, "data/polish-results.rds", compress = "bz2")
+
+## ---- wdi-pass
+polish_res <- read_rds("data/mists/polish-results.rds")
+polish_res %>%
+  group_by(.id, iteration, tol) %>%
+  summarise(iter_loss = unique(iter_loss)) %>%
+  rename(tolerance = tol) %>%
+  ggplot(aes(x = as.factor(iteration), y = iter_loss)) +
+  geom_line(aes(group = .id), size = .1, alpha = .1) +
+  facet_grid(~ tolerance, scales = "free_x", labeller = "label_both") +
+  xlab("Iteration") +
+  ylab("Loss")
+
+## ---- wdi-grid-plot
[…]

### 2021-11-04 scripts/mists.R

@@ -1,0 +1,137 @@
+## ---- mists-load
+library(tidyverse)
+library(lubridate)
+library(tsibble)
+library(imputeTS)
+library(mists)
+library(vctrs)
+theme_set(theme_bw())
+
+## ---- miss-types
+ts_airgap <- as_tsibble(tsAirgap)
+ts_airgap_full <- as_tsibble(tsAirgapComplete)
+
+# Types of temporal missings
+miss_at_time <- function(.data) {
+  idx <- pull(.data, !!index(.data))
+  init <- floor(vec_size(.data) * 0.1)
+  trend_idx <- idx[cumsum(rev(seq_len(init))) + init]
+  mutate(.data, value = ifelse(idx %in% trend_idx, NA, value))
+}
+
+miss_at_regular <- function(.data) {
+  idx <- pull(.data, !!index(.data))
+  rnd_seas <- month(sample(idx, size = 1))
+  mutate(.data, value = ifelse(month(idx) %in% rnd_seas, NA, value))
+}
+
+miss_at_occasional <- function(.data) {
+  idx <- pull(.data, !!index(.data))
+  rnd_idx <- sample(idx, size = round(vec_size(.data) * 0.1))
+  mutate(.data, value = ifelse(idx %in% rnd_idx, NA, value))
+}
+
+miss_at_runs <- function(.data) {
+  idx <- pull(.data, !!index(.data))
+  n <- vec_size(.data)
+  na_indices_lgl <- runif(n) < .1
+  na_indices <- which(na_indices_lgl)
+  na_lengths <- rpois(n = sum(na_indices_lgl), lambda = 5)
+  lst_indices <- map2(na_lengths, na_indices, function(.x, .y) 0:(.x - 1) + .y)
+  na_indices <- vec_c(!!!lst_indices)
+  rnd_run <- idx[na_indices]
+  mutate(.data, value = ifelse(idx %in% rnd_run, NA, value))
+}
+
+set.seed(2019)
+ts_airgap_time <- miss_at_time(ts_airgap_full) %>%
+  mutate(type = "Functional") %>%
+  update_tsibble(key = type)
+ts_airgap_regular <- miss_at_regular(ts_airgap_full) %>%
+  mutate(type = "Periodic") %>%
+  update_tsibble(key = type)
[…]

### 2021-11-04 scripts/tsibble-demo.R

@@ -1,0 +1,100 @@
+## ---- load-pkg
+library(tidyverse)
+library(tsibble)
+
+## ---- tb-sub
+tb_small <- read_rds("data/tsibble/tb-small.rds")
+tb_small %>%
+  filter(year > 2010) %>%
+  knitr::kable(booktabs = TRUE, caption = "(ref:tb-sub)", linesep = "") %>%
+  kableExtra::kable_styling(position = "center", latex_options= "hold_position")
+
+## ---- tb-print
+as_tsibble(tb_small, key = id(country, gender), index = year) %>%
+  filter(year > 2010) %>%
+  print(n = 5)
+
+## ---- tb-au
+tb_au <- tb_small %>%
+  filter(country == "Australia") %>%
+  group_by(year) %>%
+  summarise(count = sum(count))
+
+## ---- animate
+library(gganimate)
+reveal_group <- function(data, group){
+  group <- enquo(group)
+  data <- as_tibble(data)
+  data <- transmute(
+    data, !! group,
+    .dt = map(map(!! group, seq_len), function(groups, data) {
+      data <- filter(data, !! group %in% groups)
+      select(data, !! expr(-!! group))
+    }, data
+  ))
+  unnest(data, .dt)
+}
+types <- forcats::fct_inorder(c("Sliding", "Tiling", "Stretching"))
+slide_window <- slider(tb_au$year, .size = 5) %>%
+  map_dfr(function(x) tibble(xmin = min(x), xmax = max(x))) %>%
+  mutate(ymin = -Inf, ymax = Inf, group = row_number(), type = types[1])
+slide_mean <- tb_au %>%
+  mutate(ma = slide_dbl(count, ~ mean(.x), .size = 5, .align = "r")) %>%
+  mutate(group = pmax(0, row_number() - 4), type = types[1])
+slide_revealed <- slide_mean %>%
+  reveal_group(group)
+
+tile_window <- tiler(tb_au$year, .size = 5) %>%
+  map_dfr(function(x) tibble(xmin = min(x), xmax = max(x))) %>%
+  mutate(ymin = -Inf, ymax = Inf, type = types[2])
+tile_window <- tile_window[c(rep(1:2, each = 5), 3:4), ] %>% 
+  mutate(group = row_number())
+tile_mean <- tibble(
+    year = tile_dbl(tb_au$year, ~ median(.x), .size = 5),
[…]

### 2021-11-04 scripts/tsibble-flights.R

@@ -1,0 +1,137 @@
+## ---- load-flights-pkgs
+library(lubridate)
+library(tidyverse)
+library(tsibble)
+library(forcats)
+theme_set(theme_bw())
+
+## ---- load-flights
+flights <- read_rds("data/tsibble/flights.rds")
+
+## ---- find-duplicate
+flights %>% 
+  duplicates(key = flight_num, index = sched_dep_datetime) %>% 
+  as.data.frame()
+dup_lgl <- are_duplicated(flights, key = flight_num, 
+  index = sched_dep_datetime, from_last = TRUE)
+
+## ---- tsibble
+us_flights <- flights %>% 
+  filter(!dup_lgl) %>% 
+  as_tsibble(
+    key = flight_num, index = sched_dep_datetime,
+    regular = FALSE, validate = FALSE
+  )
+
+## ---- tsibble-header
+cat(format(us_flights)[1:2], sep = "\n")
+
+## ---- carrier-delayed
+delayed_carrier <- us_flights %>% 
+  mutate(delayed = dep_delay > 15) %>%
+  group_by(carrier) %>% 
+  index_by(year = ~ year(.)) %>% 
+  summarise(
+    Ontime = sum(delayed == 0),
+    Delayed = sum(delayed)
+  ) %>% 
+  gather(delayed, n_flights, Ontime:Delayed)
+
+## ----- carrier-mosaic
+library(ggmosaic)
+delayed_carrier %>% 
+  mutate(carrier = fct_reorder(carrier, -n_flights)) %>% 
+  ggplot() +
+    geom_mosaic(aes(x = product(carrier), fill = delayed, weight = n_flights)) +
+    scale_fill_brewer(palette = "Dark2", name = "Status") +
+    scale_x_productlist(name = "Carrier") +
+    scale_y_productlist(name = "Status") +
+    theme(legend.position = "bottom")
+
+## ---- sel-flights
+sel_flights <- us_flights %>% 
+  filter(origin %in% c("JFK", "SEA", "IAH", "KOA", "LAX"))
+
+## ---- sel-delay
+sel_delay <- sel_flights %>% 
+  group_by(origin) %>% 
+  index_by(sched_dep_date = ~ as_date(.)) %>% 
+  summarise(pct_delay = sum(dep_delay > 15) / n())
+
+## ----- sel-monthly-1
[…]

### 2021-11-04 scripts/tsibble-smart-meter.R

@@ -1,0 +1,118 @@
+## ---- load-elec
+library(tidyverse)
+library(lubridate)
+library(tsibble)
+library(fable)
+library(patchwork)
+theme_set(theme_bw())
+
+elec <- read_rds("data/tsibble/smart-meter13.rds")
+weather <- read_rds("data/tsibble/weather13.rds") %>%
+  group_by(date) %>%
+  mutate(avg_temp = mean(c(max_temp, min_temp))) %>%
+  ungroup() %>%
+  mutate(hot = if_else(avg_temp > 25, "Hot day", "Not hot day"))
+
+elec_ts <- elec %>%
+  build_tsibble(
+    key = customer_id, index = reading_datetime,
+    validate = FALSE, ordered = TRUE
+  )
+
+## ---- elec-gaps
+gap_df <- has_gaps(elec_ts)
+# sum(gap_df$.gaps) / NROW(gap_df)
+
+## ---- count-gaps
+count_na_df <- elec_ts %>%
+  count_gaps()
+
+lumped_na_df <- count_na_df %>%
+  mutate(
+    customer_id = as.factor(customer_id) %>%
+      fct_lump(50) %>% fct_reorder(.n, sum)
+  )
+
+p_49 <- lumped_na_df %>%
+  filter(customer_id != "Other") %>%
+  ggplot(aes(x = customer_id)) +
+  geom_linerange(aes(ymin = .from, ymax = .to)) +
+  geom_point(aes(y = .from), size = 0.6, shape = 4) +
+  geom_point(aes(y = .to), size = 0.6, shape = 4) +
+  coord_flip() +
+  xlab("Top customers") +
+  ylab("") +
+  theme(
+    axis.text = element_blank(),
+    axis.ticks = element_blank(),
+    legend.position = "none",
+    plot.margin = unit(c(0, 0, -1, -1), "line")
+  )
+
+p_other <- lumped_na_df %>%
+  filter(customer_id == "Other") %>%
+  ggplot(aes(x = customer_id)) +
+  geom_linerange(aes(ymin = .from, ymax = .to), alpha = 0.1) +
+  geom_point(aes(y = .from), size = 1.2, shape = 4, alpha = 0.1) +
+  geom_point(aes(y = .to), size = 1.2, shape = 4, alpha = 0.1) +
+  coord_flip() +
+  xlab("Rest") +
+  ylab("Time gaps") +
+  theme(
+    panel.grid.major.y = element_blank(),
+    axis.text.y = element_blank(),
[…]

### 2021-11-04 thesis.Rmd
gitbook update

@@
 
-# supervisors
+## supervisors
 Rob and Di are phenomenal leaders in their respective fields, who lead by example.
@@
 
-# department and PhD colleagues
-
-# R community
-
-# friends
+## department and PhD colleagues
+
+## R community
+
+## friends
 
@@
 
-# family
+## family
 Finally, a big thanks to my family for always being supportive of my choices.

### 2021-11-04 Rmd/00-d-acknowlegements.Rmd
atleast without ref runs

@@
 
-# supervisors
+supervisors
 Rob and Di are phenomenal leaders in their respective fields, who lead by example.
@@
 Thank you Di for being a fantastic female role model.
+- department and PhD colleagues
 
-# department and PhD colleagues
+- R community
 
-# R community
-
-# friends
+- friends
 
@@
 
-# family
+-  family
 Finally, a big thanks to my family for always being supportive of my choices.

### 2021-11-05 Rmd/02-gravitas.Rmd
index and gitignore files update
sometimes I do rock, Don't I? successful thesis compilation

@@
 This is consistent with the widely used grammar of graphics
-which is a framework to construct statistical graphics by relating the data space to the graphic space [@Wilkinson1999-nk, @Wickham2009pk].
+which is a framework to construct statistical graphics by relating the data space to the graphic space [@Wilkinson1999-nk], [@Wickham2009pk].
 
@@
 The Github repository,  [github.com/Sayani07/paper-gravitas](https://github.com/Sayani07/paper-gravitas), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
-This article was created with `knitr` [@knitr, @R-knitr] and `rmarkdown` [@rmarkdown, @R-rmarkdown].
+This article was created with `knitr` [@knitr], [@R-knitr] and `rmarkdown` [@rmarkdown], [@R-rmarkdown].
 

### 2021-11-05 Rmd/03-hakear.Rmd
sometimes I do rock, Don't I? successful thesis compilation

@@
 
-Cyclic temporal granularities [@Bettini1998-ed, @Gupta2021-hd] are <!--theoretical--> temporal deconstructions that define cyclic repetitions in time, e.g.
+Cyclic temporal granularities [@Bettini1998-ed], [@Gupta2021-hd] are <!--theoretical--> temporal deconstructions that define cyclic repetitions in time, e.g.
 hour-of-day, day-of-month, or regularly scheduled public holidays.
@@
 The Github repository,  [github.com/Sayani07/paper-hakear](https://github.com/Sayani07/paper-gravitas), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
-This article was created with R [@R-language], `knitr` [@knitr2015, @R-knitr] and `rmarkdown` [@rmarkdown2018, @R-rmarkdown].
+This article was created with R [@R-language], `knitr` [@knitr2015], [@R-knitr] and `rmarkdown` [@rmarkdown2018], [@R-rmarkdown].
 Graphics are produced with @Wickham2009pk.

### 2021-11-05 index.Rmd
sometimes I do rock, Don't I? successful thesis compilation

@@
 degrees: M.Stat, B.Sc (Stat Hons)
-bibliography:
-- bib/gravitas.bib
-- bib/hakear.bib
-- bib/gracsr.bib
-- bib/thesisrefs.bib
+bibliography: [bib/gravitas.bib, bib/hakear.bib, bib/gracsr.bib, bib/thesisrefs.bib]
 link-citations: yes

### 2021-11-05 Rmd/01-intro.Rmd
trying to make the gitbook work

@@
 Even analyzing the distribution of the measured variable across two cyclic granularities at once could amount to displaying many plots in search of potential  patterns.
-Thus, the first part of the research (@Gupta2020-vo) also introduces "harmony" to denote pairs of granularities that could be analyzed together and reduces the search from all possible options.
+Thus, the first part of the research also introduces "harmony" to denote pairs of granularities that could be analyzed together and reduces the search from all possible options.
 But this approach is still overwhelming for human consumption because there would still be huge number of harmonies.

### 2021-11-05 Rmd/06-appendix.Rmd
ensuring every chapter has #

--- 2021-11-04 thesis.Rmd
@@
----
-title: "Visualization and analysis of probability distributions of large temporal data"
-author: "Sayani Gupta"
-degreetype: Doctor of Philosophy
-site: bookdown::bookdown_site
-degrees: M.Stat, B.Sc (Stat Hons)
-bibliography:
-- bib/gravitas.bib
-- bib/hakear.bib
-- bib/gracsr.bib
-- bib/thesisrefs.bib
-link-citations: yes
----
-
-```{r initial, echo = FALSE, cache = FALSE, results = 'hide'}
-options("knitr.graphics.auto_pdf" = TRUE)
-library(knitr)
-opts_chunk$set(
-  warning = FALSE, message = FALSE, echo = FALSE,
-  fig.path = 'figure/', fig.align = 'center', fig.show = 'hold',
-  cache = TRUE, external = TRUE, comment = "#>", dpi = 300,
-  out.width = ifelse(is_html_output(), "100%", "\\textwidth"),
-  cache.path = ifelse(is_html_output(), 'cache-gitbook/', 'cache-pdfbook/')
-)
-hook_output <-  knit_hooks$get('output')
-knit_hooks$set(output = function(x, options) {
-  # this hook is used only when the linewidth option is not NULL
-  if (!is.null(n <- options$linewidth)) {
-    x = knitr:::split_lines(x)
-    # any lines wider than n should be wrapped
-    if (any(nchar(x) > n)) x = strwrap(x, width = n)
-    x = paste(x, collapse = '\n')
-  }
-  hook_output(x, options)
-})
-```
-
-`r if (knitr::is_latex_output()) '<!--'`
-
-# Welcome {-}
-
-This is the website for my PhD thesis at Monash University (Australia), titled "Visualization and analysis of probability distributions of large temporal data".
-
-`r Sys.Date()`
-
-`r if (knitr::is_latex_output()) '-->'`
-
-<!--chapter:end:index.Rmd-->
-
-`r if (knitr::is_html_output()) '<!--'`
-
-# Copyright notice {-}
-
-\textcopyright { }Sayani Gupta (\number\the\year).
-
-I certify that I have made all reasonable efforts to secure copyright permissions for third-party content included in this thesis and have not knowingly added copyright content to my work without the owner's permission.
-
-\newpage
-
-`r if (knitr::is_html_output()) '-->'`
-
-<!--chapter:end:Rmd//00-a-copyright.Rmd-->
-
-# Abstract {-}
-
-The research was motivated by the desire to understand the Australian smart meters data, which was collected half-hourly for two years at the household level.
-With temporal data available at ever finer scales, exploring periodicity can become overwhelming with so many possible temporal deconstructions to explore.
-Analysts are expected to comprehensively explore the many ways to view and consider temporal data.
-However, the plethora of choices and the lack of a systematic approach to do so quickly can make the task overwhelming.
-
-<!-- exploration may be required to be carried out across both finer and coarser scales to draw useful inferences about the underlying process.
-It is important to be able to navigate through all the temporal deconstructions that accommodate for periodicities to have multiple perspectives of the observed data.
-This idea aligns with the notion of EDA (Tukey 1977) which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
-Slicing and dicing the data in all possible temporal scales as suggested by EDA can be a daunting task as it leads to a -->
-<!-- myriad of possibilities. -->
-
-<!-- Analysts are expected to comprehensively explore the many ways to view and consider temporal data.
-However, the plethora of choices and the lack of a systematic approach to do so quickly can make the task overwhelming -->
-
-<!-- exploring periodicity can become overwhelming with so many possible temporal deconstructions to explore.
-Analysts are expected to comprehensively explore the many ways to view and consider temporal data.
-However, the plethora of choices and the lack of a systematic approach to do so quickly can make the task overwhelming.
--->
-
-<!-- <WHY> <WHAT> <HOW> -->
-
-This work investigates how time may be dissected, resulting in alternative data segmentation and, as a result, different visualizations that can aid in the identification of underlying patterns.
-The first contribution (Chapter \@ref(ch:gravitas)) describes classes of time deconstructions using linear and cyclic time granularities.
-It provides tools to compute possible cyclic granularities from an ordered (usually temporal) index and also a framework to systematically explore the distribution of a univariate variable conditional on two cyclic time granularities by defining "harmony".
-"harmony" denotes pairs of granularities that could be analyzed together and reduces the search from all possible options.
-This approach is still overwhelming for human consumption because there would still be a huge number of harmonies.
-Hence, the second contribution (Chapter \@ref(ch:hakear)) refines the search of informative granularities by identifying those for which the differences between the displayed distributions are greatest and rating them in order of importance of capturing maximum variation.
-The third contribution (Chapter \@ref(ch:gracsr)) builds upon the first two to provide methods for exploring heterogeneities in repetitive behavior for many households and over multiple granularities.
-It accomplishes this by providing a way to cluster time series based on probability distributions across informative cyclic granularities.
-Although we were motivated by the smart meter example, the problem and the solutions we propose are practically relevant to any temporal data observed more than once per year.
-
-<!-- there is a need to resort to clustering, which has been addressed in  It approaches the problem by computing pairwise distances between time series based on probability distributions over multiple cyclic granularities at once.
--->
-\newpage
-
-<!--chapter:end:Rmd//00-b-abstract.Rmd-->
-
-`r if (knitr::is_html_output()) '<!--'`
-
-\newpage
-
-# Declaration {-}
-
-I have not renumbered sections of submitted or published papers in order to generate a consistent presentation within the thesis.
-
-\textbf{Student name:} Sayani Gupta
-
-\textbf{Student signature:}
-
-```{r signature, out.width = "20%", fig.align = "left", eval = file.exists("img/signature.png")}
-knitr::include_graphics("img/signature.png")
-```
-
-\textbf{Date:} `r Sys.Date()`
-
-`r if (knitr::is_html_output()) '-->'`
-
-<!--chapter:end:Rmd//00-c-declaration.Rmd-->
-
-# Acknowledgements {-}
-
-## supervisors
-Rob and Di are phenomenal leaders in their respective fields, who lead by example.
-I am inspired by their creativity, wisdom, discipline, and dedication to really contribute to the society through their research.
-Thank you for constantly pushing me to improve as a programmer and researcher, and for regularly sharing best practices for conducting research.
-Looking back, I am grateful for how my thoughts and work in statistical computing, graphics, and data analytics in general have evolved over the years.
-I still have a lot to learn, but I am a much more self-reliant and independent researcher than I was when I started.
-Di has been instrumental in exposing me to the potential benefits of effective data visualization.
-As a female researcher, I am encouraged by her willingness to pursue unconventional avenues and have frequently noticed her make a conscious decision to question existing  stereotypes and biases.
-Thank you Di for being a fantastic female role model.
-
-## department and PhD colleagues
-
-## R community
-
-## friends
-
-PhD in Covid has been harder and I would not have been able to make it through this if not for the emotional support of my friends and family.
-Thank you Puwasala, Sium for being an epitome of kindness and acting as a pillar of support.
-Thank you Tushar, Samarpita and Nairita for always motivating me and believing in me when I didn't believe in myself.
-Thank you Ian for always putting things in perspective when I lost sight.
-Thank you each one of you for being available, listening to me, and not giving up on me at my lowest.
-Thanks to my housemates Anjali, Surbhi and Dulaji for the fun company and conversations that kept me sane in the covid lockdowns.
-
-## family
-Finally, a big thanks to my family for always being supportive of my choices.
-Thanks to my mum (Nupur Gupta) and dad (Arun Prasad Gupta), from whom I learnt that no matter where you start from, if you persevere and are sincere in your efforts, you can sail through any difficult situation.
-They are my constant cheerleaders for all little and big endeavours.
-Thanks to my brother (Avijit) for always having my back and inspiring me to dream bigger.
-Thank you, Juhi (sister-in-law) for instilling in me the value of organisation in all aspects of life and the importance of prioritizing my physical health from time to time.
-Thank you all for your part in supporting me in this journey.
-I love you.
-
-<!--chapter:end:Rmd//00-d-acknowlegements.Rmd-->
-
-# Preface {-}
-
-Chapter \@ref(ch:gravitas) has been accepted by the *Journal of Computational and Graphical Statistics*.
-It has won the ACEMS Business Analytics Prize in 2020.
-The accompanying R package **gravitas** is on CRAN.
-Chapter \@ref(ch:hakear) and Chapter \@ref(ch:gracsr) are yet to be submitted.
-
-### Open and reproducible research {-}
-
-This thesis is written in R Markdown [@rmarkdown] with **bookdown** [@bookdown], using **renv** [@renv] to create reproducible environments.
-The online version of this thesis is hosted at <https://sayani.netlify.app/>, powered by [Netlify](http://netlify.com).
-All materials (including the data sets and source files) required to reproduce this document can be found at the Github repository <https://github.com/Sayani07/thesis>.
-
-### License {-}
-
-<a rel="license" href="http://creativecommons.org/licenses/by-nc-sa/4.0/"><img alt="Creative Commons Licence" style="border-width:0" src="https://i.creativecommons.org/l/by-nc-sa/4.0/88x31.png" /></a>
-
-This work is licensed under a [Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International License](http://creativecommons.org/licenses/by-nc-sa/4.0/).
-
-The code used in this document is available under the [MIT license](https://opensource.org/licenses/MIT).
-
-\clearpage\pagenumbering{arabic}\setcounter{page}{0}
-
-<!--chapter:end:Rmd//00-e-preface.Rmd-->
-
-# Introduction {#ch:intro}
-
-Placeholder
-
-## Visualizing probability distributions across bivariate cyclic temporal granularities {#sec:gravitas}
-## Detecting distributional differences between temporal granularities for exploratory time series analysis {#sec:hakear}
-## Clustering time series based on probability distributions across temporal granularities {#sec:gracsr}
-## Summary
-
-<!--chapter:end:Rmd//01-intro.Rmd-->
-
-# Visualizing probability distributions across bivariate cyclic temporal granularities {#ch:gravitas}
-
-Placeholder
-
-## Introduction
-## Linear time granularities {#sec:linear-time}
-### Definitions
-### Relativities
-## Cyclic time granularities {#sec:cyclic-gran}
-### Circular granularities {#sec:circular-gran-def}
-### Quasi-circular granularities {#sec:quasi-circular-gran-def}
-### Aperiodic granularities {#sec:aperiodic-gran-def}
-### Relativities {#sec:cyclic-calendar}
-### Computation
-#### Single-to-multiple order-up {-}
-#### Multiple-to-single order-up {-}
-#### Multiple order-up quasi-circular granularities {-}
-## Data structure {#sec:data-structure}
-### Harmonies and clashes {#sec:synergy}
-### Near-clashes {#sec:near-clashes}
-## Visualization {#sec:visualization}
-### Data summarization
-### Display choices for univariate distributions
-### Comparison across sub-groups induced by conditioning {#sec:aesthetics}
-#### Levels {-}
-#### Synergy of cyclic granularities {-}
-#### Conditioning variable {-}
-## Applications {#sec:application}
-### Smart meter data of Australia {#sec:smartmeter}
-#### Cyclic granularities search and computation {-}
-#### Screening and visualizing harmonies {-}
-### T20 cricket data of Indian Premier League {#sec:cricket}
-## Discussion {#sec:discussion}
-## Acknowledgments {-}
-## Supplementary Materials
-
-<!--chapter:end:Rmd//02-gravitas.Rmd-->
-
-# Detecting distributional differences between temporal granularities for exploratory time series analysis  {#ch:hakear}
-
-Placeholder
-
-## Introduction
-## Proposed distance measure {#sec:computation-wpd}
-### Principle 
-### Notations
-### Computation
-#### Data transformation
-#### Characterising distributions
-#### Distance between distributions
-#### Within-facet and between-facet distances
-#### Tuning parameter
-#### Raw distance measure
-#### Adjusting for the number of comparisons {#sec:norm-wpd}
-## Ranking and selection of cyclic granularities {#sec:rank-wpd}
-### Selection
-### Ranking
-## Simulations {#sec:simulations}
-### Behavior of raw and adjusted distance measures
-### Choosing threshold
-### Environment
-## Application to residential smart meter dataset {#sec:application-wpd}
-## Discussion
-## Acknowledgments {-}
-## Supplementary Materials
-
-<!--chapter:end:Rmd//03-hakear.Rmd-->
-
-# Clustering time series based on probability distributions across temporal granularities {#ch:gracsr}
-
-Placeholder
-
-## Introduction
-## Clustering methodology {#sec:methodology}
-## Validation {#sec:validation}
-### Data generating processes
-### Data designs  
-#### Individual granularities
-#### Interaction of granularities
-### Visual exploration of findings
-## Case study {#sec:cases}
-### Clustering
-#### JS-based distances
-#### wpd-based distances
-## Conclusion and future work {#sec:conclusion}
-## Acknowledgments {-}
-## Supplementary Materials
-
-<!--chapter:end:Rmd//04-gracsr.Rmd-->
-
-# Conclusion and future plans {#ch:conclusion}
-
-Placeholder
-
-## Software development
-### gravitas
-### hakear 
-### gracsr
-## Future work
-### Putting all functionalities on CRAN
-### Scaling up the clustering method to incorporate large uncertainty and improved computational efficiency that comes with large data
-### Comparing our clustering method with other benchmark methods
-### Check generalizability of our methods
-## Final words
-
-<!--chapter:end:Rmd//05-conclusion.Rmd-->
-
 # (APPENDIX) Appendix {-}
 
-# Data dictionary
+## Data dictionary
 
@@
 (ref:wdi-data-dict) Data dictionary for the World Development Indicators.
-
-<!--chapter:end:Rmd//06-appendix.Rmd-->
-
-`r if (knitr::is_html_output()) '# Bibliography {-}'`
-
-<!--chapter:end:Rmd//99-references.Rmd-->

### 2021-11-05 Rmd/00-d-acknowlegements.Rmd
trying to make the gitbook work
putting reference even if unnumbered
commented for now

@@
-# Acknowledgements {-}
+# Acknowledgements {-#ch:thanks}
 
-supervisors
-Rob and Di are phenomenal leaders in their respective fields, who lead by example.
+<!-- - supervisors -->
+
+<!-- Rob and Di are phenomenal leaders in their respective fields, who lead by example.
 I am inspired by their creativity, wisdom, discipline, and dedication to really contribute to the society through their research.
@@
 Thank you Di for being a fantastic female role model.
-- department and PhD colleagues
+-->
 
-- R community
+<!-- - department and PhD colleagues -->
 
-- friends
+<!-- - R community -->
 
-PhD in Covid has been harder and I would not have been able to make it through this if not for the emotional support of my friends and family.
+<!-- - friends -->
+
+<!-- PhD in Covid has been harder and I would not have been able to make it through this if not for the emotional support of my friends and family.
 Thank you Puwasala, Sium for being an epitome of kindness and acting as a pillar of support.
@@
 Thanks to my housemates Anjali, Surbhi and Dulaji for the fun company and conversations that kept me sane in the covid lockdowns.
+-->
 
--  family
-Finally, a big thanks to my family for always being supportive of my choices.
+<!-- -  family -->
+<!-- Finally, a big thanks to my family for always being supportive of my choices.
 Thanks to my mum (Nupur Gupta) and dad (Arun Prasad Gupta), from whom I learnt that no matter where you start from, if you persevere and are sincere in your efforts, you can sail through any difficult situation.
@@
 Thank you all for your part in supporting me in this journey.
-I love you.
+I love you.  -->

### 2021-11-05 Rmd/05-conclusion.Rmd
conclusion first chapter done

@@
 
-In this thesis, I present a systematic approach for visualizing and analyzing large temporal data distributions.
-The building blocks of this framework are presented in this thesis.
-Chapter \@ref(ch:gravitas) includes tools for computing all cyclic granularities as well as a recommendation system for selecting pairs of granularities that may be effectively investigated together.
-These temporal granularities may be used to generate data visualizations to search for patterns, associations, and anomalies.
-However, when there are many granularities that can be constructed for a time period, there will also be too many possible displays to decide which might be the more interesting to display.
-Chapter \@ref(ch:hakear) presents a framework for selecting displays that are interesting, with the greatest differences between the displayed distributions, and ranking them in order of priority for capturing the most variation.
-Both of these are used for studying patterns in individual time series or comparing a few time series together.
-In Chapter \@ref(ch:gracsr), it is extended to allow for the exploration of distributions for many time series at the same time by clustering them based on probability distributions across informative cyclic granularities.
-Through the use of probability distribution, this technique is more comprehensive in recognizing clusters with recurring patterns over many important granularities and more resilient to noisy, patchy, and uneven length time series.
+In this thesis, I present a systematic approach for visualizing and analyzing distributions of large temporal data by leveraging time characteristics.
+The building blocks of this framework are presented in each chapter.
+Chapter \@ref(ch:gravitas) contains tools for computing all cyclic granularities, as well as a recommendation system for selecting pairs of granularities that can be effectively analyzed together.
+These temporal granularities can be used to generate data visualizations to look for patterns, associations, and anomalies.
+But it is difficult to decide which of the various granularities to display when there are many options.
+Chapter \@ref(ch:hakear) presents a methodology for selecting displays that are interesting, such that differences between displayed distributions are greatest.
+The methodology also ranks the displays in order of priority for capturing the most variation.
+Both \@ref(ch:gravitas) and \@ref(ch:hakear) are used for studying patterns in individual time series or comparing a few time series together.
+This is extended in Chapter @ref(ch:gracsr) to allow for the exploration of distributions for multiple time series at the same time using unsupervised clustering.
+The clustering methodology produces small groups of time series with similar distributions over multiple granularities.
+This technique is more robust to noisy, patchy, and uneven length time series because it makes use of probability distributions.
+
+<!-- is more comprehensive in recognizing clusters with recurring patterns over many important granularities and -->
 

### 2021-11-05 Rmd/05-conclusion.Rmd
software development section

@@
 
-This thesis focuses on integrating research approaches into open source R packages such as **gravitas**, **hakear**, and **gracsr**.
+This thesis focuses on integrating research approaches into open source R packages for reproducibility and ease of use in other applications.
+So a significant amount of work has been devoted to the development of R packages  **gravitas**, **hakear**, and **gracsr**, each of which correspond to a chapter presented in this thesis.
 
@@
 
-The **gravitas** package provides very general tools to compute and manipulate cyclic granularities, and to generate plots displaying distributions conditional on those granularities.
-It can be utilized in non-temporal cases for which a hierarchical structure can be construed similar to time.
-It is on CRAN.
+The **gravitas** package provides very general tools to compute and manipulate cyclic granularities and generate plots displaying distributions conditional on those granularities.
+The functions `search_gran()`, `create_gran()`, `harmony()`, `gran_advice()` and `prob_plot()` provides a workflow for an analyst to systematically explore large quantities of temporal data across different harmonies (pairs of granularities that can be analyzed together.) This package was developed as part of my internship at Google Summer of Code, 2019.
+It and has been on CRAN since January 2020.
 The website (<https://sayani07.github.io/gravitas>) includes full documentation and two vignettes about the package usage.
-There has been a grand total of 12K downloads from the RStudio mirror dating from 2020-11-01 to 2021-11-01.
+There has been a total of 12K downloads from the RStudio mirror dating from 2020-11-01 to 2021-11-01.
+This package supplements the paper corresponding to \@ref(ch:gravitas), which has won the ACEMS Business Analytics Award 2021.
+The package can be generalized to non-temporal applications for which a hierarchical structure can be construed similar to time.
 
@@
 
-The open-source R package `hakear` is available on Github (https://github.com/Sayani07/hakear) to implement ideas presented in Chapter \@ref(ch:hakear).
-Given a `tsibble` and context granularities of interest, the function `wpd()` provides support for computing the wpd for each cyclic granularities or pair of granularities and `select_harmonies()` chooses the ones with significant patterns and ranks them from highest to lowest wpd.
+The R package hakear (https://github.com/Sayani07/hakear) implements ideas presented in Chapter @ref (ch:hakear).
+The function wpd() computes the weighted pairwise distances ($wpd$) for each cyclic granularity or pair of granularities, and select harmonies() selects the ones with significant patterns and ranks them from highest to lowest $wpd$.
+These selected harmonies can be plotted using package `gravitas` for potentially interesting displays.
+Because $wpd$ is computationally expensive, this package uses the R package `parallel` to handle parallel computations.
 
@@
 The open-source R package `gracsr` is available on Github (https://github.com/Sayani07/gracsr) to implement ideas presented in Chapter \@ref(ch:gracsr).
-The package provides functions to carry out the entire clustering methodology discussed in the paper.
-It is still a work in progress and has won the ACEMS Business Analytics Prize 2021 with a prize money of AUD 3000, which would be utilized in polishing this package and preparing it for CRAN.
+The package contains functions for implementing the complete clustering methodology with choices of scaling and distance metrics discussed in the paper.
+It has received an AUD 3000 grant as part of the ACEMS Business Analytics Prize towards polishing the functions and preparing it for CRAN.
 

### 2021-11-05 Rmd/00-d-acknowlegements.Rmd
acknowledgement done

@@
 
-<!-- Rob and Di are phenomenal leaders in their respective fields, who lead by example.
-I am inspired by their creativity, wisdom, discipline, and dedication to really contribute to the society through their research.
-Thank you for constantly pushing me to improve as a programmer and researcher, and for regularly sharing best practices for conducting research.
-Looking back, I am grateful for how my thoughts and work in statistical computing, graphics, and data analytics in general have evolved over the years.
-I still have a lot to learn, but I am a much more self-reliant and independent researcher than I was when I started.
+Rob and Di are phenomenal leaders in their respective fields who lead by example.
+I am inspired by their creativity, wisdom, discipline, and dedication to contribute to the society through their research.
+Thank you for constantly pushing me to improve as a programmer and researcher and regularly sharing good practices for conducting research.
+Looking back, I am grateful for how my thoughts and work in statistical computing, graphics, and data analytics, in general, have evolved over the years.
+I still have a lot to learn, but I am a much more self-reliant and independent researcher than when I started.
 Di has been instrumental in exposing me to the potential benefits of effective data visualization.
-As a female researcher, I am encouraged by her willingness to pursue unconventional avenues and have frequently noticed her make a conscious decision to question existing  stereotypes and biases.
-Thank you Di for being a fantastic female role model.
--->
+As a female researcher, I am encouraged by her willingness to pursue unconventional avenues and have frequently noticed her make conscious and proactive choices to question existing stereotypes and biases.
+Thank you, Di, for being a fantastic female role model.
 
-<!-- - department and PhD colleagues -->
+<!-- - Department and conferences -->
 
-<!-- - R community -->
+I want to thank the Department of Econometrics and Business Statistics, Monash Business School, ARC Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS), and Data61 CSIRO for their financial support.
+Thanks to the Department, I attended a few conferences in person (before COVID hit) and had a great time networking and being inspired by many people.
+A few highlights are UseR!
+2018 (Brisbane), Young Statistician's Conference 2019 (Canberra), WOMBATS 2019 (Melbourne), and RopenSci 2019 (Sydney).
+Thank you to Gael Martin, Farshid Wahid, Klaus Ackermann, and Didier Nibbering from my PhD committee for providing such a supportive environment.
 
-<!-- - friends -->
+ <!-- - PhD friends and R community -->
+Thank you, Stephanie and Nicholas Tierney, for being so warm, kinds and welcoming when I began my PhD.
+Being a part of such an inclusive environment meant a lot to me as an international student.
+Thank you, Nicholas Spyrison and Mitchell O'Hara-Wild, for the numerous informative discussions, brainstorming sessions, as well as for generously sharing your wisdom and experience in software development.
+Thank you to each of you, NUMBATs, and the broader R community for directly and indirectly helping me countless times.
+You never fail to inspire me.
 
-<!-- PhD in Covid has been harder and I would not have been able to make it through this if not for the emotional support of my friends and family.
-Thank you Puwasala, Sium for being an epitome of kindness and acting as a pillar of support.
-Thank you Tushar, Samarpita and Nairita for always motivating me and believing in me when I didn't believe in myself.
-Thank you Ian for always putting things in perspective when I lost sight.
-Thank you each one of you for being available, listening to me, and not giving up on me at my lowest.
-Thanks to my housemates Anjali, Surbhi and Dulaji for the fun company and conversations that kept me sane in the covid lockdowns.
--->
+<!-- - Friends -->
 
-<!-- -  family -->
-<!-- Finally, a big thanks to my family for always being supportive of my choices.
+PhD in Covid has been harder, and I would not have been able to make it through this if not for the emotional support of my friends and family.
+Thank you Puwasala, Sium, and Ian, for being so kind and acting as a pillar of emotional support, especially during the last leg of my journey.
+Thank you Tushar, Samarpita and Nairita for always motivating me and believing in me when I did not believe in myself.
+Thank you to each one of you for being available, listening to me, and putting things in perspective when I lost sight.
+Thanks to my housemates Anjali, Surbhi, and Dulaji for the fun company and conversations that kept me sane in the covid lockdowns.
+
+<!-- - Family -->
+
+Finally, a big thank you to my family for always being supportive of my choices.
 Thanks to my mum (Nupur Gupta) and dad (Arun Prasad Gupta), from whom I learnt that no matter where you start from, if you persevere and are sincere in your efforts, you can sail through any difficult situation.
-They are my constant cheerleaders for all little and big endeavours.
+They are my constant cheerleaders for all little and big endeavors.
 Thanks to my brother (Avijit) for always having my back and inspiring me to dream bigger.
-Thank you, Juhi (sister-in-law) for instilling in me the value of organisation in all aspects of life and the importance of prioritizing my physical health from time to time.
+Thank you, Juhi (sister-in-law), for reminding me of the value of organization in all aspects of life and the importance of prioritizing my physical health from time to time.
 Thank you all for your part in supporting me in this journey.
-I love you.  -->
+I love you.
+
+<!-- add people who are helping you with proofreading -->

### 2021-11-06 Rmd/00-b-abstract.Rmd
putting reference even if unnumbered

@@
-# Abstract {-}
+# Abstract {- #ch:abstract}
 
-The research was motivated by the desire to understand the Australian smart meters data, which was collected half-hourly for two years at the household level.
-With temporal data available at ever finer scales, exploring periodicity can become overwhelming with so many possible temporal deconstructions to explore.
+The research was motivated by the desire to understand the Australian smart meter data, which was collected half-hourly for two years at the household level.
+With temporal data available at ever finer scales,
+exploration may be required to be carried out across both finer and coarser scales to draw useful inferences about the underlying process.
 Analysts are expected to comprehensively explore the many ways to view and consider temporal data.
-However, the plethora of choices and the lack of a systematic approach to do so quickly can make the task overwhelming.
-
-<!-- exploration may be required to be carried out across both finer and coarser scales to draw useful inferences about the underlying process.
-It is important to be able to navigate through all the temporal deconstructions that accommodate for periodicities to have multiple perspectives of the observed data.
-This idea aligns with the notion of EDA (Tukey 1977) which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
-Slicing and dicing the data in all possible temporal scales as suggested by EDA can be a daunting task as it leads to a -->
-<!-- myriad of possibilities. -->
-
-<!-- Analysts are expected to comprehensively explore the many ways to view and consider temporal data.
-However, the plethora of choices and the lack of a systematic approach to do so quickly can make the task overwhelming -->
-
-<!-- exploring periodicity can become overwhelming with so many possible temporal deconstructions to explore.
-Analysts are expected to comprehensively explore the many ways to view and consider temporal data.
-However, the plethora of choices and the lack of a systematic approach to do so quickly can make the task overwhelming.
--->
+However, the plethora of choices and the lack of a systematic approach to do so quickly can make the task daunting.
 
@@
 It provides tools to compute possible cyclic granularities from an ordered (usually temporal) index and also a framework to systematically explore the distribution of a univariate variable conditional on two cyclic time granularities by defining "harmony".
-"harmony" denotes pairs of granularities that could be analyzed together and reduces the search from all possible options.
+"harmony" denotes pairs of granularities that could be effectively analyzed together and reduces the search from all possible options.
 This approach is still overwhelming for human consumption because there would still be a huge number of harmonies.
-Hence, the second contribution (Chapter \@ref(ch:hakear)) refines the search of informative granularities by identifying those for which the differences between the displayed distributions are greatest and rating them in order of importance of capturing maximum variation.
-The third contribution (Chapter \@ref(ch:gracsr)) builds upon the first two to provide methods for exploring heterogeneities in repetitive behavior for many households and over multiple granularities.
+The second contribution (Chapter \@ref(ch:hakear)) refines the search for informative granularities by identifying those for which the differences between the displayed distributions are greatest and also rating them in order of importance of capturing maximum variation.
+The third contribution (Chapter \@ref(ch:gracsr)) builds upon the first two to provide methods for exploring heterogeneities in repetitive behavior for many time series and over multiple granularities.
 It accomplishes this by providing a way to cluster time series based on probability distributions across informative cyclic granularities.
@@
 
-<!-- there is a need to resort to clustering, which has been addressed in  It approaches the problem by computing pairwise distances between time series based on probability distributions over multiple cyclic granularities at once.
--->
 \newpage

### 2021-11-06 Rmd/00-d-acknowlegements.Rmd

@@
  <!-- - PhD friends and R community -->
-Thank you, Stephanie and Nicholas Tierney, for being so warm, kinds and welcoming when I began my PhD.
+Thank you, Stephanie and Nicholas Tierney, for being so warm, kind and welcoming when I began my PhD.
 Being a part of such an inclusive environment meant a lot to me as an international student.
-Thank you, Nicholas Spyrison and Mitchell O'Hara-Wild, for the numerous informative discussions, brainstorming sessions, as well as for generously sharing your wisdom and experience in software development.
+Thank you, Nicholas Spyrison and Mitchell O'Hara-Wild, for the numerous informative discussions, brainstorming sessions, as well as for generously sharing your expertise and experience in software development.
 Thank you to each of you, NUMBATs, and the broader R community for directly and indirectly helping me countless times.
@@
 <!-- add people who are helping you with proofreading -->
+<!-- Add Peter -->
+<!-- Although the journey is coming to an end, you wish to take these life learnings forward -->
+<!-- add Earo and Stuart that you followed their model of workflow and templates, using the power of reproducible and accessible  research -->

### 2021-11-06 Rmd/05-conclusion.Rmd

@@
 The **gravitas** package provides very general tools to compute and manipulate cyclic granularities and generate plots displaying distributions conditional on those granularities.
-The functions `search_gran()`, `create_gran()`, `harmony()`, `gran_advice()` and `prob_plot()` provides a workflow for an analyst to systematically explore large quantities of temporal data across different harmonies (pairs of granularities that can be analyzed together.) This package was developed as part of my internship at Google Summer of Code, 2019.
-It and has been on CRAN since January 2020.
+The functions `search_gran()`, `create_gran()`, `harmony()`, `gran_advice()` and `prob_plot()` provides the entire workflow for an analyst to systematically explore large quantities of temporal data across different harmonies (pairs of granularities that can be analyzed together.) This package was developed as part of my internship at Google Summer of Code, 2019.
+It has been on CRAN since January 2020.
 The website (<https://sayani07.github.io/gravitas>) includes full documentation and two vignettes about the package usage.
 There has been a total of 12K downloads from the RStudio mirror dating from 2020-11-01 to 2021-11-01.
-This package supplements the paper corresponding to \@ref(ch:gravitas), which has won the ACEMS Business Analytics Award 2021.
+This package supplements the paper corresponding to Chapter \@ref(ch:gravitas), which has won the ACEMS Business Analytics Award 2021.
 The package can be generalized to non-temporal applications for which a hierarchical structure can be construed similar to time.
@@
 
-The R package hakear (https://github.com/Sayani07/hakear) implements ideas presented in Chapter @ref (ch:hakear).
-The function wpd() computes the weighted pairwise distances ($wpd$) for each cyclic granularity or pair of granularities, and select harmonies() selects the ones with significant patterns and ranks them from highest to lowest $wpd$.
+The R package hakear (https://github.com/Sayani07/hakear) implements ideas presented in Chapter \@ref(ch:hakear).
+The function wpd() computes the weighted pairwise distances ($wpd$) for each cyclic granularity or pair of granularities, and `select_harmonies()` selects the ones with significant patterns and ranks them from highest to lowest $wpd$.
 These selected harmonies can be plotted using package `gravitas` for potentially interesting displays.
-Because $wpd$ is computationally expensive, this package uses the R package `parallel` to handle parallel computations.
+This package is reliant on parallel processing using multiple multi-core computers for faster computation of $wpd$.
 
@@
 The open-source R package `gracsr` is available on Github (https://github.com/Sayani07/gracsr) to implement ideas presented in Chapter \@ref(ch:gracsr).
-The package contains functions for implementing the complete clustering methodology with choices of scaling and distance metrics discussed in the paper.
-It has received an AUD 3000 grant as part of the ACEMS Business Analytics Prize towards polishing the functions and preparing it for CRAN.
+The package contains functions for implementing the complete clustering methodology with choices of scaling and distance metrics discussed in the chapter.
+It has received a grant (AUD 3000) as part of the ACEMS Business Analytics Prize towards polishing the functions and preparing it for CRAN.
 
@@
 
-I plan to integrate `hakear` with `gravitas` as one R package to systematically exploring few time series, whereas `gracsr` would provide the clustering framework for exploring many time series together.
-I plan to run it through https://ropensci.org/software-review/ to develop them further for more visibility and efficient usage.
+I want to combine 'hakear' and 'gravitas' into a single R package for systematically exploring distributions of a smaller number of time series, whilst 'gracsr' would offer the clustering framework for systematically exploring a large number of time series.
+I intend to run it via https://ropensci.org/software-review/ to further refine them for more exposure and efficiency.
+
+### Scaling up the methods for multivariate time series data
+
+With multiple time series available for each observation, the complexity of efficient exploration and visualization grows exponentially.
+In this case, conditional distributions include not only temporal dependency but also variables and their dependencies.
+This adds to the already high-dimensional data structures that result from studying distributions.
+Current research on visualizing multivariate data must account for time's inherent characteristics in order to efficiently explore distributions of multivariate time series data.
+
+### Comparing our clustering method with other benchmark methods
+
+Testing needs to be carried out with non-hierarchy based clustering methods to see whether the proposed distances perform better with other algorithms.
+It would also be helpful to know how much more value we obtain from considering distributions rather than aggregating time series data.
+In the case of smart meters, this could be accomplished by comparing the results while validating them against external data such as weather conditions, socioeconomic or other demographic factors of those households.
 
@@
 
-With the volume of data projected to grow in the future, potentially leading to increased variability in patterns, research is needed to understand the issues that may arise with the existing methodology and how to adapt our current algorithm.
-Computational efficiency is also critical when scaling up for analysis of huge data sets, let alone when adding features to existing highly dimensional data structures.
+With the volume of data expected to increase in the future, potentially leading to increased variability in patterns, more experiments are needed to understand what sort of problems may arise with our current methodology and how to improve our current algorithm to address those issues.
+All of this must be done while keeping computational efficiency in mind, as this is critical when scaling up for analysis of large data sets.
 
-### Comparing our clustering method with other benchmark methods
+### Develop inferential methods for the clusters
 
-Testing needs to be carried out with non-hierarchy based clustering methods to see whether these distances perform better with other algorithms.
-Also, to evaluate how much information is lost when aggregating individual level demand versus distributions.
+It is crucial to determine whether or not any detected clusters or patterns are statistically significant.
+In Chapter \@ref(ch:gracsr), we have used the permutation approach to detect if patterns are authentic or not across one or a pair of granularities.
+A similar method for finding significance across clusters is not appropriate if the observations are assumed to be independent.
+Most large temporal data sets gathered at fine temporal scales usually include several repetitive patterns.
+Therefore, the permutations used for comparisons must account that different temporal granularities have distinct conditional distributions.
 
@@
 
-We provide solutions that are realistically applicable to any temporal data observed more than once per year.
-We could just verify its value in the electricity smart meter context.
-With numerous open-source benchmark data sets accessible, it is necessary to test how well the approaches operate in various disciplines.
-
-## Final words
+We propose solutions that are realistically applicable to any temporal data collected more than once per year and may be implemented in a variety of settings.
+We could simply verify its utility in the context of smart meter electricity data.
+With the availability of several open-source benchmark data sets, it is vital to evaluate how well the methodologies perform across a variety of fields.

### 2021-11-06 Rmd/01-intro.Rmd
describing gracsr in intro

@@
 
+ <!-- This chapter contains a literature review related to linear granularities. -->
+
 ## Detecting distributional differences between temporal granularities for exploratory time series analysis {#sec:hakear}
 
-Cyclic temporal granularities, which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can be useful for measuring repetitive patterns in large univariate time series data.
-The granularities feed new approaches to exploring time series data.
-One use is to take pairs of granularities, and make plots of response values across the categories induced by the temporal deconstruction.
-However, when there are many granularities that can be constructed for a time period, there will also be too many possible displays to decide which might be the more interesting to display.
-This work proposes a new distance metric to screen and rank the possible granularities, and hence choose the most interesting ones to plot.
-The distance measure is computed for a single or pairs of cyclic granularities can can be compared across different cyclic granularities and also on a collection of time series.
-The methods are implemented in the open-source R package `hakear`.
+This work is a natural extension of Chapter \@ref(ch:gravitas).
+Many, many other displays might be built using different granularities, such as day-of-week, day-of-month, weekday/weekend, and so on.
+However, only a handful of them may be interesting, revealing major patterns.
+Determining which displays exhibit "significant" distributional differences between cyclic granularity categories and plotting only these would allow for more efficient exploration.
+Furthermore, a few of the displays in this collection will be more engaging than others.
+This chapter provides a new distance metric for selecting and ranking the multiple granularities.
+The statistical significance of potential visual discoveries is aided by selecting a threshold for the proposed numerical distance measure.
+The distance measure is computed for a single or pairs of cyclic granularities, and it can be compared across different cyclic granularities as well as a collection of time series.
+This chapter also includes a case study using residential smart meter data from Melbourne to demonstrate how the suggested methodology may be utilised to automatically find temporal granularities with significant distributional differences.
+The methods are implemented in the open-source R package 'hakear.'
 
@@
 
- With more and more time series data being collected at much finer temporal resolution, for a longer length of time, and for a larger number of individuals/entities, time series clustering research is getting a lot of traction.
-The sort of noisy, patchy, uneven, and asynchronous time series that is typical in many disciplines limits similarity searches among these lengthy time series.
-In this work, we suggest a method for overcoming these constraints by grouping time series based on probability distributions over cyclic temporal granularities.
-Cyclic granularities are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, and so on, and can be helpful for detecting repeating patterns.
-Looking at probability distributions across cyclic granularities results in an approach that is robust to missing or noisy data, aids in dimension reduction, and ensures small pockets of similar repeated behaviours.
-The proposed method was tested using a collection of residential electricity customers.
-The simulated and empirical evidence demonstrates that our method is capable of producing meaningful clusters.
+In this chapter, we look at the problem of using clustering to discover patterns in a large number of univariate time series across multiple temporal granularities.
+Time series clustering research is gaining traction as more data is collected at finer temporal resolution, over longer time periods, and for a larger number of individuals/entities.
+Many disciplines have noisy, patchy, uneven, and asynchronous time series that make it difficult to search for similarities.
+We propose a method for overcoming these constraints by calculating distances between time series based on probability distributions at various temporal granularities.
+Because they are based on probability distributions, these distances are resistant to missing or noisy data and aid in dimension reduction.
+When fed into a clustering algorithm, the distances can be used to divide large data sets into small pockets of similar repetitive behaviour.
+These subgroups can then be analysed separately or used as distinct prototype behaviours in classification problems.
+The proposed method was tested on a group of residential electricity consumers from the Australian smart meter data set to show that it can generate meaningful clusters.
+This chapter includes a brief review of the literature on traditional time series clustering and, more specifically, clustering residential smart meter data.
 

### 2021-11-06 Rmd/01-intro.Rmd
introduction - intro

@@
 
+<!-- outline the intro chapter structure -->
+
+This chapter provides an introduction to the study by first discussing the background and context, followed by the overarching research aim and objectives, each of which correspond to each chapter in this thesis.
+
+<!-- Background -->
+
+The Smart Grid, Smart City (SGSC) project (2010–2014) available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) provides half-hourly data of over 13,000 household electricity smart meters distributed unevenly from October 2011 to March 2014 for some parts in Australia.
+Because of the increasing variety of customers, larger data sets contain more uncertainty about customer behavior.
+Household behavioral patterns vary significantly due to differences in size, location, and amenities such as solar panels, central heating, and air conditioning.
+For example, some families use a dryer, while others hang their clothes to dry.
+This could be reflected in their weekly profile.
+They may vary on a monthly basis, with some customers using more air conditioners or heaters than others despite having comparable electrical equipment and weather conditions.
+Some customers are night owls, while others are morning larks, which may show up in their daily profile.
+Customers' day-off energy consumption varies depending on whether they stay at home or go outside.
+The motivation for this work comes from the desire to provide methods to better understand heterogeneity in repetitive behavior that can arise across temporal dependencies.
+
 <!-- introducing the overall field-->
@@
 This approach requires deconstructing time in various possible ways.
-Moreover, often it might be interesting to capture calendar or periodic effects like month-of-year, day-of-week or hour-of-day.
-They help us in answering questions like if certain levels of those time deconstructions are characterized by unusual/routine values of the observed variable.
+Moreover, it might often be interesting to capture calendar or periodic effects like month-of-year, day-of-week, or hour-of-day.
+They help us answer questions like if certain levels of those time deconstructions are characterized by unusual/typical values of the observed variable.
 For example, certain days of the week or months of the year are likely to be characterized by higher values.
-It is important to be able to navigate through all the temporal deconstructions that accommodate for periodicities to have multiple perspectives of the observed data.
-This idea aligns with the notion of EDA (Tukey 1977) which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
-
-<!-- outline the intro chapter structure -->
-
-This chapter will provide an introduction to the study by first discussing the background and context, followed by the research aims, objectives and questions.
-
-<!-- Background -->
-
-The motivation for this work comes from the desire to provide methods to better understand large quantities of measurements on energy usage reported by smart meters in household across Australia, and indeed many parts of the world.
-Smart meters currently provide half-hourly use in kwh for each household, from the time that they were installed, some as early as 2012.
-Households are distributed geographically, and have different demographic properties such as the existence of solar panels, central heating or air conditioning.
-The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.
-It is common to see aggregates of usage across households, total kwh used each half hour by state, for example, because energy companies need to understand maximum loads that they will have to plan ahead to accommodate.
-But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
+Visualizations using linear time granularities are often ineffective in conveying repetitive behaviors because multiple observations are squeezed together in that representation.
+Hence, we define cyclic granularities for exploring repetitive patterns.
+There could potentially be many, many cyclic granularities to explore.
+It is essential to navigate through all the options to get a comprehensive view of the repeated patterns.
+This idea aligns with the notion of EDA (Tukey 1977), which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
 
 <!-- introduce the specific problem  -->
-
 <!-- ## change in data structure as a result of transitioning from linear to cyclic deconstructions -->
 <!-- + Existing approaches, why we need to drill down, or why we want to -- probability distributions -->
-However, restructuring time in this manner leads to restructured data where each level of the time deconstructions correspond to multiple values of the observed variable.
-It is common to see aggregation or summarization of these multiple observations with a unique value to study calendar effects.
-For example, using aggregates of usage across each hour/half-hour has been common in the literature because energy companies need to plan for maximum loads on the network.
-But studying overall energy use hides the distributions of usage at finer scales, and making it more difficult to find solutions to improve energy efficiency.
-Summarizing the probability distribution of these multiple observations to capture both the shape and uncertainty could be a potential way to understand the underlying distribution of these observations.
-Studying probability distributions is likely to focus on features of the data which are not transparent through raw data or a unique summary statistic.
+
+Slicing and dicing data in all possible temporal scales, as suggested by EDA, can be a daunting task because it opens up a plethora of possibilities.
+This motivates the research presented in this thesis, which aims to provide a platform for systematically exploring periodicities in temporal data and supporting the discovery of regular patterns or anomalies, as well as the exploration of clusters of behaviors or the summarization of the behavior.
+The transition from linear to cyclic granularities results in restructured data, with each level of the time deconstructions corresponding to multiple values of the observed variable.
+For example, it is common to see aggregated usage across households, such as total kWh used each half-hour by state, because energy companies need to know the maximum loads they must plan for.
+However, studying overall energy consumption obscures usage distributions at finer scales, making it more difficult to find solutions to improve energy efficiency.
+Summarizing the probability distribution of these multiple observations to capture both shape and uncertainty could be a potential approach to understanding the underlying distribution of these observations.
+The research objectives presented in each chapter aid in the achievement of this goal.
+
+<!-- Hence, the overarching research goal is to study the periodic behavior of temporal data in a structured way by studying the probability distributions by best exploiting the characteristics of time.
+-->
 
@@
 
-Hence, the overarching research goal is to study the periodic behavior of temporal data in a structured way by studying the probability distributions by best exploiting the characteristics of time.
-Slicing and dicing the data in all possible temporal scales as suggested by EDA can be a daunting task as it leads to a myriad of possibilities.
-This inspires the research presented in this thesis, which aims to provide a platform to systematically explore periodicities in temporal data and support finding regular patterns or anomalies, explore clusters of behaviors or summarize the behavior.
-The first part of the work discusses computation of all possible combinations of cyclic time granularities and a graphical mapping such that distributions of a numeric response variable is displayed across combinations of two cyclic granularities.
+<!-- The first part of the work discusses computation of all possible combinations of cyclic time granularities and a graphical mapping such that distributions of a numeric response variable is displayed across combinations of two cyclic granularities.
 Even analyzing the distribution of the measured variable across two cyclic granularities at once could amount to displaying many plots in search of potential  patterns.
@@
 Although the motivation came through the smart meter example, this is a problem that is relevant to any temporal data observed more than once per year.
+-->
 
@@
 The distance measure is computed for a single or pairs of cyclic granularities, and it can be compared across different cyclic granularities as well as a collection of time series.
-This chapter also includes a case study using residential smart meter data from Melbourne to demonstrate how the suggested methodology may be utilised to automatically find temporal granularities with significant distributional differences.
+This chapter also includes a case study using residential smart meter data from Melbourne to demonstrate how the suggested methodology may be utilized to automatically find temporal granularities with significant distributional differences.
 The methods are implemented in the open-source R package 'hakear.'
@@
 Because they are based on probability distributions, these distances are resistant to missing or noisy data and aid in dimension reduction.
-When fed into a clustering algorithm, the distances can be used to divide large data sets into small pockets of similar repetitive behaviour.
-These subgroups can then be analysed separately or used as distinct prototype behaviours in classification problems.
+When fed into a clustering algorithm, the distances can be used to divide large data sets into small pockets of similar repetitive behavior.
+These subgroups can then be analyzed separately or used as distinct prototype behaviors in classification problems.
 The proposed method was tested on a group of residential electricity consumers from the Australian smart meter data set to show that it can generate meaningful clusters.

### 2021-11-06 Rmd/00-c-declaration.Rmd
ensuring every chapter has #
putting reference even if unnumbered
adding declaration

--- 2021-11-04 Rmd/00-c-declaration.Rmd
@@
 
-# Declaration {-}
+# Declaration {-#ch:declare}
+
+I hereby declare that this thesis contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.
+
+Three publications are included in this thesis, one of which has been accepted for publication in the Journal of Computational and Graphical Statistics (@Gupta2021-gravitas).
+Two others have not yet been submitted.
+"Exploring probability distributions across cyclic temporal granularities" is the central theme of the thesis.
+All of the papers in the thesis were conceptualized, developed, and written by myself, the student, while working in the Department of Econometrics and Business Statistics under the supervision of Professor Rob J Hyndman and Professor Dianne Cook.
+When co-authors are listed, it indicates that the work was produced through active collaboration between researchers and that their contributions to team-based research have been recognized.
+The following table details the work, including
+my and my fellow co-authors contributions:
+
+\begin{table}
+\centering\footnotesize\tabcolsep=0.12cm
+\begin{tabular}{|p{1cm}|p{2cm}|p{1.5cm}|p{3.5cm}|p{3.5cm}|p{1.5cm}|}
+\hline
+\RaggedRight\textbf{Thesis Chapter}  & \RaggedRight\textbf{Publication Title}  & \RaggedRight\textbf{Status (published, in press, accepted or returned for revision)}  & \RaggedRight\textbf{Nature and} {\%} \RaggedRight\textbf{of student contribution} & \RaggedRight\textbf{Co-author name(s) Nature and} {\%} \RaggedRight\textbf{of Co-author’s contribution} & \RaggedRight\textbf{Co-author(s), Monash student Y/N} \\ \hline
+2 & Visualizing probability distributions across bivariate cyclic temporal granularities & Published & 70\%.
+Concept and developing software and writing first draft & (1) Rob J Hyndman, input into manuscript 15\% (2) Dianne Cook, Concept and input into manuscript 15\% & \multicolumn{1}{c|}{N} \\
+\hline
+3 & Detecting distributional differences between temporal granularities
+for exploratory time series analysis & To be submitted & 80\%.
+Concept and developing software and writing first draft & (1) Rob J Hyndman, Concept and input into manuscript 10\% (2) Dianne Cook, Concept and input into manuscript 10\%  & \multicolumn{1}{c|}{N} \\
+\hline
+4 & Clustering time series based on probability distributions across temporal granularities & To be submitted & 70\%.
+Concept and developing software and writing first draft & (1) Dianne Cook, Concept and input into manuscript 20\% (2) Rob J Hyndman, concept and input into manuscript 10\% & \multicolumn{1}{c|}{N} \\
+\hline
+\end{tabular}
+\end{table}
+
+\FloatBarrier
 

### 2021-11-06 thesis.Rmd
adding declaration

--- 2021-11-04 thesis.Rmd
@@
 
-# Copyright notice {-}
+# Copyright notice {- #ch:copyright}
 
@@
 
-# Abstract {-}
-
-The research was motivated by the desire to understand the Australian smart meters data, which was collected half-hourly for two years at the household level.
-With temporal data available at ever finer scales, exploring periodicity can become overwhelming with so many possible temporal deconstructions to explore.
+# Abstract {- #ch:abstract}
+
+The research was motivated by the desire to understand the Australian smart meter data, which was collected half-hourly for two years at the household level.
+With temporal data available at ever finer scales,
+exploration may be required to be carried out across both finer and coarser scales to draw useful inferences about the underlying process.
 Analysts are expected to comprehensively explore the many ways to view and consider temporal data.
-However, the plethora of choices and the lack of a systematic approach to do so quickly can make the task overwhelming.
-
-<!-- exploration may be required to be carried out across both finer and coarser scales to draw useful inferences about the underlying process.
-It is important to be able to navigate through all the temporal deconstructions that accommodate for periodicities to have multiple perspectives of the observed data.
-This idea aligns with the notion of EDA (Tukey 1977) which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
-Slicing and dicing the data in all possible temporal scales as suggested by EDA can be a daunting task as it leads to a -->
-<!-- myriad of possibilities. -->
-
-<!-- Analysts are expected to comprehensively explore the many ways to view and consider temporal data.
-However, the plethora of choices and the lack of a systematic approach to do so quickly can make the task overwhelming -->
-
-<!-- exploring periodicity can become overwhelming with so many possible temporal deconstructions to explore.
-Analysts are expected to comprehensively explore the many ways to view and consider temporal data.
-However, the plethora of choices and the lack of a systematic approach to do so quickly can make the task overwhelming.
--->
+However, the plethora of choices and the lack of a systematic approach to do so quickly can make the task daunting.
 
@@
 It provides tools to compute possible cyclic granularities from an ordered (usually temporal) index and also a framework to systematically explore the distribution of a univariate variable conditional on two cyclic time granularities by defining "harmony".
-"harmony" denotes pairs of granularities that could be analyzed together and reduces the search from all possible options.
-This approach is still overwhelming for human consumption because there would still be a huge number of harmonies.
-Hence, the second contribution (Chapter \@ref(ch:hakear)) refines the search of informative granularities by identifying those for which the differences between the displayed distributions are greatest and rating them in order of importance of capturing maximum variation.
-The third contribution (Chapter \@ref(ch:gracsr)) builds upon the first two to provide methods for exploring heterogeneities in repetitive behavior for many households and over multiple granularities.
+"harmony" denotes pairs of granularities that could be effectively analyzed together and reduces the search from all possible options.
+This approach is still overwhelming for human consumption due of the vast amount of harmonies remaining.
+The second contribution (Chapter \@ref(ch:hakear)) refines the search for informative granularities by identifying those for which the differences between the displayed distributions are greatest and also rating them in order of importance of capturing maximum variation.
+The third contribution (Chapter \@ref(ch:gracsr)) builds upon the first two to provide methods for exploring heterogeneities in repetitive behavior for many time series and over multiple granularities.
 It accomplishes this by providing a way to cluster time series based on probability distributions across informative cyclic granularities.
@@
 
-# Declaration {-}
+# Declaration {-#ch:declare}
+
+I hereby declare that this thesis contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.
+
+Three publications are included in this thesis, one of which has been accepted for publication in the Journal of Computational and Graphical Statistics (@Gupta2021-gravitas).
+Two others have not yet been submitted.
+"Exploring probability distributions across cyclic temporal granularities" is the central theme of the thesis.
+All of the papers in the thesis were conceptualized, developed, and written by myself, the student, while working in the Department of Econometrics and Business Statistics under the supervision of Professor Rob J Hyndman and Professor Dianne Cook.
+When co-authors are listed, it indicates that the work was produced through active collaboration between researchers and that their contributions to team-based research have been recognized.
+The following table details the work, including
+my and my fellow co-authors contributions:
+
+\begin{table}
+\centering\footnotesize\tabcolsep=0.12cm
+\begin{tabular}{|p{1cm}|p{2cm}|p{1.5cm}|p{3.5cm}|p{3.5cm}|p{1.5cm}|}
+\hline
+\RaggedRight\textbf{Thesis Chapter}  & \RaggedRight\textbf{Publication Title}  & \RaggedRight\textbf{Status (published, in press, accepted or returned for revision)}  & \RaggedRight\textbf{Nature and} {\%} \RaggedRight\textbf{of student contribution} & \RaggedRight\textbf{Co-author name(s) Nature and} {\%} \RaggedRight\textbf{of Co-author’s contribution} & \RaggedRight\textbf{Co-author(s), Monash student Y/N} \\ \hline
+2 & Visualizing probability distributions across bivariate cyclic temporal granularities & Published & 70\%.
+Concept and developing software and writing first draft & (1) Rob J Hyndman, input into manuscript 15\% (2) Dianne Cook, Concept and input into manuscript 15\% & \multicolumn{1}{c|}{N} \\
+\hline
+3 & Detecting distributional differences between temporal granularities
+for exploratory time series analysis & 80\%.
+Concept and developing software and writing first draft & (1) Rob J Hyndman, Concept and input into manuscript 10\% (2) Dianne Cook, Concept and input into manuscript 10\%  & \multicolumn{1}{c|}{N} \\
+\hline
+4 & Clustering time series based on probability distributions across temporal granularities & To be submitted & 70\%.
+Concept and developing software and writing first draft & (1) Dianne Cook, Concept and input into manuscript 20\% (2) Rob J Hyndman, concept and input into manuscript 10\% & \multicolumn{1}{c|}{N} \\
+\hline
+\end{tabular}
+\end{table}
+
+\FloatBarrier
 
@@
 
-# Acknowledgements {-}
-
-# supervisors
-Rob and Di are phenomenal leaders in their respective fields, who lead by example.
-I am inspired by their creativity, wisdom, discipline, and dedication to really contribute to the society through their research.
-Thank you for constantly pushing me to improve as a programmer and researcher, and for regularly sharing best practices for conducting research.
-Looking back, I am grateful for how my thoughts and work in statistical computing, graphics, and data analytics in general have evolved over the years.
-I still have a lot to learn, but I am a much more self-reliant and independent researcher than I was when I started.
+# Acknowledgements {-#ch:thanks}
+
+<!-- - supervisors -->
+
+Rob and Di are phenomenal leaders in their respective fields who lead by example.
+I am inspired by their creativity, wisdom, discipline, and dedication to contribute to the society through their research.
+Thank you for constantly pushing me to improve as a programmer and researcher and regularly sharing good practices for conducting research.
+Looking back, I am grateful for how my thoughts and work in statistical computing, graphics, and data analytics, in general, have evolved over the years.
+I still have a lot to learn, but I am a much more self-reliant and independent researcher than when I started.
 Di has been instrumental in exposing me to the potential benefits of effective data visualization.
-As a female researcher, I am encouraged by her willingness to pursue unconventional avenues and have frequently noticed her make a conscious decision to question existing  stereotypes and biases.
-Thank you Di for being a fantastic female role model.
-
-# department and PhD colleagues
-
-# R community
-
-# friends
-
-PhD in Covid has been harder and I would not have been able to make it through this if not for the emotional support of my friends and family.
-Thank you Puwasala, Sium for being an epitome of kindness and acting as a pillar of support.
-Thank you Tushar, Samarpita and Nairita for always motivating me and believing in me when I didn't believe in myself.
-Thank you Ian for always putting things in perspective when I lost sight.
-Thank you each one of you for being available, listening to me, and not giving up on me at my lowest.
-Thanks to my housemates Anjali, Surbhi and Dulaji for the fun company and conversations that kept me sane in the covid lockdowns.
-
-# family
-Finally, a big thanks to my family for always being supportive of my choices.
+As a female researcher, I am encouraged by her willingness to pursue unconventional avenues and have frequently noticed her make conscious and proactive choices to question existing stereotypes and biases.
+Thank you, Di, for being a fantastic female role model.
+
+<!-- - Department and conferences -->
+
+I want to thank the Department of Econometrics and Business Statistics, Monash Business School, ARC Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS), and Data61 CSIRO for their financial support.
+Thanks to the Department, I attended a few conferences in person (before COVID hit) and had a great time networking and being inspired by many people.
+A few highlights are UseR!
+2018 (Brisbane), Young Statistician's Conference 2019 (Canberra), WOMBATS 2019 (Melbourne), and RopenSci 2019 (Sydney).
+Thank you to Gael Martin, Farshid Wahid, Klaus Ackermann, and Didier Nibbering from my PhD committee for providing such a supportive environment.
+
+ <!-- - PhD friends and R community -->
+Thank you, Stephanie and Nicholas Tierney, for being so warm, kind and welcoming when I began my PhD.
+Being a part of such an inclusive environment meant a lot to me as an international student.
+Thank you, Nicholas Spyrison and Mitchell O'Hara-Wild, for the numerous informative discussions, brainstorming sessions, as well as for generously sharing your expertise and experience in software development.
+Thank you to each of you, NUMBATs, and the broader R community for directly and indirectly helping me countless times.
+You never fail to inspire me.
+
+<!-- - Friends -->
+
+PhD in Covid has been harder, and I would not have been able to make it through this if not for the emotional support of my friends and family.
+Thank you Puwasala, Sium, and Ian, for being so kind and acting as a pillar of emotional support, especially during the last leg of my journey.
+Thank you Tushar, Samarpita and Nairita for always motivating me and believing in me when I did not believe in myself.
+Thank you to each one of you for being available, listening to me, and putting things in perspective when I lost sight.
+Thanks to my housemates Anjali, Surbhi, and Dulaji for the fun company and conversations that kept me sane in the covid lockdowns.
+
+<!-- - Family -->
+
+Finally, a big thank you to my family for always being supportive of my choices.
 Thanks to my mum (Nupur Gupta) and dad (Arun Prasad Gupta), from whom I learnt that no matter where you start from, if you persevere and are sincere in your efforts, you can sail through any difficult situation.
-They are my constant cheerleaders for all little and big endeavours.
+They are my constant cheerleaders for all little and big endeavors.
 Thanks to my brother (Avijit) for always having my back and inspiring me to dream bigger.
-Thank you, Juhi (sister-in-law) for instilling in me the value of organisation in all aspects of life and the importance of prioritizing my physical health from time to time.
+Thank you, Juhi (sister-in-law), for reminding me of the value of organization in all aspects of life and the importance of prioritizing my physical health from time to time.
 Thank you all for your part in supporting me in this journey.
@@
 
+<!-- add people who are helping you with proofreading -->
+<!-- Add Peter -->
+<!-- Although the journey is coming to an end, you wish to take these life learnings forward -->
+<!-- add Earo and Stuart that you followed their model of workflow and templates, using the power of reproducible and accessible  research -->
+
 <!--chapter:end:Rmd//00-d-acknowlegements.Rmd-->
 
-# Preface {-}
+# Preface {-#ch:preface}
 
@@
 
-Hence, the overarching research goal is to study the periodic behavior of temporal data in a structured way by studying the probability distributions by best exploiting the characteristics of time.
-Slicing and dicing the data in all possible temporal scales as suggested by EDA can be a daunting task as it leads to a myriad of possibilities.
-This inspires the research presented in this thesis, which aims to provide a platform to systematically explore periodicities in temporal data and support finding regular patterns or anomalies, explore clusters of behaviors or summarize the behavior.
-The first part of the work discusses computation of all possible combinations of cyclic time granularities and a graphical mapping such that distributions of a numeric response variable is displayed across combinations of two cyclic granularities.
+<!-- The first part of the work discusses computation of all possible combinations of cyclic time granularities and a graphical mapping such that distributions of a numeric response variable is displayed across combinations of two cyclic granularities.
 Even analyzing the distribution of the measured variable across two cyclic granularities at once could amount to displaying many plots in search of potential  patterns.
-Thus, the first part of the research (@Gupta2020-vo) also introduces "harmony" to denote pairs of granularities that could be analyzed together and reduces the search from all possible options.
+Thus, the first part of the research also introduces "harmony" to denote pairs of granularities that could be analyzed together and reduces the search from all possible options.
 But this approach is still overwhelming for human consumption because there would still be huge number of harmonies.
@@
 
+ <!-- This chapter contains a literature review related to linear granularities. -->
+
 ## Detecting distributional differences between temporal granularities for exploratory time series analysis {#sec:hakear}
 
-Cyclic temporal granularities, which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can be useful for measuring repetitive patterns in large univariate time series data.
-The granularities feed new approaches to exploring time series data.
-One use is to take pairs of granularities, and make plots of response values across the categories induced by the temporal deconstruction.
-However, when there are many granularities that can be constructed for a time period, there will also be too many possible displays to decide which might be the more interesting to display.
-This work proposes a new distance metric to screen and rank the possible granularities, and hence choose the most interesting ones to plot.
-The distance measure is computed for a single or pairs of cyclic granularities can can be compared across different cyclic granularities and also on a collection of time series.
-The methods are implemented in the open-source R package `hakear`.
+This work is a natural extension of Chapter \@ref(ch:gravitas).
+Many, many other displays might be built using different granularities, such as day-of-week, day-of-month, weekday/weekend, and so on.
+However, only a handful of them may be interesting, revealing major patterns.
+Determining which displays exhibit "significant" distributional differences between cyclic granularity categories and plotting only these would allow for more efficient exploration.
+Furthermore, a few of the displays in this collection will be more engaging than others.
+This chapter provides a new distance metric for selecting and ranking the multiple granularities.
+The statistical significance of potential visual discoveries is aided by selecting a threshold for the proposed numerical distance measure.
+The distance measure is computed for a single or pairs of cyclic granularities, and it can be compared across different cyclic granularities as well as a collection of time series.
+This chapter also includes a case study using residential smart meter data from Melbourne to demonstrate how the suggested methodology may be utilized to automatically find temporal granularities with significant distributional differences.
+The methods are implemented in the open-source R package 'hakear.'
 
@@
 
- With more and more time series data being collected at much finer temporal resolution, for a longer length of time, and for a larger number of individuals/entities, time series clustering research is getting a lot of traction.
-The sort of noisy, patchy, uneven, and asynchronous time series that is typical in many disciplines limits similarity searches among these lengthy time series.
-In this work, we suggest a method for overcoming these constraints by grouping time series based on probability distributions over cyclic temporal granularities.
-Cyclic granularities are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, and so on, and can be helpful for detecting repeating patterns.
-Looking at probability distributions across cyclic granularities results in an approach that is robust to missing or noisy data, aids in dimension reduction, and ensures small pockets of similar repeated behaviours.
-The proposed method was tested using a collection of residential electricity customers.
-The simulated and empirical evidence demonstrates that our method is capable of producing meaningful clusters.
+In this chapter, we look at the problem of using clustering to discover patterns in a large number of univariate time series across multiple temporal granularities.
+Time series clustering research is gaining traction as more data is collected at finer temporal resolution, over longer time periods, and for a larger number of individuals/entities.
+Many disciplines have noisy, patchy, uneven, and asynchronous time series that make it difficult to search for similarities.
+We propose a method for overcoming these constraints by calculating distances between time series based on probability distributions at various temporal granularities.
+Because they are based on probability distributions, these distances are resistant to missing or noisy data and aid in dimension reduction.
+When fed into a clustering algorithm, the distances can be used to divide large data sets into small pockets of similar repetitive behavior.
+These subgroups can then be analyzed separately or used as distinct prototype behaviors in classification problems.
+The proposed method was tested on a group of residential electricity consumers from the Australian smart meter data set to show that it can generate meaningful clusters.
+This chapter includes a brief review of the literature on traditional time series clustering and, more specifically, clustering residential smart meter data.
 
@@
 
-#### Single-to-multiple order-up {-}
+#### Single-to-multiple order-up {- #s2m}
 
@@
 
-#### Multiple-to-single order-up {-}
+#### Multiple-to-single order-up {- #m2s}
 
@@
 
-#### Multiple order-up quasi-circular granularities {-}
+#### Multiple order-up quasi-circular granularities {- #quasi}
 
@@
 
-#### Levels {-}
+#### Levels {- #levels}
 
@@
 
-#### Synergy of cyclic granularities {-}
+#### Synergy of cyclic granularities {- #synergy-cyclic}
 
@@
 
-#### Conditioning variable {-}
+#### Conditioning variable {- #conditioning}
 
@@
 
-#### Cyclic granularities search and computation {-}
+#### Cyclic granularities search and computation {- #search}
 
@@
 
-#### Screening and visualizing harmonies {-}
+#### Screening and visualizing harmonies {- #visualize}
 
@@
 
-## Acknowledgments {-}
+## Acknowledgments {- #thanksgravitas}
 
@@
 The Github repository,  [github.com/Sayani07/paper-gravitas](https://github.com/Sayani07/paper-gravitas), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
-This article was created with `knitr` [@knitr, @R-knitr] and `rmarkdown` [@rmarkdown, @R-rmarkdown].
-
-## Supplementary Materials
+This article was created with `knitr` [@knitr], [@R-knitr] and `rmarkdown` [@rmarkdown], [@R-rmarkdown].
+
+## Supplementary Materials {- #supplement-gravitas}
 
@@
 
-## Acknowledgments {-}
+## Acknowledgments {- #thankshakear}
 
@@
 The Github repository,  [github.com/Sayani07/paper-hakear](https://github.com/Sayani07/paper-gravitas), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
-This article was created with R [@R-language], `knitr` [@knitr2015, @R-knitr] and `rmarkdown` [@rmarkdown2018, @R-rmarkdown].
+This article was created with R [@R-language], `knitr` [@knitr2015], [@R-knitr] and `rmarkdown` [@rmarkdown], [@R-rmarkdown].
 Graphics are produced with @Wickham2009pk.
 
-## Supplementary Materials
+## Supplementary Materials {- #supplement-hakear}
 
@@
 
-## Acknowledgments {-}
+## Acknowledgments {- #thanksgracsr}
 
@@
 The Github repository, [github.com/Sayani07/paper-gracsr](https://github.com/Sayani07/paper-gracsr), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
-This article was created with R [@R-language], `knitr` [@knitr, @R-knitr] and `rmarkdown` [@rmarkdown, @R-rmarkdown].
+This article was created with R [@R-language], `knitr` [@knitr2015], [@R-knitr] and `rmarkdown` [@rmarkdown], [@R-rmarkdown].
 Graphics are produced with `ggplot2` [@Wickham2009pk] and GGally [@R-GGally].
 
-## Supplementary Materials
+## Supplementary Materials {- #supplement-gracsr}
 
@@
 
-The **gravitas** package provides very general tools to compute and manipulate cyclic granularities, and to generate plots displaying distributions conditional on those granularities.
-It can be utilized in non-temporal cases for which a hierarchical structure can be construed similar to time.
-It is on CRAN.
+The **gravitas** package provides very general tools to compute and manipulate cyclic granularities and generate plots displaying distributions conditional on those granularities.
+The functions `search_gran()`, `create_gran()`, `harmony()`, `gran_advice()` and `prob_plot()` provides the entire workflow for an analyst to systematically explore large quantities of temporal data across different harmonies (pairs of granularities that can be analyzed together.) This package was developed as part of my internship at Google Summer of Code, 2019.
+It has been on CRAN since January 2020.
 The website (<https://sayani07.github.io/gravitas>) includes full documentation and two vignettes about the package usage.
-There has been a grand total of 12K downloads from the RStudio mirror dating from 2020-11-01 to 2021-11-01.
+There has been a total of 12K downloads from the RStudio mirror dating from 2020-11-01 to 2021-11-01.
+This package supplements the paper corresponding to Chapter \@ref(ch:gravitas), which has won the ACEMS Business Analytics Award 2021.
+The package can be generalized to non-temporal applications for which a hierarchical structure can be construed similar to time.
 
@@
 
-The open-source R package `hakear` is available on Github (https://github.com/Sayani07/hakear) to implement ideas presented in Chapter \@ref(ch:hakear).
-Given a `tsibble` and context granularities of interest, the function `wpd()` provides support for computing the wpd for each cyclic granularities or pair of granularities and `select_harmonies()` chooses the ones with significant patterns and ranks them from highest to lowest wpd.
+The R package hakear (https://github.com/Sayani07/hakear) implements ideas presented in Chapter \@ref(ch:hakear).
+The function wpd() computes the weighted pairwise distances ($wpd$) for each cyclic granularity or pair of granularities, and `select_harmonies()` selects the ones with significant patterns and ranks them from highest to lowest $wpd$.
+These selected harmonies can be plotted using package `gravitas` for potentially interesting displays.
+This package is reliant on parallel processing using multiple multi-core computers for faster computation of $wpd$.
 
@@
 The open-source R package `gracsr` is available on Github (https://github.com/Sayani07/gracsr) to implement ideas presented in Chapter \@ref(ch:gracsr).
-The package provides functions to carry out the entire clustering methodology discussed in the paper.
-It is still a work in progress and has won the ACEMS Business Analytics Prize 2021 with a prize money of AUD 3000, which would be utilized in polishing this package and preparing it for CRAN.
+The package contains functions for implementing the complete clustering methodology with choices of scaling and distance metrics discussed in the chapter.
+It has received a grant (AUD 3000) as part of the ACEMS Business Analytics Prize towards polishing the functions and preparing it for CRAN.
 
@@
 
-With the volume of data projected to grow in the future, potentially leading to increased variability in patterns, research is needed to understand the issues that may arise with the existing methodology and how to adapt our current algorithm.
-Computational efficiency is also critical when scaling up for analysis of huge data sets, let alone when adding features to existing highly dimensional data structures.
-
-### Comparing our clustering method with other benchmark methods
-
-Testing needs to be carried out with non-hierarchy based clustering methods to see whether these distances perform better with other algorithms.
-Also, to evaluate how much information is lost when aggregating individual level demand versus distributions.
+With the volume of data expected to increase in the future, potentially leading to increased variability in patterns, more experiments are needed to understand what sort of problems may arise with our current methodology and how to improve our current algorithm to address those issues.
+All of this must be done while keeping computational efficiency in mind, as this is critical when scaling up for analysis of large data sets.
+
+### Develop inferential methods for the clusters
+
+It is crucial to determine whether or not any detected clusters or patterns are statistically significant.
+In Chapter \@ref(ch:gracsr), we have used the permutation approach to detect if patterns are authentic or not across one or a pair of granularities.
+A similar method for finding significance across clusters is not appropriate if the observations are assumed to be independent.
+Most large temporal data sets gathered at fine temporal scales usually include several repetitive patterns.
+Therefore, the permutations used for comparisons must account that different temporal granularities have distinct conditional distributions.
 
@@
 
-We provide solutions that are realistically applicable to any temporal data observed more than once per year.
-We could just verify its value in the electricity smart meter context.
-With numerous open-source benchmark data sets accessible, it is necessary to test how well the approaches operate in various disciplines.
-
-## Final words
+We propose solutions that are realistically applicable to any temporal data collected more than once per year and may be implemented in a variety of settings.
+We could simply verify its utility in the context of smart meter electricity data.
+With the availability of several open-source benchmark data sets, it is vital to evaluate how well the methodologies perform across a variety of fields.
 
@@
 
-# (APPENDIX) Appendix {-}
-
-# Data dictionary
+# (APPENDIX) Appendix {-#ch:appendix}
+
+## Data dictionary
 
@@
 
+# References {- #ch:references}
+
 <!--chapter:end:Rmd//99-references.Rmd-->

### 2021-11-06 Rmd/00-c-declaration.Rmd
added declaration (need to run the percentages through Rob and Di)

@@
+---
+output:
+  pdf_document: default
+  html_document: default
+---
 `r if (knitr::is_html_output()) '<!--'`
@@
 3 & Detecting distributional differences between temporal granularities
-for exploratory time series analysis & To be submitted & 80\%.
-Concept and developing software and writing first draft & (1) Rob J Hyndman, Concept and input into manuscript 10\% (2) Dianne Cook, Concept and input into manuscript 10\%  & \multicolumn{1}{c|}{N} \\
+for exploratory time series analysis & To be submitted & 70\%.
+Concept and developing software and writing first draft & (1) Rob J Hyndman, Concept and input into manuscript 15\% (2) Dianne Cook, Concept and input into manuscript 15\%  & \multicolumn{1}{c|}{N} \\
 \hline
@@
 \end{table}
-
-\FloatBarrier
 

### 2021-11-07 Rmd/06-appendix.Rmd
putting reference even if unnumbered
adding appendix

@@
-# (APPENDIX) Appendix {-}
+# Appendix {-#ch:appendix}
 
-## Data dictionary
+\appendix
 
-```{r wdi-data-dict}
+This is the supplementary paper corresponding to the paper "Detecting distributional differences between temporal granularities for exploratory time series analysis"
+
+```{r external-hakear-supplementary, include = FALSE, cache = FALSE}
+read_chunk("scripts/hakear.R")
+read_chunk("scripts/hakear-supplementary.R")
 ```
 
-(ref:wdi-data-dict) Data dictionary for the World Development Indicators.
+```{r load}
+```
+
+# Recalling notations
+
+Let $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ be a continuous measured variable observed across $T$ time points.
+The number of cyclic granularities considered in the display (at once) is $m$.
+Consider cyclic granularities $A$ and $B$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$.
+For $m=1$, only one of $A$ or $B$ is plotted at once and a _panel_ refers to a display of distributions of $v$ across $J$ or $K$ levels on the x-axis.
+For $m=2$, the distribution display of $v$ with $A$ placed across x-axis and $B$ across facets is referred to as a $(J, K)$ _panel_ ($J$ x-axis levels and $K$ facet levels).
+The pairwise distances between pairs ($a_jb_k,a_j'b_k'$) could be within-facets or between-facets as seen in Figure 4 of the main paper.
+The tuning parameter, used to put relative weight-age to the pairwise distances within and between facets is denoted by $\lambda$.
+Let the four elementary designs be $D_{null}$ where there is no pairwise difference in distribution of $v$ across $A$ or $B$, $D_{var_f}$ denotes the set of designs where there is difference in distribution of $v$ for $B$ and not for $A$.
+Similarly, $D_{var_x}$ denotes the set of designs where difference is observed only across $A$.
+Finally, $D_{var_{all}}$ denotes those designs for which difference is observed across both $A$ and $B$.
+The following method is deployed for generating different distributions across different combinations for non-null designs - suppose the distribution of the combination of first levels of x and facet category is $N(\mu,\sigma)$ and $\mu_{jk}$ denotes the mean of the combination $(a_jb_k)$, then $\mu_{j.} = \mu + j\omega$ (for design $D_{var_x}$) and $\mu_{.k} = \mu + k\omega$ (for design $D_{var_f}$), where $\omega$ denotes the increment in mean.
+Table \ref{tab:explain-design} shows an example of how designs are defined for a $(2, 3)$ panel using $\omega = 3$.
+$nx$ and $nfacet$ denotes the number of categories placed on x-axis and facets respectively.
+$wpd_{raw}$ and $wpd$ denote the raw and normalized weighted pairwise distances.
+
+```{r explain-design, fig.pos="ht", eval=FALSE}
+sim_varx_normal = function(nx, nfacet, mean, sd, w)
+{
+  rep(dist_normal((mean + seq(0, nx-1, by  = 1)*w), sd), nfacet)
+}
+
+sim_varf_normal = function(nx, nfacet, mean, sd, w)
+{
+  rep(dist_normal((mean + seq(0, nfacet-1, by  = 1)*w), sd), each = nx)
+}
+  sim_varall_normal = function(nx, nfacet, mean, sd, w)
+  {
+    dist_normal((mean + seq(0, (nx*nfacet - 1), by  = 1)*w), sd)
+  }
+
+mean = 0
+sd = 1
+w = 3
+nx = 2
+nfacet = 3
+
+raw_levels = expand.grid(facet = c("$b_1$", "$b_2$", "$b_3$"),
+                         x = c("$a_1$", "$a_2$"))
+
+facet = rep(c("$b_1$", "$b_2$", "$b_3$"), each = nx)
+x =  rep(c("$a_1$", "$a_2$"), nfacet)
+
+null <-  sim_varall_normal(nx, nfacet, mean, sd, w=0) %>% tibble()
+vary_f <-  sim_varf_normal(nx, nfacet, mean, sd, w) %>% tibble()
+vary_x <-  sim_varx_normal(nx, nfacet, mean, sd, w) %>% tibble()
+vary_all <-  sim_varall_normal(nx, nfacet, mean, sd, w) %>% tibble()
+
+raw_table <- bind_cols(x = x, 
+                       facet = facet, 
+                       null = null,
+                       vary_f = vary_f, 
+                       vary_x = vary_x, 
+                       vary_all = vary_all) %>% 
+  tibble() 
+names(raw_table) = c("x levels", "facet levels", "$D_{null}$",  "$D_{var_f}$", "$D_{var_x}$", "$D_{var_{all}}$" )
+raw_table %>% 
+  kable(format = "latex",
+    booktabs = TRUE, 
+    escape = FALSE,
+    caption = "Simulation setup for a panel with 3 facet levels and 2 x-axis levels for different designs starting from an initial distribution N(0, 1) for the combination $(a_1, b_1)$ and $\\omega=3$.") %>% kable_styling()
+```
+
+# Raw weighted pairwise distance
+
+## Tuning parameter
+
+For $m=1$, pairwise distances could be defined only between different categories of the cyclic granularity considered.
+So no tuning parameter is applicable for this case.
+For $m=2$, $\lambda$ might impact $wpd_{raw}$ differently depending on the value of $\omega$ and different values of $nx$, $nfacet$ and designs.
+The following simulation study sees the impact of $\lambda$ for different $\omega$, $nx$, $nfacet$ and designs.
+
+_Simulation design_
+
+Observations are generated from N(0,1) distribution for each combination of $nx$ and $nfacet$ from the following sets: $nx = nfacet = \{2, 3, 5, 7\}$ and $wpd_{raw}$ is computed for $\lambda = {0.1, 0.2, \dots, 0.9}$ under two designs $D_{var_x}$ and $D_{var_f}$ for $\omega = \{1, 8\}$  to observe how the value of $wpd_{raw}$ changes for different designs.
+Moreover, to observe for which value of $\lambda$ the two designs intersect, we generate observations from N(0,1) distribution for each combination of $nx$ and $nfacet$ from the following sets: $nx = nfacet = \{2, 3, 5, 7, 14, 20\}$ and $\omega = \{1, 2, \dots, 10\}$ under the same two designs.
+
+<!-- and the values of $wpd$ is computed for all these different values of the variables $nx$, $nfacet$, $\lambda$ and $\omega$ under two different designs $D_{var_x}$ and $D_{var_f}$ .  -->
+
+_Results_
+
+Figure \ref{fig:tuning-oneomega} shows how the value of $wpd$ changes for $\lambda = {0.1, 0.2, \dots, 0.9}$ for the two different designs $D_{var_x}$ and $D_{var_f}$ for two values of increment in mean $\omega = {1, 8}$.
+For a lower value of $\omega$, the two designs intersect at $\lambda > 0.7$ and for a higher $\omega$, the two designs intersect at $\lambda =  0.5$.
+The value of $wpd$ increases with $\lambda$ for $D_{var_x}$ and decreases with increasing $\lambda$ for $D_{var_f}$.
+\noindent Figure \ref{fig:tuning-parameter-inter} shows the value of $\lambda$ for which the two designs intersect for different values of $\omega$.
+It can be observed that as the value of $\omega$ ($\omega > 4$) increase, the value of $\lambda$ at which the two designs intersect converge is $\lambda = 0.5$ and for $\omega$ ($\omega <= 4$), designs interact between $\lambda = [0.6,0.75]$.
+Hence, $\lambda = 0.67$ is used for the rest of the paper, which is an average of the range of values $\lambda$ can take in order to have equal or more weightage for the within-facet distances.
+
+```{r tuning-oneomega, fig.cap = " $wpd_{raw}$ from two designs are plotted for different values of $\\lambda$ $nx$ and $nfacet$.
+For $\\omega = 1$, the designs intersect for $0.6<\\lambda<=0.75$, whereas for higher omega, design intersects at $\\lambda = 0.5$."}
+knitr::include_graphics(here::here("data/hakear/simulations/tuning_param/figs/fixed_omega.png"))
+```
+
+```{r tuning-parameter-inter, fig.cap ="The point of intersection of $wpd_{raw}$ values under the designs $D_{var_f}$ and $D_{var_x}$ are plotted across different $\\omega$ and $\\lambda$.
+For most panels it is observed that a common value of $\\lambda$ for which the designs interact is $0.5$ for $\\omega >= 4$, which implies any value greater than $0.5$ could be chosen to up-weigh the within-facet distances and down-weigh the between-facet distances.
+The value of $\\lambda$ is higher for  $\\omega<4$ could be anywhere between 0.6 and 0.75."}
+knitr::include_graphics(here::here("data/hakear/simulations/tuning_param/figs/intersection_plot.png"))
+```
+
+## Underlying distributions
+
+The following simulation study sees the impact of different underlying distributions on $wpd_{raw}$ for different $nx$ and $nfacet$ before and after performing Normal Quantile Transformation.
+
+_Simulation design_
+
+Observations are generated from N(0,1), N(5, 1), N(0, 5), Gamma(0.5, 1) and Gamma(2, 1) distributions for each combination of $nx$ and $nfacet$ from the following sets: $nx = nfacet = \{2, 3, 5, 7, 14, 20, 31, 50\}$ and $wpd_{raw}$ is computed with $\lambda = 0.67$ with (scenario 1) and without (scenario 2) NQT.
+
+_Results_
+
+Figure \ref{fig:gamma-diff-wotrans} shows ridge plots of $wpd_{raw}$ for a Gamma(0.5,1), Gamma(2,1) before NQT.
+It is observed that for the underlying distribution Gamma(2,1), location and scale of the distribution of wpd changes from top-left panel to bottom-right panel.
+Figure \ref{fig:gamma-diff-quantrans} shows the the distributions of $wpd$ under same underlying distributions after performing NQT.
+It is observed that within each panel, the distributions of the $wpd$ looks same, however, the distributions change from extreme top-left panel to bottom-right panels.
+Similar observations could be made in Figure \ref{fig:normal-diff-quantrans} for different underlying normal distributions  N(0,1), N(5,1) and N(0,5).
+This implies, NQT has atleast been able to make the location and scale of the distribution of $wpd_{raw}$ same for different underlying distributions.
+
+<!-- Since the measure $wpd$ is essentially set up to detect "differences" in distributions irrespective of underlying distribution, it would be ideal if it has minimal dependency on the type, location and scale of the initial distribution.
+To that end, some data pre-processing through the Normal Score Transform (NQT) has been applied in order to make most asymmetrical distributed measured variables more normal-like.
+-->
+
+```{r gamma-diff-wotrans, fig.cap="Ridge plots of raw wpd is shown for Gamma(0.5,1), Gamma(2,1) distribution without NQT.
+The densities change across different facet and x levels and also looks different for the two distributions, which implies wpd value is affected by the change in the shape paramter of the gamma distribution."}
+knitr::include_graphics(here::here("data/hakear/simulations/raw/null_design_scale/figs/gamma_ridge_nxbynfacet.png"))
+```
+
+```{r gamma-diff-quantrans, fig.cap="Ridge plots of raw wpd is shown for Gamma(0.5,1), Gamma(2,1) distribution.
+The densities change across different facet and x levels but look same for the two distributions, which implies wpd value is unaffected by the change in the shape paramter of the gamma distribution"}
+knitr::include_graphics(here::here("data/hakear/simulations/raw/null_design_quantrans/figs/diff_mean3_gamma.png"))
+```
+
+```{r normal-diff-quantrans, fig.cap = "Ridge plots of raw wpd is shown for N(0,1), N(5,1) and N(0,5) distribution.
+The densities change across different facet and x levels but look same for each panel, which implies wpd value is unaffected by the change in mean and standard deviation of the normal distribution."}
+knitr::include_graphics(here::here("data/hakear/simulations/raw/null_design_quantrans/figs/diff_mean3_normal.png"))
+```
+
+## Number of comparisons
+
+### Case: m = 1
+
+_Simulation design_
+
+\noindent Observations are generated from a N(0,1) distribution for each $nx = \{2, 3, 5, 7, 9, 14, 17, 20, 24, 31, 42, 50\}$ to cover a wide range of levels from very low to moderately high.
+$ntimes = 500$ observations are drawn for each combination of the categories, that is, for a panel with $nx=3$, $500$ observations are simulated for each of the categories.
+This design corresponds to $D_{null}$ as each combination of categories in a panel are drawn from the same distribution.
+Furthermore, the data is simulated for each of the categories $nsim=200$ times, so that the distribution of $wpd$ under $D_{null}$ could be observed.
+The values of $wpd$ is obtained for each of the panels.
+$wpd_{l, s}$ denotes the value of $wpd$ obtained for the $l^{th}$ panel and $s^{th}$ simulation.
+
+_Results_
+
+Figure \ref{fig:nxbyfacet-ridge-N01-onegran} shows ridge plots of $wpd_{raw}$ for an underlying N(0,1) distribution.
+For each panel, it could be seen that the location shifts to the right for increasing x levels.
+Across each panel, the scale of the distribution seems to change for low/moderately from lower values  to higher values of $nx$ and left tails are longer for lower facet levels.
+
+```{r nxbyfacet-ridge-N01-onegran, fig.cap="Ridge plots of raw wpd is shown for N(0,1) distribution.
+For each panel, it could be seen that the location shifts to the right for increasing x levels.
+Across each panel, the scale of the distribution seems to change for low/moderately lower values and higher values of nfacets and left tails are longer for lower facet levels."}
+knitr::include_graphics(here::here("data/hakear/simulations/supplementary/one-gran/raw/null_design_quantrans/figs/nxbyfacet_ridge_wpd_N01.png"))
+```
+
+```{r nfacetbynx-onegran, fig.cap="Ridge plots of raw wpd is shown for N(0,1) distribution. For each panel, it could be seen that the peakedness shifts for increasing x levels."}
+knitr::include_graphics(here::here("data/hakear/simulations/supplementary/one-gran/raw/null_design_quantrans/figs/nfacetbynx_ridge_wpd_N01.png"))
+```
+
+```{r density-raw-onegran, eval = FALSE}
+knitr::include_graphics(here::here("data/hakear/simulations/supplementary/one-gran/raw/null_design_quantrans/figs/nxbyfacet_density_wpd_N01.png"))
+```
+
+### Case: m = 2
+
+_Simulation design_
+
+\noindent Similarly, observations are generated from a N(0,1) distribution for each combination of $nx$ and $nfacet$ from the following sets: $nx = nfacet = \{2, 3, 5, 7, 14, 20, 31, 50\}$.
+That is, data is being generated for each of the panels $(2, 2), (2, 3), (2, 5) \dots, (50, 31), (50, 50)$.
+For each of the $64$ panels, $ntimes = 500$ observations are drawn for each combination of the categories.
+That is, if we consider a $(2, 2)$ panel, $500$ observations are generated for each of the possible subsets, namely, $\{(1, 1), (1, 2), (2, 1), (2, 2)\}$.
+
+<!-- Observations are generated from Gamma(2,1), G(0.5, 1), N(0,1), N(0, 5) and N(5, 1) distribution for each combination of $nx$ and $nfacet$ from the following sets: $nx = nfacet = \{2, 3, 5, 7, 14, 20, 31, 50\}$ to cover a wide range of levels from very low to moderately high.
+Each combination is being referred to as a _panel_.
+That is, data is being generated for each of the panels $\{nx = 2, nfacet = 2\}, \{nx = 2, nfacet = 3\}, \{nx = 2, nfacet = 5\},  \dots, \{nx = 50, nfacet = 31\}, \{nx = 50, nfacet = 50\}$.
+For each of the $64$ panels, $ntimes = 500$ observations are drawn for each combination of the categories.
+That is, if we consider the panel $\{nx = 2, nfacet = 2\}$, $500$ observations are generated for each of the combination of categories from the panel, namely, $\{(1, 1), (1, 2), (2, 1), (2, 2)\}$.
+The values -->
+<!-- of $wpd$ is obtained for each of the panels.
+The measurement variable for each combination of categories in a panel are drawn from the same distribution and hence the design corresponds to $D_{null}$ .
+Furthermore, this entire method is repeated for each panels $nsim=200$ times, so that the distribution of $wpd$ under $D_{null}$ could be observed.
+-->
+
+_Results_
+
+Figures \ref{fig:nxbyfacet-ridge-N01}, \ref{fig:nfacetbynx} and  \ref{fig:density-raw} shows the ridge plot of $wpd_{raw}$ with $nx$ as facets, $nfacet$ as facets and the density plot of $wpd_{raw}$ with $nx$ on the x-axis and $nfacet$ on the facets.
+The right-ward shift in location can be clearly observed from all figures.
+
+```{r nxbyfacet-ridge-N01, fig.cap="Ridge plots of raw wpd is shown for N(0,1) distribution with nfacets as facets.
+For each panel, it could be seen that the location shifts to the right for increasing nx levels."}
+knitr::include_graphics(here::here("data/hakear/simulations/raw/null_design_quantrans/figs/nxbyfacet_ridge_wpd_N01.png"))
+```
+
+```{r nfacetbynx, fig.cap="Ridge plots of raw wpd is shown for N(0,1) distribution with nx as facets.
+For each panel, it could be seen that the location shifts to the right for increasing nfacet levels."}
+knitr::include_graphics(here::here("data/hakear/simulations/raw/null_design_quantrans/figs/nfacetbynx_ridge_wpd_N01.png"))
+```
+
+```{r density-raw, fig.cap="Density plot of raw wpd is shown for N(0,1) distribution with nx on x-axis and nfacet on facets.
+For each panel, it could be seen that the location as well as scale shifts as we move from top-left to bottom-right corner."}
+knitr::include_graphics(here::here("data/hakear/simulations/raw/null_design_quantrans/figs/nxbyfacet_density_wpd_N01.png"))
+```
+
+# Adjusted weighted pairwise distances
+
+_Simulation design_
+
+This section shows results from the different approaches of adjusting for the number of comparisons.
+Simulation design is the same as for raw weighted pairwise distances for $m=1$ and $m=2$.
+In place of $wpd_{raw}$, $wpd_{perm}$, $wpd_{glm}$ and $wpd$ is computed for all panels.
+
+<!-- \noindent Observations are generated from a N(0,1) distribution for each $nx = \{2, 3, 5, 7, 9, 14, 17, 20, 24, 31, 42, 50\}$ to cover a wide range of levels from very low to moderately high.
+$ntimes = 500$ observations are drawn for each combination of the categories, that is, for a panel with $nx=3$, $500$ observations are simulated for each of the categories.
+This design corresponds to $D_{null}$ as each combination of categories in a panel are drawn from the same distribution.
+Furthermore, the data is simulated for each of the categories $nsim=200$ times, so that the distribution of $wpd$ under $D_{null}$ could be observed.
+The values of $wpd$ is obtained for each of the panels.
+$wpd_{l, s}$ denotes the value of $wpd$ obtained for the $l^{th}$ panel and $s^{th}$ simulation.
+-->
+
+## Permutation approach
+
+Figure \ref{fig:one-gran-nxnyfacet} and \ref{fig:norm} show the normalized ${wpd}$ for $m=1$ and $m=2$ respectively.
+The location and scale of the distribution of ${wpd}$ is similar for increasing values of $nx$ or $nfacet$.
+Due to heavy computational load, $wpd_{perm}$ for few panels with very high value of $nx$ or $nfacet$ are not computed.
+
+```{r one-gran-nxnyfacet, fig.cap="Distribution of $wpd_{perm}$ are shown for different levels for $m=1$ under the null design. The distribution for different panels have similar location and scale."}
+include_graphics(here("data/hakear/simulations/supplementary/one-gran/norm/null_design_quantrans_nperm/figs/nxbyfacet_ridge_wpd_N01.png"))
+```
+
+```{r norm, fig.cap = "Distribution of $wpd_{perm}$ are shown for different levels for $m=2$ under the null design. The distribution for different panels have similar location and scale."}
+G21_norm <- read_rds(here("data/hakear/simulations/norm/null_design_quantrans_nperm/data-agg/all_data_wpd_N01.rds"))
+
+G21_norm %>% 
+  ggplot(aes(x = value)) + 
+  geom_density(fill = "blue") +
+  facet_grid(nx~nfacet,
+             labeller = "label_both") + 
+  scale_x_continuous(breaks = scales::breaks_extended(3)) + 
+  xlab("wpd normalised using permutation approach") +    xlab(TeX("wpd_{perm}")) + theme_bw()
+
+```
+
+## Modeling approach
+
+For for $m=1$, Figure \ref{fig:glm-horizontal-onegran1} shows the scatterplot of $wpd_{raw}$ against different values of $nx$ and also the display of residuals from the model and Figure \ref{fig:glm-density-onegran} shows the distribution of $wpd_{glm}$ for different $nx$.
+For $m=2$, Figure \ref{fig:glm-horizontal-twogran1} shows the scatter plot of $wpd_{raw}$ against different values of $nx$ and $nfacet$ and also the display of residuals from the model.
+Figure \ref{fig:wpd-glm-dist-appendix} shows the distribution of $wpd_{glm}$ for different $nx$ and $nfacet$.
+Tables \ref{tab:tab-modelsm1} and \ref{tab:tab-models} show the GLM results for $m=1$ and $m=2$ respectively.
+All of these jointly show that the residuals from the model are independent of the number of comparisons.
+The location and scale of the distribution of $wpd_{glm}$ is similar for higher levels and distinctly different for lower levels.
+
+```{r tab-modelsm1}
+library(here)
+G21_onegran <- read_rds(here("data/hakear/simulations/supplementary/one-gran/raw/null_design_quantrans/data-agg/all_data_wpd_N01.rds"))
+
+G21_median_onegran  <- G21_onegran  %>% 
+  group_by(nx*nfacet) %>% 
+  summarise(actual = median(value))
+
+glm_fit_onegran  <- glm(actual ~ log(`nx * nfacet`),
+               family = Gamma(link = "inverse"),
+               data = G21_median_onegran)
+
+bind_cols(distribution = "N(0, 1)", broom::tidy(glm_fit_onegran)) %>% 
+  mutate(estimate = round(estimate, 2),
+         std.error = round(std.error, 3),
+         statistic = round(statistic, 2),
+         p.value = round(p.value, 2)) %>% kable(caption = "Results of generalised linear model to capture the relationship between $wpd_{raw}$ and number of comparisons for $m=1$.")
+
+```
+
+```{r glm-horizontal-onegran1-raw}
+
+G21 <- read_rds(here("data/hakear/simulations/supplementary/one-gran/raw/null_design_quantrans/data-agg/all_data_wpd_N01.rds"))
+
+G21_median <- G21 %>% 
+  group_by(nx*nfacet) %>% 
+  summarise(actual = median(value))
+
+glm_fit <- glm(actual ~ log(`nx * nfacet`),
+               family = Gamma(link = "inverse"),
+               data = G21_median)
+
+intercept <- glm_fit$coefficients[1]
+slope <- glm_fit$coefficients[2]
+G21_sd  = G21 %>% 
+  mutate(wpd_glm =  (value - (1/(intercept + slope*log(nx*nfacet)
+  )
+  )
+  ))
+#scale_fac <- 1/G21_sd$wpd_glm %>% sd()
+
+fitted_glm <- fitted(glm_fit, type = "response")
+residuals <- residuals.glm(glm_fit, type = "response")
+
+residual <- G21_median$actual  - fitted_glm
+
+G21_horizontal <- G21 %>% 
+  ggplot(aes(x=log(nx*nfacet),
+             y = (value - (1/(intercept + slope*log(nx*nfacet)
+             )
+             )
+             )
+  )
+  ) +
+  geom_point() + stat_summary(fun=mean, geom="line", aes(group=1), color = "blue", size = 1) + 
+  ylab(TeX("wpd_{glm} = wpd_{raw} - 1/(a  + b*log(nx*nfacet))")) + xlab("log(nx)")
+
+#ggsave("figs/G21_horizontal.png")
+```
+
+```{r glm-horizontal-onegran1, fig.cap = "$wpd_{raw}$ (points) and residuals (blue line) from the model is plotted across different number of comparisons for $m=1$.
+Residuals seem to be independent of $nx$ and have been defined as $wpd_{glm}$."}
+knitr::include_graphics(here("img/G21_horizontal.png"))
+```
+
+```{r glm-density-onegran, fig.cap = "Distribution of $wpd_{glm}$ is plotted across different $nx$ for $m=1$.
+Both shape and location of the distributions are different for lower $nx$ than for higher $nx$."}
+G21_glm_m1 <- G21 %>% 
+  mutate(wpd_glm =  (value - (1/(intercept + slope*log(nx*nfacet)
+  )
+  )
+  ),
+  wpd_glm_scaled = ((wpd_glm*261)))
+
+mylabels <- G21_glm_m1 %>% 
+  group_by(nx) %>% 
+  summarise(mean = round(mean(wpd_glm), 3), sd = round(sd(wpd_glm), 3)) %>% 
+  mutate(label_glm = paste0("(", mean, ", " ,sd, ")"))
+
+nxbyfacet_ridge <- G21_glm_m1 %>% 
+  ggplot(aes(x = wpd_glm, y = as.factor(nx))) +
+  ggridges::geom_density_ridges() +
+  xlab(TeX("wpd_{glm}")) +
+  ylab("nx") 
+  # geom_label(data = mylabels,
+  #            aes(y = as.factor(nx),
+  #                x = label_glm, 
+  #                label = label_glm))
+
+nxbyfacet_ridge
+```
+
+```{r tab-models}
+G21 <- read_rds(here("data/hakear/simulations/raw/null_design_quantrans/data-agg/all_data_wpd_Gamma21.rds"))
+G01 <- read_rds(here("data/hakear/simulations/raw/null_design_quantrans/data-agg/all_data_wpd_Gamma01.rds"))
+N01 <- read_rds(here("data/hakear/simulations/raw/null_design_quantrans/data-agg/all_data_wpd_N01.rds"))
+N05 <- read_rds(here("data/hakear/simulations/raw/null_design_quantrans/data-agg/all_data_wpd_N05.rds"))
+N51 <- read_rds(here("data/hakear/simulations/raw/null_design_quantrans/data-agg/all_data_wpd_N51.rds"))
+N55 <- read_rds(here("data/hakear/simulations/raw/null_design_quantrans/data-agg/all_data_wpd_N55.rds"))
+
+all_data <- bind_rows(G21 = G21, G01 = G01, N01= N01, N05 = N05, N51= N51, N55 = N55, .id = "group")
+
+all_median <- all_data %>% 
+  group_by(group, nx*nfacet) %>% 
+    summarise(actual = median(value))
+
+glm_model <- function(df) {
+  glm(actual ~ log(`nx * nfacet`), 
+               family = Gamma(link = "inverse"), data = df)
+}
+
+glm_fit <- all_median %>% 
+  group_by(group) %>% 
+  nest()
+
+models <- map(glm_fit$data, glm_model)
+
+by_dist <- glm_fit %>% 
+  mutate(model = map(data, glm_model))
+
+tidy_obs <- by_dist %>%
+  mutate(tidy = map(model, tidy)) %>%
+  unnest(cols = c(tidy)) %>% 
+  mutate(estimate = round(estimate, 2),
+         std.error = round(std.error, 3),
+         statistic = round(statistic, 2))%>% 
+  mutate(distribution = case_when(
+    group== "G01" ~ "Gamma(0.5, 1)",
+    group== "G21" ~ "Gamma(2, 1)",
+    group== "N01" ~ "N(0, 1)",
+    group== "N05" ~ "N(0, 5)",
+    group== "N51" ~ "N(5, 1)",
+    group== "N55" ~ "N(5, 5)",
+    TRUE~ as.character(group)
+    )) %>% ungroup() %>% 
+  select(-data, -model, - group) %>% 
+     select(distribution, everything())
+
+tidy_obs %>% kable(caption = "Results of generalised linear model to capture the relationship between $wpd_{raw}$ and number of comparisons for different underlying distribution.")
+```
+
+<!-- # ```{r} -->
+<!-- #  -->
+<!-- # G21_median <- N01 %>% -->
+<!-- #   group_by(nx*nfacet) %>% -->
+<!-- #   summarise(actual = median(value)) -->
+<!-- #  -->
+<!-- #  -->
+<!-- # glm_fit <- glm(actual ~ log(`nx * nfacet`), -->
+<!-- #                family = Gamma(link = "inverse"), -->
+<!-- #                data = G21_median) -->
+<!-- # # tidy(glm_fit) -->
+<!-- #  -->
+<!-- # ``` -->
+
+```{r glm-horizontal-twogran, fig.cap = "Residuals from the model is plotted across different $nx$ and $nfacet$.
+Residuals seem to be independent of $log(nx*nfacet)$ and have been defined as $wpd_{glm}$ in the paper."}
+
+G21 <- read_rds(here("data/hakear/simulations/raw/null_design_quantrans/data-agg/all_data_wpd_N01.rds"))
+
+G21_median <- G21 %>% 
+  group_by(nx*nfacet) %>% 
+  summarise(actual = median(value))
+
+glm_fit <- glm(actual ~ log(`nx * nfacet`),
+               family = Gamma(link = "inverse"),
+               data = G21_median)
+
+intercept <- glm_fit$coefficients[1]
+slope <- glm_fit$coefficients[2]
+G21_sd  = G21 %>% 
+  mutate(wpd_glm =  (value - (1/(intercept + slope*log(nx*nfacet)
+  )
+  )
+  ))
+#scale_fac <- 1/G21_sd$wpd_glm %>% sd()
+# checking the fit of the residuals from glm fit
+# fitted_glm <- fitted(glm_fit, type = "response")
+# residuals <- residuals.glm(glm_fit, type = "response")
+# hist(residuals)
+#h = augment(glm_fit)
+# ggplot(h) +
+#   geom_histogram(aes(x = .resid))
+# residual <- G21_median$actual  - fitted_glm
+# 
+# hist(residual)
+G21_horizontal <- G21 %>%
+  ggplot(aes(x=log(nx*nfacet),
+             y = (value - (1/(intercept + slope*log(nx*nfacet)
+             )
+             )
+             )
+  )
+  ) +
+  geom_point() + 
+  stat_summary(fun=mean,
+               geom="line",
+               aes(group=1),
+               color = "blue",
+               size = 1) +
+  ylab(TeX("wpd_{glm} = wpd_{raw} - 1/(a  + b*log(nx*nfacet))"))
+
+# broom::tidy(glm_fit) %>% kable(caption = "Results of generalised linear model to capture the relationship between $wpd_{raw}$ and number of comparisons.")
+
+#ggsave("figs/G21_horizontal_m2.png")
+```
+
+```{r glm-horizontal-twogran1, fig.cap = "$wpd_{raw}$ (points) and residuals (blue line) from the model is plotted across different number of comparisons for $m=2$.
+Residuals seem to be independent of $nx$ and have been defined as $wpd_{glm}$."}
+knitr::include_graphics(here("img/G21_horizontal_m2.png"))
+```
+
+```{r wpd-glm-dist-appendix, fig.cap = "Distribution of $wpd_{glm}$ is plotted across different $nx$ for $m=2$.
+Both shape and scale of the distributions are different for lower $nx$ than for higher $nx$."}
+G21_glm <- G21 %>% 
+  mutate(wpd_glm =  (value - (1/(intercept + slope*log(nx*nfacet)
+  )
+  )
+  ),
+  wpd_glm_scaled = ((wpd_glm*320)))
+
+#G21_glm$wpd_glm_scaled %>% sd()
+
+G21_glm %>%
+  ggplot() +
+  geom_density(aes(x = wpd_glm),
+               fill = "blue") +
+  facet_grid(nx~nfacet,
+             labeller = "label_both") +
+  theme(legend.position = "bottom")
+
+```
+
+## Combination approach
+
+Figure \ref{fig:dist-same-scale-onegran} and \ref{fig:dist-new-same-scale-link} show the distribution of $wpd_{perm}$ and $wpd_{glm-scaled}$ overlaid on each other for $m=1$ and $m=2$ respectively.
+$wpd$ takes the value of $wpd_{perm}$ for lower levels
+$(<=5)$, and $wpd_{glm-scaled}$ for higher levels so that we have similar location and scale across panels with different levels.
+
+```{r dist-same-scale-onegran, fig.cap = "The distribution of $wpd_{perm}$ and $wpd_{glm-scaled}$ are overlaid to compare the location and scale across different $nx$ for $m=1$.
+The distribution of the adjusted measure looks similar for both approaches for higher levels, but for $nx=2$ both location and scale differs."}
+G21_permutation <- read_rds(here("data/hakear/simulations/supplementary/one-gran/norm/null_design_quantrans_nperm/data-agg/all_data_wpd_N01.rds")) %>%
+  rename("wpd_permutation" = "value")
+
+G21_all_data <- G21_permutation %>% 
+  # left_join(G21_lm, by = c("nx", "nfacet", "perm_id")) %>% 
+  left_join(G21_glm_m1, by = c("nx", "nfacet", "perm_id")) %>% 
+  pivot_longer(cols = c(3, 7),
+               names_to = "type_estimate",
+               values_to = "value_estimate")
+G21_all_data$type_estimate = factor(G21_all_data$type_estimate , levels = c( "wpd_permutation", "wpd_glm_scaled"))
+
+summary_data <- G21_all_data %>% 
+  group_by(nx, nfacet, type_estimate) %>% 
+  summarise(mean = mean(value_estimate)) %>% ungroup()
+
+G21_all_data %>% 
+  filter(type_estimate %in% c("wpd_glm_scaled", "wpd_permutation")) %>% 
+  ggplot(aes(x = value_estimate)) +
+  geom_density(aes(fill = type_estimate), alpha = 0.5, size = .5) +
+  geom_vline(data = summary_data, aes(xintercept = mean, color = type_estimate)) + 
+  geom_rug(aes(color = type_estimate), length = unit(0.09,"cm"), alpha = 0.5) +
+  #coord_cartesian(clip = "off") + 
+  facet_wrap(~nx,
+             labeller = "label_both") +
+  theme_bw() + 
+  theme(legend.position = "bottom") +
+  scale_fill_manual(values = c( "#D55E00", "#0072B2")) +
+  scale_color_manual(values = c( "#D55E00", "#0072B2")) +
+  xlab("adjusted values of wpd") 
+```
+
+```{r dist-new-same-scale-link, fig.cap = "The distribution of $wpd_{perm}$ and $wpd_{glm-scaled}$ are overlaid to compare the location and scale across different $nx$ and $nfacet$ for $m=2$.
+$wpd$ takes the value of $wpd_{perm}$ for lower levels, and $wpd_{glm-scaled}$ for higher levels.
+"}
+
+G21 <- read_rds(here("data/hakear/simulations/raw/null_design_quantrans/data-agg/all_data_wpd_N01.rds"))
+G21_median <- G21 %>% 
+  group_by(nx*nfacet) %>% 
+  summarise(actual = median(value))
+glm_fit <- glm(actual ~ log(`nx * nfacet`),
+               family = Gamma(link = "inverse"),
+               data = G21_median)
+intercept <- glm_fit$coefficients[1]
+slope <- glm_fit$coefficients[2]
+
+G21_glm <- G21_glm <- G21 %>% 
+  mutate(wpd_glm =  (value - (1/(intercept + slope*log(nx*nfacet)
+  )
+  )
+  ),
+  wpd_glm_scaled = ((wpd_glm*320)))
+
+G21_permutation <- read_rds(here("data/hakear/simulations/norm/null_design_quantrans_nperm/data-agg/all_data_wpd_N01.rds")) %>%
+  rename("wpd_permutation" = "value")
+
+# G21_model_data <- G21 %>%
+#   mutate(model =
+#            ((1/value)
+#                   - intercept -
+#                     slope*log(nx*nfacet))/slope) %>%
+#   mutate(model_trans =
+#            (model - mean(model))/sd(model))
+
+# G21_model_data$model %>% summary()
+
+G21_all_data <- G21_permutation %>% 
+  # left_join(G21_lm, by = c("nx", "nfacet", "perm_id")) %>% 
+  left_join(G21_glm, by = c("nx", "nfacet", "perm_id")) %>% 
+  pivot_longer(cols = c(3, 7),
+               names_to = "type_estimate",
+               values_to = "value_estimate")
+G21_all_data$type_estimate = factor(G21_all_data$type_estimate , levels = c( "wpd_permutation", "wpd_glm_scaled"))
+
+summary_data <- G21_all_data %>% 
+  group_by(nx, nfacet, type_estimate) %>% 
+  summarise(mean = mean(value_estimate)) %>% ungroup()
+
+G21_all_data %>% 
+  filter(type_estimate %in% c("wpd_glm_scaled", "wpd_permutation")) %>% 
+  ggplot(aes(x = value_estimate)) +
+  geom_density(aes(fill = type_estimate), alpha = 0.5, size = .5) +
+  geom_vline(data = summary_data, aes(xintercept = mean, color = type_estimate)) + 
+  geom_rug(aes(color = type_estimate), length = unit(0.09,"cm"), alpha = 0.5) +
+  #coord_cartesian(clip = "off") + 
+  facet_grid(nx~nfacet,
+             labeller = "label_both") +
+  theme_bw() + 
+  theme(legend.position = "bottom") +
+  scale_fill_manual(values = c( "#D55E00", "#0072B2")) +
+  scale_color_manual(values = c( "#D55E00", "#0072B2")) +
+  xlab("adjusted values of wpd") +
+  scale_x_continuous(breaks = c(-5, -3, 0, 3, 5))
+
+```
+
+# Ranking and selecting harmonies
+
+\noindent _Simulation design_
+
+Observations are generated from a N(0,1) distribution for each combination of $nx$ and $nfacet$ from the following sets: $nx = \{3, 7, 14\}$ and $nfacet = \{2, 9, 10\}$.
+This would result in 9 panels, viz, $(3, 2), (3, 9), (3, 10), \dots, (14,9), (14, 10)$.
+Few experiments were conducted.
+In the first scenario, data for all panels are simulated using the null design $D_{null}$.
+In other scenarios, data simulated from the panel $(14, 2)$ and $(3, 10)$ are under $D_{vary_{all}}$.
+Moreover, $\omega = \{0.5, 2, 5\}$ are considered to examine if the proposed test is able to capture subtle differences and non-subtle differences when we shift from the null design.
+In the last scenario, we consider the panel $(3, 2), (7,9), (14, 10)$ to be under $D_{null}$, the panels $(7, 2), (14, 9)$ to be under $D_{var_f}$.
+$(14, 2), (3, 10)$ under $D_{var_x}$ and the rest under $D_{var_{null}}$.
+This is done to check if the consequent ranking procedure leads to designs like $D_{vary_{f}}$ rank below $D_{vary_{all}}$.
+We generate only one data set each for which these scenarios were simulated and consider this as the original data set.
+We generate $1000$ repetitions of this experiment with different seeds.
+
+\noindent _Results_
+
+Figure \ref{fig:foo} shows the distribution of $wpd$ for the simulated $9$ null designs of different levels with each harmony being rejected if the corresponding $wpd$ value is higher than $wpd_{threshold99}$.
+Although the level of significance for each test is 1%, the level of significance for multiple tests together is 10% as displayed through the red line.
+This means that 10% of the times the null designs will be tagged as interesting even when there is actually nothing interesting going on.
+
+```{r}
+all_data <- read_rds("data/hakear/simulations/supplementary/test-hpc/data-agg/all_data.rds")
+
+# computes size for 99 percentile threshold
+all_data <- all_data %>%
+  filter(!is.na(x_levels), !is.na(facet_levels))
+
+data_summary <- all_data%>% 
+  mutate(sig = if_else(significance!=99, 0, 99)) %>% 
+  group_by(seed_id) %>% 
+  summarize(sumsig = sum(sig)) %>% 
+  count(sumsig==0) %>% 
+  mutate(p_value = n/sum(n)) %>% 
+  slice(1)
+
+data_summary_lowlev <- all_data%>% 
+  filter(x_levels==3, facet_levels ==2) %>% 
+  mutate(sig = if_else(significance!=99, 0, 99)) %>% 
+  group_by(seed_id) %>% 
+  summarize(sumsig = sum(sig)) %>% 
+  count(sumsig==0) %>% 
+  mutate(p_value = n/sum(n)) %>% 
+  slice(1)
+
+# plots
+
+global_threshold <-  tibble(threshold = "89.4", value = quantile(all_data$wpd, 0.894, na.rm = TRUE))
+```
+
+<!-- $wpd_{threshold99}$ is obtained as `r round(global_threshold$value, 2)` for this experiment, which leads to a p-value of `r round(data_summary$p_value, 3)`.
+Figure \ref{fig:global-size} shows the distribution of $wpd$ obtained from this experiment with the red line denoting $wpd_{threshold99}$.
+If we split the display of distribution between $nx$ and $nfacet$, then Figure \ref{fig:local-size} shows that the probability of rejecting the null when it is actually true is higher for smaller levels, however, it is still within limits (around `r round(data_summary_lowlev$p_value, 3)`).
+-->
+
+(ref:foo) The distribution of $wpd$ for 9 null designs of different levels with each harmony being rejected if the corresponding $wpd$ value is higher than $wpd_{threshold99}$.
+Although the level of significance for each test is 1%, the level of significance for multiple tests together is 10% as displayed through the red line.
+This means 10% of the times the null designs will be tagged as interesting even when there is actually nothing interesting going on.
+
+```{r foo, fig.cap='(ref:foo)'}
+
+ggplot() + 
+  geom_histogram(data = all_data, aes(x = wpd))  + 
+  geom_vline(data = global_threshold, aes(xintercept =  value),colour = "red", size = 1.5) 
+```
+
+```{r local-size, fig.cap = "something", eval = FALSE}
+ggplot() + 
+  geom_histogram(data = all_data, aes(x = wpd))  + 
+  geom_vline(data = global_threshold, aes(xintercept =  value),colour = "red", size = 1.5) +
+  facet_grid(x_levels~facet_levels)
+```
+
+```{r hist-qq-new-appendix, fig.cap = "In panel a, the histogram of $wpd_{glm-scaled}$ is plotted.
+In part b, the QQ plot is shown with the theoretical quantiles on the x-axis and $wpd_{glm-scaled}$ quantiles on the y-axis.
+The distribution looks symmetric and looks like normal except in the tails.", eval = FALSE}
+
+```
+
+```{r quadratic, fig.cap = "$wpd$ is plotted against $nx*nfacet$ (the maximum number of pairwise comparisons) and the blue line represents the median of the multiple values for each $nx*nfacet$.
+The median increases abruptly for lower values of $nx*nfacet$ and slowly for higher $nx*nfacet$.
+Thus, the measure will have higher values for higher levels in $nx$ or $nfacet$.", eval = FALSE}
+```
+
+# Raw time series plot
+
+\noindent To get a sense of how the raw time series data looks, we plot the energy usage for $50$ sampled households is plotted along the y-axis versus time from past to future in Figure ref\{fig:raw-data-50}.
+Each of these series is associated with a single customer.
+Energy consumption for each customer is given at fine temporal resolution (every 30 minutes) for a period of 2-3 years.
+
+```{r raw-data-50, out.width = "100%", fig.cap="The raw data for 50 households are shown.
+It looks like there is a lot of missing values and unequal length of time series along with asynchronous periods for which data is observed.
+No insightful behavioral pattern could be discerned from this view other than when the customer is not at home."}
+
+knitr::include_graphics("img/raw_plot_cust.png") # look at smart-meter.R for the code
+```

### 2021-11-07 data/hakear/simulations/tuning_param/aggregate01.R

@@ -1,0 +1,48 @@
+# contains aggregated data from the file simulations/results/raw/ and tidy them up
+
+library(tidyverse)
+library(here)
+library(readr)
+library(rlang)
+
+all_files = list.files(path = "simulations/results/raw/tuning_param", 
+                       pattern = ".rds")
+
+
+names_levels <- map_dfr(all_files, 
+                        function(x){
+                          z = str_split(str_remove(x, "_dist.rds"), "_") %>% 
+                            unlist()
+                          bind_cols(nx = as.numeric(z[1]),
+                                    nfacet = as.numeric(z[2]))
+                        })
+
+len_file = read_rds(("simulations/results/raw/tuning_param/2_2_tuning_param.rds"))
+
+names_rep <- names_levels %>% slice(rep(1:n(), each = nrow(len_file)))
+
+  
+all_files_path <- paste0("simulations/results/raw/tuning_param/",
+         all_files)  
+  
+
+all_data <- lapply(1:length(all_files_path), function(x){
+    
+  data = all_files_path %>% magrittr::extract2(x) %>% 
+    readRDS()
+  
+  names = names_levels %>% magrittr::extract(x,)
+  names_rep =   names %>% slice(rep(1:n(), each = nrow(data)))
+  bind_cols(names_rep, data)
+  }) %>% bind_rows() %>% 
+    arrange(nfacet, nx)
+    
+    
+
+# run for raw MMPD files aggregation
+#write_rds(all_data, "simulations/result_report/all_data_raw.rds")
+
+# run for raw max pairwise distances files aggregation
+write_rds(all_data, "simulations/tuning_param/all_data.rds")
+# 
+# write_rds(all_data, "simulations/result_report/all_data_norm_mmpd.rds")

### 2021-11-07 data/hakear/simulations/tuning_param/index.R

@@ -1,0 +1,82 @@
+library(hakear)
+library(tidyverse)
+library(parallel)
+library(distributional)
+source("simulations/functions_simulate_designs/vary_mean.R")
+
+mean = 0
+sd = 1
+range_w =  seq(1, 10, 1)
+range_lambda = seq(0.1, 0.9, 0.05)
+simtable<-read_csv(here::here('simulations/null/sim_table.csv'))
+
+
+### changing mean with constant sd
+
+set.seed(9999)
+
+data_varx <- mclapply(seq_len(nrow(simtable)),
+         function(j){
+           scen<- j 
+           simj<-simtable[scen,] #Extract row of table
+           nfacetj<-simj$nfacet # Which nfacet level
+           nxj<-simj$nx #Which nx level
+
+data_varx <- mclapply(range_lambda, function(x){
+  data_w <- mclapply(range_w, function(w){
+    #data_y <- mclapply(range_mean, function(y){   
+      # generate the data
+      data <- sim_panel(nx = nxj,
+                        nfacet = nfacetj,
+                        ntimes = 500,
+                        sim_dist = 
+                          sim_varx_normal_varymean(nx = nxj, nfacet = nfacetj, mean = mean, sd = sd, w = w)) %>%
+        unnest(c(data))
+      
+      # compute wpd
+      wpd <- compute_pairwise_max(data, 
+                                  gran_x = "id_x",
+                                  gran_facet = "id_facet",
+                                  response = sim_data, lambda = x)
+      
+    bind_cols(omega = w, wpd = wpd)
+  })  %>% bind_rows() 
+  
+  bind_cols(lambda = x, data_w)
+}) %>% bind_rows() 
+
+
+data_varf <- mclapply(range_lambda, function(x){
+  data_w <- mclapply(range_w, function(w){
+    #data_y <- mclapply(range_mean, function(y){   
+    # generate the data
+    data <- sim_panel(nx = nxj,
+                      nfacet = nfacetj,
+                      ntimes = 500,
+                      sim_dist = 
+                        sim_varf_normal_varymean(nx = nxj, nfacet = nfacetj, mean = mean, sd = sd, w = w)) %>%
+      unnest(c(data))
+    
+    # compute wpd
+    wpd <- compute_pairwise_max(data, 
+                                gran_x = "id_x",
+                                gran_facet = "id_facet",
[…]

### 2021-11-07 data/hakear/simulations/tuning_param/makegraph02.R

@@ -1,0 +1,101 @@
+library(here)
+library(readr)
+library(tidyverse)
+library(latex2exp)
+# run for raw MMPD files aggregation
+#all_data <- read_rds("simulations/result_report/all_data_raw.rds")
+# run for raw max pairwise distances files aggregation
+all_data <- read_rds(here::here("simulations/tuning_param/all_data.rds") )%>% 
+  mutate(design = case_when(
+    design == "vary_x" ~ "var_x",
+    design == "vary_facet" ~ "var_f"
+  ))
+
+  
+# for omega 8
+
+all_data_filtered <- all_data %>% 
+  filter(nx<=7, nfacet<=7)
+
+nxbyfacet8 <- all_data_filtered %>% 
+  filter(omega == 8) %>% 
+  ggplot(aes(x = lambda, y = wpd)) +
+  geom_line(aes(colour = design), size = 1) +
+  facet_grid(nx~nfacet,
+             labeller = "label_both") + 
+  theme(legend.position = "none")  +
+  scale_color_manual(values = c("#E69F00",
+                                "#56B4E9")) +
+  theme(legend.position = "bottom") +
+  ggtitle("(b)") + theme_bw() + xlab(TeX("$\\lambda$")) +
+  ylab("raw wpd")
+
+
+# ggsave(nxbyfacet, filename = here("simulations/tuning_param/figs/", "nxbyfacet_omega8.png"))
+
+# for omega 1
+nxbyfacet1 <- all_data_filtered %>% 
+  filter(omega == 1) %>% 
+  ggplot(aes(x = lambda, y = wpd)) +
+  geom_line(aes(colour = design), size = 1) +
+  facet_grid(nx~nfacet,
+             labeller = "label_both") + 
+  theme(legend.position = "none")  +
+  scale_color_manual(values = c("#E69F00", 
+                                "#56B4E9")) +
+  theme(legend.position = "bottom")  +
+  ggtitle("(a)") + theme_bw() +  xlab(TeX("$\\lambda$"))  +
+  ylab("raw wpd")
+# ggsave(nxbyfacet, filename = here("simulations/tuning_param/figs/", "nxbyfacet_omega1.png"))
+
+# 
+# get_legend<-function(myggplot){
+#   tmp <- ggplot_gtable(ggplot_build(myggplot))
+#   leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
[…]

### 2021-11-07 scripts/hakear-supplementary.R

@@ -1,0 +1,70 @@
+## ---- load
+library(knitr)
+library(tidyverse)
+library(lubridate)
+library(lvplot)
+#library(ggridges)
+library(tsibble)
+library(gravitas)
+library(ggpubr)
+library(readr)
+library(kableExtra)
+library(distributional)
+library(ggplot2)
+library(sugrrants)
+library(here)
+library(ggplot2)
+library(patchwork)
+library(scales)
+library(GGally)
+library(viridis)
+#remotes::install_github("Sayani07/hakear", force = TRUE)
+library(hakear)
+
+##----explain-design
+sim_varx_normal = function(nx, nfacet, mean, sd, w)
+{
+  rep(dist_normal((mean + seq(0, nx-1, by  = 1)*w), sd), nfacet)
+}
+
+sim_varf_normal = function(nx, nfacet, mean, sd, w)
+{
+  rep(dist_normal((mean + seq(0, nfacet-1, by  = 1)*w), sd), each = nx)
+}
+sim_varall_normal = function(nx, nfacet, mean, sd, w)
+{
+  dist_normal((mean + seq(0, (nx*nfacet - 1), by  = 1)*w), sd)
+}
+
+mean = 0
+sd = 1
+w = 3
+nx = 2
+nfacet = 3
+
+
+raw_levels = expand.grid(facet = c("$b_1$", "$b_2$", "$b_3$"),
+                         x = c("$a_1$", "$a_2$"))
+
+facet = rep(c("$b_1$", "$b_2$", "$b_3$"), each = nx)
+x =  rep(c("$a_1$", "$a_2$"), nfacet)
+
+null <-  sim_varall_normal(nx, nfacet, mean, sd, w=0) %>% tibble()
+vary_f <-  sim_varf_normal(nx, nfacet, mean, sd, w) %>% tibble()
+vary_x <-  sim_varx_normal(nx, nfacet, mean, sd, w) %>% tibble()
+vary_all <-  sim_varall_normal(nx, nfacet, mean, sd, w) %>% tibble()
+
+
+raw_table <- bind_cols(x = x, 
+                       facet = facet, 
+                       null = null,
+                       vary_f = vary_f, 
+                       vary_x = vary_x, 
+                       vary_all = vary_all) %>% 
+  tibble() 
+names(raw_table) = c("x levels", "facet levels", "$D_{null}$",  "$D_{var_f}$", "$D_{var_x}$", "$D_{var_{all}}$" )
+raw_table %>% 
+  kable(format = "latex",
[…]

### 2021-11-07 scripts/hakear.R

@@ -20,6 +20,8 @@ library(GGally)
 library(viridis)
 #remotes::install_github("Sayani07/hakear", force = TRUE)
 library(hakear)
+library(latex2exp)
+library(broom)
 
 #library(tidyquant)
 

### 2021-11-07 Rmd/01-intro.Rmd
Changing intro for the third time - pdf

@@
 
-This chapter provides an introduction to the study by first discussing the background and context, followed by the overarching research aim and objectives, each of which correspond to each chapter in this thesis.
+<!-- This chapter provides an introduction to the study by first discussing the background and context, followed by the overarching research aim and objectives, each of which correspond to each chapter in this thesis.
+-->
 
@@
 
-The Smart Grid, Smart City (SGSC) project (2010–2014) available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) provides half-hourly data of over 13,000 household electricity smart meters distributed unevenly from October 2011 to March 2014 for some parts in Australia.
-Because of the increasing variety of customers, larger data sets contain more uncertainty about customer behavior.
-Household behavioral patterns vary significantly due to differences in size, location, and amenities such as solar panels, central heating, and air conditioning.
+The Smart Grid, Smart City (SGSC) project (2010–2014) available through the [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) provides half-hourly data of over 13,000 Australian electricity smart meter customers distributed unevenly from October 2011 to March 2014.
+Large data sets like this include significant uncertainty regarding customer behavior due to the rising variety of customers.
+Behavioral patterns vary significantly due to differences in size, location, and amenities such as solar panels, central heating, and air conditioning.
 For example, some families use a dryer, while others hang their clothes to dry.
@@
 Customers' day-off energy consumption varies depending on whether they stay at home or go outside.
-The motivation for this work comes from the desire to provide methods to better understand heterogeneity in repetitive behavior that can arise across temporal dependencies.
 
@@
 With the availability of data at more and more finer time scales, exploration of time series data may be required to be carried out across both finer and coarser scales to draw useful inferences about the underlying process.
-For example, data collected at an hourly scale could be analyzed using coarser temporal scales such as days, months or quarters.
+To reduce the complexity of time, it is typical to divide it into years, months, weeks, days, and so on in a hierarchical manner [@aigner2011visualization].
+These discrete abstractions of time are known as time granularities.
+Linear time granularities [@Bettini1998-ed] ,such as hours, days, weeks and months, respect the linear progression of time and are non-repeating.
+Cyclic temporal granularities representing cyclical repetitions in time (such as hour-of-the-day, work-day/weekend) are effective for analyzing repetitive patterns in time series data.
+
+To acquire a comprehensive view of the repeated patterns, it is necessary to navigate through all of the conceivable cyclic granularities.
+This approach is consistent with the concept of EDA [@Tukey1977-jx], which stresses the utilization of multiple perspectives on data to assist formulate hypotheses before proceeding to hypothesis testing.
+This, however, is a challenging process since it throws up a myriad of alternatives.
+Furthermore, the transition from linear to cyclic granularities results in restructured data, with each level of the temporal deconstruction corresponding to multiple values of the observed variable.
+This motivates the research presented in this thesis, which aims to provide a platform for systematically exploring probability distributions induced by these multiple observations to support the discovery of regular patterns or anomalies, as well as the exploration of clusters of behaviors or the summarization of the behavior.
+
+<!-- There could potentially be many, many ways to analyze the data across cyclic granularities. -->
+
+<!-- For example, data collected at an hourly scale could be analyzed using coarser temporal scales such as days, months or quarters.
 This approach requires deconstructing time in various possible ways.
@@
 This idea aligns with the notion of EDA (Tukey 1977), which emphasizes the use of multiple perspectives on data to help formulate hypotheses before proceeding to hypothesis testing.
+-->
 
@@
 
-Slicing and dicing data in all possible temporal scales, as suggested by EDA, can be a daunting task because it opens up a plethora of possibilities.
-This motivates the research presented in this thesis, which aims to provide a platform for systematically exploring periodicities in temporal data and supporting the discovery of regular patterns or anomalies, as well as the exploration of clusters of behaviors or the summarization of the behavior.
-The transition from linear to cyclic granularities results in restructured data, with each level of the time deconstructions corresponding to multiple values of the observed variable.
-For example, it is common to see aggregated usage across households, such as total kWh used each half-hour by state, because energy companies need to know the maximum loads they must plan for.
-However, studying overall energy consumption obscures usage distributions at finer scales, making it more difficult to find solutions to improve energy efficiency.
-Summarizing the probability distribution of these multiple observations to capture both shape and uncertainty could be a potential approach to understanding the underlying distribution of these observations.
-The research objectives presented in each chapter aid in the achievement of this goal.
+<!-- My contribution in this thesis is to explore probability distributions induced by these multiple observations to better understand heterogeneity in repetitive behavior that can arise across temporal dependencies.
+-->
+
+<!-- Summarizing the probability distribution of these multiple observations to capture both shape and uncertainty could be a potential approach to understanding the underlying distribution of these observations.
+-->
 

### 2021-11-08 Rmd/01-intro.Rmd
changed chapter 1 intro

@@
 
-  Deconstructing a time index into time granularities can assist in exploration and automated analysis of large temporal data sets.
-This paper describes classes of time deconstructions using linear and cyclic time granularities.
-Linear granularities respect the linear progression of time such as hours, days, weeks and months.
-Cyclic granularities can be circular such as hour-of-the-day, quasi-circular such as day-of-the-month, and aperiodic such as public holidays.
-The hierarchical structure of granularities creates a nested ordering: hour-of-the-day and second-of-the-minute are single-order-up.
-Hour-of-the-week is multiple-order-up, because it passes over day-of-the-week.
-Methods are provided for creating all possible granularities for a time index.
-A recommendation algorithm provides an indication whether a pair of granularities can be meaningfully examined together (a "harmony"), or when they cannot (a "clash").
-
-   Time granularities can be used to create data visualizations to explore for periodicities, associations and anomalies.
-The granularities form categorical variables (ordered or unordered) which induce groupings of the observations.
-Assuming a numeric response variable, the resulting graphics are then displays of distributions compared across combinations of categorical variables.
-
-   The methods implemented in the open source R package `gravitas` are consistent with a tidy workflow, with probability distributions examined using the range of graphics available in `ggplot2`.
-
- <!-- This chapter contains a literature review related to linear granularities. -->
+Chapter \@ref(ch:gravitas) describes classes of time deconstructions using linear and cyclic time granularities, which can be used to create data visualizations to explore periodicities, associations, and anomalies.
+It provides a formal characterization of cyclic granularities which can be circular (such as hour-of-the-day), quasi-circular (such as day-of-the-month), and aperiodic (public holidays).
+It further facilitates manipulation of single- and multiple-order-up time granularities through cyclic calendar algebra, as well as provides a recommendation algorithm to check the feasibility of creating plots for any two cyclic granularities.
+Our proposed method is also applicable to non-temporal hierarchical granularities with an underlying ordered index.
+The methods are implemented in the open-source R package `gravitas` and are consistent with a tidy workflow [@wickham2016r], with probability distributions examined using the range of graphics available in `ggplot2`[@Wickham2009pk].
 
@@
 This work is a natural extension of Chapter \@ref(ch:gravitas).
-Many, many other displays might be built using different granularities, such as day-of-week, day-of-month, weekday/weekend, and so on.
+Many, many displays might be built using cyclic granularities.
 However, only a handful of them may be interesting, revealing major patterns.
@@
 Furthermore, a few of the displays in this collection will be more engaging than others.
-This chapter provides a new distance metric for selecting and ranking the multiple granularities.
+Chapter \@ref(ch:hakear) provides a new distance metric for selecting and ranking the multiple granularities.
 The statistical significance of potential visual discoveries is aided by selecting a threshold for the proposed numerical distance measure.
@@
 
-In this chapter, we look at the problem of using clustering to discover patterns in a large number of univariate time series across multiple temporal granularities.
+In Chapter \@ref(ch:gracsr), we look at the problem of using clustering to discover patterns in a large number of univariate time series across multiple temporal granularities.
 Time series clustering research is gaining traction as more data is collected at finer temporal resolution, over longer time periods, and for a larger number of individuals/entities.
@@
 This is available as the R package **hakear**.
-The chapter @ref (ch: gracsr) provides similarity measures for comparing multiple time series.This is in the developing R package **gracsr**.
+The chapter \@ref(ch:gracsr) provides methods to explore heterogeneity in repetitive behavior for multiple time series over multiple cyclic granularities.
+This is in the developing R package **gracsr**.
 Chapter \@ref(ch:conclusion) summarizes the software tools developed for the work, and discusses some future plans.

### 2021-11-08 Rmd/00-d-acknowlegements.Rmd
add Peter

@@
 
-Rob and Di are phenomenal leaders in their respective fields who lead by example.
+I am deeply grateful to my supervisors Rob J Hyndman and Dianne Cook, who are phenomenal leaders in their respective fields and lead by example.
 I am inspired by their creativity, wisdom, discipline, and dedication to contribute to the society through their research.
@@
 Thank you, Di, for being a fantastic female role model.
+I would also like to express my gratitude to my supervisor Dr.
+Peter Toscas of Data61 CSIRO, for always being available to help with technical issues and offering feedback on practical elements of analyzing smart meter data at our monthly meetings.
 
@@
  <!-- - PhD friends and R community -->
-Thank you, Stephanie and Nicholas Tierney, for being so warm, kind and welcoming when I began my PhD.
+Thank you, Stephanie Kobakian and Nicholas Tierney, for being so warm, kind and welcoming when I began my PhD.
 Being a part of such an inclusive environment meant a lot to me as an international student.

### 2021-11-08 Rmd/05-conclusion.Rmd
minor correction

@@
 Both \@ref(ch:gravitas) and \@ref(ch:hakear) are used for studying patterns in individual time series or comparing a few time series together.
-This is extended in Chapter @ref(ch:gracsr) to allow for the exploration of distributions for multiple time series at the same time using unsupervised clustering.
+This is extended in Chapter \@ref(ch:gracsr) to allow for the exploration of distributions for multiple time series at the same time using unsupervised clustering.
 The clustering methodology produces small groups of time series with similar distributions over multiple granularities.

### 2021-11-08 Rmd/03-hakear.Rmd
trying to make the gitbook work
gravitas paper bib correction

--- 2021-11-04 Rmd/03-hakear.Rmd
@@
 
-Cyclic temporal granularities [@Bettini1998-ed, @Gupta2021-hd] are <!--theoretical--> temporal deconstructions that define cyclic repetitions in time, e.g.
+Cyclic temporal granularities [@Bettini1998-ed], [@Gupta2021-gravitas] are <!--theoretical--> temporal deconstructions that define cyclic repetitions in time, e.g.
 hour-of-day, day-of-month, or regularly scheduled public holidays.
@@
 @Dang2014-tw and @wilkinson2005graph provide potential solutions to this, where few characterizations can be used to locate anomalies in density, shape, trend, and other features in the 2D point scatters.
-In this paper, we provide a solution to narrowing down the search from $^{N_C}P_2$ plots by introducing a new distance measure that can be used to detect significant distributional differences across cyclic granularities.<!-- harmonies and why it is not enough --> This work is a natural extension of our previous work [@Gupta2021-hd] that narrows down the search from $^{N_C}P_2$ plots by identifying pairs of granularities that can be meaningfully examined together (a "harmony"), or when they cannot (a "clash").
+In this paper, we provide a solution to narrowing down the search from $^{N_C}P_2$ plots by introducing a new distance measure that can be used to detect significant distributional differences across cyclic granularities.<!-- harmonies and why it is not enough --> This work is a natural extension of our previous work [@Gupta2021-gravitas] that narrows down the search from $^{N_C}P_2$ plots by identifying pairs of granularities that can be meaningfully examined together (a "harmony"), or when they cannot (a "clash").
 However, even after excluding clashes, the list of harmonies left could be enormous for exhaustive exploration.
@@
 
-<!-- Hence, with reference to the graphical design in @Gupta2021-hd, therefore the idea would be to rate a harmony pair higher if the variation between different levels of the x-axis variable is higher on an average across all levels of the facet variables.
+<!-- Hence, with reference to the graphical design in @Gupta2021-gravitas, therefore the idea would be to rate a harmony pair higher if the variation between different levels of the x-axis variable is higher on an average across all levels of the facet variables.
 -->
@@
 The set consisting of pairs of cyclic granularities ($C_{N_C}$) will have $7_{P_2}=42$ elements which could be analyzed for detecting possible periodicities.
-The set of possible harmonies $H_{N_C}$ from $C_{N_C}$ are chosen by removing clashes using procedures described in [@Gupta2021-hd].
+The set of possible harmonies $H_{N_C}$ from $C_{N_C}$ are chosen by removing clashes using procedures described in [@Gupta2021-gravitas].
 Table \ref{tab:tab-rank-8}  shows $14$ harmony pairs that belong to $H_{N_C}$.

### 2021-11-08 _render.R
docs instead of thesis to host through github instead of Netlify

@@ -40,6 +40,6 @@ if (length(formats) == 0) {
 for (fmt in formats)
   bookdown::render_book("index.Rmd", fmt, quiet = FALSE)
 
-gif_file <- list.files("figure", "*.gif", full.names = TRUE)
-invisible(file.copy(gif_file, "_thesis/figure/animate.gif"))
+# gif_file <- list.files("figure", "*.gif", full.names = TRUE)
+# invisible(file.copy(gif_file, "_thesis/figure/animate.gif"))
 cli_alert_success("You Rock!")

### 2021-11-08 Rmd/00-e-preface.Rmd
putting reference even if unnumbered
link update

@@
-# Preface {-}
+# Preface {-#ch:preface}
 
@@
 This thesis is written in R Markdown [@rmarkdown] with **bookdown** [@bookdown], using **renv** [@renv] to create reproducible environments.
-The online version of this thesis is hosted at <https://sayani.netlify.app/>, powered by [Netlify](http://netlify.com).
-All materials (including the data sets and source files) required to reproduce this document can be found at the Github repository <https://github.com/Sayani07/thesis>.
+The online version of this thesis is hosted at <https://sayani07.github.io/thesis-sg>.
+All materials (including the data sets and source files) required to reproduce this document can be found at the Github repository <https://github.com/Sayani07/thesis-SG>.
 

### 2021-11-09 Rmd/00-c-declaration.Rmd
running after updating R to 4.1.2

@@
 
-```{r signature, out.width = "20%", fig.align = "left", eval = file.exists("img/signature.png")}
+```{r signature, out.width = "20%", fig.align = "left", eval = FALSE}
 knitr::include_graphics("img/signature.png")
 ```
+
+<!-- eval = file.exists("img/signature.png") -->
 

### 2021-11-09 Rmd/06-appendix.Rmd
adding chapter 4 supplementary

@@
 \appendix
-
-This is the supplementary paper corresponding to the paper "Detecting distributional differences between temporal granularities for exploratory time series analysis"
 
@@
 read_chunk("scripts/hakear-supplementary.R")
+knitr::read_chunk('scripts/gracsr.R')
+knitr::read_chunk('scripts/gracsr-supplementary.R')
+read_chunk(here::here("validation/simdata01.R"))
 ```
@@
 
-# Recalling notations
+# Chapter 3 supplementary
+
+This is the supplementary paper corresponding to the paper "Detecting distributional differences between temporal granularities for exploratory time series analysis"
+
+## Recalling notations
 
@@
 
-# Raw weighted pairwise distance
-
-## Tuning parameter
+## Raw weighted pairwise distance
+
+### Tuning parameter
 
@@
 
-## Underlying distributions
+### Underlying distributions
 
@@
 
-## Number of comparisons
-
-### Case: m = 1
+### Number of comparisons
+
+#### Case: m = 1
 
@@
 
-### Case: m = 2
+#### Case: m = 2
 
@@
 
-# Adjusted weighted pairwise distances
+## Adjusted weighted pairwise distances
 
@@
 
-## Permutation approach
+### Permutation approach
 
@@
 
-## Modeling approach
+### Modeling approach
 
@@
 
-## Combination approach
+### Combination approach
 
@@
 
-# Ranking and selecting harmonies
+## Ranking and selecting harmonies
 
@@
 
-# Raw time series plot
-
-\noindent To get a sense of how the raw time series data looks, we plot the energy usage for $50$ sampled households is plotted along the y-axis versus time from past to future in Figure ref\{fig:raw-data-50}.
+# Chapter 4 supplementary
+
+```{r raw-data-50, out.width = "100%", fig.cap=" The raw half-hourly energy usage for $50$ sampled households is plotted along the y-axis versus time in a linear scale.
 Each of these series is associated with a single customer.
-Energy consumption for each customer is given at fine temporal resolution (every 30 minutes) for a period of 2-3 years.
-
-```{r raw-data-50, out.width = "100%", fig.cap="The raw data for 50 households are shown.
 It looks like there is a lot of missing values and unequal length of time series along with asynchronous periods for which data is observed.
@@
 
-knitr::include_graphics("img/raw_plot_cust.png") # look at smart-meter.R for the code
-```
+knitr::include_graphics("figs/raw_plot_cust.png") # look at smart-meter.R for the code
+```
+
+(ref:missing-data) Investigating the temporal location of missing values for customers who have implicit missing values.
+There are $13,735$ customers in the data set, with 8,685 having no missing values and the remaining $5,050$ having at least one missing value.
+Each cross represents a missed observation in time, while the line connecting two dots represents continuous missingness over time.
+Missing values occur at random times and do not appear to follow a pattern, although there is a higher concentration of missing values in September and October 2012 for the majority of customers.
+This plot is inspired by @wang2020tsibble.
+
+```{r missing-data, fig.cap="(ref:missing-data)"}
+knitr::include_graphics("figs/missing-data-5050.png")
+```
+
+_Prototype selection_
+
+S1. Robust scaling is applied to each customer.
+
+S2.
+$50^{th}$ percentile for each category for each granularity is obtained for each customers.
+So we have a data structure with  $356$ rows and $(24 + 12 + 2)$ variables corresponding to $50^{th}$ percentile for each hour-of-day, month-of-year and weekend-weekday.
+
+S3. Apply principal components and restrict the results down to the first six principal components (which makes up approximately 85% of the variance explained in the data) to use with the grand tour.
+
+S4.
+Run t-SNE using the default arguments on the complete data (sets the perplexity to equal 30 and performs random initialisation).
+We then create a linked tour with t-SNE layout with R package liminal.
+
+S5.
+We inspect of the subspace generated by the set of low-dimensional projections in tour by looking for a simplex shape while the visualization moves from one basis to another.
+When we brush the corners of the simplex, we find they fall on the edge of the t-SNE point cloud.
+
+<!-- Hall, Marron, and Neeman (2005) have shown that in the extreme case of high-dimension, low-sample size data, observations are on the vertices of a simplex. -->
+
+<!-- This is because in high-dimensional data analysis the curse of dimensionality reasons that points tend to be far away from the center of the distribution and on the edge of high-dimensional space.
+Contrary to this, is that projected data tends to clump at the center.
+-->
+
+S6. These points should ideally correspond to different behavior with respect to all the variables considered while running PCA.
+
+```{r mds, out.width="100%", fig.cap="One instance of brushing in tours (right) and projecting the points in a lower dimensional tsne cloud (left).", out.height="200px"}
+knitr::include_graphics("figs/tour-tsne.png")
+```
+
+```{r prototype-data-pick}
+```
+
+```{r assemble}
+```
+
+```{r data-pick}
+```
+
+```{r all-data}
+
+```
+
+```{r clustering}
+```
+
+```{r groups-24}
+
+```
+
+```{r tsne-fit}
+```
+
+```{r tsne-plot-supplementary, fig.cap="t-SNE summary of selected $24$ prototype customers.
+Each ellipse corresponds to a group after clustering using our methodology.
+Few customers are not enclosed inside an ellipse as they are from mixed groups and there are too few points to calculate an ellipse around them.
+It is important to note that the selection of prototypes is based on only $50$ percentiles, whereas, the clustering is based on all the deciles."}
+
+```
+
+<!-- From this set we select $4$ "anchor" customers which are far apart from each other and $5$ neighboring customers for each of these anchors.
+These selections were done using the granularity $hod$ space.
+It is important to note that when we use our proposed methodologies, it is based on all dimensions `hod`, `moy` and `wkndwday`.Fig \ref{fig:mds} shows the MDS of these 356 customers in a 2D space basis their distance on individual granularities and when all of them are combined.
+Our methodolgies are run on these $24$ customers, which act as a way to evaluate the proposed methodologies.
+-->
+
+<!-- .
+Instances may be distributed in the representation space in a reasonable manner, revealing their similarities, using this instance wise discriminative learning.
+A dual-level progressive similar instance selection (DPSIS) approach could also be used -->
+
+# Interaction of granularities
+
+Consider a case in which there are only two interacting granularities of interest, $g1$ and $g2$.
+In contrast to the previous situation, when we could study distributions across $n_{g_1} + n_{g_2} = 5$ separate categories, with interaction, we must evaluate the distribution of the $n_{g_1}*n_{g_2}=6$ combination of categories.
+Consider the $4$ designs in Figure \ref{fig:interaction-gran}, where various distributions are assumed for different combinations of categories, resulting in different designs.
+Design $D1$ exhibits no change in distributions across $g1$ or $g2$, whereas Designs $D2$ and $D3$ alter across only $g1$ and $g2$, respectively.
+D4 varies across both $g1$ and $g2$ categories.
+D3 and D4 appear similar based on their relative differences across consecutive categories, but $D4$ also changes across facets, unlike $D3$, which has all facets look the same.
+
+```{r interaction-gran, fig.cap = "The distribution of simulated variable across $g1$ conditional on $g2$ is shown through boxplots for 4 designs to extend the proposed validation designs when two granularities of interest interact.
+D1 has no change in distributions across different categories of $g1$ or $g2$, while D2 and D3 change across only $g1$ and $g2$ respectively.
+D4 changes across categories of both $g1$ and $g2$."}
+
+```
+
+```{r sim-val}
+set.seed(9999)
+nx_val = 2 # number of x-axis levels
+nfacet_val = 3 # number of facet levels
+w1_val = 3 # increment in mean
+w2_val = 0 # increment in sd
+mean_val = 0 # mean of normal distribution of starting combination
+sd_val = 1 # sd of normal distribution of starting combination
+quantile_prob_val = seq(0.1, 0.9, 0.1)
+```
+
+```{r nobs50}
+ntimes_val = 300 # nobs per combination 
+```
+
+<!-- # Clustering on simulated datasets -->
+
+```{r sample-seed-few}
+sample_seed <-  seq(10, 100, 10)
+```
+
+<!-- _DGP: Generate 10 time series from each designs_  -->
+<!-- Time series are simulated from each of these designs with $50$ observations in each group.
+So we have $50$ observations each for the six combination of categories $(1,1), (1, 2), (2, 1), (2, 2), (3, 1), (3, 2)$.
+Time series are simulated for ten different seeds (from `sample_seed`) for each design.
+`data_varall-1` represents a data set from a design $D_{var_{all}}$ (distributions change across both facet and x) with a seed $10$.
+`data_varall-2` represents a data set from a design $D_{var_{all}}$ with a seed $20$ and so on.
+-->
+<!-- `data_null`, `data_varx` and `data_varf `corresponds to designs $D_{null}$, $D_{var_{x}}$ and $D_{var_{f}}$ designs respectively.    -->
+
+```{r change-index-data-few}
+change_index_data <- function(ntimes_val = NULL,
+                         nx_val = NULL,
+                         nfacet_val = NULL,
+                         sim_function = sim_varx_normal){
+
+  data <- sim_panel(
+    nx = nx_val, nfacet =  nfacet_val,
+    ntimes = ntimes_val,
+    # sim_dist = sim_varx_normal(2, 3, 5, 10, 5, -1.5)
+    sim_dist = sim_function(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
+  ) %>% unnest(data)
+
+  index_new <- map(seq_len(ntimes_val), function(i){
+    map((seq_len(nx_val*nfacet_val)), function(j)
+    {
+      value = i + (j-1)*ntimes_val
+    })
+  }) %>% unlist()
+
+  data_new = data %>%
+    ungroup() %>%
+    mutate(index_old = row_number(),
+           index_new = index_new)
+
+  y = data_new[match(index_new, data_new$index_old),]
+
+  y <- y %>%
+    mutate(time = row_number()) %>%
+    select(-c(index_old, index_new))
+
+  return(y)
+}
+```
+
+```{r data-make-few}
+data_null <- map(sample_seed, function(seed){
+  set.seed(seed)
+  change_index_data(5, 2, 3, sim_null_normal) %>%
+    compute_quantiles(quantile_prob_val)
+}) %>%
+  bind_rows(.id = "seed_id") %>%
+  mutate(data_type = "data_null")
+
+data_varf <- map(sample_seed, function(seed){
+   set.seed(seed)
+  change_index_data(5, 2, 3, sim_varf_normal)%>%
+    compute_quantiles(quantile_prob_val)
+}) %>%
+  bind_rows(.id = "seed_id") %>%
+  mutate(data_type = "data_varf")
+
+data_varx <- map(sample_seed, function(seed){
+   set.seed(seed)
+  change_index_data(5, 2, 3, sim_varx_normal)%>%
+    compute_quantiles(quantile_prob_val)
+}) %>%
+  bind_rows(.id = "seed_id") %>%
+  mutate(data_type = "data_varx")
+
+data_varall <- map(sample_seed, function(seed){
+   set.seed(seed)
+  change_index_data(5, 2, 3, sim_varall_normal)%>%
+    compute_quantiles(quantile_prob_val)
+}) %>%
+  bind_rows(.id = "seed_id") %>%
+  mutate(data_type = "data_varall")
+
+data_q <- bind_rows(data_null,
+                      data_varf,
+                      data_varx,
+                      data_varall) %>%
+  mutate(seed_id = sprintf("%02d", 
+                              as.numeric(seed_id))) %>% 
+  mutate(unique_data = paste(data_type, seed_id, sep = "-"))
+
+data_q$data_type <- factor(data_q$data_type,
+                                    levels = c("data_null" , "data_varf",
+                                               "data_varx","data_varall"))
+
+```
+
+```{r JS}
+
+```
+
+```{r JS-mydata-few}
+data_q_wide <- data_q %>%
+  select(id_facet, id_x, unique_data, sim_data_quantile) %>%
+  pivot_wider(names_from = unique_data,
+              values_from = sim_data_quantile) %>%
+  select(-c(1, 2))
+
+ndata <- data_q %>% distinct(unique_data) %>% mutate(index = row_number())
+ldata <- nrow(ndata)
+lcomb <-  nx_val*nfacet_val
+
+dist_data <- map(1:ldata, function(x){ # first data
+  map(1:ldata, function(y){ # 2nd data
+    map(1:lcomb, function(z){ # number of combinations nx*nfacet
+      JS(
+        prob = quantile_prob_val,
+        unlist(data_q_wide[z,x]),
+        unlist(data_q_wide[z,y])
+      ) %>% as_tibble()
+    })%>% bind_rows(.id = "combinations")
+  })%>% bind_rows(.id = "data_type1")
+}) %>% bind_rows(.id = "data_type2")
+
+dist_mat <- dist_data %>%
+  group_by(data_type1, data_type2) %>%
+  summarise(dist = sum(value)) %>%
+  pivot_wider(names_from = data_type2,
+              values_from = dist) %>%
+  mutate(data_type1 = as.numeric(data_type1)) %>%
+  left_join(ndata, by = c("data_type1" = "index"))
+
+dist_mat_format <- dist_mat %>%
+  ungroup() %>%
+  select(-data_type1, -unique_data)
+
+rownames(dist_mat_format) <- dist_mat$unique_data
+# write_rds(dist_mat_format, "validation/interact-4D/distmat_40series.rds")
+# write_rds(dist_mat, "validation/interact-4D/distmat_40series_withnames.rds")
+```
+
+```{r hier-clust-few, fig.cap= "A dendogram showing the branching of $40$ series with each of the $4$ designs repeated $10$ times."}
+d = stats::as.dist(dist_mat_format)
+par(cex=0.5)
+hc = stats::hclust(d,method="complete")
+plot(hc)
+rect.hclust(hc, k = 4, border = "red")
+
+```
+
+```{r mds-plot-interact-few, fig.cap= "MDS summary plot of $40$ series with each of the $4$ designs repeated $10$ times.
+It can be observed that all the series belonging to same design lie closer in this plot and all the ones in different designs are placed further away."}
+mds <- d %>%
+  cmdscale() %>%
+  as_tibble()
+
+colnames(mds) <- c("Dim.1", "Dim.2")
+rownames(mds) <- dist_mat$unique_data
+
+groups<-cutree(hc, k=4)
+
+all_data_cluster <- cbind(dist_mat_format, groups) %>%
+  cbind(mds)%>%
+  mutate(groups = as.factor(groups)) %>% as_tibble()
+
+mds_plot_10 <- ggplot(all_data_cluster,
+       aes(x = Dim.1,
+           y = Dim.2,
+       color = groups)) +
+  geom_text(label = rownames(mds),check_overlap = TRUE)+
+  geom_point(size = 1) +
+  theme_classic()+
+  theme(legend.position = "bottom") +
+  scale_color_brewer(palette = "Dark2") +
+  coord_fixed(ratio = 1)+
+  theme(text = element_text(size = 6))
+
+mds_plot_10
+```
+
+<!-- # Clustering on simulated datasets -->
+
+```{r sample-seed}
+sample_seed <-  seq(1, 100, 1)
+```
+
+<!-- _DGP: Generate 10 time series from each designs_  -->
+<!-- Time series are simulated from each of these designs with $50$ observations in each group.
+So we have $50$ observations each for the six combination of categories $(1,1), (1, 2), (2, 1), (2, 2), (3, 1), (3, 2)$.
+Time series are simulated for ten different seeds (from `sample_seed`) for each design.
+`data_varall-1` represents a data set from a design $D_{var_{all}}$ (distributions change across both facet and x) with a seed $10$.
+`data_varall-2` represents a data set from a design $D_{var_{all}}$ with a seed $20$ and so on.
+-->
+<!-- `data_null`, `data_varx` and `data_varf `corresponds to designs $D_{null}$, $D_{var_{x}}$ and $D_{var_{f}}$ designs respectively.    -->
+
+```{r change-index-data}
+change_index_data <- function(ntimes_val = NULL,
+                         nx_val = NULL,
+                         nfacet_val = NULL,
+                         sim_function = sim_varx_normal){
+
+  data <- sim_panel(
+    nx = nx_val, nfacet =  nfacet_val,
+    ntimes = ntimes_val,
+    # sim_dist = sim_varx_normal(2, 3, 5, 10, 5, -1.5)
+    sim_dist = sim_function(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
+  ) %>% unnest(data)
+
+  index_new <- map(seq_len(ntimes_val), function(i){
+    map((seq_len(nx_val*nfacet_val)), function(j)
+    {
+      value = i + (j-1)*ntimes_val
+    })
+  }) %>% unlist()
+
+  data_new = data %>%
+    ungroup() %>%
+    mutate(index_old = row_number(),
+           index_new = index_new)
+
+  y = data_new[match(index_new, data_new$index_old),]
+
+  y <- y %>%
+    mutate(time = row_number()) %>%
+    select(-c(index_old, index_new))
+
+  return(y)
+}
+```
+
+```{r data-make}
+data_null <- map(sample_seed, function(seed){
+  set.seed(sample_seed[seed])
+  change_index_data(5, 2, 3, sim_null_normal) %>%
+    compute_quantiles(quantile_prob_val)
+}) %>%
+  bind_rows(.id = "seed_id") %>%
+  mutate(data_type = "data_null")
+
+data_varf <- map(sample_seed, function(seed){
+  set.seed(sample_seed[seed])
+  change_index_data(5, 2, 3, sim_varf_normal)%>%
+    compute_quantiles(quantile_prob_val)
+}) %>%
+  bind_rows(.id = "seed_id") %>%
+  mutate(data_type = "data_varf")
+
+data_varx <- map(sample_seed, function(seed){
+  set.seed(sample_seed[seed])
+  change_index_data(5, 2, 3, sim_varx_normal)%>%
+    compute_quantiles(quantile_prob_val)
+}) %>%
+  bind_rows(.id = "seed_id") %>%
+  mutate(data_type = "data_varx")
+
+data_varall <- map(sample_seed, function(seed){
+  set.seed(sample_seed[seed])
+  change_index_data(5, 2, 3, sim_varall_normal)%>%
+    compute_quantiles(quantile_prob_val)
+}) %>%
+  bind_rows(.id = "seed_id") %>%
+  mutate(data_type = "data_varall")
+
+data_q <- bind_rows(data_null,
+                      data_varf,
+                      data_varx,
+                      data_varall) %>%
+  mutate(seed_id = sprintf("%02d", 
+                              as.numeric(seed_id))) %>% 
+  mutate(unique_data = paste(data_type, seed_id, sep = "-"))
+
+data_q$data_type <- factor(data_q$data_type,
+                                    levels = c("data_null" , "data_varf",
+                                               "data_varx","data_varall"))
+
+```
+
+```{r JS-mydata}
+data_q_wide <- data_q %>%
+  select(id_facet, id_x, unique_data, sim_data_quantile) %>%
+  pivot_wider(names_from = unique_data,
+              values_from = sim_data_quantile) %>%
+  select(-c(1, 2))
+
+ndata <- data_q %>% distinct(unique_data) %>% mutate(index = row_number())
+ldata <- nrow(ndata)
+lcomb <-  nx_val*nfacet_val
+
+dist_data <- map(1:ldata, function(x){ # first data
+  map(1:ldata, function(y){ # 2nd data
+    map(1:lcomb, function(z){ # number of combinations nx*nfacet
+      JS(
+        prob = quantile_prob_val,
+        unlist(data_q_wide[z,x]),
+        unlist(data_q_wide[z,y])
+      ) %>% as_tibble()
+    })%>% bind_rows(.id = "combinations")
+  })%>% bind_rows(.id = "data_type1")
+}) %>% bind_rows(.id = "data_type2")
+
+dist_mat <- dist_data %>%
+  group_by(data_type1, data_type2) %>%
+  summarise(dist = sum(value)) %>%
+  pivot_wider(names_from = data_type2,
+              values_from = dist) %>%
+  mutate(data_type1 = as.numeric(data_type1)) %>%
+  left_join(ndata, by = c("data_type1" = "index"))
+
+dist_mat_format <- dist_mat %>%
+  ungroup() %>%
+  select(-data_type1, -unique_data)
+
+rownames(dist_mat_format) <- dist_mat$unique_data
+# write_rds(dist_mat_format, "validation/interact-4D/distmat_400series.rds")
+# write_rds(dist_mat, "validation/interact-4D/distmat_400series_withnames.rds")
+```
+
+```{r hier-clust, fig.cap = "A dendogram showing the branching of $400$ series with each of the $4$ designs repeated $100$ times."}
+d = stats::as.dist(dist_mat_format)
+par(cex=0.5)
+hc = stats::hclust(d,method="complete")
+plot(hc)
+rect.hclust(hc, k = 4, border = "red")
+```
+
+```{r mds-plot-interact, fig.cap= "MDS summary plot of $400$ series with each of the $4$ designs repeated $100$ times.
+It can be observed that all the series belonging to same design lie closer in this plot and all the ones in different designs are placed further away."}
+mds <- d %>%
+  cmdscale() %>%
+  as_tibble()
+
+colnames(mds) <- c("Dim.1", "Dim.2")
+rownames(mds) <- dist_mat$unique_data
+
+groups<-cutree(hc, k=4)
+
+all_data_cluster <- cbind(dist_mat_format, groups) %>%
+  cbind(mds)%>%
+  mutate(groups = as.factor(groups)) %>% as_tibble()
+
+mds_plot_10 <- ggplot(all_data_cluster,
+       aes(x = Dim.1,
+           y = Dim.2,
+       color = groups)) +
+  #geom_text(label = rownames(mds),check_overlap = TRUE)+
+  geom_point(size = 0.5) +
+  theme_classic()+
+  theme(legend.position = "bottom") +
+  scale_color_brewer(palette = "Dark2") +
+  coord_fixed(ratio = 1)
+
+mds_plot_10
+```
+
+```{r heatmap-interact, fig.height=8, fig.cap="A heatmap showing the cluster summary of $400$ series.
+Deciles of the clusters are plotted along the y-axis with each group on the x-axis faceted by each combination of the interacting granularity.
+The four distinct behavior across clusters are evident in all the combinations, with the first block corresponding to $data_null$.
+The deciles of the second block changes across both facet and x-axis.
+For similar reasons, the second, third and forth blocks correspond $data_varall$, $data_varf$ and $data_varx$ respectively." }
+heatmap_raw <- data_q %>% 
+  #select(unique_data, sim_data_quantile, id_facet, id_x) %>% 
+  mutate(category = paste0("facet= ", id_facet, ",x= ", id_x)) %>% 
+  unnest(sim_data_quantile) %>% 
+  group_by(unique_data, category) %>% 
+  mutate(D = row_number()) %>% 
+  arrange(D, category) %>% 
+  left_join(dist_mat, by = c("unique_data")) %>% 
+  select(c(2:9))
+
+data2 <- cbind(dist_mat_format, groups)
+
+heatmap_final <- bind_cols(serial_data = rownames(data2) %>%
+                             as_tibble(),
+          group = data2$groups) %>%
+  left_join(heatmap_raw, by = c("value" = "unique_data")) %>% select(-id_facet, - id_x, - data_type1) %>% 
+  arrange(D, category, group)
+
+heatmap_final$data_type <- factor(heatmap_final$data_type,
+                                   levels = c("data_null" , "data_varf",
+                                              "data_varx","data_varall"))
+
+heatmap_final$category <- reorder(heatmap_final$category, heatmap_final$data_type)
+
+ggplot(heatmap_final,
+       aes(x= reorder(value, data_type), y = as.factor(D))) +
+  geom_tile(aes(fill = sim_data_quantile)) +
+  scale_fill_distiller(palette = "YlGnBu", direction = 1)+ 
+  facet_wrap(~category, ncol = 1) +
+  theme(axis.text.x = element_text(angle = 90)) +
+  ylab("Deciles") +
+  xlab("data sets") +
+  theme(legend.position = "bottom")+
+  scale_x_discrete(breaks = c("data_null" , "data_varf",
+                                              "data_varx","data_varall"))
+
+```
+
+_Clustering all 350 customers_
+
+```{r data-load-entire}
+# Read the nqt distances
+
+wkndwday <- read_rds(here("data/dist_gran_wkndwday_356cust_nqt.rds")) %>% broom::tidy()
+
+moy <- read_rds(here("data/dist_gran_moy_356cust_nqt.rds")) %>% broom::tidy()
+
+hod <- read_rds(here("data/dist_gran_hod_356cust_nqt.rds")) %>% broom::tidy()
+
+# Make the distance metrics
+
+distance <- wkndwday %>% 
+  left_join(moy, by = c("item1", "item2")) %>% 
+  left_join(hod, by = c("item1", "item2")) %>% 
+  rename("wkndwday" ="distance.x",
+         "moy" = "distance.y",
+         "hod" = "distance") %>%
+  mutate(item1 = as.integer(as.character(item1)),
+         item2 = as.integer(as.character(item2)))
+
+filtered_distance <- distance %>%
+  filter(!(item1 %in% c(8196183, 8508008, 8680538))) %>% 
+  filter(!(item2 %in% c(8196183, 8508008, 8680538)))
+
+total_distance <- filtered_distance %>% 
+  mutate(total = wkndwday/2 + moy/12 + hod/24)
+
+total_distance_wide <- total_distance %>% pivot_wider(-c(2:5), 
+                                                      names_from = item2,
+                                                      values_from = total)
+
+rownames(total_distance_wide) <- total_distance_wide$item1
+
+mds_data <- total_distance_wide %>% 
+  mutate_all(~replace(., is.na(.), 0)) %>%
+  tibble::rownames_to_column() %>%  
+  dplyr::select(-item1) %>% 
+  pivot_longer(-rowname) %>% 
+  pivot_wider(names_from=rowname, values_from=value)
+
+rownames(mds_data) <- total_distance_wide$item1
+
+df <- mds_data[-1] %>% as.matrix()
+DM <- matrix(0, ncol(mds_data), ncol(mds_data))
+DM[lower.tri(DM)] = df[lower.tri(df, diag=TRUE)] # distance metric
+f = as.dist(DM)
+
+first_lot <- mds_data %>% names()
+
+id <- c(first_lot[-1], mds_data$name[nrow(mds_data)])
+```
+
+```{r opt-clust-350, fig.cap = "Cluster separation for the 353 customers across the number of clusters is shown.
+When the cluster size changes from 17 to 18, the separation index drops sharply and then flattens out, resulting in the appearance of the elbow.
+Hence, when grouping the 353 customers, the number of clusters is taken to be 17."}
+
+# Find optimal number of clusters
+
+k = array()
+for(i in 5:50)
+{
+  group <- f %>% hclust (method = "ward.D") %>% cutree(k=i)
+  p <- cluster.stats(f, clustering = group, silhouette = TRUE)
+  k[i]=p$sindex
+}
+
+ggplot(k %>% as_tibble %>% mutate(k = row_number()), aes(x=k, y = value)) + geom_line() + scale_x_continuous(breaks = seq(2, 50, 2), minor_breaks = 1) + xlab("number of clusters") + ylab ("separation index")
+```
+
+```{r group-353}
+
+# plot(k, type = "l")
+# 6 coming as the number of clusters with maximum silwidth
+
+group <- f %>% hclust (method = "ward.D") %>% cutree(k=17)
+
+cluster_result <- bind_cols(customer_id = id, group = group) %>%
+  arrange(group)
+# %>% 
+#   mutate(divide_cust = c(rep(1, 177), rep(2, 176)))
+
+# cluster_result %>% group_by(group) %>% count()
+
+```
+
+```{r data-heatmap-hod-group-entire}
+legend_title <- "group"
+
+data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>%
+  mutate(customer_id = as.character(customer_id)) %>%
+  gracsr::scale_gran( method = "robust",
+                      response = "general_supply_kwh")
+
+data_group <- data_pick  %>% 
+  mutate(customer_id = as.character(customer_id)) %>% 
+   gracsr::scale_gran( method = "robust",
+                       response = "general_supply_kwh") %>% 
+  left_join(cluster_result, by = c("customer_id"))
+
+data_heatmap_hod_group <- quantile_gran(data_group,
+                                  gran1="hour_day",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)
+
+# data_heatmap_hod_group$group <- paste("group", data_heatmap_hod_group$group, sep = "-")
+
+hod_group_entire1 <- data_heatmap_hod_group %>% 
+  filter(group<9) %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`,
+                  group=group,
+                  fill = as.factor(group), alpha = 0.5),
+              alpha = 0.5) +
+  geom_line(aes(y = `50%`,
+                group=group, 
+                color = as.factor(group)), size = 1)+
+  facet_wrap(~group, 
+             scales = "free_y",  
+             nrow = 8,labeller = "label_both") + 
+              #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + 
+  ylab("demand (in Kwh)") + 
+  theme_bw()  +
+  scale_x_discrete(breaks = seq(1, 24, 3))+ 
+  #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme_application() +
+  scale_fill_manual(values=as.vector(okabe(n = 8)))+
+  scale_color_manual(values=as.vector(okabe(n = 8)))+
+  theme(legend.position = "bottom")
+
+hod_group_entire2 <- data_heatmap_hod_group %>% 
+  filter(group>=9) %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`,
+                  group=group,
+                  fill = as.factor(group), alpha = 0.5),
+              alpha = 0.5) +
+  geom_line(aes(y = `50%`,
+                group=group, 
+                color = as.factor(group)), size = 1)+
+  facet_wrap(~group, 
+             scales = "free_y",  
+             nrow = 9,labeller = "label_both") + 
+              #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + 
+  ylab("demand (in Kwh)") + 
+  theme_bw()  +
+  scale_x_discrete(breaks = seq(1, 24, 3))+ 
+  #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
+  theme_application() +
+  scale_fill_manual(values=as.vector(tableau20(20)))+
+  scale_color_manual(values=as.vector(tableau20(20)))+
+  theme(legend.position = "bottom")
+
+```
+
+```{r data-heatmap-moy-group-entire}
+data_heatmap_moy_group <- quantile_gran(data_group,
+                                  gran1="month_year",
+                                  quantile_prob_val = c(0.25, 0.5, 0.75),
+                                  group="group") %>% 
+  pivot_wider(names_from = quantiles, values_from = quantiles_values)
+
+data_heatmap_moy_group$category <- factor(data_heatmap_moy_group$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
+
+# data_heatmap_moy_group$group <- paste("group", data_heatmap_moy_group$group, sep = "-")
+
+moy_group_entire1 <- data_heatmap_moy_group %>% 
+  filter(group<9) %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
+  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
+  facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_both",
+             nrow = 8) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
+  ylab("demand (in Kwh)") +
+  theme_bw() + theme_application() +
+  scale_fill_manual(values=as.vector(okabe(n = 8)))+
+  scale_color_manual(values=as.vector(okabe(n = 8)))+
+  theme(legend.position = "bottom")
+
+moy_group_entire2 <- data_heatmap_moy_group %>% 
+  filter(group>=9) %>% 
+  ggplot(aes(x = category)) + 
+  geom_ribbon(aes(ymin = `25%`, 
+                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
+  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
+  facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_both",
+             nrow =  9) +
+    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
+  ylab("demand (in Kwh)") +
+  theme_bw() + theme_application() +
+  scale_fill_manual(values=as.vector(tableau20(20)))+
+  scale_color_manual(values=as.vector(tableau20(20)))+
+  theme(legend.position = "bottom")
+```
+
+```{r data-wnwd-group-entire}
+wkndwday_data <- data_group %>%
+  create_gran("wknd_wday") 
+# wkndwday_data$group <- as.factor(wkndwday_data$group)
+# %>% 
+#   create_gran("hour_day")
+
+ylim1 = boxplot.stats(wkndwday_data$general_supply_kwh)$stats[c(1, 5)]
+
+wkndwday_group_entire1 <- wkndwday_data%>% 
+  filter(group<9) %>% 
+  mutate(group = as.factor(group)) %>% 
+  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
+  #lvplot::geom_lv(aes(fill = as.factor(group)), k=5) +
+  geom_boxplot(aes(fill = group, color = group),alpha = 0.5, outlier.alpha = 0.05)+
+  #geom_boxplot(outlier.size = 1) + 
+  coord_cartesian(ylim = ylim1*1.05)+
+  #ggridges::geom_density_ridges2(aes(x = general_supply_kwh, y = wknd_wday,fill = as.factor(group))) + coord_flip() +
+#geom_boxplot(aes(fill = as.factor(group))) +
+  #scale_fill_lv() +
+ xlab("wknd-wday") + 
+  ylab("demand (in Kwh)") +
+   facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_both",
+             nrow = 8) + 
+  theme_bw() + theme_application() +
+  scale_fill_manual(values=as.vector(okabe(n = 8)))+
+  scale_color_manual(values=as.vector(okabe(n = 8)))+
+  theme(legend.position = "none")
+
+wkndwday_group_entire2 <- wkndwday_data%>% 
+  filter(group>=9) %>% 
+  mutate(group = as.factor(group)) %>% 
+  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
+  #lvplot::geom_lv(aes(fill = as.factor(group)), k=5) +
+  geom_boxplot(aes(fill = group, color = group),alpha = 0.5, outlier.alpha = 0.05)+
+  #geom_boxplot(outlier.size = 1) + 
+  coord_cartesian(ylim = ylim1*1.05)+
+  #ggridges::geom_density_ridges2(aes(x = general_supply_kwh, y = wknd_wday,fill = as.factor(group))) + coord_flip() +
+#geom_boxplot(aes(fill = as.factor(group))) +
+  #scale_fill_lv() +
+ xlab("wknd-wday") + 
+  ylab("demand (in Kwh)") +
+   facet_wrap(~group, 
+             scales = "free_y", 
+             labeller = "label_both",
+             nrow = 9) + 
+  theme_bw() + theme_application() +
+  scale_fill_manual(values=as.vector(tableau20(20)))+
+  scale_color_manual(values=as.vector(tableau20(20)))+
+  theme(legend.position = "none")
+```
+
+```{r combined-groups-js-entire, fig.cap = "The distribution of electricity demand for the clusters across hod, moy and wkndwday for the $17$ groups from $353$ customers.
+Wknd-wday variations across groups are not distinguishable, but ideally each group should have an unique combination of hod and moy.", fig.height=6}
+
+group_entire1 <- (hod_group_entire1 + moy_group_entire1 + wkndwday_group_entire1)&theme_characterisation()& theme(legend.position = 'none')
+
+group_entire2 <- (hod_group_entire2 + moy_group_entire2 + wkndwday_group_entire2)&theme_characterisation()& theme(legend.position = 'none')
+
+ggpubr::ggarrange(group_entire1, group_entire2)
+ #  
+ #  
+ #  plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')')+
+ #  plot_layout(guides = "collect")& theme(legend.position = 'none')
+ # 
+ # 
+ # (hod_group_entire/moy_group_entire/ wkndwday_group_entire) +
+ #  plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')')+
+ #  plot_layout(guides = "collect")& theme(legend.position = 'none')
+```

### 2021-11-10 Rmd/06-appendix.Rmd
adding chapter 4 supplementary material

@@
 knitr::read_chunk('scripts/gracsr-supplementary.R')
-read_chunk(here::here("validation/simdata01.R"))
 ```
@@
 
-# Chapter 3 supplementary
+# Supplementary: Chapter 3
 
@@
 
-# Chapter 4 supplementary
+# Supplementary: Chapter 4
+
+## Raw data plot
 
@@
 It looks like there is a lot of missing values and unequal length of time series along with asynchronous periods for which data is observed.
-No insightful behavioral pattern could be discerned from this view other than when the customer is not at home."}
-
-knitr::include_graphics("figs/raw_plot_cust.png") # look at smart-meter.R for the code
-```
+No insightful behavioral pattern could be discerned from this view other than when the customer is not at home.", fig.pos="h"}
+
+knitr::include_graphics("img/raw_plot_cust.png") # look at smart-meter.R for the code
+```
+
+## Missing data plot
 
@@
 
-```{r missing-data, fig.cap="(ref:missing-data)"}
-knitr::include_graphics("figs/missing-data-5050.png")
-```
-
-_Prototype selection_
+```{r missing-data, fig.cap="(ref:missing-data)", fig.pos="h"}
+knitr::include_graphics("img/missing-data-5050.png")
+```
+
+## Prototype selection
 
@@
 S4.
-Run t-SNE using the default arguments on the complete data (sets the perplexity to equal 30 and performs random initialisation).
+Run t-SNE using the default arguments on the complete data (sets the perplexity to equal 30 and performs random initialization).
 We then create a linked tour with t-SNE layout with R package liminal.
@@
 ```{r mds, out.width="100%", fig.cap="One instance of brushing in tours (right) and projecting the points in a lower dimensional tsne cloud (left).", out.height="200px"}
-knitr::include_graphics("figs/tour-tsne.png")
+knitr::include_graphics("img/tour-tsne.png")
 ```
@@
 
-# Interaction of granularities
+## Interaction of granularities
 
@@
 In contrast to the previous situation, when we could study distributions across $n_{g_1} + n_{g_2} = 5$ separate categories, with interaction, we must evaluate the distribution of the $n_{g_1}*n_{g_2}=6$ combination of categories.
-Consider the $4$ designs in Figure \ref{fig:interaction-gran}, where various distributions are assumed for different combinations of categories, resulting in different designs.
+Consider the $4$ designs in Figure \ref{fig:interaction-gran-supplement}, where various distributions are assumed for different combinations of categories, resulting in different designs.
 Design $D1$ exhibits no change in distributions across $g1$ or $g2$, whereas Designs $D2$ and $D3$ alter across only $g1$ and $g2$, respectively.
@@
 
-```{r interaction-gran, fig.cap = "The distribution of simulated variable across $g1$ conditional on $g2$ is shown through boxplots for 4 designs to extend the proposed validation designs when two granularities of interest interact.
+```{r interaction-gran-supplement, fig.cap = "The distribution of simulated variable across $g1$ conditional on $g2$ is shown through boxplots for 4 designs to extend the proposed validation designs when two granularities of interest interact.
 D1 has no change in distributions across different categories of $g1$ or $g2$, while D2 and D3 change across only $g1$ and $g2$ respectively.
@@
 ```{r JS-mydata-few}
-data_q_wide <- data_q %>%
-  select(id_facet, id_x, unique_data, sim_data_quantile) %>%
-  pivot_wider(names_from = unique_data,
-              values_from = sim_data_quantile) %>%
-  select(-c(1, 2))
-
-ndata <- data_q %>% distinct(unique_data) %>% mutate(index = row_number())
-ldata <- nrow(ndata)
-lcomb <-  nx_val*nfacet_val
-
-dist_data <- map(1:ldata, function(x){ # first data
-  map(1:ldata, function(y){ # 2nd data
-    map(1:lcomb, function(z){ # number of combinations nx*nfacet
-      JS(
-        prob = quantile_prob_val,
-        unlist(data_q_wide[z,x]),
-        unlist(data_q_wide[z,y])
-      ) %>% as_tibble()
-    })%>% bind_rows(.id = "combinations")
-  })%>% bind_rows(.id = "data_type1")
-}) %>% bind_rows(.id = "data_type2")
-
-dist_mat <- dist_data %>%
-  group_by(data_type1, data_type2) %>%
-  summarise(dist = sum(value)) %>%
-  pivot_wider(names_from = data_type2,
-              values_from = dist) %>%
-  mutate(data_type1 = as.numeric(data_type1)) %>%
-  left_join(ndata, by = c("data_type1" = "index"))
-
-dist_mat_format <- dist_mat %>%
-  ungroup() %>%
-  select(-data_type1, -unique_data)
-
-rownames(dist_mat_format) <- dist_mat$unique_data
-# write_rds(dist_mat_format, "validation/interact-4D/distmat_40series.rds")
-# write_rds(dist_mat, "validation/interact-4D/distmat_40series_withnames.rds")
+# data_q_wide <- data_q %>%
+#   select(id_facet, id_x, unique_data, sim_data_quantile) %>%
+#   pivot_wider(names_from = unique_data,
+#               values_from = sim_data_quantile) %>%
+#   select(-c(1, 2))
+# 
+# ndata <- data_q %>% distinct(unique_data) %>% mutate(index = row_number())
+# ldata <- nrow(ndata)
+# lcomb <-  nx_val*nfacet_val
+# 
+# dist_data <- map(1:ldata, function(x){ # first data
+#   map(1:ldata, function(y){ # 2nd data
+#     map(1:lcomb, function(z){ # number of combinations nx*nfacet
+#       JS(
+#         prob = quantile_prob_val,
+#         unlist(data_q_wide[z,x]),
+#         unlist(data_q_wide[z,y])
+#       ) %>% as_tibble()
+#     })%>% bind_rows(.id = "combinations")
+#   })%>% bind_rows(.id = "data_type1")
+# }) %>% bind_rows(.id = "data_type2") 
+# 
+# 
+# 
+# dist_mat <- dist_data %>%
+#   group_by(data_type1, data_type2) %>%
+#   summarise(dist = sum(value)) %>%
+#   pivot_wider(names_from = data_type2,
+#               values_from = dist) %>%
+#   mutate(data_type1 = as.numeric(data_type1)) %>%
+#   left_join(ndata, by = c("data_type1" = "index"))
+# 
+# 
+# dist_mat_format <- dist_mat %>%
+#   ungroup() %>%
+#   select(-data_type1, -unique_data)
+# 
+# 
+# rownames(dist_mat_format) <- dist_mat$unique_data
+# write_rds(dist_mat_format, "data/gracsr/validation/interact-4D/distmat_40series.rds")
+# write_rds(dist_mat, "data/gracsr/validation/interact-4D/distmat_40series_withnames.rds")
+# 
+dist_mat_format <- read_rds("data/gracsr/validation/interact-4D/distmat_40series.rds")
+dist_mat <- read_rds("data/gracsr/validation/interact-4D/distmat_40series_withnames.rds")
 ```
@@
 colnames(mds) <- c("Dim.1", "Dim.2")
-rownames(mds) <- dist_mat$unique_data
+#rownames(mds) <- dist_mat$unique_data
 
@@
        color = groups)) +
-  geom_text(label = rownames(mds),check_overlap = TRUE)+
+  geom_text(label = rownames(dist_mat_format),check_overlap = TRUE)+
   geom_point(size = 1) +
@@
 ```{r JS-mydata}
-data_q_wide <- data_q %>%
-  select(id_facet, id_x, unique_data, sim_data_quantile) %>%
-  pivot_wider(names_from = unique_data,
-              values_from = sim_data_quantile) %>%
-  select(-c(1, 2))
-
-ndata <- data_q %>% distinct(unique_data) %>% mutate(index = row_number())
-ldata <- nrow(ndata)
-lcomb <-  nx_val*nfacet_val
-
-dist_data <- map(1:ldata, function(x){ # first data
-  map(1:ldata, function(y){ # 2nd data
-    map(1:lcomb, function(z){ # number of combinations nx*nfacet
-      JS(
-        prob = quantile_prob_val,
-        unlist(data_q_wide[z,x]),
-        unlist(data_q_wide[z,y])
-      ) %>% as_tibble()
-    })%>% bind_rows(.id = "combinations")
-  })%>% bind_rows(.id = "data_type1")
-}) %>% bind_rows(.id = "data_type2")
-
-dist_mat <- dist_data %>%
-  group_by(data_type1, data_type2) %>%
-  summarise(dist = sum(value)) %>%
-  pivot_wider(names_from = data_type2,
-              values_from = dist) %>%
-  mutate(data_type1 = as.numeric(data_type1)) %>%
-  left_join(ndata, by = c("data_type1" = "index"))
-
-dist_mat_format <- dist_mat %>%
-  ungroup() %>%
-  select(-data_type1, -unique_data)
-
-rownames(dist_mat_format) <- dist_mat$unique_data
-# write_rds(dist_mat_format, "validation/interact-4D/distmat_400series.rds")
-# write_rds(dist_mat, "validation/interact-4D/distmat_400series_withnames.rds")
+# data_q_wide <- data_q %>%
+#   select(id_facet, id_x, unique_data, sim_data_quantile) %>%
+#   pivot_wider(names_from = unique_data,
+#               values_from = sim_data_quantile) %>%
+#   select(-c(1, 2))
+# 
+# ndata <- data_q %>% distinct(unique_data) %>% mutate(index = row_number())
+# ldata <- nrow(ndata)
+# lcomb <-  nx_val*nfacet_val
+# 
+# dist_data <- map(1:ldata, function(x){ # first data
+#   map(1:ldata, function(y){ # 2nd data
+#     map(1:lcomb, function(z){ # number of combinations nx*nfacet
+#       JS(
+#         prob = quantile_prob_val,
+#         unlist(data_q_wide[z,x]),
+#         unlist(data_q_wide[z,y])
+#       ) %>% as_tibble()
+#     })%>% bind_rows(.id = "combinations")
+#   })%>% bind_rows(.id = "data_type1")
+# }) %>% bind_rows(.id = "data_type2")
+# 
+# 
+# 
+# dist_mat <- dist_data %>%
+#   group_by(data_type1, data_type2) %>%
+#   summarise(dist = sum(value)) %>%
+#   pivot_wider(names_from = data_type2,
+#               values_from = dist) %>%
+#   mutate(data_type1 = as.numeric(data_type1)) %>%
+#   left_join(ndata, by = c("data_type1" = "index"))
+# 
+# 
+# dist_mat_format <- dist_mat %>%
+#   ungroup() %>%
+#   select(-data_type1, -unique_data)
+# 
+# 
+# rownames(dist_mat_format) <- dist_mat$unique_data
+# write_rds(dist_mat_format, "data/gracsr/validation/interact-4D/distmat_400series.rds")
+# write_rds(dist_mat, "data/gracsr/validation/interact-4D/distmat_400series_withnames.rds")
+dist_mat_format <- read_rds("data/gracsr/validation/interact-4D/distmat_400series.rds")
+dist_mat <- read_rds("data/gracsr/validation/interact-4D/distmat_400series_withnames.rds")
+
 ```
@@
 
-_Clustering all 350 customers_
+## Clustering 353 customers
+
+<!-- All the customers after data filtering and variable selection in the main chapter, before proceeding to instance selection. -->
 
@@
 
-wkndwday <- read_rds(here("data/dist_gran_wkndwday_356cust_nqt.rds")) %>% broom::tidy()
-
-moy <- read_rds(here("data/dist_gran_moy_356cust_nqt.rds")) %>% broom::tidy()
-
-hod <- read_rds(here("data/dist_gran_hod_356cust_nqt.rds")) %>% broom::tidy()
+wkndwday <- read_rds(here("data/gracsr/dist_gran_wkndwday_356cust_nqt.rds")) %>% broom::tidy()
+
+moy <- read_rds(here("data/gracsr/dist_gran_moy_356cust_nqt.rds")) %>% broom::tidy()
+
+hod <- read_rds(here("data/gracsr/dist_gran_hod_356cust_nqt.rds")) %>% broom::tidy()
 
@@
 
-data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>%
+data_pick <- read_rds(here::here("data/gracsr/elec_nogap_2013_clean_356cust.rds")) %>%
   mutate(customer_id = as.character(customer_id)) %>%
@@
 
-```{r combined-groups-js-entire, fig.cap = "The distribution of electricity demand for the clusters across hod, moy and wkndwday for the $17$ groups from $353$ customers.
+```{r combined-groups-js-entire-all, fig.cap = "The distribution of electricity demand for the clusters across hod, moy and wkndwday for the $17$ groups from $353$ customers.
 Wknd-wday variations across groups are not distinguishable, but ideally each group should have an unique combination of hod and moy.", fig.height=6}

### 2021-11-10 scripts/gracsr-supplementary.R

@@ -1,0 +1,255 @@
+# scripts to see how the time series for the four designs look like
+
+##----load-lib
+library(hakear)
+library(tidyr)
+library(distributional)
+library(ggplot2)
+library(patchwork)
+library(purrr)
+library(dplyr)
+library(kableExtra)
+library(tidyverse)
+library(gravitas)
+library(magrittr)
+## ----designs
+
+sim_null_normal <- function(nxj, nfacetj, mean, sd, w1 = 0, w2=0) {
+  rep(distributional::dist_normal(mu = mean, sigma = sd),
+      times = nxj * nfacetj
+  )
+}
+
+sim_varf_normal <- function(nx, nfacet, mean, sd, w1, w2) {
+  rep(dist_normal((mean + seq(0, nfacet - 1, by = 1) * w1),
+                  (sd + seq(0, nfacet - 1, by = 1) * w2)), each = nx)
+}
+
+sim_varx_normal <- function(nx, nfacet, mean, sd, w1, w2) {
+  rep(dist_normal((mean + seq(0, nx - 1, by = 1) * w1),
+                  (sd + seq(0, nx - 1, by = 1) * w2)), nfacet)
+}
+
+sim_varall_normal <- function(nx, nfacet, mean, sd, w1, w2) {
+  dist_normal((mean + seq(0, (nx *nfacet - 1), by = 1) * w1),
+              (sd + seq(0,  (nx *  nfacet - 1), by = 1) * w2))
+}
+
+
+## ----data
+
+
+sim_panel_varall <- sim_panel(
+  nx = nx_val, nfacet =  nfacet_val,
+  ntimes = ntimes_val,
+  # sim_dist = sim_varall_normal(2, 3, 5, 10, 5, -1.5)
+  sim_dist = sim_varall_normal(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
+) %>% unnest(data)
+
+sim_panel_varx <- sim_panel(
+  nx = nx_val, nfacet =  nfacet_val,
+  ntimes = ntimes_val,
+  # sim_dist = sim_varx_normal(2, 3, 5, 10, 5, -1.5)
+  sim_dist = sim_varx_normal(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
+) %>% unnest(data)
+
+sim_panel_varf <- sim_panel(
+  nx = nx_val, nfacet =  nfacet_val,
+  ntimes = ntimes_val,
[…]

### 2021-11-10 Rmd/00-b-abstract.Rmd
minor change in abstract
minor changes in abstract and intro and git update

@@
 The research was motivated by the desire to understand the Australian smart meter data, which was collected half-hourly for two years at the household level.
-With temporal data available at ever finer scales,
-exploration may be required to be carried out across both finer and coarser scales to draw useful inferences about the underlying process.
+With temporal data available at ever finer scales, exploring periodicity can become overwhelming with so many possible temporal deconstructions to explore.
 Analysts are expected to comprehensively explore the many ways to view and consider temporal data.
@@
 "harmony" denotes pairs of granularities that could be effectively analyzed together and reduces the search from all possible options.
-This approach is still overwhelming for human consumption because there would still be a huge number of harmonies.
+This approach is still overwhelming for human consumption due of the vast amount of harmonies remaining.
 The second contribution (Chapter \@ref(ch:hakear)) refines the search for informative granularities by identifying those for which the differences between the displayed distributions are greatest and also rating them in order of importance of capturing maximum variation.

### 2021-11-10 Rmd/01-intro.Rmd
minor changes in abstract and intro

@@
 Chapter \@ref(ch:gravitas) describes classes of time deconstructions using linear and cyclic time granularities, which can be used to create data visualizations to explore periodicities, associations, and anomalies.
-It provides a formal characterization of cyclic granularities which can be circular (such as hour-of-the-day), quasi-circular (such as day-of-the-month), and aperiodic (public holidays).
-It further facilitates manipulation of single- and multiple-order-up time granularities through cyclic calendar algebra, as well as provides a recommendation algorithm to check the feasibility of creating plots for any two cyclic granularities.
+It provides a formal characterization of cyclic granularities and facilitates manipulation of single- and multiple-order-up time granularities through cyclic calendar algebra, as well as provides a recommendation algorithm to check the feasibility of creating plots for any two cyclic granularities.
 Our proposed method is also applicable to non-temporal hierarchical granularities with an underlying ordered index.

### 2021-11-10 Rmd/00-c-declaration.Rmd
Rob's corrections

@@
 
-Three publications are included in this thesis, one of which has been accepted for publication in the Journal of Computational and Graphical Statistics (@Gupta2021-gravitas).
+Three publications are included in this thesis, one of which has been accepted for publication in the _Journal of Computational and Graphical Statistics_ (@Gupta2021-gravitas).
 Two others have not yet been submitted.
@@
 
-I have not renumbered sections of submitted or published papers in order to generate a consistent presentation within the thesis.
+I have renumbered sections of submitted or published papers in order to generate a consistent presentation within the thesis.
 

### 2021-11-10 Rmd/00-b-abstract.Rmd
rob's correction - abstract

@@
 
-The research was motivated by the desire to understand the Australian smart meter data, which was collected half-hourly for two years at the household level.
+The research was motivated by the desire to understand some Australian smart meter data, which was collected half-hourly for two years at the household level.
 With temporal data available at ever finer scales, exploring periodicity can become overwhelming with so many possible temporal deconstructions to explore.
@@
 It provides tools to compute possible cyclic granularities from an ordered (usually temporal) index and also a framework to systematically explore the distribution of a univariate variable conditional on two cyclic time granularities by defining "harmony".
-"harmony" denotes pairs of granularities that could be effectively analyzed together and reduces the search from all possible options.
-This approach is still overwhelming for human consumption due of the vast amount of harmonies remaining.
+A "harmony" denotes pairs of granularities that could be effectively analyzed together and reduces the search from all possible options.
+This approach is still overwhelming for human consumption due of the vast number of harmonies remaining.
 The second contribution (Chapter \@ref(ch:hakear)) refines the search for informative granularities by identifying those for which the differences between the displayed distributions are greatest and also rating them in order of importance of capturing maximum variation.

### 2021-11-10 Rmd/00-d-acknowlegements.Rmd
Rob's correction - acknowledgement

@@
 A few highlights are UseR!
-2018 (Brisbane), Young Statistician's Conference 2019 (Canberra), WOMBATS 2019 (Melbourne), and RopenSci 2019 (Sydney).
+2018 (Brisbane), Young Statistician's Conference 2019 (Canberra), WOMBAT 2019 (Melbourne), and ROpenSci 2019 (Sydney).
 Thank you to Gael Martin, Farshid Wahid, Klaus Ackermann, and Didier Nibbering from my PhD committee for providing such a supportive environment.
@@
 
-PhD in Covid has been harder, and I would not have been able to make it through this if not for the emotional support of my friends and family.
+PhD in Covid has been hard, and I would not have been able to make it through this if not for the emotional support of my friends and family.
 Thank you Puwasala, Sium, and Ian, for being so kind and acting as a pillar of emotional support, especially during the last leg of my journey.
@@
 Thank you to each one of you for being available, listening to me, and putting things in perspective when I lost sight.
-Thanks to my housemates Anjali, Surbhi, and Dulaji for the fun company and conversations that kept me sane in the covid lockdowns.
+Thanks to my housemates Anjali, Surbhi, and Dulaji for the fun company and conversations that kept me sane in the COVID lockdowns.
 

### 2021-11-10 Rmd/01-intro.Rmd
Rob's correction intro done

@@
 The Smart Grid, Smart City (SGSC) project (2010–2014) available through the [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) provides half-hourly data of over 13,000 Australian electricity smart meter customers distributed unevenly from October 2011 to March 2014.
-Large data sets like this include significant uncertainty regarding customer behavior due to the rising variety of customers.
+Large data sets like this include significant uncertainty regarding customer behavior due to a wide variety of customers.
 Behavioral patterns vary significantly due to differences in size, location, and amenities such as solar panels, central heating, and air conditioning.
@@
 <!-- introducing the overall field-->
-With the availability of data at more and more finer time scales, exploration of time series data may be required to be carried out across both finer and coarser scales to draw useful inferences about the underlying process.
+With the availability of data at finer and finer time scales, exploration of time series data may be required to be carried out across both finer and coarser scales to draw useful inferences about the underlying process.
 To reduce the complexity of time, it is typical to divide it into years, months, weeks, days, and so on in a hierarchical manner [@aigner2011visualization].
@@
 To acquire a comprehensive view of the repeated patterns, it is necessary to navigate through all of the conceivable cyclic granularities.
-This approach is consistent with the concept of EDA [@Tukey1977-jx], which stresses the utilization of multiple perspectives on data to assist formulate hypotheses before proceeding to hypothesis testing.
+This approach is consistent with the concept of EDA [@Tukey1977-jx], which stresses the utilization of multiple perspectives on data to assist with formulating hypotheses before proceeding to hypothesis testing.
 This, however, is a challenging process since it throws up a myriad of alternatives.
@@
 Chapter \@ref(ch:gravitas) describes classes of time deconstructions using linear and cyclic time granularities, which can be used to create data visualizations to explore periodicities, associations, and anomalies.
-It provides a formal characterization of cyclic granularities and facilitates manipulation of single- and multiple-order-up time granularities through cyclic calendar algebra, as well as provides a recommendation algorithm to check the feasibility of creating plots for any two cyclic granularities.
+It provides a formal characterization of cyclic granularities and facilitates manipulation of single- and multiple-order-up time granularities through cyclic calendar algebra, as well as providing a recommendation algorithm to check the feasibility of creating plots for any two cyclic granularities.
 Our proposed method is also applicable to non-temporal hierarchical granularities with an underlying ordered index.
@@
 
-This work is a natural extension of Chapter \@ref(ch:gravitas).
+Chapter \@ref(ch:hakear) is a natural extension of Chapter \@ref(ch:gravitas).
 Many, many displays might be built using cyclic granularities.
@@
 This chapter also includes a case study using residential smart meter data from Melbourne to demonstrate how the suggested methodology may be utilized to automatically find temporal granularities with significant distributional differences.
-The methods are implemented in the open-source R package 'hakear.'
+The methods are implemented in the open-source R package `hakear`.
 
@@
 Chapter \@ref(ch:gravitas) provides details of the cyclic granularities, different classes, and computation, and also its usage in exploratory time series analysis through applications.
-This is implemented in the R package **gravitas**.
+This is implemented in the R package `gravitas`.
 Chapter \@ref(ch:hakear) provides guidance on how to choose significant cyclic granularities, which are likely to have interesting patterns across its categories.
-This is available as the R package **hakear**.
-The chapter \@ref(ch:gracsr) provides methods to explore heterogeneity in repetitive behavior for multiple time series over multiple cyclic granularities.
-This is in the developing R package **gracsr**.
+This is available as the R package `hakear`.
+Chapter \@ref(ch:gracsr) provides methods to explore heterogeneity in repetitive behavior for multiple time series over multiple cyclic granularities.
+This is in the developing R package `gracsr`.
 Chapter \@ref(ch:conclusion) summarizes the software tools developed for the work, and discusses some future plans.

### 2021-11-10 Rmd/03-hakear.Rmd
updated code of paper-hakear

--- 2021-11-05 Rmd/03-hakear.Rmd
@@
 
-Cyclic temporal granularities [@Bettini1998-ed], [@Gupta2021-hd] are <!--theoretical--> temporal deconstructions that define cyclic repetitions in time, e.g.
+Cyclic temporal granularities [@Bettini1998-ed;@Gupta2021-hd] are <!--theoretical--> temporal deconstructions that define cyclic repetitions in time, e.g.
 hour-of-day, day-of-month, or regularly scheduled public holidays.
@@
 An example of an ordered granularity is day-of-week, where Tuesday is always followed by Wednesday, and so on.
-An unordered granularity example is different week types in a semester, orientation, break, exam or regular classes in an academic calendar.
-Using granularities to  explore patterns in univariate time series can be considered to be examining the distribution of the measured variable across different categories of the cyclic granularities.
-
-Figure \ref{fig:onegran-new} electricity smart meter data plotted against two granularities (hour-of-day, month-of-year).
-The data was collected on a single household in Melbourne, Australia, over a six month period, as was used in @wang2020calendar.
-The categorical variable (granularity) is mapped to the x-axis, and the distribution of response variable is displayed using both side-by-side jittered dotplots and boxplots.
-From plot (a) it can be seen that energy consumption is higher during the morning hours (5-8), when members in the household wake up, and again in the evening hours (17-20) possibly when members get back from work.
-In addition, the largest variation in energy use is in the afternoon hours (12-16), as perceived from sizes of the boxes.
-From plot (b) the variability in energy usage is higher in Jan and Feb, possibly due to the usage of air conditioners on some days.
-The median usage is highest in January, dips in February-April and rises again in May-June, although not to the height of January usage.
-This might imply that this household does not use as much energy for heating as it does for air conditioning.
-A lot of households in Victoria use gas heating and hence the heater use might not be reflected in the electricity data.
-Many, many different displays could be constructed using different granularities,  day-of-week, day-of-month, weekday/weekend, etc.
-However, only a few might be interesting, that is, reveal important patterns in energy usage.
-Determining which displays which have "significant" distributional differences between categories of the cyclic granularity, and plotting only these, would make for efficient exploration.
+An unordered granularity example is \DC{week type in an academic semester: orientation, break, exam or regular classes}.
+We can use granularities to explore patterns in univariate time series by examining the distribution of the measured variable across different categories of the cyclic granularities.
 
@@
 Daily peaks occur in morning and evening hours, indicating a working household, where members leave for and return from work.
-More volatility of usage in summer months (Jan, Feb) is probably due air conditioner use on just some days ."}
+More volatility of usage in summer months (Jan, Feb) is probably due air conditioner use on just some days.", fig.pos="!b"}
+```
+
+As a motivating example, consider Figure \ref{fig:onegran-new} which shows electricity smart meter data plotted against two granularities (hour-of-day, month-of-year).
+The data was collected on a single household in Melbourne, Australia, over a six month period, and was previously used in @wang2020calendar.
+The categorical variable (granularity) is mapped to the x-axis, and the distribution of the response variable is displayed using both side-by-side jittered dotplots and boxplots.
+From panel (a) it can be seen that energy consumption is higher during the morning hours (5--8), when members in the household wake up, and again in the evening hours (17--20) possibly when members get back from work.
+In addition, the largest variation in energy use is in the afternoon hours (12--16), as seen in the length of the boxes.
+From panel (b), it is seen that the variability in energy usage is higher in Jan and Feb, probably due to the usage of air conditioners on some days.
+The median usage is highest in January, dips in February--April and rises again in May--June, although not to the height of January usage.
+This suggests that the household does not use as much electricity for heating as it does for air conditioning.
+A lot of households in Victoria use gas heating and hence the heater use might not be reflected in the electricity data.
+
+Many different displays could be constructed using different granularities including day-of-week, day-of-month, weekday/weekend, etc.
+However, only a few might be interesting and reveal important patterns in energy usage.
+Determining which displays have "significant" distributional differences between categories of the cyclic granularity, and plotting only these, would make for efficient exploration.
+
+```{r intro_all}
+
+```
+
+```{r id2-new2, fig.cap = "Distribution of energy consumption displayed through area quantile plots across two cyclic granularities month-of-year and hour-of-day and two households.
+The black line is the median, whereas the orange band covers the 25th to 75th percentile and the green band covers the 10th to 90th percentile.
+Difference between the 90th and 75th quantiles is less for (Jan, Feb) for the first household (a), suggesting that it is a more frequent user of air conditioner than the second household (b).
+Energy consumption for in (a) changes across both granularities, whereas for (b) daily pattern stays same irrespective of the months."}
 ```
@@
 <!-- comparing one and two grans for one households -->
-\noindent Exploring the distribution of the measured variable across two cyclic granularities tends to provide more detailed information on its structure.
-For example, Figure \ref{fig:id2-new2}(a) slice down further by showing the usage distribution across hour-of-day conditional on month-of-year across two households (id 2 and 4).
+Exploring the distribution of the measured variable across two cyclic granularities provides more detailed information on its structure.
+For example, Figure \ref{fig:id2-new2}(a) shows the usage distribution across hour-of-day conditional on month-of-year across two households.
 It shows the hourly usage over a day does not remain the same across months.
-Unlike other months, the 75th and 90th percentile for all hours of the day in January are    high, pretty close, and are not characterized by a morning and evening peak.
+Unlike other months, the 75th and 90th percentile for all hours of the day in January are high, similar, and are not characterized by a morning and evening peak.
 The household in Figure \ref{fig:id2-new2}(b) has 90th percentile consumption higher in summer months relative to autumn or winter, but the 75th and 90th percentile are far apart in all months, implying that the second household resorts to air conditioning much less regularly than the first one.
@@
 <!-- The problem and its dimension -->
-\noindent Are all of these four displays in Figures \ref{fig:onegran-new} and \ref{fig:id2-new2} useful in understanding the distributional difference in energy usage?
+Are all four displays in Figures \ref{fig:onegran-new} and \ref{fig:id2-new2} useful in understanding the distributional difference in energy usage?
 Which ones are more useful than others?
@@
 When considering displays of the form in Figure \ref{fig:id2-new2},
-there are $^{N_C}P_2$ possible pairwise plots exhaustively, with one of the two cyclic granularities acting as the conditioning variable. 
-This is huge and overwhelming for human consumption even for moderately large $N_C$.
-It could be immensely useful to make the transition from all potential displays to the ones that are informative across atleast one cyclic granularity.
+there are $N_C(N_C-1)$ possible pairwise plots exhaustively, with one of the two cyclic granularities acting as the conditioning variable.
+This can be overwhelming for human consumption even for moderately large $N_C$. It is therefore  useful to identify those displays that are informative across at least one cyclic granularity.
 
@@
 
-\noindent This problem is similar to Scagnostics (Scatterplot Diagnostics) by @tukey1988computer, which is used to identify meaningful patterns in large collections of scatterplots.
+This problem is similar to Scagnostics (Scatterplot Diagnostics) by @tukey1988computer, which are used to identify meaningful patterns in large collections of scatterplots.
 Given a set of $v$ variables, there are $v(v-1)/2$ pairs of variables, and thus the same number of possible pairwise scatterplots.
-Therefore, even for small $v$, the number of scatterplots can be large, and scatterplot matrices (SPLOMs) could easily run out of pixels when presenting high-dimensional data.
-@Dang2014-tw and @wilkinson2005graph provide potential solutions to this, where few characterizations can be used to locate anomalies in density, shape, trend, and other features in the 2D point scatters.
-In this paper, we provide a solution to narrowing down the search from $^{N_C}P_2$ plots by introducing a new distance measure that can be used to detect significant distributional differences across cyclic granularities.<!-- harmonies and why it is not enough --> This work is a natural extension of our previous work [@Gupta2021-hd] that narrows down the search from $^{N_C}P_2$ plots by identifying pairs of granularities that can be meaningfully examined together (a "harmony"), or when they cannot (a "clash").
-However, even after excluding clashes, the list of harmonies left could be enormous for exhaustive exploration.
-Hence, there is a need to reduce the search even further by including only those harmonies which are informative enough.
+Therefore, even for small $v$, the number of scatterplots can be large, and scatterplot matrices (SPLOMs) can easily run out of pixels when presenting high-dimensional data.
+@Dang2014-tw and @wilkinson2005graph provide potential solutions to this, where a few characterizations can be used to locate anomalies in density, shape, trend, and other features in the 2D point scatters.
+
+In this paper, we provide a solution to narrowing down the search from $N_C(N_C-1)$ conditional distribution plots by introducing a new distance measure that can be used to detect significant distributional differences across cyclic granularities.<!-- harmonies and why it is not enough --> This work is a natural extension of our previous work [@Gupta2021-hd] which narrows down the search from $N_C(N_C-1)$ plots by identifying pairs of granularities that can be meaningfully examined together (a "harmony"), or when they cannot (a "clash").
+However, even after excluding clashes, the list of harmonies left may be too large for exhaustive exploration.
+Hence, there is a need to reduce the search even further by including only those harmonies that contain useful information.
+
 @inference and @Majumder2013-hb present methods for statistical significance testing of visual findings using human cognition as the statistical tests.
@@
 
-```{r intro_all}
-
-```
-
-```{r id2-new2, fig.cap = "Distribution of energy consumption displayed through area quantile plots across two cyclic granularities month-of-year and hour-of-day and two households.
-The black line is the median, whereas the orange band covers the 25th to 75th percentile and the green band covers the 10th to 90th percentile.
-Difference between the 90th and 75th quantiles is less for (Jan, Feb) for the first household (a), suggesting that it is a more frequent user of air conditioner than the second household (b).
-Energy consumption for in (a) changes across both granularities, whereas for (b) daily pattern stays same irrespective of the months."}
-```
-
 ```{r id4, fig.cap = "something2", eval = FALSE}
@@
 
-\noindent The article is organized as follows.
+The article is organized as follows.
 Section \ref{sec:computation-wpd} introduces a distance measure for detecting distributional difference in temporal granularities, which enables identification of patterns in the time series data;
 Section \ref{sec:rank-wpd} devises a selection criterion by choosing a threshold, which results in detection of only significantly interesting patterns.
-Section \ref{sec:simulations} provides some simulation study on the proposed methodology.
+Section \ref{sec:simulations} provides a simulation study on the proposed methodology.
 Section \ref{sec:application-wpd} presents an application to residential smart meter data in Melbourne to show how the proposed methodology can be used to automatically detect temporal granularities along which distributional differences are significant.
@@
 
-We propose a measure called Weighted Pairwise Distances ($wpd$) to detect distributional differences in the measured variable across cyclic granularities.
+We propose a measure called Weighted Pairwise Distances ($\wpd$) to detect distributional differences in the measured variable across cyclic granularities.
 
@@
 
-The principle behind the construction of $wpd$ is explained through a simple example explained in Figure \ref{fig:null4by2}.
-Each of these figures describes a panel with $2$ x-axis categories and $3$ facet levels, but with different designs.
-Figure \ref{fig:null4by2}a has all categories drawn from N(0, 1) distribution for each facet.
-It is not an interesting display particularly, as distributions do not vary across x-axis or facet categories.
-Figure \ref{fig:null4by2}b has x categories drawn from the same distribution, but across facets the distributions are $3$ standard deviations apart.
-Figure \ref{fig:null4by2}c exhibits an exact opposite situation where distribution between the x-axis categories
-are $3$ standard deviations apart, but they do not change across facets.
-Figure \ref{fig:null4by2}d takes a step further by varying the distribution across both facet and x-axis categories by  $3$ standard deviations.
-If the panels are to be ranked in order of capturing maximum variation in the measured variable from minimum to maximum, then an obvious choice would be placing (a) followed by (b), (c) and then (d).
+(ref:null4by2) An example illustrating the principle of the proposed distance measure, displaying the distribution of a normally distributed variable in four panels each with two x-axis categories and three facet levels, but with different designs.
+Panel (a) is not interesting as the distribution of the variable does not depend on x or facet categories.
+Panels (b) and (c) are more interesting than (a) since there is a change in distribution either across facets (b) or x-axis (c).
+Panel (d) is most interesting in terms of capturing structure in the variable as the distribution of the variable changes across both facet and x-axis variable.
+The value of our proposed distance measure is presented for each panel, the relative differences between which will be explained later in Section \@ref(sec:ranking).
+
+```{r null4by2, fig.cap = "(ref:null4by2)", fig.pos="!b"}
+```
+
+The principle behind the construction of $\wpd$ is explained through a simple example in Figure&nbsp;\ref{fig:null4by2}.
+Each of these figures describes a panel with two x-axis categories and three facet levels, but with different designs.
+Figure \ref{fig:null4by2}a has all categories drawn from a standard normal distribution for each facet.
+It is not a particularly interesting display, as the distributions do not vary across x-axis or facet categories.
+Figure \ref{fig:null4by2}b has x categories drawn from the same distribution, but across facets the distributions are three standard deviations apart.
+Figure \ref{fig:null4by2}c exhibits the opposite situation where distribution between the x-axis categories are three standard deviations apart, but they do not change across facets.
+In Figure \ref{fig:null4by2}d, the distribution varies across both facet and x-axis categories by three standard deviations.
+
+If the panels are to be ranked in order of capturing maximum variation in the measured variable from minimum to maximum, then an obvious choice would be (a) followed by (b), (c) and then (d).
 It might be argued that it is not clear if (b) should precede or succeed (c) in the ranking.
 Gestalt theory suggests items placed at close proximity can be compared more easily, because people assume that they are in the same group and apart from other groups.
-With this principle in mind, display (b) is considered less informative as compared to display (c) in emphasizing the distributional differences.
-Considering one cyclic granularity, we would have only two design choices similar to (a) and (c), corresponding to no difference and significant differences between categories of that cyclic granularity only.
-The proposed measure $wpd$ is constructed in a way so that it could be used to rank panels of different designs as well as test if a design is interesting.
+With this principle in mind, Panel (b) is considered less informative compared to Panel (c) in emphasizing the distributional differences.
+
+For displays showing a single cyclic granularity rather than pairs of granularities, we have only two design choices corresponding to no difference and significant differences between categories of that cyclic granularity.
+
+The proposed measure $\wpd$ is constructed in such a way that it can be used to rank panels of different designs as well as test if a design is interesting.
 This measure is aimed to be an estimate of the maximum variation in the measured variable explained by the panel.
-A higher value of  $wpd$ would indicate that the panel is interesting to look at, whereas a lower value would indicate otherwise.
+A higher value of $\wpd$ would indicate that the panel is interesting to look at, whereas a lower value would indicate otherwise.
 
@@
 Moreover, a tuning parameter specifying the weightage given to within-facet or between-facet categories can help to choose between designs like \ref{fig:null4by2}b and c.
-The maximum of all these weighted differences serves as an estimate of the maximum variation in the panel and is taken as the value of $wpd$.
+The maximum of all these weighted differences serves as an estimate of the maximum variation in the panel and is taken as the value of $\wpd$.
 -->
@@
 One of the potential ways to do this is to measure the distances between distributions of the continuous random variable measured within and between groups, weigh them basis if they are within or between groups and then take the maximum of those distances as an estimate of the maximum variation in the structure.
-This section starts with possible ways of characterizing distributions and computing distances between them and then describe in details how the measure $wpd$ is defined.
+This section starts with possible ways of characterizing distributions and computing distances between them and then describe in details how the measure $\wpd$ is defined.
 -->
@@
 For example, if there are 4 facet levels and 2 x-axis levels, data is generated in the way as described in Table \ref{}.
-Figure \ref{}b exhibits an exact opposite situation where the x-axis within facets are different but not across facets.
+Figure \ref{}b exhibits an opposite situation where the x-axis within facets are different but not across facets.
 -->
 <!-- Figure \ref{}d takes it further by varying the distribution across both facet and x-axis categories. -->
-
-```{r null4by2,fig.cap = "An example illustrating the principle of the proposed distance measure, displaying the distribution of a normally distributed variable in four panels each with $2$ x-axis categories and $3$ facet levels, but with different designs.
-Display (a) is not interesting as the distribution of the variable does not depend on x or facet categories.
-Display (b) and (c) are more interesting than (a) since there is a change in distribution either across facets (b) or x-axis (c).
-Display (d) is most interesting in terms of capturing structure in the variable as the distribution of the variable changes across both facet and x-axis variable.
-The value of our proposed distance measure is presented for each panel, the relative differences between which will be explained later in Section 3.2."}
-```
 
@@
 
-### Notations
-
-\noindent Let the number of cyclic granularities considered in the display be $m$.
+### Notation
+
+Let the number of cyclic granularities considered in the display be $m$.
 The notations and methodology are described in detail for $m=2$.
 But it can be easily extended to $m>2$.
-Consider two cyclic granularities $A$ and $B$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$ with $A$ placed across x-axis and $B$ across facets.
+Consider two cyclic granularities $A$ and $B$, such that $A = \{a_j: j = 1, 2, \dots, \nx\}$ and $B = \{b_k: k = 1, 2, \dots, \nf\}$ with $A$ placed across the x-axis and $B$ across facets.
 Let $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ be a continuous variable observed across $T$ time points.
-This  data structure with $J$ x-axis levels and $K$ facet levels is referred to as a $(J, K)$ panel.
-For example, a $(2, 3)$ panel will have cyclic granularities with 2 x-axis levels and 3 facet levels.
-Let the four elementary designs as described in Figure \ref{fig:null4by2} be $D_{null}$ (referred to as "null distribution") where there is no difference in distribution of $v$ for $A$ or $B$, $D_{var_f}$ denotes the set of designs where there is difference in distribution of $v$ for $B$ and not for $A$.
-Similarly, $D_{var_x}$ denotes the set of designs where difference is observed only across A.
-Finally, $D_{var_{all}}$ denotes those designs for which difference is observed across both $A$ and $B$.
-$m = 1$ is a special case of $m = 2$ with $J = 1$.
+This data structure with $\nx$ x-axis levels and $\nf$ facet levels is referred to as a $(\nx, \nf)$ panel.
+For example, a $(2, 3)$ panel will have cyclic granularities with two x-axis levels and three facet levels.
+Let the four elementary designs as described in Figure \ref{fig:null4by2} be $\Dnull$ (referred to as "null distribution") where there is no difference in distribution of $v$ for $A$ or $B$, $\Df$ denotes the set of designs where there is difference in distribution of $v$ for $B$ and not for $A$.
+Similarly, $\Dx$ denotes the set of designs where difference is observed only across A.
+Finally, $\Dfx$ denotes those designs for which difference is observed across both $A$ and $B$.
+We can consider a single granularity ($m = 1$) as a special case of two granularities with $\nf = 1$.
 
@@
 
-The computation of the distance measure $wpd$ for a panel involves characterizing distributions, computing distances between distributions, choosing a tuning parameter to specify the weightage for different group of distances and summarizing those weighted distances appropriately to estimate maximum variation.
-Furthermore, the data needs to be appropriately transformed to ensure that the value of $wpd$ emphasizes detection of distributional differences across categories and not across different data generating processes.
+The computation of the distance measure $\wpd$ for a panel involves characterizing distributions, computing distances between distributions, choosing a tuning parameter to specify the weight for different groups of distances and summarizing those weighted distances appropriately to estimate maximum variation.
+Furthermore, the data needs to be appropriately transformed to ensure that the value of $\wpd$ emphasizes detection of distributional differences across categories and not across different data generating processes.
 
@@
 
-<!-- is aimed to capture structure and patterns by estimating the maximum variation of the measured variable within a panel.  $wpd$ -->
+<!-- is aimed to capture structure and patterns by estimating the maximum variation of the measured variable within a panel. $\wpd$ -->
 
@@
 Moreover, a tuning parameter specifying the weightage given to within-facet or between-facet categories can help to choose between designs like \ref{fig:null4by2}b and c.
-The maximum of all these weighted differences serves as an estimate of the maximum variation in the panel and is taken as the value of $wpd$.
--->
-
-#### Data transformation
-
-The intended aim of $wpd$ is to capture differences in categories irrespective of the distribution from which the data is generated.
-Hence, as a pre-processing step, the raw data is normal-quantile transformed (NQT) (@Krzysztofowicz1997-bv), so that the quantiles of the transformed data follows a standard normal distribution.
-This sort of transformation is common in the fields of geo-statistics to make most asymmetrical distributed real world measured variables more treatable and normal-like (@Bogner2012-az).The empirical NQT involves the following steps:
- 1. The sample of measured variable $v$ is sorted from the smallest to the largest observation $v_{(1)},\dots, v_{(i)}, .., v_{(n)}$.  
- 2. The cumulative probabilities $p_{(1)},\dots, p_{(i)}, .., p_{(n)}$ are estimated using a plotting position like $i/(n + 1)$ such that $p_{(i)} = P(v \leq v_{(i)})$.  
- 3.
-Each observation $v_{(i)}$ of $v$ is transformed into observation $v^*(i) = Q^{-1}(p(i))$ of the standard normal variate $v^*$, with $Q$ denoting the standard normal distribution and $Q^{-1}$ its inverse.
-
-#### Characterising distributions
+The maximum of all these weighted differences serves as an estimate of the maximum variation in the panel and is taken as the value of $\wpd$.
+-->
+
+#### Data transformation {-}
+
+The intended aim of $\wpd$ is to capture differences in categories irrespective of the distribution from which the data is generated.
+Hence, as a pre-processing step, the raw data is normal-quantile transformed (NQT) [@Krzysztofowicz1997-bv], so that the transformed data follows a standard normal distribution.
+The empirical NQT involves the following steps:
+
+ 1. The observations of measured variable $v$ are sorted from the smallest to the largest observation $v_{(1)},\dots, v_{(n)}$.
+ 2. The cumulative probabilities $p_{(1)},\dots, p_{(n)}$ are estimated using $p_{(i)} = i/(n + 1)$  [@Hyndman1996-ty] so that $p_{(i)} = \text{Pr}(v \leq v_{(i)})$.
+ 3. Each observation $v_{(i)}$ of $v$ is transformed into $v^*(i) = \Phi^{-1}(p(i))$, with $\Phi$ denoting the standard normal distribution function.
+
+#### Characterizing distributions {-}
 
@@
 The number of observations might vary widely across subsets due to the structure of the calendar, missing observations or uneven locations of events in the time domain.
-in this paper, quantiles of $v_{jk}$'s are chosen as a way to characterize distributions for the category $(a_j, b_k)$, $\forall j\in \{1, 2, \dots, J\}, k\in \{1, 2, \dots, K\}\}$.
-The quantile of a distribution with probability $p$ is defined as $Q(p)=F^{-1}(p) = inf\{x: F(x) >p\}$, $0<p< 1$ where $F(x)$ is the distribution function.
-There are two broad approaches to quantile estimation, viz, parametric and non-parametric.
-Sample quantiles (@Hyndman1996-ty) are used for estimating population quantiles in a non-parametric setup, which is desirable because of less rigid assumptions made about the nature of the underlying distribution of the data.
-The default quantile chosen in this paper is percentiles computed for $p = {0.01, 0.02, \dots, 0.99}$, where for example, the $99^{th}$ percentile would be the value corresponding to $p=0.99$ and hence 99% of the observations would lie below that.
+In this paper, quantiles of $\{v_{jk}\}$ are chosen as a way to characterize distributions for the category $(a_j, b_k)$, $\forall j\in \{1, 2, \dots, \nx\}, k\in \{1, 2, \dots, \nf\}$.
+We use percentiles with $p = {0.01, 0.02, \dots, 0.99}$ to reduce the computational burden in summarizing distributions.
 
@@
 
-<!-- Each $v_{jk}$'s $\forall j\in \{1, 2, \dots, J\}, k\in \{1, 2, \dots, K\}\}$ are assumed to be drawn from a continuous probability distribution and have certain characteristics.
+<!-- Each $v_{jk}$'s $\forall j\in \{1, 2, \dots, \nx\}, k\in \{1, 2, \dots, \nf\}\}$ are assumed to be drawn from a continuous probability distribution and have certain characteristics.
 Shape, central tendency, and variability are the common characteristics used to describe a distribution.
@@
 
-#### Distance between distributions
+#### Distance between distributions {-}
 
@@
 The continuous version of f -divergence is given by -->
-<!-- $$D_f(P||Q) := \int q(x)f(\frac{p(x)}{q(x)})$$, where -->
+<!-- $$\Df(P||Q) := \int q(x)f(\frac{p(x)}{q(x)})$$, where -->
 <!-- $f : [0,\infty) \rightarrow R \cup \{\infty\}$ is a continuous convex function, and $f(1) = 0$. -->
 
-One of the most common ways to measure divergence between distributions is the Kullback-Leibler (KL) divergence [@Kullback1951-jy].
+A common way to measure divergence between distributions is the Kullback-Leibler (KL) divergence [@Kullback1951-jy].
 The KL divergence denoted by $D(q_1||q_2)$ is a non-symmetric measure of the difference between two probability distributions $q_1$ and $q_2$ and is interpreted as the amount of information lost when $q_2$ is used to approximate $q_1$.
-The KL divergence, however, is not symmetric and hence can not be considered as a "distance" measure.
+The KL divergence is not symmetric and hence can not be considered as a "distance" measure.
 The Jensen-Shannon divergence [@Menendez1997-in] based on the Kullback-Leibler divergence is symmetric and has a finite value.
-Hence, in this paper, the pairwise distances between the distributions of the measured variable are obtained through the square root of the Jensen-Shannon divergence, called Jensen-Shannon distance (JSD) and is defined by,
-
-$$JSD(q_1||q_2) = \frac{1}{2}D(q_1||M) + \frac{1}{2}D(q_2||M)$$
-where $M = \frac{q_1+q_2}{2}$ and 
-$D(q_1||q_2) := \int^\infty_{-\infty} q_1(x)f(\frac{q_1(x)}{q_2(x)})$ is the KL divergence between distributions $q_1$ and $q_2$.
+Hence, in this paper, the pairwise distances between the distributions of the measured variable are obtained through the square root of the Jensen-Shannon divergence, called Jensen-Shannon distance (JSD), and defined by,
+$$
+  JSD(q_1||q_2) = \frac{1}{2}D(q_1||M) + \frac{1}{2}D(q_2||M),
+$$
+where $M = \frac{q_1+q_2}{2}$ and $D(q_1||q_2) := \int^\infty_{-\infty} q_1(x)f(\frac{q_1(x)}{q_2(x)})$ is the KL divergence between distributions $q_1$ and $q_2$.
 Other common measures of distance between distributions are Hellinger distance, total variation distance and Fisher information metric.
 
-<!-- Furthermore, these distances are distributed as chi-squared with $m$ degrees of freedom (@Menendez1997-in), if the continuous distribution is being discretized with $m$ discrete values.
+<!-- Furthermore, these distances are distributed as chi-squared with $m$ degrees of freedom [@Menendez1997-in], if the continuous distribution is being discretized with $m$ discrete values.
 Taking sample percentiles to approximate the integral would mean taking $m = 99$.
@@
 
-#### Within-facet and between-facet distances
-
-\noindent Pairwise distances could be within-facets or between-facets for $m \geq 2$.
+#### Within-facet and between-facet distances {-}
+
+```{r distance-explain, fig.cap = "Within and between-facet distances shown for two cyclic granularities A and B, where A is mapped to x-axis and B is mapped to facets.
+The dotted lines represent the distances between different categories.
+Panels 1) and 2) show the between-facet distances.
+Panels 3) and 4) are used to illustrate within-facet distances when categories are un-ordered or ordered respectively.
+When categories are ordered, distances should only be considered for consecutive x-axis categories.
+Between-facet distances are distances between different facet levels for the same x-axis category, for example, distances between {($a_1$,$b_1$) and ($a_1$, $b_2$)} or {($a_1$,$b_1$) and ($a_1$, $b_3$)}.", fig.pos="!b"}
+
+```
+
+Pairwise distances could be within-facets or between-facets for $m = 2$.
 Figure \ref{fig:distance-explain} illustrates how they are defined.
@@
 Pairwise distances are between-facets when they are considered between pairs of the form $(a_{j}b_{k}, a_{j}b_{k'})$.
-Number of between-facet distances would be $^{K}C_2*J$ and number of within-facet distances are $K*(J-1)$ (ordered) and $^{J}C_2*K$ (un-ordered).
-
-#### Tuning parameter
-
-A tuning parameter specifying the weightage given to the within-facet or between-facet categories can help to balance weightage between designs like \ref{fig:null4by2}(b) and (c).
-The tuning parameters should be chosen such that $\sum_{i=1}^m \lambda_i = 1$.
-When $m=2$, following the principle of Gestalt theory, $\lambda = \frac{2}{3} = 0.67$ is chosen to put a relative weightage of $2:1$ for within-facet and between-facet distances.
-No human experiment is conducted to justify this ratio, however, typically a tuning parameter $\lambda>0.5$ would tend to upweigh the within-facet distances and that with $<0.5$ would upweigh the between-facet distances (refer to the Supplementary paper for more details).
-For $m = 1$, there are no conditioning variables or groups, and hence $\lambda = 1$.
-
-#### Raw distance measure
-
-The raw distance measure, denoted by $wpd_{raw}$, is computed after combining all the weighted distance measures appropriately.
+There are a total of ${\nf \choose 2}\nx$ between-facet distances, and ${\nx \choose 2}\nf$ within-facet distances if they are ordered and $\nf(\nx-1)$ within-facet distances if they are unordered.
+
+#### Tuning parameter {-}
+
+For displays with $m>1$ granularities, we can use a tuning parameter to specify the relative weight given to each granularity.
+In general, the tuning parameters should be chosen such that $\sum_{i=1}^m \lambda_i = 1$.
+
+Following the general principles of Gestalt theory, we wish to weight more heavily granularities that are plotted closer together.
+For $m=2$ we choose  $\lambda_x = \frac{2}{3}$ for the granularity on the x-axis and $\lambda_f = \frac13$ for the granularity mapped to facets, giving a relative weight of $2:1$ for within-facet to between-facet distances.
+No human experiment has been conducted to justify this ratio.
+Specifying $\lambda_x>0.5$ will weight within-facet distances more heavily, while $\lambda_x <0.5$ would weight the between-facet distances more heavily.
+(See the supplements for more details.).
+
+#### Raw distance measure {-}
+
+The raw distance measure, denoted by $\wpdsub{raw}$, is computed after combining all the weighted distance measures appropriately.
 First, NQT is performed on the measured variable $v_t$ to obtain $v^*_t$ (_data transformation_).
 Then, for a fixed harmony pair $(A, B)$, percentiles of $v^*_{jk}$ are computed and stored in $q_{jk}$ (_distribution characterization_).
-This is repeated for all pairs of categories of the form $(a_{j} b_{k}, a_{j'}b_{k'}): \{a_j: j = 1, 2, \dots, J\}, B = \{ b_k: k = 1, 2, \dots, K\}$.
-The pairwise distances between pairs $(a_{j} b_{k}, a_{j'}b_{k'})$  denoted by $d_{(jk, j'k')} = JSD(q_{jk}, q_{j'k'})$ is computed (_distance between distributions_).
-The pairwise distances (_Within-facet and between-facet_) are transformed using a suitable _tuning parameter_ ($0<\lambda<1$) depending on if they are within-facet($d_w$) or between-facets($d_b$) as follows:
+This is repeated for all pairs of categories of the form $(a_{j} b_{k}, a_{j'}b_{k'}): \{a_j: j = 1, 2, \dots, \nx\}, B = \{ b_k: k = 1, 2, \dots, \nf\}$.
+The pairwise distances between pairs $(a_{j} b_{k}, a_{j'}b_{k'})$ denoted by $d_{(jk, j'k')} = JSD(q_{jk}, q_{j'k'})$ are computed (_distance between distributions_).
+The pairwise distances (_within-facet and between-facet_) are transformed using a suitable _tuning parameter_ ($0<\lambda<1$) depending on if they are within-facet($d_w$) or between-facets($d_b$) as follows:
 \begin{equation}\label{def:raw-wpd}
@@
     \begin{cases}
-\lambda d_{(jk), (j'k')},& \text{if } d = d_w\\
-(1-\lambda)  d_{(jk), (j'k')},              & \text{if } d = d_b\\
+      \lambda d_{(jk), (j'k')},& \text{if } d = d_w;\\
+      (1-\lambda) d_{(jk), (j'k')},       & \text{if } d = d_b.
     \end{cases}
 \end{equation}
-
-\noindent The $wpd_{raw}$ is then computed as $$wpd = max_{j, j', k, k'}(d*_{(jk), (j'k')}) \forall j, j' \in \{1, 2, \dots, J\}, k, k' \in \{1, 2, \dots, K\}$$
-
-\noindent The statistic "maximum" is chosen to combine the weighted pairwise distances since the distance measure is aimed at capturing the maximum variation of the measured variable within a panel.
+The $\wpdsub{raw}$ is then computed as
+$$
+  \wpdsub{raw} = \max_{j, j', k, k'}(d*_{(jk), (j'k')}) \qquad\forall j, j' \in \{1, 2, \dots, \nx\},\quad k, k' \in \{1, 2, \dots, \nf\}
+$$
+The statistic "maximum" is chosen to combine the weighted pairwise distances since the distance measure is aimed at capturing the maximum variation of the measured variable within a panel.
 The statistic "maximum" is, however, affected by the number of comparisons (resulting pairwise distances).
-For example, for a $(2, 3)$ panel, there are $6$ possible subsets of obervations corresponding to the combinations $(a_1, b_1), (a_1, b_2), (a_1, b_3), (a_2 ,b_1), (a_2 ,b_2), (a_2, b_3)$, whereas for a $(2, 2)$ panel, there are only $4$ possible subsets $(a_1, b_1), (a_1, b_2), (a_2 ,b_1), (a_2 ,b_2)$.
+For example, for a $(2, 3)$ panel, there are $6$ possible subsets of observations corresponding to the combinations $(a_1, b_1), (a_1, b_2), (a_1, b_3), (a_2 ,b_1), (a_2 ,b_2), (a_2, b_3)$, whereas for a $(2, 2)$ panel, there are only $4$ possible subsets $(a_1, b_1), (a_1, b_2), (a_2 ,b_1), (a_2 ,b_2)$.
 Consequently, the measure would have higher values for the panel $(2, 3)$ as compared to $(2, 2)$, since maximum is taken over higher number of pairwise distances.
 
-#### Adjusting for the number of comparisons {#sec:norm-wpd}
-
-Ideally, it is desired that the proposed distance measure takes a higher value only if there is a significant difference between distributions across categories, and not because the number of categories $J$ or $K$ is high.
-That is, under designs like $D_{null}$, their distribution should not differ for a different number of categories.
+### Adjusting for the number of comparisons
+
+Ideally, it is desired that the proposed distance measure takes a higher value only if there is a significant difference between distributions across categories, and not because the number of categories $\nx$ or $\nf$ is high.
+That is, under designs like $\Dnull$, their distribution should not differ for a different number of categories.
 Only then the distance measure could be compared across panels with different levels.
 This calls for an adjusted measure, which normalizes for the different number of comparisons.
-We denote it by $wpd$.
+
 Two approaches for adjusting the number of comparisons are discussed, both of which are substantiated using simulations.
-The first one defines an adjusted measure $wpd_{perm}$ based on the permutation method to remove the effect of different comparisons.
-The second approach fits a model to represent the relationship between $wpd_{raw}$ and the number of comparisons and defines the adjusted measure ($wpd_{glm}$) as the residual from the model.
-
-_Permutation approach_
+The first one defines an adjusted measure $\wpdsub{perm}$ based on the permutation method to remove the effect of different comparisons.
+The second approach fits a model to represent the relationship between $\wpdsub{raw}$ and the number of comparisons and defines the adjusted measure ($\wpdsub{glm}$) as the residual from the model.
+
+#### Permutation approach {-}
 
@@
 This method accomplishes a different goal of finding the null distribution for different groups (panels in our case) and standardizing the raw values using that distribution.
-The values of $wpd_{raw}$ is computed on many ($nperm$) permuted data sets and stored in $wpd_{perm-data}$.
-Then $wpd_{perm}$ is computed as follows:
-
+The values of $\wpdsub{raw}$ is computed on many ($\nsub{perm}$) permuted data sets and stored in $\wpdsub{perm-data}$.
+Then $\wpdsub{perm}$ is computed as follows:
 \begin{align*}
-wpd_{perm} = \frac{(wpd_{raw} - \bar{wpd_{perm-data}})}{sd(wpd_{perm-data})}
+  \wpdsub{perm} = \frac{(\wpdsub{raw} - \text{mean}(\wpdsub{perm-data}))}{\text{sd}(\wpdsub{perm-data})},
 \end{align*}
-
-\noindent where $\bar{wpd_{perm-data}}$ and $sd(wpd_{perm-data})$ are the mean and standard deviation of $wpd_{perm-data}$ respectively.
-Standardizing  $wpd$ in the permutation approach ensures that the distribution of $wpd_{perm}$ under $D_{null}$ has the same $mean = 0$ and $\sigma^2_{perm} = 1$ across all comparisons.
-While this works successfully to make the location and scale similar across different $nx$ and $nfacet$, it is computationally heavy and time consuming, and hence less user friendly when being actually used in practice.
+where $\text{mean}(\wpdsub{perm-data})$ and $\text{sd}(\wpdsub{perm-data})$ are the mean and standard deviation of $\wpdsub{perm-data}$ respectively.
+Standardizing $\wpd$ in the permutation approach ensures that the distribution of $\wpdsub{perm}$ under $\Dnull$ has zero mean and unit variance across all comparisons.
+While this works successfully to make the location and scale similar across different $\nx$ and $\nf$, it is computationally heavy and time consuming, and hence less user-friendly.
 Hence, another approach to adjustment, with potentially less computational time, is proposed.
@@
 
-_Modeling approach_
-
-In this approach, a Gamma generalized linear model (GLM) for $wpd_{raw}$ is fitted with number of comparisons as the explanatory variable.
-Since, $wpd_{raw}$ is a Jensen-Shannon distance, it follows a Chi-square distribution (@Menendez1997-in), which is a special case of Gamma distribution.
-Furthermore, the mean response is bounded, since any JSD is bounded by $1$ given that base $2$ logarithm is used (@Lin1991-pt).
-Hence, by @Faraway2016-uk, an inverse link is used for the model, which is of the  form $y = a+b*log(z) + e$, where $y = wpd_{raw}$, $z = (nx*nfacet)$ is the number of groups and $e$ are idiosyncratic errors.
-Let $E(y) = \mu$ and $a + b*log(z) = g(\mu)$ where $g$ is the link function.
-Then $g(\mu)=  1/\mu$ and $\hat \mu  = 1/(\hat a + \hat b log(z))$.
-The residuals from this model $(y-\hat y) = (y-1/(\hat a + \hat b log(z)))$ would be expected to have no dependency on $z$.
-Thus, $wpd_{glm}$ is chosen as the residuals from this model and is defined as:
-$$wpd_{glm} = wpd_{raw} - 1/(\hat a + \hat b*log(nx*nfacet))$$
-The distribution of $wpd_{glm}$ under $D_{null}$ will have $mean = 0$, since it is the residuals from the model, and a constant variance $\sigma^2_{glm}$, which might not equal 1.
-
-_Combination approach_
-
-The simulation results (in Section \ref{sec:simulations}) show that the distribution of $wpd_{glm}$ under null is similar for high $nx$ and $nfacet$ (levels higher than $5$) and not so much for lower $nx$ and $nfacet$.
-Hence, a combination approach is proposed which chooses permutation approach for categories with smaller levels and modeling approach for categories with higher levels.
-This ensures that the computational load of the permutation approach is alleviated while maintaining similar null distribution across different categories.
+#### Modeling approach {-}
+
+In this approach, a Gamma generalized linear model (GLM) for $\wpdsub{raw}$ is fitted with the number of comparisons as the explanatory variable.
+Since, $\wpdsub{raw}$ is a Jensen-Shannon distance, it follows a Chi-square distribution [@Menendez1997-in], which is a special case of a Gamma distribution.
+Furthermore, the mean response is bounded, since any JSD is bounded by $1$ if a base $2$ logarithm is used [@Lin1991-pt].
+Hence, by @Faraway2016-uk, an inverse link is used for the model, which is of the form $y = a+b*\log(z) + e$, where $y = \wpdsub{raw}$, $z = (\nx*\nf)$ is the number of groups and $e$ are idiosyncratic errors.
+Let $\text{E}(y) = \mu$ and $a + b*\log(z) = g(\mu)$ where $g(\mu)= 1/\mu$ and $\hat \mu = 1/(\hat a + \hat b \log(z))$.
+The residuals from this model $(y-\hat y) = (y-1/(\hat a + \hat b \log(z)))$ would be expected to have no dependency on $z$.
+Thus, $\wpdsub{glm}$ is defined as the residuals from this model given by
+$$\wpdsub{glm} = \wpdsub{raw} - 1/(\hat a + \hat b*\log(\nx*\nf))$$
+The distribution of $\wpdsub{glm}$ under $\Dnull$ will have approximately zero mean and a constant variance (not necessarily 1).
+
+#### Combination approach {-}
+
+The simulation results (in Section \ref{sec:simulations}) show that the distribution of $\wpdsub{glm}$ under the null design is similar for high $\nx$ and $\nf$ (levels higher than $5$) but less so for lower values of $\nx$ and $\nf$.
+Hence, a combination approach is proposed where we use a permutation approach for categories with small numbers of levels, and a modeling approach for categories with higher numbers of levels.
+This ensures that the computational load of the permutation approach is alleviated while maintaining a similar null distribution across different categories.
 This approach, however, requires that the adjusted variables from the two approaches are brought to the same scale.
-We define $wpd_{glm-scaled} = wpd_{glm}*\sigma^2_{perm}/\sigma^2_{glm}$ as the transformed $wpd_{glm}$ with a similar scale as $wpd_{perm}$.
-The adjusted measure from the combination approach, denoted by $wpd$ is then defined as follows:
-
+We define $\wpdsub{glm-scaled} = \wpdsub{glm}*\sigma^2_{\text{perm}}/\sigma^2_{\text{glm}}$ as the transformed $\wpdsub{glm}$ with a similar scale as $\wpdsub{perm}$.
+The adjusted measure from the combination approach, denoted by $\wpd$ is then defined as follows:
 \begin{equation}
-wpd =
-\begin{cases}
-wpd_{perm},& \text{if } J, K <=5\\
-wpd_{glm-scaled}             & \text{otherwise}\\
+\wpd = \begin{cases}
+         \wpdsub{perm},       & \text{if $\nx, \nf \le 5$};\\
+         \wpdsub{glm-scaled}  & \text{otherwise}.\\
        \end{cases}
@@
 
-<!-- The distance measure $wpd$ between two cyclic granularities $A$ and $B$ is aimed to capture the strength of the structure by estimating the maximum within-group and between-group variations.
+<!-- The distance measure $\wpd$ between two cyclic granularities $A$ and $B$ is aimed to capture the strength of the structure by estimating the maximum within-group and between-group variations.
 Furthermore, -->
 
-<!-- The steps employed for computing $wpd$ is summarized as follows: -->
-
-<!-- The values of $wpd$ indeed depend upon the underlying distribution.
-This is also not desirable as the intended aim of $wpd$ is to capture differences in categories irrespective of the distribution from which the data is generated.
+<!-- The steps employed for computing $\wpd$ is summarized as follows: -->
+
+<!-- The values of $\wpd$ indeed depend upon the underlying distribution.
+This is also not desirable as the intended aim of $\wpd$ is to capture differences in categories irrespective of the distribution from which the data is generated.
 The steps employed for computing the distance measure is summarized as follows: -->
 
-<!-- Hence, differences in distributions of the measured variable are computed between all pairs of categories $(a_{j} b_{k}, a_{j'}b_{k'}): j = 1, 2, \dots, J\}, B = \{ b_k: k = 1, 2, \dots, K\}$.
+<!-- Hence, differences in distributions of the measured variable are computed between all pairs of categories $(a_{j} b_{k}, a_{j'}b_{k'}): j = 1, 2, \dots, \nx\}, B = \{ b_k: k = 1, 2, \dots, \nf\}$.
 Percentiles of $v_{jk}$ and $v_{j'k'}$ are computed and stored in $q_{jk}$ and $q_{j'k'}$ respectively.
@@
 Pairwise distances are between-facets ($d_{b}$) when they are considered between pairs of the form $(a_{j}b_{k}, a_{j}b_{k'})$.
-Number of between-facet distances would be $^{K}C_2*J$ and number of within-facet distances are $K*(J-1)$ (ordered) and $^{J}C_2*K$ (un-ordered).
+Number of between-facet distances would be $^{\nf}C_2*\nx$ and number of within-facet distances are $\nf*(\nx-1)$ (ordered) and $^{\nx}C_2*\nf$ (un-ordered).
 The pairwise distances $d_{(jk, j'k')}$ are transformed using a suitable tuning parameter ($0<\lambda<1$) depending on if they are $d_{b}$ or $d_{w}$ as follows: -->
 
-```{r distance-explain, fig.cap = "Within and between-facet distances shown for two cyclic granularities A and B, where A is mapped to x-axis and B is mapped to facets.
-The dotted lines represent the distances between different categories.
-Panel 1) and 2) show the between-facet distances.
-Panel 3) and 4) are used to illustrate within-facet distances when categories are un-ordered or ordered respectively.
-When categories are ordered, distances should only be considered for consecutive x-axis categories.
-Between-facet distances are distances between different facet levels for the same x-axis category, for example, distances between {($a_1$,$b_1$) and ($a_1$, $b_2$)} or {($a_1$,$b_1$) and ($a_1$, $b_3$)}."}
-
-```
-
 <!-- 5. Compute distribution of maximum distance ($M$) by shuffling the data $200$ times and finishing Steps1-4 in each case. Scale $M$ using the mean and sd of the distribution.
 
-5. Use Steps 1-4 to compute maximum distance for $\forall k \in  \{1, 2, \ldots, K\}$.
+5. Use Steps 1-4 to compute maximum distance for $\forall k \in \{1, 2, \ldots, \nf\}$.
 
@@
 
-<!-- All of these distances are then aggregated by taking the maximum from these distances to obtain $wpd$. -->
+<!-- All of these distances are then aggregated by taking the maximum from these distances to obtain $\wpd$. -->
 
@@
  <!-- old -->
-<!-- Consider two cyclic granularities $C_i$ and $C_j$, such that $C_i$ maps index set to a set $\{A_k \mid k=1,\dots,K\}$ and $C_j$ maps index set to a set $\{B_\ell \mid \ell =1,\dots,L\}$.
+<!-- Consider two cyclic granularities $C_i$ and $C_j$, such that $C_i$ maps index set to a set $\{A_k \mid k=1,\dots,\nf\}$ and $C_j$ maps index set to a set $\{B_\ell \mid \ell =1,\dots,L\}$.
 Here, $A_k$ and $B_\ell$ are the levels/categories corresponding to $C_i$ and $C_j$ respectively.
@@
 
-<!-- 5. Use Steps 1-4 to compute maximum distance for $\forall k \in  \{1, 2, \ldots, K\}$. -->
+<!-- 5. Use Steps 1-4 to compute maximum distance for $\forall k \in \{1, 2, \ldots, \nf\}$. -->
 
@@
 <!-- \begin{algorithm}[!thb] -->
-<!-- 	\caption{Calculation for a raw distance measure between two cyclic granularities $A = \{ a_j: j = 1, 2, \dots, J\}$, $B = \{ b_k: k = 1, 2, \dots, K\}$ with $A$ placed across x-axis and $B$ across facets.}\label{alg:scoreopt}.
+<!-- 	\caption{Calculation for a raw distance measure between two cyclic granularities $A = \{ a_j: j = 1, 2, \dots, \nx\}$, $B = \{ b_k: k = 1, 2, \dots, \nf\}$ with $A$ placed across x-axis and $B$ across facets.}\label{alg:scoreopt}.
 -->
 <!-- 	\begin{algorithmic}[1] -->
-<!-- 		\Procedure{RawMMPD}{$A = \{ a_j: j = 1, 2, \dots, J\}$, $B = \{ b_k: k = 1, 2, \dots, K\}$, $v = \{ v_t: t = 1, 2, \dots, T\}$}. -->
-<!-- 		\For{$k=1:K$, $j=1:J$} -->
+<!-- 		\Procedure{RawMMPD}{$A = \{ a_j: j = 1, 2, \dots, \nx\}$, $B = \{ b_k: k = 1, 2, \dots, \nf\}$, $v = \{ v_t: t = 1, 2, \dots, T\}$}. -->
+<!-- 		\For{$k=1:\nf$, $j=1:\nx$} -->
 <!-- 		\State Find distances between pairs of all possible combinations of categories $(a_jb_k,a_j'b_k')$ by computing JSD between quantiles of the measured variable $q(v)$ across these combinations.
@@
 
-<!-- ## Properties of $wpd$ -->
-
-<!-- Simulations were carried out to explore the behavior of $wpd$ under the following factors that could potentially impact the values of $wpd$: $nx$, $nfacet$, $\lambda$, $\omega$, $dist$ (normal/non-normal distributions with different location and scale), $ntimes$, and $designs$ and results are presented in two parts.
-The dependence of $wpd$ on $nx$ and $nfacet$ under $D_{null}$ is presented here, which lays the foundation for the next section.
-The rest of the results that discuss the relationship of the $wpd$ with other factors is presented in details in the Supplementary section of the paper.
-They show that the designs $D_{var_f}$ and $D_{var_x}$ intersect at $\lambda = 0.5$ and hence for up-weighing designs of the form $D_{var_x}$, $\lambda = 0.67$ has been considered for computation of $wpd$ in the rest of the paper.
--->
-
-<!-- - $nx$ (number of levels of x-axis) -->
-<!-- - $nfacet$ (number of levels of facets) -->
+<!-- ## Properties of $\wpd$ -->
+
+<!-- Simulations were carried out to explore the behavior of $\wpd$ under the following factors that could potentially impact the values of $\wpd$: $\nx$, $\nf$, $\lambda$, $\omega$, $dist$ (normal/non-normal distributions with different location and scale), $ntimes$, and $designs$ and results are presented in two parts.
+The dependence of $\wpd$ on $\nx$ and $\nf$ under $\Dnull$ is presented here, which lays the foundation for the next section.
+The rest of the results that discuss the relationship of the $\wpd$ with other factors is presented in details in the Supplementary section of the paper.
+They show that the designs $\Df$ and $D_x$ intersect at $\lambda = 0.5$ and hence for up-weighing designs of the form $D_x$, $\lambda = 0.67$ has been considered for computation of $\wpd$ in the rest of the paper.
+-->
+
+<!-- - $\nx$ (number of levels of x-axis) -->
+<!-- - $\nf$ (number of levels of facets) -->
 <!-- - $\lambda$ (tuning parameter) -->
@@
 <!-- - $n$ (sample size for each combination of categories) -->
-<!-- - $designs$ ($D_{null}$, $D_{var_f}$, $D_{var_x}$ and $D_{var_{all}}$) -->
-
-<!-- - $nsim$ (number of simulations) -->
-<!-- - $nperm$ (number of permutations of data)   -->
+<!-- - $designs$ ($\Dnull$, $\Df$, $D_x$ and $D_{fx}$) -->
+
+<!-- - $\nsub{sim}$ (number of simulations) -->
+<!-- - $\nsub{perm}$ (number of permutations of data)  -->
 
@@
 
-<!-- Observations are generated from a Gamma(2,1) distribution for each combination of $nx$ and $nfacet$ from the following sets: $nx = nfacet = \{2, 3, 5, 7, 14, 20, 31, 50\}$ to cover a wide range of levels from very low to moderately high.
+<!-- Observations are generated from a Gamma(2,1) distribution for each combination of $\nx$ and $\nf$ from the following sets: $nx = \nf = \{2, 3, 5, 7, 14, 20, 31, 50\}$ to cover a wide range of levels from very low to moderately high.
 Each combination is being referred to as a _panel_.
-That is, data is being generated for each of the panels $\{nx = 2, nfacet = 2\}, \{nx = 2, nfacet = 3\}, \{nx = 2, nfacet = 5\},  \dots, \{nx = 50, nfacet = 31\}, \{nx = 50, nfacet = 50\}$.
+That is, data is being generated for each of the panels $\{nx = 2, \nf = 2\}, \{nx = 2, \nf = 3\}, \{nx = 2, \nf = 5\}, \dots, \{nx = 50, \nf = 31\}, \{nx = 50, \nf = 50\}$.
 For each of the $64$ panels, $ntimes = 500$ observations are drawn for each combination of the categories.
-That is, if we consider the panel $\{nx = 2, nfacet = 2\}$, $500$ observations are generated for each of the combination of categories from the panel, namely, $\{(1, 1), (1, 2), (2, 1), (2, 2)\}$.
+That is, if we consider the panel $\{nx = 2, \nf = 2\}$, $500$ observations are generated for each of the combination of categories from the panel, namely, $\{(1, 1), (1, 2), (2, 1), (2, 2)\}$.
 The values -->
-<!-- of $wpd$ is obtained for each of the panels.
-This design corresponds to $D_{null}$ as each combination of categories in a panel are drawn from the same distribution.
-Furthermore, the data is simulated for each of the panels $nsim=200$ times, so that the distribution oaf $wpd$ under $D_{null}$ could be observed.
-$wpd_{l, s}$ denotes the value of $wpd$ obtained for the $l^{th}$ panel and $s^{th}$ simulation.
+<!-- of $\wpd$ is obtained for each of the panels.
+This design corresponds to $\Dnull$ as each combination of categories in a panel are drawn from the same distribution.
+Furthermore, the data is simulated for each of the panels $\nsub{sim}=200$ times, so that the distribution oaf $\wpd$ under $\Dnull$ could be observed.
+$wpd_{l, s}$ denotes the value of $\wpd$ obtained for the $l^{th}$ panel and $s^{th}$ simulation.
 -->
@@
 
-<!-- Figure \ref{fig:raw} shows the distribution of $wpd$ plotted across different $nx$ and $nfacet$ categories.
-Since under $D_{null}$, there is no difference in distributions across different categories, we expect the distance measure $wpd$ to reflect that as well and have the same distribution across categories.
+<!-- Figure \ref{fig:raw} shows the distribution of $\wpd$ plotted across different $\nx$ and $\nf$ categories.
+Since under $\Dnull$, there is no difference in distributions across different categories, we expect the distance measure $\wpd$ to reflect that as well and have the same distribution across categories.
 But Figure \ref{fig:raw} shows that both the location and scale of the distributions change across panels.
-This is not desirable under $D_{null}$ as it would mean comparisons of $wpd$ values is not appropriate across different $nx$ and $nfacet$.
-Figure \ref{fig:quadratic} shows how the median of $wpd$ varies with the total number of distances $nx*nfacet$ for each panel.
-The median increases abruptly for lower values of $nx*nfacet$ and slowly for higher $nx*nfacet$.
--->
-
-<!-- Furthermore, simulation results also show that the values of $wpd$ indeed depend upon the underlying distribution.
-This is also not desirable as the intended aim of $wpd$ is to capture differences in categories irrespective of the distribution from which the data is generated.
+This is not desirable under $\Dnull$ as it would mean comparisons of $\wpd$ values is not appropriate across different $\nx$ and $\nf$.
+Figure \ref{fig:quadratic} shows how the median of $\wpd$ varies with the total number of distances $nx*\nf$ for each panel.
+The median increases abruptly for lower values of $nx*\nf$ and slowly for higher $nx*\nf$.
+-->
+
+<!-- Furthermore, simulation results also show that the values of $\wpd$ indeed depend upon the underlying distribution.
+This is also not desirable as the intended aim of $\wpd$ is to capture differences in categories irrespective of the distribution from which the data is generated.
 -->
@@
 <!-- last writeup -->
-<!-- In the linear model approach, $wpd\in R$ was assumed, whereas, $wpd$ is a Jensen-Shannon Distance (JSD) and lies between 0 and 1 (@JSD).
+<!-- In the linear model approach, $wpd\in R$ was assumed, whereas, $\wpd$ is a Jensen-Shannon Distance (JSD) and lies between 0 and 1 [@JSD].
 Furthermore, JSD follows a Chi-square distribution, which is a special case of Gamma distribution.
@@
 The residuals from this model $(y-\hat y) = (y-1/(\hat a + \hat b log(z)))$ would be expected to have no dependency on $z$.
-Thus, $wpd_{glm}$ is chosen as the residuals from this model and is defined as: -->
-<!-- $wpd_{glm} = wpd - 1/(\hat a + \hat b*log(nx*nfacet))$. -->
+Thus, $\wpdsub{glm}$ is chosen as the residuals from this model and is defined as: -->
+<!-- $\wpdsub{glm} = wpd - 1/(\hat a + \hat b*log(nx*\nf))$. -->
 
@@
 <!-- use later -->
-<!-- The measure $wpd$ could potentially lead to comparison of the measure across different panels and also help distinguishing the interesting panels from a data set.
+<!-- The measure $\wpd$ could potentially lead to comparison of the measure across different panels and also help distinguishing the interesting panels from a data set.
 We discuss two approaches for normalization, both of which are substantiated using simulations.
@@
 
-<!-- The transformed ${wpd}$ which is normalized for the values of $nx$ and $nfacet$ is denoted by $wpd$.
-Two approaches have been employed - the first one involves a permutation method to make the distribution of the transformed $wpd$ similar for different comparisons and the second one fits a model to represent the relationship between the two variables and defines $wpd$ as the residual of the model.
+<!-- The transformed ${wpd}$ which is normalized for the values of $\nx$ and $\nf$ is denoted by $\wpd$.
+Two approaches have been employed - the first one involves a permutation method to make the distribution of the transformed $\wpd$ similar for different comparisons and the second one fits a model to represent the relationship between the two variables and defines $\wpd$ as the residual of the model.
 -->
@@
 
-<!-- Let $\{nx_{i}, i = 1, 2, \dots, nx\}$, $\{nfacet_{j}, j = 1, 2, \dots, nfacet\}$ be the set of x-axis and facet categories respectively.
-Each combination of $nx_{i}$ and $nfacet_{j}$is being referred to as a _panel_.
-Then the total number of panel is $nx*nfacet$.
-Let the total number of pairwise distances that could result in each panel be $\{z_k, k = 1, 2 , \dots, nx*nfacet\}$.
-Here, $\{z_1 = nx_1*nfacet_1\}$, $\{z_2 = nx_2*nfacet_2\}$ and $\{z_k = nx*nfacet\}$.
-Now, let $\{x_{k,l}, k = 1, 2, \dots, nx*nfacet, l = 1, 2, \dots, nsim\}$ denote the values of $wpd_{raw}$ obtained from the simulation study for $k^{th}$ panel in the $i^{th}$ simulation.
-Hence, for each of those $k$ panel, we have $nsim$ values of $wpd_{raw}$.
+<!-- Let $\{nx_{i}, i = 1, 2, \dots, nx\}$, $\{\nf_{j}, j = 1, 2, \dots, \nf\}$ be the set of x-axis and facet categories respectively.
+Each combination of $nx_{i}$ and $\nf_{j}$is being referred to as a _panel_.
+Then the total number of panel is $nx*\nf$.
+Let the total number of pairwise distances that could result in each panel be $\{z_k, k = 1, 2 , \dots, nx*\nf\}$.
+Here, $\{z_1 = nx_1*\nf_1\}$, $\{z_2 = nx_2*\nf_2\}$ and $\{z_k = nx*\nf\}$.
+Now, let $\{x_{k,l}, k = 1, 2, \dots, nx*\nf, l = 1, 2, \dots, \nsub{sim}\}$ denote the values of $\wpdsub{raw}$ obtained from the simulation study for $k^{th}$ panel in the $i^{th}$ simulation.
+Hence, for each of those $k$ panel, we have $\nsub{sim}$ values of $\wpdsub{raw}$.
 -->
@@
 
-<!-- This section reports the results of a simulation study that was carried out to evaluate the behavior of $wpd$ under different designs and other potential factors.
-The behavior of $wpd$ is explored in designs where there is in fact difference in distribution between facet categories ($D_{var_f}$) or across x-categories ($D_{var_x}$) or both ($D_{var_{all}}$).
-Using $\omega =  \{1, 2, \dots, 10\}$ and $\lambda = seq(from  = 0.1, to = 0.9, by = 0.05)$, observations are drawn from a N(0,1) distribution for each combination of $nx$ and $nfacet$ from the following sets: $nx = nfacet =  \{2, 3, 5, 7, 14, 20, 31, 50\}$.
+<!-- This section reports the results of a simulation study that was carried out to evaluate the behavior of $\wpd$ under different designs and other potential factors.
+The behavior of $\wpd$ is explored in designs where there is in fact difference in distribution between facet categories ($\Df$) or across x-categories ($D_x$) or both ($D_{fx}$).
+Using $\omega = \{1, 2, \dots, 10\}$ and $\lambda = seq(from = 0.1, to = 0.9, by = 0.05)$, observations are drawn from a N(0,1) distribution for each combination of $\nx$ and $\nf$ from the following sets: $nx = \nf = \{2, 3, 5, 7, 14, 20, 31, 50\}$.
 $ntimes = 500$ is assumed for this setup as well.
-Furthermore, to generate different distributions across different combination of facet and x levels, the following method is deployed - suppose the distribution of the combination of first levels of $x$ and $facet$ category is $N(\mu,\sigma)$ and $\mu_{jk}$ denotes the mean of the combination $(a_jb_k)$, then $\mu_{j.} = \mu + j\omega$ (for design $D_{var_x}$) and $\mu_{.k} = \mu + k\omega$ (for design $D_{var_f}$).
+Furthermore, to generate different distributions across different combination of facet and x levels, the following method is deployed - suppose the distribution of the combination of first levels of $x$ and $facet$ category is $N(\mu,\sigma)$ and $\mu_{jk}$ denotes the mean of the combination $(a_jb_k)$, then $\mu_{j.} = \mu + j\omega$ (for design $D_x$) and $\mu_{.k} = \mu + k\omega$ (for design $\Df$).
 -->
@@
 <!-- The tabulated values and graphical representations of the simulation results are provided in the Supplementary paper.
-The learning from the simulations are as follows: The values of $wpd$ is least for $D_{null}$, followed by -->
-<!-- $D_{var_f}$, $D_{var_x}$ and $D_{var_{all}}$.
-This is a desirable result since the measure $wpd$ was designed such that this relationship holds.
-Furthermore, the distribution of the measure $wpd$ does not change for different facet and x categories.
-The distribution of $wpd$ looks similar with at least the mean and standard of the distributions being uniform across panels.
-This means $wpd$ could be used to measure differences in distribution across panels.
+The learning from the simulations are as follows: The values of $\wpd$ is least for $\Dnull$, followed by -->
+<!-- $\Df$, $D_x$ and $D_{fx}$.
+This is a desirable result since the measure $\wpd$ was designed such that this relationship holds.
+Furthermore, the distribution of the measure $\wpd$ does not change for different facet and x categories.
+The distribution of $\wpd$ looks similar with at least the mean and standard of the distributions being uniform across panels.
+This means $\wpd$ could be used to measure differences in distribution across panels.
 Also, note that since the data is processed using normal-quantile-transform, this measure is independent of the initial distribution of the underlying data and hence is also comparable across different data sets.
-This is valid for the case when sample size $ntimes$ for each combination of categories is at least 30 and  $nperm$ used for computing $wpd$ is at least 100.
-More detailed results about the properties of $wpd$ could be found in the Supplementary paper.
+This is valid for the case when sample size $ntimes$ for each combination of categories is at least 30 and $\nsub{perm}$ used for computing $\wpd$ is at least 100.
+More detailed results about the properties of $\wpd$ could be found in the Supplementary paper.
 -->
@@
 A cyclic granularity is referred to as "significant" if there is a significant distributional difference of the measured variable between different categories of the harmony.
-In this section, a selection criterion to choose significant harmonies is provided, thereby eliminating all harmonies that exhibit complete randomness in the measured variable.
-The distance measure $wpd$ is used as a test statistic to test the null hypothesis that no harmony/cyclic granularity is significant.
+In this section, a selection criterion to choose significant harmonies is provided, thereby eliminating all harmonies that exhibit non-significant differences in the measured variable.
+The distance measure $\wpd$ is used as a test statistic to test the null hypothesis that no harmony/cyclic granularity is significant.
 We select only those harmonies/cyclic granularities for which the test fails.
-They are then ranked basis how well they capture variation in the measured variable.
+They are then ranked based on how well they capture variation in the measured variable.
 
@@
 
-A threshold and consequently a selection criterion is chosen using the notion of Randomization tests (@edgington2007randomization).
-The data is permuted several times and $wpd$ is computed for each of the permuted data sets to obtain the sampling distribution of $wpd$ under the null hypothesis.
-If the null hypothesis is true, then $wpd$ obtained from the original data set would be a likely value in the sampling distribution.
-But in case the null hypothesis is not true, then it is less probable that $wpd$ obtained for the original data will be from the same distribution.
-This idea is utilized to come up with a threshold for selection, denoted by $wpd_{threshold}$, defined as the $99^{th}$ percentile of the sampling distribution.
-A harmony is selected if the value of $wpd$ for that harmony is greater than the chosen threshold.
+A threshold (and consequently a selection criterion) is chosen using the notion of randomization tests [@edgington2007randomization].
+The data is permuted several times and $\wpd$ is computed for each of the permuted data sets to obtain the sampling distribution of $\wpd$ under the null hypothesis.
+If the null hypothesis is true, then $\wpd$ obtained from the original data set would be a likely value in the sampling distribution.
+But in case the null hypothesis is not true, then it is less probable that $\wpd$ obtained for the original data will be from the same distribution.
+This idea is utilized to come up with a threshold for selection, denoted by $\wpdsub{threshold}$, defined as the $99^{th}$ percentile of the sampling distribution.
+A harmony is selected if the value of $\wpd$ for that harmony is greater than the chosen threshold.
 The detailed algorithm for choosing a threshold and selection procedure (for two cyclic granularities) is listed as follows:
 
-<!-- The selection criterion is stated as follows: for $(i \in {1, 2, \dots, H_{N_C}})$, select $H_i$ if $wpd_i > wpd_{threshold}$, else reject. -->
-
-- **Input:** All harmonies of the form $\{(A, B), A = \{ a_j: j = 1, 2, \dots, J\}, B = \{ b_k: k = 1, 2, \dots, K\}\}$, $\forall (A, B) \in H_{N_C}$.
-
-- **Output:** Harmony pairs $(A, B)$ for which $wpd$ is significant.
-
-1. Fix harmony pair $(A, B)$.
-
-2. Given the measured variable; $\{v_t: t=0, 1, 2, \dots, T-1\}$, $wpd$ is computed and is represented by $wpd^{A, B}_{obs}$.
-
-3. From the original sequence a random permutation is obtained: $\{v_t^{1}: t=0, 1, 2, \dots, T-1\}$.
-
-4. $wpd$ is computed for the permuted sequence of the data and is represented by $wpd^{A, B}_{1}$.
-
-5.  Steps (3) and (4) are repeated a large number of times $M$ (e.g. $M = 200$).
-
-6. For each permutation, one $wpd^{A, B}_{i}$ is obtained. Define $wpd_{sample} = \{wpd^{A, B}_{1} wpd^{A, B}_{2}, \dots, wpd^{A, B}_{M}\}$.
-
-7. Repeat Steps (1-6) for all harmony pairs $(A, B) \in H_{N_C}$ and stored $wpd^{all}_{sample}$.
-
-8. $99^{th}$ percentiles of $wpd^{all}_{sample}$ is computed and stored in $wpd_{threshold99}$
-
-7. If $wpd^{A, B}_{obs} > wpd_{threshold99}$, harmony pair $(A, B)$ is selected, otherwise rejected.
-
-Similarly, a harmony pair $(A, B)$ could be selected if $wpd^{A, B}_{obs} > wpd_{threshold95}$ and $wpd^{A, B}_{obs} > wpd_{threshold90}$, where  $wpd_{threshold95}$ and $wpd_{threshold90}$ denote the $95^{th}$ and $90^{th}$ percentile of $wpd^{all}_{sample}$ respectively.
-A harmony selected using $99^{th}$,  $95^{th}$  and  $90^{th}$ threshold are tagged as \*\*\*, \*\*, \* respectively.
-
-### Ranking
-
-The distribution of $wpd$ is expected to be similar for all harmonies under the null hypothesis, since they have been adjusted for different number of categories for the harmonies or underlying distribution of the measured variable.
-Hence, the values of $wpd$ for different harmonies are comparable and can be used to rank the significant harmonies.
-A higher value of $wpd$ for a harmony indicates that higher maximum variation in the measured variable is captured through that harmony.
-Figure \ref{fig:null4by2} presents the results of $wpd$ from the illustrative designs in Section \ref{sec:computation-wpd}.
-The value of $wpd$ under null design (a) is the least, followed by (b), (c) and (d).
-This aligns with the principle of $wpd$, which is expected to have lowest value for null designs and highest for designs of the form $D_{var_{all}}$ (d).
-Moreover, note the relative differences in $wpd$  values between (b) and (c).
-The value of the tuning parameter $\lambda$ is set to $0.67$, which has resulted in giving more emphasis to differences in x-axis categories.
-Again consider \ref{fig:onegran-new}(a) and \ref{fig:onegran-new}(b) with a $wpd$ value of 20.5 and 145 respectively.
-This is because there is more gradual increase across hours of the day than months of the year.
-If order of the categories is not considered, they result in a $wpd$ value of $97.8$ and $161$ respectively, which follows from the fact that if we consider difference between any hours of the day, the magnitude will be much larger than if we consider difference between consecutive categories.
-Similarly, \ref{fig:onegran-new}(a) and (b) have $wpd$ values of   $110.79$ and $125.82$ respectively.
-The ranking implies that the distributional differences are more prominent for the second household, as is also seen from the bigger fluctuations in the $90^{th}$ percentile than the first household.
+<!-- The selection criterion is stated as follows: for $(i \in {1, 2, \dots, H_{N_C}})$, select $H_i$ if $wpd_i > \wpdsub{threshold}$, else reject. -->
+
+- **Input:** All harmonies of the form $\{(A, B),~ A = \{ a_j: j = 1, 2, \dots, \nx\},~ B = \{ b_k: k = 1, 2, \dots, \nf\}\}$, $\forall (A, B) \in H_{N_C}$.
+
+- **Output:** Harmony pairs $(A, B)$ for which $\wpd$ is significant.
+
+1. For each harmony pair $(A, B) \in H_{N_C}$, the following steps are taken.
+
+    a. Given the measured variable; $\{v_t: t=0, 1, 2, \dots, T-1\}$, $\wpd$ is computed and is represented by $\wpd^{A, B}_{obs}$.
+    b. For $i=1,\dots,m$, randomly permute the original time series: $\{v_t^{i}: t=0, 1, 2, \dots, T-1\}$ and compute $\wpd^{A, B}_{i}$ from $\{v_t^{i}\}$.
+    c. Define $\wpdsub{sample} = \{\wpd^{A, B}_{1}, \dots, \wpd^{A, B}_{M}\}$.
+
+2. Stack the $\wpdsub{sample}$ vectors as $\wpdsub{sample}^{\text{all}}$ and compute its $p = 100(1-\alpha)$ percentiles as $\wpdsub{threshold\emph{p}}$.
+3. If $\wpd^{A, B}_{obs} > \wpdsub{threshold\emph{p}}$, harmony pair $(A, B)$ is selected at the $1-p/100$ level, otherwise rejected.
+4. Harmonies selected using the $99^{th}$, $95^{th}$ and $90^{th}$ thresholds are tagged as \*\*\*, \*\*, \* respectively.
+
+### Ranking {#sec:ranking}
+
+The distribution of $\wpd$ is expected to be similar for all harmonies under the null hypothesis, since they have been adjusted for different number of categories for the harmonies or underlying distribution of the measured variable.
+Hence, the values of $\wpd$ for different harmonies are comparable and can be used to rank the significant harmonies.
+A higher value of $\wpd$ for a harmony indicates that higher maximum variation in the measured variable is captured through that harmony.
+
+```{r varall-new}
+
+```
+
+```{r varx-new}
+
+```
+
+```{r varf-new}
+```
+
+```{r null-new}
+
+```
+
+Figure \ref{fig:null4by2} also presents the results of $\wpd$ from the illustrative designs in Section \ref{sec:computation-wpd}.
+The value of $\wpd$ under null design (a) is the least, followed by (b), (c) and (d).
+This aligns with the principle of $\wpd$, which is expected to have lowest value for null designs and highest for designs of the form $\Dfx$ (d).
+Moreover, note the relative differences in $\wpd$ values between (b) and (c).
+The value of the tuning parameter $\lambda$ is set to 2/3, which gives greater emphasis to differences in x-axis categories than facets.
+
+Again consider \ref{fig:onegran-new}(a) and \ref{fig:onegran-new}(b) with a $\wpd$ value of 20.5 and 145 respectively.
+This is because there is a more gradual increase across hours of the day than across months of the year.
+If the order of categories is ignored, the resulting $\wpd$ values are $97.8$ and $161$ respectively, because differences between any hours of the day tend to be larger than differences only between consecutive hours.
+Similarly, \ref{fig:onegran-new}(a) and (b) have $\wpd$ values of $110.79$ and $125.82$ respectively.
+The ranking implies that the distributional differences are more prominent for the second household, as is also seen from the bigger fluctuations in the $90^{th}$ percentile than for the first household.
 
@@
 <!-- A randomization test involves calculating a test statistic, randomly shuffling the data and calculating the test statistic several times to obtain a distribution of the test statistic.
-But we will use this procedure to obtain a threshold such that harmony pairs with a $wpd$ value higher than this threshold will only be considered significant.
+But we will use this procedure to obtain a threshold such that harmony pairs with a $\wpd$ value higher than this threshold will only be considered significant.
 The process of choosing a threshold is described as follows: -->
 
-<!-- to test if there is any interesting pattern captured by the harmonies, which essentially implies if $wpd$ is significantly different from zero.
-The percentages of times the $wpd$ obtained from the permuted data is greater than or equal to the observed $wpd$ is the p-value.
+<!-- to test if there is any interesting pattern captured by the harmonies, which essentially implies if $\wpd$ is significantly different from zero.
+The percentages of times the $\wpd$ obtained from the permuted data is greater than or equal to the observed $\wpd$ is the p-value.
 The randomization test is described as follows: -->
@@
 This method is based on the generation of randomly chosen reassignments (permutations) of the data across different cyclic granularities and the computation of -->
-<!-- $wpd$ for each of these reassignments. -->
+<!-- $\wpd$ for each of these reassignments. -->
 
@@
 This method is based on the generation of randomly chosen reassignments (permutations) of the data across different cyclic granularities and the computation of -->
-<!-- $wpd$ for each of these reassignments.
-The percentages of times the theoretical distribution greater than or equal to the respective observed $wpd$ values are calculated and are used to obtain the P value.
+<!-- $\wpd$ for each of these reassignments.
+The percentages of times the theoretical distribution greater than or equal to the respective observed $\wpd$ values are calculated and are used to obtain the P value.
 The procedure for the permutation test is: -->
@@
 For each of the $64$ panels, $ntimes = 500$ observations are drawn for each combination of the categories.
-The values of $wpd$ is obtained for each of the panels for the designs $D_{null}$, $D_{var_x}$, $D_{var_f}$ and $D_{var_{all}}$.
-$wpd_{threshold}$ is computed from all of these panels together and the number of times a harmony pair $(A, B) \in H_{N_C}$ is selected when in fact it was of the design $D_{null}$ is noted.
+The values of $\wpd$ is obtained for each of the panels for the designs $\Dnull$, $D_x$, $\Df$ and $D_{fx}$.
+$\wpdsub{threshold}$ is computed from all of these panels together and the number of times a harmony pair $(A, B) \in H_{N_C}$ is selected when in fact it was of the design $\Dnull$ is noted.
 This entire process is repeated for several null data sets to see the number of times any harmony pair $(A, B) \in H_{N_C}$ is selected under null.
@@
 
-```{r varall-new}
-
-```
-
-```{r varx-new}
-
-```
-
-```{r varf-new}
-```
-
-```{r null-new}
-
-```
-
-```{r plot-all-new3, fig.cap = "$wpd$ values for the four illustrative designs are presented here.
-As expected, the value of $wpd$ under null design is the least (a), followed by (b, c and d).
-Moreover, the relative difference is interesting as c is closer to d than b, since the differences between x categories are up-weighed by design.", dependson= "null4by2"}
-
-```
-
-## Simulations {#sec:simulations}
-
-### Behavior of raw and adjusted distance measures
-
-\noindent _Simulation design_
-
-$m=1$
-
-\noindent Observations are generated from a N(0,1) distribution for a wide range of levels from very low to moderately high.
-$nx = \{2, 3, 5, 7, 9, 14, 17, 20, 24, 31, 42, 50\}$ is considered.
-$ntimes = 500$ observations are drawn for each combination of the categories, that is, for a panel with $nx=3$, $500$ observations are simulated for each of the categories.
-This design corresponds to $D_{null}$ as each combination of categories in a panel are drawn from the same distribution.
-Furthermore, the data is simulated for each of the categories $nsim=200$ times, so that the distribution of $wpd$ under $D_{null}$ could be observed.
-The values of $wpd$ is obtained for each of the panels.
-$wpd_{l, s}$ denotes the value of $wpd$ obtained for the $l^{th}$ panel and $s^{th}$ simulation.
-
-$m=2$  
-\noindent Similarly, observations are generated from a N(0,1) distribution for each combination of $nx$ and $nfacet$ from the following sets: $nx = nfacet = \{2, 3, 5, 7, 14, 20, 31, 50\}$.
-That is, data is being generated for each of the panels $(2, 2), (2, 3), (2, 5) \dots, (50, 31), (50, 50)$.
-For each of the $64$ panels, $ntimes = 500$ observations are drawn for each combination of the categories.
-That is, if we consider a $(2, 2)$ panel, $500$ observations are generated for each of the possible subsets, namely, $\{(1, 1), (1, 2), (2, 1), (2, 2)\}$.
-
-\noindent _Results_
+### Simulations {#sec:simulations}
+
+Simulations were carried out to explore the behavior of $\wpd$ as $\nx$ and $\nf$ were varied, in order to compare and evaluate different normalization approaches.
+More detailed simulation results, including an exploration of the size and power of the proposed test, are contained in the supplements.
+
+#### Simulation design {-}
+
+\noindent $m=1$
+
+\noindent Observations were generated from a N(0,1) distribution for $\nx \in \{2, 3, 5, 7, 9, 14, 17, 20, 24, 31, 42, 50\}$, with $\nsub{times} = 500$ observations drawn for each combination of categories.
+This design corresponds to $\Dnull$.
+For each of the categories, there were $\nsub{sim}=200$ replications, so that the distribution of $\wpd$ under $\Dnull$ could be observed.
+Let $\wpd_{\ell, s}$ denote the value of $\wpd$ obtained for the $\ell^{th}$ panel and $s^{th}$ simulation.
+
+\noindent $m=2$
+
+\noindent Similarly, observations were generated from a N(0,1) distribution for each combination of $\nx$ and $\nf$ from $\{2, 3, 5, 7, 14, 20, 31, 50\}$, with $\nsub{times} = 500$ observations drawn for each of the $64$ combinations.
+
+#### Results {-}
 
 Figure \ref{fig:raw} shows that both the location and scale of the distributions change across panels.
-This is not desirable under $D_{null}$ as it would mean comparisons of $wpd$ values is not appropriate across different $nx$ and $nfacet$.
-Table \ref{tab:glm-tab} gives the summary of the generalized linear model to capture the relationship between $wpd_{raw}$ and number of comparisons.
-For example, the model considered for $m = 2$ is $wpd_{l,s} = 23.4 - 0.96*log(nx * nfacet)) + e$.
-The intercepts and slopes are similar independent of the starting distributions (see supplementary paper for details) and hence the coefficients are shown for the case when observations are drawn from a N(0, 1) distribution.
-Figure \ref{fig:dist-new-same-scale-link} shows the distribution of $wpd_{perm}$ and $wpd_{glm-scaled}$ in the same scale to show that a combination approach could be used for higher values of levels to alleviate the computational time of permutation approach.
-
-```{r raw, fig.cap = "Distribution of $wpd_{raw}$ is plotted across different $nx$ and $nfacet$ categories under $D_{null}$ through density and rug plots.
+This is not desirable under $\Dnull$ as it would mean comparisons of $\wpd$ values are not appropriate across different $\nx$ and $\nf$ values.
+Table \ref{tab:glm-tab} gives the summary of a Gamma generalized linear model to capture the relationship between $\wpdsub{raw}$ and the number of comparisons.
+The intercepts and slopes are similar, independent of the underlying distributions (see supplementary paper for details) and hence the coefficients are shown for the case when observations are drawn from a N(0,1) distribution.
+Figure \ref{fig:dist-new-same-scale-link} shows the distribution of $\wpdsub{perm}$ and $\wpdsub{glm-scaled}$ on the same scale to show that a combination approach could be used for higher values of $\nx$ and $\nf$ to alleviate the computational time of the permutation approach.
+
+(ref:raw) Distribution of $\wpdsub{raw}$ is plotted across different $\nx$ and $\nf$ categories under $\Dnull$ through density and rug plots.
 Both location (blue line) and scale (orange marks) of the distribution shifts for different panels.
-This is not desirable since under null design, the distribution  is not expected to capture any differences."}
+This is not desirable since under the null design, the distribution is not expected to capture any differences.
+
+```{r raw, fig.cap = "(ref:raw)", fig.pos="!htb"}
 ```
@@
 
-```{r dist-new-same-scale-link, fig.cap = "The distribution of $wpd_{perm}$ and $wpd_{glm-scaled}$ are overlaid to compare the location and scale across different $nx$ and $nfacet$.
-$wpd_{norm}$ takes the value of $wpd_{perm}$ for lower levels, and $wpd_{glm-scaled}$ for higher levels to to alleviate the problem of computational time in permutation approaches.
-This is possible as the distribution of the adjusted measure looks similar for both appraches for higher levels."}
-```
-
-### Choosing threshold
-
-\noindent _Simulation design_
-
-Observations are generated from a N(0,1) distribution for each combination of $nx$ and $nfacet$ from the following sets: $nx = \{3, 7, 14\}$ and $nfacet = \{2, 9, 10\}$.
-This would result in 9 panels, viz, $(3, 2), (3, 9), (3, 10), \dots, (14,9), (14, 10)$.
-Few experiments were conducted.
-In the first scenario, data for all panels are simulated using the null design  $D_{null}$.
-In other scenarios, data simulated from the panel $(14, 2)$ and $(3, 10)$ are considered under $D_{vary_{all}}$.
-Moreover, $\omega = \{0.5, 2, 5\}$ are considered to examine if the proposed test is able to capture subtle differences and non-subtle differences when we shift from the null design.
-
-<!-- In the last scenario, we consider the panel $(3, 2), (7,9), (14, 10)$ to be under $D_{null}$, the panels $(7, 2), (14, 9)$ to be under $D_{var_f}$.
-$(14, 2), (3, 10)$ under $D_{var_x}$ and the rest under $D_{var_{null}}$.
+(ref:dist-new-same-scale-link) The distribution of $\wpdsub{perm}$ and $\wpdsub{glm-scaled}$ are overlaid to compare the location and scale across different $\nx$ and $\nf$.
+$\wpdsub{norm}$ takes the value of $\wpdsub{perm}$ for lower levels, and $\wpdsub{glm-scaled}$ for higher levels to to alleviate the problem of computational time in permutation approaches.
+This is possible as the distribution of the adjusted measure looks similar for both approaches for higher levels.
+
+```{r dist-new-same-scale-link, fig.cap = "(ref:dist-new-same-scale-link)"}
+```
+
+These results justify our use of the permutation approach when $\nx<5$ and $\nf<5$, and the use of the GLM otherwise.
+<!--
+## Choosing the threshold
+
+### Simulation design {-}
+
+Observations were generated from a N(0,1) distribution for each combination of $\nx$ and $\nf$ from the following sets: $nx \in \{3, 7, 14\}$ and $\nf \in \{2, 9, 10\}$, giving nine possible panels/combinations.
+In the first scenario, data for all panels were simulated using the null design $\Dnull$.
+In other scenarios, data simulated from the panel $(14, 2)$ and $(3, 10)$ used $\Dfx$ where the means of the distributions increase by $\omega$ across both x-axis and facets, where $\omega \in \{0.5, 2, 5\}$.
+This allows us to examine if the proposed test is able to capture subtle differences and obvious differences when we shift from the null design.
+
+In the last scenario, we consider the panel $(3, 2), (7,9), (14, 10)$ to be under $\Dnull$, the panels $(7, 2), (14, 9)$ to be under $\Df$.
+$(14, 2), (3, 10)$ under $D_x$ and the rest under $\Dvar{{null}}$.
 This is done to check if the consequent ranking procedure leads to designs like $D_{vary_{all}}$ to be chosen first followed by $D_{vary_{all}}$.
@@
 We generate $1000$ repetitions of this experiment with different seeds.
--->
-
-\noindent _Results_
-
-For the first scenario, size of the test is obtained as $0.1$ with $wpd_{threshold99}$ as the threshold.
-This implies that the proportion of times a panel is rejected when it is under $D_{null}$ is $0.1$.
-The level of significance for each test is 1% (as a result of choosing $99^{th}$ percentile as the threshold) and we have $9$ multiple tests.
-Hence, it is reasonable
-that the level of significance of the composite tests would be larger than the individual tests.
-We also compute the proportion of times a panel is rejected when it actually belongs to a non-null design.
-These proportions constitute to be the estimated size and power of the test.
-Power can depend on many things like sample size, number of designs that deviate from the null and extent to which from the null.
-It is found that as we increase from low to high changes from the null distribution, the power increased.
-<!---The power also increases with increased sample size, as is expected.---> The results and graphics are included in details in the Supplementary paper.
+
+### Results {-}
+
+For the first scenario, we set $\wpdsub{threshold99}$ so that the probability of rejecting each panel under $\Dnull$ is $0.01$.
+We have 9 tests, so the size of the tests overall is $1-0.99^9 \approx 0.09$.
+We also compute the proportion of times a panel is rejected when it belongs to a non-null design to obtain the power of the test.
+The detailed results and graphics are included in the Supplementary paper.
 
@@
  geom_histogram(data = all_data, aes(x = wpd)) +
-  geom_vline(data = data_summary, aes(xintercept = p_value),colour = "red") +
+ geom_vline(data = data_summary, aes(xintercept = p_value),color = "red") +
  facet_grid(x_levels~facet_levels)
@@
 ```
-
-### Environment
-
-Simulation studies were carried out to study the behavior of $wpd$, build the normalization method as well as compare and evaluate different normalization approaches.
-R version 4.0.1 (2020-06-06) is used with the platform: x86_64-apple-darwin17.0 (64-bit) running under: macOS Mojave 10.14.6 and MonaRCH, which is a next-generation High Power Computing (HPC) Cluster, addressing the needs of the Monash HPC community.
+ -->
 
@@
 
-The smart meter data set for eight households in Melbourne has been utilized to see the use of $wpd$ proposed in the paper.
-The data has been cleaned to be a `tsibble` (@wang2020tsibble) containing half-hourly electricity consumption from Jul-2019 to Dec-2019 for each of the households, which is procured by them by downloading their data from the energy supplier/retailer.
-No behavioral pattern is likely to be discerned from the line graph of energy usage over the entire period, since the plot will have too many measurements squeezed in a linear representation.
-When we zoom into the linear representation of this series in Figure \ref{fig:dotplot-8} (b) for September, some patterns are visible in terms of peaks and troughs, but we do not know if they are regular or what is their period.
-Electricity demand, in general, has a daily and weekly periodic pattern.
-However, it is not apparent from this view if all of these households have those patterns and in case they have if they are significant enough.
-Also, it is not clear if any other periodic patterns are present in any household which might have been hidden with this view.
-We start the analysis by choosing few harmonies, ranking them for each of these households, compare households to get more insights into what these rankings imply.
-Furthermore, the ranking and selection of significant harmonies is validated by analyzing the distribution of energy usage across significant harmonies.
+The smart meter data set for eight households in Melbourne were procured by downloading the data from the energy supplier/retailer.
+The data has been cleaned to form a `tsibble` [@wang2020tsibble] containing half-hourly electricity consumption from July to December 2019 for each of the households, No behavioral pattern is likely to be discerned from the time plot of energy usage over the entire period, since the plot will have too many observations squeezed in a linear representation.
+When we zoom into the September 2019 data in Figure \ref{fig:dotplot-8}(b), some patterns are visible in terms of peaks and troughs, but we do not know if they are regular or what is their period.
+
+(ref:dotplot-8) An ensemble plot with a heatmap (a), line plot (b), parallel coordinate plot (c) to demonstrate energy behavior of the households in different ways.
+Panel (b) shows the raw demand series for September to highlight the repetitive patterns of energy demand.
+Panel (a) shows $\wpd$ values across harmonies where a darker color indicates a higher ranking harmony.
+A significant harmony is shown with an asterisk.
+For example, ids 7 and 8 have significant patterns across (*hod*, *dow*) and (*dow*, *hod*).
+Panel (c) is useful for comparing households across harmonies.
+For example, for the harmony (*dow*-*hod*), ids 1 and 7 have the least and highest $\wpd$ respectively.
+
+```{r dotplot-8, fig.cap = "(ref:dotplot-8)", fig.height = 7}
+```
+
+Electricity demand, in general, has daily, weekly and annual seasonal patterns.
+However, it is not apparent from this view if all households have those patterns, or how strong they are in each case.
+It is also not clear from this view if any other periodic patterns are present in any household.
+We start the analysis by choosing a few harmonies, and ranking them for each of these households.
+The ranking and selection of significant harmonies is validated by analyzing the distribution of energy usage across significant harmonies.
 
@@
 
-_Choosing cyclic granularities of interest and removing clashes_
-
-Let $v_{i, t}$ denote the electricity demand for $i^{th}$ household for time period $t$.
-The series $v_{i, t}$ is the linear granularity corresponding to half-hour since the interval of the tsibble is 30 minutes.
+#### Choosing cyclic granularities of interest and removing clashes {-}
+
+Let $v_{i, t}$ denote the electricity demand for the $i^{th}$ household in time period $t$.
+The series $\{v_{i, 1},\dots,v_{i,T}\}$ is the linear granularity corresponding to half-hour since the interval of the tsibble is 30 minutes.
 We consider coarser linear granularities like hour, day, week and month from the commonly used Gregorian calendar.
-Considering $4$ linear granularities hour, day, week, month, the number of cyclic granularities is $N_C = (4 * 3/2) = 6$.
-We obtain cyclic granularities namely "hour_day", "hour_week", "hour_month", "day_ week", "day_month" and "week_month", read as "hour of the day", etc.
+From the four linear granularities of hour, day, week, and month, we obtain $N_C = 4\times 3/2 = 6$ cyclic granularities:  "hour_day", "hour_week", "hour_month", "day_ week", "day_month" and "week_month".
 Further, we add cyclic granularity day-type ("wknd wday") to capture weekend and weekday behavior.
-Thus, $7$ cyclic granularities are considered to be of interest.
-The set consisting of pairs of cyclic granularities ($C_{N_C}$) will have $7_{P_2}=42$ elements which could be analyzed for detecting possible periodicities.
+Thus, seven cyclic granularities are considered to be of interest.
+The pairs of cyclic granularities ($C_{N_C}$) will have $7\times 6=42$ elements.
 The set of possible harmonies $H_{N_C}$ from $C_{N_C}$ are chosen by removing clashes using procedures described in [@Gupta2021-hd].
@@
 
+```{r rank-household-mclapply-8}
+```
+
+```{r elec_select_harmony-8}
+```
+
+```{r tab-rank-8, dependson="dotplot-8"}
+```
+
  <!-- The paper also shows how the number of harmonies could be identified among by removing clashes.
@@
 
-_Selecting and Ranking harmonies for all households_
-
-$wpd_{i}$ is computed on $v_{i, t}$ for all harmony pairs $\in H_{N_C}$ and for each households $i \in \{1, 2, \dots, 8\}$.
+#### Selecting and Ranking harmonies for all households {-}
+
+$\wpd_{i}$ is computed on $v_{i, t}$ for all harmony pairs $\in H_{N_C}$ and for each household $i \in \{1, 2, \dots, 8\}$.
 The harmony pairs are then arranged in descending order and highlighted with `***`, `**` and `*` corresponding to the $99^{th}$, $95^{th}$ and $90^{th}$ percentile threshold.
@@
 The rankings are different for different households, which is a reflection of their varied behaviors.
-Most importantly, there are at most $3$ harmonies that are significant for any household.
-This is a huge reduction in the number of potential harmonies to explore closely, starting from $42$.
-
-_Detecting patterns not apparent from linear display_
-
-\noindent Figure \ref{fig:dotplot-8} helps to compare households through the heatmap (a) across harmony pairs with the cyclic granularity mapped to x-axis and facet being plotted on the x-axis and y-axis of the heatmap.
-Here $dom$, $dow$, $wdwnd$ are abbreviations for day-of-month, day-of-week, weekday/weekend and so on.
-The colors represent the value of $wpd$.
-Darker cells correspond to more significant harmony pairs.
-Also, the ones with `*` corresponds to the ones above $wpd_{threshold95}$.
-Few observations that emphasizes patterns not discernible through (b) includes - (1) id 7 and 8 have the same significant harmonies despite having very different total energy usage, (2) id 6 and 7 differ in the sense that for id 6, the difference in patterns only during weekday/weekends, whereas for id 7, all or few other days of the week are also important.
-This might be due to their flexible work routines or different day-off, (3) there are no significant periodic patterns for id 5 when we fix the threshold to $wpd_{threshold95}$.
-Note that the $wpd$ values are computed over the entire range, but the linear display in (b) is zoomed only for September, with the major and minor x-axis corresponding to weeks and days respectively.
+Most importantly, there are at most three harmonies that are significant for any household.
+This is a huge reduction in the number of potential harmonies to explore.
+
+#### Detecting patterns not apparent from linear display {-}
+
+Figure \ref{fig:dotplot-8} helps to compare households through the heatmap (a) across harmony pairs.
+Here *dom*, *dow*, *wdwnd* are abbreviations for day-of-month, day-of-week, weekday/weekend and so on.
+The colors represent the value of $\wpd$.
+Darker cells correspond to higher values of $\wpd$.
+Those with `*` correspond to $\wpd$ values above $\wpdsub{threshold95}$.
+
+We can now see some patterns that were not discernible in Panel (b), including:
+
+  1. id 7 and 8 have the same significant harmonies despite having very different total energy usage.
+  2.
+id 6 and 7 differ in the sense that for id 6, the difference in patterns only during weekday/weekends, whereas for id 7, all or few other days of the week are also important.
+This might be due to their flexible work routines or different day-off.
+  3. There are no significant periodic patterns for id 5 when we fix the threshold to $\wpdsub{threshold95}$.
+
+Note that the $\wpd$ values are computed over the entire range, but the linear display in (b) is only for September, with the major and minor x-axis corresponding to weeks and days respectively.
 
@@
 
-```{r rank-household-mclapply-8}
-```
-
-```{r elec_select_harmony-8}
-```
-
-```{r dotplot-8, fig.cap = "An ensemble plot with a heatmap (a), line plot (b), parallel coordinate plot (c) to demonstrate energy behavior of the households in different ways.
-(b) is the zoomed-in raw demand series for September to highlight the repetitive patterns of energy demand.
-(a) shows $wpd$ values across harmonies with the x variable of harmony placed across x-axis and facet variable placed across y-axis.
-The darker the colour in (a), the higher the harmony is ranked.
-Visualizing harmonies through (a) allows to view the significant cyclic granularities along the x-axis, facet or both for each household.
-For eg, ids 7 and 8 have significant patterns across (hod, dow) and (dow, hod), which was not apparent through (a).
-(c) is useful for comparing households across harmonies, for eg, for the harmony (dow-hod), ids 1 and 7 have the least and highest wpd respectively.", fig.height = 7}
-```
-
-```{r tab-rank-8}
-```
-
-_Comparing households and validating rank of harmonies_
-
-According to Figure \ref{fig:dotplot-8}(c), for the harmony pair (dow-hod), household id 7 has the greatest value of $wpd$, while id 1 has the least.
-Also, from table \ref{tab:tab-rank-8} it could be seen that the harmony pair $(dow, hod)$ is important for id 7, however it has been labeled as an inconsequential pair for id 1.
-The distribution of energy demand for both of these households, with $dow$ as the facet and $hod$ on the x-axis, may help explain the choice.
-Figure \ref{fig:gravitas-plot-8} demonstrates that for id 7, the median (black) and quartile deviation (orange) of energy consumption fluctuates throughout for most hours of the day and day of the week, while for id 1, daily patterns are consistent within weekdays and weekends.
-As a result, for id 1, it is more appropriate to examine the distributional difference solely across (dow, wdwnd), which has been rated higher in Table \ref{tab:tab-rank-8}.
+#### Comparing households and validating rank of harmonies {-}
+
+According to Figure \ref{fig:dotplot-8}(c), for the harmony pair (*dow*-*hod*), household id 7 has the greatest value of $\wpd$, while id 1 has the least.
+From Table \ref{tab:tab-rank-8} it can be seen that the harmony pair (*dow*, *hod*) is important for id 7, however it has been labeled as an inconsequential pair for id 1.
+The distribution of energy demand for both of these households, with *dow* as the facet and *hod* on the x-axis, may help explain the choice.
+Figure \ref{fig:gravitas-plot-8} demonstrates that for id 7, the median (black) and quartile deviation (orange) of energy consumption fluctuates for most hours of the day and days of the week, while for id 1, daily patterns are more consistent within weekdays and weekends.
+As a result, for id 1, it is more appropriate to examine the distributional difference solely across (*dow*, *wdwnd*), which has been rated higher in Table \ref{tab:tab-rank-8}.
 
@@
 However, the difference between morning and evening peaks are higher for weekends.
-Since $wpd$ is designed to put more weightage on within-facet differences for $\lambda>0.5$, it makes sense that the pair (hod, wdwnd) has been ranked higher than (wdwnd, hod).
+Since $\wpd$ is designed to put more weightage on within-facet differences for $\lambda>0.5$, it makes sense that the pair (hod, wdwnd) has been ranked higher than (wdwnd, hod).
 Panel c) shows the distribution of energy demand with weekday/weekend as the x-axis and week-of-month as the facets.
@@
 
-```{r gravitas-plot-8, message = FALSE, warning=FALSE,fig.cap="Comparing distribution of energy demand shown for household id 1 (a) and 7 (b) across hod in x-axis and dow in facets through quantile area plots.
-The value of $wpd$ in Table 3 suggests that the harmony pair (dow, hod) is significant for household id 7, but not for 1.
+(ref:gravitas-plot-8) Comparing distribution of energy demand shown for household id 1 (a) and 7 (b) across *hod* in x-axis and *dow* in facets through quantile area plots.
+The value of $\wpd$ in Table 3 suggests that the harmony pair (*dow*, *hod*) is significant for household id 7, but not for id 1.
 This implies that distributional differences are captured more by this harmony for id 7, which is apparent from this display with more fluctuations across median and 75th percentile for different hours of the day and day of week.
 For id 1, patterns look similar within weekdays and weekends.
-Here, the median is represented by the black line, the orange area corresponds to quartile deviation and the green area corresponds to area between $10^{th}$ and $90^{th}$ quantile."}
-
-```
-
-<!-- From Figure \ref{fig:gravitas-plot-8-id5}, it could further be observed that id5 has only one significant harmony $(hod, dow)$.
+Here, the median is represented by the black line, the orange area corresponds to quartile deviation and the green area corresponds to area between $10^{th}$ and $90^{th}$ quantile.
+
+```{r gravitas-plot-8, message = FALSE, warning=FALSE,fig.cap="(ref:gravitas-plot-8)", fig.pos="!htb"}
+
+```
+
+<!-- From Figure \ref{fig:gravitas-plot-8-id5}, it could further be observed that id5 has only one significant harmony (*hod*, *dow*).
 Apparently, $(hod, wnd/wday)$ which is an important harmony for most households is not important for this one.
@@
 With temporal data available at finer scales, exploring time series has become overwhelming with so many possible granularities to explore.
-A common solution is to aggregate and look at the patterns across usual granularities like hour-of-day or day-of-week, but there is no way to know the “interesting” granularities a priori.
+A common solution is to aggregate and look at the patterns across the usual granularities such as hour-of-day or day-of-week, but there is no way to know the “interesting” granularities a priori.
 A huge number of displays need to be analyzed or we might end up missing informative granularities.
@@
 
-The Australian authors thank the ARC Centre of Excellence for Mathematical and Statistical Frontiers [(ACEMS)](https://acems.org.au/home) for supporting this research.
+The authors thank the ARC Centre of Excellence for Mathematical and Statistical Frontiers [(ACEMS)](https://acems.org.au/home) for supporting this research.
 Sayani Gupta was partially funded by [Data61 CSIRO](https://data61.csiro.au/) during her PhD.
-The Github repository,  [github.com/Sayani07/paper-hakear](https://github.com/Sayani07/paper-gravitas), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
+The Github repository, [github.com/Sayani07/paper-hakear](https://github.com/Sayani07/paper-hakear), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
 This article was created with R [@R-language], `knitr` [@knitr2015], [@R-knitr] and `rmarkdown` [@rmarkdown2018], [@R-rmarkdown].
-Graphics are produced with @Wickham2009pk.
+Graphics are produced with `ggplot2` [@Wickham2009pk].
 
@@
 \noindent
-**R-package:** The open-source R [@R-language] package `hakear` is available on Github (https://github.com/Sayani07/hakear) to implement ideas presented in this paper.
+**R-package:** The open-source R package `hakear` is available on Github (https://github.com/Sayani07/hakear) to implement ideas presented in this paper.

### 2021-11-10 scripts/hakear.R

@@ -1,6 +1,7 @@
-## ---- load
+## ---- load 
 library(knitr)
 library(tidyverse)
+library(broom)
 library(lubridate)
 library(lvplot)
 #library(ggridges)
@@ -17,44 +18,45 @@ library(ggplot2)
 library(patchwork)
 library(scales)
 library(GGally)
-library(viridis)
+#library(viridis)
 #remotes::install_github("Sayani07/hakear", force = TRUE)
 library(hakear)
-library(latex2exp)
-library(broom)
+
+# OkabeIto color blind friendly scale
+okabeito <- c("#D55E00", "#0072B2","#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442", "#333333")
 
 #library(tidyquant)
 
 ## ---- calendar-elec
-elec <- read_rds(here("data/hakear/elec.rds")) %>%
+elec <- read_rds(here("data/hakear/elec.rds")) %>% 
   filter(date >= ymd("20180101"), date < ymd("20180701"))
 rdbl <- c("Weekday" = "#d7191c", "Weekend" = "#2c7bb6")
 
-elec <- elec %>%
+elec <- elec %>% 
   mutate(
     wday = wday(date, label = TRUE, week_start = 1),
     weekday = if_else(wday %in% c("Sat", "Sun"), "Weekend", "Weekday")
   )
-# p_cal_elec <- elec %>%
-#   filter(id %in% c(2, 4)) %>%
-#   frame_calendar(x = time, y = kwh, date = date, nrow = 1) %>%
+# p_cal_elec <- elec %>% 
+#   filter(id %in% c(2, 4)) %>% 
+#   frame_calendar(x = time, y = kwh, date = date, nrow = 1) %>% 
 #   ggplot(aes(x = .time, y = .kwh, group = date)) +
 #   geom_line(aes(colour = as.factor(id)), size = 0.5) +
 #   scale_colour_brewer(name = "", palette = "Dark2", direction = 1) +
 #   facet_grid(id ~ ., labeller = label_both) +
 #   theme(legend.position = "bottom")
 # prettify(p_cal_elec, size = 2.5, label.padding = unit(0.1, "lines"))
-#
+# 
 # simtable_new <- simtable %>% filter(nfacet==50 & nx == 20 | nfacet==50 & nx == 31 | nfacet==50 & nx == 50 | nfacet==31 & nx == 50)
 
 
[…]

### 2021-11-10 Rmd/05-conclusion.Rmd
Rob's coorection all - except for "assumed to be independent"

@@
 The methodology also ranks the displays in order of priority for capturing the most variation.
-Both \@ref(ch:gravitas) and \@ref(ch:hakear) are used for studying patterns in individual time series or comparing a few time series together.
+The ideas in Chpaters \@ref(ch:gravitas) and \@ref(ch:hakear) can be used for studying patterns in individual time series or comparing a few time series together.
 This is extended in Chapter \@ref(ch:gracsr) to allow for the exploration of distributions for multiple time series at the same time using unsupervised clustering.
@@
 This thesis focuses on integrating research approaches into open source R packages for reproducibility and ease of use in other applications.
-So a significant amount of work has been devoted to the development of R packages  **gravitas**, **hakear**, and **gracsr**, each of which correspond to a chapter presented in this thesis.
+So a significant amount of work has been devoted to the development of R packages  `gravitas`, `hakear`, and `gracsr`, each of which corresponds to a chapter presented in this thesis.
 
@@
 
-I want to combine 'hakear' and 'gravitas' into a single R package for systematically exploring distributions of a smaller number of time series, whilst 'gracsr' would offer the clustering framework for systematically exploring a large number of time series.
+I want to combine `hakear` and `gravitas` into a single R package for systematically exploring distributions of a smaller number of time series, whilst `gracsr` would offer the clustering framework for systematically exploring a large number of time series.
 I intend to run it via https://ropensci.org/software-review/ to further refine them for more exposure and efficiency.
@@
 It is crucial to determine whether or not any detected clusters or patterns are statistically significant.
-In Chapter \@ref(ch:gracsr), we have used the permutation approach to detect if patterns are authentic or not across one or a pair of granularities.
+In Chapter \@ref(ch:gracsr), we have used a permutation approach to detect if patterns are authentic or not across one or a pair of granularities.
 A similar method for finding significance across clusters is not appropriate if the observations are assumed to be independent.

### 2021-11-10 Rmd/03-hakear.Rmd
updated hakear chapter

@@
 
-Cyclic temporal granularities [@Bettini1998-ed;@Gupta2021-hd] are <!--theoretical--> temporal deconstructions that define cyclic repetitions in time, e.g.
+Cyclic temporal granularities [@Bettini1998-ed;@Gupta2021-gravitas] are <!--theoretical--> temporal deconstructions that define cyclic repetitions in time, e.g.
 hour-of-day, day-of-month, or regularly scheduled public holidays.
@@
 
-In this paper, we provide a solution to narrowing down the search from $N_C(N_C-1)$ conditional distribution plots by introducing a new distance measure that can be used to detect significant distributional differences across cyclic granularities.<!-- harmonies and why it is not enough --> This work is a natural extension of our previous work [@Gupta2021-hd] which narrows down the search from $N_C(N_C-1)$ plots by identifying pairs of granularities that can be meaningfully examined together (a "harmony"), or when they cannot (a "clash").
+In this paper, we provide a solution to narrowing down the search from $N_C(N_C-1)$ conditional distribution plots by introducing a new distance measure that can be used to detect significant distributional differences across cyclic granularities.<!-- harmonies and why it is not enough --> This work is a natural extension of our previous work [@Gupta2021-gravitas] which narrows down the search from $N_C(N_C-1)$ plots by identifying pairs of granularities that can be meaningfully examined together (a "harmony"), or when they cannot (a "clash").
 However, even after excluding clashes, the list of harmonies left may be too large for exhaustive exploration.
@@
 The pairs of cyclic granularities ($C_{N_C}$) will have $7\times 6=42$ elements.
-The set of possible harmonies $H_{N_C}$ from $C_{N_C}$ are chosen by removing clashes using procedures described in [@Gupta2021-hd].
+The set of possible harmonies $H_{N_C}$ from $C_{N_C}$ are chosen by removing clashes using procedures described in [@Gupta2021-gravitas].
 Table \ref{tab:tab-rank-8} shows $14$ harmony pairs that belong to $H_{N_C}$.
@@
 The Github repository, [github.com/Sayani07/paper-hakear](https://github.com/Sayani07/paper-hakear), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
-This article was created with R [@R-language], `knitr` [@knitr2015], [@R-knitr] and `rmarkdown` [@rmarkdown2018], [@R-rmarkdown].
+This article was created with R [@R-language], `knitr` [@knitr], [@R-knitr] and `rmarkdown` [@rmarkdown], [@R-rmarkdown].
 Graphics are produced with `ggplot2` [@Wickham2009pk].

### 2021-11-10 Rmd/05-conclusion.Rmd
spell correction

@@
 The methodology also ranks the displays in order of priority for capturing the most variation.
-The ideas in Chpaters \@ref(ch:gravitas) and \@ref(ch:hakear) can be used for studying patterns in individual time series or comparing a few time series together.
+The ideas in Chapters \@ref(ch:gravitas) and \@ref(ch:hakear) can be used for studying patterns in individual time series or comparing a few time series together.
 This is extended in Chapter \@ref(ch:gracsr) to allow for the exploration of distributions for multiple time series at the same time using unsupervised clustering.

### 2021-11-11 Rmd/05-conclusion.Rmd
structural changes proposed in the meeting today

@@
-# Conclusion and future plans {#ch:conclusion}
+# Conclusion {#ch:conclusion}
+
+This chapter summarizes the thesis content, covering original contributions, software development, and future ideas.
+
+## Original contributions
 
@@
 
-## Future work
+## Future ideas
 
-### Putting all functionalities on CRAN
+<!-- ### Putting all functionalities on CRAN -->
 
-I want to combine `hakear` and `gravitas` into a single R package for systematically exploring distributions of a smaller number of time series, whilst `gracsr` would offer the clustering framework for systematically exploring a large number of time series.
+<!-- I want to combine `hakear` and `gravitas` into a single R package for systematically exploring distributions of a smaller number of time series, whilst `gracsr` would offer the clustering framework for systematically exploring a large number of time series.
 I intend to run it via https://ropensci.org/software-review/ to further refine them for more exposure and efficiency.
+-->
 
@@
 
-### Comparing our clustering method with other benchmark methods
+<!-- ### Comparing our clustering method with other benchmark methods -->
 
-Testing needs to be carried out with non-hierarchy based clustering methods to see whether the proposed distances perform better with other algorithms.
+<!-- Testing needs to be carried out with non-hierarchy based clustering methods to see whether the proposed distances perform better with other algorithms.
 It would also be helpful to know how much more value we obtain from considering distributions rather than aggregating time series data.
 In the case of smart meters, this could be accomplished by comparing the results while validating them against external data such as weather conditions, socioeconomic or other demographic factors of those households.
+-->
 
@@
 
-### Check generalizability of our methods
+### Extend methods to other fields
 

### 2021-11-12 Rmd/02-gravitas.Rmd
trying to make the gitbook work
making spelling consistent with USenglish and make fig.height =6 for compressed figure

@@
 
-```{r allFig, fig.pos="!p", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap= "(ref:allFig)", fig.show = 'hold'}
+```{r allFig, fig.pos="!p", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap= "(ref:allFig)", fig.show = 'hold', fig.height=6}
 ```
@@
 
-## Supplementary Materials
+## Supplementary Materials {- #supplement-gravitas}
 

### 2021-11-12 Rmd/04-gracsr.Rmd
sometimes I do rock, Don't I? successful thesis compilation
trying to make the gitbook work
making spelling consistent with USenglish and make fig.height =6 for compressed figure

@@
 Cyclic granularities like hour-of-the-day, work-day/weekend, month-of-the-year and so on, are useful for finding repeated patterns in the data.
-Looking at probability distributions across cyclic granularities serves two purposes: (a) "Probability distributions" characterise the inherent temporal data structure of these large unequal-length time series and are robust to missing or noisy data.
+Looking at probability distributions across cyclic granularities serves two purposes: (a) "Probability distributions" characterize the inherent temporal data structure of these large unequal-length time series and are robust to missing or noisy data.
 (b) Using probability distributions over "cyclic granularities" ensures small pockets of similar "repeated" behaviors.
@@
 Differences in Weekend/Weekday between groups are not discernible, implying that it may not be a relevant variable in distinguishing various clusters unless maybe conditioned by `moy` or `hod`.
-It may be interesting to compare these two plots to verify if the summarized distributions across groups correctly characterised the groupings.
+It may be interesting to compare these two plots to verify if the summarized distributions across groups correctly characterized the groupings.
 If it has, then the majority of the group's members should share a similar profile.
@@
 We chose the optimal number of clusters using (@Hennig2014-ah) as $3$.
-A parallel coordinate plot with the three significant cyclic granularities is used to characterise the groups here.
+A parallel coordinate plot with the three significant cyclic granularities is used to characterize the groups here.
 The variables are sorted according to their separation across classes (rather than their overall variation between classes).
@@
 Depending on the goal of the clustering, these distance metrics, when fed into a hierarchical clustering algorithm using Ward's linkage, yield meaningful clusters.
-Probability distributions provide an intuitive method to characterise noisy, patchy, long, and unequal-length time series data.
-Distributions over cyclic granularities help to characterise the formed clusters in terms of their repeating behavior over these cyclic granularities.
+Probability distributions provide an intuitive method to characterize noisy, patchy, long, and unequal-length time series data.
+Distributions over cyclic granularities help to characterize the formed clusters in terms of their repeating behavior over these cyclic granularities.
 Furthermore, unlike earlier efforts that group customers based on behavior across only one cyclic granularity (such as hour-of-day), our method is more comprehensive in detecting clusters with repeated patterns at all relevant granularities.

### 2021-11-12 Rmd/05-conclusion.Rmd
changing gravitas conclusion

@@
 This technique is more robust to noisy, patchy, and uneven length time series because it makes use of probability distributions.
+
+Chapter \@ref(ch:gravitas)
+
+This thesis presents a framework for visualizing and analyzing distributions of large temporal data by leveraging time characteristics.
+The building blocks of this framework are presented in each chapter.
+Exploratory time series analysis involves many iterations of finding and summarizing temporal dependencies.
+It is typical to divide time into years, months, weeks, days and so on to carry out inferences both across finer and coarser scales.
+The formalization of these temporal deconstructions (granularities) in literature are limited to linear time granularities like hours, days, weeks, and months which respect the linear progression of time and are non-repeating.
+Cyclic granularities which are repeating in nature are useful for finding patterns in temporal data.
+They can be circular, quasi-circular, or aperiodic.
+Examples of circular granularities are hour of the day and day of the week; an example of a quasi-circular granularity is day of the month; examples of aperiodic granularities are public holidays and school holidays.
+Time deconstructions can also be based on time hierarchy.
+Thus, single-order-up granularities like second of minute or multiple-order-up granularities like second of hour can be conceived.
+The definitions and rules defined in the literature for linear granularities are inadequate for describing different categories of cyclic granularities.
+Chapter \@ref(ch:gravitas) provides a formal characterization of cyclic granularities and provides tools to classify and compute possible cyclic granularities from an ordered (usually temporal) index.
+It also facilitates manipulation of single- and multiple-order-up time granularities through cyclic calendar algebra.
+
+Visualizing probability distributions conditional on one or more cyclic granularities is an indispensable tool for exploration.
+However, firstly, there could be too many cyclic granularities to look at manually for comprehensive exploration and secondly, not all pairs of granularities can be effectively explored together.
+This chapter helps to list down the possible cyclic granularities that can be explored given a considered hierarchy and provides a recommendation whether a pair of granularities can be meaningfully plotted or analyzed together based on the synergy and levels of the cyclic granularities.
 

### 2021-11-12 Rmd/03-hakear.Rmd
adding original contributions

@@
 If the order of categories is ignored, the resulting $\wpd$ values are $97.8$ and $161$ respectively, because differences between any hours of the day tend to be larger than differences only between consecutive hours.
-Similarly, \ref{fig:onegran-new}(a) and (b) have $\wpd$ values of $110.79$ and $125.82$ respectively.
+Similarly, Figures \ref{fig:id2-new2}(a) and (b) have $\wpd$ values of $110.79$ and $125.82$ respectively.
 The ranking implies that the distributional differences are more prominent for the second household, as is also seen from the bigger fluctuations in the $90^{th}$ percentile than for the first household.
@@
 The smart meter data set for eight households in Melbourne were procured by downloading the data from the energy supplier/retailer.
-The data has been cleaned to form a `tsibble` [@wang2020tsibble] containing half-hourly electricity consumption from July to December 2019 for each of the households, No behavioral pattern is likely to be discerned from the time plot of energy usage over the entire period, since the plot will have too many observations squeezed in a linear representation.
+The data has been cleaned to form a `tsibble` [@wang2020tsibble] containing half-hourly electricity consumption from July to December 2019 for each of the households.
+No behavioral pattern is likely to be discerned from the time plot of energy usage over the entire period, since the plot will have too many observations squeezed in a linear representation.
 When we zoom into the September 2019 data in Figure \ref{fig:dotplot-8}(b), some patterns are visible in terms of peaks and troughs, but we do not know if they are regular or what is their period.

### 2021-11-12 Rmd/05-conclusion.Rmd
adding original contributions

--- 2021-11-11 Rmd/05-conclusion.Rmd
@@
 
+This thesis presents a framework for visualizing and analyzing distributions of large temporal data by leveraging time characteristics.
+The building blocks of this framework are presented in each main chapter.
 This chapter summarizes the thesis content, covering original contributions, software development, and future ideas.
@@
 
-In this thesis, I present a systematic approach for visualizing and analyzing distributions of large temporal data by leveraging time characteristics.
+<!-- In this thesis, I present a systematic approach for visualizing and analyzing distributions of large temporal data by leveraging time characteristics.
 The building blocks of this framework are presented in each chapter.
@@
 This technique is more robust to noisy, patchy, and uneven length time series because it makes use of probability distributions.
+-->
+
+Exploratory time series analysis entails numerous iterations of identifying and summarizing temporal dependencies.
+It is common practice to divide time into years, months, weeks, days, and so on in order to make inferences at both finer and coarser scales.
+In the literature, the formalization of these temporal deconstructions (granularities) is limited to linear time granularities such as hours, days, weeks, and months that respect the linear progression of time and are non-repeating.
+Cyclic granularities that are repeating in nature are useful for finding patterns in temporal data.
+They can be circular, quasi-circular, or aperiodic in nature.
+Hour of the day and day of the week are examples of circular granularities; the day of the month is an example of a quasi-circular granularity; and public holidays and school holidays are examples of aperiodic granularities.
+Additionally, time deconstructions can be based on a time hierarchy.
+Thus, single-order-up granularities such as second of minute or multiple-order-up granularities such as second of hour can be envisioned.
+The definitions and rules defined in the literature for linear granularities are insufficient for describing various types of cyclic granularities.
+Chapter \@ref(ch:gravitas) provides a formal characterization of cyclic granularities as well as tools for classifying and computing potential cyclic granularities from an ordered (usually temporal) index.
+It also allows for the manipulation of single- and multiple-order-up time granularities via cyclic calendar algebra.
+
+Visualizing probability distributions conditional on one or more cyclic granularities is a powerful exploration tool.
+However, there may be too many cyclic granularities to look at manually for comprehensive exploration, and not all pairs of granularities can be effectively explored together.
+Chapter \@ref(ch:gravitas) also provides a recommendation on whether a pair of granularities can be meaningfully plotted or analyzed together (a "harmony") or when they cannot (a "clash" or "near-clash").
+
+Cyclic granularities could be used to create a wide range of displays.
+And, when there are numerous granularities to choose from, deciding which one to display can be difficult.
+However, only a few of them may be useful in revealing major patterns.
+In Chapter \@ref(ch:hakear), the search for informative granularities by selecting "significant" granularities.
+A cyclic granularity is referred to as "significant" if there is a significant distributional difference of the measured variable between different categories.
+Chapter \@ref(ch:hakear) defines a distance measure to quantify the distributional differences between displayed distributions.
+A higher value of the distance measure for a cyclic granularity or harmony implies that they could be interesting for further investigation, whereas a low value indicates that nothing noteworthy is unfolding.
+A threshold and, consequently, a selection criterion are chosen using a permutation test such that cyclic granularities with significant values of the distance measure are selected.
+In addition, the distance metric has been appropriately adjusted, allowing it to be compared not only across cyclic granularities with different numbers of categories but also across a set of time series.
+As a result, it can also be used to rank the displays according to their ability to capture the greatest amount of variation across one or multiple time series.
+
+The ideas in Chapters \@ref(ch:gravitas) and \@ref(ch:hakear) can be used for studying patterns in individual time series or comparing a few time series together.
+This is extended in Chapter \@ref(ch:gracsr) to allow for the exploration of distributions for multiple time series at the same time using unsupervised clustering.
+In the time series clustering literature, probability distributions across cyclic granularities have not been considered in determining similarity.
+However, such a similarity measure can be useful for characterizing the inherent temporal data structure of long, unequal-length time series in a way that is resistant to missing or noisy data while allowing for the detection of similar repeated patterns.
+Chapter \@ref(ch:gracsr) proposes two approaches for calculating distances between time series based on probability distributions across cyclic granularities.
+The first approach considers two time series to be similar if the distributions of each category of one or more cyclic granularities are similar.
+The second approach considers two time series to be similar if they have a similar significance of patterns across different granularities.
+A similar significance does not imply a similar pattern, which is where this technique varies from the former.
+When the distances from these approaches are fed into a hierarchical clustering algorithm, they yield small groups of time series with similar distributions or significance over multiple granularities.
+Our method is capable of producing useful clusters for both approaches, as demonstrated by testing on a range of validation data designs and a sample of residential smart meter consumers.
 

### 2021-11-12 Rmd/05-conclusion.Rmd
minor changes in original contribution

@@
 The definitions and rules defined in the literature for linear granularities are insufficient for describing various types of cyclic granularities.
-Chapter \@ref(ch:gravitas) provides a formal characterization of cyclic granularities as well as tools for classifying and computing potential cyclic granularities from an ordered (usually temporal) index.
+Chapter \@ref(ch:gravitas) provides a formal characterization of cyclic granularities as well as tools for classifying and computing potential cyclic granularities from an ordered temporal and non-temporal index.
 It also allows for the manipulation of single- and multiple-order-up time granularities via cyclic calendar algebra.
@@
 And, when there are numerous granularities to choose from, deciding which one to display can be difficult.
-However, only a few of them may be useful in revealing major patterns.
-In Chapter \@ref(ch:hakear), the search for informative granularities by selecting "significant" granularities.
+Moreover, only a few of them may be useful in revealing major patterns.
+In Chapter \@ref(ch:hakear), the search for informative granularities is facilitated by selecting "significant" granularities.
 A cyclic granularity is referred to as "significant" if there is a significant distributional difference of the measured variable between different categories.
-Chapter \@ref(ch:hakear) defines a distance measure to quantify the distributional differences between displayed distributions.
+Chapter \@ref(ch:hakear) defines a distance measure to quantify these distributional differences.
 A higher value of the distance measure for a cyclic granularity or harmony implies that they could be interesting for further investigation, whereas a low value indicates that nothing noteworthy is unfolding.

### 2021-11-12 Rmd/05-conclusion.Rmd
non-temporal index added

@@
 The definitions and rules defined in the literature for linear granularities are insufficient for describing various types of cyclic granularities.
-Chapter \@ref(ch:gravitas) provides a formal characterization of cyclic granularities as well as tools for classifying and computing potential cyclic granularities from an ordered temporal and non-temporal index.
+Chapter \@ref(ch:gravitas) provides a formal characterization of cyclic granularities as well as tools for classifying and computing potential cyclic granularities from an ordered temporal index.
 It also allows for the manipulation of single- and multiple-order-up time granularities via cyclic calendar algebra.
+The approach is generalizable to non-temporal hierarchical granularities with an ordered index.
 
@@
 
+### Computing aperiodic cyclic granularities
+
+Aperiodic cyclic granularities capture repetitions of aperiodic linear granularities.
+All of these are recurring events, but with non-periodic patterns which can still be visualized to understand repetitive behavior.
+The computation of cyclic aperiodic granularities requires first computing aperiodic linear granularities.
+Aperiodic events may be created using R packages such as `almanac`(Vaughan, 2020) and `gs`(Laird-Smith, 2020).
+The functions in `gravitas` package need to integrate with them to support computation of aperiodic cyclic granularities.
+
+### Clustering
+
+Scaling up the clustering method to incorporate large uncertainty and improved computational efficiency that comes with large data
+
+With the volume of data expected to increase in the future, potentially leading to increased variability in patterns, more experiments are needed to understand what sort of problems may arise with our current methodology and how to improve our current algorithm to address those issues.
+All of this must be done while keeping computational efficiency in mind, as this is critical when scaling up for analysis of large data sets.
+
 ### Scaling up the methods for multivariate time series data
@@
 This adds to the already high-dimensional data structures that result from studying distributions.
-Current research on visualizing multivariate data must account for time's inherent characteristics in order to efficiently explore distributions of multivariate time series data.
+This big problem can be tackled by first incorporating time’s inherent characteristics while visualizing multivariate time series data for one individual.
+Then a similar approach of unsupervised clustering can be used to group multiple time series over multiple time granularities and multiple variables.
+This is a similar approach followed in this thesis for handling univariate time series.
 
@@
 -->
-
-### Scaling up the clustering method to incorporate large uncertainty and improved computational efficiency that comes with large data
-
-With the volume of data expected to increase in the future, potentially leading to increased variability in patterns, more experiments are needed to understand what sort of problems may arise with our current methodology and how to improve our current algorithm to address those issues.
-All of this must be done while keeping computational efficiency in mind, as this is critical when scaling up for analysis of large data sets.
 

### 2021-11-13 Rmd/05-conclusion.Rmd
another go at conclusions and future work

@@
 
-### Computing aperiodic cyclic granularities
+<!-- ### Computing aperiodic cyclic granularities -->
 
-Aperiodic cyclic granularities capture repetitions of aperiodic linear granularities.
+<!-- Aperiodic cyclic granularities capture repetitions of aperiodic linear granularities.
 All of these are recurring events, but with non-periodic patterns which can still be visualized to understand repetitive behavior.
@@
 The functions in `gravitas` package need to integrate with them to support computation of aperiodic cyclic granularities.
+-->
 
-### Clustering
+#### Non-stationary time series
 
-Scaling up the clustering method to incorporate large uncertainty and improved computational efficiency that comes with large data
+Because the time series are observed over a short period of time (1-3 years), they are assumed to be stationary.
+But it is possible that the distributions change over time, even over a short period.
+For the smart meter example, the distribution for a customer moving to a different house or changing electrical equipment can change drastically.
+To detect these dynamic changes, non-stationarity in time series has to be incorporated while visualizing distributions and also computing distances for two non-stationary time series or one stationary and another non-stationary time series.
 
-With the volume of data expected to increase in the future, potentially leading to increased variability in patterns, more experiments are needed to understand what sort of problems may arise with our current methodology and how to improve our current algorithm to address those issues.
-All of this must be done while keeping computational efficiency in mind, as this is critical when scaling up for analysis of large data sets.
+#### Structured missingness
+
+It is possible that data for a whole category of cyclic granularity is unavailable or that there are insufficient observations to compute distributions.
+For example, a customer does not have data for a particular day of the week or month throughout their observation period.
+While visualizing probability distributions across categories in Chapters \@ref(ch:gravitas) and \@ref(ch:hakear), this can be indicated by displaying dot plots instead of summarizing distributions.
+But the distances in Chapter \@ref(ch:gracsr) can not handle missing observations if they are structured like this.
+The distances have to be designed to incorporate customers with structural missingness and more research is needed to fully comprehend its implications while visually characterizing them.
+
+#### Anomalies
+
+Characterizing clusters with varied or outlying customers can result in patterns that do not represent the group.
+Moreover, integrating heterogeneous consumers may result in visually identical end clusters, which are potentially not useful.
+An appropriate anomaly detection approach could be applied to filter the anomalies, and then a decision has to be made regarding if the monsters should form a group or be assigned to the closest group through a classification method.
+
+### Develop inferential methods for the clusters
+
+It is important to determine whether or not any detected clusters are statistically significant.
+In Chapter \@ref(ch:gracsr), we have used a permutation approach to detect if patterns are authentic or not across one or a pair of granularities.
+A similar approach could be used for finding the significance of clusters, but permutations used for comparisons must adjust for the fact that different temporal granularities have distinct conditional distributions.
+
+#### Scaling up for larger uncertainty and computational efficiency
+
+Larger data sets with more uncertainty complicate matters, with potentially more problems than already stated above.
+More research is needed to understand what other sorts of problems may arise with the current methodology and possible ways to address them.
+All of this must be done while keeping computational efficiency in mind, as this is critical when scaling up for the analysis of large data sets.
 
@@
 
+Another possible extension would be to create a similar framework for visualizing and analyzing multivariate time series data.
 With multiple time series available for each observation, the complexity of efficient exploration and visualization grows exponentially.
@@
 This adds to the already high-dimensional data structures that result from studying distributions.
-This big problem can be tackled by first incorporating time’s inherent characteristics while visualizing multivariate time series data for one individual.
+This big problem can be tackled by first incorporating time’s inherent characteristics while visualizing one or a few multivariate time series data.
 Then a similar approach of unsupervised clustering can be used to group multiple time series over multiple time granularities and multiple variables.
-This is a similar approach followed in this thesis for handling univariate time series.
+This is a similar approach to that followed in this thesis for handling univariate time series.
 
@@
 
-### Develop inferential methods for the clusters
+<!-- ### Extend methods to other fields -->
 
-It is crucial to determine whether or not any detected clusters or patterns are statistically significant.
-In Chapter \@ref(ch:gracsr), we have used a permutation approach to detect if patterns are authentic or not across one or a pair of granularities.
-A similar method for finding significance across clusters is not appropriate if the observations are assumed to be independent.
-Most large temporal data sets gathered at fine temporal scales usually include several repetitive patterns.
-Therefore, the permutations used for comparisons must account that different temporal granularities have distinct conditional distributions.
-
-### Extend methods to other fields
-
-We propose solutions that are realistically applicable to any temporal data collected more than once per year and may be implemented in a variety of settings.
+<!-- We propose solutions that are realistically applicable to any temporal data collected more than once per year and may be implemented in a variety of settings.
 We could simply verify its utility in the context of smart meter electricity data.
 With the availability of several open-source benchmark data sets, it is vital to evaluate how well the methodologies perform across a variety of fields.
+-->

### 2021-11-13 Rmd/05-conclusion.Rmd
another go at conclusion and future work

@@
 
-The **gravitas** package provides very general tools to compute and manipulate cyclic granularities and generate plots displaying distributions conditional on those granularities.
+The `gravitas` package provides very general tools to compute and manipulate cyclic granularities and generate plots displaying distributions conditional on those granularities.
 The functions `search_gran()`, `create_gran()`, `harmony()`, `gran_advice()` and `prob_plot()` provides the entire workflow for an analyst to systematically explore large quantities of temporal data across different harmonies (pairs of granularities that can be analyzed together.) This package was developed as part of my internship at Google Summer of Code, 2019.
@@
 
-The R package hakear (https://github.com/Sayani07/hakear) implements ideas presented in Chapter \@ref(ch:hakear).
+The R package `hakear` (https://github.com/Sayani07/hakear) implements ideas presented in Chapter \@ref(ch:hakear).
 The function wpd() computes the weighted pairwise distances ($wpd$) for each cyclic granularity or pair of granularities, and `select_harmonies()` selects the ones with significant patterns and ranks them from highest to lowest $wpd$.
@@
 
-#### Non-stationary time series
+**Non-stationary time series**
 
@@
 
-#### Structured missingness
+**Structured missingness**
 
@@
 But the distances in Chapter \@ref(ch:gracsr) can not handle missing observations if they are structured like this.
-The distances have to be designed to incorporate customers with structural missingness and more research is needed to fully comprehend its implications while visually characterizing them.
+More research is needed to design a distance that can incorporate customers with structural missingness and also comprehend its implications while visually characterizing them.
 
-#### Anomalies
+**Anomalies**
 
@@
 Moreover, integrating heterogeneous consumers may result in visually identical end clusters, which are potentially not useful.
-An appropriate anomaly detection approach could be applied to filter the anomalies, and then a decision has to be made regarding if the monsters should form a group or be assigned to the closest group through a classification method.
+An appropriate anomaly detection approach could be applied to filter the anomalies, and then a decision has to be made regarding if the anomalies should form a group or be assigned to the closest group through a classification method.
 
-### Develop inferential methods for the clusters
+**Develop inferential methods for the clusters**
 
@@
 
-#### Scaling up for larger uncertainty and computational efficiency
+**Scaling up for larger uncertainty and computational efficiency**
 
@@
 
-### Scaling up the methods for multivariate time series data
+**Scaling up the methods for multivariate time series data**
 
@@
 This big problem can be tackled by first incorporating time’s inherent characteristics while visualizing one or a few multivariate time series data.
-Then a similar approach of unsupervised clustering can be used to group multiple time series over multiple time granularities and multiple variables.
-This is a similar approach to that followed in this thesis for handling univariate time series.
+Unsupervised clustering can then be used to group multiple time series across multiple time granularities and variables.
+This is a method similar to the one used in this thesis for dealing with univariate time series.
 

### 2021-11-14 Rmd/00-b-abstract.Rmd
ch: to ch-

@@
-# Abstract {- #ch:abstract}
+# Abstract {- #ch-abstract}
 
@@
 This work investigates how time may be dissected, resulting in alternative data segmentation and, as a result, different visualizations that can aid in the identification of underlying patterns.
-The first contribution (Chapter \@ref(ch:gravitas)) describes classes of time deconstructions using linear and cyclic time granularities.
+The first contribution (Chapter \@ref(ch-gravitas)) describes classes of time deconstructions using linear and cyclic time granularities.
 It provides tools to compute possible cyclic granularities from an ordered (usually temporal) index and also a framework to systematically explore the distribution of a univariate variable conditional on two cyclic time granularities by defining "harmony".
@@
 This approach is still overwhelming for human consumption due of the vast number of harmonies remaining.
-The second contribution (Chapter \@ref(ch:hakear)) refines the search for informative granularities by identifying those for which the differences between the displayed distributions are greatest and also rating them in order of importance of capturing maximum variation.
-The third contribution (Chapter \@ref(ch:gracsr)) builds upon the first two to provide methods for exploring heterogeneities in repetitive behavior for many time series and over multiple granularities.
+The second contribution (Chapter \@ref(ch-hakear)) refines the search for informative granularities by identifying those for which the differences between the displayed distributions are greatest and also rating them in order of importance of capturing maximum variation.
+The third contribution (Chapter \@ref(ch-gracsr)) builds upon the first two to provide methods for exploring heterogeneities in repetitive behavior for many time series and over multiple granularities.
 It accomplishes this by providing a way to cluster time series based on probability distributions across informative cyclic granularities.

### 2021-11-14 Rmd/00-e-preface.Rmd
ch: to ch-

@@
-# Preface {-#ch:preface}
+# Preface {-#ch-preface}
 
-Chapter \@ref(ch:gravitas) has been accepted by the *Journal of Computational and Graphical Statistics*.
+Chapter \@ref(ch-gravitas) has been accepted by the *Journal of Computational and Graphical Statistics*.
 It has won the ACEMS Business Analytics Prize in 2020.
 The accompanying R package **gravitas** is on CRAN.
-Chapter \@ref(ch:hakear) and Chapter \@ref(ch:gracsr) are yet to be submitted.
+Chapter \@ref(ch-hakear) and Chapter \@ref(ch-gracsr) are yet to be submitted.
 

### 2021-11-14 Rmd/01-intro.Rmd
structural changes proposed in the meeting today
ch: to ch-

@@
-# Introduction {#ch:intro}
+# Introduction {#ch-intro}
 
@@
 
-Chapter \@ref(ch:gravitas) describes classes of time deconstructions using linear and cyclic time granularities, which can be used to create data visualizations to explore periodicities, associations, and anomalies.
+Chapter \@ref(ch-gravitas) describes classes of time deconstructions using linear and cyclic time granularities, which can be used to create data visualizations to explore periodicities, associations, and anomalies.
 It provides a formal characterization of cyclic granularities and facilitates manipulation of single- and multiple-order-up time granularities through cyclic calendar algebra, as well as providing a recommendation algorithm to check the feasibility of creating plots for any two cyclic granularities.
@@
 
-Chapter \@ref(ch:hakear) is a natural extension of Chapter \@ref(ch:gravitas).
+Chapter \@ref(ch-hakear) is a natural extension of Chapter \@ref(ch-gravitas).
 Many, many displays might be built using cyclic granularities.
@@
 Furthermore, a few of the displays in this collection will be more engaging than others.
-Chapter \@ref(ch:hakear) provides a new distance metric for selecting and ranking the multiple granularities.
+Chapter \@ref(ch-hakear) provides a new distance metric for selecting and ranking the multiple granularities.
 The statistical significance of potential visual discoveries is aided by selecting a threshold for the proposed numerical distance measure.
@@
 
-In Chapter \@ref(ch:gracsr), we look at the problem of using clustering to discover patterns in a large number of univariate time series across multiple temporal granularities.
+In Chapter \@ref(ch-gracsr), we look at the problem of using clustering to discover patterns in a large number of univariate time series across multiple temporal granularities.
 Time series clustering research is gaining traction as more data is collected at finer temporal resolution, over longer time periods, and for a larger number of individuals/entities.
@@
 
-## Summary
+## Thesis structure
 
 The thesis is structured as follows.
-Chapter \@ref(ch:gravitas) provides details of the cyclic granularities, different classes, and computation, and also its usage in exploratory time series analysis through applications.
+Chapter \@ref(ch-gravitas) provides details of the cyclic granularities, different classes, and computation, and also its usage in exploratory time series analysis through applications.
 This is implemented in the R package `gravitas`.
-Chapter \@ref(ch:hakear) provides guidance on how to choose significant cyclic granularities, which are likely to have interesting patterns across its categories.
+Chapter \@ref(ch-hakear) provides guidance on how to choose significant cyclic granularities, which are likely to have interesting patterns across its categories.
 This is available as the R package `hakear`.
-Chapter \@ref(ch:gracsr) provides methods to explore heterogeneity in repetitive behavior for multiple time series over multiple cyclic granularities.
+Chapter \@ref(ch-gracsr) provides methods to explore heterogeneity in repetitive behavior for multiple time series over multiple cyclic granularities.
 This is in the developing R package `gracsr`.
-Chapter \@ref(ch:conclusion) summarizes the software tools developed for the work, and discusses some future plans.
+Chapter \@ref(ch-conclusion) summarizes the software tools developed for the work, and discusses some future plans.

### 2021-11-14 Rmd/05-conclusion.Rmd
ch: to ch-

@@
-# Conclusion {#ch:conclusion}
+# Conclusion {#ch-conclusion}
 
@@
 The building blocks of this framework are presented in each chapter.
-Chapter \@ref(ch:gravitas) contains tools for computing all cyclic granularities, as well as a recommendation system for selecting pairs of granularities that can be effectively analyzed together.
+Chapter \@ref(ch-gravitas) contains tools for computing all cyclic granularities, as well as a recommendation system for selecting pairs of granularities that can be effectively analyzed together.
 These temporal granularities can be used to generate data visualizations to look for patterns, associations, and anomalies.
 But it is difficult to decide which of the various granularities to display when there are many options.
-Chapter \@ref(ch:hakear) presents a methodology for selecting displays that are interesting, such that differences between displayed distributions are greatest.
+Chapter \@ref(ch-hakear) presents a methodology for selecting displays that are interesting, such that differences between displayed distributions are greatest.
 The methodology also ranks the displays in order of priority for capturing the most variation.
-The ideas in Chapters \@ref(ch:gravitas) and \@ref(ch:hakear) can be used for studying patterns in individual time series or comparing a few time series together.
-This is extended in Chapter \@ref(ch:gracsr) to allow for the exploration of distributions for multiple time series at the same time using unsupervised clustering.
+The ideas in Chapters \@ref(ch-gravitas) and \@ref(ch-hakear) can be used for studying patterns in individual time series or comparing a few time series together.
+This is extended in Chapter \@ref(ch-gracsr) to allow for the exploration of distributions for multiple time series at the same time using unsupervised clustering.
 The clustering methodology produces small groups of time series with similar distributions over multiple granularities.
@@
 The definitions and rules defined in the literature for linear granularities are insufficient for describing various types of cyclic granularities.
-Chapter \@ref(ch:gravitas) provides a formal characterization of cyclic granularities as well as tools for classifying and computing potential cyclic granularities from an ordered temporal index.
+Chapter \@ref(ch-gravitas) provides a formal characterization of cyclic granularities as well as tools for classifying and computing potential cyclic granularities from an ordered temporal index.
 It also allows for the manipulation of single- and multiple-order-up time granularities via cyclic calendar algebra.
@@
 However, there may be too many cyclic granularities to look at manually for comprehensive exploration, and not all pairs of granularities can be effectively explored together.
-Chapter \@ref(ch:gravitas) also provides a recommendation on whether a pair of granularities can be meaningfully plotted or analyzed together (a "harmony") or when they cannot (a "clash" or "near-clash").
+Chapter \@ref(ch-gravitas) also provides a recommendation on whether a pair of granularities can be meaningfully plotted or analyzed together (a "harmony") or when they cannot (a "clash" or "near-clash").
 
@@
 Moreover, only a few of them may be useful in revealing major patterns.
-In Chapter \@ref(ch:hakear), the search for informative granularities is facilitated by selecting "significant" granularities.
+In Chapter \@ref(ch-hakear), the search for informative granularities is facilitated by selecting "significant" granularities.
 A cyclic granularity is referred to as "significant" if there is a significant distributional difference of the measured variable between different categories.
-Chapter \@ref(ch:hakear) defines a distance measure to quantify these distributional differences.
+Chapter \@ref(ch-hakear) defines a distance measure to quantify these distributional differences.
 A higher value of the distance measure for a cyclic granularity or harmony implies that they could be interesting for further investigation, whereas a low value indicates that nothing noteworthy is unfolding.
@@
 
-The ideas in Chapters \@ref(ch:gravitas) and \@ref(ch:hakear) can be used for studying patterns in individual time series or comparing a few time series together.
-This is extended in Chapter \@ref(ch:gracsr) to allow for the exploration of distributions for multiple time series at the same time using unsupervised clustering.
+The ideas in Chapters \@ref(ch-gravitas) and \@ref(ch-hakear) can be used for studying patterns in individual time series or comparing a few time series together.
+This is extended in Chapter \@ref(ch-gracsr) to allow for the exploration of distributions for multiple time series at the same time using unsupervised clustering.
 In the time series clustering literature, probability distributions across cyclic granularities have not been considered in determining similarity.
 However, such a similarity measure can be useful for characterizing the inherent temporal data structure of long, unequal-length time series in a way that is resistant to missing or noisy data while allowing for the detection of similar repeated patterns.
-Chapter \@ref(ch:gracsr) proposes two approaches for calculating distances between time series based on probability distributions across cyclic granularities.
+Chapter \@ref(ch-gracsr) proposes two approaches for calculating distances between time series based on probability distributions across cyclic granularities.
 The first approach considers two time series to be similar if the distributions of each category of one or more cyclic granularities are similar.
@@
 There has been a total of 12K downloads from the RStudio mirror dating from 2020-11-01 to 2021-11-01.
-This package supplements the paper corresponding to Chapter \@ref(ch:gravitas), which has won the ACEMS Business Analytics Award 2021.
+This package supplements the paper corresponding to Chapter \@ref(ch-gravitas), which has won the ACEMS Business Analytics Award 2021.
 The package can be generalized to non-temporal applications for which a hierarchical structure can be construed similar to time.
@@
 
-The R package `hakear` (https://github.com/Sayani07/hakear) implements ideas presented in Chapter \@ref(ch:hakear).
+The R package `hakear` (https://github.com/Sayani07/hakear) implements ideas presented in Chapter \@ref(ch-hakear).
 The function wpd() computes the weighted pairwise distances ($wpd$) for each cyclic granularity or pair of granularities, and `select_harmonies()` selects the ones with significant patterns and ranks them from highest to lowest $wpd$.
@@
 
-The open-source R package `gracsr` is available on Github (https://github.com/Sayani07/gracsr) to implement ideas presented in Chapter \@ref(ch:gracsr).
+The open-source R package `gracsr` is available on Github (https://github.com/Sayani07/gracsr) to implement ideas presented in Chapter \@ref(ch-gracsr).
 The package contains functions for implementing the complete clustering methodology with choices of scaling and distance metrics discussed in the chapter.
@@
 For example, a customer does not have data for a particular day of the week or month throughout their observation period.
-While visualizing probability distributions across categories in Chapters \@ref(ch:gravitas) and \@ref(ch:hakear), this can be indicated by displaying dot plots instead of summarizing distributions.
-But the distances in Chapter \@ref(ch:gracsr) can not handle missing observations if they are structured like this.
+While visualizing probability distributions across categories in Chapters \@ref(ch-gravitas) and \@ref(ch-hakear), this can be indicated by displaying dot plots instead of summarizing distributions.
+But the distances in Chapter \@ref(ch-gracsr) can not handle missing observations if they are structured like this.
 More research is needed to design a distance that can incorporate customers with structural missingness and also comprehend its implications while visually characterizing them.
@@
 It is important to determine whether or not any detected clusters are statistically significant.
-In Chapter \@ref(ch:gracsr), we have used a permutation approach to detect if patterns are authentic or not across one or a pair of granularities.
+In Chapter \@ref(ch-gracsr), we have used a permutation approach to detect if patterns are authentic or not across one or a pair of granularities.
 A similar approach could be used for finding the significance of clusters, but permutations used for comparisons must adjust for the fact that different temporal granularities have distinct conditional distributions.

### 2021-11-15 Rmd/02-gravitas.Rmd
ch: to ch-
paper-gracsr update

@@
-# Visualizing probability distributions across bivariate cyclic temporal granularities {#ch:gravitas}
+# Visualizing probability distributions across bivariate cyclic temporal granularities {#ch-gravitas}
 
@@
 
-\noindent In the Mayan calendar (Table \@ref(tab:tab-mayan)), kin-of-tun or kin-of-baktun are examples of multiple-order-up granularities and single-order-up granularities are kin-of-uinal, uinal-of-tun etc.
+\noindent In the Mayan calendar (Table \@ref{tab:tab-mayan}), kin-of-tun or kin-of-baktun are examples of multiple-order-up granularities and single-order-up granularities are kin-of-uinal, uinal-of-tun etc.
 

### 2021-11-15 Rmd/04-gracsr.Rmd
ch: to ch-
paper-gracsr update

@@
-# Clustering time series based on probability distributions across temporal granularities {#ch:gracsr}
+# Clustering time series based on probability distributions across temporal granularities {#ch-gracsr}
 
-With more and more time series data being collected at much finer temporal resolution, for a longer length of time, and for a larger number of individuals/entities, time series clustering research is getting a lot of traction.
-Long, noisy, patchy, uneven, and asynchronous time series are common in many fields, limiting similarity searches or lowering method efficiency when clustering is based on a distance metric.
-In this work, we suggest two approaches for obtaining similarity between time series based on probability distributions over cyclic temporal granularities for distance-based clustering approaches.
-Cyclic granularities like hour-of-the-day, work-day/weekend, month-of-the-year and so on, are useful for finding repeated patterns in the data.
-Looking at probability distributions across cyclic granularities serves two purposes: (a) "Probability distributions" characterize the inherent temporal data structure of these large unequal-length time series and are robust to missing or noisy data.
-(b) Using probability distributions over "cyclic granularities" ensures small pockets of similar "repeated" behaviors.
-Our method is capable of producing useful clusters, as demonstrated by testing on validation data designs and a sample of residential smart meter consumers.
+Clustering is a potential approach for Organizing large collections of times series into small homogeneous groups, but a difficult step is determining an appropriate metric to measure similarity between time series.
+The similarity metric needs to be capable of accommodating long, noisy, and asynchronous time series and also capture cyclical patterns.
+In this paper, two approaches for measuring distances between time series are presented, based on probability distributions over cyclic temporal granularities.
+Both are compatible with a variety of clustering algorithms.
+Cyclic granularities like hour-of-the-day, work-day/weekend, month-of-the-year, are useful for finding repeated patterns in the data.
+Measuring similarity based on probability distributions across cyclic granularities serves two purposes: (a) characterizing the inherent temporal data structure of long, unequal-length time series in a manner robust to missing or noisy data; (b) small pockets of similar repeated behaviors can be captured.
+This approach is capable of producing useful clusters, as demonstrated on validation data designs and a sample of residential smart meter records.
 
-```{r external3, include = FALSE, cache = FALSE}
-knitr::read_chunk('scripts/gracsr.R')
+```{r external-gracsr, include = FALSE, cache = FALSE}
+#knitr::read_chunk('script/main.R')
+knitr::read_chunk('scripts/gracsr/mylibrary.R')
+knitr::read_chunk('scripts/gracsr/mytheme.R')
+knitr::read_chunk('scripts/gracsr/application.R')
+knitr::read_chunk('scripts/gracsr/application.R')
+knitr::read_chunk('scripts/gracsr/validation.R')
 ```
@@
 ```{r theme_characterisation}
+```
+
+```{r mytheme-application2}
+
+```
+
+```{r mytheme-application3}
+
 ```
@@
 Defining an appropriate distance measure for the raw time series may be a difficult task since it must take into account noise, variable lengths of time series, asynchronous time series, different scales, and missing data.
-Commonly used Distance-based similarity measures as suggested by a decade review of time series clustering approaches (@Aghabozorgi2015-ct) are Euclidean, Pearson's correlation coefficient and related distances, Dynamic Time Warping, Autocorrelation, Short time series distance, Piecewise regularization, cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances (@liao2007clustering) but on a vector time series data.
+Commonly used distance-based similarity measures as suggested by a decade review of time series clustering approaches (@Aghabozorgi2015-ct) are Euclidean, Pearson's correlation coefficient and related distances, Dynamic Time Warping (DTW), Autocorrelation, Short time series distance, Piecewise regularization, cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances (@liao2007clustering) but on a vector time series data.
 Among these alternatives, Euclidean distances have high performance but need the same length of data over the same period, resulting in information loss regardless of whether it is on raw data or a smaller collection of features.
@@
 
-\noindent We consider the problem of clustering a large number of univariate time series of continuous values which are available at fine temporal scales, being motivated by the residential smart meter data.
-These time series data are long (with more and more data collected at finer resolutions), are asynchronous, with varying time lengths for different houses and missing observations and characterized by noisy and patchy behavior that can quickly become overwhelming and hard to interpret, requiring summarizing the large number of customers into pockets of similar energy behavior.
-Choosing probability distributions seem to be a natural way to analyze these types of data sets since they are robust to uneven length, missing data, or noise.
-This paper proposes two approaches for obtaining pairwise similarities based on Jensen-Shannon distances between probability distributions across significant cyclic granularities.
-Cyclic temporal granularities [@Gupta2021-hakear], which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can be useful for measuring repetitive patterns in large univariate time series data.
-The resulting clusters are expected to group customers that have similar repetitive behaviors across each of the interesting cyclic granularities.
-Below are some of the benefits of our method:
+\noindent This work is motivated by a need to cluster a large collection of residential smart meter data, so that households can be grouped into similar energy usage patterns.
+These can be considered to be univariate time series of continuous values which are available at fine temporal scales.
+These time series data are long (with more and more data collected at finer resolutions), are asynchronous, with varying time lengths for different houses and sporadic missing values.
+Using probability distributions is a natural way to analyze this types of data because they are robust to uneven length, missing data, or noise.
+This paper proposes two approaches for obtaining pairwise similarities based on Jensen-Shannon distances between probability distributions across a selection of cyclic granularities.
+Cyclic temporal granularities [@Gupta2021-hakear], which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can measure repetitive patterns in large univariate time series data.
+The resulting clusters are expected to group customers that have similar repetitive behaviors across cyclic granularities.
+The benefits of this approach are as follows.
 
-- When using probability distributions, data does not have to be the same length or observed during the exact same time period (unless there is a structural pattern);
+- When using probability distributions, data does not have to be the same length or observed during the exact same time period (unless there is a structural pattern).
 
@@
 
-- Jensen-Shannon distances evaluate the distance between two distributions rather than raw data, which is less sensitive to missing observations and outliers than other conventional distance methods;
+- Jensen-Shannon distances evaluate the distance between two distributions rather than raw data, which is less sensitive to missing observations and outliers than other conventional distance methods.
 
-- While most clustering algorithms produce clusters similar across just one temporal granularity, this technique takes a broader approach to the problem, attempting to group observations with similar distributions across all interesting cyclic granularities;
+- While most clustering algorithms produce clusters similar across just one temporal granularity, this technique takes a broader approach to the problem, attempting to group observations with similar distributions across all interesting cyclic granularities.
 
@@
 
-_Background and motivation_
+<!--_Background and motivation_-->
 
-Large spatio-temporal data sets, both from open and administrative sources, offer up a world of possibilities for research.
-One such data sets for Australia is the Smart Grid, Smart City (SGSC) project (2010–2014) available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) The project provides half-hourly data of over 13,000 household electricity smart meters distributed unevenly from October 2011 to March 2014.
-Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.
-Households vary in size, location, and amenities such as solar panels, central heating, and air conditioning.
+The primary application of this work is data from the Smart Grid, Smart City (SGSC) project (2010–2014) available through the [Department of the Environment and Energy](https://data.gov.au/data/organization/doee).
+Half-hourly measurements of usage for more than 13,000 household electricity smart meters is provided from from October 2011 to March 2014.
+<!--Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.--> Households vary in size, location, and amenities such as solar panels, central heating, and air conditioning.
 The behavioral patterns differ amongst customers due to many temporal dependencies.
 Some households use a dryer, while others dry their clothes on a line.
-Their weekly profile may reflect this.
+Their weekly usage profile may reflect this.
 They may vary monthly, with some customers using more air conditioners or heaters than others, while having equivalent electrical equipment and weather circumstances.
 Some customers are night owls, while others are morning larks.
-Day-off energy use varies depending on whether customers stay home or go outside.
-Age, lifestyle, family composition, building attributes, weather, availability of diverse electrical equipment, among other factors, make the task of properly segmenting customers into comparable energy behavior a fascinating one.
-This challenge is worsened when all we know about our consumers is their energy use history (@Ushakova2020-rl).
-To safeguard the customers' privacy, it is probable that such information is not accessible.
-Also, energy suppliers may not always update client information, such as property features, in a timely manner.
-Thus, there is a growing need to have research that examines how much energy usage heterogeneity can be found in smart meter data and what are some of the most common power consumption patterns, rather than explaining why consumption differs.
+Daily energy usage varies depending on whether customers stay home or work away from home.
+Age, lifestyle, family composition, building attributes, weather, availability of diverse electrical equipment, among other factors, make the task of properly segmenting customers into comparable energy behavior complex.
+The challenge is to be able to cluster consumers into these type of expected patterns, and other unexpected patterns, using only their energy usage history (@Ushakova2020-rl).
+<!--To safeguard the customers' privacy, it is probable that such information is not accessible.
+Also, energy suppliers may not always update client information, such as property features, in a timely manner.--> There is a growing need to have methods that can examine the energy usage heterogeneity observed in smart meter data and what are some of the most common power consumption patterns.
 
-_Related work_
+<!-- _Related work_-->
 
-A multitude of papers have emerged around smart meter time series clustering for deepening our knowledge of consumption patterns.
+There is a growing body of literature focused on time series clustering related to smart meter data.
 @Tureczek2017-pb conducted a systematic study of over $2100$ peer-reviewed papers on smart meter data analytics.
-None of the $34$ articles chosen for their emphasis use Australian smart meter data.
-The most often used algorithm is K-Means.
-Using K-Means without considering time series structure or correlation results in inefficient clusters.
-Principal Component Analysis (PCA) or Self-Organizing Maps (SOM) eliminate correlation patterns and decrease feature space, but lose interpretability.
-To reduce dimensionality, several studies use principal component analysis or factor analysis to pre-process smart-meter data before clustering (@Ndiaye2011-pf).
-Other algorithms utilized in the literature include k-means variants, hierarchical approaches, and greedy k-medoids.
+<!--None of the $34$ articles chosen for their emphasis use Australian smart meter data.--> The most often used algorithm is $k$-means [@Rhodes2014-it].
+$k$-means can be made to perform better by explicitly incorporating time series features such as correlation or cyclic patterns rather than performing it on raw data.
+To reduce dimensionality, several studies use principal component analysis (PCA) or factor analysis to pre-process smart-meter data before clustering (@Ndiaye2011-pf).
+PCA eliminates correlation patterns and decreases feature space, but loses interpretability.
+Other algorithms utilized in the literature include $k$-means variants, hierarchical clustering, and greedy $k$-medoids.
 Time series data, such as smart meter data, are not well-suited to any of the techniques mentioned in @Tureczek2017-pb.
-Only one study [@ozawa2016determining] identified time series characteristics using Fourier transformation, which converts data from time to frequency and then uses K-Means to cluster by greatest frequency.
-@Motlagh2019-yj suggests that the time feature extraction is limited by the type of noisy, patchy, and unequal time-series common in residential and addresses model-based clustering by transforming the series into other other objects such as structure or set of parameters which can be more easily characterized and clustered.
+Only one study [@ozawa2016determining] identified time series characteristics by first conducting a Fourier transformation, to convert data from time to frequency domain, followed by $k$-means to cluster by greatest frequency.
+@Motlagh2019-yj suggests that the time feature extraction is limited by the type of noisy, patchy, and unequal time-series common in residential customers and addresses model-based clustering by transforming the series into other objects such as structure or set of parameters which can be more easily characterized and clustered.
 [@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
 @Melnykov2013-sp discusses how outliers, noisy observations and scattered observations can complicate estimating mixture model parameters and hence the partitions.
-To our knowledge, none of the methods focus on exploring heterogeneity in repetitive behaviors based on the dynamics of multiple temporal dependencies using probability distributions.
+None of these methods focuses on exploring heterogeneity in repetitive patterns based on the dynamics of multiple temporal dependencies using probability distributions, which forms the basis of the methodology reported here.
 
@@
 
-The remainder of the paper is organized as follows: Section&nbsp;\ref{sec:methodology} provides the clustering methodology.
-Section&nbsp;\ref{sec:validation} shows data designs to validate our methods.
-Section&nbsp;\ref{sec:cases} discusses the application of the method to a subset of the real data.
-Finally, we summarize our results and discuss possible future directions in Section&nbsp;\ref{sec:conclusion}.
+This paper is organized as follows.
+Section~\ref{sec:methodology} provides the clustering methodology.
+Section~\ref{sec:validation} shows data designs to validate our methods.
+Section~\ref{sec:application-gracsr} discusses the application of the method to a subset of the real data.
+Finally, we summarize our results and discuss possible future directions in Section~\ref{sec:discussion-gracsr}.
 
@@
 
-The foundation of our method is unsupervised clustering algorithms based exclusively on the time-series data.
-The proposed methodology aims to leverage the intrinsic data structure hidden within cyclic temporal granularities.
-The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
-<!--So going back to the smart meter example, $f_i(v)$ is the distribution of customer $i$ and $v$ is electricity demand.--> In this work, instead of considering the probability distributions of the linear time series, we assume it across different categories of any cyclic granularity.
+<!--The foundation of our method is unsupervised clustering algorithms based exclusively on the time-series data.
+The proposed methods leverage the intrinsic data structure hidden within cyclic temporal granularities.--> The existing work on clustering probability distributions assumes we have an independent and identically distributed samples $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
+<!--So going back to the smart meter example, $f_i(v)$ is the distribution of customer $i$ and $v$ is electricity demand.--> In our approach, instead of considering the probability distributions of the linear time series, we compare them across different categories of a cyclic granularity.
 We can consider categories of an individual cyclic granularity ($A$) or combination of categories for two interacting granularities ($A*B$) to have a distribution, where $A, B$ are defined as $A = \{a_j: j=1, 2, \dots J\}$ and $B = \{b_k: k = 1, 2, \dots K\}$.
-For example, let us consider two cyclic granularities $A$ and $B$ representing hour-of-day and day-of-week.
+For example, let us consider two cyclic granularities, $A$ and $B$, representing hour-of-day and day-of-week, repsectively.
 Then $A = \{0, 1, 2, \dots, 23\}$ and $B = \{Mon, Tue, Wed, \dots, Sun\}$.
@@
 Hence clustering these customers is equivalent to clustering these collections of conditional univariate probability distributions.
-Towards this goal, we need to decide how to measure similarities between collections of univariate probability distributions.
-There are multiple ways to measure similarities depending on the aim of the analysis.
+Towards this goal, the next step is to decide how to measure distances between collections of univariate probability distributions.
+<!--There are multiple ways to measure similarities depending on the aim of the analysis.
 <!--This paper focuses on looking at the (dis) similarity between underlying distributions that may have resulted in different patterns across different cyclic temporal granularities, that eventually have resulted in the (dis) similarity between time series.
---> This paper considers a two approaches for finding distances between time series.
-Both of these approaches may be useful in a practical context and, depending on the data set, may or may not propose the same customer classification.
-The obtained distances could be fed into a clustering algorithm to break large data sets into subgroups that can then be analyzed separately.
+--> Here, we describe two approaches for finding distances between time series.
+Both of these approaches may be useful in a practical context, and produce very different but equally useful customer groupings.
+The distances can be supplied to any usual clustering algorithm, including $k$-means or hierarchical clustering, to group observations into a smaller more homogeneous collection.
 <!--These clusters may be commonly associated with real-world data segmentation.
-However, since the data is unlabeled a priori, more information is required to corroborate this.--> The methodology is explained in the Figure \ref{fig:flowchart} and each element of the pipeline is discussed.
+However, since the data is unlabeled a priori, more information is required to corroborate this.--> The flow of the procedures is illustrated in Figure \ref{fig:flowchart}.
 
-```{r flowchart, fig.cap = "Flow chart illustrating the pipeline for methodology", out.width="100%"}
+```{r flowchart, fig.cap = "Flow chart illustrating the pipeline for out method for clustering time series.", out.width="100%"}
 knitr::include_graphics("img/flowchart.png")
@@
 
-- _Find significant granularities or harmonies_
+### Selecting granularities
 
-[@Gupta2021-hakear] proposes a method for choosing significant cyclic granularities and harmonies (interacting granularities that can be studied together) [@Gupta2021-gravitas], which is used in this work.
-We define "significant" granularities as those with significant distributional differences across categories.
-It is preferable to consider those since it is more probable that there will be some intriguing repeated behavior worth investigating.
-It should be noted that all of the observations in the study may not have the same set of important granularities.
-The following is a method for generating a list ($S_c$) of significant granularities for all observations:
+[@Gupta2021-hakear] provides a method for determining the significance of a cyclic granularity, and a ranking of multiple cyclic granularities.
+(This extends to harmonies, pairs of granularities that might interact with each other.) We define "significant" granularities as those with significant distributional differences across at least one category.
+The reason for subsetting granularities in this way, is that, clustering algorithms perform badly in the presence of nuisance variables.
+Granularities that do not have some difference between categories are likely to be nuisance variables.
+It should be noted that all of the time series in a collection may not have the same set of significant granularities.
+This is the approach for managing this generate a subset ($S_c$) of significant granularities across a collection of time series:
 
- (a) Remove granularities from the comprehensive list that are insignificant for all observations.
+ (a) Remove granularities from the comprehensive list that are not significant for any time series.
 
- (b) select only the granularities that are significant for the majority of observations.
+ (b) Select only the granularities that are significant for the majority of time series.
 
-There will be observations in both cases where one or a few selected granularities are boring.
+<!--There will be time series in both cases where one or a few selected granularities are boring.
 Even in that case, having this group of observations with no interesting patterns at a granularity that regularly discovers patterns may be valuable.
-If, on the other hand, the granularities under examination are truly important for a group of data, unique patterns may be identified while clustering them.
+If, on the other hand, the granularities under examination are truly important for a group of data, unique patterns may be identified while clustering them.-->
 
-- _Data transformation_
+### Data transformation
 
-Observations often have a somewhat skewed time series distribution and their ranges might vary greatly.
+The shape and scale of the distribution of the measured variable (e.g.
+energy usage) affects distance calculations.
+Skewed distributions need to be symmetrized.
+Scales of individuals needs to be standardized, because clustering is to select similar patterns, not magnitude of usage.
+(Organizing individuals based on magnitude can achieved simply by sorting on a statistic like the average value across time.) <!--Observations often have a somewhat skewed time series distribution and their ranges might vary greatly.
 Statistical transformations are employed to bring all of them to the same range or normalize each observation.
-This is important because we are not interested in trivial clusters that vary in magnitudes, but rather in uncovering comparable patterns of distributional differences between categories.
-For the JS-based approaches, two data transformation techniques are utilized viz, Normal-Quantile Transform (NQT) and Robust scaling.
-NQT is a built-in transformation for computing $wpd$, which is the foundation of wpd-based distances.
+This is important because we are not interested in trivial clusters that vary in magnitudes, but rather in uncovering comparable patterns of distributional differences between categories.--> For the JS-based approaches, two data transformation techniques are recommended, normal-quantile transform (NQT) and robust scaling (RS).
+NQT is used in @@Gupta2021-hakear prior sorting granularities.
+<!-- $wpd$, which is the foundation of wpd-based distances.
+XX wpd was not yet introduced-->
 
-_Robust scaling_ The normalized $i^{th}$ observation is denoted by $v_{norm} = \frac{v_t - q_{0.5}}{q_{0.75}-q_{0.25}}$, where $v_t$ is the actual value at the $t^{th}$ time point and $q_{0.25}$, $q_{0.5}$ and $q_{0.75}$ are the $25^{th}$, $50^{th}$ and $75^{th}$ percentile of the time series for the $i^{th}$ observation.
-$v_{norm}$ has zero mean and median, as well as a standard deviation of one, with the outliers having same relative connections to other values.
+- RS: The normalized $i^{th}$ observation is denoted by $v_{norm} = \frac{v_t - q_{0.5}}{q_{0.75}-q_{0.25}}$, where $v_t$ is the actual value at the $t^{th}$ time point and $q_{0.25}$, $q_{0.5}$ and $q_{0.75}$ are the $25^{th}$, $50^{th}$ and $75^{th}$ percentile of the time series for the $i^{th}$ observation.
+Note that, $v_{norm}$ has zero mean and median, but otherwise the shape does not change.
 
-_Normal-Quantile transform_ The raw data for all observations is individually normal-quantile transformed (NQT) [@Krzysztofowicz1997-bv], so that the transformed data follows a standard normal distribution.
-NQT will make the skewed distributions bell-shaped.
-As a result, determining which raw distribution was used is difficult using the modified distribution.
-Multimodality is also concealed or reversed.
-As a result, while displaying transformed data using NQT, one must exercise caution.
-But NQT is a useful transformation that often improves clustering performance.
+- NQT: The raw data for all observations is individually transformed [@Krzysztofowicz1997-bv], so that the transformed data follows a standard normal distribution.
+NQT will symmetrize skewed distributions.
+<!--As a result, determining which raw distribution was used is difficult using the modified distribution.--> A drawback is that any multimodality will be concealed.
+This, this should be checked prior to applying NQT.
 
@@
 
-- _Data pre-preprocessing_
+### Data pre-preprocessing
 
-We start with a "tsibble" (@wang2020tsibble) data structure with an index variable representing inherent ordering from past to present and a key variable that defines observational units over time.
+Computationally in R, the data is assumed to be a "tsibble object" (@wang2020tsibble) equipped with an index variable representing inherent ordering from past to present and a key variable that defines observational units over time.
 The measured variable for an observation is a time-indexed sequence of values.
@@
 A shuffle of the raw sequence may represent hourly consumption throughout a day, a week, or a year.
-Cyclic granularities can be expressed in terms of the index set in the "tsibble" data structure.
-But the data structure changes while transporting from linear to cyclic scale of time as multiple observations now correspond to each category and induce a probability distribution.
-Directly computing Jensen-Shannon distances between the entire probability distributions can be very costly.
-Hence, in this paper, quantiles are chosen to characterize the probability distributions.
-So, in the final data structure, each category of a cyclic granularity corresponds to a list of numbers which is essentially a few chosen quantiles.
+Cyclic granularities can be expressed in terms of the index set in the tsibble data structure.
 
-- _Distance metrics_
+The data object will change when cyclic granularities are computed, as multiple observations will be categorized into levels of the granularity, thus inducing multiple probability distributions.
+Directly computing Jensen-Shannon distances between the entire probability distributions can be computationally intensive.
+Thus it is recommended that quantiles are used to characterize the probability distributions.
+In the final data object, each category of a cyclic granularity corresponds to a list of numbers which is composed of a few quantiles.
+
+### Distance metrics
 
@@
 This needs to be done in a way such that the resulting metric is a distance metric, and could be fed into the clustering algorithm.
-Two types of distance metric is considered:
+Two types of distance metrics are considered:
 
@@
 
-**JS-based distances**
+#### JS-based distances
 
@@
 
-\noindent The sum of the distances between two observations $x$ and $y$ in terms of cyclic granularity $A$ is defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity $A$) or $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$.
+The sum of the distances between two observations $x$ and $y$ in terms of cyclic granularity $A$ is defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity $A$) or $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$).
 After determining the distance between two series in terms of one granularity, we must combine them to produce a distance based on all significant granularities.
@@
 
-**wpd-based distances**
+#### wpd-based distances
 
@@
 $wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction and is estimated by the maximum pairwise distances between consecutive categories normalized by appropriate parameters.
-A higher value of $wpd$ indicates that some interesting pattern is expected, whereas a lower value would indicate otherwise.
+A higher value of $wpd$ indicates that some interesting patterns are expected, whereas a lower value would indicate otherwise.
 
-\noindent Once we have chosen $wpd$ as a relevant feature for characterizing the distributions across one cyclic granularity, we have to decide how we combine differences between the multiple features (corresponding to multiple granularities) into a single number.
+Once we have chosen $wpd$ as a relevant feature for characterizing the distributions across one cyclic granularity, we have to decide how we combine differences between the multiple features (corresponding to multiple granularities) into a single number.
 The Euclidean distance between them is chosen, with the granularities acting as variables and $wpd$ representing the value under each variable.
@@
 
-- _Clustering algorithm_
+### Clustering
+
+#### Number of clusters
+
+Different approaches to determining number of clusters are required for varied applications and research objectives.
+It is often beneficial to investigate numerous cluster statistics in order to arrive at a number of clusters, since clustering has a variety of objectives (e.g., between-cluster separation, within-cluster homogeneity, representation by centroids, etc.), which could be in conflict with one another.
+Some typical ways of determining the number of clusters in the literature (e.g., the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@dunn1973fuzzy]) are inspired by defining and balancing "within-cluster homogeneity" and "between-cluster separation".
+The Cluster separation index ($sindex$) proposed by @Hennig2014-ah is used as a guide to select the number of clusters in Section \ref{sec:validation}.
+$sindex$ [@R-fpc] is a compromise between minimum cluster separation and average cluster separation and is less sensitive to a single or few ambiguous points.
+$sindex$ cannot be optimized over the number of clusters, since increasing the number of clusters reduces it.
+The number of clusters required is determined by the slope going from steep to shallow (an elbow).
+
+.<!--determined by calculating the distance between each observation and the cluster to which it does not belong.
+Then, the least 10% of these distances are averaged to only evaluate places near the cluster edges.-->
+
+#### Algorithm
 
@@
 
-- _Characterization of clusters_
+#### Characterization of clusters
 
-\noindent Cluster characterization is an important element of cluster analysis.
+Cluster characterization is an important element of cluster analysis.
 @Cook2007-qe provides several methods for characterizing clusters.
@@
 Displaying cluster statistics is useful when we have larger problems and it is difficult to read the parallel coordinate plots due to congestion.
-All of MDS, PCA and t-SNE use a distance or dissimilarity matrix to construct a reduced-dimension space representation, their goals are diverse.
+All of MDS, PCA and t-SNE use a distance or dissimilarity matrix to construct a reduced-dimension space representation, but their goals are diverse.
 Multidimensional scaling (@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space.
@@
 A continuous measured variable $v$ of length $T$ indexed by $\{0, 1, \dots T-1\}$ is simulated such that it follows the structure across $g1$, $g2$ and $g3$.
-We constructed independent replications of all data designs $R = \{25, 250, 500\}$ to investigate if our proposed clustering method can discover distinct designs in small, medium, and big number of series.
+We constructed independent replications of all data designs $R = \{25, 250, 500\}$ to investigate if our proposed clustering method can discover distinct designs in small, medium, and big numbers of series.
 All designs employ $T=\{300, 1000, 5000\}$ sample sizes to evaluate small, medium, and large-sized series.
@@
 So three scenarios are considered to accommodate that.
+Figure \ref{fig:range-parameter} shows the range of parameters considered for each scenario.
+
+```{r range-parameter, fig.cap="Range of parameters for different scenarios"}
+knitr::include_graphics("img/simulation_table.png")
+```
 
@@
 
-### Data generating processes
+### Data generation
 
@@
 
-_Scenario (a): All significant granularities_
+_Scenario (S1): All granularities significant_
 
@@
 
-\noindent Consider the instance where $g1$, $g2$, and $g3$ all contribute to design distinction.
+Consider the instance where $g1$, $g2$, and $g3$ all contribute to design distinction.
 This means that each granularity will have significantly different patterns at least across one of the designs to be clustered.
@@
 
-_Scenario (b): Few significant granularities_
+_Scenario (S2): Few significant granularities_
 
-\noindent This is the case where one granularity will remain the same across all designs.
+This is the case where one granularity will remain the same across all designs.
 We consider the case where the distribution of $v$ varies across $g2$ levels for all designs, across $g3$ levels for a few designs, and $g1$ does not vary across designs.
@@
 
-_Scenario (c): One significant granularity_
+_Scenario (S3): Only one significant granularity_
 
-\noindent Only one granularity is responsible for identifying the designs in this case.
-This is depicted in Figure \ref{fig:gran2and1-clubbed} (right) where only $g3$ affects the designs significantly.
+Only one granularity is responsible for identifying the designs in this case. This is depicted in Figure \ref{fig:gran2and1-clubbed} (right) where only $g3$ affects the designs significantly.
 
@@
 
-```{r plot-3gran-new, out.width="100%", fig.cap="The linear (left) and cyclic (right) representation of the simulated variable is shown.
-Each row represents a design in Scenario (a).
-In this scenario, all of $g1$, $g2$ and $g3$ changes across at least one design.
-Also, it is not possible to comprehend these differences in patterns just by looking at or considering the linear representation."}
+```{r plot-3gran-new, out.width="100%", fig.cap = "The linear (left) and cyclic (right) representation is shown under scenario S1 using line plots and boxplots respectively.
+Each row represents a design.
+Distributions of categories across $g1$, $g2$ and $g3$ change across at least one design as can be observed in the cyclic representation.
+It is not possible to comprehend these structural differences in patterns just by looking at or considering the linear representation."}
 
@@
 
-```{r gran2and1-clubbed, fig.cap=" Plots (b) and (c) correspond to Design (b) and (c) respectively.
-In (b) $g2$, $g3$ changes across atleast one design but $g1$ remains constant.
-Only $g3$ changes across different designs in (c)."}
+```{r gran2and1-clubbed, fig.cap=" Boxplots showing distributions of categories across different designs (rows) and granularities (columns) for scenarios S2 and S3.
+In S2, $g2$, $g3$ change across at least one design but $g1$ remains constant.
+Only $g3$ changes across different designs in S3."}
 #gran2_change  + gran1_change + plot_layout(widths = c(1,1))
@@
 Consider a group that has a different weekday and weekend behavior in the summer but not in the winter.
-This type of combined behavior across granularities can be discovered by evaluating the distribution across combinations of categories for different interacting granularities (Weekend/Weekday and month-of-year in this example).
+This type of combined behavior across granularities can be discovered by evaluating the distribution across combinations of categories for different interacting granularities (weekend/weekday and month-of-year in this example).
 As a result, in this scenario, we analyze a combination of categories generated from different distributions.
-Consider a case in which there are only two interacting granularities of interest, $g1$ and $g2$.
-In contrast to the previous situation, when we could study distributions across $n_{g_1} + n_{g_2} = 5$ separate categories, with interaction, we must evaluate the distribution of the $n_{g_1}*n_{g_2}=6$ combination of categories.
-Consider the $4$ designs in Figure \ref{fig:interaction-gran}, where various distributions are assumed for different combinations of categories, resulting in different designs.
-Design $D1$ exhibits no change in distributions across $g1$ or $g2$, whereas Designs $D2$ and $D3$ alter across only $g1$ and $g2$, respectively.
-D4 varies across both $g1$ and $g2$ categories.
-D3 and D4 appear similar based on their relative differences across consecutive categories, but $D4$ also changes across facets, unlike $D3$, which has all facets look the same.
-
-```{r interaction-gran, fig.cap = "Distribution of the simulated variable across $g1$ conditional on $g2$ is shown through boxplots for 4 designs.
-D1 has no change in distributions across different categories of $g1$ or $g2$, while D2 and D3 change across only $g1$ and $g2$ respectively.
-D4 changes across categories of both $g1$ and $g2$."}
-
-```
+Display of design and related results can be found in supplementary.
 
@@
 
-### Visual exploration of findings
+### Visual exploration of results
 
-All of the approaches were fitted to each data design and for each combination of the considered parameters.
+All of the approaches were fitted to each data design and to each combination of the considered parameters.
 The formed clusters have to match the design, be well separated, and have minimal intra-cluster variation.
 It is possible to study these desired clustering traits visually in a more comprehensive way than just looking at index values.
-So we use MDS and parallel coordinate graphs to demonstrate the findings:
+MDS and parallel coordinate graphs are used to demonstrate the findings, as well as an index value plot to provide direction on the number of clusters.
+JS-based approaches corresponding to NQT and RS are referred to as JS-NQT and JS-RS respectively.
+In the following plots, results for JS-NQT are reported.
+For similar results with JS-RS or wpd-based distances, please refer to the supplementary paper.
 
-- In Figure \ref{fig:mds-plot-validation}, we tried to see how separated our clusters are.
-We observe that in all scenarios and for different mean differences, clusters are separated.
+- Figure \ref{fig:sindex-plot-validation} shows $sindex$ plotted against the number of clusters ($k$) for the range of mean differences (rows) under the different scenarios (columns).
+This can be used to determine the number of clusters for each scenario.
+When $sindex$ for each scenario are examined, it appears that $k = 5, 4, 4$ is justified for scenarios S1, S2, and S3, respectively, given the sharp decrease in $sindex$ from that value of $k$.
+Thus, the number of clusters corresponds to the number of designs that were originally considered in each scenario.
+
+- Figure \ref{fig:mds-plot-validation} shows separation of our clusters.
+It can be observed that in all scenarios and for different mean differences, clusters are separated.
 However, the separation increases with an increase in mean differences across scenarios.
@@
 The lines are not colored by group since the purpose is to highlight the contribution of the factors to categorization rather than class separation.
-The first plot shows that no variable stands out in the clustering, but the following two designs show that {g1} and {g1, g2} have very low inter cluster distances, meaning that they did not contribute to the clustering.
-It is worth noting that these facts correspond to our original assumptions when developing the scenarios, which incorporate distributional differences over three (a), two (b), and one (c) significant granularities.
-Hence, Figure \ref{fig:parcoord-sim} (a), (b), and (c) validate the construction of scenarios (a), (b), and (c) respectively.
+Panel S1 shows that no variable stands out in the clustering, but the following two panels show that {g1} and {g1, g2} have very low inter-cluster distances, meaning that they did not contribute to the clustering.
+It is worth noting that these facts correspond to our original assumptions when developing the scenarios, which incorporate distributional differences over three (S1), two (S2), and one (S3) significant granularities.
+Hence, Figure \ref{fig:parcoord-sim} (S1), (S2), and (S3) validate the construction of scenarios (S1), (S2), and (S3) respectively.
 
- - The js-robust and wpd methods perform worse for $nT=300$, then improve for higher $nT$ evaluated in the study.
-Although, a complete year of data is the minimum requirement to capture distributional differences in winter and summer profiles, for example.
+ - The JS-RS and wpd-based methods perform worse for $nT=300$, then improve for higher $nT$ evaluated in the study.
+However, a complete year of data is the minimum requirement to capture distributional differences in winter and summer profiles, for example.
 Even if the data is only available for a month, $nT$ with half-hourly data is expected to be at least $1000$.
-As a result, as long as the performance is promising for higher $nT=300$, this is not a challenge.
+As a result, as long as the performance is promising for higher $nT$, this is not a challenge.
 
-- In our study sample, the method js-nqt outperforms the method js-robust for smaller differences between categories. More testing, however, is required to corroborate this.
+- In our study sample, the method JS-NQT outperforms the method JS-RS for smaller differences between categories. More testing, however, is required to corroborate this.
 
-For more detailed results, please refer to the supplementary paper.
-The code for creating these data designs and running the methodologies is available at (https://github.com/Sayani07/paper-gracsR/Validation).
-
-```{r parcoord-sim, fig.cap = "The parallel coordinate plot illustrates the total inter-cluster distances for granularities $g1$, $g2$, and $g3$.
-One line in the figure depicts the inter-cluster distances for a single simulation scenario.
-While the first plot indicates that no variable stands out during clustering, the next two designs demonstrate that {g1} and {g1, g2} have extremely low inter-cluster distances, indicating that they did not contribute to clustering.
-It is worth emphasising that these facts are consistent with our initial assumptions when designing the scenarios and (a), (b), and (c) correspond to Scenario (a), (b) and (c) respectively."}
+```{r sindex-data-validation}
 
@@
 
-```{r mds-plot-validation, fig.cap = "Relative positions of clusters corresponding to different scenarios (columns) for different values of mean differences between categories (rows) are shown using the first two dimensions of MDS.
-It can be observed that clusters become more compact and separated for higher mean differences between categories across all designs.
-Between designs, separation is least prominent corresponding to scenario (c) where only granularity is responsible for the clusters."}
+```{r sindex-plot-validation, fig.cap = "The cluster separation index (sindex) is plotted as a function of the number of clusters for the range of mean differences (rows) under the different scenarios (columns).
+S1 has a sharp decrease in sindex from 5 to 6, whereas S2 and S3 have a decrease from 4 to 5.
+As a result, the number of clusters considered for S1, S2, and S3 is 5, 4, 4, respectively.
+This corresponds to the number of designs taken into account in each scenario."}
+```
+
+```{r mds-data-validation}
+
+```
+
+```{r mds-plot-validation, fig.cap = "MDS summary plots to illustrate the cluster sparation for the range of mean differences (rows) under the different scenarios (columns).
+It can be observed that clusters become more compact and separated for higher mean differences between categories across all scenarios.
+Between scenarios, separation is least prominent corresponding to Scenario (S3) where only one granularity is responsible for distinguishing the clusters."}
+
+```
+
+```{r parcoord-sim, fig.cap = "The parallel coordinate figure depicts total inter-cluster distances for the granularities $g1$, $g2$, and $g3$.
+The inter-cluster distances for a single simulation scenario are represented by one line in the figure.
+While panel S1 shows that no variable stands out during clustering, panels S2 and S3 show that g1 and (g1, g2) had smaller inter-cluster distances, indicating that they did not contribute to clustering.
+These facts are consistent with our initial assumptions when building the scenarios, and S1, S2, S3 correspond to scenarios with three, two, and one significant granularity, respectively."}
 
@@
 (c) demonstrates that intercluster distances are only significant for the variable $g3$, meaning that the clusters formed are primarily due to $g3$.
-This corresponds to our design in Scenario (c), in which distributional differences were implemented solely for g3.
+This corresponds to our design in Scenario (S3), in which distributional differences were implemented solely for g3.
 (b) demonstrates the role of g2 and g3 in clustering.
@@
 
-## Case study {#sec:cases}
+## Application {#sec:application-gracsr}
 
 The use of our methodology is illustrated on smart meter energy usage for a sample of customers from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) which was available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) and Data61 CSIRO.
-It contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.
-In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual.
+It contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.<!--In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual.
 We do not learn about these repetitive behaviors from the linear view because too many measurements all squeezed in that representation.
-Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior.
-The raw data for these consumers is of unequal length, with varying start and finish dates.
+Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior.--> The raw data for these consumers is of unequal length, with varying start and finish dates.
 Because our proposed methods evaluate probability distributions rather than raw data, neither of these data features would pose any threat to our methodology unless they contained any structure or systematic patterns.
 Additionally, there were missing values in the database but further investigation revealed that there is no structure in the missingness (see Supplementary paper for raw data features and missingness).
-The study begins by subsetting a data set along all dimensions of interest using data filtering and prototyping.
-By grouping the prototypes using our methods and assessing their meaning, the study hopes to unravel some of the heterogeneities observed in energy usage data.
+
+Clustering huge data sets presents a slew of difficulties.
+As a result, we dissect the larger problem and test our solutions on a small sample of prototype customers.
+To do this, data is first filtered to generate a clean sample, and then significant cyclic granularities (variables) for them are chosen.
+This is described in Section \ref{sec:datafilter}.
+The clean set is subsequently subsampled along all dimensions of interest, to ensure that the sampled data reveals some patterns across at least one specified variable, which is described in Section \ref{sec:prototype}.
+By grouping the prototypes using our methods in Section \ref{sec:clustering} and assessing their meaning, the study hopes to unravel some of the heterogeneities observed in energy usage data.
 Because our application does not employ additional customer data, we cannot explain why consumption varies, but rather try to identify how it varies.
 
-_Data filtering and variable selection_
+### Data filtering and variable selection {#sec:datafilter}
 
@@
 - Obtain $wpd$ for all cyclic granularities considered for these customers.
-It was found that `hod` (hour-of-day), `moy` (month-of-year) and `wkndwday` (weekend/weekday) are coming out to be significant for most customers.
+It was found that `hod` (hour-of-day), `moy` (month-of-year) and `wkndwday` (weekend/weekday) are turning out to be significant for most customers.
 We use these three granularities while clustering.
@@
 
-_Prototype selection_
+### Prototype selection {#sec:prototype}
 
@@
 This is analogous to subsampling the population along all dimensions of interest such that the sampled data represents the primary features of the underlying distribution.
-Instance selection in unsupervised learning has received little attention in the literature, yet it could be a useful tool for evaluating model or method performance.
+Instance selection in unsupervised learning has received little attention in the literature.
+However, it could be a useful tool for evaluating model or method performance.
 There are several ways to approach prototype selection.
-Following @Fan2021-bq's idea of picking related examples (neighbors) for each instance (anchor), we can first use any dimensionality reduction techniques like MDS or PCA to project the data into a 2D space.
-Then pick a few "anchor" customers who are far apart in 2D space and pick a few neighbors for each.
+Following the idea of @Fan2021-bq of picking related examples (neighbors) for each instance (anchor), we can first use any dimensionality reduction techniques like MDS or PCA to project the data into a 2D space.
+Then pick a few "anchor" customers far apart in 2D space and pick a few neighbors for each.
 Unfortunately, this does not ensure that consumers with significant patterns across all variables are chosen.
-Tours can reveal variable separation that was hidden in a single variable display better than static projections.
-Hence we perform a linked tour with a t-SNE layout using the R package `liminal` (@R-liminal) to identify customers who are more likely to have distinct patterns across the variables studied.
-(Refer to Supplementary article for further details).
-Figure \ref{fig:hod-ind-group} shows the distribution across `hod` (a), `moy`(b) and `wnwd` (c) for the set of chosen $24$ customers that were chosen.
-Few of these customers have similar distribution across `moy` and some are similar in their `hod` distribution.
+Tours can reveal variable separation that gets hidden in a single variable display better than static projections.
+Hence, we perform a linked tour with a t-SNE layout using the R package `liminal` (@R-liminal) to identify customers who are more likely to have distinct patterns across the variables studied.
+(Refer to the supplementary article for further details).
+Figure \ref{fig:hod-ind-group-png} shows the distribution across `hod`, `moy` and `wnwd` for the set of chosen $24$ customers that were chosen.
+The customers are split into batches of $12$ in (a) and (b).
+Each row in (a) and that in (b) represents the profile of one customer across different variables.
+<!--Few of these customers have similar distribution across `moy` and some are similar in their `hod` distribution.
+-->
 
@@
 
-```{r hod-moy-wkndwday plots, fig.cap="The distribution across moy, hod and wkndwday for the selected designs.
+<!-- # ```{r hod-moy-wkndwday plots, fig.cap="The distribution across moy, hod and wkndwday for the selected designs.
 Few are similar in their hod pattern, while others are similar in moy behavior.
-Some customers have distinct behavior as compared to all other customers.For example, although patterns across wkndwday do not look distinctly different for most households, there is one household for whom weekend behavior is standing out from the rest."}
-```
+Some customers have distinct behavior as compared to all other customers.For example, although patterns across wkndwday do not look distinctly different for most households, there is one household for whom weekend behavior is standing out from the rest."} -->
+<!-- # ``` -->
 
-### Clustering
+### Clustering {#sec:clustering}
 
 The $24$ prototypes <!--and the data subset consisting of $353$ customers--> are clustered using the methodology described in Section \ref{sec:methodology} and results are reported below.
-In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$.
-All customers with the same color represent the same clustered groups.
-Groups by JS-based distances and wpd-based distances are colored differently as they represent different groupings.
-The plotting scales are not displayed since we want to emphasize comparable shapes rather than scales.
+In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$ percentiles.<!--All customers with the same color represent the same clustered groups.
+--> Groups are colored differently whenever they represent different groupings.
+While plotting the clusters, RS-based scaling is applied to each customer, so that no customer dominates the shape of the summarized clusters because of magnitude.
+The plotting scales are not displayed since we want to emphasize comparable patterns rather than scales.
 The idea is that a customer in a cluster may have low total energy usage, but their behavior may be quite similar to a customer with high usage with respect to distributional pattern or significance across cyclic granularities.
@@
 
-For clustering based on JS-based distances, we chose the optimal number of clusters using (@Hennig2014-ah) as $5$.
-The groupings are shown in Figure \ref{fig:hod-ind-group}.
-Customers with the same color represent the same clustered groups.
-Our methodology is useful for grouping similar distributions over `hod` and `moy` and they are placed closely for easy comparison.
+The $sindex$ is plotted against the number of clusters for clustering based on JS-based (JS-NQT) distances in Figure \ref{fig:opt-cluster-tsne-jsd}.
+The $sindex$ falls from around $0.56$ when $k=3$ to $0.54$ when $k$ increases.
+In our study, we use $k=4, 5$.
+Figure \ref{fig:groups-4and5} depicts the summarized distributions across $4$ and  $5$ clusters in (a) and (b) respectively, and assists in characterizing each cluster.
+A-2, Groups A-3, and A-4 have profiles that correspond to B-2, B-3, and B-5, respectively.
+B-2 (id: 4-9 in Figure \ref{fig:hod-ind-group-png}) and B-5 (id: 21-24) have a `hod` pattern with a typical morning and evening peak, whereas the other groups have no distinct pattern or high variability throughout the day.
+A-1 is subdivided into B-1 and B-4, each of which has a distinct shape across `moy` and `wnwd`.
+B-4 (id: 16-20) is distinguished by higher energy consumption in the first few months of the year and greater variability in weekend usage.
+B-1 (id: 1-3) has higher consumption in the middle of the year (winter months) and a similar weekday-weekend inter-quartile range.
+When $k = 4$ is used, these two groups merge to form A-1, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals.
+It may be worthwhile to compare Figures \ref{fig:opt-cluster-tsne-jsd} and \ref{fig:hod-ind-group-png} to see if the summarized distributions across groups accurately characterized the groupings.
+If it has, then the majority of the members of the group should have a similar profile.
+
+<!-- Our methodology is useful for grouping similar distributions over `hod` and `moy` and they are placed closely for easy comparison.
 Few groups have mixed patterns across `hod` and `moy`, but few have all customers in the group having a similar profile.
-Figure \ref{fig:combined-groups-js} shows the summarized distributions across $5$ groups and assists us in characterizing each cluster.
-It shows Groups $2$ and $4$ have `hod` pattern with a typical morning and evening peak, whereas groups $1$, $3$, and $5$ show a `moy` pattern with higher usage in winter months.
-Differences in Weekend/Weekday between groups are not discernible, implying that it may not be a relevant variable in distinguishing various clusters unless maybe conditioned by `moy` or `hod`.
-It may be interesting to compare these two plots to verify if the summarized distributions across groups correctly characterized the groupings.
-If it has, then the majority of the group's members should share a similar profile.
+-->
 
@@
 
-```{r opt-clusters}
-```
-
 ```{r groups-24}
@@
 
-```{r hod-ind-group, fig.cap = "The distribution of selected customers over hod (a), moy (b), and wnwd (c) for the $24$ selected customers is shown.
-In each case, the same colour denotes the same group using js-based clustering and are placed together to facilitate comparison.
-Individuals with similar distributional differences across categories of `hod` or `moy` are grouped together, with some exceptions.
-" }
+```{r data-pick-robust}
+
+```
+
+```{r hod-moy-wkndwday-data}
+
+```
+
+```{r opt-clusters-jsd}
+
+```
+
+```{r data-pick-wpd}
+```
+
+```{r tsne-fit}
+```
+
+```{r tsne-df}
+```
+
+```{r tsne-xy}
+
+```
+
+```{r opt-cluster-tsne-jsd, fig.cap = "(a) shows t-SNE summary of the $24$ selected customers chosen through prototype selection, (b) and (c) shows a plot of sindex for different cluster size for clustering based on JS-based distances and wpd-distances respectively.
+The plot suggests optimal number of clusters as $4$ and $7$ respectively."}
+#include_graphics("img/opt_clust_parenthesis.png")
+```
+
+```{r hod-ind-group}
+```
+
+```{r hod-ind-group-png, fig.cap = "The distribution of electricity demand across individual customers over multiple granularities hod, moy, and wnwd are shown for the $24$ selected customers using quantile and box plots.
+They are split into batches of $12$ in (a) and (b), with each row in (a) or (b) representing a customer.
+Each customer is identified by their profile across all the granularities.
+A few of these customers have similar distributions across moy and some are similar in their hod distributions.
+Distributional differences across categories of wnwd are not discernable except for a couple of customers (id: 16 and 18)."}
 ```
@@
 
-```{r data-heatmap-hod-group}
+```{r data-heatmap-hod-group-new}
 ```
 
-```{r data-heatmap-moy-group}
+```{r data-heatmap-moy-group-new}
 
@@
 
-```{r combined-groups-js, fig.cap = "The distribution of electricity demand over hod (a), moy (b), and wnwd*hod (c).
-Groups 2 and 5 appear to have a hod pattern among its members, whereas groups 1, 3, and 5 appear to have a moy pattern.
-Weekend/weekday differences across groups are not discernible, showing that it is not a critical variable for clustering on its own.
-It is probable that the distribution over Weekend/Weekday will differ conditional on the categories of hod or moy.
-(c) demonstrates that the distribution is not different, except for the tails.
-It is useful to compare the summarised distributions of groups to those of individuals in order to establish that the majority of individuals in the group share the same characteristics."}
+```{r combined-groups-js, fig.cap = "The distribution of electricity demand across groups over hod (a), moy (b), and hod (conditional on wnwd) (c) are shown through quantile plots.
+Group 1 (1-3), Group 2 (4-9), Group 3 (10 - 15), Group 4 (16- 20), Group 5 (21 - 24).
+Groups 2 and 5 appear to have a hod pattern among its members (morning and evening peak), whereas groups 1, 3, and 5 appear to have a moy pattern (higher usage in mid-year).
+With their conditional dependencies on hod or moy categories, wnwd differences may not be discernible on their own.
+(c) demonstrates that the distribution of hod conditional on wnwd is also not different, except for the tails.
+This might be because most of these chosen customers do not have a wnwd pattern.
+It is useful to compare the raw distributions with the summarized distributions to establish that the majority of individuals in the group share the same characteristics.
+To facilitate this comparison, the individual distributions are sorted by their JS-based groupings.", eval = FALSE}
+
+```
+
+<!-- The widest within-cluster gap (widest_gap) is high till $k=5$ and then have a gradual decrease. Number of -->
+
+```{r opt-clusters, fig.cap=" Different cluster statistics plotted as a function of number of clusters for JS-based (a) and wpd-based (b) approaches.
+avg within is not too useful for revealing any sharp change in average distance between clusters with increasing number of clusters.
+For (a), sindex sees a sharp fall at $k = 4$ and then do not decrease until $k=16$ and the widest gap is high till $k=5$.
+For (b), sindex sees a fall at $k=3$ and then at $k=7$ and widest gap has a sharp fall at $k=6$.", eval = FALSE}
+#knitr::include_graphics("img/opt_clusters.png")
+```
+
+<!-- GroupA and Group B may represent different set of customers as they are groupings based on different number of clusters. -->
+
+```{r groups-4and5, fig.cap = "For $k = 4$ (a) and $k=5$ (b), the distribution of electricity demand across groups over hod, moy, and wnwd is shown.
+Groups A-2, A-3, and A-4 profiles correspond to Groups B-2, B-3, and B-5, respectively.
+A-1 is split into B-1 and B-4, each of which has a distinct shape across moy and wnwd.
+B-4 (id: 16-20) is characterized by higher energy consumption in the first few months of the year and more variation in weekend usage.
+B-1 (id:1-3) is distinguished by higher consumption in the middle of the year (winter months) and similar weekday-weekend inter-quartile range.
+When $k = 4$ is used, these two groups merge to form A-1, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals in the group.", fig.height=6}
+
 ```
@@
 
-We chose the optimal number of clusters using (@Hennig2014-ah) as $3$.
-A parallel coordinate plot with the three significant cyclic granularities is used to characterize the groups here.
+In Figure \ref{fig:opt-cluster-tsne-jsd}, the $sindex$ is plotted against the number of clusters for clustering based on wpd-based distances.
+The scales are very different from those used in JS-based clustering.
+The $sindex$ falls from around $20$ when $k=3$ to $19$ when $k=4$ and further decreases to $17.5$ with $k>7$.
+Thus, there are few choices available for the number of clusters.
+The groupings with $k=4$ is shown through a parallel coordinate plot with the three significant cyclic granularities.
 The variables are sorted according to their separation across classes (rather than their overall variation between classes).
-This means that $moy$ is the most important variable in distinguishing the groups followed by $hod$ and $wnwd$.
-There is only one customer who has significant $wpd$ across $wnwd$ and stands out from the rest of the customers.
-Group 3 has a higher wpd for hod than moy or wkndwday.
-Group 2 has the most distinct pattern across moy.
-Group 1 is a mixed group that has strong patterns on at least one of the three variables.
-The findings vary from js-based clustering, yet it is a helpful grouping.
+This means that $moy$ is the most important variable in distinguishing the groups, followed by $hod$ and $wnwd$.
+The two customers (id: 16 and 18) that have significant differences between $wnwd$ categories (in Figure \ref{fig:hod-ind-group-png}) stand out from the rest of the customers.
+Group-4 has a higher $wpd$ for `hod` than `moy` or `wkndwday`.
+Group-2 and Group-3 have the most distinct patterns across `moy`.
+Group-1 is a mixed group that has strong patterns on at least one of the three variables.
+When $k=3$ is used, Group-2 and Group-3 are merged because their patterns relative to `moy`, `hod` and `wnwd` are the same.
+The findings vary from JS-based clustering, yet it is a helpful grouping.
 
-```{r data-pick-wpd}
+```{r wpd-clustering}
+
 ```
 
-```{r parcoord, fig.cap="Each of the 24 customers is represented by a parallel coordinate plot (a) with three wpd-based groupings.
-The plot shows that moy is the most important variable in identifying clusters, whereas wkdn-wday is the least significant and has the least fluctuation.
-One particular customer with high wpd across wknwday stands out in this display.
-Group 3 has a higher wpd for hod than moy or wkndwday.
-Group 2 has most discernible pattern across moy.
-Group 1 is a mixed group with strong patterns on atleast one of the three variables.
-All of these could be observed from the plot or the table (b) which shows median wpd values for each group.", message=FALSE, warning=FALSE}
-
+```{r parcoord-application-gracsr, fig.cap="Each of the $24$ customers is represented by a parallel coordinate plot (a) with four wpd-based groupings.
+Plot (b) which shows median $wpd$ values for each group.
+The plot shows that moy is the most important variable in identifying clusters, whereas wnwd is the least significant and has the least fluctuation.
+Two customers (id: 16 and 18) with high $wpd$ across wnwd stand out in this display.
+Group-4 has a higher $wpd$ for hod than moy or wnwd.
+Group-2 and Group-3 have the most discernible pattern across moy.
+Group-1 is a mixed group with strong patterns across at least one of the three variables.
+All of these could be observed from the plot or the table (b).", message=FALSE, warning=FALSE, fig.height=6}
+#include_graphics("img/parcoord-application.png")
 ```
@@
 Furthermore, combining heterogeneous customers may result in similar-looking final clusters that are not effective for visually differentiating them.
-It is also worth noting that the Weekend/Weekday behavior in the given case does not characterize any cluster.
+It is also worth noting that the weekend/weekday behavior in the given case does not characterize most selected customers.
 This, however, will not be true for all of the customers in the data set.
@@
 
-## Conclusion and future work {#sec:conclusion}
+## Discussion {#sec:discussion-gracsr}
 
@@
 
-There are few areas to extend this research.
+There are a few areas to extend this research.
 First, larger data sets with more uncertainty complicate matters, as is true for any clustering task.
@@
 Thirdly, it is possible that for a few customers, data for some categories from the list of considered significant granularities are missing.
-In our application, we have removed those customers and done the analysis but the metrics used should be able to incorporate those customers in the clustering by handling their missing categories.
+In our application, we have removed those customers and done the analysis but the metrics used should be able to incorporate those customers with such structured missingness.
 Finally, $wpd$ is computationally heavy even under parallel computation.
@@
 
-## Acknowledgments {- #thanksgracsr}
+## Acknowledgments {-}
 
@@
 The Github repository, [github.com/Sayani07/paper-gracsr](https://github.com/Sayani07/paper-gracsr), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
-This article was created with R [@R-language], `knitr` [@knitr2015], [@R-knitr] and `rmarkdown` [@rmarkdown], [@R-rmarkdown].
-Graphics are produced with `ggplot2` [@Wickham2009pk] and GGally [@R-GGally].
+This article was created with R [@R-language], `knitr` [@knitr; @R-knitr] and `rmarkdown` [@rmarkdown; @R-rmarkdown].
+Graphics are produced with `ggplot2` [@Wickham2009pk] and `GGally` [@R-GGally].
 
-## Supplementary Materials {- #supplement-gracsr}
+## Supplementary Materials
 
@@
 
-\noindent
-**Supplementary paper:** Additional tables, graphics and and R code to reproduce it (paper-supplementary.pdf, paper-supplementary.Rmd)
+**Supplementary paper:** Additional tables, graphics and R code to reproduce it (paper-supplementary.pdf, paper-supplementary.Rmd).
+The code for creating validation designs and running the methodologies is available at (https://github.com/Sayani07/paper-gracsR/Validation).
 
-\noindent
 **R-package:** To implement the ideas provided in this research, the open-source R package 'gracsr' is available on Github (https://github.com/Sayani07/gracsr).

### 2021-11-15 scripts/gracsr/application.R

@@ -1,0 +1,841 @@
+##### application
+
+## ----prototype-data-pick-------------------------------------------------------
+quantile_prob_graph <- c(0.25, 0.5, 0.75)
+
+data_pick_one <- c(8541744, 9355808, 8603880, 8619309, 10542667) %>%
+  as_tibble() %>%
+  set_names("customer_id")
+# data_pick_two <- c(8688242, 8643837, 8184707, 10534355, 8684420) %>% as_tibble%>% set_names("customer_id")
+data_pick_three <- c(9792072, 8589936, 8454235, 10692366, 8603828) %>%
+  as_tibble() %>%
+  set_names("customer_id")
+data_pick_four <- c(8618759, 8291696, 10357256, 8290374) %>%
+  as_tibble() %>%
+  set_names("customer_id")
+data_pick_five <- c(9044864, 8642053, 10534367, 9021526, 11162275) %>%
+  as_tibble() %>%
+  set_names("customer_id")
+data_pick_six <- c(8221762, 8273636, 10359424, 8232822, 11450499) %>%
+  as_tibble() %>%
+  set_names("customer_id")
+
+
+
+## ----assemble------------------------------------------------------------------
+data_pick_cust <- bind_rows(
+  data_pick_one,
+  # data_pick_two,
+  data_pick_three,
+  data_pick_four,
+  data_pick_five,
+  data_pick_six,
+  .id = "design"
+) %>%
+  mutate(customer_id = as.character(customer_id))
+
+## ----all-data-----------------------------------------------------------------
+data_pick <- read_rds(here::here("data/gracsr/elec_nogap_2013_clean_356cust.rds")) %>%
+  mutate(customer_id = as.character(customer_id)) %>%
+  dplyr::filter(customer_id %in% data_pick_cust$customer_id) %>%
+  gracsr::scale_gran(
+    method = "nqt",
+    response = "general_supply_kwh"
+  )
+
+
+
+## ----clustering---------------------------------------------------------------
+hod <- suppressMessages(data_pick %>%
+  dist_gran(
+    gran1 = "hour_day",
+    response = "general_supply_kwh"
+  ))
+
+moy <- suppressMessages(data_pick %>%
+  dist_gran(
+    gran1 = "month_year",
[…]

### 2021-11-15 scripts/gracsr/mylibrary.R

@@ -1,0 +1,34 @@
+## ---- load-lib
+
+#theme_set(theme_bw())
+library(ggplot2)
+library(lubridate)
+library(gravitas)
+library(gracsr)
+library(ggdendro)
+library(dplyr)
+library(readr)
+library(visdat)
+library(ggplot2)
+library(tidyverse)
+library(naniar)
+library(here)
+library(tsibble)
+library(knitr)
+library(patchwork)
+library(GGally)
+library(distributional)
+library(viridis)
+library(forcats)
+library(tidyr)
+library(purrr)
+library(stringr)
+library(pals)
+library(Rtsne)
+library(fpc)
+library(cluster)
+library(ggpubr)
+library(grid)
+library(ggrepel)
+
+

### 2021-11-15 scripts/gracsr/mytheme.R

@@ -1,0 +1,98 @@
+##----theme-validation--------------------------------------------------------------------
+theme_validation <- function() {
+  theme_bw() +
+    theme(
+      strip.text = element_text(size = 8, 
+                                margin = margin(b = 0, t = 1)),
+      plot.margin = margin(0, 0, 0, 0, "cm"),
+      axis.title.y = element_blank(),
+      #axis.text.x = element_blank(),
+      axis.text.y = element_blank(),
+      axis.ticks = element_blank(),
+      legend.position = "bottom"
+    )
+}
+
+
+##----theme_characterisation--------------------------------------------------------------
+
+theme_characterisation <- function() {
+  
+  theme_bw() + # seeting theme
+    theme(strip.text = element_text(size = 10,
+                                    margin = margin(b = 0, t = 0))) +
+    # narrow facet space
+    theme(axis.title.y=element_blank(),
+          axis.text.y=element_blank(),
+          axis.ticks.y=element_blank()) + # no axis ticks 
+    theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
+    theme(axis.text.x = element_text(angle=90, hjust=1, size = 10)) + 
+    # rotate the x-axis text
+    theme(legend.position = "bottom")+
+    theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
+    theme(axis.text.x = element_text(size=5)) +
+    theme(axis.text = element_text(size = 12))  
+}
+
+
+##----mytheme-application----------------------------------------------------------------
+theme_application <- function() {
+  
+  theme_light() + # setting theme
+    theme(strip.text = element_text(size = 10,
+                                    margin = margin(b = 0, t = 0)), 
+          strip.text.x = element_text(size = 8)) + # narrow facet space
+    theme(axis.title.y=element_blank(),
+          axis.text.y=element_blank(),
+          axis.ticks.y=element_blank()) + # no axis ticks 
+    theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
+    theme(axis.ticks.x = element_blank(),
+          axis.text.x = element_blank()) + # no x-axis labels to further reduce the gap between facets
+    #theme(axis.text.x = element_text(angle=90, hjust=1, size = 9)) + # rotate the x-axis text
+    theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
+    #theme(axis.text.x = element_text(size=5)) +
+    theme(strip.background = element_blank(),
[…]

### 2021-11-15 scripts/gracsr/supplementary.R

@@ -1,0 +1,875 @@
+## ----raw-data-50
+
+knitr::include_graphics("figs/raw_plot_cust.png") # look at smart-meter.R for the code
+
+
+## ----missing-data
+
+knitr::include_graphics("figs/missing-data-5050.png")
+
+
+## ----mds
+
+
+knitr::include_graphics("figs/tour-tsne.png")
+
+
+
+## ----sim-val
+
+set.seed(9999)
+nx_val <- 2 # number of x-axis levels
+nfacet_val <- 3 # number of facet levels
+w1_val <- 3 # increment in mean
+w2_val <- 0 # increment in sd
+mean_val <- 0 # mean of normal distribution of starting combination
+sd_val <- 1 # sd of normal distribution of starting combination
+quantile_prob_val <- seq(0.1, 0.9, 0.1)
+
+## ----nobs50
+
+ntimes_val <- 300
+
+### ----sample-seed-few
+
+sample_seed <- seq(10, 100, 10)
+
+
+#### ----change-index-data-few
+
+change_index_data <- function(ntimes_val = NULL,
+                              nx_val = NULL,
+                              nfacet_val = NULL,
+                              sim_function = sim_varx_normal) {
+  data <- sim_panel(
+    nx = nx_val, nfacet = nfacet_val,
+    ntimes = ntimes_val,
+    # sim_dist = sim_varx_normal(2, 3, 5, 10, 5, -1.5)
+    sim_dist = sim_function(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
+  ) %>% unnest(data)
+
+
+  index_new <- map(seq_len(ntimes_val), function(i) {
+    map((seq_len(nx_val * nfacet_val)), function(j) {
+      value <- i + (j - 1) * ntimes_val
+    })
+  }) %>% unlist()
+
+  data_new <- data %>%
+    ungroup() %>%
+    mutate(
+      index_old = row_number(),
+      index_new = index_new
+    )
+
+  y <- data_new[match(index_new, data_new$index_old), ]
+
+  y <- y %>%
+    mutate(time = row_number()) %>%
+    select(-c(index_old, index_new))
+
+  return(y)
+}
+
+## ---- data-make-few
+
+data_null <- map(sample_seed, function(seed) {
+  set.seed(seed)
[…]

### 2021-11-15 scripts/gracsr/validation.R

@@ -1,0 +1,1027 @@
+## Validation
+
+## ----tab-distribution-------------------------------------------------------------------
+granularity <- c("g1", "g2", "g3")
+alternate_design <- c(
+  "g10 ~ N(0, 1), g11 ~ N(2, 1)",
+  "g21 ~ N(2, 1), g22 ~ N(1, 1), g23 ~ N(0, 1)",
+  "g31 ~ N(0, 1), g32 ~ N(1, 1), g33 ~ N(2, 1),
+                      g34 ~ N(1, 1), g35 ~ N(0, 1)"
+)
+
+tab_dist <- tibble::tibble(
+  granularity = granularity,
+  `Varying distributions` = alternate_design
+)
+
+
+
+## ----tab-design-------------------------------------------------------------------------
+design <- c("design-1", "design-2", "design-3", "design-4", "design-5")
+g1 <- c("fixed", "vary", "fixed", "fixed", "vary")
+g2 <- c("fixed", "fixed", "vary", "fixed", "vary")
+g3 <- c("fixed", "fixed", "fixed", "vary", "vary")
+table <- tibble(design, g1, g2, g3)
+
+
+
+## ----tab-dist-design--------------------------------------------------------------------
+
+knitr::kable(
+  list(tab_dist, tibble(design, g1, g2, g3)),
+  caption = "For S1, distributions of different categories when they vary (displayed on top). If distributions are fixed, they are set to N(0, 1). The various distributions across categories result in five designs (displayed below).",
+  booktabs = TRUE, valign = "t"
+)
+
+
+
+## ----generate-design-3change------------------------------------------------------------
+generate_design <- function(t, mu1, mu2, mu3) {
+  t <- seq(0, t, 1)
+  g1 <- t %% 2
+  g2 <- t %% 3
+  g3 <- t %% 5
+
+  # null design
+  g1_dnull <- rep(rep(0, each = length(unique(g1))), length.out = length(t))
+  g2_dnull <- rep(rep(0, each = length(unique(g2))), length.out = length(t))
+  g3_dnull <- rep(rep(0, each = length(unique(g3))), length.out = length(t))
+
+  # mean changing across categories in varying ways
+
[…]

### 2021-11-15 Rmd/00-d-acknowlegements.Rmd
ch: to ch-
addressed Rob's all comments + acknowledgement add

@@
-# Acknowledgements {-#ch:thanks}
+# Acknowledgements {-#ch-thanks}
 
@@
 I would also like to express my gratitude to my supervisor Dr.
-Peter Toscas of Data61 CSIRO, for always being available to help with technical issues and offering feedback on practical elements of analyzing smart meter data at our monthly meetings.
+Peter Toscas of Data61 CSIRO, for always being available to help and offering feedback on practical elements of analyzing smart meter data at our monthly meetings.
 
@@
 
-PhD in Covid has been hard, and I would not have been able to make it through this if not for the emotional support of my friends and family.
+PhD in COVID has been hard, and I would not have been able to make it through this if not for the emotional support of my friends and family.
 Thank you Puwasala, Sium, and Ian, for being so kind and acting as a pillar of emotional support, especially during the last leg of my journey.
@@
 Thank you to each one of you for being available, listening to me, and putting things in perspective when I lost sight.
-Thanks to my housemates Anjali, Surbhi, and Dulaji for the fun company and conversations that kept me sane in the COVID lockdowns.
+Thanks to my housemates Anjali, Surbhi, and Dulaji for the fun company, food and conversations that kept me sane in the COVID lockdowns.
 
@@
 
-Finally, a big thank you to my family for always being supportive of my choices.
+A big thank you to my family for always being supportive of my choices.
 Thanks to my mum (Nupur Gupta) and dad (Arun Prasad Gupta), from whom I learnt that no matter where you start from, if you persevere and are sincere in your efforts, you can sail through any difficult situation.
@@
 Thank you, Juhi (sister-in-law), for reminding me of the value of organization in all aspects of life and the importance of prioritizing my physical health from time to time.
-Thank you all for your part in supporting me in this journey.
+Thank you for making me feel at home away from home, Somdatta (cousin).
+Lastly, thank you to my one-year-old nephew (Rayan) and niece (Mehr) for being the ultimate stress relievers.
+Thank you all for being a part of this journey with me.
 I love you.
@@
 <!-- add people who are helping you with proofreading -->
-<!-- Add Peter -->
 <!-- Although the journey is coming to an end, you wish to take these life learnings forward -->

### 2021-11-15 Rmd/04-gracsr.Rmd
addressed Rob's all comments + acknowledgement add

@@
 <!-- time series clustering and its challenges -->
-Time-series clustering is the process of unsupervised partitioning of $n$ time-series data into $k$ ($k<n$) meaningful groups such that homogeneous time-series are grouped together based on a certain similarity measure.
-The time-series features, length of time-series, representation technique, and, of course, the purpose of clustering time-series all influence the suitable similarity measure or distance metric to a meaningful level.
+Time series clustering is the process of unsupervised partitioning of $n$ time series data into $k$ ($k<n$) meaningful groups such that homogeneous time series are grouped together based on a certain similarity measure.
+The time series features, length of time series, representation technique, and, of course, the purpose of clustering time series all influence the suitable similarity measure or distance metric to a meaningful level.
 The three primary methods to time series clustering (@liao2005clustering) are algorithms that operate directly with distances or raw data points in the time or frequency domain (distance-based), with features derived from raw data (feature-based), or indirectly with models constructed from raw data (model-based).
@@
 Only one study [@ozawa2016determining] identified time series characteristics by first conducting a Fourier transformation, to convert data from time to frequency domain, followed by $k$-means to cluster by greatest frequency.
-@Motlagh2019-yj suggests that the time feature extraction is limited by the type of noisy, patchy, and unequal time-series common in residential customers and addresses model-based clustering by transforming the series into other objects such as structure or set of parameters which can be more easily characterized and clustered.
+@Motlagh2019-yj suggests that the time feature extraction is limited by the type of noisy, patchy, and unequal time series common in residential customers and addresses model-based clustering by transforming the series into other objects such as structure or set of parameters which can be more easily characterized and clustered.
 [@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
@@
 <!-- This work -->
-<!-- This is similar to a stochastic approach (@Motlagh2019-yj) to clustering, which proposes interpreting electricity demand as a random process and extracting time-series characteristics, or a model of the series, to enable unsupervised clustering.
+<!-- This is similar to a stochastic approach (@Motlagh2019-yj) to clustering, which proposes interpreting electricity demand as a random process and extracting time series characteristics, or a model of the series, to enable unsupervised clustering.
 Unsupervised clustering is only as good as the features that are extracted/selected or the distance metrics that were utilized.
@@
 
-<!--The foundation of our method is unsupervised clustering algorithms based exclusively on the time-series data.
+<!--The foundation of our method is unsupervised clustering algorithms based exclusively on the time series data.
 The proposed methods leverage the intrinsic data structure hidden within cyclic temporal granularities.--> The existing work on clustering probability distributions assumes we have an independent and identically distributed samples $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.

### 2021-11-15 Rmd/03-hakear.Rmd
ch: to ch-
proofread paper-hakear done

@@
-# Detecting distributional differences between temporal granularities for exploratory time series analysis  {#ch:hakear}
+# Detecting distributional differences between temporal granularities for exploratory time series analysis  {#ch-hakear}
 
@@
 This work proposes a new distance metric to screen and rank the possible granularities, and hence choose the most interesting ones to plot.
-The distance measure is computed for a single or pairs of cyclic granularities can can be compared across different cyclic granularities and also on a collection of time series.
+The distance measure is computed for a single or pairs of cyclic granularities that can be compared across different cyclic granularities and also on a collection of time series.
 The methods are implemented in the open-source R package `hakear`.
@@
 Daily peaks occur in morning and evening hours, indicating a working household, where members leave for and return from work.
-More volatility of usage in summer months (Jan, Feb) is probably due air conditioner use on just some days.", fig.pos="!b"}
+More volatility of usage in summer months (Jan, Feb) is probably due to air conditioner use on just some days.", fig.pos="!b"}
 ```
@@
 The categorical variable (granularity) is mapped to the x-axis, and the distribution of the response variable is displayed using both side-by-side jittered dotplots and boxplots.
-From panel (a) it can be seen that energy consumption is higher during the morning hours (5--8), when members in the household wake up, and again in the evening hours (17--20) possibly when members get back from work.
+From panel (a) it can be seen that energy consumption is higher during the morning hours (5--8), when members in the household wake up, and again in the evening hours (17--20), possibly when members get back from work.
 In addition, the largest variation in energy use is in the afternoon hours (12--16), as seen in the length of the boxes.
@@
 The black line is the median, whereas the orange band covers the 25th to 75th percentile and the green band covers the 10th to 90th percentile.
-Difference between the 90th and 75th quantiles is less for (Jan, Feb) for the first household (a), suggesting that it is a more frequent user of air conditioner than the second household (b).
-Energy consumption for in (a) changes across both granularities, whereas for (b) daily pattern stays same irrespective of the months."}
+Difference between the 90th and 75th quantiles is less for (Jan, Feb) for the first household (a), suggesting that it is a more frequent user of air conditioners than the second household (b).
+Energy consumption for (a) changes across both granularities, whereas for (b) daily pattern stays same irrespective of the months."}
 ```
@@
 Therefore, even for small $v$, the number of scatterplots can be large, and scatterplot matrices (SPLOMs) can easily run out of pixels when presenting high-dimensional data.
-@Dang2014-tw and @wilkinson2005graph provide potential solutions to this, where a few characterizations can be used to locate anomalies in density, shape, trend, and other features in the 2D point scatters.
+@Dang2014-tw; @wilkinson2005graph provide potential solutions to this, where a few characterizations can be used to locate anomalies in density, shape, trend, and other features in the 2D point scatters.
 
@@
 
-@inference and @Majumder2013-hb present methods for statistical significance testing of visual findings using human cognition as the statistical tests.
+@inference; @Majumder2013-hb present methods for statistical significance testing of visual findings using human cognition as the statistical tests.
 In this paper, the visual discovery of distributional differences is facilitated by choosing a threshold for the proposed numerical distance measure, eventually selecting only those cyclic granularities for which the distributional differences are sufficient to make it an interesting display.
@@
 The Jensen-Shannon divergence [@Menendez1997-in] based on the Kullback-Leibler divergence is symmetric and has a finite value.
-Hence, in this paper, the pairwise distances between the distributions of the measured variable are obtained through the square root of the Jensen-Shannon divergence, called Jensen-Shannon distance (JSD), and defined by,
+Hence, in this paper, the pairwise distances between the distributions of the measured variable are obtained through the square root of the Jensen-Shannon divergence, called Jensen-Shannon distance (JSD), and defined by
 $$
@@
 When categories are ordered, distances should only be considered for consecutive x-axis categories.
-Between-facet distances are distances between different facet levels for the same x-axis category, for example, distances between {($a_1$,$b_1$) and ($a_1$, $b_2$)} or {($a_1$,$b_1$) and ($a_1$, $b_3$)}.", fig.pos="!b"}
+Between-facet distances are distances between different facet levels for the same x-axis category; for example, distances between {($a_1$,$b_1$) and ($a_1$, $b_2$)} or {($a_1$,$b_1$) and ($a_1$, $b_3$)}.", fig.pos="!b"}
 
@@
 Specifying $\lambda_x>0.5$ will weight within-facet distances more heavily, while $\lambda_x <0.5$ would weight the between-facet distances more heavily.
-(See the supplements for more details.).
+(See the supplements for more details.)
 
@@
 That is, under designs like $\Dnull$, their distribution should not differ for a different number of categories.
-Only then the distance measure could be compared across panels with different levels.
+Only then could the distance measure be compared across panels with different levels.
 This calls for an adjusted measure, which normalizes for the different number of comparisons.
@@
 This method accomplishes a different goal of finding the null distribution for different groups (panels in our case) and standardizing the raw values using that distribution.
-The values of $\wpdsub{raw}$ is computed on many ($\nsub{perm}$) permuted data sets and stored in $\wpdsub{perm-data}$.
+The values of $\wpdsub{raw}$ are computed on many ($\nsub{perm}$) permuted data sets and stored in $\wpdsub{perm-data}$.
 Then $\wpdsub{perm}$ is computed as follows:
 \begin{align*}
-  \wpdsub{perm} = \frac{(\wpdsub{raw} - \text{mean}(\wpdsub{perm-data}))}{\text{sd}(\wpdsub{perm-data})},
+  \wpdsub{perm} = \frac{(\wpdsub{raw} - \text{mean}(\wpdsub{perm-data}))}{\text{sd}(\wpdsub{perm-data})}
 \end{align*}
@@
 
-(ref:dist-new-same-scale-link) The distribution of $\wpdsub{perm}$ and $\wpdsub{glm-scaled}$ are overlaid to compare the location and scale across different $\nx$ and $\nf$.
+(ref:dist-new-same-scale-link) The distributions of $\wpdsub{perm}$ and $\wpdsub{glm-scaled}$ are overlaid to compare the location and scale across different $\nx$ and $\nf$.
 $\wpdsub{norm}$ takes the value of $\wpdsub{perm}$ for lower levels, and $\wpdsub{glm-scaled}$ for higher levels to to alleviate the problem of computational time in permutation approaches.
@@
 
-The smart meter data set for eight households in Melbourne were procured by downloading the data from the energy supplier/retailer.
+The smart meter data set for eight households in Melbourne was procured by downloading the data from the energy supplier/retailer.
 The data has been cleaned to form a `tsibble` [@wang2020tsibble] containing half-hourly electricity consumption from July to December 2019 for each of the households.
@@
 Let $v_{i, t}$ denote the electricity demand for the $i^{th}$ household in time period $t$.
-The series $\{v_{i, 1},\dots,v_{i,T}\}$ is the linear granularity corresponding to half-hour since the interval of the tsibble is 30 minutes.
+The series $\{v_{i, 1},\dots,v_{i,T}\}$ is the linear granularity corresponding to half-hour since the interval of the `tsibble` is 30 minutes.
 We consider coarser linear granularities like hour, day, week and month from the commonly used Gregorian calendar.
@@
 The pairs of cyclic granularities ($C_{N_C}$) will have $7\times 6=42$ elements.
-The set of possible harmonies $H_{N_C}$ from $C_{N_C}$ are chosen by removing clashes using procedures described in [@Gupta2021-gravitas].
+The set of possible harmonies $H_{N_C}$ from $C_{N_C}$ are chosen by removing clashes using procedures described in @Gupta2021-gravitas.
 Table \ref{tab:tab-rank-8} shows $14$ harmony pairs that belong to $H_{N_C}$.
@@
 
-#### Selecting and Ranking harmonies for all households {-}
+#### Selecting and ranking harmonies for all households {-}
 
@@
   2.
-id 6 and 7 differ in the sense that for id 6, the difference in patterns only during weekday/weekends, whereas for id 7, all or few other days of the week are also important.
+id 6 and 7 differ in the sense that for id 6, the difference in patterns is only during weekday/weekends, whereas for id 7 all or few other days of the week are also important.
 This might be due to their flexible work routines or different day-off.
@@
 According to Figure \ref{fig:dotplot-8}(c), for the harmony pair (*dow*-*hod*), household id 7 has the greatest value of $\wpd$, while id 1 has the least.
-From Table \ref{tab:tab-rank-8} it can be seen that the harmony pair (*dow*, *hod*) is important for id 7, however it has been labeled as an inconsequential pair for id 1.
+From Table \ref{tab:tab-rank-8} it can be seen that the harmony pair (*dow*, *hod*) is important for id 7; however, it has been labeled as an inconsequential pair for id 1.
 The distribution of energy demand for both of these households, with *dow* as the facet and *hod* on the x-axis, may help explain the choice.
@@
 A huge number of displays need to be analyzed or we might end up missing informative granularities.
-This work refines the search of informative granularities by identifying those for which the differences between the displayed distributions are greatest and rating them in order of importance of capturing maximum variation.
+This work refines the search for informative granularities by identifying those for which the differences between the displayed distributions are greatest and rating them in order of importance of capturing maximum variation.
 
@@
 The Github repository, [github.com/Sayani07/paper-hakear](https://github.com/Sayani07/paper-hakear), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
-This article was created with R [@R-language], `knitr` [@knitr], [@R-knitr] and `rmarkdown` [@rmarkdown], [@R-rmarkdown].
+This article was created with R [@R-language], `knitr` [@knitr; @R-knitr] and `rmarkdown` [@rmarkdown; @R-rmarkdown].
 Graphics are produced with `ggplot2` [@Wickham2009pk].
@@
 \noindent
-**Simulation results and scripts:** All simulation table, graphics and and R code to reproduce it (paper-supplementary.pdf, paper-supplementary.Rmd)
+**Supplementary paper:** All simulation tables, graphics and and R codes to reproduce the supplementary paper (paper-supplementary.Rmd, paper-supplementary.pdf).
 

### 2021-11-15 scripts/hakear.R

@@ -1067,7 +1067,7 @@ elec_rank <- elec_sig_split %>%
 
 elec_rank %>% 
   kable(format = "markdown", 
-        caption = "Ranking of harmonies for the eight households with significance marked for different thresholds. Rankings are different and at most three harmonies are significant for any household. The number of harmonies to explore are reduced from 42 to 3.")
+        caption = "Ranking of harmonies for the eight households with significance marked for different thresholds. Rankings are different and at most three harmonies are significant for any household. The number of harmonies to explore is reduced from 42 to 3.")
 
 
 

### 2021-11-15 Rmd/00-d-acknowlegements.Rmd
paper-gracsr done

@@
  <!-- - PhD friends and R community -->
-Thank you, Stephanie Kobakian and Nicholas Tierney, for being so warm, kind and welcoming when I began my PhD.
+Thank you, Puwasala Gamakumara, Stephanie Kobakian and Nicholas Tierney, for being so warm, kind and welcoming when I began my PhD.
 Being a part of such an inclusive environment meant a lot to me as an international student.

### 2021-11-15 Rmd/02-gravitas.Rmd
paper-gracsr done

@@
 Although we were motivated by the smart meter example, the problem and the solutions we propose are practically relevant to any temporal data observed more than once per year.
-In a broader sense, it could be even suitable for data observed by years, decades, and centuries as might be in weather or astronomical data.
+In a broader sense, it could be even suitable for data observed by years, decades, and centuries as might be the case in weather or astronomical data.
 
-This work provides tools for systematically exploring bivariate granularities within the tidy workflow (@wickham2016r). In particular, we
+This work provides tools for systematically exploring bivariate granularities within the tidy workflow [@wickham2016r]. In particular, we
 
@@
 Hour, day, week, month and year are linear granularities and can also be considered to be time domains.
-These are ordered with ordering guided by integers and hence is unidirectional and non-repeating.
+These are ordered with ordering guided by integers and hence are unidirectional and non-repeating.
 Hours could also be considered the index set, and a bottom granularity.
@@
 Then, $G_1 \preceq H_1$, but $G_1 \ntrianglelefteq H_1$.
-Further, with $G_2$ denoting "days" and $H_2$ denoting "business-week", $G_2 \npreceq H_2$, but $G_2 \trianglelefteq H_2$, since each business-week can be expressed as an union of some days, but Saturdays and Sundays are not subset of any business-week.
+Further, with $G_2$ denoting "days" and $H_2$ denoting "business-week", $G_2 \npreceq H_2$, but $G_2 \trianglelefteq H_2$, since each business-week can be expressed as an union of some days, but Saturdays and Sundays are not subsets of any business-week.
 Moreover, with  $H_3$ denoting "public holidays", $G_1 \npreceq H_3$ and $G_1 \ntrianglelefteq H_3$.
@@
 
-If $G$ groups into $H$, it would imply that any granule $H(i)$ is the union of some granules of $G$, for example, $G(a_1), G(a_2), \dots, G(a_k)$.
+If $G$ groups into $H$, it would imply that any granule $H(i)$ is the union of some granules of $G$; for example, $G(a_1), G(a_2), \dots, G(a_k)$.
 Condition (2) in Definition \ref{def:periodic} implies that if $H(i + R) \neq \emptyset$, then $H(i + R) = \bigcup (G(a_1 + P), G(a_2 + P), \dots, G(a_k + P))$, resulting in a "periodic" pattern of the composition of $H$ using granules of $G$.
 In this pattern, each granule of $H$ is shifted by $P$ granules of $G$.
-$P$ is called the \textbf{Period} (@Bettini2000-qk).
+$P$ is called the \textbf{period} (@Bettini2000-qk).
 
@@
 
-With two linear granularities $G$ and $H$, if $G$ *groups into* or *finer than* $H$ then $G$ is of lower order than $H$.
-For example, if the bottom granularity is hour, then granularity $day$ will have lower order than $week$ since each day consist of fewer hours than each week.
+With two linear granularities $G$ and $H$, if $G$ *groups into* or is *finer than* $H$ then $G$ is of lower order than $H$.
+For example, if the bottom granularity is hour, then granularity $day$ will have lower order than $week$ since each day consists of fewer hours than each week.
 
@@
 \begin{table}[ht]
+\caption{Examples of circular granularities with bottom granularity minutes.
+Circular granularity $C_i$ relates two linear granularities one of which groups periodically into the other with regular mapping and period $P_i$.
+Circular granularities can be expressed using modular arithmetic due to their regular mapping.}
 \begin{center}
@@
 \end{center}
-\caption{Examples of circular granularities with bottom granularity minutes.
-Circular granularity $C_i$ relates two linear granularities one of which groups periodically into the other with regular mapping and period $P_i$.
-Circular granularities can be expressed using modular arithmetic due to their regular mapping.}
 \label{tab:definitions}
@@
 \begin{table}[ht]
+\caption{Examples of quasi-circular granularities relating two linear granularities with irregular mapping leading to several possible period lengths.}
 \centering
@@
 \end{tabular}
-\caption{Examples of quasi-circular granularities relating two linear granularities with irregular mapping leading to several possible period lengths.}
 \label{tab:quasi}
@@
 Examples include public holidays which repeat every year, but there is no reasonably small span of time within which their behavior remains constant.
-A classic example is Easter (in the western tradition) whose dates repeat only after 5.7 million years [@Reingold2001-kf].
+A classic example is Easter (in the Western tradition) whose dates repeat only after 5.7 million years [@Reingold2001-kf].
 In Australia, if a standard public holiday falls on a weekend, a substitute public holiday will sometimes be observed on the first non-weekend day (usually Monday) after the weekend.
@@
 The linear display shows the granularities days ($B$), weeks ($G$), semester days ($B'$), and stages of a semester ($M$) indexed over a linear representation of time.
-The granules of $B'$ is only defined for days when the semester is running.
+The granules of $B'$ are only defined for days when the semester is running.
 Here a semester spans 18 weeks and 2 days, and consists of 6 stages.
@@
 
-, where $z$ is the index of the bottom granularity $kin$.
+where $z$ is the index of the bottom granularity $kin$.
 
@@
 \end{equation}
-where, $T_w$ is such that $G_{m'}(w) = \bigcup_{z \in T_w}G_{\ell}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
+where $T_w$ is such that $G_{m'}(w) = \bigcup_{z \in T_w}G_{\ell}$ and $\vert T_w \vert$ is the cardinality of set $T_w$.
 
@@
 This is consistent with the widely used grammar of graphics
-which is a framework to construct statistical graphics by relating the data space to the graphic space [@Wilkinson1999-nk], [@Wickham2009pk].
+which is a framework to construct statistical graphics by relating the data space to the graphic space [@Wilkinson1999-nk; @Wickham2009pk].
 
@@
 For displaying the distribution of a continuous univariate variable, many options are available.
-Displays based on descriptive statistics include box plots [@Tukey1977-jx] and its variants such as notched box plots [@McGill1978-hg] or other variations as mentioned in @boxplots.
+Displays based on descriptive statistics include boxplots [@Tukey1977-jx] and its variants such as notched boxplots [@McGill1978-hg] or other variations as mentioned in @boxplots.
 They also include line or area quantile plots which can display any quantiles and not only quartiles like in a boxplot.
-Plots based on kernel density estimates include violin plots [@Hintze1998-zi], summary plot [@Potter2010-qc], ridge line plots [@R-ggridges], and highest density region (HDR) plots [@Hyndman1996-ft].
-The less commonly used Letter-Value plots [@Hofmann2017-sg] is midway between boxplots and density plots.
-Letter values are order statistics with specific depths, for example, the median ($M$) is a letter value that divides the data set into halves.
+Plots based on kernel density estimates include violin plots [@Hintze1998-zi], summary plots [@Potter2010-qc], ridge line plots [@R-ggridges], and highest density region (HDR) plots [@Hyndman1996-ft].
+The less commonly used letter-value plots [@Hofmann2017-sg] is midway between boxplots and density plots.
+Letter values are order statistics with specific depths; for example, the median ($M$) is a letter value that divides the data set into halves.
 Each of the next letter values splits the remaining parts into two separate regions so that the fourths ($F$), eighths ($E$), sixteenths ($D$), etc.
@@
 For example, arrangement in **a** makes it easier to compare across weekday type (x-axis) within a quarter (facet).
-It can be seen that in quarter 2, there is more mass occupied the lower tail on the weekends (letter value E corresponding to tail area 1/8) relative to that of the weekdays (letter value D 1/16), which corresponds to more days with lower energy use in this period.
+It can be seen that in quarter 2, more mass occupied the lower tail on the weekends (letter value E corresponding to tail area 1/8) relative to that of the weekdays (letter value D 1/16), which corresponds to more days with lower energy use in this period.
 
@@
 
-Another question that can be addressed is if good fielding or bowling (defending) in the previous over affects the scoring rate in the subsequent over?
+Another question that can be addressed is if good fielding or bowling (defending) in the previous over affects the scoring rate in the subsequent over.
 To measure the defending quality, we use an indicator function on dismissals (1 if there was at least one wicket in the previous over, 0 otherwise).
@@
 This seems to be unaffected by the over of the innings (the faceting variable).
-This might be because the new batsman needs to "play himself in" or the dismissals lead the (not-dismissed) batsman to adopt a more defensive playstyle.
+This might be because the new batsman needs to "play himself in" or the dismissals lead the (not-dismissed) batsman to adopt a more defensive play style.
 Run rates will also vary depending on which player is facing the next over and when the wicket falls in the previous over.
@@
 
-(ref:cricex) Examining distribution of runs per by innings, over of the innings and number of wickets in previous innings.
+(ref:cricex) Examining distribution of runs per innings, overs of the innings and number of wickets in previous innings.
 Plot **a** displays distribution using letter value plots.
@@
 A missing piece in the package `gravitas` is the computation of cyclic aperiodic granularities which would require computing aperiodic linear granularities first.
-A few R packages including `almanac`[@R-almanac] and `gs`[@R-gs] provide the tools to create recurring aperiodic events.
+A few R packages including `almanac` [@R-almanac] and `gs` [@R-gs] provide the tools to create recurring aperiodic events.
 These functions can be used with the `gravitas` package to accommodate aperiodic cyclic granularities.
@@
 The Australian authors thank the ARC Centre of Excellence for Mathematical and Statistical Frontiers [(ACEMS)](https://acems.org.au/home) for supporting this research.
-Thanks to [Data61 CSIRO](https://data61.csiro.au/) for partially funding Sayani's research and Dr Peter Toscas for providing useful inputs on improving the analysis of the smart meter application.
+Thanks to [Data61 CSIRO](https://data61.csiro.au/) for partially funding Sayani's research and Dr.
+Peter Toscas for providing useful inputs on improving the analysis of the smart meter application.
 We would also like to thank Nicholas Spyrison for many useful discussions, sketching figures and feedback on the manuscript.
@@
 The Github repository,  [github.com/Sayani07/paper-gravitas](https://github.com/Sayani07/paper-gravitas), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
-This article was created with `knitr` [@knitr], [@R-knitr] and `rmarkdown` [@rmarkdown], [@R-rmarkdown].
+This article was created with `knitr` [@knitr; @R-knitr] and `rmarkdown` [@rmarkdown; @R-rmarkdown].
 
@@
 **R-package:** The ideas presented in this article have been implemented in the open-source R [@R-language] package `gravitas` [@R-gravitas], available from CRAN.
-The R-package facilitates manipulation of single and multiple-order-up time granularities through cyclic calendar algebra, checks feasibility of creating plots or drawing inferences for any two cyclic granularities by providing list of harmonies and recommends possible visual summaries through factors described in the article.
+The R-package facilitates manipulation of single and multiple-order-up time granularities through cyclic calendar algebra, checks feasibility of creating plots or drawing inferences for any two cyclic granularities by providing list of harmonies, and recommends possible visual summaries through factors described in the article.
 Version 0.1.3 of the package was used for the results presented in the article and is available on Github (https://github.com/Sayani07/gravitas).

### 2021-11-16 Rmd/01-intro.Rmd
intro minor rephrasing

@@
 The Smart Grid, Smart City (SGSC) project (2010–2014) available through the [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) provides half-hourly data of over 13,000 Australian electricity smart meter customers distributed unevenly from October 2011 to March 2014.
-Large data sets like this include significant uncertainty regarding customer behavior due to a wide variety of customers.
+The wide variety of customers means that there will be large variance in behavior, leading to greater uncertainty in the data.
 Behavioral patterns vary significantly due to differences in size, location, and amenities such as solar panels, central heating, and air conditioning.
@@
 Many, many displays might be built using cyclic granularities.
-However, only a handful of them may be interesting, revealing major patterns.
-Determining which displays exhibit "significant" distributional differences between cyclic granularity categories and plotting only these would allow for more efficient exploration.
+However, only a handful of them may reveal major patterns of interest.
+Identifying the displays which exhibit "significant" distributional differences and plotting only these would allow for more efficient exploration.
 Furthermore, a few of the displays in this collection will be more engaging than others.
@@
 Because they are based on probability distributions, these distances are resistant to missing or noisy data and aid in dimension reduction.
-When fed into a clustering algorithm, the distances can be used to divide large data sets into small pockets of similar repetitive behavior.
+When fed into a clustering algorithm, the distances can be used to divide large data sets into small pockets of similar repetitive behaviors.
 These subgroups can then be analyzed separately or used as distinct prototype behaviors in classification problems.
@@
 This is in the developing R package `gracsr`.
-Chapter \@ref(ch-conclusion) summarizes the software tools developed for the work, and discusses some future plans.
+Chapter \@ref(ch-conclusion) summarizes the software tools developed for the work, and discusses future plans.

### 2021-11-16 Rmd/00-b-abstract.Rmd
Update 00-b-abstract.Rmd

@@
 The research was motivated by the desire to understand some Australian smart meter data, which was collected half-hourly for two years at the household level.
-With temporal data available at ever finer scales, exploring periodicity can become overwhelming with so many possible temporal deconstructions to explore.
+As temporal data is made available at ever finer scales, exploring periodicity can become overwhelming with many possible temporal deconstructions to explore.
 Analysts are expected to comprehensively explore the many ways to view and consider temporal data.
@@
 The first contribution (Chapter \@ref(ch-gravitas)) describes classes of time deconstructions using linear and cyclic time granularities.
-It provides tools to compute possible cyclic granularities from an ordered (usually temporal) index and also a framework to systematically explore the distribution of a univariate variable conditional on two cyclic time granularities by defining "harmony".
+It provides tools to compute possible cyclic granularities from an ordered (usually temporal) index and also a framework to systematically explore the distribution of a univariate variable conditional on two cyclic time granularities by defining "harmonies".
 A "harmony" denotes pairs of granularities that could be effectively analyzed together and reduces the search from all possible options.

### 2021-11-16 Rmd/02-gravitas.Rmd
Update 02-gravitas.Rmd

@@
 
-In \autoref{fig:bothcust}b, the black line is the median, whereas the purple (narrow) band covers the 25th to 75th percentile, the orange (middle) band covers the 10th to 90th percentile, and the green (broad) band covers the 1st to 99th percentile.
+In \autoref{fig:bothcust}b, the black line is the median, the purple (narrow) band covers the 25th to 75th percentile, the orange (middle) band covers the 10th to 90th percentile, and the green (broad) band covers the 1st to 99th percentile.
 The first facet represents the weekday behavior while the second facet displays the weekend behavior; energy consumption across each hour of the day is shown inside each facet.

### 2021-11-17 Rmd/01-intro.Rmd

@@
 These discrete abstractions of time are known as time granularities.
-Linear time granularities [@Bettini1998-ed] ,such as hours, days, weeks and months, respect the linear progression of time and are non-repeating.
+Linear time granularities [@Bettini1998-ed], such as hours, days, weeks and months, respect the linear progression of time and are non-repeating.
 Cyclic temporal granularities representing cyclical repetitions in time (such as hour-of-the-day, work-day/weekend) are effective for analyzing repetitive patterns in time series data.
@@
 Chapter \@ref(ch-hakear) is a natural extension of Chapter \@ref(ch-gravitas).
-Many, many displays might be built using cyclic granularities.
+Many displays might be built using cyclic granularities.
 However, only a handful of them may reveal major patterns of interest.
@@
 The statistical significance of potential visual discoveries is aided by selecting a threshold for the proposed numerical distance measure.
-The distance measure is computed for a single or pairs of cyclic granularities, and it can be compared across different cyclic granularities as well as a collection of time series.
-This chapter also includes a case study using residential smart meter data from Melbourne to demonstrate how the suggested methodology may be utilized to automatically find temporal granularities with significant distributional differences.
+The distance measure is computed for a single, or pairs of, cyclic granularities, and it can be compared across different cyclic granularities as well as a collection of time series.
+This chapter also includes a case study using residential smart meter data from household in Melbourne, Australia to demonstrate how the suggested methodology may be utilized to automatically find temporal granularities with significant distributional differences.
 The methods are implemented in the open-source R package `hakear`.
@@
 When fed into a clustering algorithm, the distances can be used to divide large data sets into small pockets of similar repetitive behaviors.
-These subgroups can then be analyzed separately or used as distinct prototype behaviors in classification problems.
+These smaller groups can then be analyzed separately or used as distinct prototype behaviors in classification problems.
 The proposed method was tested on a group of residential electricity consumers from the Australian smart meter data set to show that it can generate meaningful clusters.

### 2021-11-17 Rmd/04-gracsr.Rmd
dunn index reference update

@@
 It is often beneficial to investigate numerous cluster statistics in order to arrive at a number of clusters, since clustering has a variety of objectives (e.g., between-cluster separation, within-cluster homogeneity, representation by centroids, etc.), which could be in conflict with one another.
-Some typical ways of determining the number of clusters in the literature (e.g., the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@dunn1973fuzzy]) are inspired by defining and balancing "within-cluster homogeneity" and "between-cluster separation".
+Some typical ways of determining the number of clusters in the literature (e.g., the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@Dunn1973-gk]) are inspired by defining and balancing "within-cluster homogeneity" and "between-cluster separation".
 The Cluster separation index ($sindex$) proposed by @Hennig2014-ah is used as a guide to select the number of clusters in Section \ref{sec:validation}.

### 2021-11-17 Rmd/03-hakear.Rmd
Update 03-hakear.Rmd

@@
 The categorical variable (granularity) is mapped to the x-axis, and the distribution of the response variable is displayed using both side-by-side jittered dotplots and boxplots.
-From panel (a) it can be seen that energy consumption is higher during the morning hours (5--8), when members in the household wake up, and again in the evening hours (17--20), possibly when members get back from work.
+From Figure \ref{fig:onegran-new} (a) it can be seen that energy consumption is higher during the morning hours (5--8), when members in the household wake up, and again in the evening hours (17--20), possibly when members get back from work.
 In addition, the largest variation in energy use is in the afternoon hours (12--16), as seen in the length of the boxes.
-From panel (b), it is seen that the variability in energy usage is higher in Jan and Feb, probably due to the usage of air conditioners on some days.
+From panel Figure \ref{fig:onegran-new} (b), it is seen that the variability in energy usage is higher in Jan and Feb, probably due to the usage of air conditioners on some days.
 The median usage is highest in January, dips in February--April and rises again in May--June, although not to the height of January usage.
@@
 
-In this paper, we provide a solution to narrowing down the search from $N_C(N_C-1)$ conditional distribution plots by introducing a new distance measure that can be used to detect significant distributional differences across cyclic granularities.<!-- harmonies and why it is not enough --> This work is a natural extension of our previous work [@Gupta2021-gravitas] which narrows down the search from $N_C(N_C-1)$ plots by identifying pairs of granularities that can be meaningfully examined together (a "harmony"), or when they cannot (a "clash").
+In this paper, we provide a solution to narrowing down the search from $N_C(N_C-1)$ conditional distribution plots by introducing a new distance measure that can be used to detect significant distributional differences across cyclic granularities.<!-- harmonies and why it is not enough --> This work is a natural extension of our previous work [@Gupta2021-gravitas] which narrows down the search from $N_C(N_C-1)$ plots by identifying "harmonies", pairs of granularities that can be meaningfully examined together, or "clashes" when they cannot.
 However, even after excluding clashes, the list of harmonies left may be too large for exhaustive exploration.
@@
 Let the number of cyclic granularities considered in the display be $m$.
-The notations and methodology are described in detail for $m=2$.
-But it can be easily extended to $m>2$.
+The notations and methodology are described in detail for $m=2$, but this can be easily extended to $m>2$.
 Consider two cyclic granularities $A$ and $B$, such that $A = \{a_j: j = 1, 2, \dots, \nx\}$ and $B = \{b_k: k = 1, 2, \dots, \nf\}$ with $A$ placed across the x-axis and $B$ across facets.

### 2021-11-17 Rmd/04-gracsr.Rmd
Update 04-gracsr.Rmd

@@
 
-Clustering is a potential approach for Organizing large collections of times series into small homogeneous groups, but a difficult step is determining an appropriate metric to measure similarity between time series.
+Clustering is a potential approach for organizing large collections of times series into small homogeneous groups.
+However, it is difficult to determine an appropriate metric to measure similarity between time series.
 The similarity metric needs to be capable of accommodating long, noisy, and asynchronous time series and also capture cyclical patterns.
@@
 The primary application of this work is data from the Smart Grid, Smart City (SGSC) project (2010–2014) available through the [Department of the Environment and Energy](https://data.gov.au/data/organization/doee).
-Half-hourly measurements of usage for more than 13,000 household electricity smart meters is provided from from October 2011 to March 2014.
+Half-hourly measurements of usage for more than 13,000 household electricity smart meters is provided from October 2011 to March 2014.
 <!--Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.--> Households vary in size, location, and amenities such as solar panels, central heating, and air conditioning.
 The behavioral patterns differ amongst customers due to many temporal dependencies.
-Some households use a dryer, while others dry their clothes on a line.
-Their weekly usage profile may reflect this.
-They may vary monthly, with some customers using more air conditioners or heaters than others, while having equivalent electrical equipment and weather circumstances.
+Some households dry their clothes on a line, others use a dryer.
+Their patterns may vary monthly, with some customers using more air conditioning or heating than others, while having equivalent electrical equipment and weather circumstances.
 Some customers are night owls, while others are morning larks.
-Daily energy usage varies depending on whether customers stay home or work away from home.
-Age, lifestyle, family composition, building attributes, weather, availability of diverse electrical equipment, among other factors, make the task of properly segmenting customers into comparable energy behavior complex.
-The challenge is to be able to cluster consumers into these type of expected patterns, and other unexpected patterns, using only their energy usage history (@Ushakova2020-rl).
+Daily energy usage varies depending on whether customers stay home or work out of the home.
+Age, lifestyle, family composition, building attributes, weather, availability of diverse electrical equipment, among other factors, contribute to the complex task of properly segmenting customers into groups with comparable energy behavior.
+The challenge is to be able to cluster consumers into these groups of expected patterns, and other unexpected patterns, using only their energy usage history (@Ushakova2020-rl).
 <!--To safeguard the customers' privacy, it is probable that such information is not accessible.
-Also, energy suppliers may not always update client information, such as property features, in a timely manner.--> There is a growing need to have methods that can examine the energy usage heterogeneity observed in smart meter data and what are some of the most common power consumption patterns.
+Also, energy suppliers may not always update client information, such as property features, in a timely manner.--> There is a growing need to have methods that can examine the energy usage heterogeneity observed in smart meter data and identify the most common power consumption patterns.
 
@@
 @Motlagh2019-yj suggests that the time feature extraction is limited by the type of noisy, patchy, and unequal time series common in residential customers and addresses model-based clustering by transforming the series into other objects such as structure or set of parameters which can be more easily characterized and clustered.
-[@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
+@chicco2010renyi addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
 @Melnykov2013-sp discusses how outliers, noisy observations and scattered observations can complicate estimating mixture model parameters and hence the partitions.
@@
 This paper is organized as follows.
-Section~\ref{sec:methodology} provides the clustering methodology.
-Section~\ref{sec:validation} shows data designs to validate our methods.
-Section~\ref{sec:application-gracsr} discusses the application of the method to a subset of the real data.
+Section \ref{sec:methodology} provides the clustering methodology.
+Section \ref{sec:validation} shows data designs to validate our methods.
+Section \ref{sec:application-gracsr} discusses the application of the method to a subset of the real data.
 Finally, we summarize our results and discuss possible future directions in Section~\ref{sec:discussion-gracsr}.
@@
 
-[@Gupta2021-hakear] provides a method for determining the significance of a cyclic granularity, and a ranking of multiple cyclic granularities.
+@Gupta2021-hakear provides a method for determining the significance of a cyclic granularity, and a ranking of multiple cyclic granularities.
 (This extends to harmonies, pairs of granularities that might interact with each other.) We define "significant" granularities as those with significant distributional differences across at least one category.
@@
 Skewed distributions need to be symmetrized.
-Scales of individuals needs to be standardized, because clustering is to select similar patterns, not magnitude of usage.
-(Organizing individuals based on magnitude can achieved simply by sorting on a statistic like the average value across time.) <!--Observations often have a somewhat skewed time series distribution and their ranges might vary greatly.
+Scales of individuals need to be standardized, because the goal of clustering is to select similar patterns, not magnitude of usage.
+Organizing individuals based on magnitude may be helpful, but can be achieved simply by sorting on a statistic like the average value across time.
+<!--Observations often have a somewhat skewed time series distribution and their ranges might vary greatly.
 Statistical transformations are employed to bring all of them to the same range or normalize each observation.
 This is important because we are not interested in trivial clusters that vary in magnitudes, but rather in uncovering comparable patterns of distributional differences between categories.--> For the JS-based approaches, two data transformation techniques are recommended, normal-quantile transform (NQT) and robust scaling (RS).
-NQT is used in @@Gupta2021-hakear prior sorting granularities.
+NQT is used in @Gupta2021-hakear prior sorting granularities.
 <!-- $wpd$, which is the foundation of wpd-based distances.
@@
 <!--As a result, determining which raw distribution was used is difficult using the modified distribution.--> A drawback is that any multimodality will be concealed.
-This, this should be checked prior to applying NQT.
+This should be checked prior to applying NQT.
 
@@
 
-The total (dis) similarity between each pair of customers is obtained by combining the distances between the collections of conditional distributions.
-This needs to be done in a way such that the resulting metric is a distance metric, and could be fed into the clustering algorithm.
+The total (dis)similarity between each pair of customers is obtained by combining the distances between the collections of conditional distributions.
+This needs to be done in a way such that the resulting metric is a distance metric, that can be fed into the clustering algorithm.
 Two types of distance metrics are considered:
@@
 
-Compute weighted pairwise distances $wpd$ (@Gupta2021-hakear) for all considered granularities for all observations.
+This metric computes weighted pairwise distances $wpd$ (@Gupta2021-hakear) for all considered granularities for all observations.
 $wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction and is estimated by the maximum pairwise distances between consecutive categories normalized by appropriate parameters.
@@
 Different approaches to determining number of clusters are required for varied applications and research objectives.
-It is often beneficial to investigate numerous cluster statistics in order to arrive at a number of clusters, since clustering has a variety of objectives (e.g., between-cluster separation, within-cluster homogeneity, representation by centroids, etc.), which could be in conflict with one another.
-Some typical ways of determining the number of clusters in the literature (e.g., the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@Dunn1973-gk]) are inspired by defining and balancing "within-cluster homogeneity" and "between-cluster separation".
+It is often beneficial to investigate numerous cluster statistics in order to arrive at a number of clusters, since clustering has a variety of objectives (e.g.
+between-cluster separation, within-cluster homogeneity, representation by centroids, etc.), which could be in conflict with one another.
+Some typical ways of determining the number of clusters in the literature are inspired by defining and balancing "within-cluster homogeneity" and "between-cluster separation" (e.g., the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@dunn1973fuzzy]).
 The Cluster separation index ($sindex$) proposed by @Hennig2014-ah is used as a guide to select the number of clusters in Section \ref{sec:validation}.
@@
 
-.<!--determined by calculating the distance between each observation and the cluster to which it does not belong.
+<!--determined by calculating the distance between each observation and the cluster to which it does not belong.
 Then, the least 10% of these distances are averaged to only evaluate places near the cluster edges.-->
@@
 So a mean difference of $\mu = \{1, 2, 5\}$ between categories is considered.
-The performance of the approaches varies with the number of granularities which has interesting patterns across its categories.
-So three scenarios are considered to accommodate that.
+The performance of the approaches varies with the number of granularities which has interesting patterns across its categories, so three scenarios are considered to accommodate that.
 Figure \ref{fig:range-parameter} shows the range of parameters considered for each scenario.

### 2021-11-17 Rmd/05-conclusion.Rmd
done with Rob's comments

@@
 
-This thesis presents a framework for visualizing and analyzing distributions of large temporal data by leveraging time characteristics.
-The building blocks of this framework are presented in each main chapter.
-This chapter summarizes the thesis content, covering original contributions, software development, and future ideas.
+This thesis presents methods for visualizing and analyzing distributions of large temporal data by deconstructing time into temporal granularities.
+This chapter summarizes the thesis content, outlines the original contributions, software development, and possible directions for future research.
 
@@
 They can be circular, quasi-circular, or aperiodic in nature.
-Hour of the day and day of the week are examples of circular granularities; the day of the month is an example of a quasi-circular granularity; and public holidays and school holidays are examples of aperiodic granularities.
+Hour-of-the-day and day-of-the-week are examples of circular granularities; the day-of-the-month is an example of a quasi-circular granularity; and public holidays and school holidays are examples of aperiodic granularities.
 Additionally, time deconstructions can be based on a time hierarchy.
@@
 
-This thesis focuses on integrating research approaches into open source R packages for reproducibility and ease of use in other applications.
+This thesis focuses on translating research approaches into open source R packages for reproducibility and ease of use in other applications.
 So a significant amount of work has been devoted to the development of R packages  `gravitas`, `hakear`, and `gracsr`, each of which corresponds to a chapter presented in this thesis.
@@
 The R package `hakear` (https://github.com/Sayani07/hakear) implements ideas presented in Chapter \@ref(ch-hakear).
-The function wpd() computes the weighted pairwise distances ($wpd$) for each cyclic granularity or pair of granularities, and `select_harmonies()` selects the ones with significant patterns and ranks them from highest to lowest $wpd$.
+The function `wpd()` computes the weighted pairwise distances ($wpd$) for each cyclic granularity or pair of granularities, and `select_harmonies()` selects the ones with significant patterns and ranks them from highest to lowest $wpd$.
 These selected harmonies can be plotted using package `gravitas` for potentially interesting displays.
@@
 It is possible that data for a whole category of cyclic granularity is unavailable or that there are insufficient observations to compute distributions.
-For example, a customer does not have data for a particular day of the week or month throughout their observation period.
+For example, a customer may not have data for a particular day of the week or month throughout their observation period.
 While visualizing probability distributions across categories in Chapters \@ref(ch-gravitas) and \@ref(ch-hakear), this can be indicated by displaying dot plots instead of summarizing distributions.
@@
 Moreover, integrating heterogeneous consumers may result in visually identical end clusters, which are potentially not useful.
-An appropriate anomaly detection approach could be applied to filter the anomalies, and then a decision has to be made regarding if the anomalies should form a group or be assigned to the closest group through a classification method.
+An appropriate anomaly detection approach could be applied to filter the anomalies, and then a decision has to be made whether the anomalies should form a group or be assigned to the closest group through a classification method.
 
@@
 
-Larger data sets with more uncertainty complicate matters, with potentially more problems than already stated above.
+Larger data sets with more uncertainty may complicate matters, with potentially more problems than already stated above.
 More research is needed to understand what other sorts of problems may arise with the current methodology and possible ways to address them.

### 2021-11-17 Rmd/05-conclusion.Rmd
Section 5.1, re-organise to group chapters
software rewrite using Di's suggestions

@@
 The definitions and rules defined in the literature for linear granularities are insufficient for describing various types of cyclic granularities.
+
 Chapter \@ref(ch-gravitas) provides a formal characterization of cyclic granularities as well as tools for classifying and computing potential cyclic granularities from an ordered temporal index.
@@
 The approach is generalizable to non-temporal hierarchical granularities with an ordered index.
-
 Visualizing probability distributions conditional on one or more cyclic granularities is a powerful exploration tool.
@@
 The `gravitas` package provides very general tools to compute and manipulate cyclic granularities and generate plots displaying distributions conditional on those granularities.
-The functions `search_gran()`, `create_gran()`, `harmony()`, `gran_advice()` and `prob_plot()` provides the entire workflow for an analyst to systematically explore large quantities of temporal data across different harmonies (pairs of granularities that can be analyzed together.) This package was developed as part of my internship at Google Summer of Code, 2019.
+The functions `search_gran()`, `create_gran()`, `harmony()`, `gran_advice()` and `prob_plot()` provides the entire workflow for an analyst to systematically explore large quantities of temporal data across different harmonies (pairs of granularities that can be analyzed together).
+This package was developed as part of my internship at Google Summer of Code, 2019.
 It has been on CRAN since January 2020.
@@
 
-The R package `hakear` (https://github.com/Sayani07/hakear) implements ideas presented in Chapter \@ref(ch-hakear).
-The function `wpd()` computes the weighted pairwise distances ($wpd$) for each cyclic granularity or pair of granularities, and `select_harmonies()` selects the ones with significant patterns and ranks them from highest to lowest $wpd$.
-These selected harmonies can be plotted using package `gravitas` for potentially interesting displays.
+The R package `hakear` (https://github.com/Sayani07/hakear) provides tools for selecting and sorting significant cyclic granularities.
+The function `wpd()` computes the weighted pairwise distances ($wpd$) for each cyclic granularity or pair of granularities, and the function `select_harmonies()` chooses those with significant patterns and ranks them from highest to lowest $wpd$.
 This package is reliant on parallel processing using multiple multi-core computers for faster computation of $wpd$.
+The selected harmonies can be plotted using package `gravitas` for potentially interesting displays.
+Currently, `hakear` implements ideas presented in Chapter \@ref(ch-hakear), but it will be integrated with `gravitas` in the future to explore distributions of a smaller number of time series.
 
@@
 
-The open-source R package `gracsr` is available on Github (https://github.com/Sayani07/gracsr) to implement ideas presented in Chapter \@ref(ch-gracsr).
-The package contains functions for implementing the complete clustering methodology with choices of scaling and distance metrics discussed in the chapter.
-It has received a grant (AUD 3000) as part of the ACEMS Business Analytics Prize towards polishing the functions and preparing it for CRAN.
+The R package `gracsr` (https://github.com/Sayani07/gracsr) has functions for exploring a large number of time series using the clustering methodology described in Chapter \@ref(ch-gracsr).
+The workflow begins with the function `scale gran()`, which may be used to scale individual series using NQT/RS.
+The distances for the JS and WPD approaches are computed in the second phase of the workflow using `dist gran()`/`dist wpd()`.
+The distances can then be used to do clustering with `clust gran()`.
+The package has received a grant (AUD 3000) as part of the ACEMS Business Analytics Prize towards polishing the functions and preparing it for CRAN.
+
+### Computational resources
+Simulation studies were carried out to study the behavior of $wpd$, build the normalization method as well as compare and evaluate different normalization approaches in \@ref(ch-hakear).
+In \@ref(ch-gracsr), our methods were tested on several data designs with different parameters to evaluate their performance.
+$wpd$ is computationally heavy for cyclic granularities with smaller levels.
+Moreover, JS and WPD approaches are also computationally intensive when run on large number of customers.
+Hence, most of the scripts used to run these studies use parallel processing for better computational speed.
+R version 4.0.1 (2020-06-06) is utilized on the platform x86 64-apple-darwin17.0 (64-bit) operating on macOS Mojave 10.14.6, as well as the High Performance Computing (HPC) resources provided by [MonARCH](https://docs.monarch.erc.monash.edu/MonARCH/aboutMonArch.html).
+
+<!-- The package contains functions for implementing the complete clustering methodology with choices of scaling and distance metrics discussed in the chapter. It -->
 

### 2021-11-17 Rmd/05-conclusion.Rmd
Di's comments on conclusion incorporated

@@
 
-<!-- ### Putting all functionalities on CRAN -->
-
-<!-- I want to combine `hakear` and `gravitas` into a single R package for systematically exploring distributions of a smaller number of time series, whilst `gracsr` would offer the clustering framework for systematically exploring a large number of time series.
-I intend to run it via https://ropensci.org/software-review/ to further refine them for more exposure and efficiency.
--->
-
-<!-- ### Computing aperiodic cyclic granularities -->
-
-<!-- Aperiodic cyclic granularities capture repetitions of aperiodic linear granularities.
-All of these are recurring events, but with non-periodic patterns which can still be visualized to understand repetitive behavior.
-The computation of cyclic aperiodic granularities requires first computing aperiodic linear granularities.
-Aperiodic events may be created using R packages such as `almanac`(Vaughan, 2020) and `gs`(Laird-Smith, 2020).
-The functions in `gravitas` package need to integrate with them to support computation of aperiodic cyclic granularities.
--->
+There are a few areas to extend this research.
+The first two explore catering to time series data, where the design of our framework is limited.
+Handling non-stationary time series data or data with structured missingness, for example.
+Second, scaling up for more uncertainty is beneficial since huge data sets with greater uncertainty may have more difficulties than a few time series.
+It is essential to recognize the difficulties of analyzing clustering results in such a setting.
+To that end, it would be interesting to investigate how to deal with anomalies or determine the significance of the clusters.
+A third possibility is to focus on expanding the scope of the framework to include multivariate time series data.
 

### 2021-11-17 scripts/gravitas.R
x axis labels overlap in paper-gravitas fixed

@@ -234,8 +234,8 @@ cust2_box <- smart_meter10 %>%
   ggtitle("") +
   scale_x_discrete(labels = c("wday", "wend")) +
   scale_y_sqrt() +
-  theme(axis.text.x = element_text(size = 7)) +
   theme_minimal() +
+  theme(axis.text.x = element_text(size = 7, angle=90))  +
   theme(panel.border = element_rect(colour = "#E0E0E0", fill = NA)
   )
 

### 2021-11-17 Rmd/05-conclusion.Rmd
conclusion proofread

@@
 ### Computational resources
-Simulation studies were carried out to study the behavior of $wpd$, build the normalization method as well as compare and evaluate different normalization approaches in \@ref(ch-hakear).
-In \@ref(ch-gracsr), our methods were tested on several data designs with different parameters to evaluate their performance.
+Simulation studies were carried out to study the behavior of $wpd$, build the normalization method as well as compare and evaluate different normalization approaches in Chapter \@ref(ch-hakear).
+In Chapter \@ref(ch-gracsr), our methods were tested on several data designs with different parameters to evaluate their performance.
 $wpd$ is computationally heavy for cyclic granularities with smaller levels.
-Moreover, JS and WPD approaches are also computationally intensive when run on large number of customers.
+JS and WPD approaches are also computationally intensive when run on large number of customers.
 Hence, most of the scripts used to run these studies use parallel processing for better computational speed.

### 2021-11-17 scripts/hakear.R
y label change to avoid overlap

@@ -70,7 +70,7 @@ p1 <- id2_tsibble %>%
   ggtitle("(a)") + theme(
     strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + 
   scale_colour_brewer(name = "", palette = "PiYG") + 
-  ylab ("energy consumption (in kwh)")
+  ylab ("energy usage [kWh]")
 
 
 p2 <- id4_tsibble %>%
@@ -82,7 +82,7 @@ p2 <- id4_tsibble %>%
             quantile_prob = c(0.1, 0.25,0.5,0.75, 0.9)) + 
   ggtitle("(b)") + theme(
     strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))+
-  ylab ("energy consumption (in kwh)")
+  ylab ("energy usage [kWh]")
 
 
 
@@ -157,7 +157,7 @@ id2_tsibble_hd <- elec %>%
   geom_boxplot(width = 0.6, outlier.colour = "black", outlier.alpha = 0.5, fill = "#CC79A7", colour =  "#0072B2") + scale_y_log10() +
   geom_jitter(alpha = 0.04, colour = "#E69F00") +
   ggtitle("(a)")+ 
-  ylab ("energy consumption (in kwh)")
+  ylab ("energy usage [kWh]")
 
 
 
@@ -169,7 +169,7 @@ id2_tsibble_dw <- elec %>%
   geom_boxplot(width = 0.6, outlier.colour = "black", outlier.alpha = 0.5, fill = "#CC79A7", colour =  "#0072B2") + scale_y_log10()+
   geom_jitter(alpha = 0.04, colour = "#E69F00") +
   ggtitle("(b)")+ 
-  ylab ("energy consumption (in kwh)")
+  ylab ("energy usage [kWh]")
 
 
 
@@ -767,7 +767,7 @@ elec_raw <- elec %>%
   ggplot() +
   geom_density(aes(x = kwh)) +
   facet_wrap(~id, scales = "free_y", ncol = 1) +
-  xlab("demand (in kwh)") + ggtitle("a") +
+  xlab("demand [kWh]") + ggtitle("a") +
   theme(
     strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))
 
@@ -776,7 +776,7 @@ elec_raw <- elec %>%
 #   ggplot() +
 #   geom_density(aes(x = kwh_transformed))+
[…]

### 2021-11-18 Rmd/01-intro.Rmd

--- 2021-11-16 Rmd/01-intro.Rmd
@@
 To acquire a comprehensive view of the repeated patterns, it is necessary to navigate through all of the conceivable cyclic granularities.
-This approach is consistent with the concept of EDA [@Tukey1977-jx], which stresses the utilization of multiple perspectives on data to assist with formulating hypotheses before proceeding to hypothesis testing.
-This, however, is a challenging process since it throws up a myriad of alternatives.
+This approach is consistent with the concept of exploratory data analysis (EDA) [@Tukey1977-jx], which stresses the utilization of multiple perspectives on data to assist with formulating hypotheses before proceeding to formal inferences or modelling.
+This, however, is a challenging process since it throws up a myriad of possible hypotheses.
 Furthermore, the transition from linear to cyclic granularities results in restructured data, with each level of the temporal deconstruction corresponding to multiple values of the observed variable.
@@
 Our proposed method is also applicable to non-temporal hierarchical granularities with an underlying ordered index.
-The methods are implemented in the open-source R package `gravitas` and are consistent with a tidy workflow [@wickham2016r], with probability distributions examined using the range of graphics available in `ggplot2`[@Wickham2009pk].
+The methods are implemented in the open-source R package `gravitas` and are consistent with a tidy workflow [@wickham2016r], with probability distributions examined using the range of graphics available in `ggplot2` [@Wickham2009pk].
 
@@
 The distance measure is computed for a single or pairs of cyclic granularities, and it can be compared across different cyclic granularities as well as a collection of time series.
-This chapter also includes a case study using residential smart meter data from Melbourne to demonstrate how the suggested methodology may be utilized to automatically find temporal granularities with significant distributional differences.
+The method is demonstrated with a case study using residential smart meter data from Melbourne where temporal granularities with significant distributional differences are automatically found.
 The methods are implemented in the open-source R package `hakear`.

### 2021-11-18 Rmd/05-conclusion.Rmd

@@
 
-The `gravitas` package provides very general tools to compute and manipulate cyclic granularities and generate plots displaying distributions conditional on those granularities.
+The `gravitas` package provides general-purpose functions to compute and manipulate cyclic granularities and generate plots displaying distributions conditional on those granularities.
 The functions `search_gran()`, `create_gran()`, `harmony()`, `gran_advice()` and `prob_plot()` provides the entire workflow for an analyst to systematically explore large quantities of temporal data across different harmonies (pairs of granularities that can be analyzed together).
@@
 The R package `hakear` (https://github.com/Sayani07/hakear) provides tools for selecting and sorting significant cyclic granularities.
-The function `wpd()` computes the weighted pairwise distances ($wpd$) for each cyclic granularity or pair of granularities, and the function `select_harmonies()` chooses those with significant patterns and ranks them from highest to lowest $wpd$.
-This package is reliant on parallel processing using multiple multi-core computers for faster computation of $wpd$.
+The function `wpd()` computes the weighted pairwise distances (WPD) for each cyclic granularity or pair of granularities, and the function `select_harmonies()` chooses those with significant patterns and ranks them from highest to lowest WPD.
+This package is reliant on parallel processing using multiple multi-core computers for faster computation of WPD.
 The selected harmonies can be plotted using package `gravitas` for potentially interesting displays.
@@
 The R package `gracsr` (https://github.com/Sayani07/gracsr) has functions for exploring a large number of time series using the clustering methodology described in Chapter \@ref(ch-gracsr).
-The workflow begins with the function `scale gran()`, which may be used to scale individual series using NQT/RS.
-The distances for the JS and WPD approaches are computed in the second phase of the workflow using `dist gran()`/`dist wpd()`.
-The distances can then be used to do clustering with `clust gran()`.
+The workflow begins with the function `scale_gran()`, which may be used to scale individual series using NQT/RS.
+The distances for the JS and WPD approaches are computed in the second phase of the workflow using `dist_gran()`/`dist_wpd()`.
+The distances can then be used to do clustering with `clust_gran()`.
 The package has received a grant (AUD 3000) as part of the ACEMS Business Analytics Prize towards polishing the functions and preparing it for CRAN.

### 2021-11-18 Rmd/00-c-declaration.Rmd
ch: to ch-
declaration percentage and wording

@@
 
-# Declaration {-#ch:declare}
+# Declaration {-#ch-declare}
 
@@
 2 & Visualizing probability distributions across bivariate cyclic temporal granularities & Published & 70\%.
-Concept and developing software and writing first draft & (1) Rob J Hyndman, input into manuscript 15\% (2) Dianne Cook, Concept and input into manuscript 15\% & \multicolumn{1}{c|}{N} \\
+Concept, methodology development, writing first draft and software development & (1) Rob J Hyndman, Concept, methodology and software development, writing 12\% (2) Dianne Cook, Concept, methodology and software development, writing 12\% (3) Antony Unwin, Concept and software development 6\% & \multicolumn{1}{c|}{N} \\
 \hline
 3 & Detecting distributional differences between temporal granularities
-for exploratory time series analysis & To be submitted & 70\%.
-Concept and developing software and writing first draft & (1) Rob J Hyndman, Concept and input into manuscript 15\% (2) Dianne Cook, Concept and input into manuscript 15\%  & \multicolumn{1}{c|}{N} \\
+for exploratory time series analysis & Working Paper & 80\%.
+Concept, methodology development, writing first draft and software development & (1) Rob J Hyndman, Concept, methodology development and writing 10\% (2) Dianne Cook, Concept, methodology development and writing 10\%  & \multicolumn{1}{c|}{N} \\
 \hline
-4 & Clustering time series based on probability distributions across temporal granularities & To be submitted & 70\%.
-Concept and developing software and writing first draft & (1) Dianne Cook, Concept and input into manuscript 20\% (2) Rob J Hyndman, concept and input into manuscript 10\% & \multicolumn{1}{c|}{N} \\
+4 & Clustering time series based on probability distributions across temporal granularities & To be submitted & 80\%.
+Concept, methodology development, writing first draft and software development & (1) Dianne Cook, Concept, methodology development and writing 15\% (2) Rob J Hyndman, Concept, methodology development and writing 5\% & \multicolumn{1}{c|}{N} \\
 \hline

### 2021-11-18 Rmd/00-d-acknowlegements.Rmd
done with Rob's comments
thank Steff, Emi, Nick T and Nick S in acknowledgement for proofreading and add fav quote from Rob.

@@
 
-I want to thank the Department of Econometrics and Business Statistics, Monash Business School, ARC Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS), and Data61 CSIRO for their financial support.
+I want to thank the Department of Econometrics and Business Statistics, Monash Business School, ARC Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS), Data61 CSIRO for their financial support.
 Thanks to the Department, I attended a few conferences in person (before COVID hit) and had a great time networking and being inspired by many people.
@@
 Thank you, Nicholas Spyrison and Mitchell O'Hara-Wild, for the numerous informative discussions, brainstorming sessions, as well as for generously sharing your expertise and experience in software development.
+A special thanks goes to Stephanie, Emi Tanaka, Nicholas Tierney, and Nicholas Spyrison for taking the time out of their busy schedules to proofread parts of my thesis.
 Thank you to each of you, NUMBATs, and the broader R community for directly and indirectly helping me countless times.
@@
 
-PhD in COVID has been hard, and I would not have been able to make it through this if not for the emotional support of my friends and family.
+Doing a PhD in COVID has been hard, and I would not have been able to make it through this if not for the emotional support of my friends and family.
 Thank you Puwasala, Sium, and Ian, for being so kind and acting as a pillar of emotional support, especially during the last leg of my journey.
@@
 Thank you, Juhi (sister-in-law), for reminding me of the value of organization in all aspects of life and the importance of prioritizing my physical health from time to time.
-Thank you for making me feel at home away from home, Somdatta (cousin).
-Lastly, thank you to my one-year-old nephew (Rayan) and niece (Mehr) for being the ultimate stress relievers.
-Thank you all for being a part of this journey with me.
-I love you.
+Lastly, thank you to my one-year-old nephew (Rayan) and niece (Mehr) for being the ultimate stress relievers. Thank you all for being a part of this journey with me. I love you.
 
+<!-- Thank you for making me feel at home away from home, Somdatta (cousin). -->
 <!-- add people who are helping you with proofreading -->
@@
 <!-- add Earo and Stuart that you followed their model of workflow and templates, using the power of reproducible and accessible  research -->
+
+Throughout my PhD candidature, I have had the great privilege of working with and being inspired by my advisers and many other people, which resulted in a great lot of learning in both academics and life.
+Although the journey is coming to a close, I want to carry these life lessons forward, and I would like to conclude with one of my favorites from Rob.
+"Be kind; remember you are collaborating with your future self."
+s

### 2021-11-18 Rmd/05-conclusion.Rmd
Future work update

--- 2021-11-17 Rmd/05-conclusion.Rmd
@@
 The R package `gracsr` (https://github.com/Sayani07/gracsr) has functions for exploring a large number of time series using the clustering methodology described in Chapter \@ref(ch-gracsr).
-The workflow begins with the function `scale gran()`, which may be used to scale individual series using NQT/RS.
-The distances for the JS and WPD approaches are computed in the second phase of the workflow using `dist gran()`/`dist wpd()`.
-The distances can then be used to do clustering with `clust gran()`.
+The workflow begins with the function `scale_gran()`, which may be used to scale individual series using NQT/RS.
+The distances for the JS and WPD approaches are computed in the second phase of the workflow using `dist_gran()`/`dist_ wpd()`.
+The distances can then be used to do clustering with `clust_gran()`.
 The package has received a grant (AUD 3000) as part of the ACEMS Business Analytics Prize towards polishing the functions and preparing it for CRAN.
@@
 
-## Future ideas
+## Limitations and future ideas
 
-There are a few areas to extend this research.
+We address few limitations of the current framework that might serve as natural next steps for this work and some potential short and long-term aims in future ideas.
+
+### Limitations
+
+The time series are observed over a short period of time (1-3 years) in the motivating example of this research.
+Hence, they are assumed to be stationary.
+But it is possible that the distributions change over time, even over a short period.
+For the smart meter example, the distribution for a customer moving to a different house or changing electrical equipment may change drastically.
+To detect these dynamic changes, non-stationarity in time series has to be incorporated while visualizing distributions and also computing distances for two non-stationary time series or one stationary and another non-stationary time series.
+
+Additionally, it is possible that data for a whole category of cyclic granularity is unavailable or that there are insufficient observations to compute distributions.
+For example, a customer may not have data for a particular day of the week or month throughout their observation period.
+While visualizing probability distributions across categories in Chapters \@ref(ch-gravitas) and \@ref(ch-hakear), this can be indicated by displaying dot plots instead of summarizing distributions.
+But the distances in Chapter \@ref(ch-gracsr) can not handle missing observations if they are structured like this.
+We would like to be able to think about designing a distance metric that can incorporate customers with structural missingness and also comprehend its implications while visually characterizing them.
+Another related direction is how to manage the swarm of nuisance variables produced by transforming the time series to cyclic granularities.
+Because the design of our existing framework treats each level of granularity as a variable, it is important to identify levels of granularity that do not contribute to clustering and remove them from the distance metric in order to improve the performance of the clustering algorithm.
+
+### Future ideas
+
+The standard methodologies provide contradicting recommendations for the optimal number of clusters when there are nuisance variables (no contribution to clustering) or nuisance observations (inlying and outlying observations falling between clusters).
+As a result, a more realistic short-term goal should be to test the persistence of clustering solutions in the presence of nuisance variables for the chosen number of clusters.
+This can be done by introducing slightly different samples into the study and observing how the clustering methods handle increased heterogeneity.
+
+Another possible direction is to reduce computational time so that the proposed methods are easily scalable to many customers.
+The distance measure, $wpd$ is computationally heavy even under parallel computation.
+Moreover, while computing distances between time series, the proposed methods compute all possible pairwise distances, which acts as a computational barrier.
+Faster nearest-neighbor search algorithms can be employed here to decrease the computational load.
+
+A longer-term goal would be to create a similar framework for visualizing and analyzing multivariate time series data.
+With multiple time series available for each observation, the complexity of efficient exploration and visualization grows exponentially.
+In this case, conditional distributions include not only temporal dependency but also variables and their dependencies.
+This adds to the already high-dimensional data structures that result from studying distributions.
+This big problem can be tackled by first incorporating time’s inherent characteristics while visualizing one or a few multivariate time series data.
+Unsupervised clustering can then be used to group multiple time series across multiple time granularities and variables.
+This is a method similar to the one used in this thesis for dealing with univariate time series.
+
+<!-- There are a few areas to extend this research.
 The first two explore catering to time series data, where the design of our framework is limited.
@@
 A third possibility is to focus on expanding the scope of the framework to include multivariate time series data.
+-->
 
-**Non-stationary time series**
+<!-- **Non-stationary time series** -->
 
-Because the time series are observed over a short period of time (1-3 years), they are assumed to be stationary.
-But it is possible that the distributions change over time, even over a short period.
-For the smart meter example, the distribution for a customer moving to a different house or changing electrical equipment can change drastically.
-To detect these dynamic changes, non-stationarity in time series has to be incorporated while visualizing distributions and also computing distances for two non-stationary time series or one stationary and another non-stationary time series.
+<!-- **Structured missingness** -->
 
-**Structured missingness**
+<!-- **Anomalies** -->
 
-It is possible that data for a whole category of cyclic granularity is unavailable or that there are insufficient observations to compute distributions.
-For example, a customer may not have data for a particular day of the week or month throughout their observation period.
-While visualizing probability distributions across categories in Chapters \@ref(ch-gravitas) and \@ref(ch-hakear), this can be indicated by displaying dot plots instead of summarizing distributions.
-But the distances in Chapter \@ref(ch-gracsr) can not handle missing observations if they are structured like this.
-More research is needed to design a distance that can incorporate customers with structural missingness and also comprehend its implications while visually characterizing them.
-
-**Anomalies**
-
-Characterizing clusters with varied or outlying customers can result in patterns that do not represent the group.
+<!-- Characterizing clusters with varied or outlying customers can result in patterns that do not represent the group.
 Moreover, integrating heterogeneous consumers may result in visually identical end clusters, which are potentially not useful.
 An appropriate anomaly detection approach could be applied to filter the anomalies, and then a decision has to be made whether the anomalies should form a group or be assigned to the closest group through a classification method.
+-->
 
-**Develop inferential methods for the clusters**
-
-It is important to determine whether or not any detected clusters are statistically significant.
-In Chapter \@ref(ch-gracsr), we have used a permutation approach to detect if patterns are authentic or not across one or a pair of granularities.
-A similar approach could be used for finding the significance of clusters, but permutations used for comparisons must adjust for the fact that different temporal granularities have distinct conditional distributions.
-
-**Scaling up for larger uncertainty and computational efficiency**
-
-Larger data sets with more uncertainty may complicate matters, with potentially more problems than already stated above.
-More research is needed to understand what other sorts of problems may arise with the current methodology and possible ways to address them.
-All of this must be done while keeping computational efficiency in mind, as this is critical when scaling up for the analysis of large data sets.
-
-**Scaling up the methods for multivariate time series data**
-
-Another possible extension would be to create a similar framework for visualizing and analyzing multivariate time series data.
-With multiple time series available for each observation, the complexity of efficient exploration and visualization grows exponentially.
-In this case, conditional distributions include not only temporal dependency but also variables and their dependencies.
-This adds to the already high-dimensional data structures that result from studying distributions.
-This big problem can be tackled by first incorporating time’s inherent characteristics while visualizing one or a few multivariate time series data.
-Unsupervised clustering can then be used to group multiple time series across multiple time granularities and variables.
-This is a method similar to the one used in this thesis for dealing with univariate time series.
+s
 

### 2021-11-18 Rmd/04-gracsr.Rmd
Supplementary "Material" to "material"
no section supplementary

--- 2021-11-17 Rmd/04-gracsr.Rmd
@@
 
-Clustering is a potential approach for Organizing large collections of times series into small homogeneous groups, but a difficult step is determining an appropriate metric to measure similarity between time series.
+Clustering is a potential approach for organizing large collections of time series into small homogeneous groups, but a difficult step is determining an appropriate metric to measure similarity between time series.
 The similarity metric needs to be capable of accommodating long, noisy, and asynchronous time series and also capture cyclical patterns.
@@
 Both are compatible with a variety of clustering algorithms.
-Cyclic granularities like hour-of-the-day, work-day/weekend, month-of-the-year, are useful for finding repeated patterns in the data.
+Cyclic granularities like hour-of-the-day, work-day/weekend, and month-of-the-year, are useful for finding repeated patterns in the data.
 Measuring similarity based on probability distributions across cyclic granularities serves two purposes: (a) characterizing the inherent temporal data structure of long, unequal-length time series in a manner robust to missing or noisy data; (b) small pockets of similar repeated behaviors can be captured.
@@
 #knitr::read_chunk('script/main.R')
-knitr::read_chunk('scripts/gracsr/mylibrary.R')
-knitr::read_chunk('scripts/gracsr/mytheme.R')
-knitr::read_chunk('scripts/gracsr/application.R')
+knitr::read_chunk('scripts/gracsr/library.R')
+knitr::read_chunk('scripts/gracsr/theme.R')
 knitr::read_chunk('scripts/gracsr/application.R')
@@
 The time series features, length of time series, representation technique, and, of course, the purpose of clustering time series all influence the suitable similarity measure or distance metric to a meaningful level.
-The three primary methods to time series clustering (@liao2005clustering) are algorithms that operate directly with distances or raw data points in the time or frequency domain (distance-based), with features derived from raw data (feature-based), or indirectly with models constructed from raw data (model-based).
+The three primary methods to time series clustering [@liao2005clustering] are algorithms that operate directly with distances or raw data points in the time or frequency domain (distance-based), with features derived from raw data (feature-based), or indirectly with models constructed from raw data (model-based).
 The efficacy of distance-based techniques is highly dependent on the distance measure utilized.
 Defining an appropriate distance measure for the raw time series may be a difficult task since it must take into account noise, variable lengths of time series, asynchronous time series, different scales, and missing data.
-Commonly used distance-based similarity measures as suggested by a decade review of time series clustering approaches (@Aghabozorgi2015-ct) are Euclidean, Pearson's correlation coefficient and related distances, Dynamic Time Warping (DTW), Autocorrelation, Short time series distance, Piecewise regularization, cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances (@liao2007clustering) but on a vector time series data.
+Commonly used distance-based similarity measures as suggested by a decade review of time series clustering approaches [@Aghabozorgi2015-ct] are Euclidean, Pearson's correlation coefficient and related distances, Dynamic Time Warping (DTW), Autocorrelation, Short time series distance, Piecewise regularization, cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances [@liao2007clustering] but on vector time series data.
 Among these alternatives, Euclidean distances have high performance but need the same length of data over the same period, resulting in information loss regardless of whether it is on raw data or a smaller collection of features.
-DTW works well with time series of different lengths (@corradini2001dynamic), but it is incapable of handling missing observations.
-Surprisingly, probability distributions, which may reflect the inherent temporal structure of a time series have not been considered in determining time series similarity.
-
-\noindent This work is motivated by a need to cluster a large collection of residential smart meter data, so that households can be grouped into similar energy usage patterns.
+DTW works well with time series of different lengths [@corradini2001dynamic], but it is incapable of handling missing observations.
+Surprisingly, probability distributions, which may reflect the inherent temporal structure of a time series, have not been considered in determining time series similarity.
+
+\noindent This work is motivated by a need to cluster a large collection of residential smart meter data, so that customers can be grouped into similar energy usage patterns.
 These can be considered to be univariate time series of continuous values which are available at fine temporal scales.
 These time series data are long (with more and more data collected at finer resolutions), are asynchronous, with varying time lengths for different houses and sporadic missing values.
-Using probability distributions is a natural way to analyze this types of data because they are robust to uneven length, missing data, or noise.
+Using probability distributions is a natural way to analyze these types of data because they are robust to uneven length, missing data, or noise.
 This paper proposes two approaches for obtaining pairwise similarities based on Jensen-Shannon distances between probability distributions across a selection of cyclic granularities.
-Cyclic temporal granularities [@Gupta2021-hakear], which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can measure repetitive patterns in large univariate time series data.
+Cyclic temporal granularities [@Gupta2021-hakear], which are temporal deconstructions of a time period into units such as hour-of-the-day or work-day/weekend, can measure repetitive patterns in large univariate time series data.
 The resulting clusters are expected to group customers that have similar repetitive behaviors across cyclic granularities.
@@
 The primary application of this work is data from the Smart Grid, Smart City (SGSC) project (2010–2014) available through the [Department of the Environment and Energy](https://data.gov.au/data/organization/doee).
-Half-hourly measurements of usage for more than 13,000 household electricity smart meters is provided from from October 2011 to March 2014.
-<!--Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.--> Households vary in size, location, and amenities such as solar panels, central heating, and air conditioning.
+Half-hourly measurements of usage for more than 13,000 electricity smart meter customers is provided from October 2011 to March 2014.
+<!--Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.--> Customers vary in size, location, and amenities such as solar panels, central heating, and air conditioning.
 The behavioral patterns differ amongst customers due to many temporal dependencies.
-Some households use a dryer, while others dry their clothes on a line.
+Some customers use a dryer, while others dry their clothes on a line.
 Their weekly usage profile may reflect this.
@@
 Age, lifestyle, family composition, building attributes, weather, availability of diverse electrical equipment, among other factors, make the task of properly segmenting customers into comparable energy behavior complex.
-The challenge is to be able to cluster consumers into these type of expected patterns, and other unexpected patterns, using only their energy usage history (@Ushakova2020-rl).
+The challenge is to be able to cluster consumers into these types of expected patterns, and other unexpected patterns, using only their energy usage history [@Ushakova2020-rl].
 <!--To safeguard the customers' privacy, it is probable that such information is not accessible.
@@
 
-There is a growing body of literature focused on time series clustering related to smart meter data.
+There is an extensive body of literature focused on time series clustering related to smart meter data.
 @Tureczek2017-pb conducted a systematic study of over $2100$ peer-reviewed papers on smart meter data analytics.
@@
 $k$-means can be made to perform better by explicitly incorporating time series features such as correlation or cyclic patterns rather than performing it on raw data.
-To reduce dimensionality, several studies use principal component analysis (PCA) or factor analysis to pre-process smart-meter data before clustering (@Ndiaye2011-pf).
+To reduce dimensionality, several studies use principal component analysis (PCA) or factor analysis to pre-process smart-meter data before clustering [@Ndiaye2011-pf].
 PCA eliminates correlation patterns and decreases feature space, but loses interpretability.
@@
 This paper is organized as follows.
-Section~\ref{sec:methodology} provides the clustering methodology.
-Section~\ref{sec:validation} shows data designs to validate our methods.
-Section~\ref{sec:application-gracsr} discusses the application of the method to a subset of the real data.
-Finally, we summarize our results and discuss possible future directions in Section~\ref{sec:discussion-gracsr}.
+Section \ref{sec:methodology} provides the clustering methodology.
+Section \ref{sec:validation} shows data designs to validate our methods.
+Section \ref{sec:application-gracsr} discusses the application of the method to a subset of the real data.
+Finally, we summarize our results and discuss possible future directions in Section \ref{sec:discussion}.
 
@@
 
-<!-- A model-based clustering works by transforming the series into other other objects such as structure or set of parameters which can be more easily characterized and clustered (@Motlagh2019-yj).
+<!-- A model-based clustering works by transforming the series into other other objects such as structure or set of parameters which can be more easily characterized and clustered [@Motlagh2019-yj].
 [@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
@@
 <!-- This work -->
-<!-- This is similar to a stochastic approach (@Motlagh2019-yj) to clustering, which proposes interpreting electricity demand as a random process and extracting time series characteristics, or a model of the series, to enable unsupervised clustering.
+<!-- This is similar to a stochastic approach [@Motlagh2019-yj) to clustering, which proposes interpreting electricity demand as a random process and extracting time series characteristics, or a model of the series, to enable unsupervised clustering.
 Unsupervised clustering is only as good as the features that are extracted/selected or the distance metrics that were utilized.
@@
 <!--The foundation of our method is unsupervised clustering algorithms based exclusively on the time series data.
-The proposed methods leverage the intrinsic data structure hidden within cyclic temporal granularities.--> The existing work on clustering probability distributions assumes we have an independent and identically distributed samples $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
+The proposed methods leverage the intrinsic data structure hidden within cyclic temporal granularities.--> The existing work on clustering probability distributions assumes we have independent and identically distributed samples $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
 <!--So going back to the smart meter example, $f_i(v)$ is the distribution of customer $i$ and $v$ is electricity demand.--> In our approach, instead of considering the probability distributions of the linear time series, we compare them across different categories of a cyclic granularity.
-We can consider categories of an individual cyclic granularity ($A$) or combination of categories for two interacting granularities ($A*B$) to have a distribution, where $A, B$ are defined as $A = \{a_j: j=1, 2, \dots J\}$ and $B = \{b_k: k = 1, 2, \dots K\}$.
-For example, let us consider two cyclic granularities, $A$ and $B$, representing hour-of-day and day-of-week, repsectively.
+We can consider categories of an individual cyclic granularity ($A$) or combination of categories for two interacting granularities ($A, B$) to have a distribution, where $A$ and $B$ are defined as $A = \{a_j: j=1, 2, \dots J\}$ and $B = \{b_k: k = 1, 2, \dots K\}$.
+For example, let us consider two cyclic granularities, $A$ and $B$, representing hour-of-day and day-of-week, respectively.
 Then $A = \{0, 1, 2, \dots, 23\}$ and $B = \{Mon, Tue, Wed, \dots, Sun\}$.
 In case individual granularities are considered, there are $J = 24$ distributions of the form $f_{i,j}(v)$ or $K = 7$ distributions of the form $f_{i,k}(v)$ for each customer $i$.
-In case of interaction, $J*K=168$ distributions of the form $f_{i,j, k}(v)$ could be conceived for each customer $i$.
+In case of interaction, $J \times K=168$ distributions of the form $f_{i,j, k}(v)$ could be conceived for each customer $i$.
 Hence clustering these customers is equivalent to clustering these collections of conditional univariate probability distributions.
@@
 
-[@Gupta2021-hakear] provides a method for determining the significance of a cyclic granularity, and a ranking of multiple cyclic granularities.
+@Gupta2021-hakear provide a distance measure ($wpd$) for determining the significance of a cyclic granularity, and a ranking of multiple cyclic granularities.
 (This extends to harmonies, pairs of granularities that might interact with each other.) We define "significant" granularities as those with significant distributional differences across at least one category.
@@
 Granularities that do not have some difference between categories are likely to be nuisance variables.
-It should be noted that all of the time series in a collection may not have the same set of significant granularities.
-This is the approach for managing this generate a subset ($S_c$) of significant granularities across a collection of time series:
+It should be noted that all of the time series in a collection may not have the same set of significant granularities.<!--This is the approach for managing this generate a subset ($S_c$) of significant granularities across a collection of time series:--> This is the approach for generating a subset ($S_c$) of significant granularities across a collection of time series:
 
@@
 Skewed distributions need to be symmetrized.
-Scales of individuals needs to be standardized, because clustering is to select similar patterns, not magnitude of usage.
-(Organizing individuals based on magnitude can achieved simply by sorting on a statistic like the average value across time.) <!--Observations often have a somewhat skewed time series distribution and their ranges might vary greatly.
+Scales of individuals need to be standardized, because clustering is to select similar patterns, not magnitude of usage.
+(Organizing individuals based on magnitude can be achieved simply by sorting on a statistic like the average value across time.) <!--Observations often have a somewhat skewed time series distribution and their ranges might vary greatly.
 Statistical transformations are employed to bring all of them to the same range or normalize each observation.
-This is important because we are not interested in trivial clusters that vary in magnitudes, but rather in uncovering comparable patterns of distributional differences between categories.--> For the JS-based approaches, two data transformation techniques are recommended, normal-quantile transform (NQT) and robust scaling (RS).
-NQT is used in @@Gupta2021-hakear prior sorting granularities.
-<!-- $wpd$, which is the foundation of wpd-based distances.
-XX wpd was not yet introduced-->
-
-- RS: The normalized $i^{th}$ observation is denoted by $v_{norm} = \frac{v_t - q_{0.5}}{q_{0.75}-q_{0.25}}$, where $v_t$ is the actual value at the $t^{th}$ time point and $q_{0.25}$, $q_{0.5}$ and $q_{0.75}$ are the $25^{th}$, $50^{th}$ and $75^{th}$ percentile of the time series for the $i^{th}$ observation.
-Note that, $v_{norm}$ has zero mean and median, but otherwise the shape does not change.
+This is important because we are not interested in trivial clusters that vary in magnitudes, but rather in uncovering comparable patterns of distributional differences between categories.--> For the JS-based approaches, two data transformation techniques are recommended, normal-quantile transform (NQT) and robust scaling (RS).<!-- NQT is used in @Gupta2021-hakear prior sorting granularities.-- $wpd$, which is the foundation of wpd-based distances.
+XX wpd was not yet introduced--> While @Gupta2021-hakear already use NQT when computing $wpd$, it could be useful to standardize it for the selected set of significant granularities prior to computing the distances.
+
+- RS: The normalized $i^{th}$ observation is denoted by $v_{norm} = \frac{v_t - q_{0.5}}{q_{0.75}-q_{0.25}}$, where $v_t$ is the actual value at the $t^{th}$ time point and $q_{0.25}$, $q_{0.5}$ and $q_{0.75}$ are the $25^{th}$, $50^{th}$ and $75^{th}$ percentiles of the time series for the $i^{th}$ observation.
+Note that $v_{norm}$ has zero mean and median, but otherwise the shape does not change.
 
@@
 
-Computationally in R, the data is assumed to be a "tsibble object" (@wang2020tsibble) equipped with an index variable representing inherent ordering from past to present and a key variable that defines observational units over time.
+Computationally in R, the data is assumed to be a "tsibble object" [@wang2020tsibble] equipped with an index variable representing inherent ordering from past to present and a key variable that defines observational units over time.
 The measured variable for an observation is a time-indexed sequence of values.
@@
 This distance metric considers two time series to be similar if the distributions of each category of an individual cyclic granularity or combination of categories for interacting cyclic granularities are similar.
-In this study, the distribution for each category is characterized using deciles (can potentially consider any list of quantiles), and the distances between distributions are calculated using the Jensen-Shannon distances (@Menendez1997-in), which are symmetric and thus could be used as a distance measure.
-
-The sum of the distances between two observations $x$ and $y$ in terms of cyclic granularity $A$ is defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity $A$) or $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$).
+In this study, the distribution for each category is characterized using deciles (can potentially consider any list of quantiles), and the distances between distributions are calculated using the Jensen-Shannon distances [@Menendez1997-in], which are symmetric and thus could be used as a distance measure.
+
+The sum of the distances between two observations $x$ and $y$ in terms of a cyclic granularity $A$ is defined as $$S^A_{x,y} = \sum_{j\in A} D(x_j,y_j)$$ where $D$ is the Jensen-Shannon distances, $x_j$ is the set of quantiles over the values filtered by $j^{th}$ level of granularity $A$ for observation $x$ (similar for $y$).
+
+The sum of the distances between two observations $x$ and $y$ in terms of a pair of cyclic granularities $(A, B)$ is defined as $$S^{A,B}_{x,y} = \sum_{(j, k) \in (A, B)} D(x_{jk},y_{jk})$$ $x_{jk}$ is the set of quantiles over the values filtered by the combination of $j^{th}$ level of granularity $A$ and $k^{th}$ level of granularity $B$ for observation $x$ (similar for $y$).
+
 After determining the distance between two series in terms of one granularity, we must combine them to produce a distance based on all significant granularities.
-When combining distances from individual $L$ cyclic granularities $C_l$ with $n_l$ levels, $$S_{x, y} = \sum_lS^{C_l}_{x,y}/n_l$$ is employed, which is also a distance metric since it is the sum of JS distances.
+When combining distances from individual $L$ cyclic granularities $C_l$ with $n_l$ levels, $$S_{x, y} = \sum_{l \in L}S^{C_l}_{x,y}/n_l$$ is employed, which is also a distance metric since it is the sum of JS distances.
 This approach is expected to yield groups, such that the variation in observations within each group is in magnitude rather than distributional pattern, while the variation between groups is only in distributional pattern across categories.
@@
 
-Compute weighted pairwise distances $wpd$ (@Gupta2021-hakear) for all considered granularities for all observations.
-$wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction and is estimated by the maximum pairwise distances between consecutive categories normalized by appropriate parameters.
+We compute weighted pairwise distances $wpd$ [@Gupta2021-hakear] for all considered granularities for all observations.
+$wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction.
+It is estimated by the maximum pairwise distances between distributions across consecutive categories normalized by appropriate parameters.
 A higher value of $wpd$ indicates that some interesting patterns are expected, whereas a lower value would indicate otherwise.
@@
 If we have $H_{N_C}$ harmonies in the harmony table, then for each household we have a vector of $wpd_{{s}}$ of $H_{N_C}$ elements with each element corresponding to one harmony.
-We aim to have pockets of households showing similar periodic behavior by considering $wpd$ values for different harmonies and some time series features.
+We aim to have pockets of customers showing similar periodic behavior by considering $wpd$ values for different harmonies and some time series features.
 The features should also characterize probability distributions of different household.
@@
 
-Different approaches to determining number of clusters are required for varied applications and research objectives.
-It is often beneficial to investigate numerous cluster statistics in order to arrive at a number of clusters, since clustering has a variety of objectives (e.g., between-cluster separation, within-cluster homogeneity, representation by centroids, etc.), which could be in conflict with one another.
-Some typical ways of determining the number of clusters in the literature (e.g., the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@Dunn1973-gk]) are inspired by defining and balancing "within-cluster homogeneity" and "between-cluster separation".
-The Cluster separation index ($sindex$) proposed by @Hennig2014-ah is used as a guide to select the number of clusters in Section \ref{sec:validation}.
-$sindex$ [@R-fpc] is a compromise between minimum cluster separation and average cluster separation and is less sensitive to a single or few ambiguous points.
-$sindex$ cannot be optimized over the number of clusters, since increasing the number of clusters reduces it.
-The number of clusters required is determined by the slope going from steep to shallow (an elbow).
+Determining the number of clusters is typically a difficult task.
+Many metrics have been defined for choosing clusters.
+Most metrics for choosing the optimal number of clusters are based on comparing distances between observations within a class to those distances between observations between classes, which makes the assumption that there are some separated clusters.
+Some common procedures include the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@Dunn1973-gk] and the separation index ($sindex$) [@Hennig2019-eq; @Hennig2014-ah].
+These are constructed by balancing within-cluster homogeneity and between-cluster separation.
+
+All of the common approaches can give contradictory suggestions for the optimal number of clusters, particularly when the data does not naturally break into groups, or in the presence of nuisance variables (no contribution to clustering) or nuisance observations (inlying and outlying observations falling between clusters).
+There is no one best metric, which is perhaps a reason why so many metrics exist.
+
+In this work, we have chosen to use $sindex$.
+It is a very simple but effective metric.
+This is computed by averaging the smallest 10% of inter-cluster distances.
+It is relatively robust to nuisance observations.
+The value of $sindex$ always decreases, and sharp drops in value indicate candidates for the optimal number of clusters.
+The number of clusters corresponding to the value **before the drop** is the recommendation.
 
@@
 
-Cluster characterization is an important element of cluster analysis.
-@Cook2007-qe provides several methods for characterizing clusters.
-_Parallel coordinate plots_ (@wegman1990hyperdimensional), _Scatterplot matrix_, _Displaying cluster statistics_ (@dasu2005grouping), _MDS_ (@borg2005modern), PCA, _t-SNE_(@R-tsne), _Tour_ (@wickham2011tourr) are some of the graphical approaches used in this study.
-A Parallel Coordinates Plot features parallel axes for each variable and each axis is linked by lines.
+Cluster characterization is an important final stage of a cluster analysis.
+The primary purpose is to compare the homogeneity within a cluster to the heterogeneity of clusters.
+This can be done numerically, by tabulating cluster means and standard deviations [@dasu2005grouping], and visually using methods for graphics multivariate data.
+@Cook2007-qe provide visual examples using both tours [@asimov1985grand] and parallel coordinate plots [@wegman1990hyperdimensional].
+Dimension reduction techniques like principal component analysis [@jolliffe2016principal], multidimensional scaling (MDS) [@borg2005modern], t-distributed stochastic neighbor embedding (t-SNE) [@van2008visualizing] and linear discriminant analysis (LDA) [@fisher1936use] are also useful.
+<!-- A Parallel Coordinates Plot features parallel axes for each variable and each axis is linked by lines.
 Changing the axes may reveal patterns or relationships between variables for categorical variables.
@@
 All of MDS, PCA and t-SNE use a distance or dissimilarity matrix to construct a reduced-dimension space representation, but their goals are diverse.
-Multidimensional scaling (@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space.
+Multidimensional scaling [@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space.
 The t-SNE embedding will compress data points that are close in high-dimensional space.
@@
 For JS-based distances, plotting cluster statistics is beneficial for characterization and variable importance could be displayed through parallel coordinate plots.
+-->
 
@@
 
-<!-- (a) _Parallel coordinate plots_ (@wegman1990hyperdimensional) are often used to visualize high-dimensional and multivariate data, allowing visual grouping and pattern detection.
+<!-- (a) _Parallel coordinate plots_ [@wegman1990hyperdimensional) are often used to visualize high-dimensional and multivariate data, allowing visual grouping and pattern detection.
 A Parallel Coordinates Plot features parallel axes for each variable.
@@
 
-<!-- (c) _Displaying cluster statistics_ are useful when we have larger problems and it is difficult to read the Parallel coordinate plots due to congestion. (@dasu2005grouping) -->
+<!-- (c) _Displaying cluster statistics_ are useful when we have larger problems and it is difficult to read the Parallel coordinate plots due to congestion. [@dasu2005grouping) -->
 
@@
 PCA seeks to retain data variance.
-Multidimensional scaling (@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space.
+Multidimensional scaling [@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space.
 t-SNE, on the other hand, is concerned with preserving neighborhood data points.
@@
 
-<!-- This part of the work uses R packages `GGally` (@R-GGally), `Rtsne` (@R-tsne), `ggplot2` (Wickham2009pk), `tour` (@wickham2011tourr), `stats` (@R-language). -->
+<!-- This part of the work uses R packages `GGally` [@R-GGally), `Rtsne` [@R-tsne), `ggplot2` (Wickham2009pk), `tour` [@wickham2011tourr), `stats` [@R-language). -->
 
@@
 <!-- 1. Compute quantiles of distributions across each category of the cyclic granularity -->
-<!-- 2. Compute JS distance between households for each each category of the cyclic granularity -->
-<!-- 3. Total distance between households computed as sum of JS distances for all hours -->
+<!-- 2. Compute JS distance between customers for each each category of the cyclic granularity -->
+<!-- 3. Total distance between customers computed as sum of JS distances for all hours -->
 <!-- 4. Cluster using this distance with hierarchical clustering algorithm (method "Ward.D") -->
@@
 
-To validate our clustering methods, we spiked many attributes in the data to create different data designs.
-<!--to see where one method works better than the other and where they might give us the same outcome --or the effect of missing data <!---and trends on the proposed methods.--> Three circular granularities $g1$, $g2$ and $g3$ are considered with categories denoted by $\{g10,g11\}$, $\{g20, g21, g22\}$ and $\{g30, g31, g32, g33, g34\}$ and levels $n_{g_1}=2$, $n_{g_2}=3$ and $n_{g_3}=5$.
+To validate our clustering methods, we have created several different data designs containing different granularity features.
+<!--to see where one method works better than the other and where they might give us the same outcome --or the effect of missing data <!---and trends on the proposed methods.--> There are three circular granularities $g1$, $g2$ and $g3$ with categories denoted by $\{g10,g11\}$, $\{g20, g21, g22\}$ and $\{g30, g31, g32, g33, g34\}$ and levels $n_{g_1}=2$, $n_{g_2}=3$ and $n_{g_3}=5$.
 These categories could be integers or some more meaningful labels.
@@
 So three scenarios are considered to accommodate that.
-Figure \ref{fig:range-parameter} shows the range of parameters considered for each scenario.
-
-```{r range-parameter, fig.cap="Range of parameters for different scenarios"}
-knitr::include_graphics("img/simulation_table.png")
+Table \ref{tab:range-parameter} shows the range of parameters considered for each scenario.
+
+<!-- ```{r range-parameter, fig.cap="The range of parameters used for the validation study, for the three different scenarios, number of simulations for each design, differences between means across granularities and series lengths.
+"} -->
+<!-- knitr::include_graphics("img/simulation_table.png") -->
+<!-- ``` -->
+
+```{r range-parameter}
+scenario = c("S1", "S2","S3")
+designs = c(5, 4, 4)
+niter = list (c(25, 20, 500),
+              c(25, 20, 500),
+              c(25, 20, 500))
+mu = list (c(1, 2, 5),
+           c(1, 2, 5),
+           c(1, 2, 5))
+nT = list (c(300, 1000, 5000),
+           c(300, 1000, 5000),
+           c(300, 1000, 5000))
+
+tibble::tibble(scenario, designs, niter, mu, nT) %>% rlang::set_names("scenario", "designs", "R", "$\\mu$", "T") %>% 
+  #tibble::tibble(scenario, " designs" = designs, "R" =niter,  "$\\mu$" = mu, "T"=nT) %>% 
+  knitr::kable(format = "latex",
+             escape = FALSE,
+             caption = "The range of parameters used for the validation study, for the three different scenarios, number of simulations ($R$) for each design, differences between means ($\\mu$) across granularities and series lengths ($T$).")%>%  kableExtra::collapse_rows(columns = 3:5)
 ```
@@
 
-_Scenario (S1): All granularities significant_
-
+**Scenario 1 (S1) - All granularities significant:**
 <!-- Consider a case where all the three granularities $g1$, $g2$ and $g3$ would be responsible for making the designs distinct.
@@
 -->
-
 Consider the instance where $g1$, $g2$, and $g3$ all contribute to design distinction.
@@
 
-_Scenario (S2): Few significant granularities_
-
+**Scenario 2 (S2) - Few significant granularities:**
 This is the case where one granularity will remain the same across all designs.
@@
 
-_Scenario (S3): Only one significant granularity_
-
+**Scenario 3 (S3) - Only one significant granularity:**
 Only one granularity is responsible for identifying the designs in this case. This is depicted in Figure \ref{fig:gran2and1-clubbed} (right) where only $g3$ affects the designs significantly.
@@
 As a result, in this scenario, we analyze a combination of categories generated from different distributions.
-Display of design and related results can be found in supplementary.
+Display of design and related results can be found in supplementary material.
 
@@
 The formed clusters have to match the design, be well separated, and have minimal intra-cluster variation.
-It is possible to study these desired clustering traits visually in a more comprehensive way than just looking at index values.
-MDS and parallel coordinate graphs are used to demonstrate the findings, as well as an index value plot to provide direction on the number of clusters.
+<!--It is possible to study these desired clustering traits visually in a more comprehensive way than just looking at index values.--> MDS and parallel coordinate graphs are used to demonstrate the findings, as well as an index value plot to provide direction on the number of clusters.
 JS-based approaches corresponding to NQT and RS are referred to as JS-NQT and JS-RS respectively.
-In the following plots, results for JS-NQT are reported.
-For similar results with JS-RS or wpd-based distances, please refer to the supplementary paper.
-
-- Figure \ref{fig:sindex-plot-validation} shows $sindex$ plotted against the number of clusters ($k$) for the range of mean differences (rows) under the different scenarios (columns).
+In the following plots, results for JS-NQT are reported, and results with JS-RS or wpd-based distances are in the supplementary material.
+
+Figure \ref{fig:sindex-plot-validation} shows $sindex$ plotted against the number of clusters ($k$) for the range of mean differences (rows) under the different scenarios (columns).
 This can be used to determine the number of clusters for each scenario.
-When $sindex$ for each scenario are examined, it appears that $k = 5, 4, 4$ is justified for scenarios S1, S2, and S3, respectively, given the sharp decrease in $sindex$ from that value of $k$.
+When $sindex$ for each scenario are examined, it appears that $k = \{5, 4, 4\}$ is justified for scenarios S1, S2, and S3, respectively, given the sharp decrease in $sindex$ from that value of $k$.
 Thus, the number of clusters corresponds to the number of designs that were originally considered in each scenario.
 
-- Figure \ref{fig:mds-plot-validation} shows separation of our clusters.
+Figure \ref{fig:mds-plot-validation} shows separation of our clusters.
 It can be observed that in all scenarios and for different mean differences, clusters are separated.
@@
 
-- Figure \ref{fig:parcoord-sim} depicts a parallel coordinate plot with the vertical bar showing total inter-cluster distances with regard to granularities $g1$, $g2$, and $g3$ for all simulation settings and scenarios.
+Figure \ref{fig:parcoord-sim} depicts a parallel coordinate plot with the vertical bar showing total inter-cluster distances with regard to granularities $g1$, $g2$, and $g3$ for all simulation settings and scenarios.
 So one line in the figure shows the inter-cluster distances for one simulation setting and scenarios vary across facets.
@@
 
- - The JS-RS and wpd-based methods perform worse for $nT=300$, then improve for higher $nT$ evaluated in the study.
+The JS-RS and wpd-based methods perform worse for $nT=300$, then improve for higher $nT$ evaluated in the study.
 However, a complete year of data is the minimum requirement to capture distributional differences in winter and summer profiles, for example.
@@
 
-- In our study sample, the method JS-NQT outperforms the method JS-RS for smaller differences between categories. More testing, however, is required to corroborate this.
+In our study sample, the method JS-NQT outperforms the method JS-RS for smaller differences between categories. More testing, however, would be needed to to be confident in this conclusion.
 
@@
 
-```{r sindex-plot-validation, fig.cap = "The cluster separation index (sindex) is plotted as a function of the number of clusters for the range of mean differences (rows) under the different scenarios (columns).
-S1 has a sharp decrease in sindex from 5 to 6, whereas S2 and S3 have a decrease from 4 to 5.
-As a result, the number of clusters considered for S1, S2, and S3 is 5, 4, 4, respectively.
-This corresponds to the number of designs taken into account in each scenario."}
+```{r sindex-plot-validation, fig.cap = "Choosing optimal cluster number across the range of scenarios and mean differences used in the validation study, using the cluster separation index (sindex) for the JS-NQT.
+S1 has a sharp decrease in sindex from 5 to 6, whereas S2 and S3 have a decrease from 4 to 5, especially when mean difference is large, providing the recommended number of clusters to be 5, 4, 4, respectively.
+This precisely reflects the structure in designs that we would hope the clustering could recover."}
 ```
@@
 
-```{r parcoord-sim, fig.cap = "The parallel coordinate figure depicts total inter-cluster distances for the granularities $g1$, $g2$, and $g3$.
-The inter-cluster distances for a single simulation scenario are represented by one line in the figure.
-While panel S1 shows that no variable stands out during clustering, panels S2 and S3 show that g1 and (g1, g2) had smaller inter-cluster distances, indicating that they did not contribute to clustering.
-These facts are consistent with our initial assumptions when building the scenarios, and S1, S2, S3 correspond to scenarios with three, two, and one significant granularity, respectively."}
+```{r parcoord-sim, fig.cap = "Exploring the contribution of granularities in the clustering for scenarios S1, S2, S3, using parallel coordinate plots.
+Inter-cluster distances are displayed vertically.
+All three granularities $g1$, $g2$, and $g3$ have high inter-cluster distances for S1, suggesting all are important.
+In S2 $g1$ and in S3 both $g1$ and $g2$ have smaller inter-cluster distances, indicating that they did not contribute to clustering."}
 
@@
 
-The use of our methodology is illustrated on smart meter energy usage for a sample of customers from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) which was available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) and Data61 CSIRO.
-It contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.<!--In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual.
+Clustering with the new distances is illustrated on the smart meter energy usage for a sample of customers from @smart-meter.
+The full data contains half-hourly general supply in kWh for 13,735 customers, resulting in 344,518,791 observations in total.<!--In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual.
 We do not learn about these repetitive behaviors from the linear view because too many measurements all squeezed in that representation.
-Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior.--> The raw data for these consumers is of unequal length, with varying start and finish dates.
-Because our proposed methods evaluate probability distributions rather than raw data, neither of these data features would pose any threat to our methodology unless they contained any structure or systematic patterns.
-Additionally, there were missing values in the database but further investigation revealed that there is no structure in the missingness (see Supplementary paper for raw data features and missingness).
-
-Clustering huge data sets presents a slew of difficulties.
+Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior.--> The raw data for these consumers is of unequal length, with varying starting and end dates.
+Additionally, there were missing values in many series.
+(The supplementary material contains details from checking for systematic missingness.) Because our proposed methods evaluate probability distributions rather than raw data, these data issues are not problematic, unless there is any systematic structure related to granularities.
+
+Huge data sets present more complications for clustering.
+Clustering algorithms work well when there are well-separated clusters, with no nuisance variables or nuisance observations.
+When converting a series to granularities, many variables (each level of a granularity) are generated, possibly creating a slew of nuisance variables.
+Some customers may have a mix of energy use patterns, which could be considered nuisance observations located between major clusters.
+For this reason, we have chosen to select a small group of customers with relatively distinct and different patterns in order to illustrate the clustering more simply.
+Figure \ref{fig:hod-ind-group-png} shows the distribution across $hod$, $moy$ and $wnwd$ for the set of $24$ customers used to illustrate clustering.
+The customers are displayed in two columns of $12$ for space reasons.
+Each row, of each column, represents the profile of a single customer across different variables.
+Each customer is associated with an identifier of the form [$a$-$b$], where $a \in \{1, 2, \dots, 24\}$ represents the customer-prototype id and $b \in \{1, 2, \dots, 5\}$ indicates the label of the prototype in which a customer was placed.
+This is often a good approach to tackling a big analysis task, to start with a simpler task.
+The approach, however, is applicable to all customers.
+
 As a result, we dissect the larger problem and test our solutions on a small sample of prototype customers.
-To do this, data is first filtered to generate a clean sample, and then significant cyclic granularities (variables) for them are chosen.
-This is described in Section \ref{sec:datafilter}.
-The clean set is subsequently subsampled along all dimensions of interest, to ensure that the sampled data reveals some patterns across at least one specified variable, which is described in Section \ref{sec:prototype}.
-By grouping the prototypes using our methods in Section \ref{sec:clustering} and assessing their meaning, the study hopes to unravel some of the heterogeneities observed in energy usage data.
-Because our application does not employ additional customer data, we cannot explain why consumption varies, but rather try to identify how it varies.
+To do this, data is first filtered to generate a small sample, and then significant cyclic granularities (variables) for them are chosen (as described in Section \ref{sec:datafilter}).
+The sample set is subsequently examined along all dimensions of interest, to ensure that they reveal some patterns across at least one specified variable (as described in Section \ref{sec:prototype}).
+<!--By grouping the prototypes using our methods in Section \ref{sec:clustering} and assessing their meaning, the study hopes to unravel some of the heterogeneities observed in energy usage data.--> Because the data does not contain additional customer characteristics, we cannot explain why consumption varies, but can only identify how it varies.
 
@@
 
-- Choose a smaller subset of randomly selected $600$ customers with no implicit missing values for 2013.
-
-- Obtain $wpd$ for all cyclic granularities considered for these customers.
-It was found that `hod` (hour-of-day), `moy` (month-of-year) and `wkndwday` (weekend/weekday) are turning out to be significant for most customers.
+The steps for customer filtering and variable selection were:
+
+1. Choose a smaller subset of randomly selected $600$ customers with no implicit missing values for 2013.
+2.
+Obtain $wpd$ for all cyclic granularities considered for these customers.
+It was found that $hod$ (hour-of-day), $moy$ (month-of-year) and $wnwd$ (weekend/weekday) are significant for most customers.
 We use these three granularities while clustering.
-
-- Remove customers whose data for an entire category of `hod`, `moy` or `wnwd` is empty.
+3.
+Remove customers whose data for an entire category of $hod$, $moy$ or $wnwd$ is empty.
 For example, a customer who does not have data for an entire month is excluded because their monthly behavior cannot be analyzed.
-
-- Remove customers whose energy consumption is 0 in all deciles.
+4.
+Remove customers whose energy consumption is 0 in all deciles.
 These are the clients whose consumption is likely to remain essentially flat and with no intriguing repeated patterns that we are interested in studying.
 
-### Prototype selection {#sec:prototype}
+### Selecting prototypes {#sec:prototype}
 
 <!-- Why instance selection -->
-Supervised learning uses a training set of known information to categorize new events through instance selection.
-Instance selection (@olvera2010review) is a method of rejecting instances that are not helpful for classification.
-This is analogous to subsampling the population along all dimensions of interest such that the sampled data represents the primary features of the underlying distribution.
-Instance selection in unsupervised learning has received little attention in the literature.
-However, it could be a useful tool for evaluating model or method performance.
-There are several ways to approach prototype selection.
-Following the idea of @Fan2021-bq of picking related examples (neighbors) for each instance (anchor), we can first use any dimensionality reduction techniques like MDS or PCA to project the data into a 2D space.
-Then pick a few "anchor" customers far apart in 2D space and pick a few neighbors for each.
-Unfortunately, this does not ensure that consumers with significant patterns across all variables are chosen.
-Tours can reveal variable separation that gets hidden in a single variable display better than static projections.
-Hence, we perform a linked tour with a t-SNE layout using the R package `liminal` (@R-liminal) to identify customers who are more likely to have distinct patterns across the variables studied.
-(Refer to the supplementary article for further details).
-Figure \ref{fig:hod-ind-group-png} shows the distribution across `hod`, `moy` and `wnwd` for the set of chosen $24$ customers that were chosen.
-The customers are split into batches of $12$ in (a) and (b).
-Each row in (a) and that in (b) represents the profile of one customer across different variables.
-<!--Few of these customers have similar distribution across `moy` and some are similar in their `hod` distribution.
+
+It is common to filter data prior to fitting a supervised classification model using instance selection [@olvera2010review] which removes observations that might impede the model building.
+For clustering, this is analogous to identifying and removing nuisance observations.
+Prototype selection is more severe than instance selection, because only a handful of cases is selected.
+@Cutler1994-hm proposed a method called archetypal analysis which has inspired this approach but the procedure we have used follows @Fan2021-bq.
+First, dimension reduction such as t-SNE, MDS or PCA is used to project the data into a 2D space.
+Second, a few "anchor" customers far apart in 2D space are selected.
+Additional close neighbors to the anchors are selected.
+To check the selections relative to the full set of variables, we used a tour linked to a t-SNE layout using the R package `liminal` [@R-liminal].
+This ensured that the final sample of clustered customers were also far apart in the high-dimensional space.
+(See the supplementary materials for further details.)<!--Few of these customers have similar distribution across $moy$ and some are similar in their $hod$ distribution.
 -->
@@
 <!-- # ```{r hod-moy-wkndwday plots, fig.cap="The distribution across moy, hod and wkndwday for the selected designs.
-Few are similar in their hod pattern, while others are similar in moy behavior.
-Some customers have distinct behavior as compared to all other customers.For example, although patterns across wkndwday do not look distinctly different for most households, there is one household for whom weekend behavior is standing out from the rest."} -->
+Few are similar in their hod pattern, while others are similar in $moy$ behavior.
+Some customers have distinct behavior as compared to all other customers.For example, although patterns across wkndwday do not look distinctly different for most customers, there is one household for whom weekend behavior is standing out from the rest."} -->
 <!-- # ``` -->
 
-### Clustering {#sec:clustering}
-
-The $24$ prototypes <!--and the data subset consisting of $353$ customers--> are clustered using the methodology described in Section \ref{sec:methodology} and results are reported below.
-In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$ percentiles.<!--All customers with the same color represent the same clustered groups.
---> Groups are colored differently whenever they represent different groupings.
-While plotting the clusters, RS-based scaling is applied to each customer, so that no customer dominates the shape of the summarized clusters because of magnitude.
+### Clustering results {#sec:clustering}
+
+Clustering of the $24$ prototypes <!--and the data subset consisting of $353$ customers--> was conducted with all three distances, JS-NQT, JS-RS and WPD, and is summarized in Figures \ref{fig:opt-cluster-tsne-jsd}, \ref{fig:groups-4and5} and \ref{fig:summary-plot-wpd}.
+The t-SNE visualization suggests that there are four well-separated clusters.
+It is possible that because the representation is only 2D, the fifth group from the original prototype selection is distinctly different in high dimensions.
+The $sindex$ plots for the three methods indicate some disagreement: JS-NQT suggests $3$, JS-RS suggests $2$ or $5$ and WPD suggests $3$ or $5$.
+JS-RS would appear to match the original prototypes with the five cluster solution, but it actually differs.
+Even though the $sindex$ for JS-NQT suggests three clusters, the five cluster solution more closely matches the original prototypes.
+The WPD clustering provides a different grouping of the customers, and even though it disagrees with the original prototypes it is a useful grouping.<!--While plotting the clusters, RS-based scaling is applied to each customer, so that no customer dominates the shape of the summarized clusters because of magnitude.
 The plotting scales are not displayed since we want to emphasize comparable patterns rather than scales.
-The idea is that a customer in a cluster may have low total energy usage, but their behavior may be quite similar to a customer with high usage with respect to distributional pattern or significance across cyclic granularities.
-
-#### JS-based distances
-
-The $sindex$ is plotted against the number of clusters for clustering based on JS-based (JS-NQT) distances in Figure \ref{fig:opt-cluster-tsne-jsd}.
+The idea is that a customer in a cluster may have low total energy usage, but their behavior may be quite similar to a customer with high usage with respect to distributional pattern or significance across cyclic granularities.-->
+
+<!--### JS-NQT and JS-RS-->
+
+<!--Figure \ref{fig:opt-cluster-tsne-jsd} The $sindex$ is plotted against the number of clusters for clustering based on JS-based (JS-NQT) distances in Figure \ref{fig:opt-cluster-tsne-jsd}.
 The $sindex$ falls from around $0.56$ when $k=3$ to $0.54$ when $k$ increases.
-In our study, we use $k=4, 5$.
-Figure \ref{fig:groups-4and5} depicts the summarized distributions across $4$ and  $5$ clusters in (a) and (b) respectively, and assists in characterizing each cluster.
-A-2, Groups A-3, and A-4 have profiles that correspond to B-2, B-3, and B-5, respectively.
-B-2 (id: 4-9 in Figure \ref{fig:hod-ind-group-png}) and B-5 (id: 21-24) have a `hod` pattern with a typical morning and evening peak, whereas the other groups have no distinct pattern or high variability throughout the day.
-A-1 is subdivided into B-1 and B-4, each of which has a distinct shape across `moy` and `wnwd`.
+In our study, we use $k=4, 5$.--> Figure \ref{fig:groups-4and5} displays the summarized distributions across $4$ and $5$ clusters from JS-NQT clustering in (a) and (b) respectively, and helps to characterize each cluster.
+In the quantile plots the line represents the median, and the region shows the area between the $25^{th}$ and $75^{th}$ percentiles.
+The only difference between the four and five cluster solution is that A-4 divides further into B-4 and B-5.
+This additional division makes a clearer clustering, because it resolves the heterogeneity in $moy$ creating a group (B-5, customers 1-3) which has a winter peak in usage, and a group (B-4, customers 16-20) which has a start of the year peak in usage.
+B-2 (customers 4-9) and B-1 (customers 21-24) have distinctive $hod$ patterns but are both heterogeneous in $moy$ and $wnwd$.
+B-3 (customers 10-15) has peak usage at the end of the year, but is heterogeneous on $hod$ and $wnwd$.
+This clustering almost agrees with the clusters visible in the t-SNE plot.
+
+<!-- A-4 is subdivided into B-4 and B-5, each of which has a distinct shape across $moy$ and $wnwd$.
 B-4 (id: 16-20) is distinguished by higher energy consumption in the first few months of the year and greater variability in weekend usage.
-B-1 (id: 1-3) has higher consumption in the middle of the year (winter months) and a similar weekday-weekend inter-quartile range.
-When $k = 4$ is used, these two groups merge to form A-1, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals.
+B-5 (id: 1-3) has higher consumption in the middle of the year (winter months) and a similar weekday-weekend inter-quartile range.
+When $k = 4$ is used, these two groups merge to form A-4, which has a $moy$ profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals.
 It may be worthwhile to compare Figures \ref{fig:opt-cluster-tsne-jsd} and \ref{fig:hod-ind-group-png} to see if the summarized distributions across groups accurately characterized the groupings.
-If it has, then the majority of the members of the group should have a similar profile.
-
-<!-- Our methodology is useful for grouping similar distributions over `hod` and `moy` and they are placed closely for easy comparison.
-Few groups have mixed patterns across `hod` and `moy`, but few have all customers in the group having a similar profile.
--->
+If it has, then the majority of the members of the group should have a similar profile.-->
+
+<!-- Our methodology is useful for grouping similar distributions over $hod$ and $moy$ and they are placed closely for easy comparison.
+Few groups have mixed patterns across $hod$ and $moy$, but few have all customers in the group having a similar profile.-->
 
@@
 
+```{r opt_clusters-wpd}
+```
+
+```{r summary-group-wpd}
+```
+
 ```{r tsne-fit}
@@
 
-```{r opt-cluster-tsne-jsd, fig.cap = "(a) shows t-SNE summary of the $24$ selected customers chosen through prototype selection, (b) and (c) shows a plot of sindex for different cluster size for clustering based on JS-based distances and wpd-distances respectively.
-The plot suggests optimal number of clusters as $4$ and $7$ respectively."}
+```{r all-data-rs}
+
+```
+
+```{r clustering-rs}
+
+```
+
+```{r opt-clusters-jsd-rs}
+
+```
+
+```{r opt-clusters-all}
+
+```
+
+```{r hod-ind-group}
+```
+
+```{r hod-ind-group-png, fig.cap = "The distribution of electricity demand across individual customers over three granularities $hod$, $moy$, and $wnwd$ are shown for the $24$ selected customers using quantile and box plots.
+They are split into batches of $12$ in (a) and (b), with each row in (a) or (b) representing a customer.
+The number indicates a unique customer id and a prototype id."}
+```
+
+```{r opt-cluster-tsne-jsd, fig.cap = "Clustering summaries: (a) t-SNE computed on the $24$ selected customers, and (b) separation index ($sindex$) for $2$-$10$ clusters using JS-NQT, JS-RS and WPD.
+Various choices in number of clusters would be recommended.
+Four clusters are visible in t-SNE, although it might hide a fifth cluster because dimension reduction to 2D may be insufficient to see the difference.
+JS-NQT suggests $3$, JS-RS suggests $2$ or $5$ and WPD suggests $3$ or $5$."}
 #include_graphics("img/opt_clust_parenthesis.png")
@@
 
-```{r hod-ind-group}
-```
-
-```{r hod-ind-group-png, fig.cap = "The distribution of electricity demand across individual customers over multiple granularities hod, moy, and wnwd are shown for the $24$ selected customers using quantile and box plots.
-They are split into batches of $12$ in (a) and (b), with each row in (a) or (b) representing a customer.
-Each customer is identified by their profile across all the granularities.
-A few of these customers have similar distributions across moy and some are similar in their hod distributions.
-Distributional differences across categories of wnwd are not discernable except for a couple of customers (id: 16 and 18)."}
-```
-
 <!-- It appears that the aim of grouping comparable distributions over considered variables has been accomplished to some extent. -->
@@
 
-```{r combined-groups-js, fig.cap = "The distribution of electricity demand across groups over hod (a), moy (b), and hod (conditional on wnwd) (c) are shown through quantile plots.
-Group 1 (1-3), Group 2 (4-9), Group 3 (10 - 15), Group 4 (16- 20), Group 5 (21 - 24).
-Groups 2 and 5 appear to have a hod pattern among its members (morning and evening peak), whereas groups 1, 3, and 5 appear to have a moy pattern (higher usage in mid-year).
-With their conditional dependencies on hod or moy categories, wnwd differences may not be discernible on their own.
-(c) demonstrates that the distribution of hod conditional on wnwd is also not different, except for the tails.
-This might be because most of these chosen customers do not have a wnwd pattern.
+```{r combined-groups-js, fig.cap = "The distribution of electricity demand across clusters over $hod$ (a), $moy$ (b), and $wnwd$ (c) are shown through quantile plots.
+Group 1 (customers 1-3), Group 2 (customers 4-9), Group 3 (customers 10-15), Group 4 (customers 16-20), Group 5 (customers 21-24).
+Groups 2 and 5 appear to have a $hod$ pattern among its members (morning and evening peak), whereas groups 1, 3, and 5 appear to have a $moy$ pattern (higher usage in mid-year).
+With their conditional dependencies on $hod$ or $moy$ categories, $wnwd$ differences may not be discernible on their own.
+(c) demonstrates that the distribution of $hod$ conditional on $wnwd$ is also not different, except for the tails.
+This might be because most of these chosen customers do not have a $wnwd$ pattern.
 It is useful to compare the raw distributions with the summarized distributions to establish that the majority of individuals in the group share the same characteristics.
@@
 
-```{r groups-4and5, fig.cap = "For $k = 4$ (a) and $k=5$ (b), the distribution of electricity demand across groups over hod, moy, and wnwd is shown.
-Groups A-2, A-3, and A-4 profiles correspond to Groups B-2, B-3, and B-5, respectively.
-A-1 is split into B-1 and B-4, each of which has a distinct shape across moy and wnwd.
-B-4 (id: 16-20) is characterized by higher energy consumption in the first few months of the year and more variation in weekend usage.
-B-1 (id:1-3) is distinguished by higher consumption in the middle of the year (winter months) and similar weekday-weekend inter-quartile range.
-When $k = 4$ is used, these two groups merge to form A-1, which has a moy profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals in the group.", fig.height=6}
-
-```
-
-#### wpd-based distances
-
-In Figure \ref{fig:opt-cluster-tsne-jsd}, the $sindex$ is plotted against the number of clusters for clustering based on wpd-based distances.
+```{r groups-4and5, fig.cap = "Summary plots for four and five clusters from JS-NQT showing the distribution of electricity demand combined for all members over $hod$, $moy$, and $wnwd$.
+Groups A-1 (customers 21-24), A-2 (customers 4-9), and A-3 (customers 10-15) profiles correspond to Groups B-1, B-2, and B-3, respectively.
+Cluster A-4 splits into B-4 (customers 16-20) and B-5 (customers 1-3) to produce the five clusters, which better resolves the $moy$ distribution.", fig.height=5}
+
+```
+
+```{r summary-plot-wpd, fig.cap = "Summary plots for three (a) and five (b) clusters from WPD approach showing the $wpd$ values of each customers across $hod$, $moy$, and $wnwd$ through a parallel coordinate plot.
+P-1 and P-2 groups correspond to Q-1 and Q-2, respectively.
+Cluster P-3 is subdivided into Q-3, Q-4, and Q-5.
+P-1 (customers 16, 18) is distinguished by high $wpd$ on $wnwd$ values.
+P-2 has lower $wpd$ than $moy$ and $wnwd$ for $hod$.
+P-3 operates in the opposite way as P-2, with larger $wpd$ for $hod$ in comparison to $moy$ and $wnwd$.
+For the $5$ cluster solution, this group is divided into Q3-Q5, which are distinct due to their different relative significance of $moy$ and $wnwd$.", fig.height=6}
+```
+
+```{r summary-table-wpd}
+
+```
+
+<!--### WPD-->
+
+<!-- In Figure \ref{fig:opt-cluster-tsne-jsd}, the $sindex$ is plotted against the number of clusters for clustering based on wpd-based distances.
 The scales are very different from those used in JS-based clustering.
@@
 Thus, there are few choices available for the number of clusters.
-The groupings with $k=4$ is shown through a parallel coordinate plot with the three significant cyclic granularities.
+Figure \ref{fig:summary-plot-wpd} whos the wpd values for the three different granularities of the 24 customers, coloured by the 6 cluster solution.
+XXX REWORK THIS WHEN NEW CLUSTER RESULTS ARE SHOWN<!--The groupings with $k=4$ is shown through a parallel coordinate plot with the three significant cyclic granularities.
 The variables are sorted according to their separation across classes (rather than their overall variation between classes).
@@
 The two customers (id: 16 and 18) that have significant differences between $wnwd$ categories (in Figure \ref{fig:hod-ind-group-png}) stand out from the rest of the customers.
-Group-4 has a higher $wpd$ for `hod` than `moy` or `wkndwday`.
-Group-2 and Group-3 have the most distinct patterns across `moy`.
+Group-4 has a higher $wpd$ for $hod$ than $moy$ or $wnwd$.
+Group-2 and Group-3 have the most distinct patterns across $moy$.
 Group-1 is a mixed group that has strong patterns on at least one of the three variables.
-When $k=3$ is used, Group-2 and Group-3 are merged because their patterns relative to `moy`, `hod` and `wnwd` are the same.
-The findings vary from JS-based clustering, yet it is a helpful grouping.
+When $k=3$ is used, Group-2 and Group-3 are merged because their patterns relative to $moy$, $hod$ and $wnwd$ are the same.
+The findings vary from JS-based clustering, yet it is a helpful grouping.-->
+
+Figure \ref{fig:summary-plot-wpd} shows the $wpd$ values of the $24$ customers over $hod$, $moy$, and $wnwd$ , colored by $3$ (a) and $5$ (b) clusters from WPD clustering using a parallel coordinate plot.
+The variables ($wpd$ for different granularities) are
+standardized prior to clustering using WPD.
+In the display, the variables are sorted according to their separation across groups.
+This means that $wnwd$ is the most important variable in distinguishing the groups, followed by $hod$ and $moy$ for both (a) and (b).
+Groups P-1 and P-2 correspond to Q-1 and Q-2 respectively.
+Cluster P-3 splits into Q-3, Q-4 and Q-5.
+Customers 16 and 18 are characterized by unusual high values of $wpd$ on $wnwd$ compared to the rest of the customers and hence form a group.
+This could again be verified from Figure \ref{fig:hod-ind-group-png}, where these were the only two customers with a difference in their $wnwd$ behavior.
+They are represented by P-1.
+P-2 has lower $wpd$ for $hod$ than $moy$ and $wnwd$.
+P-3 behaves opposite to P-2 with higher $wpd$ for $hod$ compared to $moy$ and $wnwd$.
+These are the customers who have some significant pattern across $hod$ and this can again be validated by looking at Figure \ref{fig:hod-ind-group-png}.
+For a $5$ cluster solution, this group gets split into Q-3, Q-4 and Q-5 characterizing different relative significance of $moy$ and $wnwd$.
+For example, Q-4 and Q-5 have almost no pattern across $moy$, but Q-3 has a $moy$ pattern and thus it is reasonable to split them.
+The patterns could be different, but they are significant.
+Q-4 and Q-5 are separated because of their different significance of $hod$.
+All of these can also be verified from Table \ref{tab:summary-table-wpd} which shows the cluster summaries with members and median values of $wpd$ for the three variables.
+
+<!-- $hod$ has very low median values of $wpd$ for the members in the cluster, implying at least half of these customers do not have an interesting $hod$ pattern. -->
 
@@
 
-```{r parcoord-application-gracsr, fig.cap="Each of the $24$ customers is represented by a parallel coordinate plot (a) with four wpd-based groupings.
-Plot (b) which shows median $wpd$ values for each group.
-The plot shows that moy is the most important variable in identifying clusters, whereas wnwd is the least significant and has the least fluctuation.
-Two customers (id: 16 and 18) with high $wpd$ across wnwd stand out in this display.
-Group-4 has a higher $wpd$ for hod than moy or wnwd.
-Group-2 and Group-3 have the most discernible pattern across moy.
-Group-1 is a mixed group with strong patterns across at least one of the three variables.
-All of these could be observed from the plot or the table (b).", message=FALSE, warning=FALSE, fig.height=6}
-#include_graphics("img/parcoord-application.png")
-```
-
-Things become far more complicated when we consider a larger data set with more uncertainty, as they do with any clustering problem.
+<!-- # ```{r summary-plot-wpd, fig.cap="Summary of the four cluster results from WPD, shown as a parallel coordinate plot and table.
+XXX TO BE REVISED WHEN DIFFERENT CLUSTER SOLUTION IS SHOWN", message=FALSE, warning=FALSE, fig.height=6} -->
+<!-- # #include_graphics("img/summary-plot-wpd.png") -->
+<!-- # ``` -->
+
+In summary, none of the methods captured the five original prototypes exactly.
+JS-NQT was almost identical, though, but WPD produced quite a different grouping.
+This is quite a reasonable result and illustrates both the difficulties of clustering to obtain a particularly expected grouping and the ability to learn unexpected patterns in the data.
+It is possible that the JS-based distances were distracted by the presence of nuisance variables, levels of the granularities that do not contribute to clustering.
+This would also be supported by the results of the validation study, where clustering was less effective in S2 and S3, where only some granularities had differences between levels.
+Clustering using WPD is expected to produce quite different results because it will group only by overall value of a granularity, not a particular pattern.
+A cluster summary like Figure \ref{fig:groups-4and5} is not possible because there may be different but equally interesting patterns (e.g.
+high evening $hod$ and high daytime $hod$) in the same cluster.
+Simply it provides information that across a collection of customers this specific cluster has interesting patterns in a granularity (e.g.
+$hod$).
+One would need to post-process the cluster to separate specific patterns.
+
+<!--Things become far more complicated when we consider a larger data set with more uncertainty, as they do with any clustering problem.
 Summarizing distributions across clusters with varied or outlying customers can result in a shape that does not represent the group.
@@
 If more extensive prototype selection is used, resulting in more comprehensive prototypes in the data set, this method might be used to classify the entire data set into these prototype behaviors.
-However, the goal of this section was to have a few customers that have significant patterns over one or more cyclic granularities, apply our methodology to cluster them, and demonstrate that the method produces useful clusters.
+However, the goal of this section was to have a few customers that have significant patterns over one or more cyclic granularities, apply our methodology to cluster them, and demonstrate that the method produces useful clusters.-->
 
@@
 The Monash eResearch Centre and eSolutions-Study Support Services supported this research in part through the resource usage of the MonARCH HPC Cluster.
-The Github repository, [github.com/Sayani07/paper-gracsr](https://github.com/Sayani07/paper-gracsr), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
+The Github repository, [github.com/Sayani07/paper-gracsr](https://github.com/Sayani07/paper-gracsr), contains all materials required to reproduce this article and the code is also available online in the supplementary materials.
 This article was created with R [@R-language], `knitr` [@knitr; @R-knitr] and `rmarkdown` [@rmarkdown; @R-rmarkdown].
@@
 
-## Supplementary Materials
-
-**Data and scripts:** Data sets and R code to reproduce all figures in this article (main.R).
-
-**Supplementary paper:** Additional tables, graphics and R code to reproduce it (paper-supplementary.pdf, paper-supplementary.Rmd).
-The code for creating validation designs and running the methodologies is available at (https://github.com/Sayani07/paper-gracsR/Validation).
-
-**R-package:** To implement the ideas provided in this research, the open-source R package 'gracsr' is available on Github (https://github.com/Sayani07/gracsr).
+## Supplementary materials {- #supplement-gracsr}
+
+**Data and scripts:** Data sets and R codes to reproduce all figures in this article  
+(validation.R, application.R)
+
+\noindent
+**Supplementary paper:** Additional graphics and R codes to reproduce it  
+(paper-supplementary.pdf, paper-supplementary.Rmd, supplementary.R)
+
+\noindent
+**R-package:** To implement the ideas provided in this research, the open-source R package `gracsr` is available on Github (https://github.com/Sayani07/gracsr).

### 2021-11-18 Rmd/05-conclusion.Rmd
no section supplementary

@@
 
-We address few limitations of the current framework that might serve as natural next steps for this work and some potential short and long-term aims in future ideas.
+We address few limitations of the current framework that might serve as natural next steps for this work and some potential short- and long-term aims in future ideas.
 
@@
 
-s
-
 <!-- ### Comparing our clustering method with other benchmark methods -->

### 2021-11-18 scripts/gracsr/application.R

@@ -35,7 +35,7 @@ data_pick_cust <- bind_rows(
   mutate(customer_id = as.character(customer_id))
 
 ## ----all-data-----------------------------------------------------------------
-data_pick <- read_rds(here::here("data/gracsr/elec_nogap_2013_clean_356cust.rds")) %>%
+data_pick <- readr::read_rds(here::here("data/gracsr/elec_nogap_2013_clean_356cust.rds")) %>%
   mutate(customer_id = as.character(customer_id)) %>%
   dplyr::filter(customer_id %in% data_pick_cust$customer_id) %>%
   gracsr::scale_gran(
@@ -69,7 +69,7 @@ distance_jsd <- as.dist(distance)
 
 ## ----opt-clusters-jsd---------------------------------------------------------
 
-all_index <- map_dfr(2:10, function(x) {
+all_index_nqt <- map_dfr(2:10, function(x) {
   group <- distance_jsd %>%
     hclust(method = "ward.D") %>%
     cutree(k = x)
@@ -78,22 +78,23 @@ all_index <- map_dfr(2:10, function(x) {
     k = x,
     sindex = p$sindex
   )
-})
+})%>% mutate(method = "JS-NQT")
 
-opt_clusters <- all_index %>% ggplot(aes(
-  x = k,
-  y = sindex
-)) +
-  geom_line() +
-  scale_x_continuous(
-    breaks = seq(2, 10, 1),
-    minor_breaks = 1
-  ) +
-  theme_bw() +
-  ylab("sindex") +
-  xlab("number of clusters") +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8))
-
+# 
+# opt_clusters <- all_index %>% ggplot(aes(
+#   x = k,
+#   y = sindex
+# )) +
+#   geom_line() +
+#   scale_x_continuous(
+#     breaks = seq(2, 10, 1),
+#     minor_breaks = 1
+#   ) +
+#   theme_bw() +
+#   ylab("sindex") +
+#   xlab("number of clusters") +
+#   theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8))
+# 
 
 
 ## ----groups-24----------------------------------------------------------------
@@ -106,8 +107,14 @@ cluster_result <- suppressMessages(distance_jsd %>%
 cluster_result_id <- cluster_result %>%
[…]

### 2021-11-18 scripts/gracsr/library.R

@@ -1,0 +1,36 @@
+## ---- load-lib
+
+#theme_set(theme_bw())
+library(ggplot2)
+library(lubridate)
+library(gravitas)
+library(gracsr)
+library(ggdendro)
+library(dplyr)
+library(readr)
+library(visdat)
+library(ggplot2)
+library(tidyverse)
+library(naniar)
+library(here)
+library(tsibble)
+library(knitr)
+library(patchwork)
+library(GGally)
+library(distributional)
+library(viridis)
+library(forcats)
+library(tidyr)
+library(purrr)
+library(stringr)
+library(pals)
+library(Rtsne)
+library(fpc)
+library(cluster)
+library(ggpubr)
+library(grid)
+library(ggrepel)
+library(kableExtra)
+library(rlang)
+
+

### 2021-11-18 scripts/gracsr/mylibrary.R

@@ -1,34 +1,0 @@
-## ---- load-lib
-
-#theme_set(theme_bw())
-library(ggplot2)
-library(lubridate)
-library(gravitas)
-library(gracsr)
-library(ggdendro)
-library(dplyr)
-library(readr)
-library(visdat)
-library(ggplot2)
-library(tidyverse)
-library(naniar)
-library(here)
-library(tsibble)
-library(knitr)
-library(patchwork)
-library(GGally)
-library(distributional)
-library(viridis)
-library(forcats)
-library(tidyr)
-library(purrr)
-library(stringr)
-library(pals)
-library(Rtsne)
-library(fpc)
-library(cluster)
-library(ggpubr)
-library(grid)
-library(ggrepel)
-
-

### 2021-11-18 scripts/gracsr/mytheme.R

@@ -1,98 +1,0 @@
-##----theme-validation--------------------------------------------------------------------
-theme_validation <- function() {
-  theme_bw() +
-    theme(
-      strip.text = element_text(size = 8, 
-                                margin = margin(b = 0, t = 1)),
-      plot.margin = margin(0, 0, 0, 0, "cm"),
-      axis.title.y = element_blank(),
-      #axis.text.x = element_blank(),
-      axis.text.y = element_blank(),
-      axis.ticks = element_blank(),
-      legend.position = "bottom"
-    )
-}
-
-
-##----theme_characterisation--------------------------------------------------------------
-
-theme_characterisation <- function() {
-  
-  theme_bw() + # seeting theme
-    theme(strip.text = element_text(size = 10,
-                                    margin = margin(b = 0, t = 0))) +
-    # narrow facet space
-    theme(axis.title.y=element_blank(),
-          axis.text.y=element_blank(),
-          axis.ticks.y=element_blank()) + # no axis ticks 
-    theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
-    theme(axis.text.x = element_text(angle=90, hjust=1, size = 10)) + 
-    # rotate the x-axis text
-    theme(legend.position = "bottom")+
-    theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
-    theme(axis.text.x = element_text(size=5)) +
-    theme(axis.text = element_text(size = 12))  
-}
-
-
-##----mytheme-application----------------------------------------------------------------
-theme_application <- function() {
-  
-  theme_light() + # setting theme
-    theme(strip.text = element_text(size = 10,
-                                    margin = margin(b = 0, t = 0)), 
-          strip.text.x = element_text(size = 8)) + # narrow facet space
-    theme(axis.title.y=element_blank(),
-          axis.text.y=element_blank(),
-          axis.ticks.y=element_blank()) + # no axis ticks 
-    theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
-    theme(axis.ticks.x = element_blank(),
-          axis.text.x = element_blank()) + # no x-axis labels to further reduce the gap between facets
-    #theme(axis.text.x = element_text(angle=90, hjust=1, size = 9)) + # rotate the x-axis text
-    theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
-    #theme(axis.text.x = element_text(size=5)) +
-    theme(strip.background = element_blank(),
[…]

### 2021-11-18 scripts/gracsr/theme.R

@@ -1,0 +1,98 @@
+##----theme-validation--------------------------------------------------------------------
+theme_validation <- function() {
+  theme_bw() +
+    theme(
+      strip.text = element_text(size = 8, 
+                                margin = margin(b = 0, t = 1)),
+      plot.margin = margin(0, 0, 0, 0, "cm"),
+      axis.title.y = element_blank(),
+      #axis.text.x = element_blank(),
+      axis.text.y = element_blank(),
+      axis.ticks = element_blank(),
+      legend.position = "bottom"
+    )
+}
+
+
+##----theme_characterisation--------------------------------------------------------------
+
+theme_characterisation <- function() {
+  
+  theme_bw() + # seeting theme
+    theme(strip.text = element_text(size = 10,
+                                    margin = margin(b = 0, t = 0))) +
+    # narrow facet space
+    theme(axis.title.y=element_blank(),
+          axis.text.y=element_blank(),
+          axis.ticks.y=element_blank()) + # no axis ticks 
+    theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
+    theme(axis.text.x = element_text(angle=90, hjust=1, size = 10)) + 
+    # rotate the x-axis text
+    theme(legend.position = "bottom")+
+    theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
+    theme(axis.text.x = element_text(size=8)) +
+    theme(axis.text = element_text(size = 8))  
+}
+
+
+##----mytheme-application----------------------------------------------------------------
+theme_application <- function() {
+  
+  theme_light() + # setting theme
+    theme(strip.text = element_text(size = 10,
+                                    margin = margin(b = 0, t = 0)), 
+          strip.text.x = element_text(size = 8)) + # narrow facet space
+    theme(axis.title.y=element_blank(),
+          axis.text.y=element_blank(),
+          axis.ticks.y=element_blank()) + # no axis ticks 
+    theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
+    theme(axis.ticks.x = element_blank(),
+          axis.text.x = element_blank()) + # no x-axis labels to further reduce the gap between facets
+    #theme(axis.text.x = element_text(angle=90, hjust=1, size = 9)) + # rotate the x-axis text
+    theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
+    #theme(axis.text.x = element_text(size=5)) +
+    theme(strip.background = element_blank(),
[…]

### 2021-11-18 scripts/gracsr/validation.R

@@ -31,7 +31,8 @@ knitr::kable(
   list(tab_dist, tibble(design, g1, g2, g3)),
   caption = "For S1, distributions of different categories when they vary (displayed on top). If distributions are fixed, they are set to N(0, 1). The various distributions across categories result in five designs (displayed below).",
   booktabs = TRUE, valign = "t"
-)
+) %>%
+  kable_styling(latex_options = "hold_position")
 
 
 
@@ -1022,6 +1023,46 @@ mds_plot_10 <- ggplot(
   scale_color_brewer(palette = "Dark2") +
   ylab("mds1") +
   xlab("mds2") +
+  theme(axis.text = element_text(size = 8))+
   theme(aspect.ratio = 1)
 
 mds_plot_10
+
+
+##----wpd-sindex-validation----------------------------------------
+data1 <- sindex_data(read_rds("validation/wpd/3gran_change_5D/data_validation_25.rds")) %>% mutate(diff = 1, design = "S1")
+data2 <- sindex_data(read_rds("validation/wpd/3gran_change_5D/data_validation_26.rds")) %>% mutate(diff = 2, design = "S1")
+data3 <- sindex_data(read_rds("validation/wpd/3gran_change_5D/data_validation_27.rds")) %>% mutate(diff = 5, design = "S1")
+
+data4 <- sindex_data(read_rds("validation/wpd/2gran_change_4D/data_validation_7.rds")) %>% mutate(diff = 1, design = "S2")
+data5 <- sindex_data(read_rds("validation/wpd/2gran_change_4D/data_validation_17.rds")) %>% mutate(diff = 2, design = "S2")
+data6 <- sindex_data(read_rds("validation/wpd/2gran_change_4D/data_validation_9.rds")) %>% mutate(diff = 5, design = "S2")
+
+
+data7 <- sindex_data(read_rds("validation/wpd/1gran_change_5D/data_validation_16.rds")) %>% mutate(diff = 1, design = "S3")
+data8 <- sindex_data(read_rds("validation/wpd/1gran_change_5D/data_validation_17.rds")) %>% mutate(diff = 2, design = "S3")
+data9 <- sindex_data(read_rds("validation/wpd/1gran_change_5D/data_validation_18.rds")) %>% mutate(diff = 5, design = "S3")
+
+all_data <- bind_rows(data1, data2, data3, data4, data5, data6, data7, data8, data9) %>% rename("Scenario" = "design")
+
+
+
[…]

### 2021-11-19 Rmd/04-gracsr.Rmd
Typos and consistent notation

@@
 
-The primary application of this work is data from the Smart Grid, Smart City (SGSC) project (2010–2014) available through the [Department of the Environment and Energy](https://data.gov.au/data/organization/doee).
-Half-hourly measurements of usage for more than 13,000 electricity smart meter customers is provided from October 2011 to March 2014.
+The primary application of this work is data from the Smart Grid, Smart City (SGSC) project (2010–2014) available through @smart-meter.
+Half-hourly measurements of usage for more than 13,000 electricity smart meter customers are provided from October 2011 to March 2014.
 <!--Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.--> Customers vary in size, location, and amenities such as solar panels, central heating, and air conditioning.
@@
 
-@Gupta2021-hakear provide a distance measure ($wpd$) for determining the significance of a cyclic granularity, and a ranking of multiple cyclic granularities.
+@Gupta2021-hakear provide a distance measure ($\wpd$) for determining the significance of a cyclic granularity, and a ranking of multiple cyclic granularities.
 (This extends to harmonies, pairs of granularities that might interact with each other.) We define "significant" granularities as those with significant distributional differences across at least one category.
@@
 Statistical transformations are employed to bring all of them to the same range or normalize each observation.
-This is important because we are not interested in trivial clusters that vary in magnitudes, but rather in uncovering comparable patterns of distributional differences between categories.--> For the JS-based approaches, two data transformation techniques are recommended, normal-quantile transform (NQT) and robust scaling (RS).<!-- NQT is used in @Gupta2021-hakear prior sorting granularities.-- $wpd$, which is the foundation of wpd-based distances.
-XX wpd was not yet introduced--> While @Gupta2021-hakear already use NQT when computing $wpd$, it could be useful to standardize it for the selected set of significant granularities prior to computing the distances.
+This is important because we are not interested in trivial clusters that vary in magnitudes, but rather in uncovering comparable patterns of distributional differences between categories.--> For the JS-based approaches, two data transformation techniques are recommended, normal-quantile transform (NQT) and robust scaling (RS).<!-- NQT is used in @Gupta2021-hakear prior sorting granularities.-- $\wpd$, which is the foundation of wpd-based distances.
+XX wpd was not yet introduced--> While @Gupta2021-hakear already use NQT when computing $\wpd$, it could be useful to standardize it for the selected set of significant granularities prior to computing the distances.
 
@@
 
-We compute weighted pairwise distances $wpd$ [@Gupta2021-hakear] for all considered granularities for all observations.
-$wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction.
+We compute weighted pairwise distances $\wpd$ [@Gupta2021-hakear] for all considered granularities for all observations.
+$\wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction.
 It is estimated by the maximum pairwise distances between distributions across consecutive categories normalized by appropriate parameters.
-A higher value of $wpd$ indicates that some interesting patterns are expected, whereas a lower value would indicate otherwise.
-
-Once we have chosen $wpd$ as a relevant feature for characterizing the distributions across one cyclic granularity, we have to decide how we combine differences between the multiple features (corresponding to multiple granularities) into a single number.
-The Euclidean distance between them is chosen, with the granularities acting as variables and $wpd$ representing the value under each variable.
-With this approach, we should expect the observations with similar $wpd$ values to be clustered together.
+A higher value of $\wpd$ indicates that some interesting patterns are expected, whereas a lower value would indicate otherwise.
+
+Once we have chosen $\wpd$ as a relevant feature for characterizing the distributions across one cyclic granularity, we have to decide how we combine differences between the multiple features (corresponding to multiple granularities) into a single number.
+The Euclidean distance between them is chosen, with the granularities acting as variables and $\wpd$ representing the value under each variable.
+With this approach, we should expect the observations with similar $\wpd$ values to be clustered together.
 Thus, this approach is useful for grouping observations that have a similar significance of patterns across different granularities.
@@
 If we have $H_{N_C}$ harmonies in the harmony table, then for each household we have a vector of $wpd_{{s}}$ of $H_{N_C}$ elements with each element corresponding to one harmony.
-We aim to have pockets of customers showing similar periodic behavior by considering $wpd$ values for different harmonies and some time series features.
+We aim to have pockets of customers showing similar periodic behavior by considering $\wpd$ values for different harmonies and some time series features.
 The features should also characterize probability distributions of different household.
@@
 Most metrics for choosing the optimal number of clusters are based on comparing distances between observations within a class to those distances between observations between classes, which makes the assumption that there are some separated clusters.
-Some common procedures include the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@Dunn1973-gk] and the separation index ($sindex$) [@Hennig2019-eq; @Hennig2014-ah].
+Some common procedures include the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@Dunn1973-gk] and the separation index (*sindex*) [@Hennig2019-eq; @Hennig2014-ah].
 These are constructed by balancing within-cluster homogeneity and between-cluster separation.
@@
 
-In this work, we have chosen to use $sindex$.
+In this work, we have chosen to use *sindex*.
 It is a very simple but effective metric.
@@
 It is relatively robust to nuisance observations.
-The value of $sindex$ always decreases, and sharp drops in value indicate candidates for the optimal number of clusters.
+The value of *sindex* always decreases, and sharp drops in value indicate candidates for the optimal number of clusters.
 The number of clusters corresponding to the value **before the drop** is the recommendation.
@@
 
-Figure \ref{fig:sindex-plot-validation} shows $sindex$ plotted against the number of clusters ($k$) for the range of mean differences (rows) under the different scenarios (columns).
+Figure \ref{fig:sindex-plot-validation} shows *sindex* plotted against the number of clusters ($k$) for the range of mean differences (rows) under the different scenarios (columns).
 This can be used to determine the number of clusters for each scenario.
-When $sindex$ for each scenario are examined, it appears that $k = \{5, 4, 4\}$ is justified for scenarios S1, S2, and S3, respectively, given the sharp decrease in $sindex$ from that value of $k$.
+When *sindex* for each scenario are examined, it appears that $k = \{5, 4, 4\}$ is justified for scenarios S1, S2, and S3, respectively, given the sharp decrease in *sindex* from that value of $k$.
 Thus, the number of clusters corresponds to the number of designs that were originally considered in each scenario.
@@
 
-```{r mds-plot-validation, fig.cap = "MDS summary plots to illustrate the cluster sparation for the range of mean differences (rows) under the different scenarios (columns).
+```{r mds-plot-validation, fig.cap = "MDS summary plots to illustrate the cluster separation for the range of mean differences (rows) under the different scenarios (columns).
 It can be observed that clusters become more compact and separated for higher mean differences between categories across all scenarios.
@@
 For this reason, we have chosen to select a small group of customers with relatively distinct and different patterns in order to illustrate the clustering more simply.
-Figure \ref{fig:hod-ind-group-png} shows the distribution across $hod$, $moy$ and $wnwd$ for the set of $24$ customers used to illustrate clustering.
+Figure \ref{fig:hod-ind-group-png} shows the distribution across *hod*, *moy* and *wnwd* for the set of $24$ customers used to illustrate clustering.
 The customers are displayed in two columns of $12$ for space reasons.
@@
 2.
-Obtain $wpd$ for all cyclic granularities considered for these customers.
-It was found that $hod$ (hour-of-day), $moy$ (month-of-year) and $wnwd$ (weekend/weekday) are significant for most customers.
+Obtain $\wpd$ for all cyclic granularities considered for these customers.
+It was found that *hod* (hour-of-day), *moy* (month-of-year) and *wnwd* (weekend/weekday) are significant for most customers.
 We use these three granularities while clustering.
 3.
-Remove customers whose data for an entire category of $hod$, $moy$ or $wnwd$ is empty.
+Remove customers whose data for an entire category of *hod*, *moy* or *wnwd* is empty.
 For example, a customer who does not have data for an entire month is excluded because their monthly behavior cannot be analyzed.
@@
 This ensured that the final sample of clustered customers were also far apart in the high-dimensional space.
-(See the supplementary materials for further details.)<!--Few of these customers have similar distribution across $moy$ and some are similar in their $hod$ distribution.
+(See the supplementary materials for further details.)<!--Few of these customers have similar distribution across *moy* and some are similar in their *hod* distribution.
 -->
@@
 <!-- # ```{r hod-moy-wkndwday plots, fig.cap="The distribution across moy, hod and wkndwday for the selected designs.
-Few are similar in their hod pattern, while others are similar in $moy$ behavior.
+Few are similar in their hod pattern, while others are similar in *moy* behavior.
 Some customers have distinct behavior as compared to all other customers.For example, although patterns across wkndwday do not look distinctly different for most customers, there is one household for whom weekend behavior is standing out from the rest."} -->
@@
 It is possible that because the representation is only 2D, the fifth group from the original prototype selection is distinctly different in high dimensions.
-The $sindex$ plots for the three methods indicate some disagreement: JS-NQT suggests $3$, JS-RS suggests $2$ or $5$ and WPD suggests $3$ or $5$.
+The *sindex* plots for the three methods indicate some disagreement: JS-NQT suggests $3$, JS-RS suggests $2$ or $5$ and WPD suggests $3$ or $5$.
 JS-RS would appear to match the original prototypes with the five cluster solution, but it actually differs.
-Even though the $sindex$ for JS-NQT suggests three clusters, the five cluster solution more closely matches the original prototypes.
+Even though the *sindex* for JS-NQT suggests three clusters, the five cluster solution more closely matches the original prototypes.
 The WPD clustering provides a different grouping of the customers, and even though it disagrees with the original prototypes it is a useful grouping.<!--While plotting the clusters, RS-based scaling is applied to each customer, so that no customer dominates the shape of the summarized clusters because of magnitude.
@@
 
-<!--Figure \ref{fig:opt-cluster-tsne-jsd} The $sindex$ is plotted against the number of clusters for clustering based on JS-based (JS-NQT) distances in Figure \ref{fig:opt-cluster-tsne-jsd}.
-The $sindex$ falls from around $0.56$ when $k=3$ to $0.54$ when $k$ increases.
+<!--Figure \ref{fig:opt-cluster-tsne-jsd} The *sindex* is plotted against the number of clusters for clustering based on JS-based (JS-NQT) distances in Figure \ref{fig:opt-cluster-tsne-jsd}.
+The *sindex* falls from around $0.56$ when $k=3$ to $0.54$ when $k$ increases.
 In our study, we use $k=4, 5$.--> Figure \ref{fig:groups-4and5} displays the summarized distributions across $4$ and $5$ clusters from JS-NQT clustering in (a) and (b) respectively, and helps to characterize each cluster.
@@
 The only difference between the four and five cluster solution is that A-4 divides further into B-4 and B-5.
-This additional division makes a clearer clustering, because it resolves the heterogeneity in $moy$ creating a group (B-5, customers 1-3) which has a winter peak in usage, and a group (B-4, customers 16-20) which has a start of the year peak in usage.
-B-2 (customers 4-9) and B-1 (customers 21-24) have distinctive $hod$ patterns but are both heterogeneous in $moy$ and $wnwd$.
-B-3 (customers 10-15) has peak usage at the end of the year, but is heterogeneous on $hod$ and $wnwd$.
+This additional division makes a clearer clustering, because it resolves the heterogeneity in *moy* creating a group (B-5, customers 1-3) which has a winter peak in usage, and a group (B-4, customers 16-20) which has a start of the year peak in usage.
+B-2 (customers 4-9) and B-1 (customers 21-24) have distinctive *hod* patterns but are both heterogeneous in *moy* and *wnwd*.
+B-3 (customers 10-15) has peak usage at the end of the year, but is heterogeneous on *hod* and *wnwd*.
 This clustering almost agrees with the clusters visible in the t-SNE plot.
 
-<!-- A-4 is subdivided into B-4 and B-5, each of which has a distinct shape across $moy$ and $wnwd$.
+<!-- A-4 is subdivided into B-4 and B-5, each of which has a distinct shape across *moy* and *wnwd*.
 B-4 (id: 16-20) is distinguished by higher energy consumption in the first few months of the year and greater variability in weekend usage.
 B-5 (id: 1-3) has higher consumption in the middle of the year (winter months) and a similar weekday-weekend inter-quartile range.
-When $k = 4$ is used, these two groups merge to form A-4, which has a $moy$ profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals.
+When $k = 4$ is used, these two groups merge to form A-4, which has a *moy* profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals.
 It may be worthwhile to compare Figures \ref{fig:opt-cluster-tsne-jsd} and \ref{fig:hod-ind-group-png} to see if the summarized distributions across groups accurately characterized the groupings.
@@
 
-<!-- Our methodology is useful for grouping similar distributions over $hod$ and $moy$ and they are placed closely for easy comparison.
-Few groups have mixed patterns across $hod$ and $moy$, but few have all customers in the group having a similar profile.-->
+<!-- Our methodology is useful for grouping similar distributions over *hod* and *moy* and they are placed closely for easy comparison.
+Few groups have mixed patterns across *hod* and *moy*, but few have all customers in the group having a similar profile.-->
 
@@
 
-```{r hod-ind-group-png, fig.cap = "The distribution of electricity demand across individual customers over three granularities $hod$, $moy$, and $wnwd$ are shown for the $24$ selected customers using quantile and box plots.
+```{r hod-ind-group-png, fig.cap = "The distribution of electricity demand across individual customers over three granularities *hod*, *moy*, and *wnwd* are shown for the $24$ selected customers using quantile and box plots.
 They are split into batches of $12$ in (a) and (b), with each row in (a) or (b) representing a customer.
@@
 
-```{r opt-cluster-tsne-jsd, fig.cap = "Clustering summaries: (a) t-SNE computed on the $24$ selected customers, and (b) separation index ($sindex$) for $2$-$10$ clusters using JS-NQT, JS-RS and WPD.
+```{r opt-cluster-tsne-jsd, fig.cap = "Clustering summaries: (a) t-SNE computed on the $24$ selected customers, and (b) separation index (*sindex*) for $2$-$10$ clusters using JS-NQT, JS-RS and WPD.
 Various choices in number of clusters would be recommended.
@@
 
-```{r combined-groups-js, fig.cap = "The distribution of electricity demand across clusters over $hod$ (a), $moy$ (b), and $wnwd$ (c) are shown through quantile plots.
+```{r combined-groups-js, fig.cap = "The distribution of electricity demand across clusters over *hod* (a), *moy* (b), and *wnwd* (c) are shown through quantile plots.
 Group 1 (customers 1-3), Group 2 (customers 4-9), Group 3 (customers 10-15), Group 4 (customers 16-20), Group 5 (customers 21-24).
-Groups 2 and 5 appear to have a $hod$ pattern among its members (morning and evening peak), whereas groups 1, 3, and 5 appear to have a $moy$ pattern (higher usage in mid-year).
-With their conditional dependencies on $hod$ or $moy$ categories, $wnwd$ differences may not be discernible on their own.
-(c) demonstrates that the distribution of $hod$ conditional on $wnwd$ is also not different, except for the tails.
-This might be because most of these chosen customers do not have a $wnwd$ pattern.
+Groups 2 and 5 appear to have a *hod* pattern among its members (morning and evening peak), whereas groups 1, 3, and 5 appear to have a *moy* pattern (higher usage in mid-year).
+With their conditional dependencies on *hod* or *moy* categories, *wnwd* differences may not be discernible on their own.
+(c) demonstrates that the distribution of *hod* conditional on *wnwd* is also not different, except for the tails.
+This might be because most of these chosen customers do not have a *wnwd* pattern.
 It is useful to compare the raw distributions with the summarized distributions to establish that the majority of individuals in the group share the same characteristics.
@@
 
-```{r groups-4and5, fig.cap = "Summary plots for four and five clusters from JS-NQT showing the distribution of electricity demand combined for all members over $hod$, $moy$, and $wnwd$.
+```{r groups-4and5, fig.cap = "Summary plots for four and five clusters from JS-NQT showing the distribution of electricity demand combined for all members over *hod*, *moy*, and *wnwd*.
 Groups A-1 (customers 21-24), A-2 (customers 4-9), and A-3 (customers 10-15) profiles correspond to Groups B-1, B-2, and B-3, respectively.
-Cluster A-4 splits into B-4 (customers 16-20) and B-5 (customers 1-3) to produce the five clusters, which better resolves the $moy$ distribution.", fig.height=5}
-
-```
-
-```{r summary-plot-wpd, fig.cap = "Summary plots for three (a) and five (b) clusters from WPD approach showing the $wpd$ values of each customers across $hod$, $moy$, and $wnwd$ through a parallel coordinate plot.
+Cluster A-4 splits into B-4 (customers 16-20) and B-5 (customers 1-3) to produce the five clusters, which better resolves the *moy* distribution.", fig.height=5}
+
+```
+
+(ref:summary-plot-wpd) "Summary plots for three (a) and five (b) clusters from WPD approach showing the $\wpd$ values of each customers across *hod*, *moy*, and *wnwd* through a parallel coordinate plot.
 P-1 and P-2 groups correspond to Q-1 and Q-2, respectively.
 Cluster P-3 is subdivided into Q-3, Q-4, and Q-5.
-P-1 (customers 16, 18) is distinguished by high $wpd$ on $wnwd$ values.
-P-2 has lower $wpd$ than $moy$ and $wnwd$ for $hod$.
-P-3 operates in the opposite way as P-2, with larger $wpd$ for $hod$ in comparison to $moy$ and $wnwd$.
-For the $5$ cluster solution, this group is divided into Q3-Q5, which are distinct due to their different relative significance of $moy$ and $wnwd$.", fig.height=6}
+P-1 (customers 16, 18) is distinguished by high $\wpd$ on *wnwd* values.
+P-2 has lower $\wpd$ than *moy* and *wnwd* for *hod*.
+P-3 operates in the opposite way as P-2, with larger $\wpd$ for *hod* in comparison to *moy* and *wnwd*.
+For the $5$ cluster solution, this group is divided into Q3-Q5, which are distinct due to their different relative significance of *moy* and *wnwd*."
+
+```{r summary-plot-wpd, fig.cap = "(ref:summary-plot-wpd)", fig.height=6}
 ```
@@
 
-<!-- In Figure \ref{fig:opt-cluster-tsne-jsd}, the $sindex$ is plotted against the number of clusters for clustering based on wpd-based distances.
+<!-- In Figure \ref{fig:opt-cluster-tsne-jsd}, the *sindex* is plotted against the number of clusters for clustering based on wpd-based distances.
 The scales are very different from those used in JS-based clustering.
-The $sindex$ falls from around $20$ when $k=3$ to $19$ when $k=4$ and further decreases to $17.5$ with $k>7$.
+The *sindex* falls from around $20$ when $k=3$ to $19$ when $k=4$ and further decreases to $17.5$ with $k>7$.
 Thus, there are few choices available for the number of clusters.
@@
 The variables are sorted according to their separation across classes (rather than their overall variation between classes).
-This means that $moy$ is the most important variable in distinguishing the groups, followed by $hod$ and $wnwd$.
-The two customers (id: 16 and 18) that have significant differences between $wnwd$ categories (in Figure \ref{fig:hod-ind-group-png}) stand out from the rest of the customers.
-Group-4 has a higher $wpd$ for $hod$ than $moy$ or $wnwd$.
-Group-2 and Group-3 have the most distinct patterns across $moy$.
+This means that *moy* is the most important variable in distinguishing the groups, followed by *hod* and *wnwd*.
+The two customers (id: 16 and 18) that have significant differences between *wnwd* categories (in Figure \ref{fig:hod-ind-group-png}) stand out from the rest of the customers.
+Group-4 has a higher $\wpd$ for *hod* than *moy* or *wnwd*.
+Group-2 and Group-3 have the most distinct patterns across *moy*.
 Group-1 is a mixed group that has strong patterns on at least one of the three variables.
-When $k=3$ is used, Group-2 and Group-3 are merged because their patterns relative to $moy$, $hod$ and $wnwd$ are the same.
+When $k=3$ is used, Group-2 and Group-3 are merged because their patterns relative to *moy*, *hod* and *wnwd* are the same.
 The findings vary from JS-based clustering, yet it is a helpful grouping.-->
 
-Figure \ref{fig:summary-plot-wpd} shows the $wpd$ values of the $24$ customers over $hod$, $moy$, and $wnwd$ , colored by $3$ (a) and $5$ (b) clusters from WPD clustering using a parallel coordinate plot.
-The variables ($wpd$ for different granularities) are
+Figure \ref{fig:summary-plot-wpd} shows the $\wpd$ values of the $24$ customers over *hod*, *moy*, and *wnwd* , colored by $3$ (a) and $5$ (b) clusters from WPD clustering using a parallel coordinate plot.
+The variables ($\wpd$ for different granularities) are
 standardized prior to clustering using WPD.
 In the display, the variables are sorted according to their separation across groups.
-This means that $wnwd$ is the most important variable in distinguishing the groups, followed by $hod$ and $moy$ for both (a) and (b).
+This means that *wnwd* is the most important variable in distinguishing the groups, followed by *hod* and *moy* for both (a) and (b).
 Groups P-1 and P-2 correspond to Q-1 and Q-2 respectively.
 Cluster P-3 splits into Q-3, Q-4 and Q-5.
-Customers 16 and 18 are characterized by unusual high values of $wpd$ on $wnwd$ compared to the rest of the customers and hence form a group.
-This could again be verified from Figure \ref{fig:hod-ind-group-png}, where these were the only two customers with a difference in their $wnwd$ behavior.
+Customers 16 and 18 are characterized by unusual high values of $\wpd$ on *wnwd* compared to the rest of the customers and hence form a group.
+This could again be verified from Figure \ref{fig:hod-ind-group-png}, where these were the only two customers with a difference in their *wnwd* behavior.
 They are represented by P-1.
-P-2 has lower $wpd$ for $hod$ than $moy$ and $wnwd$.
-P-3 behaves opposite to P-2 with higher $wpd$ for $hod$ compared to $moy$ and $wnwd$.
-These are the customers who have some significant pattern across $hod$ and this can again be validated by looking at Figure \ref{fig:hod-ind-group-png}.
-For a $5$ cluster solution, this group gets split into Q-3, Q-4 and Q-5 characterizing different relative significance of $moy$ and $wnwd$.
-For example, Q-4 and Q-5 have almost no pattern across $moy$, but Q-3 has a $moy$ pattern and thus it is reasonable to split them.
+P-2 has lower $\wpd$ for *hod* than *moy* and *wnwd*.
+P-3 behaves opposite to P-2 with higher $\wpd$ for *hod* compared to *moy* and *wnwd*.
+These are the customers who have some significant pattern across *hod* and this can again be validated by looking at Figure \ref{fig:hod-ind-group-png}.
+For a $5$ cluster solution, this group gets split into Q-3, Q-4 and Q-5 characterizing different relative significance of *moy* and *wnwd*.
+For example, Q-4 and Q-5 have almost no pattern across *moy*, but Q-3 has a *moy* pattern and thus it is reasonable to split them.
 The patterns could be different, but they are significant.
-Q-4 and Q-5 are separated because of their different significance of $hod$.
-All of these can also be verified from Table \ref{tab:summary-table-wpd} which shows the cluster summaries with members and median values of $wpd$ for the three variables.
-
-<!-- $hod$ has very low median values of $wpd$ for the members in the cluster, implying at least half of these customers do not have an interesting $hod$ pattern. -->
+Q-4 and Q-5 are separated because of their different significance of *hod*.
+All of these can also be verified from Table \ref{tab:summary-table-wpd} which shows the cluster summaries with members and median values of $\wpd$ for the three variables.
+
+<!-- *hod* has very low median values of $\wpd$ for the members in the cluster, implying at least half of these customers do not have an interesting *hod* pattern. -->
 
@@
 A cluster summary like Figure \ref{fig:groups-4and5} is not possible because there may be different but equally interesting patterns (e.g.
-high evening $hod$ and high daytime $hod$) in the same cluster.
+high evening *hod* and high daytime *hod*) in the same cluster.
 Simply it provides information that across a collection of customers this specific cluster has interesting patterns in a granularity (e.g.
-$hod$).
+*hod*).
 One would need to post-process the cluster to separate specific patterns.
@@
 In our application, we have removed those customers and done the analysis but the metrics used should be able to incorporate those customers with such structured missingness.
-Finally, $wpd$ is computationally heavy even under parallel computation.
+Finally, $\wpd$ is computationally heavy even under parallel computation.
 Future work can make the computations more efficient so that they are easily scalable to a large number of customers.

### 2021-11-19 Rmd/05-conclusion.Rmd
Typos and consistent notation

@@
 The R package `hakear` (https://github.com/Sayani07/hakear) provides tools for selecting and sorting significant cyclic granularities.
-The function `wpd()` computes the weighted pairwise distances ($wpd$) for each cyclic granularity or pair of granularities, and the function `select_harmonies()` chooses those with significant patterns and ranks them from highest to lowest $wpd$.
-This package is reliant on parallel processing using multiple multi-core computers for faster computation of $wpd$.
+The function `wpd()` computes the weighted pairwise distances ($\wpd$) for each cyclic granularity or pair of granularities, and the function `select_harmonies()` chooses those with significant patterns and ranks them from highest to lowest $\wpd$.
+This package is reliant on parallel processing using multiple multi-core computers for faster computation of $\wpd$.
 The selected harmonies can be plotted using package `gravitas` for potentially interesting displays.
@@
 ### Computational resources
-Simulation studies were carried out to study the behavior of $wpd$, build the normalization method as well as compare and evaluate different normalization approaches in Chapter \@ref(ch-hakear).
+Simulation studies were carried out to study the behavior of $\wpd$, build the normalization method as well as compare and evaluate different normalization approaches in Chapter \@ref(ch-hakear).
 In Chapter \@ref(ch-gracsr), our methods were tested on several data designs with different parameters to evaluate their performance.
-$wpd$ is computationally heavy for cyclic granularities with smaller levels.
+$\wpd$ is computationally heavy for cyclic granularities with smaller levels.
 JS and WPD approaches are also computationally intensive when run on large number of customers.
@@
 
-We address few limitations of the current framework that might serve as natural next steps for this work and some potential short- and long-term aims in future ideas.
+We address several limitations of the current framework that might serve as natural next steps for this work and some potential short- and long-term aims in future ideas.
 
@@
 
-The time series are observed over a short period of time (1-3 years) in the motivating example of this research.
-Hence, they are assumed to be stationary.
+The time series are observed over a short period of time (1--3 years) in the motivating example of this research, and they are assumed to be stationary.
 But it is possible that the distributions change over time, even over a short period.
@@
 Another possible direction is to reduce computational time so that the proposed methods are easily scalable to many customers.
-The distance measure, $wpd$ is computationally heavy even under parallel computation.
+The distance measure, $\wpd$ is computationally heavy even under parallel computation.
 Moreover, while computing distances between time series, the proposed methods compute all possible pairwise distances, which acts as a computational barrier.

### 2021-11-19 Rmd/01-intro.Rmd
update bib file
Typos and consistent notation

--- 2021-11-16 Rmd/01-intro.Rmd
@@
 <!-- To tame the complexity of time, breaking it into years, months, weeks, days and so on in a hierarchical manner is a common way to relate data to time.
-Such dicrete human made abstractions of time can be thought of as time granularities.(Aigner et al., 2008).
+Such discrete human made abstractions of time can be thought of as time granularities.(Aigner et al., 2008).
 -->
@@
 These discrete abstractions of time are known as time granularities.
-Linear time granularities [@Bettini1998-ed] ,such as hours, days, weeks and months, respect the linear progression of time and are non-repeating.
+Linear time granularities [@Bettini1998-ed],such as hours, days, weeks and months, respect the linear progression of time and are non-repeating.
 Cyclic temporal granularities representing cyclical repetitions in time (such as hour-of-the-day, work-day/weekend) are effective for analyzing repetitive patterns in time series data.

### 2021-11-19 Rmd/01-intro.Rmd
check link works
Emi's pull changes to intro - can't merge because of conflict

--- 2021-11-18 Rmd/01-intro.Rmd
@@
 To acquire a comprehensive view of the repeated patterns, it is necessary to navigate through all of the conceivable cyclic granularities.
-This approach is consistent with the concept of exploratory data analysis (EDA) [@Tukey1977-jx], which stresses the utilization of multiple perspectives on data to assist with formulating hypotheses before proceeding to formal inferences or modelling.
+This approach is consistent with the concept of exploratory data analysis EDA [@Tukey1977-jx], which stresses the utilization of multiple perspectives on data to assist with formulating hypotheses before proceeding to formal inferences or modelling.
 This, however, is a challenging process since it throws up a myriad of possible hypotheses.
@@
 The distance measure is computed for a single or pairs of cyclic granularities, and it can be compared across different cyclic granularities as well as a collection of time series.
-The method is demonstrated with a case study using residential smart meter data from Melbourne where temporal granularities with significant distributional differences are automatically found.
+This chapter also includes a case study using residential smart meter data from Melbourne to demonstrate how the suggested methodology may be utilized to automatically find temporal granularities with significant distributional differences.
 The methods are implemented in the open-source R package `hakear`.

### 2021-11-19 Rmd/01-intro.Rmd
minor typos while proofreading

@@
 To acquire a comprehensive view of the repeated patterns, it is necessary to navigate through all of the conceivable cyclic granularities.
-This approach is consistent with the concept of exploratory data analysis EDA [@Tukey1977-jx], which stresses the utilization of multiple perspectives on data to assist with formulating hypotheses before proceeding to formal inferences or modelling.
+This approach is consistent with the concept of exploratory data analysis EDA [@Tukey1977-jx], which stresses the utilization of multiple perspectives on data to assist with formulating hypotheses before proceeding to formal inferences or modeling.
 This, however, is a challenging process since it throws up a myriad of possible hypotheses.

### 2021-11-19 Rmd/02-gravitas.Rmd
x axis labels overlap in paper-gravitas fixed
Supplementary "Material" to "material"
update bib file
minor typos while proofreading

@@
 However, the definitions and rules defined are inadequate for describing non-linear granularities.
-Hence, there is a need to define some new time granularities, that can be useful in visualizations.
+Hence, there is a need to define some new time granularities that can be useful in visualizations.
 Cyclic time granularities can be circular, quasi-circular or aperiodic.
@@
 In this pattern, each granule of $H$ is shifted by $P$ granules of $G$.
-$P$ is called the \textbf{period} (@Bettini2000-qk).
+$P$ is called the \textbf{period} [@Bettini2000-qk].
 
@@
 \caption{Examples of circular granularities with bottom granularity minutes.
-Circular granularity $C_i$ relates two linear granularities one of which groups periodically into the other with regular mapping and period $P_i$.
+Circular granularity $C_i$ relates two linear granularities, one of which groups periodically into the other with regular mapping and period $P_i$.
 Circular granularities can be expressed using modular arithmetic due to their regular mapping.}
@@
 
-```{r bothcust, fig.cap = "(ref:bothcust)", fig.pos="!p"}
+```{r bothcust, fig.cap = "(ref:bothcust)", fig.pos="!p",fig.height=5}
 ```
@@
 
-## Supplementary Materials {- #supplement-gravitas}
+## Supplementary materials {- #supplement-gravitas}
 

### 2021-11-19 Rmd/04-gracsr.Rmd
minor typos while proofreading

@@
 
-```{r flowchart, fig.cap = "Flow chart illustrating the pipeline for out method for clustering time series.", out.width="100%"}
+```{r flowchart, fig.cap = "Flow chart illustrating the pipeline for our method for clustering time series.", out.width="100%"}
 knitr::include_graphics("img/flowchart.png")
@@
 (This extends to harmonies, pairs of granularities that might interact with each other.) We define "significant" granularities as those with significant distributional differences across at least one category.
-The reason for subsetting granularities in this way, is that, clustering algorithms perform badly in the presence of nuisance variables.
+The reason for subsetting granularities in this way is that clustering algorithms perform badly in the presence of nuisance variables.
 Granularities that do not have some difference between categories are likely to be nuisance variables.
@@
 P-3 operates in the opposite way as P-2, with larger $\wpd$ for *hod* in comparison to *moy* and *wnwd*.
-For the $5$ cluster solution, this group is divided into Q3-Q5, which are distinct due to their different relative significance of *moy* and *wnwd*."
+For the $5$ cluster solution, this group is divided into Q-3, Q-4, Q-5, which are distinct due to their different relative significance of *moy* and *wnwd*."
 
@@
 In summary, none of the methods captured the five original prototypes exactly.
-JS-NQT was almost identical, though, but WPD produced quite a different grouping.
+JS-NQT was almost identical, but WPD produced quite a different grouping.
 This is quite a reasonable result and illustrates both the difficulties of clustering to obtain a particularly expected grouping and the ability to learn unexpected patterns in the data.

### 2021-11-19 Rmd/04-gracsr.Rmd
fig reference caption remove quote

@@
 
-(ref:summary-plot-wpd) "Summary plots for three (a) and five (b) clusters from WPD approach showing the $\wpd$ values of each customers across *hod*, *moy*, and *wnwd* through a parallel coordinate plot.
+(ref:summary-plot-wpd) Summary plots for three (a) and five (b) clusters from WPD approach showing the $\wpd$ values of each customers across *hod*, *moy*, and *wnwd* through a parallel coordinate plot.
 P-1 and P-2 groups correspond to Q-1 and Q-2, respectively.
@@
 P-3 operates in the opposite way as P-2, with larger $\wpd$ for *hod* in comparison to *moy* and *wnwd*.
-For the $5$ cluster solution, this group is divided into Q-3, Q-4, Q-5, which are distinct due to their different relative significance of *moy* and *wnwd*."
+For the $5$ cluster solution, this group is divided into Q-3, Q-4, Q-5, which are distinct due to their different relative significance of *moy* and *wnwd*.
 

### 2021-11-19 Rmd/04-gracsr.Rmd
Fig 4.8 caption

@@
 
-```{r opt-cluster-tsne-jsd, fig.cap = "Clustering summaries: (a) t-SNE computed on the $24$ selected customers, and (b) separation index (*sindex*) for $2$-$10$ clusters using JS-NQT, JS-RS and WPD.
+(ref:opt-cluster-tsne-jsd) Clustering summaries: (a) t-SNE computed on the $24$ selected customers, and (b) separation index (*sindex*) for $2$-$10$ clusters using JS-NQT, JS-RS and WPD.
 Various choices in number of clusters would be recommended.
 Four clusters are visible in t-SNE, although it might hide a fifth cluster because dimension reduction to 2D may be insufficient to see the difference.
-JS-NQT suggests $3$, JS-RS suggests $2$ or $5$ and WPD suggests $3$ or $5$."}
+JS-NQT suggests $3$, JS-RS suggests $2$ or $5$ and WPD suggests $3$ or $5$.
+
+```{r opt-cluster-tsne-jsd, fig.cap = "(ref:opt-cluster-tsne-jsd)"}
 #include_graphics("img/opt_clust_parenthesis.png")

### 2021-11-19 Rmd/04-gracsr.Rmd
fixed notational problem in caption

@@
 
-```{r hod-ind-group-png, fig.cap = "The distribution of electricity demand across individual customers over three granularities *hod*, *moy*, and *wnwd* are shown for the $24$ selected customers using quantile and box plots.
+(ref:hod-ind-group-png) The distribution of electricity demand across individual customers over three granularities *hod*, *moy*, and *wnwd* are shown for the $24$ selected customers using quantile and box plots.
 They are split into batches of $12$ in (a) and (b), with each row in (a) or (b) representing a customer.
-The number indicates a unique customer id and a prototype id."}
+The number indicates a unique customer id and a prototype id.
+
+```{r hod-ind-group-png, fig.cap = "(ref:hod-ind-group-png)"}
 ```
@@
 
-```{r groups-4and5, fig.cap = "Summary plots for four and five clusters from JS-NQT showing the distribution of electricity demand combined for all members over *hod*, *moy*, and *wnwd*.
+(ref:groups-4and5) Summary plots for four and five clusters from JS-NQT showing the distribution of electricity demand combined for all members over *hod*, *moy*, and *wnwd*.
 Groups A-1 (customers 21-24), A-2 (customers 4-9), and A-3 (customers 10-15) profiles correspond to Groups B-1, B-2, and B-3, respectively.
-Cluster A-4 splits into B-4 (customers 16-20) and B-5 (customers 1-3) to produce the five clusters, which better resolves the *moy* distribution.", fig.height=5}
+Cluster A-4 splits into B-4 (customers 16-20) and B-5 (customers 1-3) to produce the five clusters, which better resolves the *moy* distribution.
+
+```{r groups-4and5, fig.cap = "(ref:groups-4and5)", fig.height=5}
 

### 2021-11-23 Rmd/00-c-declaration.Rmd
minor typos while proofreading
declaration with signature

@@
 When co-authors are listed, it indicates that the work was produced through active collaboration between researchers and that their contributions to team-based research have been recognized.
-The following table details the work, including
-my and my fellow co-authors contributions:
+The following table details the work, including my and my fellow co-authors' contributions.
 
@@
 
-```{r signature, out.width = "20%", fig.align = "left", eval = FALSE}
+```{r signature, out.width = "20%", fig.align = "left"}
 knitr::include_graphics("img/signature.png")

### 2021-11-24 Rmd/00-c-declaration.Rmd
updated declaration file

@@
 
-# Declaration {-#ch-declare}
+# Thesis including published works declaration {-#ch-declare}
 
@@
 Concept, methodology development, writing first draft and software development & (1) Rob J Hyndman, Concept, methodology and software development, writing 12\% (2) Dianne Cook, Concept, methodology and software development, writing 12\% (3) Antony Unwin, Concept and software development 6\% & \multicolumn{1}{c|}{N} \\
-\hline
-3 & Detecting distributional differences between temporal granularities
-for exploratory time series analysis & Working Paper & 80\%.
-Concept, methodology development, writing first draft and software development & (1) Rob J Hyndman, Concept, methodology development and writing 10\% (2) Dianne Cook, Concept, methodology development and writing 10\%  & \multicolumn{1}{c|}{N} \\
-\hline
-4 & Clustering time series based on probability distributions across temporal granularities & To be submitted & 80\%.
-Concept, methodology development, writing first draft and software development & (1) Dianne Cook, Concept, methodology development and writing 15\% (2) Rob J Hyndman, Concept, methodology development and writing 5\% & \multicolumn{1}{c|}{N} \\
 \hline
@@
 `r if (knitr::is_html_output()) '-->'`
+
+\vspace{2em}
+\noindent
+The undersigned hereby certify that the above declaration correctly reflects the nature and extent of the student's co-authors' contribution to this work.
+In instances where I am not the responsible author I have consulted with the responsible author to agree on the respective contributions of the authors.
+
+\textbf{Main Supervisor name:} Rob J Hyndman
+
+\textbf{Main Supervisor signature:}
+
+\textbf{Date:} `r Sys.Date()`

### 2021-11-24 Rmd/00-e-preface.Rmd
check link works
updated declaration file

@@
 
-Chapter \@ref(ch-gravitas) has been accepted by the *Journal of Computational and Graphical Statistics*.
+Chapter \@ref(ch-gravitas) has been accepted by the _Journal of Computational and Graphical Statistics_.
 It has won the ACEMS Business Analytics Prize in 2020.
-The accompanying R package **gravitas** is on CRAN.
+The accompanying R package `gravitas` is on CRAN.
 Chapter \@ref(ch-hakear) and Chapter \@ref(ch-gracsr) are yet to be submitted.
@@
 
-This thesis is written in R Markdown [@rmarkdown] with **bookdown** [@bookdown], using **renv** [@renv] to create reproducible environments.
-The online version of this thesis is hosted at <https://sayani07.github.io/thesis-sg>.
+This thesis is written in R Markdown [@rmarkdown] with `bookdown` [@bookdown], using `renv` [@renv] to create reproducible environments.
+The online version of this thesis is hosted at <https://sayani07.github.io/thesis-SG>.
 All materials (including the data sets and source files) required to reproduce this document can be found at the Github repository <https://github.com/Sayani07/thesis-SG>.

### 2022-02-02 Rmd/00-c-declaration.Rmd
date changed to 19th November
updated changes working on comments

@@
----
-output:
-  pdf_document: default
-  html_document: default
----
-`r if (knitr::is_html_output()) '<!--'`
+# Declaration {-#ch-declare}
 
-\newpage
+This thesis is an original work of my research and contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.
 
-# Thesis including published works declaration {-#ch-declare}
-
-I hereby declare that this thesis contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.
-
-Three publications are included in this thesis, one of which has been accepted for publication in the _Journal of Computational and Graphical Statistics_ (@Gupta2021-gravitas).
-Two others have not yet been submitted.
-"Exploring probability distributions across cyclic temporal granularities" is the central theme of the thesis.
-All of the papers in the thesis were conceptualized, developed, and written by myself, the student, while working in the Department of Econometrics and Business Statistics under the supervision of Professor Rob J Hyndman and Professor Dianne Cook.
-When co-authors are listed, it indicates that the work was produced through active collaboration between researchers and that their contributions to team-based research have been recognized.
-The following table details the work, including my and my fellow co-authors' contributions.
-
-\begin{table}
-\centering\footnotesize\tabcolsep=0.12cm
-\begin{tabular}{|p{1cm}|p{2cm}|p{1.5cm}|p{3.5cm}|p{3.5cm}|p{1.5cm}|}
-\hline
-\RaggedRight\textbf{Thesis Chapter}  & \RaggedRight\textbf{Publication Title}  & \RaggedRight\textbf{Status (published, in press, accepted or returned for revision)}  & \RaggedRight\textbf{Nature and} {\%} \RaggedRight\textbf{of student contribution} & \RaggedRight\textbf{Co-author name(s) Nature and} {\%} \RaggedRight\textbf{of Co-author’s contribution} & \RaggedRight\textbf{Co-author(s), Monash student Y/N} \\ \hline
-2 & Visualizing probability distributions across bivariate cyclic temporal granularities & Published & 70\%.
-Concept, methodology development, writing first draft and software development & (1) Rob J Hyndman, Concept, methodology and software development, writing 12\% (2) Dianne Cook, Concept, methodology and software development, writing 12\% (3) Antony Unwin, Concept and software development 6\% & \multicolumn{1}{c|}{N} \\
-\hline
-\end{tabular}
-\end{table}
-
-I have renumbered sections of submitted or published papers in order to generate a consistent presentation within the thesis.
-
-\textbf{Student name:} Sayani Gupta
+\vspace{4em}
 
@@
 
-<!-- eval = file.exists("img/signature.png") -->
+\textbf{Student name:} Sayani Gupta
 
-\textbf{Date:} `r Sys.Date()`
-
-`r if (knitr::is_html_output()) '-->'`
-
-\vspace{2em}
-\noindent
-The undersigned hereby certify that the above declaration correctly reflects the nature and extent of the student's co-authors' contribution to this work.
-In instances where I am not the responsible author I have consulted with the responsible author to agree on the respective contributions of the authors.
-
-\textbf{Main Supervisor name:} Rob J Hyndman
-
-\textbf{Main Supervisor signature:}
-
-\textbf{Date:} `r Sys.Date()`
+\textbf{Date:} `r as.Date("2021-11-19")`

### 2022-02-02 Rmd/00-e-preface.Rmd
updated changes working on comments

@@
 
-Chapter \@ref(ch-gravitas) has been accepted by the _Journal of Computational and Graphical Statistics_.
+Chapter \@ref(ch-gravitas) has been published online at the [_Journal of Computational and Graphical Statistics_](https://www.tandfonline.com/eprint/NXKCSZHIBFXCJD2NKB68/full?target=10.1080/10618600.2021.1938588).
 It has won the ACEMS Business Analytics Prize in 2020.

### 2022-02-02 Rmd/02-gravitas.Rmd
updated changes working on comments

@@
 
-\noindent In the Mayan calendar (Table \@ref{tab:tab-mayan}), kin-of-tun or kin-of-baktun are examples of multiple-order-up granularities and single-order-up granularities are kin-of-uinal, uinal-of-tun etc.
+\noindent In the Mayan calendar (\autoref{tab:tab-mayan}), kin-of-tun or kin-of-baktun are examples of multiple-order-up granularities and single-order-up granularities are kin-of-uinal, uinal-of-tun etc.
 

### 2022-02-02 Rmd/04-gracsr.Rmd
updated changes working on comments

@@
 The time series features, length of time series, representation technique, and, of course, the purpose of clustering time series all influence the suitable similarity measure or distance metric to a meaningful level.
-The three primary methods to time series clustering [@liao2005clustering] are algorithms that operate directly with distances or raw data points in the time or frequency domain (distance-based), with features derived from raw data (feature-based), or indirectly with models constructed from raw data (model-based).
+The three primary methods of time series clustering [@liao2005clustering] are algorithms that operate directly with distances or raw data points in the time or frequency domain (distance-based), with features derived from raw data (feature-based), or indirectly with models constructed from raw data (model-based).
 The efficacy of distance-based techniques is highly dependent on the distance measure utilized.
 Defining an appropriate distance measure for the raw time series may be a difficult task since it must take into account noise, variable lengths of time series, asynchronous time series, different scales, and missing data.
-Commonly used distance-based similarity measures as suggested by a decade review of time series clustering approaches [@Aghabozorgi2015-ct] are Euclidean, Pearson's correlation coefficient and related distances, Dynamic Time Warping (DTW), Autocorrelation, Short time series distance, Piecewise regularization, cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances [@liao2007clustering] but on vector time series data.
+Commonly used distance-based similarity measures as suggested by a review of time series clustering approaches [@Aghabozorgi2015-ct] are Euclidean, Pearson's correlation coefficient and related distances, Dynamic Time Warping (DTW), Autocorrelation, Short time series distance, Piecewise regularization, cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances [@liao2007clustering] but on vector time series data.
 Among these alternatives, Euclidean distances have high performance but need the same length of data over the same period, resulting in information loss regardless of whether it is on raw data or a smaller collection of features.

### 2022-02-02 Rmd/03-hakear.Rmd
Supplementary "Material" to "material"
no section supplementary
comments hakear-Catherine

--- 2021-11-15 Rmd/03-hakear.Rmd
@@
 The distance measure is computed for a single or pairs of cyclic granularities that can be compared across different cyclic granularities and also on a collection of time series.
-The methods are implemented in the open-source R package `hakear`.
+The methods are implemented in the open-source R package [`hakear`](https://github.com/Sayani07/hakear).
 
@@
 Difference between the 90th and 75th quantiles is less for (Jan, Feb) for the first household (a), suggesting that it is a more frequent user of air conditioners than the second household (b).
-Energy consumption for (a) changes across both granularities, whereas for (b) daily pattern stays same irrespective of the months."}
+Distribution of energy usage for (a) changes across both granularities, whereas for (b) daily pattern stays same irrespective of the months."}
 ```
@@
 This problem is similar to Scagnostics (Scatterplot Diagnostics) by @tukey1988computer, which are used to identify meaningful patterns in large collections of scatterplots.
-Given a set of $v$ variables, there are $v(v-1)/2$ pairs of variables, and thus the same number of possible pairwise scatterplots.
+Given a set of $v$ variables, there are $p(p-1)/2$ pairs of variables, and thus the same number of possible pairwise scatterplots.
 Therefore, even for small $v$, the number of scatterplots can be large, and scatterplot matrices (SPLOMs) can easily run out of pixels when presenting high-dimensional data.
@@
 The article is organized as follows.
-Section \ref{sec:computation-wpd} introduces a distance measure for detecting distributional difference in temporal granularities, which enables identification of patterns in the time series data;
+Section \ref{sec:computation-wpd} introduces a distance measure for detecting distributional difference in temporal granularities for a continuous univariate dependent variable.
+This enables identification of patterns in the time series data;
 Section \ref{sec:rank-wpd} devises a selection criterion by choosing a threshold, which results in detection of only significantly interesting patterns.
@@
 
-## Supplementary Materials
+## Supplementary materials {- #supplement-hakear}
 

### 2022-02-02 report/reponse.Rmd
comments hakear-Catherine

---
title: Amendments summary
author: Sayani Gupta
output:
  MonashEBSTemplates::memo:
    includes:
      in_header: preamble2.tex
fontfamily: mathpazo
---

I thank my examiners for their thorough and constructive comments. The point by point description of changes are below: the examiners' comments are in red and my response is in black.

## Professor Catherine Hurley {-}

I appreciate all the comments and suggestions, and will incorporate them into future work. Specific comments are addressed as follows:

### Chapter 2

> Q1. In Figure 2.1 on the right hand side it should be t/24, t/(24 × 7)

> Q2. Last sentence of Section 2.3.4: fix the table reference

Done.

> Q3. In Figure 2.4, the outlier glyphs are in my opinion too big and dark and attract too much attention.
The log scale for the y-axis should be mentioned in the caption.

> Q4. The violin plot in Figure 2.5c is barely recognisable, and comparison of the distributions is challenging.
The plot deserves a bit more space.

### Chapter 3

> Q1. On page 33 there is a mention of the hakear package, without a reference.

It appears in the abstract of the paper. 
For the thesis, I have added it since the abstract acts as a summary to the chapter.

> Q2. In the Introduction state the methods are for continuous variables.

Added "for a continuous univariate dependent variable" in the last paragraph of Introduction.

> Q3. In Figure 3.1, the points are a bit hard to see, so it looks like whiskers extend beyond the data.

Changed to theme_bw() instead of default and outlier.alpha (0.5 ->0.1) and geom_jitter(alpha = 0.04 -> 0.06)

> Q4. On page 35, refer to months consistently using short or long names.

No change. Refer as months or month_year depending on the context.

> Q5. On page 35 v refers to the number of variables, in Section 3.3.3 it refers to the variable.

Changed. On p35, $v$ changed to $p$, where $p$ represents the number of variables. This is analogous to many time granularities and their pairwise combinations and $v$ represents the univariate measured variable for which distributions are constructed for different time granularities and their pairs.

> Q6. In the Figure 3.2 caption, I do not understand the comment beginning “Difference between the 90th. . . ”.
Also for the sentence “Energy consumption for (a)..”, it would clarify to insert the word “median”.

No change.
Changed to "distribution of energy usage"

> Q6. On the top of page 38, it should say “the distribution means are three standard deviations apart”.

> Q7. Maybe include a reference for gestalt theory.

> Q8. In Section 3.2.2, I find the references to the null distribution confusing. Surelt that is the design in
Figure 3.3(a) only?

> Q9. In table 3.1 Nc is the number of cyclic granularities, whereas in the text (first sentence of 3.3.2 and
3.3.3) it is m.

> Q10. Add v to table 3.1.

Q42. In the first sentence on the top of page 42, ordered and unordered are mixed up. Have you considered
the setting where the facet variable levels are ordered? For the within-facet ordered distances, you
could consider a distance measure that respects circular order, or choose the start level appropriate to
the display.

Q43. I like Figure 3.4. In (3), the dotted arcs only connect to a1 which might be misleading.

Q44. In the pairwise distance measure on page 43, is there any adjustment made for varying numbers of
observations across the levels of A and B? For example in Figure 3.3(d) if there are few values at
xlevel=1 the comparison shown is less interesting

Q45. The equations on page 44 could be tidied up. There are extra () on the definition of wpdperm and on
the residual definition at the bottom of the page. Use × not  for the equations.

Q46. Is there a practical reason why wpdglm is not working as expected for lower nx and nf ?

Q47. In the 3.3.1 algorithm, there is m, and then M.

Q48. In Table 3.2, I would suggest separating the m = 1 and m = 2 tables.

Q49. The presentation of the material on the simulation study on page 48 could be improved. Where is
the notation wpdl,s used? Figure 3.5 shows the m = 2 results only. The text alludes to a simulation
involving different underlying distributions, but this is not mentioned again.

Q410: In Figure 3.5 the axis tick labels should be smaller. The blue and orange marks are hard to see. Maybe
show fewer panels?

Q411: In Figure 3.6 the axis tick labels should be smaller. The blue and orange are hard to see. The caption
should refer to the rug. Maybe show fewer panels?

Q412: In Figure 3.7 (a) the heatmaps need id labels. The grey color is missing from the legend.Maybe use a
different colour in the heatmap for the significant comparisons. State the threshold for significance in
the caption. In Table 3.3 the caption should explain the threshold. Maybe use colour instead of stars
to indicate significance?

Q413: The link in the Acknowledgements https://github.com/Sayani07/paper-hakear is not available. Neither
are the supplementary materials. I understand this is for the paper version, I mention it for completeness.

Q51:In Section 4.1, line 5 “method of time series clustering”.

Q52: Page 59, fourth bullet point. I found these sentences confusing.

Q53: In the material in the bottom of page 59, make it clear from the outset you only have data on energy
use, not on property size, location, family size and so on.

Q54. End of Chapter 4.1: incorrect reference to Section 2.7.

Q55. I found Figure 4.1 very useful. In each box, would be helpful to put in “or”, in places where only one of
the steps listed is performed, eq Normal quantile transform or Robust scaling. As there are many steps
in the algorithm, it would be helpful to the reader to label the pipeline steps and to refer back to them
in the text. The text in the Data pre-processing box does not make sense to me.

Q56. The section describing RS and NQT is confusing. It states RS is applied to each time series separately.
Is NQT also applied to each observation separately? How does “it could be useful to standardize it for
the selected set of significant granularities prior to computing the distances” relate to the following
bullet points?

Q57. Page 64 “D is the Jensen-Shannon distance”.

Q58. In Table 4.1, the R column should have 250, not 20.

Q59. The description of the data generation on page 67 and Table 4.2 is confusing. Maybe state the
distribution of each vt.

Q510. Where is μ in Table 4.2?

Q511. In Section 4.3.3 what are the values of R and T?

Q512. Figure 4.5 caption: MDS summary plots of what?

Q513: Figure 4.7 is missing labels (a), (b) (c).

Q514: Why do you chose to summarize 4 and 5 clusters from JS-NQT when sindex suggests 3 clusters?

Q515: The reference to the stationarity assumption in the Discussion needs clarification.

Q516: Page 81 “can not” should be “cannot”

Q517: Again, the computational burden could do with more Discussion. At present, if I were to use this
method on my data, what kind of sizes are realistic to work with?

### 2022-02-02 scripts/hakear.R

@@ -154,10 +154,11 @@ id2_tsibble_hd <- elec %>%
   as_tsibble(index = date_time) %>% 
   create_gran("hour_day") %>% 
   ggplot(aes(x = hour_day, y = kwh)) +
-  geom_boxplot(width = 0.6, outlier.colour = "black", outlier.alpha = 0.5, fill = "#CC79A7", colour =  "#0072B2") + scale_y_log10() +
-  geom_jitter(alpha = 0.04, colour = "#E69F00") +
+  geom_boxplot(width = 0.6, outlier.colour = "black", outlier.alpha = 0.1, fill = "#CC79A7", colour =  "#0072B2") + scale_y_log10() +
+  geom_jitter(alpha = 0.06, colour = "#E69F00") +
   ggtitle("(a)")+ 
-  ylab ("energy usage [kWh]")
+  ylab ("energy usage [kWh]") +
+  theme_bw()
 
 
 
@@ -166,10 +167,11 @@ id2_tsibble_dw <- elec %>%
   as_tsibble(index = date_time) %>% 
   create_gran("month_year") %>% 
   ggplot(aes(x = month_year, y = kwh)) +
-  geom_boxplot(width = 0.6, outlier.colour = "black", outlier.alpha = 0.5, fill = "#CC79A7", colour =  "#0072B2") + scale_y_log10()+
-  geom_jitter(alpha = 0.04, colour = "#E69F00") +
+  geom_boxplot(width = 0.6, outlier.colour = "black", outlier.alpha = 0.1, fill = "#CC79A7", colour =  "#0072B2") + scale_y_log10()+
+  geom_jitter(alpha = 0.05, colour = "#E69F00") +
   ggtitle("(b)")+ 
-  ylab ("energy usage [kWh]")
+  ylab ("energy usage [kWh]") +
+  theme_bw()
 
 
 

### 2022-02-02 Rmd/03-hakear.Rmd
till comment 13, chapter 3 - Catherine

@@
 It is not a particularly interesting display, as the distributions do not vary across x-axis or facet categories.
-Figure \ref{fig:null4by2}b has x categories drawn from the same distribution, but across facets the distributions are three standard deviations apart.
+Figure \ref{fig:null4by2}b has x categories drawn from the same distribution, but across facets the distribution means are three standard deviations apart.
 Figure \ref{fig:null4by2}c exhibits the opposite situation where distribution between the x-axis categories are three standard deviations apart, but they do not change across facets.
@@
 Panels 1) and 2) show the between-facet distances.
-Panels 3) and 4) are used to illustrate within-facet distances when categories are un-ordered or ordered respectively.
+Within-facet distances are illustrated in Panels 3) (when categories are un-ordered, shown only with respect to a1) and Panel 4) (when categories are ordered).
 When categories are ordered, distances should only be considered for consecutive x-axis categories.
@@
 Pairwise distances are between-facets when they are considered between pairs of the form $(a_{j}b_{k}, a_{j}b_{k'})$.
-There are a total of ${\nf \choose 2}\nx$ between-facet distances, and ${\nx \choose 2}\nf$ within-facet distances if they are ordered and $\nf(\nx-1)$ within-facet distances if they are unordered.
+There are a total of ${\nf \choose 2}\nx$ between-facet distances, and ${\nx \choose 2}\nf$ within-facet distances if they are unordered and $\nf(\nx-1)$ within-facet distances if they are ordered.
 

### 2022-02-02 report/reponse.Rmd
till comment 13, chapter 3 - Catherine

@@
 
+Done.
+
 > Q7. Maybe include a reference for gestalt theory.
 
-> Q8. In Section 3.2.2, I find the references to the null distribution confusing. Surelt that is the design in
+To do.
+
+> Q8. In Section 3.2.2, I find the references to the null distribution confusing. Surely that is the design in
 Figure 3.3(a) only?
+
+No change.
+Null distributions refer to the case when there is no significant distributional differences between categories of one or more granularities.
 
@@
 
+Add $m$ to table 3.1
+$m$ = number of cyclic granularities to display together
+
+$N_c$ refers to the total number of contextual cyclic granularities.
+$m$ refers to the number of cyclic granularities we are considering together in the display.
+For example, contextual cyclic granularities could be $hour_day$, $day_week$ and $month_year$ and we want to visualize any one granularity at a time.
+So $N_c = 3$ and $m=1$.
+
 > Q10. Add v to table 3.1.
 
-Q42. In the first sentence on the top of page 42, ordered and unordered are mixed up. Have you considered
+Done.
+Added v:continuous univariate measured variable
+
+> Q11. In the first sentence on the top of page 42, ordered and unordered are mixed up. Have you considered
 the setting where the facet variable levels are ordered? For the within-facet ordered distances, you
@@
 
-Q43. I like Figure 3.4. In (3), the dotted arcs only connect to a1 which might be misleading.
+first part: Done.
+second part: To do
 
-Q44. In the pairwise distance measure on page 43, is there any adjustment made for varying numbers of
+> Q12. I like Figure 3.4. In (3), the dotted arcs only connect to a1 which might be misleading.
+
+Changed.
+Caption changed to incorporate that.
+Within-facet distances are illustrated in Panels 3) (when categories are un-ordered, shown only with respect to a1) and Panel 4) (when categories are ordered)
+
+> Q13. In the pairwise distance measure on page 43, is there any adjustment made for varying numbers of
 observations across the levels of A and B? For example in Figure 3.3(d) if there are few values at
@@
 
-Q45. The equations on page 44 could be tidied up. There are extra () on the definition of wpdperm and on
+No change.
+We assume we have enough observations for each level or combination of levels to compute distributions. As long as that holds, no adjustment made for varying number of observations.
+
+> Q14. The equations on page 44 could be tidied up. There are extra () on the definition of wpdperm and on
 the residual definition at the bottom of the page. Use × not  for the equations.

### 2022-02-02 scripts/hakear.R

@@ -408,10 +408,10 @@ p_varx <- sim_panel_varx %>%
 ## ---- notations
 ## 
 d1 <- tibble (
-  variable = c("$N_C$", "$H_{N_C}$", "$\\nx$", "$\\nf$" ,
+  variable = c("$N_C$", "$H_{N_C}$", "m", "$\\nx$", "$\\nf$" ,
                "$\\lambda$", "$\\omega$", "$\\wpd$", "$\\wpdsub{norm}$",
                "$\\nsub{perm}$", "$\\nsub{sim}$"),
-  description = c("number of cyclic granularities", "set of harmonies",
+  description = c("number of cyclic granularities", "set of harmonies", "number of cyclic granularities to display together",
                   "number of x-axis categories", "number of facet categories",
                   "tuning parameter", "increment (mean or sd)",
                   "raw weighted pairwise distance", "normalized weighted pairwise distance",
@@ -423,13 +423,15 @@ d2 <- tibble(variable =  c(
   "$\\Dnull$",
   "$\\Df$",
   "$\\Dx$",
-  "$\\Dfx$"),
+  "$\\Dfx$",
+  "v"),
   description = c(
     "threshold for significance",
     "null design with no distributional difference across categories",
     "design with distributional difference only across facets categories", 
     "design with distributional difference only across x-axis categories", 
-    "design with distributional difference across both facet and x-axis"))
+    "design with distributional difference across both facet and x-axis",
+    "continuous univariate measured variable"))
 
 
 d <- bind_rows(d1, d2)

### 2022-02-02 Rmd/03-hakear.Rmd
chapter 3 catherine comments done

@@
 \begin{align*}
-  \wpdsub{perm} = \frac{(\wpdsub{raw} - \text{mean}(\wpdsub{perm-data}))}{\text{sd}(\wpdsub{perm-data})}
+  \wpdsub{perm} = \frac{\wpdsub{raw} - \text{mean}(\wpdsub{perm-data})}{\text{sd}(\wpdsub{perm-data})}
 \end{align*}
@@
 Furthermore, the mean response is bounded, since any JSD is bounded by $1$ if a base $2$ logarithm is used [@Lin1991-pt].
-Hence, by @Faraway2016-uk, an inverse link is used for the model, which is of the form $y = a+b*\log(z) + e$, where $y = \wpdsub{raw}$, $z = (\nx*\nf)$ is the number of groups and $e$ are idiosyncratic errors.
-Let $\text{E}(y) = \mu$ and $a + b*\log(z) = g(\mu)$ where $g(\mu)= 1/\mu$ and $\hat \mu = 1/(\hat a + \hat b \log(z))$.
+Hence, by @Faraway2016-uk, an inverse link is used for the model, which is of the form $y = a+b\times\log(z) + e$, where $y = \wpdsub{raw}$, $z = (\nx \times \nf)$ is the number of groups and $e$ are idiosyncratic errors.
+Let $\text{E}(y) = \mu$ and $a + b\times\log(z) = g(\mu)$ where $g(\mu)= 1/\mu$ and $\hat \mu = 1/(\hat a + \hat b \log(z))$.
 The residuals from this model $(y-\hat y) = (y-1/(\hat a + \hat b \log(z)))$ would be expected to have no dependency on $z$.
 Thus, $\wpdsub{glm}$ is defined as the residuals from this model given by
-$$\wpdsub{glm} = \wpdsub{raw} - 1/(\hat a + \hat b*\log(\nx*\nf))$$
+$$\wpdsub{glm} = \wpdsub{raw} - 1/(\hat a + \hat b\times\log(\nx\times\nf))$$
 The distribution of $\wpdsub{glm}$ under $\Dnull$ will have approximately zero mean and a constant variance (not necessarily 1).
@@
 This approach, however, requires that the adjusted variables from the two approaches are brought to the same scale.
-We define $\wpdsub{glm-scaled} = \wpdsub{glm}*\sigma^2_{\text{perm}}/\sigma^2_{\text{glm}}$ as the transformed $\wpdsub{glm}$ with a similar scale as $\wpdsub{perm}$.
+We define $\wpdsub{glm-scaled} = \wpdsub{glm}\times \sigma^2_{\text{perm}}/\sigma^2_{\text{glm}}$ as the transformed $\wpdsub{glm}$ with a similar scale as $\wpdsub{perm}$.
 The adjusted measure from the combination approach, denoted by $\wpd$ is then defined as follows:
@@
     a. Given the measured variable; $\{v_t: t=0, 1, 2, \dots, T-1\}$, $\wpd$ is computed and is represented by $\wpd^{A, B}_{obs}$.
-    b. For $i=1,\dots,m$, randomly permute the original time series: $\{v_t^{i}: t=0, 1, 2, \dots, T-1\}$ and compute $\wpd^{A, B}_{i}$ from $\{v_t^{i}\}$.
+    b. For $i=1,\dots,M$, randomly permute the original time series: $\{v_t^{i}: t=0, 1, 2, \dots, T-1\}$ and compute $\wpd^{A, B}_{i}$ from $\{v_t^{i}\}$.
     c. Define $\wpdsub{sample} = \{\wpd^{A, B}_{1}, \dots, \wpd^{A, B}_{M}\}$.

### 2022-02-02 report/reponse.Rmd
chapter 3 catherine comments done

@@
 
-> Q6. On the top of page 38, it should say “the distribution means are three standard deviations apart”.
+> Q7. On the top of page 38, it should say “the distribution means are three standard deviations apart”.
 
@@
 
-> Q7. Maybe include a reference for gestalt theory.
+> Q8. Maybe include a reference for gestalt theory.
 
@@
 
-> Q8. In Section 3.2.2, I find the references to the null distribution confusing. Surely that is the design in
+> Q9. In Section 3.2.2, I find the references to the null distribution confusing. Surely that is the design in
 Figure 3.3(a) only?
@@
 
-> Q9. In table 3.1 Nc is the number of cyclic granularities, whereas in the text (first sentence of 3.3.2 and
+> Q10. In table 3.1 Nc is the number of cyclic granularities, whereas in the text (first sentence of 3.3.2 and
 3.3.3) it is m.
@@
 
-> Q10. Add v to table 3.1.
+> Q11. Add v to table 3.1.
 
@@
 
-> Q11. In the first sentence on the top of page 42, ordered and unordered are mixed up. Have you considered
+> Q12. In the first sentence on the top of page 42, ordered and unordered are mixed up. Have you considered
 the setting where the facet variable levels are ordered? For the within-facet ordered distances, you
@@
 
-> Q12. I like Figure 3.4. In (3), the dotted arcs only connect to a1 which might be misleading.
+> Q13. I like Figure 3.4. In (3), the dotted arcs only connect to a1 which might be misleading.
 
@@
 
-> Q13. In the pairwise distance measure on page 43, is there any adjustment made for varying numbers of
+> Q14. In the pairwise distance measure on page 43, is there any adjustment made for varying numbers of
 observations across the levels of A and B? For example in Figure 3.3(d) if there are few values at
@@
 
-> Q14. The equations on page 44 could be tidied up. There are extra () on the definition of wpdperm and on
+> Q15. The equations on page 44 could be tidied up. There are extra () on the definition of wpdperm and on
 the residual definition at the bottom of the page. Use × not  for the equations.
 
-Q46. Is there a practical reason why wpdglm is not working as expected for lower nx and nf ?
+Done.
 
-Q47. In the 3.3.1 algorithm, there is m, and then M.
+> Q16. Is there a practical reason why wpdglm is not working as expected for lower nx and nf ?
 
-Q48. In Table 3.2, I would suggest separating the m = 1 and m = 2 tables.
+To do.
 
-Q49. The presentation of the material on the simulation study on page 48 could be improved. Where is
+> Q17. In the 3.3.1 algorithm, there is m, and then M.
+
+Changed.
+1.b) for $i \in {1, 2, \dots, M}$.
+
+> Q18. In Table 3.2, I would suggest separating the m = 1 and m = 2 tables.
+
+Not changed.
+
+> Q19. The presentation of the material on the simulation study on page 48 could be improved. Where is
 the notation wpdl,s used? Figure 3.5 shows the m = 2 results only. The text alludes to a simulation
@@
 
-Q410: In Figure 3.5 the axis tick labels should be smaller. The blue and orange marks are hard to see. Maybe
+first part: How?
+second part: After NQT, results are pretty similar to an underlying normal distribution.
+
+> Q20. In Figure 3.5 the axis tick labels should be smaller. The blue and orange marks are hard to see. Maybe
 show fewer panels?
 
-Q411: In Figure 3.6 the axis tick labels should be smaller. The blue and orange are hard to see. The caption
+To do.
+
+> Q21. In Figure 3.6 the axis tick labels should be smaller. The blue and orange are hard to see. The caption
 should refer to the rug. Maybe show fewer panels?
 
-Q412: In Figure 3.7 (a) the heatmaps need id labels. The grey color is missing from the legend.Maybe use a
+To do.
+
+> Q22. In Figure 3.7 (a) the heatmaps need id labels. The grey color is missing from the legend. Maybe use a
 different colour in the heatmap for the significant comparisons. State the threshold for significance in
@@
 
-Q413: The link in the Acknowledgements https://github.com/Sayani07/paper-hakear is not available. Neither
+To do.
+
+> Q23. The link in the Acknowledgements https://github.com/Sayani07/paper-hakear is not available. Neither
 are the supplementary materials. I understand this is for the paper version, I mention it for completeness.
 
-Q51:In Section 4.1, line 5 “method of time series clustering”.
+Changed.
+The link works. The repository has now been made public.
 
-Q52: Page 59, fourth bullet point. I found these sentences confusing.
+### Chapter 4
 
-Q53: In the material in the bottom of page 59, make it clear from the outset you only have data on energy
+> Q1. In Section 4.1, line 5 “method of time series clustering”.
+
+Done.
+
+> Q2. Page 59, fourth bullet point. I found these sentences confusing.
+
+To do.
+
+> Q3. In the material in the bottom of page 59, make it clear from the outset you only have data on energy
 use, not on property size, location, family size and so on.
 
-Q54. End of Chapter 4.1: incorrect reference to Section 2.7.
+To do.
 
-Q55. I found Figure 4.1 very useful. In each box, would be helpful to put in “or”, in places where only one of
+> Q4. End of Chapter 4.1: incorrect reference to Section 2.7.
+
+> Q5. I found Figure 4.1 very useful. In each box, would be helpful to put in “or”, in places where only one of
 the steps listed is performed, eq Normal quantile transform or Robust scaling. As there are many steps
@@
 
-Q56. The section describing RS and NQT is confusing. It states RS is applied to each time series separately.
+> Q6. The section describing RS and NQT is confusing. It states RS is applied to each time series separately.
 Is NQT also applied to each observation separately? How does “it could be useful to standardize it for
@@
 
-Q57. Page 64 “D is the Jensen-Shannon distance”.
+> Q7. Page 64 “D is the Jensen-Shannon distance”.
 
-Q58. In Table 4.1, the R column should have 250, not 20.
+> Q8. In Table 4.1, the R column should have 250, not 20.
 
-Q59. The description of the data generation on page 67 and Table 4.2 is confusing. Maybe state the
+> Q9. The description of the data generation on page 67 and Table 4.2 is confusing. Maybe state the
 distribution of each vt.
 
-Q510. Where is μ in Table 4.2?
+> Q10. Where is μ in Table 4.2?
 
-Q511. In Section 4.3.3 what are the values of R and T?
+> Q11. In Section 4.3.3 what are the values of R and T?
 
-Q512. Figure 4.5 caption: MDS summary plots of what?
+> Q12. Figure 4.5 caption: MDS summary plots of what?
 
-Q513: Figure 4.7 is missing labels (a), (b) (c).
+> Q13. Figure 4.7 is missing labels (a), (b) (c).
 
-Q514: Why do you chose to summarize 4 and 5 clusters from JS-NQT when sindex suggests 3 clusters?
+> Q14. Why do you chose to summarize 4 and 5 clusters from JS-NQT when sindex suggests 3 clusters?
 
-Q515: The reference to the stationarity assumption in the Discussion needs clarification.
+> Q15. The reference to the stationarity assumption in the Discussion needs clarification.
 
-Q516: Page 81 “can not” should be “cannot”
+> Q16. Page 81 “can not” should be “cannot”
 
-Q517: Again, the computational burden could do with more Discussion. At present, if I were to use this
+> Q17. Again, the computational burden could do with more Discussion. At present, if I were to use this
 method on my data, what kind of sizes are realistic to work with?

### 2022-02-03 Rmd/04-gracsr.Rmd
chapter 3 catherine comments done
comments Chapter 4 Symanzik

@@
 Only one study [@ozawa2016determining] identified time series characteristics by first conducting a Fourier transformation, to convert data from time to frequency domain, followed by $k$-means to cluster by greatest frequency.
-@Motlagh2019-yj suggests that the time feature extraction is limited by the type of noisy, patchy, and unequal time series common in residential customers and addresses model-based clustering by transforming the series into other objects such as structure or set of parameters which can be more easily characterized and clustered.
-@chicco2010renyi addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
-@Melnykov2013-sp discusses how outliers, noisy observations and scattered observations can complicate estimating mixture model parameters and hence the partitions.
+@Motlagh2019-yj suggested that the time feature extraction is limited by the type of noisy, patchy, and unequal time series common in residential customers and addresses model-based clustering by transforming the series into other objects such as structure or set of parameters which can be more easily characterized and clustered.
+@chicco2010renyi addressed information theory-based clustering such as Shannon or Renyi entropy and its variations.
+@Melnykov2013-sp discussed how outliers, noisy observations and scattered observations can complicate estimating mixture model parameters and hence the partitions.
 None of these methods focuses on exploring heterogeneity in repetitive patterns based on the dynamics of multiple temporal dependencies using probability distributions, which forms the basis of the methodology reported here.
@@
 Section \ref{sec:application-gracsr} discusses the application of the method to a subset of the real data.
-Finally, we summarize our results and discuss possible future directions in Section \ref{sec:discussion}.
+Finally, we summarize our results and discuss possible future directions in Section \ref{sec:discussion-gracsr}.
 
@@
 <!--These clusters may be commonly associated with real-world data segmentation.
-However, since the data is unlabeled a priori, more information is required to corroborate this.--> The flow of the procedures is illustrated in Figure \ref{fig:flowchart}.
+However, since the data is unlabeled a priori, more information is required to corroborate this.--> The flow of the procedures is illustrated in Figure \ref{fig:flowchart} and is further described in the following subsections.
 
@@
 designs = c(5, 4, 4)
-niter = list (c(25, 20, 500),
-              c(25, 20, 500),
-              c(25, 20, 500))
+niter = list (c(25, 250, 500),
+              c(25, 250, 500),
+              c(25, 250, 500))
 mu = list (c(1, 2, 5),
@@
 We consider the case where the distribution of $v$ varies across $g2$ levels for all designs, across $g3$ levels for a few designs, and $g1$ does not vary across designs.
-The proposed design is shown in Figure \ref{fig:gran2and1-clubbed}(b).
+The proposed design is shown in Figure \ref{fig:gran2and1-clubbed}(right).
 
@@
 For the smart meter example, the distribution for a customer moving to a different house or changing electrical equipment can change drastically.
-Our current approach can not detect these dynamic changes.
+Our current approach cannot detect these dynamic changes.
 Thirdly, it is possible that for a few customers, data for some categories from the list of considered significant granularities are missing.
@@
 
-**Data and scripts:** Data sets and R codes to reproduce all figures in this article  
-(validation.R, application.R)
+**Data and scripts:** Data sets and R codes to reproduce all figures in this article ([validation.R](https://github.com/Sayani07/paper-gracsR/blob/master/script/validation.R), [application.R](https://github.com/Sayani07/paper-gracsR/blob/master/script/application.R))
 
@@
 **Supplementary paper:** Additional graphics and R codes to reproduce it  
-(paper-supplementary.pdf, paper-supplementary.Rmd, supplementary.R)
+([paper-supplementary.pdf](https://github.com/Sayani07/paper-gracsR/blob/master/paper-supplementary.pdf), [paper-supplementary.Rmd](https://github.com/Sayani07/paper-gracsR/blob/master/paper-supplementary.Rmd), [supplementary.R](https://github.com/Sayani07/paper-gracsR/blob/master/script/supplementary.R))
 

### 2022-02-03 scripts/gracsr/application.R

@@ -211,7 +211,8 @@ hod_ind_group1 <- data_hod %>%
   ggrepel::geom_label_repel(aes(label = label, y = `50%`),
     nudge_x = -1,
     na.rm = TRUE, max.overlaps = 12
-  )
+  ) + plot_annotation("a")
+
 
 
 
@@ -253,7 +254,7 @@ hod_ind_group2 <- data_hod %>%
   ggrepel::geom_label_repel(aes(label = label, y = `50%`),
     nudge_x = -1,
     na.rm = TRUE
-  )
+  )+ plot_annotation("b")
 
 
 moy_ind_group1 <- data_moy %>%
@@ -830,7 +831,7 @@ data_table <- bind_rows(data_table3,
 data_table %>% 
 knitr::kable(format = "latex",
              escape = FALSE,
-             caption = "Summary table from WPD clusters showing median $wpd$ values ($moy$, $hod$, $wnwd$), cluster size ($nobs$) and the list of the customer-prototype id for each cluster with $3$ and $5$ number of clusters ($k$).")%>% 
+             caption = "Summary table from WPD clusters showing median $wpd$ values ($moy$, $hod$, $wnwd$), cluster size ($nobs$) and the list of the customer-prototype id for each cluster with $3$ and $5$ number of clusters ($k$). It is to be noted that $P-1$, $Q-1$ and $P-2$, $Q-2$ are identical and $P-3$ gets split into $Q-3$, $Q-4$ and $Q-5$.")%>% 
   kableExtra::collapse_rows(columns = 1)
 
 ## ----summary-plot-wpd------------------------------------------------------------

### 2022-02-03 report/response.Rmd
added response document

---
title: Amendments summary
author: Sayani Gupta
output:
  MonashEBSTemplates::memo:
    includes:
      in_header: preamble2.tex
fontfamily: mathpazo
---

I thank my examiners for their thorough and constructive comments. The point by point description of changes are below: the examiners' comments are in red and my response is in black.

# Professor Catherine Hurley

I appreciate all the comments and suggestions, and will incorporate them into future work. Specific comments are addressed as follows:

## Chapter 2

> Q1. In Figure 2.1 on the right hand side it should be t/24, t/(24 × 7)

> Q2. Last sentence of Section 2.3.4: fix the table reference

Done.

> Q3. In Figure 2.4, the outlier glyphs are in my opinion too big and dark and attract too much attention.
The log scale for the y-axis should be mentioned in the caption.

> Q4. The violin plot in Figure 2.5c is barely recognisable, and comparison of the distributions is challenging.
The plot deserves a bit more space.

## Chapter 3

> Q1. On page 33 there is a mention of the hakear package, without a reference.

It appears in the abstract of the paper. 
For the thesis, I have added it since the abstract acts as a summary to the chapter.

> Q2. In the Introduction state the methods are for continuous variables.

Added "for a continuous univariate dependent variable" in the last paragraph of Introduction.

> Q3. In Figure 3.1, the points are a bit hard to see, so it looks like whiskers extend beyond the data.

Changed to theme_bw() instead of default and outlier.alpha (0.5 ->0.1) and geom_jitter(alpha = 0.04 -> 0.06)

> Q4. On page 35, refer to months consistently using short or long names.

No change. Refer as months or month_year depending on the context.

> Q5. On page 35 v refers to the number of variables, in Section 3.3.3 it refers to the variable.

Changed. On p35, $v$ changed to $p$, where $p$ represents the number of variables. This is analogous to many time granularities and their pairwise combinations and $v$ represents the univariate measured variable for which distributions are constructed for different time granularities and their pairs.

> Q6. In the Figure 3.2 caption, I do not understand the comment beginning “Difference between the 90th. . . ”.
Also for the sentence “Energy consumption for (a)..”, it would clarify to insert the word “median”.

No change.
Changed to "distribution of energy usage"

> Q7. On the top of page 38, it should say “the distribution means are three standard deviations apart”.

Done.

> Q8. Maybe include a reference for gestalt theory.

To do.

> Q9. In Section 3.2.2, I find the references to the null distribution confusing. Surely that is the design in
Figure 3.3(a) only?

No change.
Null distributions refer to the case when there is no significant distributional differences between categories of one or more granularities.

> Q10. In table 3.1 Nc is the number of cyclic granularities, whereas in the text (first sentence of 3.3.2 and
3.3.3) it is m.

Add $m$ to table 3.1
$m$ = number of cyclic granularities to display together

$N_c$ refers to the total number of contextual cyclic granularities. $m$ refers to the number of cyclic granularities we are considering together in the display. For example, contextual cyclic granularities could be $hour_day$, $day_week$ and $month_year$ and we want to visualize any one granularity at a time. So $N_c = 3$ and $m=1$.

> Q11. Add v to table 3.1.

Done.
Added v:continuous univariate measured variable

> Q12. In the first sentence on the top of page 42, ordered and unordered are mixed up. Have you considered
the setting where the facet variable levels are ordered? For the within-facet ordered distances, you
could consider a distance measure that respects circular order, or choose the start level appropriate to
the display.

first part: Done.
second part: To do

> Q13. I like Figure 3.4. In (3), the dotted arcs only connect to a1 which might be misleading.

Changed.
Caption changed to incorporate that.
Within-facet distances are illustrated in Panels 3) (when categories are un-ordered, shown only with respect to a1) and Panel 4) (when categories are ordered)

> Q14. In the pairwise distance measure on page 43, is there any adjustment made for varying numbers of
observations across the levels of A and B? For example in Figure 3.3(d) if there are few values at
xlevel=1 the comparison shown is less interesting

No change.
We assume we have enough observations for each level or combination of levels to compute distributions. As long as that holds, no adjustment made for varying number of observations.

> Q15. The equations on page 44 could be tidied up. There are extra () on the definition of wpdperm and on
the residual definition at the bottom of the page. Use $\times$ for the equations.

Done.

> Q16. Is there a practical reason why wpdglm is not working as expected for lower nx and nf ?

To do.

> Q17. In the 3.3.1 algorithm, there is m, and then M.

Changed.
1.b) for $i \in {1, 2, \dots, M}$.

> Q18. In Table 3.2, I would suggest separating the m = 1 and m = 2 tables.

Not changed.

> Q19. The presentation of the material on the simulation study on page 48 could be improved. Where is
the notation wpdl,s used? Figure 3.5 shows the m = 2 results only. The text alludes to a simulation
involving different underlying distributions, but this is not mentioned again.

first part: How?
second part: After NQT, results are pretty similar to an underlying normal distribution.

> Q20. In Figure 3.5 the axis tick labels should be smaller. The blue and orange marks are hard to see. Maybe
show fewer panels?

To do.

> Q21. In Figure 3.6 the axis tick labels should be smaller. The blue and orange are hard to see. The caption
should refer to the rug. Maybe show fewer panels?

To do.

> Q22. In Figure 3.7 (a) the heatmaps need id labels. The grey color is missing from the legend. Maybe use a
different colour in the heatmap for the significant comparisons. State the threshold for significance in
the caption. In Table 3.3 the caption should explain the threshold. Maybe use colour instead of stars
to indicate significance?

To do.

> Q23. The link in the Acknowledgements https://github.com/Sayani07/paper-hakear is not available. Neither
are the supplementary materials. I understand this is for the paper version, I mention it for completeness.

Changed.
The link works. The repository has now been made public.

### Chapter 4

> Q1. In Section 4.1, line 5 “method of time series clustering”.

Done.

> Q2. Page 59, fourth bullet point. I found these sentences confusing.

To do.

> Q3. In the material in the bottom of page 59, make it clear from the outset you only have data on energy
use, not on property size, location, family size and so on.

To do.

> Q4. End of Chapter 4.1: incorrect reference to Section 2.7.

Done.

> Q5. I found Figure 4.1 very useful. In each box, would be helpful to put in “or”, in places where only one of
the steps listed is performed, eq Normal quantile transform or Robust scaling. As there are many steps
in the algorithm, it would be helpful to the reader to label the pipeline steps and to refer back to them
in the text. The text in the Data pre-processing box does not make sense to me.

To do? (Or not)

> Q6. The section describing RS and NQT is confusing. It states RS is applied to each time series separately.
Is NQT also applied to each observation separately? How does “it could be useful to standardize it for
the selected set of significant granularities prior to computing the distances” relate to the following
bullet points?

> Q7. Page 64 “D is the Jensen-Shannon distance”.

What is the question? I have mentioned that D stands for Jensen-Shannon distances.

> Q8. In Table 4.1, the R column should have 250, not 20.

Done.

> Q9. The description of the data generation on page 67 and Table 4.2 is confusing. Maybe state the
distribution of each vt.

To do.

> Q10. Where is $\mu$ in Table 4.2?

 $\mu$ is the difference between means considered for consecutive categories.

> Q11. In Section 4.3.3 what are the values of R and T?

> Q12. Figure 4.5 caption: MDS summary plots of what?

> Q13. Figure 4.7 is missing labels (a), (b) (c).

> Q14. Why do you chose to summarize 4 and 5 clusters from JS-NQT when sindex suggests 3 clusters?

> Q15. The reference to the stationarity assumption in the Discussion needs clarification.

> Q16. Page 81 “can not” should be “cannot”

Done.

> Q17. Again, the computational burden could do with more Discussion. At present, if I were to use this
method on my data, what kind of sizes are realistic to work with?

# Prof Juergen Symanzik

## Chapter 4

> Q1. There seems to be some contradiction on p. 60. One sentence states: Tureczek and Nielsen (2017) conducted a systematic study of over 2100 peer-reviewed papers on smart meter data analytics. Another sentence states: Time series data, such as smart meter data, are not well-suited to any of the techniques mentioned in Tureczek and Nielsen (2017).

> Q2. Use a consistent style and use past tense throughout when you summarize what was previously published.

Done.

> Q3. The last cross{reference at the end of the Introduction should be to Section 4.5 (and not Section 2.7).

Done.

> Q4. At the end of p. 61, the sentence 
The flow of the procedures is illustrated in Figure 4.1.
should be extended with "and is further described in the following subsections." Also match subsection headings and headers in Figure 4.1, e.g., Selecting granularities vs. Find significant granularities and more.

first part: Done
second part: Discuss (The diagram to be changed as per text or reverse)

> Q5. The text on p. 68 speaks of Figure 4.3(b) and Figure 4.3 (right). Either add letters a and b or speak of left and right.

Done.

>Q6. Similar to Chapter 3, supplementary material should be specified in more detail, e.g., on p. 68, p. 70, p. 71, and p. 75.

Done. Added links in the supplementary materials section at the end of the chapter.

> Q7. In Figure 4.5, it is hard to see which groups are overplotting, in particular for S3. To better reveal this graphically, use different (open) glyphs in addition to different colors. In particular, a + for group 1 and an open circle for group 5 may help to better distinguish the groups (and possibly $\times$, open triangle, and open box to be used for groups 2 to 4).

Discuss.

> Q8. Apparently, in Figure 4.6, the vertical axis is not standardized to [0; 1] as it is frequently done for parallel coordinate plots. Therefore, it would make sense to display an actual vertical axis in each of the three graphs.

Discuss.

> Q9. Figure 4.7 speaks of (a) and (b) in the caption, but those letters do not appear in the figure. Either add them or refer to left and right. Moreover, in previous chapters, you used orange and green for the quartiles and 10th/90th quantile. Why not using the
same colors and quantiles here? If this becomes too confusing for a reader, then at least, indicate in the caption which quantiles are covered by the gray areas.

first part: Done.
second part: Discuss

> Q10. Be consistent across chapters. For example, in Table 3.3, you use wdwnd. On p. 72 and in Figure 4.7, you use wnwd. Adjust across all chapters. Check whether other abbreviations need to be standardized as well.

Ok. To do in hakear.

> Q11. Similar to Figure 4.9, it would help the reader that cluster P-3 is visually represented via three somewhat similar colors for Q-3, Q-4, and Q-5. Changing from red to blue/purple initially hides this information. Also arrange the legend from P-1 to P-3 and Q-1 to Q-5 to make it more obvious that P-1 & Q-1 and P-2 & Q-2 are identic

Discuss.

> Q12. It would help the reader to also mention in the caption of Table 4.3 that P-1 & Q-1 and P-2 & Q-2 are identical. I first looked at Figure 4.9 and then at Table 4.3, but if read in the other order, I likely would have been surprised, to see identical values in some of the table rows (without any further explanation).

Done. (but not showing in pdf check again)

> Q13. For the thesis, the work is adequate and meaningful. For a journal paper, I would like to see some of the points outlined in the Discussion being addressed. What happens if this method is applied to the data from the 13,000 customers that were introduced earlier on? As a reader of a journal paper, I would rather like to see limitations of the method proposed here, e.g., if it does not scale up then why it most likely does not scale up. Otherwise, it is hard to assess whether it is worthwhile to try this method for one's own data.

Discuss.

### 2022-02-03 report/response.Rmd
response document updated

@@
 ---
-title: Amendments summary
+title: Response document
 author: Sayani Gupta
@@
 
+Discuss before change.
+
 > Q4. The violin plot in Figure 2.5c is barely recognisable, and comparison of the distributions is challenging.
 The plot deserves a bit more space.
+
+Discuss before change.
 
@@
 
+Discuss.
 Changed to theme_bw() instead of default and outlier.alpha (0.5 ->0.1) and geom_jitter(alpha = 0.04 -> 0.06)
@@
 
+Discuss.
 No change.
@@
 first part: Done.
-second part: To do
+second part: Discuss.
 
@@
 
+Discuss.
 To do.
@@
 
+Discuss.
 Not changed.
@@
 first part: How?
-second part: After NQT, results are pretty similar to an underlying normal distribution.
+second part: After NQT, results are pretty similar to an underlying normal distribution. (Discuss)
 
@@
 
-To do.
+To do. Discuss.
 
@@
 different colour in the heatmap for the significant comparisons. State the threshold for significance in
-the caption. In Table 3.3 the caption should explain the threshold. Maybe use colour instead of stars
-to indicate significance?
+the caption. In Table 3.3 the caption should explain the threshold. Maybe use colour instead of stars to indicate significance?
 
-To do.
+To do. Discuss.
 
@@
 
-### Chapter 4
+## Chapter 4
 
@@
 # Prof Juergen Symanzik
+
+## Chapter 3
+
+> Q1.
+In Figure 3.1, it would help the reader to mention in the figure caption that a log-scale has been used for the vertical axis.
+Moreover, it might be helpful to add a ticmark label at 10 kWh and add ticmarks representing each fraction of 10 on the log-scale and not only the mid{point between two ticmark labels.
+Also, reminding the reader that summer months are in January and February in Australia would be helpful.
+When looking at the figure (prior to reading the main text), my first interpretation for the increased energy usage in January and February was because of (electric) heating | and not cooling.
+
+Can do.
+
+> Q2. Readers may not be familiar with administrative units in Australia. So, instead of speaking of Victoria on p. 35, simply speak of Melbourne (as this seems to be the source of the data).
+
+Can do.
+
+> Q3.
+In Figure 3.2, two households are being compared.
+It would be much easier for most readers if the graphs use a common scale, thus following the small multiple principle.
+Both vertical axes should be extend to 1.5 kWh and a ticmark label should be placed there as well.
+
+Can do.
+
+> Q4. In general, I like to see ticmark labels close to the extrema of the shown data points. So, in Figure 3.3a, there should be additional ticmark labels at -3 and +3, in 3.3b at -5
+and +10, and so on.
+
+Can do.
+
+> Q5.
+Before introducing a new distance measure in Section 3.2, I would like to see a literature review of existing distance measures | and their limitations.
+Why is it necessary to introduce a new distance measure here?
+This is partially addressed on p.
+41, but that should be placed earlier in the text.
+One clarifying question: Which density is represented by f in the equation?
+And shouldn't there be a dx at the end of the integral?
+Also, the distance measures that are mentioned at the end of that paragraph should be explained in one or two sentences.
+
+Discuss.
+
+> Q6. For the data transformation steps on p. 40, a small table, say with a sample of 5 values from an exponential distribution, might be helpful to better explain each of the three steps.
+
+Discuss.
+
+> Q7. A few editorial corrections are necessary, e.g.,
+Dang and Wilkinson (2014); Wilkinson, Anand, and Grossman (2005) provide misses an \and" between the two references. The same holds for Buja et al. (2009); Majumder, Hofmann, and Cook (2013) present.
+I would leave it to the author to check for similar omissions. Moreover, these should be past tense: "provided" and "presented".
+
+Done. "and" needed?
+
+> Q8. The term Gestalt theory is mentioned a few times (p. 38 & p. 42), but it is never supported by a reference.
+
+> Q9. It would be helpful to add a specific link to a subsection in a cross-reference such as (See the supplements for more details.) on p. 42. Same on p. 47 and p. 48.
+
+Discuss. (Not adding link in supplementary section enough?)
+
+> Q10.
+On p.
+45 and in (3.2), it is not immediately clear whether nx; nf simultaneously have to be less than or equal to 5 so that wpdperm is being used - or whether only one of them has to be less than 5.
+Just a verbal clarification is needed.
+A practical scenario where this applies might be months (12) and weekdays/weekend (2).
+
+Discuss.
+
+> Q11. On p. 47, Again consider 3.1(a) and 3.1(b) seems to miss the word "Figure".
+
+Done.
+
+> Q12. In Table 3.2, listing a p-value as 0 never is a good idea. List it as < 0:01 or < 0:0001 or any other meaningful threshold that matches the number of your simulation runs.
+
+> Q13.
+The Results section and Figure 3.5 need some clarifications.
+The text states: Figure 3.5 shows that both the location and scale of the distributions change across panels.
+I suppose this figure relates to m = 2, but this is not stated in the main text.
+Table 3.2
+lists both, m = 1 and m = 2. Moreover, in Figure 3.5, the y{axis ticmark labels overplot and the nx values are partially cut off.
+
+Discuss for changes. Can do.
+
+> Q14.
+There are similar problems with the labels in Figure 3.6.
+Moreover, the two overlaid curves are very hard to distinguish.
+Would it help to change colors and transparency, or draw the boundary of the curves with the specific colors (and not in black)?
+This may require some experimentation.
+Finally, using a differently colored background (say light yellow) would be helpful to visually demonstrate where wpdperm is being used.
+For the legend, use the terms as in equation (3.2) and not the full word.
+
+Discuss
+
+> Q15. The last paragraph on p. 48 uses $<$ while equation (3.2) uses $\leqslant$. Which one is correct?
+
+Can do.
+
+> Q16. Match the text on p. 50 and introduce the abbreviations from Figure 3.7 and Table 3.3, e.g., "hod" likely is "hour day" and so on. Introducing these abbreviations on p. 53 comes
+too late.
+
+Can do.
+
+> Q17.
+In Figure 3.7a, it is not clear what the eight heatmaps represent.
+Do they belong to id 1, ..., id 8 ?
+Some explanation in the text or listing the id information on the left would be helpful.
+Similarly, on p.
+53, the text states: id 7 and 8 have the same significant harmonies.
+I do not know where to look at the heatmaps in Figure 3.7a to find this information (and what is mentioned under items 2.
+and 3.
+in the text).
+Some explanation and labeling is
+needed here.
+
+discuss.
+
+> Q18.
+How were the harmony pairs selected and sorted in Figure 3.7 and Table 3.3?
+headmaps show 25 pairs, the parallel coordinate plot shows only 14 (same as Table 3.3).
+However, the last two are sorted differently, "hod wdwnd" (the first row in the table) is neither the first nor the last row in the parallel coordinate plot.
+
+Can do.
+
+> Q19. On p. 53, you use the term inconsequential.This hasn't been introduced. Do you mean "not significant" here?
+
+Can do.
+
+> Q20. Apparently, Figure 3.8 uses a logscale again. Adjust similar to my comments for Figure 3.1.
+
+Can do.
+
+> Q21. Are parts a and b in Figure 3.8 correctly labeled and also matching with the figure caption? The caption reads For id 1, patterns look similar within weekdays and weekends.
+Visually, this seems to be the case for id 7 (in part b). This also visually matches Figure 3.7b.
+
+Discuss.
+
+> Q22.
+I didn't ask earlier, but how were the eight households for Section 3.4 being selected?
+Was this based on a random sample, or manually do obtain rather different energy usage patterns?
+This should be answered earlier on, but this also should be addressed in the Discussion: What can be expected when working with the full data set of Chapter 2, i.e., about 13,000.
+You indicate:
+A future direction of work is to be able to explore and compare many individuals/subjects together for similar patterns across significant granularities.
+What would be computationally (and visually) feasible as 13,000 is a few orders of magnitude
+higher than the number of eight households used in this example? Clearly, no full answer is expected here, but rather some speculation of what could possibly be done in
+the future (or what is done in the next chapter).
+
+Discuss.
+
+> Q23.The Supplementary materials section mentions data and scripts and a supplementary paper.
+For the thesis, it would be helpful to indicate where these can be located on github.
+There is only a link for the R package.
+
+Can do.
 

### 2022-02-03 Rmd/03-hakear.Rmd
response document updated

@@
 Therefore, even for small $v$, the number of scatterplots can be large, and scatterplot matrices (SPLOMs) can easily run out of pixels when presenting high-dimensional data.
-@Dang2014-tw; @wilkinson2005graph provide potential solutions to this, where a few characterizations can be used to locate anomalies in density, shape, trend, and other features in the 2D point scatters.
+@Dang2014-tw; @wilkinson2005graph provided potential solutions to this, where a few characterizations can be used to locate anomalies in density, shape, trend, and other features in the 2D point scatters.
 
@@
 
-@inference; @Majumder2013-hb present methods for statistical significance testing of visual findings using human cognition as the statistical tests.
+@inference; @Majumder2013-hb presented methods for statistical significance testing of visual findings using human cognition as the statistical tests.
 In this paper, the visual discovery of distributional differences is facilitated by choosing a threshold for the proposed numerical distance measure, eventually selecting only those cyclic granularities for which the distributional differences are sufficient to make it an interesting display.
@@
 
-Again consider \ref{fig:onegran-new}(a) and \ref{fig:onegran-new}(b) with a $\wpd$ value of 20.5 and 145 respectively.
+Again consider Figures \ref{fig:onegran-new}(a) and \ref{fig:onegran-new}(b) with a $\wpd$ value of 20.5 and 145 respectively.
 This is because there is a more gradual increase across hours of the day than across months of the year.

### 2022-02-07 Rmd/02-gravitas.Rmd
Figure 2.4 Catherine

@@
 
-(ref:allFig) Distribution of energy consumption displayed as letter value plots, illustrating harmonies and clashes, and how mappings change emphasis: **a** weekday/weekend faceted by quarter-of-year produces a harmony, **b** quarter-of-year faceted by weekday/weekend produces a harmony, **c** month-of-year faceted by quarter-of-year produces a clash, as indicated by the empty sets and white space.
+(ref:allFig) Distribution of energy consumption displayed on a logarithmic scale as letter value plots, illustrating harmonies and clashes, and how mappings change emphasis: **a** weekday/weekend faceted by quarter-of-year produces a harmony, **b** quarter-of-year faceted by weekday/weekend produces a harmony, **c** month-of-year faceted by quarter-of-year produces a clash, as indicated by the empty sets and white space.
 Placement within a facet should be done for primary comparisons.

### 2022-02-07 thesis.Rmd
Figure 2.4 Catherine

@@
 
-# Copyright notice {- #ch:copyright}
+# Copyright notice {- #ch-copyright}
 
@@
 
-# Abstract {- #ch:abstract}
-
-The research was motivated by the desire to understand the Australian smart meter data, which was collected half-hourly for two years at the household level.
-With temporal data available at ever finer scales,
-exploration may be required to be carried out across both finer and coarser scales to draw useful inferences about the underlying process.
+# Abstract {- #ch-abstract}
+
+The research was motivated by the desire to understand some Australian smart meter data, which was collected half-hourly for two years at the household level.
+With temporal data available at ever finer scales, exploring periodicity can become overwhelming with so many possible temporal deconstructions to explore.
 Analysts are expected to comprehensively explore the many ways to view and consider temporal data.
@@
 This work investigates how time may be dissected, resulting in alternative data segmentation and, as a result, different visualizations that can aid in the identification of underlying patterns.
-The first contribution (Chapter \@ref(ch:gravitas)) describes classes of time deconstructions using linear and cyclic time granularities.
+The first contribution (Chapter \@ref(ch-gravitas)) describes classes of time deconstructions using linear and cyclic time granularities.
 It provides tools to compute possible cyclic granularities from an ordered (usually temporal) index and also a framework to systematically explore the distribution of a univariate variable conditional on two cyclic time granularities by defining "harmony".
-"harmony" denotes pairs of granularities that could be effectively analyzed together and reduces the search from all possible options.
-This approach is still overwhelming for human consumption due of the vast amount of harmonies remaining.
-The second contribution (Chapter \@ref(ch:hakear)) refines the search for informative granularities by identifying those for which the differences between the displayed distributions are greatest and also rating them in order of importance of capturing maximum variation.
-The third contribution (Chapter \@ref(ch:gracsr)) builds upon the first two to provide methods for exploring heterogeneities in repetitive behavior for many time series and over multiple granularities.
+A "harmony" denotes pairs of granularities that could be effectively analyzed together and reduces the search from all possible options.
+This approach is still overwhelming for human consumption due to the vast number of harmonies remaining.
+The second contribution (Chapter \@ref(ch-hakear)) refines the search for informative granularities by identifying those for which the differences between the displayed distributions are greatest and also rating them in order of importance of capturing maximum variation.
+The third contribution (Chapter \@ref(ch-gracsr)) builds upon the first two to provide methods for exploring heterogeneities in repetitive behavior for many time series and over multiple granularities.
 It accomplishes this by providing a way to cluster time series based on probability distributions across informative cyclic granularities.
@@
 
-`r if (knitr::is_html_output()) '<!--'`
-
-\newpage
-
-# Declaration {-#ch:declare}
-
-I hereby declare that this thesis contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.
-
-Three publications are included in this thesis, one of which has been accepted for publication in the Journal of Computational and Graphical Statistics (@Gupta2021-gravitas).
-Two others have not yet been submitted.
-"Exploring probability distributions across cyclic temporal granularities" is the central theme of the thesis.
-All of the papers in the thesis were conceptualized, developed, and written by myself, the student, while working in the Department of Econometrics and Business Statistics under the supervision of Professor Rob J Hyndman and Professor Dianne Cook.
-When co-authors are listed, it indicates that the work was produced through active collaboration between researchers and that their contributions to team-based research have been recognized.
-The following table details the work, including
-my and my fellow co-authors contributions:
-
-\begin{table}
-\centering\footnotesize\tabcolsep=0.12cm
-\begin{tabular}{|p{1cm}|p{2cm}|p{1.5cm}|p{3.5cm}|p{3.5cm}|p{1.5cm}|}
-\hline
-\RaggedRight\textbf{Thesis Chapter}  & \RaggedRight\textbf{Publication Title}  & \RaggedRight\textbf{Status (published, in press, accepted or returned for revision)}  & \RaggedRight\textbf{Nature and} {\%} \RaggedRight\textbf{of student contribution} & \RaggedRight\textbf{Co-author name(s) Nature and} {\%} \RaggedRight\textbf{of Co-author’s contribution} & \RaggedRight\textbf{Co-author(s), Monash student Y/N} \\ \hline
-2 & Visualizing probability distributions across bivariate cyclic temporal granularities & Published & 70\%.
-Concept and developing software and writing first draft & (1) Rob J Hyndman, input into manuscript 15\% (2) Dianne Cook, Concept and input into manuscript 15\% & \multicolumn{1}{c|}{N} \\
-\hline
-3 & Detecting distributional differences between temporal granularities
-for exploratory time series analysis & 80\%.
-Concept and developing software and writing first draft & (1) Rob J Hyndman, Concept and input into manuscript 10\% (2) Dianne Cook, Concept and input into manuscript 10\%  & \multicolumn{1}{c|}{N} \\
-\hline
-4 & Clustering time series based on probability distributions across temporal granularities & To be submitted & 70\%.
-Concept and developing software and writing first draft & (1) Dianne Cook, Concept and input into manuscript 20\% (2) Rob J Hyndman, concept and input into manuscript 10\% & \multicolumn{1}{c|}{N} \\
-\hline
-\end{tabular}
-\end{table}
-
-\FloatBarrier
-
-I have not renumbered sections of submitted or published papers in order to generate a consistent presentation within the thesis.
+# Declaration {-#ch-declare}
+
+This thesis is an original work of my research and contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.
+
+\vspace{4em}
+
+\textbf{Student signature:}
+
+```{r signature, out.width = "20%", fig.align = "left"}
+knitr::include_graphics("img/signature.png")
+```
 
@@
 
-\textbf{Student signature:}
-
-```{r signature, out.width = "20%", fig.align = "left", eval = file.exists("img/signature.png")}
-knitr::include_graphics("img/signature.png")
-```
-
-\textbf{Date:} `r Sys.Date()`
-
-`r if (knitr::is_html_output()) '-->'`
+\textbf{Date:} `r as.Date("2021-11-19")`
 
@@
 
-# Acknowledgements {-#ch:thanks}
+# Acknowledgements {-#ch-thanks}
 
@@
 Thank you, Di, for being a fantastic female role model.
+I would also like to express my gratitude to my supervisor Dr.
+Peter Toscas of Data61 CSIRO, for always being available to help and offering feedback on practical elements of analyzing smart meter data at our monthly meetings.
 
@@
 
-I want to thank the Department of Econometrics and Business Statistics, Monash Business School, ARC Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS), and Data61 CSIRO for their financial support.
+I want to thank the Department of Econometrics and Business Statistics, Monash Business School, ARC Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS), Data61 CSIRO for their financial support.
 Thanks to the Department, I attended a few conferences in person (before COVID hit) and had a great time networking and being inspired by many people.
 A few highlights are UseR!
-2018 (Brisbane), Young Statistician's Conference 2019 (Canberra), WOMBATS 2019 (Melbourne), and RopenSci 2019 (Sydney).
+2018 (Brisbane), Young Statistician's Conference 2019 (Canberra), WOMBAT 2019 (Melbourne), and ROpenSci 2019 (Sydney).
 Thank you to Gael Martin, Farshid Wahid, Klaus Ackermann, and Didier Nibbering from my PhD committee for providing such a supportive environment.
@@
  <!-- - PhD friends and R community -->
-Thank you, Stephanie and Nicholas Tierney, for being so warm, kind and welcoming when I began my PhD.
+Thank you, Puwasala Gamakumara, Stephanie Kobakian and Nicholas Tierney, for being so warm, kind and welcoming when I began my PhD.
 Being a part of such an inclusive environment meant a lot to me as an international student.
 Thank you, Nicholas Spyrison and Mitchell O'Hara-Wild, for the numerous informative discussions, brainstorming sessions, as well as for generously sharing your expertise and experience in software development.
+A special thanks goes to Stephanie, Emi Tanaka, Nicholas Tierney, and Nicholas Spyrison for taking the time out of their busy schedules to proofread parts of my thesis.
 Thank you to each of you, NUMBATs, and the broader R community for directly and indirectly helping me countless times.
@@
 
-PhD in Covid has been harder, and I would not have been able to make it through this if not for the emotional support of my friends and family.
+Doing a PhD in COVID has been hard, and I would not have been able to make it through this if not for the emotional support of my friends and family.
 Thank you Puwasala, Sium, and Ian, for being so kind and acting as a pillar of emotional support, especially during the last leg of my journey.
@@
 Thank you to each one of you for being available, listening to me, and putting things in perspective when I lost sight.
-Thanks to my housemates Anjali, Surbhi, and Dulaji for the fun company and conversations that kept me sane in the covid lockdowns.
+Thanks to my housemates Anjali, Surbhi, and Dulaji for the fun company, food and conversations that kept me sane in the COVID lockdowns.
 
@@
 Thank you, Juhi (sister-in-law), for reminding me of the value of organization in all aspects of life and the importance of prioritizing my physical health from time to time.
-Thank you all for your part in supporting me in this journey.
-I love you.
-
+Lastly, thank you to my one-year-old nephew (Rayan) and niece (Mehr) for being the ultimate stress relievers. Thank you all for being a part of this journey with me. I love you.
+
+<!-- Thank you for making me feel at home away from home, Somdatta (cousin). -->
 <!-- add people who are helping you with proofreading -->
-<!-- Add Peter -->
 <!-- Although the journey is coming to an end, you wish to take these life learnings forward -->
@@
 
+Throughout my PhD candidature, I have had the great privilege of working with and being inspired by my advisers and many other people, which resulted in a great lot of learning in both academics and life.
+Although the journey is coming to a close, I want to carry these life lessons forward, and I would like to conclude with one of my favorites from Rob.
+"Be kind; remember you are collaborating with your future self."
+
 <!--chapter:end:Rmd//00-d-acknowlegements.Rmd-->
 
-# Preface {-#ch:preface}
-
-Chapter \@ref(ch:gravitas) has been accepted by the *Journal of Computational and Graphical Statistics*.
+# Preface {-#ch-preface}
+
+Chapter \@ref(ch-gravitas) has been published online at the [_Journal of Computational and Graphical Statistics_](https://www.tandfonline.com/eprint/NXKCSZHIBFXCJD2NKB68/full?target=10.1080/10618600.2021.1938588).
 It has won the ACEMS Business Analytics Prize in 2020.
-The accompanying R package **gravitas** is on CRAN.
-Chapter \@ref(ch:hakear) and Chapter \@ref(ch:gracsr) are yet to be submitted.
+The accompanying R package `gravitas` is on CRAN.
+Chapter \@ref(ch-hakear) and Chapter \@ref(ch-gracsr) are yet to be submitted.
 
@@
 
-This thesis is written in R Markdown [@rmarkdown] with **bookdown** [@bookdown], using **renv** [@renv] to create reproducible environments.
-The online version of this thesis is hosted at <https://sayani.netlify.app/>, powered by [Netlify](http://netlify.com).
-All materials (including the data sets and source files) required to reproduce this document can be found at the Github repository <https://github.com/Sayani07/thesis>.
+This thesis is written in R Markdown [@rmarkdown] with `bookdown` [@bookdown], using `renv` [@renv] to create reproducible environments.
+The online version of this thesis is hosted at <https://sayani07.github.io/thesis-SG>.
+All materials (including the data sets and source files) required to reproduce this document can be found at the Github repository <https://github.com/Sayani07/thesis-SG>.
 
@@
 
-# Introduction {#ch:intro}
+# Introduction {#ch-intro}
 
@@
 
-The Smart Grid, Smart City (SGSC) project (2010–2014) available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) provides half-hourly data of over 13,000 household electricity smart meters distributed unevenly from October 2011 to March 2014 for some parts in Australia.
-Because of the increasing variety of customers, larger data sets contain more uncertainty about customer behavior.
-Household behavioral patterns vary significantly due to differences in size, location, and amenities such as solar panels, central heating, and air conditioning.
+The Smart Grid, Smart City (SGSC) project (2010–2014) available through the [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) provides half-hourly data of over 13,000 Australian electricity smart meter customers distributed unevenly from October 2011 to March 2014.
+The wide variety of customers means that there will be large variance in behavior, leading to greater uncertainty in the data.
+Behavioral patterns vary significantly due to differences in size, location, and amenities such as solar panels, central heating, and air conditioning.
 For example, some families use a dryer, while others hang their clothes to dry.
@@
 Customers' day-off energy consumption varies depending on whether they stay at home or go outside.
-The motivation for this work comes from the desire to provide methods to better understand heterogeneity in repetitive behavior that can arise across temporal dependencies.
 
 <!-- introducing the overall field-->
-With the availability of data at more and more finer time scales, exploration of time series data may be required to be carried out across both finer and coarser scales to draw useful inferences about the underlying process.
-For example, data collected at an hourly scale could be analyzed using coarser temporal scales such as days, months or quarters.
+With the availability of data at finer and finer time scales, exploration of time series data may be required to be carried out across both finer and coarser scales to draw useful inferences about the underlying process.
+To reduce the complexity of time, it is typical to divide it into years, months, weeks, days, and so on in a hierarchical manner [@aigner2011visualization].
+These discrete abstractions of time are known as time granularities.
+Linear time granularities [@Bettini1998-ed], such as hours, days, weeks and months, respect the linear progression of time and are non-repeating.
+Cyclic temporal granularities representing cyclical repetitions in time (such as hour-of-the-day, work-day/weekend) are effective for analyzing repetitive patterns in time series data.
+
+To acquire a comprehensive view of the repeated patterns, it is necessary to navigate through all of the conceivable cyclic granularities.
+This approach is consistent with the concept of exploratory data analysis EDA [@Tukey1977-jx], which stresses the utilization of multiple perspectives on data to assist with formulating hypotheses before proceeding to formal inferences or modeling.
+This, however, is a challenging process since it throws up a myriad of possible hypotheses.
+Furthermore, the transition from linear to cyclic granularities results in restructured data, with each level of the temporal deconstruction corresponding to multiple values of the observed variable.
+This motivates the research presented in this thesis, which aims to provide a platform for systematically exploring probability distributions induced by these multiple observations to support the discovery of regular patterns or anomalies, as well as the exploration of clusters of behaviors or the summarization of the behavior.
+
+<!-- There could potentially be many, many ways to analyze the data across cyclic granularities. -->
+
+<!-- For example, data collected at an hourly scale could be analyzed using coarser temporal scales such as days, months or quarters.
 This approach requires deconstructing time in various possible ways.
@@
 ## Visualizing probability distributions across bivariate cyclic temporal granularities {#sec:gravitas}
+
+Chapter \@ref(ch-gravitas) describes classes of time deconstructions using linear and cyclic time granularities, which can be used to create data visualizations to explore periodicities, associations, and anomalies.
+It provides a formal characterization of cyclic granularities and facilitates manipulation of single- and multiple-order-up time granularities through cyclic calendar algebra, as well as providing a recommendation algorithm to check the feasibility of creating plots for any two cyclic granularities.
+Our proposed method is also applicable to non-temporal hierarchical granularities with an underlying ordered index.
+The methods are implemented in the open-source R package `gravitas` and are consistent with a tidy workflow [@wickham2016r], with probability distributions examined using the range of graphics available in `ggplot2` [@Wickham2009pk].
+
+## Detecting distributional differences between temporal granularities for exploratory time series analysis {#sec:hakear}
+
+Chapter \@ref(ch-hakear) is a natural extension of Chapter \@ref(ch-gravitas).
+Many displays might be built using cyclic granularities.
+However, only a handful of them may reveal major patterns of interest.
+Identifying the displays which exhibit "significant" distributional differences and plotting only these would allow for more efficient exploration.
+Furthermore, a few of the displays in this collection will be more engaging than others.
+Chapter \@ref(ch-hakear) provides a new distance metric for selecting and ranking the multiple granularities.
+The statistical significance of potential visual discoveries is aided by selecting a threshold for the proposed numerical distance measure.
+The distance measure is computed for a single or pairs of cyclic granularities, and it can be compared across different cyclic granularities as well as a collection of time series.
+This chapter also includes a case study using residential smart meter data from Melbourne to demonstrate how the suggested methodology may be utilized to automatically find temporal granularities with significant distributional differences.
+The methods are implemented in the open-source R package `hakear`.
+
+## Clustering time series based on probability distributions across temporal granularities {#sec:gracsr}
+
+In Chapter \@ref(ch-gracsr), we look at the problem of using clustering to discover patterns in a large number of univariate time series across multiple temporal granularities.
+Time series clustering research is gaining traction as more data is collected at finer temporal resolution, over longer time periods, and for a larger number of individuals/entities.
+Many disciplines have noisy, patchy, uneven, and asynchronous time series that make it difficult to search for similarities.
+We propose a method for overcoming these constraints by calculating distances between time series based on probability distributions at various temporal granularities.
+Because they are based on probability distributions, these distances are resistant to missing or noisy data and aid in dimension reduction.
+When fed into a clustering algorithm, the distances can be used to divide large data sets into small pockets of similar repetitive behaviors.
+These subgroups can then be analyzed separately or used as distinct prototype behaviors in classification problems.
+The proposed method was tested on a group of residential electricity consumers from the Australian smart meter data set to show that it can generate meaningful clusters.
+This chapter includes a brief review of the literature on traditional time series clustering and, more specifically, clustering residential smart meter data.
+
+<!-- Time series clustering research is gaining a lot of importance with more and more time series data recorded at much finer temporal resolution, for a longer period of time and for a larger number of individuals/entities.Similarity searches amongst these long time series are often limited by the type of noisy, patchy, unequal and asynchronous time series common in many fields.
+In this paper we propose a strategy to alleviate these limitations by  clustering time series based on probability distributions across cyclic temporal granularities.
+Cyclic granularities are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, which can be useful for measuring repetitive patterns in large univariate time series data.
+Thus, looking at the probability distributions across cyclic granularity leads to a method which is robust to missing or noisy data, helps in dimension reduction while ensuring small pockets of similar repetitive behaviors.
+The method is applied to electricity smart meter dat -->
+<!-- The suggested approach was evaluated on a set of benchmark time series on residential electricity customers. The empirical results show that our approach is able to yield meaningful clusters. -->
+
+## Thesis structure
+
+The thesis is structured as follows.
+Chapter \@ref(ch-gravitas) provides details of the cyclic granularities, different classes, and computation, and also its usage in exploratory time series analysis through applications.
+This is implemented in the R package `gravitas`.
+Chapter \@ref(ch-hakear) provides guidance on how to choose significant cyclic granularities, which are likely to have interesting patterns across its categories.
+This is available as the R package `hakear`.
+Chapter \@ref(ch-gracsr) provides methods to explore heterogeneity in repetitive behavior for multiple time series over multiple cyclic granularities.
+This is in the developing R package `gracsr`.
+Chapter \@ref(ch-conclusion) summarizes the software tools developed for the work, and discusses future plans.
+
+<!--chapter:end:Rmd//01-intro.Rmd-->
+
+# Visualizing probability distributions across bivariate cyclic temporal granularities {#ch-gravitas}
 
@@
 
- <!-- This chapter contains a literature review related to linear granularities. -->
-
-## Detecting distributional differences between temporal granularities for exploratory time series analysis {#sec:hakear}
-
-This work is a natural extension of Chapter \@ref(ch:gravitas).
-Many, many other displays might be built using different granularities, such as day-of-week, day-of-month, weekday/weekend, and so on.
-However, only a handful of them may be interesting, revealing major patterns.
-Determining which displays exhibit "significant" distributional differences between cyclic granularity categories and plotting only these would allow for more efficient exploration.
-Furthermore, a few of the displays in this collection will be more engaging than others.
-This chapter provides a new distance metric for selecting and ranking the multiple granularities.
-The statistical significance of potential visual discoveries is aided by selecting a threshold for the proposed numerical distance measure.
-The distance measure is computed for a single or pairs of cyclic granularities, and it can be compared across different cyclic granularities as well as a collection of time series.
-This chapter also includes a case study using residential smart meter data from Melbourne to demonstrate how the suggested methodology may be utilized to automatically find temporal granularities with significant distributional differences.
-The methods are implemented in the open-source R package 'hakear.'
-
-## Clustering time series based on probability distributions across temporal granularities {#sec:gracsr}
-
-In this chapter, we look at the problem of using clustering to discover patterns in a large number of univariate time series across multiple temporal granularities.
-Time series clustering research is gaining traction as more data is collected at finer temporal resolution, over longer time periods, and for a larger number of individuals/entities.
-Many disciplines have noisy, patchy, uneven, and asynchronous time series that make it difficult to search for similarities.
-We propose a method for overcoming these constraints by calculating distances between time series based on probability distributions at various temporal granularities.
-Because they are based on probability distributions, these distances are resistant to missing or noisy data and aid in dimension reduction.
-When fed into a clustering algorithm, the distances can be used to divide large data sets into small pockets of similar repetitive behavior.
-These subgroups can then be analyzed separately or used as distinct prototype behaviors in classification problems.
-The proposed method was tested on a group of residential electricity consumers from the Australian smart meter data set to show that it can generate meaningful clusters.
-This chapter includes a brief review of the literature on traditional time series clustering and, more specifically, clustering residential smart meter data.
-
-<!-- Time series clustering research is gaining a lot of importance with more and more time series data recorded at much finer temporal resolution, for a longer period of time and for a larger number of individuals/entities.Similarity searches amongst these long time series are often limited by the type of noisy, patchy, unequal and asynchronous time series common in many fields.
-In this paper we propose a strategy to alleviate these limitations by  clustering time series based on probability distributions across cyclic temporal granularities.
-Cyclic granularities are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, which can be useful for measuring repetitive patterns in large univariate time series data.
-Thus, looking at the probability distributions across cyclic granularity leads to a method which is robust to missing or noisy data, helps in dimension reduction while ensuring small pockets of similar repetitive behaviors.
-The method is applied to electricity smart meter dat -->
-<!-- The suggested approach was evaluated on a set of benchmark time series on residential electricity customers. The empirical results show that our approach is able to yield meaningful clusters. -->
-
-## Summary
-
-The thesis is structured as follows.
-Chapter \@ref(ch:gravitas) provides details of the cyclic granularities, different classes, and computation, and also its usage in exploratory time series analysis through applications.
-This is implemented in the R package **gravitas**.
-Chapter \@ref(ch:hakear) provides guidance on how to choose significant cyclic granularities, which are likely to have interesting patterns across its categories.
-This is available as the R package **hakear**.
-The chapter @ref (ch: gracsr) provides similarity measures for comparing multiple time series.This is in the developing R package **gracsr**.
-Chapter \@ref(ch:conclusion) summarizes the software tools developed for the work, and discusses some future plans.
-
-<!--chapter:end:Rmd//01-intro.Rmd-->
-
-# Visualizing probability distributions across bivariate cyclic temporal granularities {#ch:gravitas}
-
-Deconstructing a time index into time granularities can assist in exploration and automated analysis of large temporal data sets.
-This paper describes classes of time deconstructions using linear and cyclic time granularities.
-Linear granularities respect the linear progression of time such as hours, days, weeks and months.
-Cyclic granularities can be circular such as hour-of-the-day, quasi-circular such as day-of-the-month, and aperiodic such as public holidays.
-The hierarchical structure of granularities creates a nested ordering: hour-of-the-day and second-of-the-minute are single-order-up.
-Hour-of-the-week is multiple-order-up, because it passes over day-of-the-week.
-Methods are provided for creating all possible granularities for a time index.
-A recommendation algorithm provides an indication whether a pair of granularities can be meaningfully examined together (a "harmony"), or when they cannot (a "clash").
-
- Time granularities can be used to create data visualizations to explore for periodicities, associations and anomalies.
-The granularities form categorical variables (ordered or unordered) which induce groupings of the observations.
-Assuming a numeric response variable, the resulting graphics are then displays of distributions compared across combinations of categorical variables.
-
- The methods implemented in the open source R package `gravitas` are consistent with a tidy workflow, with probability distributions examined using the range of graphics available in `ggplot2`.
-
 ```{r external, include = FALSE, cache = FALSE}
@@
 Although we were motivated by the smart meter example, the problem and the solutions we propose are practically relevant to any temporal data observed more than once per year.
-In a broader sense, it could be even suitable for data observed by years, decades, and centuries as might be in weather or astronomical data.
-
-This work provides tools for systematically exploring bivariate granularities within the tidy workflow (@wickham2016r). In particular, we
+In a broader sense, it could be even suitable for data observed by years, decades, and centuries as might be the case in weather or astronomical data.
+
+This work provides tools for systematically exploring bivariate granularities within the tidy workflow [@wickham2016r]. In particular, we
 
@@
 
-If $G$ groups into $H$, it would imply that any granule $H(i)$ is the union of some granules of $G$, for example, $G(a_1), G(a_2), \dots, G(a_k)$.
+If $G$ groups into $H$, it would imply that any granule $H(i)$ is the union of some granules of $G$; for example, $G(a_1), G(a_2), \dots, G(a_k)$.
 Condition (2) in Definition \ref{def:periodic} implies that if $H(i + R) \neq \emptyset$, then $H(i + R) = \bigcup (G(a_1 + P), G(a_2 + P), \dots, G(a_k + P))$, resulting in a "periodic" pattern of the composition of $H$ using granules of $G$.
 In this pattern, each granule of $H$ is shifted by $P$ granules of $G$.
-$P$ is called the \textbf{Period} (@Bettini2000-qk).
+$P$ is called the \textbf{period} [@Bettini2000-qk].
 
@@
 \begin{table}[ht]
+\caption{Examples of circular granularities with bottom granularity minutes.
+Circular granularity $C_i$ relates two linear granularities, one of which groups periodically into the other with regular mapping and period $P_i$.
+Circular granularities can be expressed using modular arithmetic due to their regular mapping.}
 \begin{center}
@@
 
-\noindent In the Mayan calendar (Table \@ref(tab:tab-mayan)), kin-of-tun or kin-of-baktun are examples of multiple-order-up granularities and single-order-up granularities are kin-of-uinal, uinal-of-tun etc.
+\noindent In the Mayan calendar (\autoref{tab:tab-mayan}), kin-of-tun or kin-of-baktun are examples of multiple-order-up granularities and single-order-up granularities are kin-of-uinal, uinal-of-tun etc.
 
@@
 
-```{r allFig, fig.pos="!p", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap= "(ref:allFig)", fig.show = 'hold'}
-```
-
-(ref:allFig) Distribution of energy consumption displayed as letter value plots, illustrating harmonies and clashes, and how mappings change emphasis: **a** weekday/weekend faceted by quarter-of-year produces a harmony, **b** quarter-of-year faceted by weekday/weekend produces a harmony, **c** month-of-year faceted by quarter-of-year produces a clash, as indicated by the empty sets and white space.
+```{r allFig, fig.pos="!p", fig.align= 'left',echo=FALSE, eval=TRUE, warning=FALSE,message = FALSE, fig.cap= "(ref:allFig)", fig.show = 'hold', fig.height=6}
+```
+
+(ref:allFig) Distribution of energy consumption displayed on a logarithmic scale as letter value plots, illustrating harmonies and clashes, and how mappings change emphasis: **a** weekday/weekend faceted by quarter-of-year produces a harmony, **b** quarter-of-year faceted by weekday/weekend produces a harmony, **c** month-of-year faceted by quarter-of-year produces a clash, as indicated by the empty sets and white space.
 Placement within a facet should be done for primary comparisons.
 For example, arrangement in **a** makes it easier to compare across weekday type (x-axis) within a quarter (facet).
-It can be seen that in quarter 2, there is more mass occupied the lower tail on the weekends (letter value E corresponding to tail area 1/8) relative to that of the weekdays (letter value D 1/16), which corresponds to more days with lower energy use in this period.
+It can be seen that in quarter 2, more mass occupied the lower tail on the weekends (letter value E corresponding to tail area 1/8) relative to that of the weekdays (letter value D 1/16), which corresponds to more days with lower energy use in this period.
 
@@
 The Github repository,  [github.com/Sayani07/paper-gravitas](https://github.com/Sayani07/paper-gravitas), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
-This article was created with `knitr` [@knitr], [@R-knitr] and `rmarkdown` [@rmarkdown], [@R-rmarkdown].
-
-## Supplementary Materials {- #supplement-gravitas}
+This article was created with `knitr` [@knitr; @R-knitr] and `rmarkdown` [@rmarkdown; @R-rmarkdown].
+
+## Supplementary materials {- #supplement-gravitas}
 
@@
 
-# Detecting distributional differences between temporal granularities for exploratory time series analysis  {#ch:hakear}
+# Detecting distributional differences between temporal granularities for exploratory time series analysis  {#ch-hakear}
 
@@
 This work proposes a new distance metric to screen and rank the possible granularities, and hence choose the most interesting ones to plot.
-The distance measure is computed for a single or pairs of cyclic granularities can can be compared across different cyclic granularities and also on a collection of time series.
-The methods are implemented in the open-source R package `hakear`.
+The distance measure is computed for a single or pairs of cyclic granularities that can be compared across different cyclic granularities and also on a collection of time series.
+The methods are implemented in the open-source R package [`hakear`](https://github.com/Sayani07/hakear).
 
@@
 
-Cyclic temporal granularities [@Bettini1998-ed], [@Gupta2021-hd] are <!--theoretical--> temporal deconstructions that define cyclic repetitions in time, e.g.
+Cyclic temporal granularities [@Bettini1998-ed;@Gupta2021-gravitas] are <!--theoretical--> temporal deconstructions that define cyclic repetitions in time, e.g.
 hour-of-day, day-of-month, or regularly scheduled public holidays.
@@
 Daily peaks occur in morning and evening hours, indicating a working household, where members leave for and return from work.
-More volatility of usage in summer months (Jan, Feb) is probably due air conditioner use on just some days ."}
+More volatility of usage in summer months (Jan, Feb) is probably due to air conditioner use on just some days.", fig.pos="!b"}
+```
+
+As a motivating example, consider Figure \ref{fig:onegran-new} which shows electricity smart meter data plotted against two granularities (hour-of-day, month-of-year).
+The data was collected on a single household in Melbourne, Australia, over a six month period, and was previously used in @wang2020calendar.
+The categorical variable (granularity) is mapped to the x-axis, and the distribution of the response variable is displayed using both side-by-side jittered dotplots and boxplots.
+From panel (a) it can be seen that energy consumption is higher during the morning hours (5--8), when members in the household wake up, and again in the evening hours (17--20), possibly when members get back from work.
+In addition, the largest variation in energy use is in the afternoon hours (12--16), as seen in the length of the boxes.
+From panel (b), it is seen that the variability in energy usage is higher in Jan and Feb, probably due to the usage of air conditioners on some days.
+The median usage is highest in January, dips in February--April and rises again in May--June, although not to the height of January usage.
+This suggests that the household does not use as much electricity for heating as it does for air conditioning.
+A lot of households in Victoria use gas heating and hence the heater use might not be reflected in the electricity data.
+
+Many different displays could be constructed using different granularities including day-of-week, day-of-month, weekday/weekend, etc.
+However, only a few might be interesting and reveal important patterns in energy usage.
+Determining which displays have "significant" distributional differences between categories of the cyclic granularity, and plotting only these, would make for efficient exploration.
+
+```{r intro_all}
+
+```
+
+```{r id2-new2, fig.cap = "Distribution of energy consumption displayed through area quantile plots across two cyclic granularities month-of-year and hour-of-day and two households.
+The black line is the median, whereas the orange band covers the 25th to 75th percentile and the green band covers the 10th to 90th percentile.
+Difference between the 90th and 75th quantiles is less for (Jan, Feb) for the first household (a), suggesting that it is a more frequent user of air conditioners than the second household (b).
+Distribution of energy usage for (a) changes across both granularities, whereas for (b) daily pattern stays same irrespective of the months."}
 ```
@@
 
-\noindent This problem is similar to Scagnostics (Scatterplot Diagnostics) by @tukey1988computer, which is used to identify meaningful patterns in large collections of scatterplots.
-Given a set of $v$ variables, there are $v(v-1)/2$ pairs of variables, and thus the same number of possible pairwise scatterplots.
-Therefore, even for small $v$, the number of scatterplots can be large, and scatterplot matrices (SPLOMs) could easily run out of pixels when presenting high-dimensional data.
-@Dang2014-tw and @wilkinson2005graph provide potential solutions to this, where few characterizations can be used to locate anomalies in density, shape, trend, and other features in the 2D point scatters.
-In this paper, we provide a solution to narrowing down the search from $^{N_C}P_2$ plots by introducing a new distance measure that can be used to detect significant distributional differences across cyclic granularities.<!-- harmonies and why it is not enough --> This work is a natural extension of our previous work [@Gupta2021-hd] that narrows down the search from $^{N_C}P_2$ plots by identifying pairs of granularities that can be meaningfully examined together (a "harmony"), or when they cannot (a "clash").
-However, even after excluding clashes, the list of harmonies left could be enormous for exhaustive exploration.
-Hence, there is a need to reduce the search even further by including only those harmonies which are informative enough.
-@inference and @Majumder2013-hb present methods for statistical significance testing of visual findings using human cognition as the statistical tests.
+This problem is similar to Scagnostics (Scatterplot Diagnostics) by @tukey1988computer, which are used to identify meaningful patterns in large collections of scatterplots.
+Given a set of $v$ variables, there are $p(p-1)/2$ pairs of variables, and thus the same number of possible pairwise scatterplots.
+Therefore, even for small $v$, the number of scatterplots can be large, and scatterplot matrices (SPLOMs) can easily run out of pixels when presenting high-dimensional data.
+@Dang2014-tw; @wilkinson2005graph provided potential solutions to this, where a few characterizations can be used to locate anomalies in density, shape, trend, and other features in the 2D point scatters.
+
+In this paper, we provide a solution to narrowing down the search from $N_C(N_C-1)$ conditional distribution plots by introducing a new distance measure that can be used to detect significant distributional differences across cyclic granularities.<!-- harmonies and why it is not enough --> This work is a natural extension of our previous work [@Gupta2021-gravitas] which narrows down the search from $N_C(N_C-1)$ plots by identifying pairs of granularities that can be meaningfully examined together (a "harmony"), or when they cannot (a "clash").
+However, even after excluding clashes, the list of harmonies left may be too large for exhaustive exploration.
+Hence, there is a need to reduce the search even further by including only those harmonies that contain useful information.
+
+@inference; @Majumder2013-hb presented methods for statistical significance testing of visual findings using human cognition as the statistical tests.
 In this paper, the visual discovery of distributional differences is facilitated by choosing a threshold for the proposed numerical distance measure, eventually selecting only those cyclic granularities for which the distributional differences are sufficient to make it an interesting display.
@@
 
-\noindent The article is organized as follows.
-Section \ref{sec:computation-wpd} introduces a distance measure for detecting distributional difference in temporal granularities, which enables identification of patterns in the time series data;
+The article is organized as follows.
+Section \ref{sec:computation-wpd} introduces a distance measure for detecting distributional difference in temporal granularities for a continuous univariate dependent variable.
+This enables identification of patterns in the time series data;
 Section \ref{sec:rank-wpd} devises a selection criterion by choosing a threshold, which results in detection of only significantly interesting patterns.
-Section \ref{sec:simulations} provides some simulation study on the proposed methodology.
+Section \ref{sec:simulations} provides a simulation study on the proposed methodology.
 Section \ref{sec:application-wpd} presents an application to residential smart meter data in Melbourne to show how the proposed methodology can be used to automatically detect temporal granularities along which distributional differences are significant.
@@
 
-The principle behind the construction of $wpd$ is explained through a simple example explained in Figure \ref{fig:null4by2}.
-Each of these figures describes a panel with $2$ x-axis categories and $3$ facet levels, but with different designs.
-Figure \ref{fig:null4by2}a has all categories drawn from N(0, 1) distribution for each facet.
-It is not an interesting display particularly, as distributions do not vary across x-axis or facet categories.
-Figure \ref{fig:null4by2}b has x categories drawn from the same distribution, but across facets the distributions are $3$ standard deviations apart.
-Figure \ref{fig:null4by2}c exhibits an exact opposite situation where distribution between the x-axis categories
-are $3$ standard deviations apart, but they do not change across facets.
-Figure \ref{fig:null4by2}d takes a step further by varying the distribution across both facet and x-axis categories by  $3$ standard deviations.
-If the panels are to be ranked in order of capturing maximum variation in the measured variable from minimum to maximum, then an obvious choice would be placing (a) followed by (b), (c) and then (d).
+(ref:null4by2) An example illustrating the principle of the proposed distance measure, displaying the distribution of a normally distributed variable in four panels each with two x-axis categories and three facet levels, but with different designs.
+Panel (a) is not interesting as the distribution of the variable does not depend on x or facet categories.
+Panels (b) and (c) are more interesting than (a) since there is a change in distribution either across facets (b) or x-axis (c).
+Panel (d) is most interesting in terms of capturing structure in the variable as the distribution of the variable changes across both facet and x-axis variable.
+The value of our proposed distance measure is presented for each panel, the relative differences between which will be explained later in Section \@ref(sec:ranking).
+
+```{r null4by2, fig.cap = "(ref:null4by2)", fig.pos="!b"}
+```
+
+The principle behind the construction of $\wpd$ is explained through a simple example in Figure&nbsp;\ref{fig:null4by2}.
+Each of these figures describes a panel with two x-axis categories and three facet levels, but with different designs.
+Figure \ref{fig:null4by2}a has all categories drawn from a standard normal distribution for each facet.
+It is not a particularly interesting display, as the distributions do not vary across x-axis or facet categories.
+Figure \ref{fig:null4by2}b has x categories drawn from the same distribution, but across facets the distribution means are three standard deviations apart.
+Figure \ref{fig:null4by2}c exhibits the opposite situation where distribution between the x-axis categories are three standard deviations apart, but they do not change across facets.
+In Figure \ref{fig:null4by2}d, the distribution varies across both facet and x-axis categories by three standard deviations.
+
+If the panels are to be ranked in order of capturing maximum variation in the measured variable from minimum to maximum, then an obvious choice would be (a) followed by (b), (c) and then (d).
 It might be argued that it is not clear if (b) should precede or succeed (c) in the ranking.
 Gestalt theory suggests items placed at close proximity can be compared more easily, because people assume that they are in the same group and apart from other groups.
-With this principle in mind, display (b) is considered less informative as compared to display (c) in emphasizing the distributional differences.
-Considering one cyclic granularity, we would have only two design choices similar to (a) and (c), corresponding to no difference and significant differences between categories of that cyclic granularity only.
-The proposed measure $wpd$ is constructed in a way so that it could be used to rank panels of different designs as well as test if a design is interesting.
+With this principle in mind, Panel (b) is considered less informative compared to Panel (c) in emphasizing the distributional differences.
+
+For displays showing a single cyclic granularity rather than pairs of granularities, we have only two design choices corresponding to no difference and significant differences between categories of that cyclic granularity.
+
+The proposed measure $\wpd$ is constructed in such a way that it can be used to rank panels of different designs as well as test if a design is interesting.
 This measure is aimed to be an estimate of the maximum variation in the measured variable explained by the panel.
-A higher value of  $wpd$ would indicate that the panel is interesting to look at, whereas a lower value would indicate otherwise.
+A higher value of $\wpd$ would indicate that the panel is interesting to look at, whereas a lower value would indicate otherwise.
 
@@
 The continuous version of f -divergence is given by -->
-<!-- $$D_f(P||Q) := \int q(x)f(\frac{p(x)}{q(x)})$$, where -->
+<!-- $$\Df(P||Q) := \int q(x)f(\frac{p(x)}{q(x)})$$, where -->
 <!-- $f : [0,\infty) \rightarrow R \cup \{\infty\}$ is a continuous convex function, and $f(1) = 0$. -->
 
-One of the most common ways to measure divergence between distributions is the Kullback-Leibler (KL) divergence [@Kullback1951-jy].
+A common way to measure divergence between distributions is the Kullback-Leibler (KL) divergence [@Kullback1951-jy].
 The KL divergence denoted by $D(q_1||q_2)$ is a non-symmetric measure of the difference between two probability distributions $q_1$ and $q_2$ and is interpreted as the amount of information lost when $q_2$ is used to approximate $q_1$.
-The KL divergence, however, is not symmetric and hence can not be considered as a "distance" measure.
+The KL divergence is not symmetric and hence can not be considered as a "distance" measure.
 The Jensen-Shannon divergence [@Menendez1997-in] based on the Kullback-Leibler divergence is symmetric and has a finite value.
-Hence, in this paper, the pairwise distances between the distributions of the measured variable are obtained through the square root of the Jensen-Shannon divergence, called Jensen-Shannon distance (JSD) and is defined by,
-
-$$JSD(q_1||q_2) = \frac{1}{2}D(q_1||M) + \frac{1}{2}D(q_2||M)$$
-where $M = \frac{q_1+q_2}{2}$ and 
-$D(q_1||q_2) := \int^\infty_{-\infty} q_1(x)f(\frac{q_1(x)}{q_2(x)})$ is the KL divergence between distributions $q_1$ and $q_2$.
+Hence, in this paper, the pairwise distances between the distributions of the measured variable are obtained through the square root of the Jensen-Shannon divergence, called Jensen-Shannon distance (JSD), and defined by
+$$
+  JSD(q_1||q_2) = \frac{1}{2}D(q_1||M) + \frac{1}{2}D(q_2||M),
+$$
+where $M = \frac{q_1+q_2}{2}$ and $D(q_1||q_2) := \int^\infty_{-\infty} q_1(x)f(\frac{q_1(x)}{q_2(x)})$ is the KL divergence between distributions $q_1$ and $q_2$.
 Other common measures of distance between distributions are Hellinger distance, total variation distance and Fisher information metric.
 
-<!-- Furthermore, these distances are distributed as chi-squared with $m$ degrees of freedom (@Menendez1997-in), if the continuous distribution is being discretized with $m$ discrete values.
+<!-- Furthermore, these distances are distributed as chi-squared with $m$ degrees of freedom [@Menendez1997-in], if the continuous distribution is being discretized with $m$ discrete values.
 Taking sample percentiles to approximate the integral would mean taking $m = 99$.
@@
 
-#### Within-facet and between-facet distances
-
-\noindent Pairwise distances could be within-facets or between-facets for $m \geq 2$.
+#### Within-facet and between-facet distances {-}
+
+```{r distance-explain, fig.cap = "Within and between-facet distances shown for two cyclic granularities A and B, where A is mapped to x-axis and B is mapped to facets.
+The dotted lines represent the distances between different categories.
+Panels 1) and 2) show the between-facet distances.
+Within-facet distances are illustrated in Panels 3) (when categories are un-ordered, shown only with respect to a1) and Panel 4) (when categories are ordered).
+When categories are ordered, distances should only be considered for consecutive x-axis categories.
+Between-facet distances are distances between different facet levels for the same x-axis category; for example, distances between {($a_1$,$b_1$) and ($a_1$, $b_2$)} or {($a_1$,$b_1$) and ($a_1$, $b_3$)}.", fig.pos="!b"}
+
+```
+
+Pairwise distances could be within-facets or between-facets for $m = 2$.
 Figure \ref{fig:distance-explain} illustrates how they are defined.
@@
 Pairwise distances are between-facets when they are considered between pairs of the form $(a_{j}b_{k}, a_{j}b_{k'})$.
-Number of between-facet distances would be $^{K}C_2*J$ and number of within-facet distances are $K*(J-1)$ (ordered) and $^{J}C_2*K$ (un-ordered).
-
-#### Tuning parameter
-
-A tuning parameter specifying the weightage given to the within-facet or between-facet categories can help to balance weightage between designs like \ref{fig:null4by2}(b) and (c).
-The tuning parameters should be chosen such that $\sum_{i=1}^m \lambda_i = 1$.
-When $m=2$, following the principle of Gestalt theory, $\lambda = \frac{2}{3} = 0.67$ is chosen to put a relative weightage of $2:1$ for within-facet and between-facet distances.
-No human experiment is conducted to justify this ratio, however, typically a tuning parameter $\lambda>0.5$ would tend to upweigh the within-facet distances and that with $<0.5$ would upweigh the between-facet distances (refer to the Supplementary paper for more details).
-For $m = 1$, there are no conditioning variables or groups, and hence $\lambda = 1$.
-
-#### Raw distance measure
-
-The raw distance measure, denoted by $wpd_{raw}$, is computed after combining all the weighted distance measures appropriately.
+There are a total of ${\nf \choose 2}\nx$ between-facet distances, and ${\nx \choose 2}\nf$ within-facet distances if they are unordered and $\nf(\nx-1)$ within-facet distances if they are ordered.
+
+#### Tuning parameter {-}
+
+For displays with $m>1$ granularities, we can use a tuning parameter to specify the relative weight given to each granularity.
+In general, the tuning parameters should be chosen such that $\sum_{i=1}^m \lambda_i = 1$.
+
+Following the general principles of Gestalt theory, we wish to weight more heavily granularities that are plotted closer together.
+For $m=2$ we choose  $\lambda_x = \frac{2}{3}$ for the granularity on the x-axis and $\lambda_f = \frac13$ for the granularity mapped to facets, giving a relative weight of $2:1$ for within-facet to between-facet distances.
+No human experiment has been conducted to justify this ratio.
+Specifying $\lambda_x>0.5$ will weight within-facet distances more heavily, while $\lambda_x <0.5$ would weight the between-facet distances more heavily.
+(See the supplements for more details.)
+
+#### Raw distance measure {-}
+
+The raw distance measure, denoted by $\wpdsub{raw}$, is computed after combining all the weighted distance measures appropriately.
 First, NQT is performed on the measured variable $v_t$ to obtain $v^*_t$ (_data transformation_).
 Then, for a fixed harmony pair $(A, B)$, percentiles of $v^*_{jk}$ are computed and stored in $q_{jk}$ (_distribution characterization_).
-This is repeated for all pairs of categories of the form $(a_{j} b_{k}, a_{j'}b_{k'}): \{a_j: j = 1, 2, \dots, J\}, B = \{ b_k: k = 1, 2, \dots, K\}$.
-The pairwise distances between pairs $(a_{j} b_{k}, a_{j'}b_{k'})$  denoted by $d_{(jk, j'k')} = JSD(q_{jk}, q_{j'k'})$ is computed (_distance between distributions_).
-The pairwise distances (_Within-facet and between-facet_) are transformed using a suitable _tuning parameter_ ($0<\lambda<1$) depending on if they are within-facet($d_w$) or between-facets($d_b$) as follows:
+This is repeated for all pairs of categories of the form $(a_{j} b_{k}, a_{j'}b_{k'}): \{a_j: j = 1, 2, \dots, \nx\}, B = \{ b_k: k = 1, 2, \dots, \nf\}$.
+The pairwise distances between pairs $(a_{j} b_{k}, a_{j'}b_{k'})$ denoted by $d_{(jk, j'k')} = JSD(q_{jk}, q_{j'k'})$ are computed (_distance between distributions_).
+The pairwise distances (_within-facet and between-facet_) are transformed using a suitable _tuning parameter_ ($0<\lambda<1$) depending on if they are within-facet($d_w$) or between-facets($d_b$) as follows:
 \begin{equation}\label{def:raw-wpd}
@@
     \begin{cases}
-\lambda d_{(jk), (j'k')},& \text{if } d = d_w\\
-(1-\lambda)  d_{(jk), (j'k')},              & \text{if } d = d_b\\
+      \lambda d_{(jk), (j'k')},& \text{if } d = d_w;\\
+      (1-\lambda) d_{(jk), (j'k')},       & \text{if } d = d_b.
     \end{cases}
 \end{equation}
-
-\noindent The $wpd_{raw}$ is then computed as $$wpd = max_{j, j', k, k'}(d*_{(jk), (j'k')}) \forall j, j' \in \{1, 2, \dots, J\}, k, k' \in \{1, 2, \dots, K\}$$
-
-\noindent The statistic "maximum" is chosen to combine the weighted pairwise distances since the distance measure is aimed at capturing the maximum variation of the measured variable within a panel.
+The $\wpdsub{raw}$ is then computed as
+$$
+  \wpdsub{raw} = \max_{j, j', k, k'}(d*_{(jk), (j'k')}) \qquad\forall j, j' \in \{1, 2, \dots, \nx\},\quad k, k' \in \{1, 2, \dots, \nf\}
+$$
+The statistic "maximum" is chosen to combine the weighted pairwise distances since the distance measure is aimed at capturing the maximum variation of the measured variable within a panel.
 The statistic "maximum" is, however, affected by the number of comparisons (resulting pairwise distances).
-For example, for a $(2, 3)$ panel, there are $6$ possible subsets of obervations corresponding to the combinations $(a_1, b_1), (a_1, b_2), (a_1, b_3), (a_2 ,b_1), (a_2 ,b_2), (a_2, b_3)$, whereas for a $(2, 2)$ panel, there are only $4$ possible subsets $(a_1, b_1), (a_1, b_2), (a_2 ,b_1), (a_2 ,b_2)$.
+For example, for a $(2, 3)$ panel, there are $6$ possible subsets of observations corresponding to the combinations $(a_1, b_1), (a_1, b_2), (a_1, b_3), (a_2 ,b_1), (a_2 ,b_2), (a_2, b_3)$, whereas for a $(2, 2)$ panel, there are only $4$ possible subsets $(a_1, b_1), (a_1, b_2), (a_2 ,b_1), (a_2 ,b_2)$.
 Consequently, the measure would have higher values for the panel $(2, 3)$ as compared to $(2, 2)$, since maximum is taken over higher number of pairwise distances.
 
-#### Adjusting for the number of comparisons {#sec:norm-wpd}
-
-Ideally, it is desired that the proposed distance measure takes a higher value only if there is a significant difference between distributions across categories, and not because the number of categories $J$ or $K$ is high.
-That is, under designs like $D_{null}$, their distribution should not differ for a different number of categories.
-Only then the distance measure could be compared across panels with different levels.
+### Adjusting for the number of comparisons
+
+Ideally, it is desired that the proposed distance measure takes a higher value only if there is a significant difference between distributions across categories, and not because the number of categories $\nx$ or $\nf$ is high.
+That is, under designs like $\Dnull$, their distribution should not differ for a different number of categories.
+Only then could the distance measure be compared across panels with different levels.
 This calls for an adjusted measure, which normalizes for the different number of comparisons.
-We denote it by $wpd$.
+
 Two approaches for adjusting the number of comparisons are discussed, both of which are substantiated using simulations.
-The first one defines an adjusted measure $wpd_{perm}$ based on the permutation method to remove the effect of different comparisons.
-The second approach fits a model to represent the relationship between $wpd_{raw}$ and the number of comparisons and defines the adjusted measure ($wpd_{glm}$) as the residual from the model.
-
-_Permutation approach_
+The first one defines an adjusted measure $\wpdsub{perm}$ based on the permutation method to remove the effect of different comparisons.
+The second approach fits a model to represent the relationship between $\wpdsub{raw}$ and the number of comparisons and defines the adjusted measure ($\wpdsub{glm}$) as the residual from the model.
+
+#### Permutation approach {-}
 
@@
 This method accomplishes a different goal of finding the null distribution for different groups (panels in our case) and standardizing the raw values using that distribution.
-The values of $wpd_{raw}$ is computed on many ($nperm$) permuted data sets and stored in $wpd_{perm-data}$.
-Then $wpd_{perm}$ is computed as follows:
-
+The values of $\wpdsub{raw}$ are computed on many ($\nsub{perm}$) permuted data sets and stored in $\wpdsub{perm-data}$.
+Then $\wpdsub{perm}$ is computed as follows:
 \begin{align*}
-wpd_{perm} = \frac{(wpd_{raw} - \bar{wpd_{perm-data}})}{sd(wpd_{perm-data})}
+  \wpdsub{perm} = \frac{\wpdsub{raw} - \text{mean}(\wpdsub{perm-data})}{\text{sd}(\wpdsub{perm-data})}
 \end{align*}
-
-\noindent where $\bar{wpd_{perm-data}}$ and $sd(wpd_{perm-data})$ are the mean and standard deviation of $wpd_{perm-data}$ respectively.
-Standardizing  $wpd$ in the permutation approach ensures that the distribution of $wpd_{perm}$ under $D_{null}$ has the same $mean = 0$ and $\sigma^2_{perm} = 1$ across all comparisons.
-While this works successfully to make the location and scale similar across different $nx$ and $nfacet$, it is computationally heavy and time consuming, and hence less user friendly when being actually used in practice.
+where $\text{mean}(\wpdsub{perm-data})$ and $\text{sd}(\wpdsub{perm-data})$ are the mean and standard deviation of $\wpdsub{perm-data}$ respectively.
+Standardizing $\wpd$ in the permutation approach ensures that the distribution of $\wpdsub{perm}$ under $\Dnull$ has zero mean and unit variance across all comparisons.
+While this works successfully to make the location and scale similar across different $\nx$ and $\nf$, it is computationally heavy and time consuming, and hence less user-friendly.
 Hence, another approach to adjustment, with potentially less computational time, is proposed.
@@
 
-_Modeling approach_
-
-In this approach, a Gamma generalized linear model (GLM) for $wpd_{raw}$ is fitted with number of comparisons as the explanatory variable.
-Since, $wpd_{raw}$ is a Jensen-Shannon distance, it follows a Chi-square distribution (@Menendez1997-in), which is a special case of Gamma distribution.
-Furthermore, the mean response is bounded, since any JSD is bounded by $1$ given that base $2$ logarithm is used (@Lin1991-pt).
-Hence, by @Faraway2016-uk, an inverse link is used for the model, which is of the  form $y = a+b*log(z) + e$, where $y = wpd_{raw}$, $z = (nx*nfacet)$ is the number of groups and $e$ are idiosyncratic errors.
-Let $E(y) = \mu$ and $a + b*log(z) = g(\mu)$ where $g$ is the link function.
-Then $g(\mu)=  1/\mu$ and $\hat \mu  = 1/(\hat a + \hat b log(z))$.
-The residuals from this model $(y-\hat y) = (y-1/(\hat a + \hat b log(z)))$ would be expected to have no dependency on $z$.
-Thus, $wpd_{glm}$ is chosen as the residuals from this model and is defined as:
-$$wpd_{glm} = wpd_{raw} - 1/(\hat a + \hat b*log(nx*nfacet))$$
-The distribution of $wpd_{glm}$ under $D_{null}$ will have $mean = 0$, since it is the residuals from the model, and a constant variance $\sigma^2_{glm}$, which might not equal 1.
-
-_Combination approach_
-
-The simulation results (in Section \ref{sec:simulations}) show that the distribution of $wpd_{glm}$ under null is similar for high $nx$ and $nfacet$ (levels higher than $5$) and not so much for lower $nx$ and $nfacet$.
-Hence, a combination approach is proposed which chooses permutation approach for categories with smaller levels and modeling approach for categories with higher levels.
-This ensures that the computational load of the permutation approach is alleviated while maintaining similar null distribution across different categories.
+#### Modeling approach {-}
+
+In this approach, a Gamma generalized linear model (GLM) for $\wpdsub{raw}$ is fitted with the number of comparisons as the explanatory variable.
+Since, $\wpdsub{raw}$ is a Jensen-Shannon distance, it follows a Chi-square distribution [@Menendez1997-in], which is a special case of a Gamma distribution.
+Furthermore, the mean response is bounded, since any JSD is bounded by $1$ if a base $2$ logarithm is used [@Lin1991-pt].
+Hence, by @Faraway2016-uk, an inverse link is used for the model, which is of the form $y = a+b\times\log(z) + e$, where $y = \wpdsub{raw}$, $z = (\nx \times \nf)$ is the number of groups and $e$ are idiosyncratic errors.
+Let $\text{E}(y) = \mu$ and $a + b\times\log(z) = g(\mu)$ where $g(\mu)= 1/\mu$ and $\hat \mu = 1/(\hat a + \hat b \log(z))$.
+The residuals from this model $(y-\hat y) = (y-1/(\hat a + \hat b \log(z)))$ would be expected to have no dependency on $z$.
+Thus, $\wpdsub{glm}$ is defined as the residuals from this model given by
+$$\wpdsub{glm} = \wpdsub{raw} - 1/(\hat a + \hat b\times\log(\nx\times\nf))$$
+The distribution of $\wpdsub{glm}$ under $\Dnull$ will have approximately zero mean and a constant variance (not necessarily 1).
+
+#### Combination approach {-}
+
+The simulation results (in Section \ref{sec:simulations}) show that the distribution of $\wpdsub{glm}$ under the null design is similar for high $\nx$ and $\nf$ (levels higher than $5$) but less so for lower values of $\nx$ and $\nf$.
+Hence, a combination approach is proposed where we use a permutation approach for categories with small numbers of levels, and a modeling approach for categories with higher numbers of levels.
+This ensures that the computational load of the permutation approach is alleviated while maintaining a similar null distribution across different categories.
 This approach, however, requires that the adjusted variables from the two approaches are brought to the same scale.
-We define $wpd_{glm-scaled} = wpd_{glm}*\sigma^2_{perm}/\sigma^2_{glm}$ as the transformed $wpd_{glm}$ with a similar scale as $wpd_{perm}$.
-The adjusted measure from the combination approach, denoted by $wpd$ is then defined as follows:
-
+We define $\wpdsub{glm-scaled} = \wpdsub{glm}\times \sigma^2_{\text{perm}}/\sigma^2_{\text{glm}}$ as the transformed $\wpdsub{glm}$ with a similar scale as $\wpdsub{perm}$.
+The adjusted measure from the combination approach, denoted by $\wpd$ is then defined as follows:
 \begin{equation}
-wpd =
-\begin{cases}
-wpd_{perm},& \text{if } J, K <=5\\
-wpd_{glm-scaled}             & \text{otherwise}\\
+\wpd = \begin{cases}
+         \wpdsub{perm},       & \text{if $\nx, \nf \le 5$};\\
+         \wpdsub{glm-scaled}  & \text{otherwise}.\\
        \end{cases}
@@
 
-A threshold and consequently a selection criterion is chosen using the notion of Randomization tests (@edgington2007randomization).
-The data is permuted several times and $wpd$ is computed for each of the permuted data sets to obtain the sampling distribution of $wpd$ under the null hypothesis.
-If the null hypothesis is true, then $wpd$ obtained from the original data set would be a likely value in the sampling distribution.
-But in case the null hypothesis is not true, then it is less probable that $wpd$ obtained for the original data will be from the same distribution.
-This idea is utilized to come up with a threshold for selection, denoted by $wpd_{threshold}$, defined as the $99^{th}$ percentile of the sampling distribution.
-A harmony is selected if the value of $wpd$ for that harmony is greater than the chosen threshold.
+A threshold (and consequently a selection criterion) is chosen using the notion of randomization tests [@edgington2007randomization].
+The data is permuted several times and $\wpd$ is computed for each of the permuted data sets to obtain the sampling distribution of $\wpd$ under the null hypothesis.
+If the null hypothesis is true, then $\wpd$ obtained from the original data set would be a likely value in the sampling distribution.
+But in case the null hypothesis is not true, then it is less probable that $\wpd$ obtained for the original data will be from the same distribution.
+This idea is utilized to come up with a threshold for selection, denoted by $\wpdsub{threshold}$, defined as the $99^{th}$ percentile of the sampling distribution.
+A harmony is selected if the value of $\wpd$ for that harmony is greater than the chosen threshold.
 The detailed algorithm for choosing a threshold and selection procedure (for two cyclic granularities) is listed as follows:
 
-<!-- The selection criterion is stated as follows: for $(i \in {1, 2, \dots, H_{N_C}})$, select $H_i$ if $wpd_i > wpd_{threshold}$, else reject. -->
-
-- **Input:** All harmonies of the form $\{(A, B), A = \{ a_j: j = 1, 2, \dots, J\}, B = \{ b_k: k = 1, 2, \dots, K\}\}$, $\forall (A, B) \in H_{N_C}$.
-
-- **Output:** Harmony pairs $(A, B)$ for which $wpd$ is significant.
-
-1. Fix harmony pair $(A, B)$.
-
-2. Given the measured variable; $\{v_t: t=0, 1, 2, \dots, T-1\}$, $wpd$ is computed and is represented by $wpd^{A, B}_{obs}$.
-
-3. From the original sequence a random permutation is obtained: $\{v_t^{1}: t=0, 1, 2, \dots, T-1\}$.
-
-4. $wpd$ is computed for the permuted sequence of the data and is represented by $wpd^{A, B}_{1}$.
-
-5.  Steps (3) and (4) are repeated a large number of times $M$ (e.g. $M = 200$).
-
-6. For each permutation, one $wpd^{A, B}_{i}$ is obtained. Define $wpd_{sample} = \{wpd^{A, B}_{1} wpd^{A, B}_{2}, \dots, wpd^{A, B}_{M}\}$.
-
-7. Repeat Steps (1-6) for all harmony pairs $(A, B) \in H_{N_C}$ and stored $wpd^{all}_{sample}$.
-
-8. $99^{th}$ percentiles of $wpd^{all}_{sample}$ is computed and stored in $wpd_{threshold99}$
-
-7. If $wpd^{A, B}_{obs} > wpd_{threshold99}$, harmony pair $(A, B)$ is selected, otherwise rejected.
-
-Similarly, a harmony pair $(A, B)$ could be selected if $wpd^{A, B}_{obs} > wpd_{threshold95}$ and $wpd^{A, B}_{obs} > wpd_{threshold90}$, where  $wpd_{threshold95}$ and $wpd_{threshold90}$ denote the $95^{th}$ and $90^{th}$ percentile of $wpd^{all}_{sample}$ respectively.
-A harmony selected using $99^{th}$,  $95^{th}$  and  $90^{th}$ threshold are tagged as \*\*\*, \*\*, \* respectively.
-
-### Ranking
-
-The distribution of $wpd$ is expected to be similar for all harmonies under the null hypothesis, since they have been adjusted for different number of categories for the harmonies or underlying distribution of the measured variable.
-Hence, the values of $wpd$ for different harmonies are comparable and can be used to rank the significant harmonies.
-A higher value of $wpd$ for a harmony indicates that higher maximum variation in the measured variable is captured through that harmony.
-Figure \ref{fig:null4by2} presents the results of $wpd$ from the illustrative designs in Section \ref{sec:computation-wpd}.
-The value of $wpd$ under null design (a) is the least, followed by (b), (c) and (d).
-This aligns with the principle of $wpd$, which is expected to have lowest value for null designs and highest for designs of the form $D_{var_{all}}$ (d).
-Moreover, note the relative differences in $wpd$  values between (b) and (c).
-The value of the tuning parameter $\lambda$ is set to $0.67$, which has resulted in giving more emphasis to differences in x-axis categories.
-Again consider \ref{fig:onegran-new}(a) and \ref{fig:onegran-new}(b) with a $wpd$ value of 20.5 and 145 respectively.
-This is because there is more gradual increase across hours of the day than months of the year.
-If order of the categories is not considered, they result in a $wpd$ value of $97.8$ and $161$ respectively, which follows from the fact that if we consider difference between any hours of the day, the magnitude will be much larger than if we consider difference between consecutive categories.
-Similarly, \ref{fig:onegran-new}(a) and (b) have $wpd$ values of   $110.79$ and $125.82$ respectively.
-The ranking implies that the distributional differences are more prominent for the second household, as is also seen from the bigger fluctuations in the $90^{th}$ percentile than the first household.
+<!-- The selection criterion is stated as follows: for $(i \in {1, 2, \dots, H_{N_C}})$, select $H_i$ if $wpd_i > \wpdsub{threshold}$, else reject. -->
+
+- **Input:** All harmonies of the form $\{(A, B),~ A = \{ a_j: j = 1, 2, \dots, \nx\},~ B = \{ b_k: k = 1, 2, \dots, \nf\}\}$, $\forall (A, B) \in H_{N_C}$.
+
+- **Output:** Harmony pairs $(A, B)$ for which $\wpd$ is significant.
+
+1. For each harmony pair $(A, B) \in H_{N_C}$, the following steps are taken.
+
+    a. Given the measured variable; $\{v_t: t=0, 1, 2, \dots, T-1\}$, $\wpd$ is computed and is represented by $\wpd^{A, B}_{obs}$.
+    b. For $i=1,\dots,M$, randomly permute the original time series: $\{v_t^{i}: t=0, 1, 2, \dots, T-1\}$ and compute $\wpd^{A, B}_{i}$ from $\{v_t^{i}\}$.
+    c. Define $\wpdsub{sample} = \{\wpd^{A, B}_{1}, \dots, \wpd^{A, B}_{M}\}$.
+
+2. Stack the $\wpdsub{sample}$ vectors as $\wpdsub{sample}^{\text{all}}$ and compute its $p = 100(1-\alpha)$ percentiles as $\wpdsub{threshold\emph{p}}$.
+3. If $\wpd^{A, B}_{obs} > \wpdsub{threshold\emph{p}}$, harmony pair $(A, B)$ is selected at the $1-p/100$ level, otherwise rejected.
+4. Harmonies selected using the $99^{th}$, $95^{th}$ and $90^{th}$ thresholds are tagged as \*\*\*, \*\*, \* respectively.
+
+### Ranking {#sec:ranking}
+
+The distribution of $\wpd$ is expected to be similar for all harmonies under the null hypothesis, since they have been adjusted for different number of categories for the harmonies or underlying distribution of the measured variable.
+Hence, the values of $\wpd$ for different harmonies are comparable and can be used to rank the significant harmonies.
+A higher value of $\wpd$ for a harmony indicates that higher maximum variation in the measured variable is captured through that harmony.
+
+```{r varall-new}
+
+```
+
+```{r varx-new}
+
+```
+
+```{r varf-new}
+```
+
+```{r null-new}
+
+```
+
+Figure \ref{fig:null4by2} also presents the results of $\wpd$ from the illustrative designs in Section \ref{sec:computation-wpd}.
+The value of $\wpd$ under null design (a) is the least, followed by (b), (c) and (d).
+This aligns with the principle of $\wpd$, which is expected to have lowest value for null designs and highest for designs of the form $\Dfx$ (d).
+Moreover, note the relative differences in $\wpd$ values between (b) and (c).
+The value of the tuning parameter $\lambda$ is set to 2/3, which gives greater emphasis to differences in x-axis categories than facets.
+
+Again consider Figures \ref{fig:onegran-new}(a) and \ref{fig:onegran-new}(b) with a $\wpd$ value of 20.5 and 145 respectively.
+This is because there is a more gradual increase across hours of the day than across months of the year.
+If the order of categories is ignored, the resulting $\wpd$ values are $97.8$ and $161$ respectively, because differences between any hours of the day tend to be larger than differences only between consecutive hours.
+Similarly, Figures \ref{fig:id2-new2}(a) and (b) have $\wpd$ values of $110.79$ and $125.82$ respectively.
+The ranking implies that the distributional differences are more prominent for the second household, as is also seen from the bigger fluctuations in the $90^{th}$ percentile than for the first household.
 
@@
 
-```{r dist-new-same-scale-link, fig.cap = "The distribution of $wpd_{perm}$ and $wpd_{glm-scaled}$ are overlaid to compare the location and scale across different $nx$ and $nfacet$.
-$wpd_{norm}$ takes the value of $wpd_{perm}$ for lower levels, and $wpd_{glm-scaled}$ for higher levels to to alleviate the problem of computational time in permutation approaches.
-This is possible as the distribution of the adjusted measure looks similar for both appraches for higher levels."}
-```
-
-### Choosing threshold
-
-\noindent _Simulation design_
-
-Observations are generated from a N(0,1) distribution for each combination of $nx$ and $nfacet$ from the following sets: $nx = \{3, 7, 14\}$ and $nfacet = \{2, 9, 10\}$.
-This would result in 9 panels, viz, $(3, 2), (3, 9), (3, 10), \dots, (14,9), (14, 10)$.
-Few experiments were conducted.
-In the first scenario, data for all panels are simulated using the null design  $D_{null}$.
-In other scenarios, data simulated from the panel $(14, 2)$ and $(3, 10)$ are considered under $D_{vary_{all}}$.
-Moreover, $\omega = \{0.5, 2, 5\}$ are considered to examine if the proposed test is able to capture subtle differences and non-subtle differences when we shift from the null design.
-
-<!-- In the last scenario, we consider the panel $(3, 2), (7,9), (14, 10)$ to be under $D_{null}$, the panels $(7, 2), (14, 9)$ to be under $D_{var_f}$.
-$(14, 2), (3, 10)$ under $D_{var_x}$ and the rest under $D_{var_{null}}$.
+(ref:dist-new-same-scale-link) The distributions of $\wpdsub{perm}$ and $\wpdsub{glm-scaled}$ are overlaid to compare the location and scale across different $\nx$ and $\nf$.
+$\wpdsub{norm}$ takes the value of $\wpdsub{perm}$ for lower levels, and $\wpdsub{glm-scaled}$ for higher levels to to alleviate the problem of computational time in permutation approaches.
+This is possible as the distribution of the adjusted measure looks similar for both approaches for higher levels.
+
+```{r dist-new-same-scale-link, fig.cap = "(ref:dist-new-same-scale-link)"}
+```
+
+These results justify our use of the permutation approach when $\nx<5$ and $\nf<5$, and the use of the GLM otherwise.
+<!--
+## Choosing the threshold
+
+### Simulation design {-}
+
+Observations were generated from a N(0,1) distribution for each combination of $\nx$ and $\nf$ from the following sets: $nx \in \{3, 7, 14\}$ and $\nf \in \{2, 9, 10\}$, giving nine possible panels/combinations.
+In the first scenario, data for all panels were simulated using the null design $\Dnull$.
+In other scenarios, data simulated from the panel $(14, 2)$ and $(3, 10)$ used $\Dfx$ where the means of the distributions increase by $\omega$ across both x-axis and facets, where $\omega \in \{0.5, 2, 5\}$.
+This allows us to examine if the proposed test is able to capture subtle differences and obvious differences when we shift from the null design.
+
+In the last scenario, we consider the panel $(3, 2), (7,9), (14, 10)$ to be under $\Dnull$, the panels $(7, 2), (14, 9)$ to be under $\Df$.
+$(14, 2), (3, 10)$ under $D_x$ and the rest under $\Dvar{{null}}$.
 This is done to check if the consequent ranking procedure leads to designs like $D_{vary_{all}}$ to be chosen first followed by $D_{vary_{all}}$.
@@
 
-The smart meter data set for eight households in Melbourne has been utilized to see the use of $wpd$ proposed in the paper.
-The data has been cleaned to be a `tsibble` (@wang2020tsibble) containing half-hourly electricity consumption from Jul-2019 to Dec-2019 for each of the households, which is procured by them by downloading their data from the energy supplier/retailer.
-No behavioral pattern is likely to be discerned from the line graph of energy usage over the entire period, since the plot will have too many measurements squeezed in a linear representation.
-When we zoom into the linear representation of this series in Figure \ref{fig:dotplot-8} (b) for September, some patterns are visible in terms of peaks and troughs, but we do not know if they are regular or what is their period.
-Electricity demand, in general, has a daily and weekly periodic pattern.
-However, it is not apparent from this view if all of these households have those patterns and in case they have if they are significant enough.
-Also, it is not clear if any other periodic patterns are present in any household which might have been hidden with this view.
-We start the analysis by choosing few harmonies, ranking them for each of these households, compare households to get more insights into what these rankings imply.
-Furthermore, the ranking and selection of significant harmonies is validated by analyzing the distribution of energy usage across significant harmonies.
+The smart meter data set for eight households in Melbourne was procured by downloading the data from the energy supplier/retailer.
+The data has been cleaned to form a `tsibble` [@wang2020tsibble] containing half-hourly electricity consumption from July to December 2019 for each of the households.
+No behavioral pattern is likely to be discerned from the time plot of energy usage over the entire period, since the plot will have too many observations squeezed in a linear representation.
+When we zoom into the September 2019 data in Figure \ref{fig:dotplot-8}(b), some patterns are visible in terms of peaks and troughs, but we do not know if they are regular or what is their period.
+
+(ref:dotplot-8) An ensemble plot with a heatmap (a), line plot (b), parallel coordinate plot (c) to demonstrate energy behavior of the households in different ways.
+Panel (b) shows the raw demand series for September to highlight the repetitive patterns of energy demand.
+Panel (a) shows $\wpd$ values across harmonies where a darker color indicates a higher ranking harmony.
+A significant harmony is shown with an asterisk.
+For example, ids 7 and 8 have significant patterns across (*hod*, *dow*) and (*dow*, *hod*).
+Panel (c) is useful for comparing households across harmonies.
+For example, for the harmony (*dow*-*hod*), ids 1 and 7 have the least and highest $\wpd$ respectively.
+
+```{r dotplot-8, fig.cap = "(ref:dotplot-8)", fig.height = 7}
+```
+
+Electricity demand, in general, has daily, weekly and annual seasonal patterns.
+However, it is not apparent from this view if all households have those patterns, or how strong they are in each case.
+It is also not clear from this view if any other periodic patterns are present in any household.
+We start the analysis by choosing a few harmonies, and ranking them for each of these households.
+The ranking and selection of significant harmonies is validated by analyzing the distribution of energy usage across significant harmonies.
 
@@
 
-_Choosing cyclic granularities of interest and removing clashes_
-
-Let $v_{i, t}$ denote the electricity demand for $i^{th}$ household for time period $t$.
-The series $v_{i, t}$ is the linear granularity corresponding to half-hour since the interval of the tsibble is 30 minutes.
+#### Choosing cyclic granularities of interest and removing clashes {-}
+
+Let $v_{i, t}$ denote the electricity demand for the $i^{th}$ household in time period $t$.
+The series $\{v_{i, 1},\dots,v_{i,T}\}$ is the linear granularity corresponding to half-hour since the interval of the `tsibble` is 30 minutes.
 We consider coarser linear granularities like hour, day, week and month from the commonly used Gregorian calendar.
-Considering $4$ linear granularities hour, day, week, month, the number of cyclic granularities is $N_C = (4 * 3/2) = 6$.
-We obtain cyclic granularities namely "hour_day", "hour_week", "hour_month", "day_ week", "day_month" and "week_month", read as "hour of the day", etc.
+From the four linear granularities of hour, day, week, and month, we obtain $N_C = 4\times 3/2 = 6$ cyclic granularities:  "hour_day", "hour_week", "hour_month", "day_ week", "day_month" and "week_month".
 Further, we add cyclic granularity day-type ("wknd wday") to capture weekend and weekday behavior.
-Thus, $7$ cyclic granularities are considered to be of interest.
-The set consisting of pairs of cyclic granularities ($C_{N_C}$) will have $7_{P_2}=42$ elements which could be analyzed for detecting possible periodicities.
-The set of possible harmonies $H_{N_C}$ from $C_{N_C}$ are chosen by removing clashes using procedures described in [@Gupta2021-hd].
+Thus, seven cyclic granularities are considered to be of interest.
+The pairs of cyclic granularities ($C_{N_C}$) will have $7\times 6=42$ elements.
+The set of possible harmonies $H_{N_C}$ from $C_{N_C}$ are chosen by removing clashes using procedures described in @Gupta2021-gravitas.
 Table \ref{tab:tab-rank-8} shows $14$ harmony pairs that belong to $H_{N_C}$.
+
+```{r rank-household-mclapply-8}
+```
+
+```{r elec_select_harmony-8}
+```
+
+```{r tab-rank-8, dependson="dotplot-8"}
+```
 
@@
 
-_Selecting and Ranking harmonies for all households_
-
-$wpd_{i}$ is computed on $v_{i, t}$ for all harmony pairs $\in H_{N_C}$ and for each households $i \in \{1, 2, \dots, 8\}$.
+#### Selecting and ranking harmonies for all households {-}
+
+$\wpd_{i}$ is computed on $v_{i, t}$ for all harmony pairs $\in H_{N_C}$ and for each household $i \in \{1, 2, \dots, 8\}$.
 The harmony pairs are then arranged in descending order and highlighted with `***`, `**` and `*` corresponding to the $99^{th}$, $95^{th}$ and $90^{th}$ percentile threshold.
@@
 The rankings are different for different households, which is a reflection of their varied behaviors.
-Most importantly, there are at most $3$ harmonies that are significant for any household.
-This is a huge reduction in the number of potential harmonies to explore closely, starting from $42$.
-
-_Detecting patterns not apparent from linear display_
-
-\noindent Figure \ref{fig:dotplot-8} helps to compare households through the heatmap (a) across harmony pairs with the cyclic granularity mapped to x-axis and facet being plotted on the x-axis and y-axis of the heatmap.
-Here $dom$, $dow$, $wdwnd$ are abbreviations for day-of-month, day-of-week, weekday/weekend and so on.
-The colors represent the value of $wpd$.
-Darker cells correspond to more significant harmony pairs.
-Also, the ones with `*` corresponds to the ones above $wpd_{threshold95}$.
-Few observations that emphasizes patterns not discernible through (b) includes - (1) id 7 and 8 have the same significant harmonies despite having very different total energy usage, (2) id 6 and 7 differ in the sense that for id 6, the difference in patterns only during weekday/weekends, whereas for id 7, all or few other days of the week are also important.
-This might be due to their flexible work routines or different day-off, (3) there are no significant periodic patterns for id 5 when we fix the threshold to $wpd_{threshold95}$.
-Note that the $wpd$ values are computed over the entire range, but the linear display in (b) is zoomed only for September, with the major and minor x-axis corresponding to weeks and days respectively.
+Most importantly, there are at most three harmonies that are significant for any household.
+This is a huge reduction in the number of potential harmonies to explore.
+
+#### Detecting patterns not apparent from linear display {-}
+
+Figure \ref{fig:dotplot-8} helps to compare households through the heatmap (a) across harmony pairs.
+Here *dom*, *dow*, *wdwnd* are abbreviations for day-of-month, day-of-week, weekday/weekend and so on.
+The colors represent the value of $\wpd$.
+Darker cells correspond to higher values of $\wpd$.
+Those with `*` correspond to $\wpd$ values above $\wpdsub{threshold95}$.
+
+We can now see some patterns that were not discernible in Panel (b), including:
+
+  1. id 7 and 8 have the same significant harmonies despite having very different total energy usage.
+  2.
+id 6 and 7 differ in the sense that for id 6, the difference in patterns is only during weekday/weekends, whereas for id 7 all or few other days of the week are also important.
+This might be due to their flexible work routines or different day-off.
+  3. There are no significant periodic patterns for id 5 when we fix the threshold to $\wpdsub{threshold95}$.
+
+Note that the $\wpd$ values are computed over the entire range, but the linear display in (b) is only for September, with the major and minor x-axis corresponding to weeks and days respectively.
 
@@
 
-```{r rank-household-mclapply-8}
-```
-
-```{r elec_select_harmony-8}
-```
-
-```{r dotplot-8, fig.cap = "An ensemble plot with a heatmap (a), line plot (b), parallel coordinate plot (c) to demonstrate energy behavior of the households in different ways.
-(b) is the zoomed-in raw demand series for September to highlight the repetitive patterns of energy demand.
-(a) shows $wpd$ values across harmonies with the x variable of harmony placed across x-axis and facet variable placed across y-axis.
-The darker the colour in (a), the higher the harmony is ranked.
-Visualizing harmonies through (a) allows to view the significant cyclic granularities along the x-axis, facet or both for each household.
-For eg, ids 7 and 8 have significant patterns across (hod, dow) and (dow, hod), which was not apparent through (a).
-(c) is useful for comparing households across harmonies, for eg, for the harmony (dow-hod), ids 1 and 7 have the least and highest wpd respectively.", fig.height = 7}
-```
-
-```{r tab-rank-8}
-```
-
-_Comparing households and validating rank of harmonies_
-
-According to Figure \ref{fig:dotplot-8}(c), for the harmony pair (dow-hod), household id 7 has the greatest value of $wpd$, while id 1 has the least.
-Also, from table \ref{tab:tab-rank-8} it could be seen that the harmony pair $(dow, hod)$ is important for id 7, however it has been labeled as an inconsequential pair for id 1.
-The distribution of energy demand for both of these households, with $dow$ as the facet and $hod$ on the x-axis, may help explain the choice.
-Figure \ref{fig:gravitas-plot-8} demonstrates that for id 7, the median (black) and quartile deviation (orange) of energy consumption fluctuates throughout for most hours of the day and day of the week, while for id 1, daily patterns are consistent within weekdays and weekends.
-As a result, for id 1, it is more appropriate to examine the distributional difference solely across (dow, wdwnd), which has been rated higher in Table \ref{tab:tab-rank-8}.
+#### Comparing households and validating rank of harmonies {-}
+
+According to Figure \ref{fig:dotplot-8}(c), for the harmony pair (*dow*-*hod*), household id 7 has the greatest value of $\wpd$, while id 1 has the least.
+From Table \ref{tab:tab-rank-8} it can be seen that the harmony pair (*dow*, *hod*) is important for id 7; however, it has been labeled as an inconsequential pair for id 1.
+The distribution of energy demand for both of these households, with *dow* as the facet and *hod* on the x-axis, may help explain the choice.
+Figure \ref{fig:gravitas-plot-8} demonstrates that for id 7, the median (black) and quartile deviation (orange) of energy consumption fluctuates for most hours of the day and days of the week, while for id 1, daily patterns are more consistent within weekdays and weekends.
+As a result, for id 1, it is more appropriate to examine the distributional difference solely across (*dow*, *wdwnd*), which has been rated higher in Table \ref{tab:tab-rank-8}.
 
@@
 With temporal data available at finer scales, exploring time series has become overwhelming with so many possible granularities to explore.
-A common solution is to aggregate and look at the patterns across usual granularities like hour-of-day or day-of-week, but there is no way to know the “interesting” granularities a priori.
+A common solution is to aggregate and look at the patterns across the usual granularities such as hour-of-day or day-of-week, but there is no way to know the “interesting” granularities a priori.
 A huge number of displays need to be analyzed or we might end up missing informative granularities.
-This work refines the search of informative granularities by identifying those for which the differences between the displayed distributions are greatest and rating them in order of importance of capturing maximum variation.
+This work refines the search for informative granularities by identifying those for which the differences between the displayed distributions are greatest and rating them in order of importance of capturing maximum variation.
 
@@
 
-## Acknowledgments {- #thankshakear}
-
-The Australian authors thank the ARC Centre of Excellence for Mathematical and Statistical Frontiers [(ACEMS)](https://acems.org.au/home) for supporting this research.
+## Acknowledgments {-}
+
+The authors thank the ARC Centre of Excellence for Mathematical and Statistical Frontiers [(ACEMS)](https://acems.org.au/home) for supporting this research.
 Sayani Gupta was partially funded by [Data61 CSIRO](https://data61.csiro.au/) during her PhD.
-The Github repository,  [github.com/Sayani07/paper-hakear](https://github.com/Sayani07/paper-gravitas), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
-This article was created with R [@R-language], `knitr` [@knitr2015], [@R-knitr] and `rmarkdown` [@rmarkdown], [@R-rmarkdown].
-Graphics are produced with @Wickham2009pk.
-
-## Supplementary Materials {- #supplement-hakear}
+The Github repository, [github.com/Sayani07/paper-hakear](https://github.com/Sayani07/paper-hakear), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
+This article was created with R [@R-language], `knitr` [@knitr; @R-knitr] and `rmarkdown` [@rmarkdown; @R-rmarkdown].
+Graphics are produced with `ggplot2` [@Wickham2009pk].
+
+## Supplementary materials {- #supplement-hakear}
 
@@
 \noindent
-**Simulation results and scripts:** All simulation table, graphics and and R code to reproduce it (paper-supplementary.pdf, paper-supplementary.Rmd)
+**Supplementary paper:** All simulation tables, graphics and and R codes to reproduce the supplementary paper (paper-supplementary.Rmd, paper-supplementary.pdf).
 
 \noindent
-**R-package:** The open-source R [@R-language] package `hakear` is available on Github (https://github.com/Sayani07/hakear) to implement ideas presented in this paper.
+**R-package:** The open-source R package `hakear` is available on Github (https://github.com/Sayani07/hakear) to implement ideas presented in this paper.
 
@@
 
-# Clustering time series based on probability distributions across temporal granularities {#ch:gracsr}
-
-With more and more time series data being collected at much finer temporal resolution, for a longer length of time, and for a larger number of individuals/entities, time series clustering research is getting a lot of traction.
-Long, noisy, patchy, uneven, and asynchronous time series are common in many fields, limiting similarity searches or lowering method efficiency when clustering is based on a distance metric.
-In this work, we suggest two approaches for obtaining similarity between time series based on probability distributions over cyclic temporal granularities for distance-based clustering approaches.
-Cyclic granularities like hour-of-the-day, work-day/weekend, month-of-the-year and so on, are useful for finding repeated patterns in the data.
-Looking at probability distributions across cyclic granularities serves two purposes: (a) "Probability distributions" characterise the inherent temporal data structure of these large unequal-length time series and are robust to missing or noisy data.
-(b) Using probability distributions over "cyclic granularities" ensures small pockets of similar "repeated" behaviors.
-Our method is capable of producing useful clusters, as demonstrated by testing on validation data designs and a sample of residential smart meter consumers.
-
-```{r external3, include = FALSE, cache = FALSE}
-knitr::read_chunk('scripts/gracsr.R')
+# Clustering time series based on probability distributions across temporal granularities {#ch-gracsr}
+
+Clustering is a potential approach for organizing large collections of time series into small homogeneous groups, but a difficult step is determining an appropriate metric to measure similarity between time series.
+The similarity metric needs to be capable of accommodating long, noisy, and asynchronous time series and also capture cyclical patterns.
+In this paper, two approaches for measuring distances between time series are presented, based on probability distributions over cyclic temporal granularities.
+Both are compatible with a variety of clustering algorithms.
+Cyclic granularities like hour-of-the-day, work-day/weekend, and month-of-the-year, are useful for finding repeated patterns in the data.
+Measuring similarity based on probability distributions across cyclic granularities serves two purposes: (a) characterizing the inherent temporal data structure of long, unequal-length time series in a manner robust to missing or noisy data; (b) small pockets of similar repeated behaviors can be captured.
+This approach is capable of producing useful clusters, as demonstrated on validation data designs and a sample of residential smart meter records.
+
+```{r external-gracsr, include = FALSE, cache = FALSE}
+#knitr::read_chunk('script/main.R')
+knitr::read_chunk('scripts/gracsr/library.R')
+knitr::read_chunk('scripts/gracsr/theme.R')
+knitr::read_chunk('scripts/gracsr/application.R')
+knitr::read_chunk('scripts/gracsr/validation.R')
 ```
@@
 
+```{r mytheme-application2}
+
+```
+
+```{r mytheme-application3}
+
+```
+
 ```{r theme-validation}
@@
 <!-- time series clustering and its challenges -->
-Time-series clustering is the process of unsupervised partitioning of $n$ time-series data into $k$ ($k<n$) meaningful groups such that homogeneous time-series are grouped together based on a certain similarity measure.
-The time-series features, length of time-series, representation technique, and, of course, the purpose of clustering time-series all influence the suitable similarity measure or distance metric to a meaningful level.
-The three primary methods to time series clustering (@liao2005clustering) are algorithms that operate directly with distances or raw data points in the time or frequency domain (distance-based), with features derived from raw data (feature-based), or indirectly with models constructed from raw data (model-based).
+Time series clustering is the process of unsupervised partitioning of $n$ time series data into $k$ ($k<n$) meaningful groups such that homogeneous time series are grouped together based on a certain similarity measure.
+The time series features, length of time series, representation technique, and, of course, the purpose of clustering time series all influence the suitable similarity measure or distance metric to a meaningful level.
+The three primary methods of time series clustering [@liao2005clustering] are algorithms that operate directly with distances or raw data points in the time or frequency domain (distance-based), with features derived from raw data (feature-based), or indirectly with models constructed from raw data (model-based).
 The efficacy of distance-based techniques is highly dependent on the distance measure utilized.
 Defining an appropriate distance measure for the raw time series may be a difficult task since it must take into account noise, variable lengths of time series, asynchronous time series, different scales, and missing data.
-Commonly used Distance-based similarity measures as suggested by a decade review of time series clustering approaches (@Aghabozorgi2015-ct) are Euclidean, Pearson's correlation coefficient and related distances, Dynamic Time Warping, Autocorrelation, Short time series distance, Piecewise regularization, cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances (@liao2007clustering) but on a vector time series data.
+Commonly used distance-based similarity measures as suggested by a review of time series clustering approaches [@Aghabozorgi2015-ct] are Euclidean, Pearson's correlation coefficient and related distances, Dynamic Time Warping (DTW), Autocorrelation, Short time series distance, Piecewise regularization, cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances [@liao2007clustering] but on vector time series data.
 Among these alternatives, Euclidean distances have high performance but need the same length of data over the same period, resulting in information loss regardless of whether it is on raw data or a smaller collection of features.
-DTW works well with time series of different lengths (@corradini2001dynamic), but it is incapable of handling missing observations.
-Surprisingly, probability distributions, which may reflect the inherent temporal structure of a time series have not been considered in determining time series similarity.
-
-\noindent We consider the problem of clustering a large number of univariate time series of continuous values which are available at fine temporal scales, being motivated by the residential smart meter data.
-These time series data are long (with more and more data collected at finer resolutions), are asynchronous, with varying time lengths for different houses and missing observations and characterized by noisy and patchy behavior that can quickly become overwhelming and hard to interpret, requiring summarizing the large number of customers into pockets of similar energy behavior.
-Choosing probability distributions seem to be a natural way to analyze these types of data sets since they are robust to uneven length, missing data, or noise.
-This paper proposes two approaches for obtaining pairwise similarities based on Jensen-Shannon distances between probability distributions across significant cyclic granularities.
-Cyclic temporal granularities [@Gupta2021-hakear], which are temporal deconstructions of a time period into units such as hour-of-the-day, work-day/weekend, can be useful for measuring repetitive patterns in large univariate time series data.
-The resulting clusters are expected to group customers that have similar repetitive behaviors across each of the interesting cyclic granularities.
-Below are some of the benefits of our method:
-
-- When using probability distributions, data does not have to be the same length or observed during the exact same time period (unless there is a structural pattern);
+DTW works well with time series of different lengths [@corradini2001dynamic], but it is incapable of handling missing observations.
+Surprisingly, probability distributions, which may reflect the inherent temporal structure of a time series, have not been considered in determining time series similarity.
+
+\noindent This work is motivated by a need to cluster a large collection of residential smart meter data, so that customers can be grouped into similar energy usage patterns.
+These can be considered to be univariate time series of continuous values which are available at fine temporal scales.
+These time series data are long (with more and more data collected at finer resolutions), are asynchronous, with varying time lengths for different houses and sporadic missing values.
+Using probability distributions is a natural way to analyze these types of data because they are robust to uneven length, missing data, or noise.
+This paper proposes two approaches for obtaining pairwise similarities based on Jensen-Shannon distances between probability distributions across a selection of cyclic granularities.
+Cyclic temporal granularities [@Gupta2021-hakear], which are temporal deconstructions of a time period into units such as hour-of-the-day or work-day/weekend, can measure repetitive patterns in large univariate time series data.
+The resulting clusters are expected to group customers that have similar repetitive behaviors across cyclic granularities.
+The benefits of this approach are as follows.
+
+- When using probability distributions, data does not have to be the same length or observed during the exact same time period (unless there is a structural pattern).
 
@@
 
-- Jensen-Shannon distances evaluate the distance between two distributions rather than raw data, which is less sensitive to missing observations and outliers than other conventional distance methods;
-
-- While most clustering algorithms produce clusters similar across just one temporal granularity, this technique takes a broader approach to the problem, attempting to group observations with similar distributions across all interesting cyclic granularities;
+- Jensen-Shannon distances evaluate the distance between two distributions rather than raw data, which is less sensitive to missing observations and outliers than other conventional distance methods.
+
+- While most clustering algorithms produce clusters similar across just one temporal granularity, this technique takes a broader approach to the problem, attempting to group observations with similar distributions across all interesting cyclic granularities.
 
@@
 
-_Background and motivation_
-
-Large spatio-temporal data sets, both from open and administrative sources, offer up a world of possibilities for research.
-One such data sets for Australia is the Smart Grid, Smart City (SGSC) project (2010–2014) available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) The project provides half-hourly data of over 13,000 household electricity smart meters distributed unevenly from October 2011 to March 2014.
-Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.
-Households vary in size, location, and amenities such as solar panels, central heating, and air conditioning.
+<!--_Background and motivation_-->
+
+The primary application of this work is data from the Smart Grid, Smart City (SGSC) project (2010–2014) available through @smart-meter.
+Half-hourly measurements of usage for more than 13,000 electricity smart meter customers are provided from October 2011 to March 2014.
+<!--Larger data sets include greater uncertainty about customer behavior due to growing variety of customers.--> Customers vary in size, location, and amenities such as solar panels, central heating, and air conditioning.
 The behavioral patterns differ amongst customers due to many temporal dependencies.
-Some households use a dryer, while others dry their clothes on a line.
-Their weekly profile may reflect this.
+Some customers use a dryer, while others dry their clothes on a line.
+Their weekly usage profile may reflect this.
 They may vary monthly, with some customers using more air conditioners or heaters than others, while having equivalent electrical equipment and weather circumstances.
 Some customers are night owls, while others are morning larks.
-Day-off energy use varies depending on whether customers stay home or go outside.
-Age, lifestyle, family composition, building attributes, weather, availability of diverse electrical equipment, among other factors, make the task of properly segmenting customers into comparable energy behavior a fascinating one.
-This challenge is worsened when all we know about our consumers is their energy use history (@Ushakova2020-rl).
-To safeguard the customers' privacy, it is probable that such information is not accessible.
-Also, energy suppliers may not always update client information, such as property features, in a timely manner.
-Thus, there is a growing need to have research that examines how much energy usage heterogeneity can be found in smart meter data and what are some of the most common power consumption patterns, rather than explaining why consumption differs.
-
-_Related work_
-
-A multitude of papers have emerged around smart meter time series clustering for deepening our knowledge of consumption patterns.
+Daily energy usage varies depending on whether customers stay home or work away from home.
+Age, lifestyle, family composition, building attributes, weather, availability of diverse electrical equipment, among other factors, make the task of properly segmenting customers into comparable energy behavior complex.
+The challenge is to be able to cluster consumers into these types of expected patterns, and other unexpected patterns, using only their energy usage history [@Ushakova2020-rl].
+<!--To safeguard the customers' privacy, it is probable that such information is not accessible.
+Also, energy suppliers may not always update client information, such as property features, in a timely manner.--> There is a growing need to have methods that can examine the energy usage heterogeneity observed in smart meter data and what are some of the most common power consumption patterns.
+
+<!-- _Related work_-->
+
+There is an extensive body of literature focused on time series clustering related to smart meter data.
 @Tureczek2017-pb conducted a systematic study of over $2100$ peer-reviewed papers on smart meter data analytics.
-None of the $34$ articles chosen for their emphasis use Australian smart meter data.
-The most often used algorithm is K-Means.
-Using K-Means without considering time series structure or correlation results in inefficient clusters.
-Principal Component Analysis (PCA) or Self-Organizing Maps (SOM) eliminate correlation patterns and decrease feature space, but lose interpretability.
-To reduce dimensionality, several studies use principal component analysis or factor analysis to pre-process smart-meter data before clustering (@Ndiaye2011-pf).
-Other algorithms utilized in the literature include k-means variants, hierarchical approaches, and greedy k-medoids.
+<!--None of the $34$ articles chosen for their emphasis use Australian smart meter data.--> The most often used algorithm is $k$-means [@Rhodes2014-it].
+$k$-means can be made to perform better by explicitly incorporating time series features such as correlation or cyclic patterns rather than performing it on raw data.
+To reduce dimensionality, several studies use principal component analysis (PCA) or factor analysis to pre-process smart-meter data before clustering [@Ndiaye2011-pf].
+PCA eliminates correlation patterns and decreases feature space, but loses interpretability.
+Other algorithms utilized in the literature include $k$-means variants, hierarchical clustering, and greedy $k$-medoids.
 Time series data, such as smart meter data, are not well-suited to any of the techniques mentioned in @Tureczek2017-pb.
-Only one study [@ozawa2016determining] identified time series characteristics using Fourier transformation, which converts data from time to frequency and then uses K-Means to cluster by greatest frequency.
-@Motlagh2019-yj suggests that the time feature extraction is limited by the type of noisy, patchy, and unequal time-series common in residential and addresses model-based clustering by transforming the series into other other objects such as structure or set of parameters which can be more easily characterized and clustered.
-[@chicco2010renyi] addresses information theory-based clustering such as Shannon or Renyi entropy and its variations.
-@Melnykov2013-sp discusses how outliers, noisy observations and scattered observations can complicate estimating mixture model parameters and hence the partitions.
-To our knowledge, none of the methods focus on exploring heterogeneity in repetitive behaviors based on the dynamics of multiple temporal dependencies using probability distributions.
+Only one study [@ozawa2016determining] identified time series characteristics by first conducting a Fourier transformation, to convert data from time to frequency domain, followed by $k$-means to cluster by greatest frequency.
+@Motlagh2019-yj suggested that the time feature extraction is limited by the type of noisy, patchy, and unequal time series common in residential customers and addresses model-based clustering by transforming the series into other objects such as structure or set of parameters which can be more easily characterized and clustered.
+@chicco2010renyi addressed information theory-based clustering such as Shannon or Renyi entropy and its variations.
+@Melnykov2013-sp discussed how outliers, noisy observations and scattered observations can complicate estimating mixture model parameters and hence the partitions.
+None of these methods focuses on exploring heterogeneity in repetitive patterns based on the dynamics of multiple temporal dependencies using probability distributions, which forms the basis of the methodology reported here.
 
@@
 
-The remainder of the paper is organized as follows: Section&nbsp;\ref{sec:methodology} provides the clustering methodology.
-Section&nbsp;\ref{sec:validation} shows data designs to validate our methods.
-Section&nbsp;\ref{sec:cases} discusses the application of the method to a subset of the real data.
-Finally, we summarize our results and discuss possible future directions in Section&nbsp;\ref{sec:conclusion}.
+This paper is organized as follows.
+Section \ref{sec:methodology} provides the clustering methodology.
+Section \ref{sec:validation} shows data designs to validate our methods.
+Section \ref{sec:application-gracsr} discusses the application of the method to a subset of the real data.
+Finally, we summarize our results and discuss possible future directions in Section \ref{sec:discussion-gracsr}.
 
@@
 <!-- This work -->
-<!-- This is similar to a stochastic approach (@Motlagh2019-yj) to clustering, which proposes interpreting electricity demand as a random process and extracting time-series characteristics, or a model of the series, to enable unsupervised clustering.
+<!-- This is similar to a stochastic approach [@Motlagh2019-yj) to clustering, which proposes interpreting electricity demand as a random process and extracting time series characteristics, or a model of the series, to enable unsupervised clustering.
 Unsupervised clustering is only as good as the features that are extracted/selected or the distance metrics that were utilized.
@@
 
-The foundation of our method is unsupervised clustering algorithms based exclusively on the time-series data.
-The proposed methodology aims to leverage the intrinsic data structure hidden within cyclic temporal granularities.
-The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
-<!--So going back to the smart meter example, $f_i(v)$ is the distribution of customer $i$ and $v$ is electricity demand.--> In this work, instead of considering the probability distributions of the linear time series, we assume it across different categories of any cyclic granularity.
-We can consider categories of an individual cyclic granularity ($A$) or combination of categories for two interacting granularities ($A*B$) to have a distribution, where $A, B$ are defined as $A = \{a_j: j=1, 2, \dots J\}$ and $B = \{b_k: k = 1, 2, \dots K\}$.
-For example, let us consider two cyclic granularities $A$ and $B$ representing hour-of-day and day-of-week.
+<!--The foundation of our method is unsupervised clustering algorithms based exclusively on the time series data.
+The proposed methods leverage the intrinsic data structure hidden within cyclic temporal granularities.--> The existing work on clustering probability distributions assumes we have independent and identically distributed samples $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points.
+<!--So going back to the smart meter example, $f_i(v)$ is the distribution of customer $i$ and $v$ is electricity demand.--> In our approach, instead of considering the probability distributions of the linear time series, we compare them across different categories of a cyclic granularity.
+We can consider categories of an individual cyclic granularity ($A$) or combination of categories for two interacting granularities ($A, B$) to have a distribution, where $A$ and $B$ are defined as $A = \{a_j: j=1, 2, \dots J\}$ and $B = \{b_k: k = 1, 2, \dots K\}$.
+For example, let us consider two cyclic granularities, $A$ and $B$, representing hour-of-day and day-of-week, respectively.
 Then $A = \{0, 1, 2, \dots, 23\}$ and $B = \{Mon, Tue, Wed, \dots, Sun\}$.
 In case individual granularities are considered, there are $J = 24$ distributions of the form $f_{i,j}(v)$ or $K = 7$ distributions of the form $f_{i,k}(v)$ for each customer $i$.
-In case of interaction, $J*K=168$ distributions of the form $f_{i,j, k}(v)$ could be conceived for each customer $i$.
+In case of interaction, $J \times K=168$ distributions of the form $f_{i,j, k}(v)$ could be conceived for each customer $i$.
 Hence clustering these customers is equivalent to clustering these collections of conditional univariate probability distributions.
-Towards this goal, we need to decide how to measure similarities between collections of univariate probability distributions.
-There are multiple ways to measure similarities depending on the aim of the analysis.
+Towards this goal, the next step is to decide how to measure distances between collections of univariate probability distributions.
+<!--There are multiple ways to measure similarities depending on the aim of the analysis.
 <!--This paper focuses on looking at the (dis) similarity between underlying distributions that may have resulted in different patterns across different cyclic temporal granularities, that eventually have resulted in the (dis) similarity between time series.
---> This paper considers a two approaches for finding distances between time series.
-Both of these approaches may be useful in a practical context and, depending on the data set, may or may not propose the same customer classification.
-The obtained distances could be fed into a clustering algorithm to break large data sets into subgroups that can then be analyzed separately.
+--> Here, we describe two approaches for finding distances between time series.
+Both of these approaches may be useful in a practical context, and produce very different but equally useful customer groupings.
+The distances can be supplied to any usual clustering algorithm, including $k$-means or hierarchical clustering, to group observations into a smaller more homogeneous collection.
 <!--These clusters may be commonly associated with real-world data segmentation.
-However, since the data is unlabeled a priori, more information is required to corroborate this.--> The methodology is explained in the Figure \ref{fig:flowchart} and each element of the pipeline is discussed.
-
-```{r flowchart, fig.cap = "Flow chart illustrating the pipeline for methodology", out.width="100%"}
+However, since the data is unlabeled a priori, more information is required to corroborate this.--> The flow of the procedures is illustrated in Figure \ref{fig:flowchart} and is further described in the following subsections.
+
+```{r flowchart, fig.cap = "Flow chart illustrating the pipeline for our method for clustering time series.", out.width="100%"}
 knitr::include_graphics("img/flowchart.png")
@@
 
-- _Find significant granularities or harmonies_
-
-[@Gupta2021-hakear] proposes a method for choosing significant cyclic granularities and harmonies (interacting granularities that can be studied together) [@Gupta2021-gravitas], which is used in this work.
-We define "significant" granularities as those with significant distributional differences across categories.
-It is preferable to consider those since it is more probable that there will be some intriguing repeated behavior worth investigating.
-It should be noted that all of the observations in the study may not have the same set of important granularities.
-The following is a method for generating a list ($S_c$) of significant granularities for all observations:
-
- (a) Remove granularities from the comprehensive list that are insignificant for all observations.
-
- (b) select only the granularities that are significant for the majority of observations.
-
-There will be observations in both cases where one or a few selected granularities are boring.
+### Selecting granularities
+
+@Gupta2021-hakear provide a distance measure ($\wpd$) for determining the significance of a cyclic granularity, and a ranking of multiple cyclic granularities.
+(This extends to harmonies, pairs of granularities that might interact with each other.) We define "significant" granularities as those with significant distributional differences across at least one category.
+The reason for subsetting granularities in this way is that clustering algorithms perform badly in the presence of nuisance variables.
+Granularities that do not have some difference between categories are likely to be nuisance variables.
+It should be noted that all of the time series in a collection may not have the same set of significant granularities.<!--This is the approach for managing this generate a subset ($S_c$) of significant granularities across a collection of time series:--> This is the approach for generating a subset ($S_c$) of significant granularities across a collection of time series:
+
+ (a) Remove granularities from the comprehensive list that are not significant for any time series.
+
+ (b) Select only the granularities that are significant for the majority of time series.
+
+<!--There will be time series in both cases where one or a few selected granularities are boring.
 Even in that case, having this group of observations with no interesting patterns at a granularity that regularly discovers patterns may be valuable.
-If, on the other hand, the granularities under examination are truly important for a group of data, unique patterns may be identified while clustering them.
-
-- _Data transformation_
-
-Observations often have a somewhat skewed time series distribution and their ranges might vary greatly.
+If, on the other hand, the granularities under examination are truly important for a group of data, unique patterns may be identified while clustering them.-->
+
+### Data transformation
+
+The shape and scale of the distribution of the measured variable (e.g.
+energy usage) affects distance calculations.
+Skewed distributions need to be symmetrized.
+Scales of individuals need to be standardized, because clustering is to select similar patterns, not magnitude of usage.
+(Organizing individuals based on magnitude can be achieved simply by sorting on a statistic like the average value across time.) <!--Observations often have a somewhat skewed time series distribution and their ranges might vary greatly.
 Statistical transformations are employed to bring all of them to the same range or normalize each observation.
-This is important because we are not interested in trivial clusters that vary in magnitudes, but rather in uncovering comparable patterns of distributional differences between categories.
-For the JS-based approaches, two data transformation techniques are utilized viz, Normal-Quantile Transform (NQT) and Robust scaling.
-NQT is a built-in transformation for computing $wpd$, which is the foundation of wpd-based distances.
-
-_Robust scaling_ The normalized $i^{th}$ observation is denoted by $v_{norm} = \frac{v_t - q_{0.5}}{q_{0.75}-q_{0.25}}$, where $v_t$ is the actual value at the $t^{th}$ time point and $q_{0.25}$, $q_{0.5}$ and $q_{0.75}$ are the $25^{th}$, $50^{th}$ and $75^{th}$ percentile of the time series for the $i^{th}$ observation.
-$v_{norm}$ has zero mean and median, as well as a standard deviation of one, with the outliers having same relative connections to other values.
-
-_Normal-Quantile transform_ The raw data for all observations is individually normal-quantile transformed (NQT) [@Krzysztofowicz1997-bv], so that the transformed data follows a standard normal distribution.
-NQT will make the skewed distributions bell-shaped.
-As a result, determining which raw distribution was used is difficult using the modified distribution.
-Multimodality is also concealed or reversed.
-As a result, while displaying transformed data using NQT, one must exercise caution.
-But NQT is a useful transformation that often improves clustering performance.
+This is important because we are not interested in trivial clusters that vary in magnitudes, but rather in uncovering comparable patterns of distributional differences between categories.--> For the JS-based approaches, two data transformation techniques are recommended, normal-quantile transform (NQT) and robust scaling (RS).<!-- NQT is used in @Gupta2021-hakear prior sorting granularities.-- $\wpd$, which is the foundation of wpd-based distances.
+XX wpd was not yet introduced--> While @Gupta2021-hakear already use NQT when computing $\wpd$, it could be useful to standardize it for the selected set of significant granularities prior to computing the distances.
+
+- RS: The normalized $i^{th}$ observation is denoted by $v_{norm} = \frac{v_t - q_{0.5}}{q_{0.75}-q_{0.25}}$, where $v_t$ is the actual value at the $t^{th}$ time point and $q_{0.25}$, $q_{0.5}$ and $q_{0.75}$ are the $25^{th}$, $50^{th}$ and $75^{th}$ percentiles of the time series for the $i^{th}$ observation.
+Note that $v_{norm}$ has zero mean and median, but otherwise the shape does not change.
+
+- NQT: The raw data for all observations is individually transformed [@Krzysztofowicz1997-bv], so that the transformed data follows a standard normal distribution.
+NQT will symmetrize skewed distributions.
+<!--As a result, determining which raw distribution was used is difficult using the modified distribution.--> A drawback is that any multimodality will be concealed.
+This should be checked prior to applying NQT.
 
@@
 
-- _Data pre-preprocessing_
-
-We start with a "tsibble" (@wang2020tsibble) data structure with an index variable representing inherent ordering from past to present and a key variable that defines observational units over time.
+### Data pre-preprocessing
+
+Computationally in R, the data is assumed to be a "tsibble object" [@wang2020tsibble] equipped with an index variable representing inherent ordering from past to present and a key variable that defines observational units over time.
 The measured variable for an observation is a time-indexed sequence of values.
@@
 A shuffle of the raw sequence may represent hourly consumption throughout a day, a week, or a year.
-Cyclic granularities can be expressed in terms of the index set in the "tsibble" data structure.
-But the data structure changes while transporting from linear to cyclic scale of time as multiple observations now correspond to each category and induce a probability distribution.
-Directly computing Jensen-Shannon distances between the entire probability distributions can be very costly.
-Hence, in this paper, quantiles are chosen to characterize the probability distributions.
-So, in the final data structure, each category of a cyclic granularity corresponds to a list of numbers which is essentially a few chosen quantiles.
-
-- _Distance metrics_
+Cyclic granularities can be expressed in terms of the index set in the tsibble data structure.
+
+The data object will change when cyclic granularities are computed, as multiple observations will be categorized into levels of the granularity, thus inducing multiple probability distributions.
+Directly computing Jensen-Shannon distances between the entire probability distributions can be computationally intensive.
+Thus it is recommended that quantiles are used to characterize the probability distributions.
+In the final data object, each category of a cyclic granularity corresponds to a list of numbers which is composed of a few quantiles.
+
+### Distance metrics
 
@@
 
-**JS-based distances**
+#### JS-based distances
 
 This distance metric considers two time series to be similar if the distributions of each category of an individual cyclic granularity or combination of categories for interacting cyclic granularities are similar.
-In this study, the distribution for each category is characterized using deciles (can potentially consider any list of quantiles), and the distances between distributions are calculated using the Jensen-Shannon distances (@Menendez1997-in), which are symmetric and thus could be used as a distance measure.
-
-\noindent The sum of the distances between two observations $x$ and $y$ in terms of cyclic granularity $A$ is defined as $$S^A_{x,y} = \sum_{j} D_{x,y}(A)$$ (sum of distances between each category $j$ of cyclic granularity $A$) or $$S^{A*B}_{x,y} = \sum_j \sum_k D_{x,y}(A, B)$$ (sum of distances between each combination of categories $(j, k)$ of the harmony $(A, B)$.
+In this study, the distribution for each category is characterized using deciles (can potentially consider any list of quantiles), and the distances between distributions are calculated using the Jensen-Shannon distances [@Menendez1997-in], which are symmetric and thus could be used as a distance measure.
+
+The sum of the distances between two observations $x$ and $y$ in terms of a cyclic granularity $A$ is defined as $$S^A_{x,y} = \sum_{j\in A} D(x_j,y_j)$$ where $D$ is the Jensen-Shannon distances, $x_j$ is the set of quantiles over the values filtered by $j^{th}$ level of granularity $A$ for observation $x$ (similar for $y$).
+
+The sum of the distances between two observations $x$ and $y$ in terms of a pair of cyclic granularities $(A, B)$ is defined as $$S^{A,B}_{x,y} = \sum_{(j, k) \in (A, B)} D(x_{jk},y_{jk})$$ $x_{jk}$ is the set of quantiles over the values filtered by the combination of $j^{th}$ level of granularity $A$ and $k^{th}$ level of granularity $B$ for observation $x$ (similar for $y$).
+
 After determining the distance between two series in terms of one granularity, we must combine them to produce a distance based on all significant granularities.
-When combining distances from individual $L$ cyclic granularities $C_l$ with $n_l$ levels, $$S_{x, y} = \sum_lS^{C_l}_{x,y}/n_l$$ is employed, which is also a distance metric since it is the sum of JS distances.
+When combining distances from individual $L$ cyclic granularities $C_l$ with $n_l$ levels, $$S_{x, y} = \sum_{l \in L}S^{C_l}_{x,y}/n_l$$ is employed, which is also a distance metric since it is the sum of JS distances.
 This approach is expected to yield groups, such that the variation in observations within each group is in magnitude rather than distributional pattern, while the variation between groups is only in distributional pattern across categories.
 
-**wpd-based distances**
-
-Compute weighted pairwise distances $wpd$ (@Gupta2021-hakear) for all considered granularities for all observations.
-$wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction and is estimated by the maximum pairwise distances between consecutive categories normalized by appropriate parameters.
-A higher value of $wpd$ indicates that some interesting pattern is expected, whereas a lower value would indicate otherwise.
-
-\noindent Once we have chosen $wpd$ as a relevant feature for characterizing the distributions across one cyclic granularity, we have to decide how we combine differences between the multiple features (corresponding to multiple granularities) into a single number.
-The Euclidean distance between them is chosen, with the granularities acting as variables and $wpd$ representing the value under each variable.
-With this approach, we should expect the observations with similar $wpd$ values to be clustered together.
+#### wpd-based distances
+
+We compute weighted pairwise distances $\wpd$ [@Gupta2021-hakear] for all considered granularities for all observations.
+$\wpd$ is designed to capture the maximum variation in the measured variable explained by an individual cyclic granularity or their interaction.
+It is estimated by the maximum pairwise distances between distributions across consecutive categories normalized by appropriate parameters.
+A higher value of $\wpd$ indicates that some interesting patterns are expected, whereas a lower value would indicate otherwise.
+
+Once we have chosen $\wpd$ as a relevant feature for characterizing the distributions across one cyclic granularity, we have to decide how we combine differences between the multiple features (corresponding to multiple granularities) into a single number.
+The Euclidean distance between them is chosen, with the granularities acting as variables and $\wpd$ representing the value under each variable.
+With this approach, we should expect the observations with similar $\wpd$ values to be clustered together.
 Thus, this approach is useful for grouping observations that have a similar significance of patterns across different granularities.
@@
 If we have $H_{N_C}$ harmonies in the harmony table, then for each household we have a vector of $wpd_{{s}}$ of $H_{N_C}$ elements with each element corresponding to one harmony.
-We aim to have pockets of households showing similar periodic behavior by considering $wpd$ values for different harmonies and some time series features.
+We aim to have pockets of customers showing similar periodic behavior by considering $\wpd$ values for different harmonies and some time series features.
 The features should also characterize probability distributions of different household.
@@
 
-- _Clustering algorithm_
+### Clustering
+
+#### Number of clusters
+
+Determining the number of clusters is typically a difficult task.
+Many metrics have been defined for choosing clusters.
+Most metrics for choosing the optimal number of clusters are based on comparing distances between observations within a class to those distances between observations between classes, which makes the assumption that there are some separated clusters.
+Some common procedures include the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@Dunn1973-gk] and the separation index (*sindex*) [@Hennig2019-eq; @Hennig2014-ah].
+These are constructed by balancing within-cluster homogeneity and between-cluster separation.
+
+All of the common approaches can give contradictory suggestions for the optimal number of clusters, particularly when the data does not naturally break into groups, or in the presence of nuisance variables (no contribution to clustering) or nuisance observations (inlying and outlying observations falling between clusters).
+There is no one best metric, which is perhaps a reason why so many metrics exist.
+
+In this work, we have chosen to use *sindex*.
+It is a very simple but effective metric.
+This is computed by averaging the smallest 10% of inter-cluster distances.
+It is relatively robust to nuisance observations.
+The value of *sindex* always decreases, and sharp drops in value indicate candidates for the optimal number of clusters.
+The number of clusters corresponding to the value **before the drop** is the recommendation.
+
+<!--determined by calculating the distance between each observation and the cluster to which it does not belong.
+Then, the least 10% of these distances are averaged to only evaluate places near the cluster edges.-->
+
+#### Algorithm
 
@@
 
-- _Characterization of clusters_
-
-\noindent Cluster characterization is an important element of cluster analysis.
-@Cook2007-qe provides several methods for characterizing clusters.
-_Parallel coordinate plots_ (@wegman1990hyperdimensional), _Scatterplot matrix_, _Displaying cluster statistics_ (@dasu2005grouping), _MDS_ (@borg2005modern), PCA, _t-SNE_(@R-tsne), _Tour_ (@wickham2011tourr) are some of the graphical approaches used in this study.
-A Parallel Coordinates Plot features parallel axes for each variable and each axis is linked by lines.
+#### Characterization of clusters
+
+Cluster characterization is an important final stage of a cluster analysis.
+The primary purpose is to compare the homogeneity within a cluster to the heterogeneity of clusters.
+This can be done numerically, by tabulating cluster means and standard deviations [@dasu2005grouping], and visually using methods for graphics multivariate data.
+@Cook2007-qe provide visual examples using both tours [@asimov1985grand] and parallel coordinate plots [@wegman1990hyperdimensional].
+Dimension reduction techniques like principal component analysis [@jolliffe2016principal], multidimensional scaling (MDS) [@borg2005modern], t-distributed stochastic neighbor embedding (t-SNE) [@van2008visualizing] and linear discriminant analysis (LDA) [@fisher1936use] are also useful.
+<!-- A Parallel Coordinates Plot features parallel axes for each variable and each axis is linked by lines.
 Changing the axes may reveal patterns or relationships between variables for categorical variables.
@@
 Displaying cluster statistics is useful when we have larger problems and it is difficult to read the parallel coordinate plots due to congestion.
-All of MDS, PCA and t-SNE use a distance or dissimilarity matrix to construct a reduced-dimension space representation, their goals are diverse.
-Multidimensional scaling (@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space.
+All of MDS, PCA and t-SNE use a distance or dissimilarity matrix to construct a reduced-dimension space representation, but their goals are diverse.
+Multidimensional scaling [@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space.
 The t-SNE embedding will compress data points that are close in high-dimensional space.
@@
 So three scenarios are considered to accommodate that.
+Table \ref{tab:range-parameter} shows the range of parameters considered for each scenario.
+
+<!-- ```{r range-parameter, fig.cap="The range of parameters used for the validation study, for the three different scenarios, number of simulations for each design, differences between means across granularities and series lengths.
+"} -->
+<!-- knitr::include_graphics("img/simulation_table.png") -->
+<!-- ``` -->
+
+```{r range-parameter}
+scenario = c("S1", "S2","S3")
+designs = c(5, 4, 4)
+niter = list (c(25, 250, 500),
+              c(25, 250, 500),
+              c(25, 250, 500))
+mu = list (c(1, 2, 5),
+           c(1, 2, 5),
+           c(1, 2, 5))
+nT = list (c(300, 1000, 5000),
+           c(300, 1000, 5000),
+           c(300, 1000, 5000))
+
+tibble::tibble(scenario, designs, niter, mu, nT) %>% rlang::set_names("scenario", "designs", "R", "$\\mu$", "T") %>% 
+  #tibble::tibble(scenario, " designs" = designs, "R" =niter,  "$\\mu$" = mu, "T"=nT) %>% 
+  knitr::kable(format = "latex",
+             escape = FALSE,
+             caption = "The range of parameters used for the validation study, for the three different scenarios, number of simulations ($R$) for each design, differences between means ($\\mu$) across granularities and series lengths ($T$).")%>%  kableExtra::collapse_rows(columns = 3:5)
+```
 
@@
 
-_Scenario (a): All significant granularities_
-
+**Scenario 1 (S1) - All granularities significant:**
 <!-- Consider a case where all the three granularities $g1$, $g2$ and $g3$ would be responsible for making the designs distinct.
@@
 -->
-
-\noindent Consider the instance where $g1$, $g2$, and $g3$ all contribute to design distinction.
+Consider the instance where $g1$, $g2$, and $g3$ all contribute to design distinction.
 This means that each granularity will have significantly different patterns at least across one of the designs to be clustered.
@@
 
-_Scenario (b): Few significant granularities_
-
-\noindent This is the case where one granularity will remain the same across all designs.
+**Scenario 2 (S2) - Few significant granularities:**
+This is the case where one granularity will remain the same across all designs.
 We consider the case where the distribution of $v$ varies across $g2$ levels for all designs, across $g3$ levels for a few designs, and $g1$ does not vary across designs.
-The proposed design is shown in Figure \ref{fig:gran2and1-clubbed}(b).
-
-_Scenario (c): One significant granularity_
-
-\noindent Only one granularity is responsible for identifying the designs in this case.
-This is depicted in Figure \ref{fig:gran2and1-clubbed} (right) where only $g3$ affects the designs significantly.
+The proposed design is shown in Figure \ref{fig:gran2and1-clubbed}(right).
+
+**Scenario 3 (S3) - Only one significant granularity:**
+Only one granularity is responsible for identifying the designs in this case. This is depicted in Figure \ref{fig:gran2and1-clubbed} (right) where only $g3$ affects the designs significantly.
 
@@
 Consider a group that has a different weekday and weekend behavior in the summer but not in the winter.
-This type of combined behavior across granularities can be discovered by evaluating the distribution across combinations of categories for different interacting granularities (Weekend/Weekday and month-of-year in this example).
+This type of combined behavior across granularities can be discovered by evaluating the distribution across combinations of categories for different interacting granularities (weekend/weekday and month-of-year in this example).
 As a result, in this scenario, we analyze a combination of categories generated from different distributions.
-Consider a case in which there are only two interacting granularities of interest, $g1$ and $g2$.
-In contrast to the previous situation, when we could study distributions across $n_{g_1} + n_{g_2} = 5$ separate categories, with interaction, we must evaluate the distribution of the $n_{g_1}*n_{g_2}=6$ combination of categories.
-Consider the $4$ designs in Figure \ref{fig:interaction-gran}, where various distributions are assumed for different combinations of categories, resulting in different designs.
-Design $D1$ exhibits no change in distributions across $g1$ or $g2$, whereas Designs $D2$ and $D3$ alter across only $g1$ and $g2$, respectively.
-D4 varies across both $g1$ and $g2$ categories.
-D3 and D4 appear similar based on their relative differences across consecutive categories, but $D4$ also changes across facets, unlike $D3$, which has all facets look the same.
-
-```{r interaction-gran, fig.cap = "Distribution of the simulated variable across $g1$ conditional on $g2$ is shown through boxplots for 4 designs.
-D1 has no change in distributions across different categories of $g1$ or $g2$, while D2 and D3 change across only $g1$ and $g2$ respectively.
-D4 changes across categories of both $g1$ and $g2$."}
-
-```
+Display of design and related results can be found in supplementary material.
 
@@
 
-### Visual exploration of findings
-
-All of the approaches were fitted to each data design and for each combination of the considered parameters.
+### Visual exploration of results
+
+All of the approaches were fitted to each data design and to each combination of the considered parameters.
 The formed clusters have to match the design, be well separated, and have minimal intra-cluster variation.
-It is possible to study these desired clustering traits visually in a more comprehensive way than just looking at index values.
-So we use MDS and parallel coordinate graphs to demonstrate the findings:
-
-- In Figure \ref{fig:mds-plot-validation}, we tried to see how separated our clusters are.
-We observe that in all scenarios and for different mean differences, clusters are separated.
+<!--It is possible to study these desired clustering traits visually in a more comprehensive way than just looking at index values.--> MDS and parallel coordinate graphs are used to demonstrate the findings, as well as an index value plot to provide direction on the number of clusters.
+JS-based approaches corresponding to NQT and RS are referred to as JS-NQT and JS-RS respectively.
+In the following plots, results for JS-NQT are reported, and results with JS-RS or wpd-based distances are in the supplementary material.
+
+Figure \ref{fig:sindex-plot-validation} shows *sindex* plotted against the number of clusters ($k$) for the range of mean differences (rows) under the different scenarios (columns).
+This can be used to determine the number of clusters for each scenario.
+When *sindex* for each scenario are examined, it appears that $k = \{5, 4, 4\}$ is justified for scenarios S1, S2, and S3, respectively, given the sharp decrease in *sindex* from that value of $k$.
+Thus, the number of clusters corresponds to the number of designs that were originally considered in each scenario.
+
+Figure \ref{fig:mds-plot-validation} shows separation of our clusters.
+It can be observed that in all scenarios and for different mean differences, clusters are separated.
 However, the separation increases with an increase in mean differences across scenarios.
@@
 
-- Figure \ref{fig:parcoord-sim} depicts a parallel coordinate plot with the vertical bar showing total inter-cluster distances with regard to granularities $g1$, $g2$, and $g3$ for all simulation settings and scenarios.
+Figure \ref{fig:parcoord-sim} depicts a parallel coordinate plot with the vertical bar showing total inter-cluster distances with regard to granularities $g1$, $g2$, and $g3$ for all simulation settings and scenarios.
 So one line in the figure shows the inter-cluster distances for one simulation setting and scenarios vary across facets.
 The lines are not colored by group since the purpose is to highlight the contribution of the factors to categorization rather than class separation.
-The first plot shows that no variable stands out in the clustering, but the following two designs show that {g1} and {g1, g2} have very low inter cluster distances, meaning that they did not contribute to the clustering.
-It is worth noting that these facts correspond to our original assumptions when developing the scenarios, which incorporate distributional differences over three (a), two (b), and one (c) significant granularities.
-Hence, Figure \ref{fig:parcoord-sim} (a), (b), and (c) validate the construction of scenarios (a), (b), and (c) respectively.
-
- - The js-robust and wpd methods perform worse for $nT=300$, then improve for higher $nT$ evaluated in the study.
-Although, a complete year of data is the minimum requirement to capture distributional differences in winter and summer profiles, for example.
+Panel S1 shows that no variable stands out in the clustering, but the following two panels show that {g1} and {g1, g2} have very low inter-cluster distances, meaning that they did not contribute to the clustering.
+It is worth noting that these facts correspond to our original assumptions when developing the scenarios, which incorporate distributional differences over three (S1), two (S2), and one (S3) significant granularities.
+Hence, Figure \ref{fig:parcoord-sim} (S1), (S2), and (S3) validate the construction of scenarios (S1), (S2), and (S3) respectively.
+
+The JS-RS and wpd-based methods perform worse for $nT=300$, then improve for higher $nT$ evaluated in the study.
+However, a complete year of data is the minimum requirement to capture distributional differences in winter and summer profiles, for example.
 Even if the data is only available for a month, $nT$ with half-hourly data is expected to be at least $1000$.
-As a result, as long as the performance is promising for higher $nT=300$, this is not a challenge.
-
-- In our study sample, the method js-nqt outperforms the method js-robust for smaller differences between categories. More testing, however, is required to corroborate this.
-
-For more detailed results, please refer to the supplementary paper.
-The code for creating these data designs and running the methodologies is available at (https://github.com/Sayani07/paper-gracsR/Validation).
-
-```{r parcoord-sim, fig.cap = "The parallel coordinate plot illustrates the total inter-cluster distances for granularities $g1$, $g2$, and $g3$.
-One line in the figure depicts the inter-cluster distances for a single simulation scenario.
-While the first plot indicates that no variable stands out during clustering, the next two designs demonstrate that {g1} and {g1, g2} have extremely low inter-cluster distances, indicating that they did not contribute to clustering.
-It is worth emphasising that these facts are consistent with our initial assumptions when designing the scenarios and (a), (b), and (c) correspond to Scenario (a), (b) and (c) respectively."}
-
-```
-
-```{r mds-plot-validation, fig.cap = "Relative positions of clusters corresponding to different scenarios (columns) for different values of mean differences between categories (rows) are shown using the first two dimensions of MDS.
-It can be observed that clusters become more compact and separated for higher mean differences between categories across all designs.
-Between designs, separation is least prominent corresponding to scenario (c) where only granularity is responsible for the clusters."}
+As a result, as long as the performance is promising for higher $nT$, this is not a challenge.
+
+In our study sample, the method JS-NQT outperforms the method JS-RS for smaller differences between categories. More testing, however, would be needed to to be confident in this conclusion.
+
+```{r sindex-data-validation}
+
+```
+
+```{r sindex-plot-validation, fig.cap = "Choosing optimal cluster number across the range of scenarios and mean differences used in the validation study, using the cluster separation index (sindex) for the JS-NQT.
+S1 has a sharp decrease in sindex from 5 to 6, whereas S2 and S3 have a decrease from 4 to 5, especially when mean difference is large, providing the recommended number of clusters to be 5, 4, 4, respectively.
+This precisely reflects the structure in designs that we would hope the clustering could recover."}
+```
+
+```{r mds-data-validation}
+
+```
+
+```{r mds-plot-validation, fig.cap = "MDS summary plots to illustrate the cluster separation for the range of mean differences (rows) under the different scenarios (columns).
+It can be observed that clusters become more compact and separated for higher mean differences between categories across all scenarios.
+Between scenarios, separation is least prominent corresponding to Scenario (S3) where only one granularity is responsible for distinguishing the clusters."}
+
+```
+
+```{r parcoord-sim, fig.cap = "Exploring the contribution of granularities in the clustering for scenarios S1, S2, S3, using parallel coordinate plots.
+Inter-cluster distances are displayed vertically.
+All three granularities $g1$, $g2$, and $g3$ have high inter-cluster distances for S1, suggesting all are important.
+In S2 $g1$ and in S3 both $g1$ and $g2$ have smaller inter-cluster distances, indicating that they did not contribute to clustering."}
 
@@
 
-## Case study {#sec:cases}
-
-The use of our methodology is illustrated on smart meter energy usage for a sample of customers from [SGSC consumer trial data](https://data.gov.au/data/dataset/smart-grid-smart-city-customer-trial-data) which was available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) and Data61 CSIRO.
-It contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.
-In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual.
+## Application {#sec:application-gracsr}
+
+Clustering with the new distances is illustrated on the smart meter energy usage for a sample of customers from @smart-meter.
+The full data contains half-hourly general supply in kWh for 13,735 customers, resulting in 344,518,791 observations in total.<!--In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual.
 We do not learn about these repetitive behaviors from the linear view because too many measurements all squeezed in that representation.
-Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior.
-The raw data for these consumers is of unequal length, with varying start and finish dates.
-Because our proposed methods evaluate probability distributions rather than raw data, neither of these data features would pose any threat to our methodology unless they contained any structure or systematic patterns.
-Additionally, there were missing values in the database but further investigation revealed that there is no structure in the missingness (see Supplementary paper for raw data features and missingness).
-The study begins by subsetting a data set along all dimensions of interest using data filtering and prototyping.
-By grouping the prototypes using our methods and assessing their meaning, the study hopes to unravel some of the heterogeneities observed in energy usage data.
-Because our application does not employ additional customer data, we cannot explain why consumption varies, but rather try to identify how it varies.
-
-_Data filtering and variable selection_
-
-- Choose a smaller subset of randomly selected $600$ customers with no implicit missing values for 2013.
-
-- Obtain $wpd$ for all cyclic granularities considered for these customers.
-It was found that `hod` (hour-of-day), `moy` (month-of-year) and `wkndwday` (weekend/weekday) are coming out to be significant for most customers.
+Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior.--> The raw data for these consumers is of unequal length, with varying starting and end dates.
+Additionally, there were missing values in many series.
+(The supplementary material contains details from checking for systematic missingness.) Because our proposed methods evaluate probability distributions rather than raw data, these data issues are not problematic, unless there is any systematic structure related to granularities.
+
+Huge data sets present more complications for clustering.
+Clustering algorithms work well when there are well-separated clusters, with no nuisance variables or nuisance observations.
+When converting a series to granularities, many variables (each level of a granularity) are generated, possibly creating a slew of nuisance variables.
+Some customers may have a mix of energy use patterns, which could be considered nuisance observations located between major clusters.
+For this reason, we have chosen to select a small group of customers with relatively distinct and different patterns in order to illustrate the clustering more simply.
+Figure \ref{fig:hod-ind-group-png} shows the distribution across *hod*, *moy* and *wnwd* for the set of $24$ customers used to illustrate clustering.
+The customers are displayed in two columns of $12$ for space reasons.
+Each row, of each column, represents the profile of a single customer across different variables.
+Each customer is associated with an identifier of the form [$a$-$b$], where $a \in \{1, 2, \dots, 24\}$ represents the customer-prototype id and $b \in \{1, 2, \dots, 5\}$ indicates the label of the prototype in which a customer was placed.
+This is often a good approach to tackling a big analysis task, to start with a simpler task.
+The approach, however, is applicable to all customers.
+
+As a result, we dissect the larger problem and test our solutions on a small sample of prototype customers.
+To do this, data is first filtered to generate a small sample, and then significant cyclic granularities (variables) for them are chosen (as described in Section \ref{sec:datafilter}).
+The sample set is subsequently examined along all dimensions of interest, to ensure that they reveal some patterns across at least one specified variable (as described in Section \ref{sec:prototype}).
+<!--By grouping the prototypes using our methods in Section \ref{sec:clustering} and assessing their meaning, the study hopes to unravel some of the heterogeneities observed in energy usage data.--> Because the data does not contain additional customer characteristics, we cannot explain why consumption varies, but can only identify how it varies.
+
+### Data filtering and variable selection {#sec:datafilter}
+
+The steps for customer filtering and variable selection were:
+
+1. Choose a smaller subset of randomly selected $600$ customers with no implicit missing values for 2013.
+2.
+Obtain $\wpd$ for all cyclic granularities considered for these customers.
+It was found that *hod* (hour-of-day), *moy* (month-of-year) and *wnwd* (weekend/weekday) are significant for most customers.
 We use these three granularities while clustering.
-
-- Remove customers whose data for an entire category of `hod`, `moy` or `wnwd` is empty.
+3.
+Remove customers whose data for an entire category of *hod*, *moy* or *wnwd* is empty.
 For example, a customer who does not have data for an entire month is excluded because their monthly behavior cannot be analyzed.
-
-- Remove customers whose energy consumption is 0 in all deciles.
+4.
+Remove customers whose energy consumption is 0 in all deciles.
 These are the clients whose consumption is likely to remain essentially flat and with no intriguing repeated patterns that we are interested in studying.
 
-_Prototype selection_
+### Selecting prototypes {#sec:prototype}
 
 <!-- Why instance selection -->
-Supervised learning uses a training set of known information to categorize new events through instance selection.
-Instance selection (@olvera2010review) is a method of rejecting instances that are not helpful for classification.
-This is analogous to subsampling the population along all dimensions of interest such that the sampled data represents the primary features of the underlying distribution.
-Instance selection in unsupervised learning has received little attention in the literature, yet it could be a useful tool for evaluating model or method performance.
-There are several ways to approach prototype selection.
-Following @Fan2021-bq's idea of picking related examples (neighbors) for each instance (anchor), we can first use any dimensionality reduction techniques like MDS or PCA to project the data into a 2D space.
-Then pick a few "anchor" customers who are far apart in 2D space and pick a few neighbors for each.
-Unfortunately, this does not ensure that consumers with significant patterns across all variables are chosen.
-Tours can reveal variable separation that was hidden in a single variable display better than static projections.
-Hence we perform a linked tour with a t-SNE layout using the R package `liminal` (@R-liminal) to identify customers who are more likely to have distinct patterns across the variables studied.
-(Refer to Supplementary article for further details).
-Figure \ref{fig:hod-ind-group} shows the distribution across `hod` (a), `moy`(b) and `wnwd` (c) for the set of chosen $24$ customers that were chosen.
-Few of these customers have similar distribution across `moy` and some are similar in their `hod` distribution.
+
+It is common to filter data prior to fitting a supervised classification model using instance selection [@olvera2010review] which removes observations that might impede the model building.
+For clustering, this is analogous to identifying and removing nuisance observations.
+Prototype selection is more severe than instance selection, because only a handful of cases is selected.
+@Cutler1994-hm proposed a method called archetypal analysis which has inspired this approach but the procedure we have used follows @Fan2021-bq.
+First, dimension reduction such as t-SNE, MDS or PCA is used to project the data into a 2D space.
+Second, a few "anchor" customers far apart in 2D space are selected.
+Additional close neighbors to the anchors are selected.
+To check the selections relative to the full set of variables, we used a tour linked to a t-SNE layout using the R package `liminal` [@R-liminal].
+This ensured that the final sample of clustered customers were also far apart in the high-dimensional space.
+(See the supplementary materials for further details.)<!--Few of these customers have similar distribution across *moy* and some are similar in their *hod* distribution.
+-->
 
@@
 
-```{r hod-moy-wkndwday plots, fig.cap="The distribution across moy, hod and wkndwday for the selected designs.
-Few are similar in their hod pattern, while others are similar in moy behavior.
-Some customers have distinct behavior as compared to all other customers.For example, although patterns across wkndwday do not look distinctly different for most households, there is one household for whom weekend behavior is standing out from the rest."}
-```
-
-### Clustering
-
-The $24$ prototypes <!--and the data subset consisting of $353$ customers--> are clustered using the methodology described in Section \ref{sec:methodology} and results are reported below.
-In the following plots, the median is shown by a line, and the shaded region shows the area between the  $25^{th}$ and $75^{th}$.
-All customers with the same color represent the same clustered groups.
-Groups by JS-based distances and wpd-based distances are colored differently as they represent different groupings.
-The plotting scales are not displayed since we want to emphasize comparable shapes rather than scales.
-The idea is that a customer in a cluster may have low total energy usage, but their behavior may be quite similar to a customer with high usage with respect to distributional pattern or significance across cyclic granularities.
-
-#### JS-based distances
-
-For clustering based on JS-based distances, we chose the optimal number of clusters using (@Hennig2014-ah) as $5$.
-The groupings are shown in Figure \ref{fig:hod-ind-group}.
-Customers with the same color represent the same clustered groups.
-Our methodology is useful for grouping similar distributions over `hod` and `moy` and they are placed closely for easy comparison.
-Few groups have mixed patterns across `hod` and `moy`, but few have all customers in the group having a similar profile.
-Figure \ref{fig:combined-groups-js} shows the summarized distributions across $5$ groups and assists us in characterizing each cluster.
-It shows Groups $2$ and $4$ have `hod` pattern with a typical morning and evening peak, whereas groups $1$, $3$, and $5$ show a `moy` pattern with higher usage in winter months.
-Differences in Weekend/Weekday between groups are not discernible, implying that it may not be a relevant variable in distinguishing various clusters unless maybe conditioned by `moy` or `hod`.
-It may be interesting to compare these two plots to verify if the summarized distributions across groups correctly characterised the groupings.
-If it has, then the majority of the group's members should share a similar profile.
+<!-- # ```{r hod-moy-wkndwday plots, fig.cap="The distribution across moy, hod and wkndwday for the selected designs.
+Few are similar in their hod pattern, while others are similar in *moy* behavior.
+Some customers have distinct behavior as compared to all other customers.For example, although patterns across wkndwday do not look distinctly different for most customers, there is one household for whom weekend behavior is standing out from the rest."} -->
+<!-- # ``` -->
+
+### Clustering results {#sec:clustering}
+
+Clustering of the $24$ prototypes <!--and the data subset consisting of $353$ customers--> was conducted with all three distances, JS-NQT, JS-RS and WPD, and is summarized in Figures \ref{fig:opt-cluster-tsne-jsd}, \ref{fig:groups-4and5} and \ref{fig:summary-plot-wpd}.
+The t-SNE visualization suggests that there are four well-separated clusters.
+It is possible that because the representation is only 2D, the fifth group from the original prototype selection is distinctly different in high dimensions.
+The *sindex* plots for the three methods indicate some disagreement: JS-NQT suggests $3$, JS-RS suggests $2$ or $5$ and WPD suggests $3$ or $5$.
+JS-RS would appear to match the original prototypes with the five cluster solution, but it actually differs.
+Even though the *sindex* for JS-NQT suggests three clusters, the five cluster solution more closely matches the original prototypes.
+The WPD clustering provides a different grouping of the customers, and even though it disagrees with the original prototypes it is a useful grouping.<!--While plotting the clusters, RS-based scaling is applied to each customer, so that no customer dominates the shape of the summarized clusters because of magnitude.
+The plotting scales are not displayed since we want to emphasize comparable patterns rather than scales.
+The idea is that a customer in a cluster may have low total energy usage, but their behavior may be quite similar to a customer with high usage with respect to distributional pattern or significance across cyclic granularities.-->
+
+<!--### JS-NQT and JS-RS-->
+
+<!--Figure \ref{fig:opt-cluster-tsne-jsd} The *sindex* is plotted against the number of clusters for clustering based on JS-based (JS-NQT) distances in Figure \ref{fig:opt-cluster-tsne-jsd}.
+The *sindex* falls from around $0.56$ when $k=3$ to $0.54$ when $k$ increases.
+In our study, we use $k=4, 5$.--> Figure \ref{fig:groups-4and5} displays the summarized distributions across $4$ and $5$ clusters from JS-NQT clustering in (a) and (b) respectively, and helps to characterize each cluster.
+In the quantile plots the line represents the median, and the region shows the area between the $25^{th}$ and $75^{th}$ percentiles.
+The only difference between the four and five cluster solution is that A-4 divides further into B-4 and B-5.
+This additional division makes a clearer clustering, because it resolves the heterogeneity in *moy* creating a group (B-5, customers 1-3) which has a winter peak in usage, and a group (B-4, customers 16-20) which has a start of the year peak in usage.
+B-2 (customers 4-9) and B-1 (customers 21-24) have distinctive *hod* patterns but are both heterogeneous in *moy* and *wnwd*.
+B-3 (customers 10-15) has peak usage at the end of the year, but is heterogeneous on *hod* and *wnwd*.
+This clustering almost agrees with the clusters visible in the t-SNE plot.
+
+<!-- A-4 is subdivided into B-4 and B-5, each of which has a distinct shape across *moy* and *wnwd*.
+B-4 (id: 16-20) is distinguished by higher energy consumption in the first few months of the year and greater variability in weekend usage.
+B-5 (id: 1-3) has higher consumption in the middle of the year (winter months) and a similar weekday-weekend inter-quartile range.
+When $k = 4$ is used, these two groups merge to form A-4, which has a *moy* profile of higher usage in both the beginning and middle of the year, which is not representative of the individuals.
+It may be worthwhile to compare Figures \ref{fig:opt-cluster-tsne-jsd} and \ref{fig:hod-ind-group-png} to see if the summarized distributions across groups accurately characterized the groupings.
+If it has, then the majority of the members of the group should have a similar profile.-->
+
+<!-- Our methodology is useful for grouping similar distributions over *hod* and *moy* and they are placed closely for easy comparison.
+Few groups have mixed patterns across *hod* and *moy*, but few have all customers in the group having a similar profile.-->
 
@@
 
-```{r hod-ind-group, fig.cap = "The distribution of selected customers over hod (a), moy (b), and wnwd (c) for the $24$ selected customers is shown.
-In each case, the same colour denotes the same group using js-based clustering and are placed together to facilitate comparison.
-Individuals with similar distributional differences across categories of `hod` or `moy` are grouped together, with some exceptions.
-" }
+```{r data-pick-robust}
+
+```
+
+```{r hod-moy-wkndwday-data}
+
+```
+
+```{r opt-clusters-jsd}
+
+```
+
+```{r data-pick-wpd}
+```
+
+```{r opt_clusters-wpd}
+```
+
+```{r summary-group-wpd}
+```
+
+```{r tsne-fit}
+```
+
+```{r tsne-df}
+```
+
+```{r tsne-xy}
+
+```
+
+```{r all-data-rs}
+
+```
+
+```{r clustering-rs}
+
+```
+
+```{r opt-clusters-jsd-rs}
+
+```
+
+```{r opt-clusters-all}
+
+```
+
+```{r hod-ind-group}
+```
+
+(ref:hod-ind-group-png) The distribution of electricity demand across individual customers over three granularities *hod*, *moy*, and *wnwd* are shown for the $24$ selected customers using quantile and box plots.
+They are split into batches of $12$ in (a) and (b), with each row in (a) or (b) representing a customer.
+The number indicates a unique customer id and a prototype id.
+
+```{r hod-ind-group-png, fig.cap = "(ref:hod-ind-group-png)"}
+```
+
+(ref:opt-cluster-tsne-jsd) Clustering summaries: (a) t-SNE computed on the $24$ selected customers, and (b) separation index (*sindex*) for $2$-$10$ clusters using JS-NQT, JS-RS and WPD.
+Various choices in number of clusters would be recommended.
+Four clusters are visible in t-SNE, although it might hide a fifth cluster because dimension reduction to 2D may be insufficient to see the difference.
+JS-NQT suggests $3$, JS-RS suggests $2$ or $5$ and WPD suggests $3$ or $5$.
+
+```{r opt-cluster-tsne-jsd, fig.cap = "(ref:opt-cluster-tsne-jsd)"}
+#include_graphics("img/opt_clust_parenthesis.png")
 ```
@@
 
-```{r data-heatmap-hod-group}
-```
-
-```{r data-heatmap-moy-group}
+```{r data-heatmap-hod-group-new}
+```
+
+```{r data-heatmap-moy-group-new}
 
@@
 
-```{r combined-groups-js, fig.cap = "The distribution of electricity demand over hod (a), moy (b), and wnwd*hod (c).
-Groups 2 and 5 appear to have a hod pattern among its members, whereas groups 1, 3, and 5 appear to have a moy pattern.
-Weekend/weekday differences across groups are not discernible, showing that it is not a critical variable for clustering on its own.
-It is probable that the distribution over Weekend/Weekday will differ conditional on the categories of hod or moy.
-(c) demonstrates that the distribution is not different, except for the tails.
-It is useful to compare the summarised distributions of groups to those of individuals in order to establish that the majority of individuals in the group share the same characteristics."}
-```
-
-#### wpd-based distances
-
-We chose the optimal number of clusters using (@Hennig2014-ah) as $3$.
-A parallel coordinate plot with the three significant cyclic granularities is used to characterise the groups here.
+```{r combined-groups-js, fig.cap = "The distribution of electricity demand across clusters over *hod* (a), *moy* (b), and *wnwd* (c) are shown through quantile plots.
+Group 1 (customers 1-3), Group 2 (customers 4-9), Group 3 (customers 10-15), Group 4 (customers 16-20), Group 5 (customers 21-24).
+Groups 2 and 5 appear to have a *hod* pattern among its members (morning and evening peak), whereas groups 1, 3, and 5 appear to have a *moy* pattern (higher usage in mid-year).
+With their conditional dependencies on *hod* or *moy* categories, *wnwd* differences may not be discernible on their own.
+(c) demonstrates that the distribution of *hod* conditional on *wnwd* is also not different, except for the tails.
+This might be because most of these chosen customers do not have a *wnwd* pattern.
+It is useful to compare the raw distributions with the summarized distributions to establish that the majority of individuals in the group share the same characteristics.
+To facilitate this comparison, the individual distributions are sorted by their JS-based groupings.", eval = FALSE}
+
+```
+
+<!-- The widest within-cluster gap (widest_gap) is high till $k=5$ and then have a gradual decrease. Number of -->
+
+```{r opt-clusters, fig.cap=" Different cluster statistics plotted as a function of number of clusters for JS-based (a) and wpd-based (b) approaches.
+avg within is not too useful for revealing any sharp change in average distance between clusters with increasing number of clusters.
+For (a), sindex sees a sharp fall at $k = 4$ and then do not decrease until $k=16$ and the widest gap is high till $k=5$.
+For (b), sindex sees a fall at $k=3$ and then at $k=7$ and widest gap has a sharp fall at $k=6$.", eval = FALSE}
+#knitr::include_graphics("img/opt_clusters.png")
+```
+
+<!-- GroupA and Group B may represent different set of customers as they are groupings based on different number of clusters. -->
+
+(ref:groups-4and5) Summary plots for four and five clusters from JS-NQT showing the distribution of electricity demand combined for all members over *hod*, *moy*, and *wnwd*.
+Groups A-1 (customers 21-24), A-2 (customers 4-9), and A-3 (customers 10-15) profiles correspond to Groups B-1, B-2, and B-3, respectively.
+Cluster A-4 splits into B-4 (customers 16-20) and B-5 (customers 1-3) to produce the five clusters, which better resolves the *moy* distribution.
+
+```{r groups-4and5, fig.cap = "(ref:groups-4and5)", fig.height=5}
+
+```
+
+(ref:summary-plot-wpd) Summary plots for three (a) and five (b) clusters from WPD approach showing the $\wpd$ values of each customers across *hod*, *moy*, and *wnwd* through a parallel coordinate plot.
+P-1 and P-2 groups correspond to Q-1 and Q-2, respectively.
+Cluster P-3 is subdivided into Q-3, Q-4, and Q-5.
+P-1 (customers 16, 18) is distinguished by high $\wpd$ on *wnwd* values.
+P-2 has lower $\wpd$ than *moy* and *wnwd* for *hod*.
+P-3 operates in the opposite way as P-2, with larger $\wpd$ for *hod* in comparison to *moy* and *wnwd*.
+For the $5$ cluster solution, this group is divided into Q-3, Q-4, Q-5, which are distinct due to their different relative significance of *moy* and *wnwd*.
+
+```{r summary-plot-wpd, fig.cap = "(ref:summary-plot-wpd)", fig.height=6}
+```
+
+```{r summary-table-wpd}
+
+```
+
+<!--### WPD-->
+
+<!-- In Figure \ref{fig:opt-cluster-tsne-jsd}, the *sindex* is plotted against the number of clusters for clustering based on wpd-based distances.
+The scales are very different from those used in JS-based clustering.
+The *sindex* falls from around $20$ when $k=3$ to $19$ when $k=4$ and further decreases to $17.5$ with $k>7$.
+Thus, there are few choices available for the number of clusters.
+Figure \ref{fig:summary-plot-wpd} whos the wpd values for the three different granularities of the 24 customers, coloured by the 6 cluster solution.
+XXX REWORK THIS WHEN NEW CLUSTER RESULTS ARE SHOWN<!--The groupings with $k=4$ is shown through a parallel coordinate plot with the three significant cyclic granularities.
 The variables are sorted according to their separation across classes (rather than their overall variation between classes).
-This means that $moy$ is the most important variable in distinguishing the groups followed by $hod$ and $wnwd$.
-There is only one customer who has significant $wpd$ across $wnwd$ and stands out from the rest of the customers.
-Group 3 has a higher wpd for hod than moy or wkndwday.
-Group 2 has the most distinct pattern across moy.
-Group 1 is a mixed group that has strong patterns on at least one of the three variables.
-The findings vary from js-based clustering, yet it is a helpful grouping.
-
-```{r data-pick-wpd}
-```
-
-```{r parcoord, fig.cap="Each of the 24 customers is represented by a parallel coordinate plot (a) with three wpd-based groupings.
-The plot shows that moy is the most important variable in identifying clusters, whereas wkdn-wday is the least significant and has the least fluctuation.
-One particular customer with high wpd across wknwday stands out in this display.
-Group 3 has a higher wpd for hod than moy or wkndwday.
-Group 2 has most discernible pattern across moy.
-Group 1 is a mixed group with strong patterns on atleast one of the three variables.
-All of these could be observed from the plot or the table (b) which shows median wpd values for each group.", message=FALSE, warning=FALSE}
-
-```
-
-Things become far more complicated when we consider a larger data set with more uncertainty, as they do with any clustering problem.
+This means that *moy* is the most important variable in distinguishing the groups, followed by *hod* and *wnwd*.
+The two customers (id: 16 and 18) that have significant differences between *wnwd* categories (in Figure \ref{fig:hod-ind-group-png}) stand out from the rest of the customers.
+Group-4 has a higher $\wpd$ for *hod* than *moy* or *wnwd*.
+Group-2 and Group-3 have the most distinct patterns across *moy*.
+Group-1 is a mixed group that has strong patterns on at least one of the three variables.
+When $k=3$ is used, Group-2 and Group-3 are merged because their patterns relative to *moy*, *hod* and *wnwd* are the same.
+The findings vary from JS-based clustering, yet it is a helpful grouping.-->
+
+Figure \ref{fig:summary-plot-wpd} shows the $\wpd$ values of the $24$ customers over *hod*, *moy*, and *wnwd* , colored by $3$ (a) and $5$ (b) clusters from WPD clustering using a parallel coordinate plot.
+The variables ($\wpd$ for different granularities) are
+standardized prior to clustering using WPD.
+In the display, the variables are sorted according to their separation across groups.
+This means that *wnwd* is the most important variable in distinguishing the groups, followed by *hod* and *moy* for both (a) and (b).
+Groups P-1 and P-2 correspond to Q-1 and Q-2 respectively.
+Cluster P-3 splits into Q-3, Q-4 and Q-5.
+Customers 16 and 18 are characterized by unusual high values of $\wpd$ on *wnwd* compared to the rest of the customers and hence form a group.
+This could again be verified from Figure \ref{fig:hod-ind-group-png}, where these were the only two customers with a difference in their *wnwd* behavior.
+They are represented by P-1.
+P-2 has lower $\wpd$ for *hod* than *moy* and *wnwd*.
+P-3 behaves opposite to P-2 with higher $\wpd$ for *hod* compared to *moy* and *wnwd*.
+These are the customers who have some significant pattern across *hod* and this can again be validated by looking at Figure \ref{fig:hod-ind-group-png}.
+For a $5$ cluster solution, this group gets split into Q-3, Q-4 and Q-5 characterizing different relative significance of *moy* and *wnwd*.
+For example, Q-4 and Q-5 have almost no pattern across *moy*, but Q-3 has a *moy* pattern and thus it is reasonable to split them.
+The patterns could be different, but they are significant.
+Q-4 and Q-5 are separated because of their different significance of *hod*.
+All of these can also be verified from Table \ref{tab:summary-table-wpd} which shows the cluster summaries with members and median values of $\wpd$ for the three variables.
+
+<!-- *hod* has very low median values of $\wpd$ for the members in the cluster, implying at least half of these customers do not have an interesting *hod* pattern. -->
+
+```{r wpd-clustering}
+
+```
+
+<!-- # ```{r summary-plot-wpd, fig.cap="Summary of the four cluster results from WPD, shown as a parallel coordinate plot and table.
+XXX TO BE REVISED WHEN DIFFERENT CLUSTER SOLUTION IS SHOWN", message=FALSE, warning=FALSE, fig.height=6} -->
+<!-- # #include_graphics("img/summary-plot-wpd.png") -->
+<!-- # ``` -->
+
+In summary, none of the methods captured the five original prototypes exactly.
+JS-NQT was almost identical, but WPD produced quite a different grouping.
+This is quite a reasonable result and illustrates both the difficulties of clustering to obtain a particularly expected grouping and the ability to learn unexpected patterns in the data.
+It is possible that the JS-based distances were distracted by the presence of nuisance variables, levels of the granularities that do not contribute to clustering.
+This would also be supported by the results of the validation study, where clustering was less effective in S2 and S3, where only some granularities had differences between levels.
+Clustering using WPD is expected to produce quite different results because it will group only by overall value of a granularity, not a particular pattern.
+A cluster summary like Figure \ref{fig:groups-4and5} is not possible because there may be different but equally interesting patterns (e.g.
+high evening *hod* and high daytime *hod*) in the same cluster.
+Simply it provides information that across a collection of customers this specific cluster has interesting patterns in a granularity (e.g.
+*hod*).
+One would need to post-process the cluster to separate specific patterns.
+
+<!--Things become far more complicated when we consider a larger data set with more uncertainty, as they do with any clustering problem.
 Summarizing distributions across clusters with varied or outlying customers can result in a shape that does not represent the group.
 Furthermore, combining heterogeneous customers may result in similar-looking final clusters that are not effective for visually differentiating them.
-It is also worth noting that the Weekend/Weekday behavior in the given case does not characterize any cluster.
+It is also worth noting that the weekend/weekday behavior in the given case does not characterize most selected customers.
 This, however, will not be true for all of the customers in the data set.
 If more extensive prototype selection is used, resulting in more comprehensive prototypes in the data set, this method might be used to classify the entire data set into these prototype behaviors.
-However, the goal of this section was to have a few customers that have significant patterns over one or more cyclic granularities, apply our methodology to cluster them, and demonstrate that the method produces useful clusters.
-
-## Conclusion and future work {#sec:conclusion}
+However, the goal of this section was to have a few customers that have significant patterns over one or more cyclic granularities, apply our methodology to cluster them, and demonstrate that the method produces useful clusters.-->
+
+## Discussion {#sec:discussion-gracsr}
 
@@
 For the smart meter example, the distribution for a customer moving to a different house or changing electrical equipment can change drastically.
-Our current approach can not detect these dynamic changes.
+Our current approach cannot detect these dynamic changes.
 Thirdly, it is possible that for a few customers, data for some categories from the list of considered significant granularities are missing.
-In our application, we have removed those customers and done the analysis but the metrics used should be able to incorporate those customers in the clustering by handling their missing categories.
-Finally, $wpd$ is computationally heavy even under parallel computation.
+In our application, we have removed those customers and done the analysis but the metrics used should be able to incorporate those customers with such structured missingness.
+Finally, $\wpd$ is computationally heavy even under parallel computation.
 Future work can make the computations more efficient so that they are easily scalable to a large number of customers.
@@
 The Monash eResearch Centre and eSolutions-Study Support Services supported this research in part through the resource usage of the MonARCH HPC Cluster.
-The Github repository, [github.com/Sayani07/paper-gracsr](https://github.com/Sayani07/paper-gracsr), contains all materials required to reproduce this article and the code is also available online in the supplemental materials.
-This article was created with R [@R-language], `knitr` [@knitr2015], [@R-knitr] and `rmarkdown` [@rmarkdown], [@R-rmarkdown].
-Graphics are produced with `ggplot2` [@Wickham2009pk] and GGally [@R-GGally].
-
-## Supplementary Materials {- #supplement-gracsr}
-
-**Data and scripts:** Data sets and R code to reproduce all figures in this article (main.R).
+The Github repository, [github.com/Sayani07/paper-gracsr](https://github.com/Sayani07/paper-gracsr), contains all materials required to reproduce this article and the code is also available online in the supplementary materials.
+This article was created with R [@R-language], `knitr` [@knitr; @R-knitr] and `rmarkdown` [@rmarkdown; @R-rmarkdown].
+Graphics are produced with `ggplot2` [@Wickham2009pk] and `GGally` [@R-GGally].
+
+## Supplementary materials {- #supplement-gracsr}
+
+**Data and scripts:** Data sets and R codes to reproduce all figures in this article ([validation.R](https://github.com/Sayani07/paper-gracsR/blob/master/script/validation.R), [application.R](https://github.com/Sayani07/paper-gracsR/blob/master/script/application.R))
 
 \noindent
-**Supplementary paper:** Additional tables, graphics and and R code to reproduce it (paper-supplementary.pdf, paper-supplementary.Rmd)
+**Supplementary paper:** Additional graphics and R codes to reproduce it  
+([paper-supplementary.pdf](https://github.com/Sayani07/paper-gracsR/blob/master/paper-supplementary.pdf), [paper-supplementary.Rmd](https://github.com/Sayani07/paper-gracsR/blob/master/paper-supplementary.Rmd), [supplementary.R](https://github.com/Sayani07/paper-gracsR/blob/master/script/supplementary.R))
 
 \noindent
-**R-package:** To implement the ideas provided in this research, the open-source R package 'gracsr' is available on Github (https://github.com/Sayani07/gracsr).
+**R-package:** To implement the ideas provided in this research, the open-source R package `gracsr` is available on Github (https://github.com/Sayani07/gracsr).
 
@@
 
-# Conclusion and future plans {#ch:conclusion}
-
-In this thesis, I present a systematic approach for visualizing and analyzing distributions of large temporal data by leveraging time characteristics.
+# Conclusion {#ch-conclusion}
+
+This thesis presents methods for visualizing and analyzing distributions of large temporal data by deconstructing time into temporal granularities.
+This chapter summarizes the thesis content, outlines the original contributions, software development, and possible directions for future research.
+
+## Original contributions
+
+<!-- In this thesis, I present a systematic approach for visualizing and analyzing distributions of large temporal data by leveraging time characteristics.
 The building blocks of this framework are presented in each chapter.
-Chapter \@ref(ch:gravitas) contains tools for computing all cyclic granularities, as well as a recommendation system for selecting pairs of granularities that can be effectively analyzed together.
+Chapter \@ref(ch-gravitas) contains tools for computing all cyclic granularities, as well as a recommendation system for selecting pairs of granularities that can be effectively analyzed together.
 These temporal granularities can be used to generate data visualizations to look for patterns, associations, and anomalies.
 But it is difficult to decide which of the various granularities to display when there are many options.
-Chapter \@ref(ch:hakear) presents a methodology for selecting displays that are interesting, such that differences between displayed distributions are greatest.
+Chapter \@ref(ch-hakear) presents a methodology for selecting displays that are interesting, such that differences between displayed distributions are greatest.
 The methodology also ranks the displays in order of priority for capturing the most variation.
-Both \@ref(ch:gravitas) and \@ref(ch:hakear) are used for studying patterns in individual time series or comparing a few time series together.
-This is extended in Chapter @ref(ch:gracsr) to allow for the exploration of distributions for multiple time series at the same time using unsupervised clustering.
+The ideas in Chapters \@ref(ch-gravitas) and \@ref(ch-hakear) can be used for studying patterns in individual time series or comparing a few time series together.
+This is extended in Chapter \@ref(ch-gracsr) to allow for the exploration of distributions for multiple time series at the same time using unsupervised clustering.
 The clustering methodology produces small groups of time series with similar distributions over multiple granularities.
 This technique is more robust to noisy, patchy, and uneven length time series because it makes use of probability distributions.
+-->
+
+Exploratory time series analysis entails numerous iterations of identifying and summarizing temporal dependencies.
+It is common practice to divide time into years, months, weeks, days, and so on in order to make inferences at both finer and coarser scales.
+In the literature, the formalization of these temporal deconstructions (granularities) is limited to linear time granularities such as hours, days, weeks, and months that respect the linear progression of time and are non-repeating.
+Cyclic granularities that are repeating in nature are useful for finding patterns in temporal data.
+They can be circular, quasi-circular, or aperiodic in nature.
+Hour-of-the-day and day-of-the-week are examples of circular granularities; the day-of-the-month is an example of a quasi-circular granularity; and public holidays and school holidays are examples of aperiodic granularities.
+Additionally, time deconstructions can be based on a time hierarchy.
+Thus, single-order-up granularities such as second of minute or multiple-order-up granularities such as second of hour can be envisioned.
+The definitions and rules defined in the literature for linear granularities are insufficient for describing various types of cyclic granularities.
+
+Chapter \@ref(ch-gravitas) provides a formal characterization of cyclic granularities as well as tools for classifying and computing potential cyclic granularities from an ordered temporal index.
+It also allows for the manipulation of single- and multiple-order-up time granularities via cyclic calendar algebra.
+The approach is generalizable to non-temporal hierarchical granularities with an ordered index.
+Visualizing probability distributions conditional on one or more cyclic granularities is a powerful exploration tool.
+However, there may be too many cyclic granularities to look at manually for comprehensive exploration, and not all pairs of granularities can be effectively explored together.
+Chapter \@ref(ch-gravitas) also provides a recommendation on whether a pair of granularities can be meaningfully plotted or analyzed together (a "harmony") or when they cannot (a "clash" or "near-clash").
+
+Cyclic granularities could be used to create a wide range of displays.
+And, when there are numerous granularities to choose from, deciding which one to display can be difficult.
+Moreover, only a few of them may be useful in revealing major patterns.
+In Chapter \@ref(ch-hakear), the search for informative granularities is facilitated by selecting "significant" granularities.
+A cyclic granularity is referred to as "significant" if there is a significant distributional difference of the measured variable between different categories.
+Chapter \@ref(ch-hakear) defines a distance measure to quantify these distributional differences.
+A higher value of the distance measure for a cyclic granularity or harmony implies that they could be interesting for further investigation, whereas a low value indicates that nothing noteworthy is unfolding.
+A threshold and, consequently, a selection criterion are chosen using a permutation test such that cyclic granularities with significant values of the distance measure are selected.
+In addition, the distance metric has been appropriately adjusted, allowing it to be compared not only across cyclic granularities with different numbers of categories but also across a set of time series.
+As a result, it can also be used to rank the displays according to their ability to capture the greatest amount of variation across one or multiple time series.
+
+The ideas in Chapters \@ref(ch-gravitas) and \@ref(ch-hakear) can be used for studying patterns in individual time series or comparing a few time series together.
+This is extended in Chapter \@ref(ch-gracsr) to allow for the exploration of distributions for multiple time series at the same time using unsupervised clustering.
+In the time series clustering literature, probability distributions across cyclic granularities have not been considered in determining similarity.
+However, such a similarity measure can be useful for characterizing the inherent temporal data structure of long, unequal-length time series in a way that is resistant to missing or noisy data while allowing for the detection of similar repeated patterns.
+Chapter \@ref(ch-gracsr) proposes two approaches for calculating distances between time series based on probability distributions across cyclic granularities.
+The first approach considers two time series to be similar if the distributions of each category of one or more cyclic granularities are similar.
+The second approach considers two time series to be similar if they have a similar significance of patterns across different granularities.
+A similar significance does not imply a similar pattern, which is where this technique varies from the former.
+When the distances from these approaches are fed into a hierarchical clustering algorithm, they yield small groups of time series with similar distributions or significance over multiple granularities.
+Our method is capable of producing useful clusters for both approaches, as demonstrated by testing on a range of validation data designs and a sample of residential smart meter consumers.
 
@@
 
-This thesis focuses on integrating research approaches into open source R packages for reproducibility and ease of use in other applications.
-So a significant amount of work has been devoted to the development of R packages  **gravitas**, **hakear**, and **gracsr**, each of which correspond to a chapter presented in this thesis.
+This thesis focuses on translating research approaches into open source R packages for reproducibility and ease of use in other applications.
+So a significant amount of work has been devoted to the development of R packages  `gravitas`, `hakear`, and `gracsr`, each of which corresponds to a chapter presented in this thesis.
 
@@
 
-The **gravitas** package provides very general tools to compute and manipulate cyclic granularities and generate plots displaying distributions conditional on those granularities.
-The functions `search_gran()`, `create_gran()`, `harmony()`, `gran_advice()` and `prob_plot()` provides the entire workflow for an analyst to systematically explore large quantities of temporal data across different harmonies (pairs of granularities that can be analyzed together.) This package was developed as part of my internship at Google Summer of Code, 2019.
+The `gravitas` package provides very general tools to compute and manipulate cyclic granularities and generate plots displaying distributions conditional on those granularities.
+The functions `search_gran()`, `create_gran()`, `harmony()`, `gran_advice()` and `prob_plot()` provides the entire workflow for an analyst to systematically explore large quantities of temporal data across different harmonies (pairs of granularities that can be analyzed together).
+This package was developed as part of my internship at Google Summer of Code, 2019.
 It has been on CRAN since January 2020.
@@
 
-The R package hakear (https://github.com/Sayani07/hakear) implements ideas presented in Chapter \@ref(ch:hakear).
-The function wpd() computes the weighted pairwise distances ($wpd$) for each cyclic granularity or pair of granularities, and `select_harmonies()` selects the ones with significant patterns and ranks them from highest to lowest $wpd$.
-These selected harmonies can be plotted using package `gravitas` for potentially interesting displays.
-This package is reliant on parallel processing using multiple multi-core computers for faster computation of $wpd$.
+The R package `hakear` (https://github.com/Sayani07/hakear) provides tools for selecting and sorting significant cyclic granularities.
+The function `wpd()` computes the weighted pairwise distances ($\wpd$) for each cyclic granularity or pair of granularities, and the function `select_harmonies()` chooses those with significant patterns and ranks them from highest to lowest $\wpd$.
+This package is reliant on parallel processing using multiple multi-core computers for faster computation of $\wpd$.
+The selected harmonies can be plotted using package `gravitas` for potentially interesting displays.
+Currently, `hakear` implements ideas presented in Chapter \@ref(ch-hakear), but it will be integrated with `gravitas` in the future to explore distributions of a smaller number of time series.
 
@@
 
-The open-source R package `gracsr` is available on Github (https://github.com/Sayani07/gracsr) to implement ideas presented in Chapter \@ref(ch:gracsr).
-The package contains functions for implementing the complete clustering methodology with choices of scaling and distance metrics discussed in the chapter.
-It has received a grant (AUD 3000) as part of the ACEMS Business Analytics Prize towards polishing the functions and preparing it for CRAN.
-
-## Future work
-
-### Putting all functionalities on CRAN
-
-I want to combine 'hakear' and 'gravitas' into a single R package for systematically exploring distributions of a smaller number of time series, whilst 'gracsr' would offer the clustering framework for systematically exploring a large number of time series.
-I intend to run it via https://ropensci.org/software-review/ to further refine them for more exposure and efficiency.
-
-### Scaling up the methods for multivariate time series data
-
+The R package `gracsr` (https://github.com/Sayani07/gracsr) has functions for exploring a large number of time series using the clustering methodology described in Chapter \@ref(ch-gracsr).
+The workflow begins with the function `scale_gran()`, which may be used to scale individual series using NQT/RS.
+The distances for the JS and WPD approaches are computed in the second phase of the workflow using `dist_gran()`/`dist_ wpd()`.
+The distances can then be used to do clustering with `clust_gran()`.
+The package has received a grant (AUD 3000) as part of the ACEMS Business Analytics Prize towards polishing the functions and preparing it for CRAN.
+
+### Computational resources
+Simulation studies were carried out to study the behavior of $\wpd$, build the normalization method as well as compare and evaluate different normalization approaches in Chapter \@ref(ch-hakear).
+In Chapter \@ref(ch-gracsr), our methods were tested on several data designs with different parameters to evaluate their performance.
+$\wpd$ is computationally heavy for cyclic granularities with smaller levels.
+JS and WPD approaches are also computationally intensive when run on large number of customers.
+Hence, most of the scripts used to run these studies use parallel processing for better computational speed.
+R version 4.0.1 (2020-06-06) is utilized on the platform x86 64-apple-darwin17.0 (64-bit) operating on macOS Mojave 10.14.6, as well as the High Performance Computing (HPC) resources provided by [MonARCH](https://docs.monarch.erc.monash.edu/MonARCH/aboutMonArch.html).
+
+<!-- The package contains functions for implementing the complete clustering methodology with choices of scaling and distance metrics discussed in the chapter. It -->
+
+## Limitations and future ideas
+
+We address several limitations of the current framework that might serve as natural next steps for this work and some potential short- and long-term aims in future ideas.
+
+### Limitations
+
+The time series are observed over a short period of time (1--3 years) in the motivating example of this research, and they are assumed to be stationary.
+But it is possible that the distributions change over time, even over a short period.
+For the smart meter example, the distribution for a customer moving to a different house or changing electrical equipment may change drastically.
+To detect these dynamic changes, non-stationarity in time series has to be incorporated while visualizing distributions and also computing distances for two non-stationary time series or one stationary and another non-stationary time series.
+
+Additionally, it is possible that data for a whole category of cyclic granularity is unavailable or that there are insufficient observations to compute distributions.
+For example, a customer may not have data for a particular day of the week or month throughout their observation period.
+While visualizing probability distributions across categories in Chapters \@ref(ch-gravitas) and \@ref(ch-hakear), this can be indicated by displaying dot plots instead of summarizing distributions.
+But the distances in Chapter \@ref(ch-gracsr) can not handle missing observations if they are structured like this.
+We would like to be able to think about designing a distance metric that can incorporate customers with structural missingness and also comprehend its implications while visually characterizing them.
+Another related direction is how to manage the swarm of nuisance variables produced by transforming the time series to cyclic granularities.
+Because the design of our existing framework treats each level of granularity as a variable, it is important to identify levels of granularity that do not contribute to clustering and remove them from the distance metric in order to improve the performance of the clustering algorithm.
+
+### Future ideas
+
+The standard methodologies provide contradicting recommendations for the optimal number of clusters when there are nuisance variables (no contribution to clustering) or nuisance observations (inlying and outlying observations falling between clusters).
+As a result, a more realistic short-term goal should be to test the persistence of clustering solutions in the presence of nuisance variables for the chosen number of clusters.
+This can be done by introducing slightly different samples into the study and observing how the clustering methods handle increased heterogeneity.
+
+Another possible direction is to reduce computational time so that the proposed methods are easily scalable to many customers.
+The distance measure, $\wpd$ is computationally heavy even under parallel computation.
+Moreover, while computing distances between time series, the proposed methods compute all possible pairwise distances, which acts as a computational barrier.
+Faster nearest-neighbor search algorithms can be employed here to decrease the computational load.
+
+A longer-term goal would be to create a similar framework for visualizing and analyzing multivariate time series data.
 With multiple time series available for each observation, the complexity of efficient exploration and visualization grows exponentially.
@@
 This adds to the already high-dimensional data structures that result from studying distributions.
-Current research on visualizing multivariate data must account for time's inherent characteristics in order to efficiently explore distributions of multivariate time series data.
-
-### Comparing our clustering method with other benchmark methods
-
-Testing needs to be carried out with non-hierarchy based clustering methods to see whether the proposed distances perform better with other algorithms.
+This big problem can be tackled by first incorporating time’s inherent characteristics while visualizing one or a few multivariate time series data.
+Unsupervised clustering can then be used to group multiple time series across multiple time granularities and variables.
+This is a method similar to the one used in this thesis for dealing with univariate time series.
+
+<!-- There are a few areas to extend this research.
+The first two explore catering to time series data, where the design of our framework is limited.
+Handling non-stationary time series data or data with structured missingness, for example.
+Second, scaling up for more uncertainty is beneficial since huge data sets with greater uncertainty may have more difficulties than a few time series.
+It is essential to recognize the difficulties of analyzing clustering results in such a setting.
+To that end, it would be interesting to investigate how to deal with anomalies or determine the significance of the clusters.
+A third possibility is to focus on expanding the scope of the framework to include multivariate time series data.
+-->
+
+<!-- **Non-stationary time series** -->
+
+<!-- **Structured missingness** -->
+
+<!-- **Anomalies** -->
+
+<!-- Characterizing clusters with varied or outlying customers can result in patterns that do not represent the group.
+Moreover, integrating heterogeneous consumers may result in visually identical end clusters, which are potentially not useful.
+An appropriate anomaly detection approach could be applied to filter the anomalies, and then a decision has to be made whether the anomalies should form a group or be assigned to the closest group through a classification method.
+-->
+
+<!-- ### Comparing our clustering method with other benchmark methods -->
+
+<!-- Testing needs to be carried out with non-hierarchy based clustering methods to see whether the proposed distances perform better with other algorithms.
 It would also be helpful to know how much more value we obtain from considering distributions rather than aggregating time series data.
 In the case of smart meters, this could be accomplished by comparing the results while validating them against external data such as weather conditions, socioeconomic or other demographic factors of those households.
-
-### Scaling up the clustering method to incorporate large uncertainty and improved computational efficiency that comes with large data
-
-With the volume of data expected to increase in the future, potentially leading to increased variability in patterns, more experiments are needed to understand what sort of problems may arise with our current methodology and how to improve our current algorithm to address those issues.
-All of this must be done while keeping computational efficiency in mind, as this is critical when scaling up for analysis of large data sets.
-
-### Develop inferential methods for the clusters
-
-It is crucial to determine whether or not any detected clusters or patterns are statistically significant.
-In Chapter \@ref(ch:gracsr), we have used the permutation approach to detect if patterns are authentic or not across one or a pair of granularities.
-A similar method for finding significance across clusters is not appropriate if the observations are assumed to be independent.
-Most large temporal data sets gathered at fine temporal scales usually include several repetitive patterns.
-Therefore, the permutations used for comparisons must account that different temporal granularities have distinct conditional distributions.
-
-### Check generalizability of our methods
-
-We propose solutions that are realistically applicable to any temporal data collected more than once per year and may be implemented in a variety of settings.
+-->
+
+<!-- ### Extend methods to other fields -->
+
+<!-- We propose solutions that are realistically applicable to any temporal data collected more than once per year and may be implemented in a variety of settings.
 We could simply verify its utility in the context of smart meter electricity data.
 With the availability of several open-source benchmark data sets, it is vital to evaluate how well the methodologies perform across a variety of fields.
+-->
 
@@
 
-# (APPENDIX) Appendix {-#ch:appendix}
-
-## Data dictionary
-
-```{r wdi-data-dict}
-```
-
-(ref:wdi-data-dict) Data dictionary for the World Development Indicators.
-
-<!--chapter:end:Rmd//06-appendix.Rmd-->
-
 `r if (knitr::is_html_output()) '# Bibliography {-}'`
 
-# References {- #ch:references}
-
 <!--chapter:end:Rmd//99-references.Rmd-->

### 2022-02-07 scripts/gravitas.R

@@ -64,7 +64,7 @@ scene1 <- VIC %>%
   ggplot(aes(x = wknd_wday,
              y = general_supply_kwh)) +
   facet_wrap(~quarter_year, nrow = 1) +
-  geom_lv(aes(fill = ..LV..), k = 5) +
+  geom_lv(aes(fill = ..LV..), k = 5, outlier.size = 0.5, outlier.alpha = 0.1) +
   ylab("") +
   xlab("") +
   scale_fill_brewer(type = "seq", direction = -1) +
@@ -86,7 +86,7 @@ scene2 <- VIC %>%
     "wknd_wday", "quarter_year",
     response = "general_supply_kwh",
     plot_type = "lv",
-    symmetric = FALSE
+    symmetric = FALSE, outlier.size = 0.5, outlier.alpha = 0.1
   ) +
   ylab("") +
   xlab("quarters of the year") +
@@ -106,7 +106,8 @@ scene3 <- VIC %>%
     "quarter_year", "month_year",
     response = "general_supply_kwh",
     plot_type = "lv",
-    symmetric = FALSE
+    symmetric = FALSE,
+    outlier.size = 0.5, outlier.alpha = 0.1
   ) +
   ylab("") +
   xlab("months of the year") +

### 2022-02-07 scripts/hakear.R

@@ -942,16 +942,17 @@ heatplot <- elec_sig_split %>%
   scale_fill_gradient(low = "white",high = "#ba5370") +
   #scale_fill_manual(palette = "Dark2") +
   #scale_colour_manual(values = c("white","red")) + 
-  theme(legend.position = "none") +
+  theme(legend.position = "right") +
   coord_fixed() + 
   guides(fill = guide_legend()) +
   theme_void() +
   theme_gray(base_size = 12, base_family = "Times") +
-  theme(panel.grid = element_blank(),
-        axis.text.x = element_text(angle = 60, hjust=1),legend.position = "bottom") + ggtitle("a") +
   theme(
-    strip.background = element_blank(),
-    strip.text.y  = element_blank(), plot.margin = unit(c(0, -2, 0, 0), "cm")) + ggtitle("(a)") + xlab("x") + ylab("facet") 
+        axis.text.x = element_text(angle = 60, hjust=1),
+        legend.position = "bottom") + ggtitle("a") +
+  theme(
+    #strip.background = element_blank(),
+    strip.text.y  = element_blank(), plot.margin = unit(c(0, -2, 0, 0), "cm")) + ggtitle("(a)") + xlab("x") + ylab("facet")
 
 elec <- read_rds(here("data/hakear/elec_all-8.rds")) %>% 
   dplyr::filter(date(reading_datetime) >= ymd("20190701"), date(reading_datetime) < ymd("20191231"), meter_id==1) %>% 
@@ -1007,7 +1008,8 @@ elec_zoom <-  elec %>%
         panel.grid.minor.x =  element_line(colour = "#D3D3D3")) +
   theme(
     strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) +
-  scale_color_discrete(type = okabeito)
+  scale_color_discrete(type = okabeito) +
+  theme(axis.ticks.y=element_blank())
 
 
 
@@ -1048,13 +1050,16 @@ parcoord <- GGally::ggparcoord(data_pcp,
   ylab("wpd")
 
 
-(heatplot + 
-    theme(legend.position = "none") +
-    facet_grid(id~.) +
-    elec_zoom +
-    theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))) +  
+((heatplot + 
+    #theme(legend.position = "none") +
+    facet_wrap(~id, ncol = 1, strip.position = "right")) +
+    (elec_zoom +
[…]

### 2022-02-07 Rmd/02-gravitas.Rmd
added gestalt theory comments

@@
 When $C_i$ is mapped to the $x$ position and $C_j$ to facets, then the $A_k$ levels are juxtaposed and each $B_\ell$ represents a group/facet.
-Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups.
+Gestalt theory (@wertheimer1938gestalt) suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups.
 Hence, in this case, the $A_k$'s are compared against each other within each group.

### 2022-02-07 scripts/gravitas.R

@@ -228,7 +228,7 @@ cust2_box <- smart_meter10 %>%
   prob_plot(
     "hour_day", "wknd_wday",
     response = "general_supply_kwh",
-    plot_type = "boxplot"
+    plot_type = "boxplot", outlier.size = 0.5, outlier.alpha = 0.1
   ) +
   xlab("") +
   ylab("") +

### 2022-02-09 report/response.Rmd
Ch 4 Q1 Juergen

@@
 
+## Chapter 1
+
+Perhaps though, this would be a good place to detail the co-authorship arrangements for the succeeding chapters, and for the student’s contribution to be stated clearly.
+
 ## Chapter 2
@@
 
-Discuss before change.
+Done. 
+The outliers are made smaller and transparent and caption changed.
 
@@
 
-Discuss before change.
+Not change.
+Due to the space constraint in the paper.
 
@@
 
+No change.
 It appears in the abstract of the paper.
 For the thesis, I have added it since the abstract acts as a summary to the chapter.
+The package reference is mentioned in the Supplementary section of the paper.
+Moreover, the plan is to bring all the functions from `gravitas` to `hakear` when we publish it.
 
@@
 
+Done.
 Added "for a continuous univariate dependent variable" in the last paragraph of Introduction.
@@
 
-Discuss.
+Done.
 Changed to theme_bw() instead of default and outlier.alpha (0.5 ->0.1) and geom_jitter(alpha = 0.04 -> 0.06)
@@
 
-No change. Refer as months or month_year depending on the context.
+No change. 
+Refer as months or month_year depending on the context.
 
@@
 
-Changed.
+Done.
 On p35, $v$ changed to $p$, where $p$ represents the number of variables.
@@
 
-To do.
+Added.
+@wertheimer1938gestalt
 
@@
 
-Discuss.
+CHECK
 No change.
@@
 3.3.3) it is m.
+
+Done.
 
@@
 first part: Done.
-second part: Discuss.
+second part: This is a good suggestion. We will include that into future work.
 
@@
 
-Changed.
+Done.
 Caption changed to incorporate that.
@@
 
-Discuss.
+CHECK.
 No change.
@@
 
-> Q15. The equations on page 44 could be tidied up. There are extra () on the definition of wpdperm and on
-the residual definition at the bottom of the page. Use $\times$ for the equations.
+> Q15. The equations on page 44 could be tidied up. There are extra () on the definition of wpdperm and on the residual definition at the bottom of the page. Use $\times$ for the equations.
 
@@
 
-> Q16. Is there a practical reason why wpdglm is not working as expected for lower nx and nf ?
+> Q16. Is there a practical reason why wpdglm is not working as expected for lower nx and nf?
 
-Discuss.
-To do.
+CHECK.
+No practical reason but a more theoretical one.
+For lower $n_x$ and $n_f$, the number of combinations $n_xn_f$ is low such that the max do not have an assymptotic distribution which can be assumed as Normal by CLT.
+With higher combinations, the distribution of the each combination could be assumed to be normal.
 
@@
 
-Discuss.
 Not changed.
 
-> Q19. The presentation of the material on the simulation study on page 48 could be improved. Where is
-the notation wpdl,s used? Figure 3.5 shows the m = 2 results only. The text alludes to a simulation
-involving different underlying distributions, but this is not mentioned again.
+> Q19.
+The presentation of the material on the simulation study on page 48 could be improved.
+Where is the notation $wpd_{l,s}$ used?
+Figure 3.5 shows the m = 2 results only.
+The text alludes to a simulation involving different underlying distributions, but this is not mentioned again.
 
+DISCUSS
 first part: How?
@@
 
-To do. Discuss.
+Done.
 
@@
 
-To do. Discuss.
+Done.
 
-> Q22. In Figure 3.7 (a) the heatmaps need id labels. The grey color is missing from the legend. Maybe use a
-different colour in the heatmap for the significant comparisons. State the threshold for significance in
-the caption. In Table 3.3 the caption should explain the threshold. Maybe use colour instead of stars to indicate significance?
+> Q22.
+In Figure 3.7 (a) the heatmaps need id labels.
+The grey color is missing from the legend.
+Maybe use a different colour in the heatmap for the significant comparisons.
+State the threshold for significance in the caption.
+In Table 3.3 the caption should explain the threshold.
+Maybe use colour instead of stars to indicate significance?
 
@@
 
+MDS plot summary of the data sets from the validation study.
+
 > Q13. Figure 4.7 is missing labels (a), (b) (c).
 
+Done. Added a, b representing 12 customers (rows) in each of them.
+
 > Q14. Why do you chose to summarize 4 and 5 clusters from JS-NQT when sindex suggests 3 clusters?
+
+It is told in the "number of clusters" section that the number of optimal number of clusters is a difficult exercise.
+The idea was to show how decreasing the number of clusters can hide distinct patterns.
 
@@
 # Prof Juergen Symanzik
+
+## Chapter 1
+
+> Q1.
+While the author hints at other sources of similar data, I think it would be worthwhile to give the reader some further specific ideas here in the Introduction where similar complex data may occur, e.g., data from traffic sensors, movement data (turnstiles in metro stations or public buildings), and even more traditional measurements such as temperatures or precipitation over
+time that could benefit from the methods and software described in this thesis.
+
+> Q2. It may also be helpful to state what is not be covered in this thesis: The spatial component of such data where an additional component could be latitude and longitude or some areal code,
+such as the post code or some administrative units.
+Extending the work from this thesis with spatial proximity methods for clustering (https://doi.org/10.1016/j.cageo.2011.12.017) [assuming that households in certain residential neighborhoods may have similar energy usage patterns] and trying spatial visualization methods (if possible) such as glyph maps (https:
+//onlinelibrary.wiley.com/doi/abs/10.1002/env.2152) should be mentioned somewhere as a consideration for future work | but that may become a future new thesis by itself.
+
+> Q3. Finally, it would be helpful to indicate here (and in each chapter) what the names or abbreviations for the three R packages (gravitas, hakear, and gracsr) represent. Even the full R package
+title of "gravitas: Explore Probability Distributions for Bivariate Temporal Granularities" does not provide a full answer what the "tas" component means (I am assuming that Gravi stands for Granularities Visualization).
+I am even more lost with hakear.
+
+## Chapter 2
+
+> Q1. One reference to a table (Table 2.3) does not resolve and appears as @ref{tab:tab-mayan} in the main text on p. 16.
+
+Done.
+
+> Q2.
+how to define and handle pay day (or loan day) which often is defined as the 1st (and/or possibly 15th) work day of a month, resulting in some aperiodic granularity as some of these days across the year could initially fall on a weekend, but even more in a few
+cases, may coincide with a public holiday (and thus shifting pay day by 2 or 3 days).
+Similarly, extended weekends that could stretch from Fridays to Sundays or from Saturdays to Mondays (and possibly even over a 4{day period) might be worth some discussion, in particular as a case of variable{length aperiodic granularity.
+This latter granularity may be of particular interest when working with temporal data with an economic (travel, restaurants, etc.) or entertainment (movie theaters, casinos, etc.) aspect.
 
@@
 
-Discuss. (Not adding link in supplementary section enough?)
+Done. (Adding link in supplementary section enough?)
 
@@
 
-Discuss.
+Done.
+We have assumed both nx; nf simultaneously to be less than or equal to 5
 
@@
 
-Can do.
+Done.
+$\leqslant$ for both.
 
@@
 
-Can do.
+Done.
+Changed to "insignificant"
 
@@
 
-Can do.
+Done.
 

### 2022-02-09 Rmd/03-hakear.Rmd
Ch 4 Q1 Juergen

@@
 
-These results justify our use of the permutation approach when $\nx<5$ and $\nf<5$, and the use of the GLM otherwise.
+These results justify our use of the permutation approach when $\nx \leq 5$ and $\nf \leq 5$, and the use of the GLM otherwise.
 <!--
@@
 According to Figure \ref{fig:dotplot-8}(c), for the harmony pair (*dow*-*hod*), household id 7 has the greatest value of $\wpd$, while id 1 has the least.
-From Table \ref{tab:tab-rank-8} it can be seen that the harmony pair (*dow*, *hod*) is important for id 7; however, it has been labeled as an inconsequential pair for id 1.
+From Table \ref{tab:tab-rank-8} it can be seen that the harmony pair (*dow*, *hod*) is important for id 7; however, it has been labeled as an insignificant pair for id 1.
 The distribution of energy demand for both of these households, with *dow* as the facet and *hod* on the x-axis, may help explain the choice.
@@
 
-(ref:gravitas-plot-8) Comparing distribution of energy demand shown for household id 1 (a) and 7 (b) across *hod* in x-axis and *dow* in facets through quantile area plots.
+(ref:gravitas-plot-8) Comparing distribution of energy demand shown for household id 1 (a) and 7 (b) on logarithmic scale across *hod* in x-axis and *dow* in facets through quantile area plots.
 The value of $\wpd$ in Table 3 suggests that the harmony pair (*dow*, *hod*) is significant for household id 7, but not for id 1.

### 2022-02-09 Rmd/04-gracsr.Rmd
Ch 4 Q1 Juergen
Ch 4 all done Juergen

@@
 Other algorithms utilized in the literature include $k$-means variants, hierarchical clustering, and greedy $k$-medoids.
-Time series data, such as smart meter data, are not well-suited to any of the techniques mentioned in @Tureczek2017-pb.
+Many techniques mentioned in @Tureczek2017-pb fail to recognize smart meter readings as a data type with a temporal component [@Tureczek2018-ha].
 Only one study [@ozawa2016determining] identified time series characteristics by first conducting a Fourier transformation, to convert data from time to frequency domain, followed by $k$-means to cluster by greatest frequency.
@@
 ```{r flowchart, fig.cap = "Flow chart illustrating the pipeline for our method for clustering time series.", out.width="100%"}
-knitr::include_graphics("img/flowchart.png")
+knitr::include_graphics("img/flowchart-comment.png")
 
@@
 The number indicates a unique customer id and a prototype id.
+In each of the plots, the line
+represents the median, and the grey region shows the area between the 25th and 75th percentiles.
 

### 2022-02-09 scripts/gracsr.R

@@ -1064,13 +1064,14 @@ mds_plot_10 <- ggplot(all_data,
                       aes(x = Dim.1,
                           y = Dim.2,
                           color = group)) +
-  geom_point(size = 2, alpha = 0.5, shape = 3) +
+  geom_point(size = 2, alpha = 0.5, aes(shape = group)) +
   facet_grid(diff~Scenario, labeller = "label_both") +
   #geom_text(check_overlap = TRUE)  +
   theme_bw()+
   theme(legend.position = "bottom") +
   scale_color_brewer(palette = "Dark2")+
-  ylab("mds1") + xlab("mds2")
+  ylab("mds1") + xlab("mds2") +
+  scale_shape_manual(values = c(3, 4, 2, 0, 1)) 
 
 mds_plot_10
 

### 2022-02-10 report/response.Rmd
update response document with questions to discuss

@@
 
+DISCUSS
+
 ## Chapter 2
@@
 
+. DISCUSS
 Not change.
@@
 
-CHECK
+. DISCUSS
 No change.
@@
 
-CHECK.
+. DISCUSS
 No change.
@@
 
-CHECK.
+. DISCUSS
 No practical reason but a more theoretical one.
@@
 
-DISCUSS
+. DISCUSS
 first part: How?
@@
 
-To do. Discuss.
+To do. DISCUSS.
 
@@
 To do? (Or not)
+Partially done. Need to add "or"
 
@@
 
+. DISCUSS
+
 > Q15. The reference to the stationarity assumption in the Discussion needs clarification.
+
+. DISCUSS
 
@@
 
+. DISCUSS
 # Prof Juergen Symanzik
@@
 
+. DISCUSS
+
 > Q2. It may also be helpful to state what is not be covered in this thesis: The spatial component of such data where an additional component could be latitude and longitude or some areal code,
@@
 
+. DISCUSS
 > Q3. Finally, it would be helpful to indicate here (and in each chapter) what the names or abbreviations for the three R packages (gravitas, hakear, and gracsr) represent. Even the full R package
@@
 This latter granularity may be of particular interest when working with temporal data with an economic (travel, restaurants, etc.) or entertainment (movie theaters, casinos, etc.) aspect.
+
+DISCUSS
 
@@
 
-Discuss.
+DISCUSS
 
@@
 
-Done. "and" needed?
+Done.
+. DISCUSS "and" needed?
 
@@
 
+Done.
+Rephrased to make it clear that most techniques mentioned in Tureczek and Nielsen (2017) do not treat smart meter data as a data type with temporal component.
+
 > Q2. Use a consistent style and use past tense throughout when you summarize what was previously published.
@@
 
-> Q3. The last cross{reference at the end of the Introduction should be to Section 4.5 (and not Section 2.7).
+> Q3. The last cross-reference at the end of the Introduction should be to Section 4.5 (and not Section 2.7).
 
@@
 
-first part: Done
-second part: Discuss (The diagram to be changed as per text or reverse)
+Done.
 
@@
 
-Done. Added links in the supplementary materials section at the end of the chapter.
+Done. 
+Added links in the supplementary materials section at the end of the chapter.
 
@@
 
-Discuss.
+Done.
+Not showing in the plot, check again.
 
@@
 
-Discuss.
+No change.
+Showing the vertical axis would not give us any more information than it is already giving.
 
@@
 Moreover, in previous chapters, you used orange and green for the quartiles and 10th/90th quantile.
-Why not using the
-same colors and quantiles here? If this becomes too confusing for a reader, then at least, indicate in the caption which quantiles are covered by the gray areas.
+Why not using the same colors and quantiles here?
+If this becomes too confusing for a reader, then at least, indicate in the caption which quantiles are covered by the gray areas.
 
-first part: Done.
-second part: Discuss
+Done.
 
@@
 
-Done. (but not showing in pdf check again)
+Done.
 
@@
 
-Discuss.
+We will include it in the journal paper.
+
+## Bibliography
+
+> Q1. Capitalize all nouns, e.g., in
+Time-series clustering - A decade review.
+
+CHECK?
+Google scholar citation appears like this, is it important?
+
+> Q2. Use full{length journal names, i.e., do not use abbreviations, e.g., in Inf. Syst.
+
+Done.
+Fixed all similar problems.
+
+> Q5. Make sure that URLs are clickable and lead to the correct web page, e.g.,
+robjhyndman.com/publications/hakear does not resolve correctly.
+
+Done.
+All URLs checked and they are clickable.
+
+> Q6. For Gupta, S, RJ Hyndman, D Cook, and A Unwin (2021), I would suggest to list the DOI as https://doi.org/10.1080/10618600.2021.1938588 as part of the reference.
+
+Done.
+
+> Q7. No need to list the ISBN, e.g., in Xie, Y (2016).
+
+Done.

### 2022-02-10 report/response.Rmd
update response document with hints for  discussion quesstions

@@
 
-. DISCUSS
 Not change.
-Due to the space constraint in the paper.
+Figure is already published
 
@@
 No change.
+Data in Fig 3.3 (a) is consistent with the null distribution.
 Null distributions refer to the case when there is no significant distributional differences between categories of one or more granularities.
@@
 
+Add it somewhere, need to have reasonable number of observations, otherwise collapse categories (To do)
+
 > Q15. The equations on page 44 could be tidied up. There are extra () on the definition of wpdperm and on the residual definition at the bottom of the page. Use $\times$ for the equations.
@@
 
-. DISCUSS
+To do. Need to mention it in the paper.
 No practical reason but a more theoretical one.
@@
 first part: How?
-second part: After NQT, results are pretty similar to an underlying normal distribution. (Discuss)
+second part: The results are made available in the supplementary paper (Discuss)
+
+Mention m=1 results are same as m=2, although important not included in the paper but in the supplements. Make subsections for m=1 and 2.
+TO DO
 
@@
 In Table 3.3 the caption should explain the threshold.
-Maybe use colour instead of stars to indicate significance?
+(Maybe use colour instead of stars to indicate significance?)
 
@@
 
+Visual summary to show why 3 clusters are inadequate for this data set.
+It is pretty normal in a cluster analysis to check few $k$ before and after the optimal $k$.
+
 . DISCUSS
@@
 
-. DISCUSS
+The conditional distributions are constant. Don't talk about the stationary time series. If there are jumps, then out methods can't model it.
 
@@
 
-DISCUSS
+Very good point. Paper is already published. We would incorporate this in our future work.
 
@@
 DISCUSS
+(Existing distance measure to obtain differences between categories like in ANOVA, limitation: only capture one measure of central tendency
+Cite some anova here as lit review?
 
@@
 Done.
-. DISCUSS "and" needed?
+Made it past tense.
 
@@
 We will include it in the journal paper. 
+Split into a "Limitation" and "Discussion" section separately detailing the drawbacks.
 
@@
 
-CHECK?
-Google scholar citation appears like this, is it important?
+To do. Make it consistent sentence case.
 

### 2022-02-16 report/response.Rmd
response update

@@
 
-To do.
+Done.
+Rephrased to make it clear.
 
@@
 
-To do.
+Done.
+Added the statement in page 60.
 
@@
 
-> Q5. I found Figure 4.1 very useful. In each box, would be helpful to put in “or”, in places where only one of
-the steps listed is performed, eq Normal quantile transform or Robust scaling. As there are many steps
-in the algorithm, it would be helpful to the reader to label the pipeline steps and to refer back to them
-in the text. The text in the Data pre-processing box does not make sense to me.
+> Q5.
+I found Figure 4.1 very useful.
+In each box, would be helpful to put in “or”, in places where only one of the steps listed is performed, eq Normal quantile transform or Robust scaling.
+As there are many steps
+in the algorithm, it would be helpful to the reader to label the pipeline steps and to refer back to them in the text. The text in the Data pre-processing box does not make sense to me.
 
@@
 
-> Q6. The section describing RS and NQT is confusing. It states RS is applied to each time series separately.
-Is NQT also applied to each observation separately? How does “it could be useful to standardize it for
-the selected set of significant granularities prior to computing the distances” relate to the following
-bullet points?
+> Q6.
+The section describing RS and NQT is confusing.
+It states RS is applied to each time series separately.
+Is NQT also applied to each observation separately?
+How does “it could be useful to standardize it for
+the selected set of significant granularities prior to computing the distances” relate to the following bullet points?
+
+Done.
+Clarification: Each observation (time series) is scaled independently in both JS-NQT and JS-RS.
+For the WPD approach, each observation might be scaled or not scaled, resulting in usable groups in either case.
 
@@
 
-> Q21. Are parts a and b in Figure 3.8 correctly labeled and also matching with the figure caption? The caption reads For id 1, patterns look similar within weekdays and weekends.
-Visually, this seems to be the case for id 7 (in part b). This also visually matches Figure 3.7b.
+> Q21.
+Are parts a and b in Figure 3.8 correctly labeled and also matching with the figure caption?
+The caption reads For id 1, patterns look similar within weekdays and weekends.
+Visually, this seems to be the case for id 7 (in part b).
+This also visually matches Figure 3.7b.
 
-Discuss.
+Caption rephrased to make it more clear.
+
+For id: 7, there is more distributional difference between consecutive hours of the day compared to id:1.
+Also all the days of the week are different from each other in id:7.
+id:1 has similar behavior across all days within both weekdays and weekends.
+For this reason, the wpd for id:7 has a higher value than id:1 as the similar behavior within weekdays and weekends balances out the difference between weekdays and weekends.
+The display with hod vs wdwnd would have shown higher values of wpd for id:1 (as can be verified from Table 3).
 

### 2022-02-16 Rmd/04-gracsr.Rmd
bibliography fix
response update

@@
 
-- It is reasonable to define a time series based on its degree of trend and seasonality, and to take these characteristics into account while clustering it.
-The modification of the data structure by taking into account probability distributions across cyclic granularities assures that there is no trend and that seasonal variations are handled independently.
-As a result, there is no need to de-trend or de-seasonalize the data before applying the clustering method.
+- It is fair to describe a time series based on its degree of trend and seasonality, and to cluster it based on these features.
+The addition of probability distributions across cyclic granularities to the data structure ensures, there is no need to de-trend or de-seasonalize the data prior to using the clustering algorithm.
 For similar reasons, there is no need to exclude holiday or weekend routines.
@@
 Age, lifestyle, family composition, building attributes, weather, availability of diverse electrical equipment, among other factors, make the task of properly segmenting customers into comparable energy behavior complex.
-The challenge is to be able to cluster consumers into these types of expected patterns, and other unexpected patterns, using only their energy usage history [@Ushakova2020-rl].
+When there is no further customer data available, such as property type, location, or family size, the problem is to cluster customers into these sorts of predicted patterns, as well as other unexpected patterns, using just their energy consumption history [@Ushakova2020-rl].
 <!--To safeguard the customers' privacy, it is probable that such information is not accessible.
@@
 In each of the plots, the line
-represents the median, and the grey region shows the area between the 25th and 75th percentiles.
+represents the median, and the gray region shows the area between the 25th and 75th percentiles.
 

### 2022-02-17 report/response.Rmd
Ch4 Q14 Catherine

@@
 
-> Q14. In the pairwise distance measure on page 43, is there any adjustment made for varying numbers of
-observations across the levels of A and B? For example in Figure 3.3(d) if there are few values at
-xlevel=1 the comparison shown is less interesting
+> Q14.
+In the pairwise distance measure on page 43, is there any adjustment made for varying numbers of observations across the levels of A and B?
+For example in Figure 3.3(d) if there are few values at xlevel=1 the comparison shown is less interesting
 
@@
 
-> Q9. The description of the data generation on page 67 and Table 4.2 is confusing. Maybe state the
-distribution of each vt.
+> Q9. The description of the data generation on page 67 and Table 4.2 is confusing. Maybe state the distribution of each vt.
 
@@
 
- $\mu$ is the difference between means considered for consecutive categories.
+Done.
+Clarification: $\mu$ is the difference between means considered for consecutive categories.
 
 > Q11. In Section 4.3.3 what are the values of R and T?
+
+Done.
+Clarification: All
 
@@
 
-It is told in the "number of clusters" section that the number of optimal number of clusters is a difficult exercise.
-The idea was to show how decreasing the number of clusters can hide distinct patterns.
-
-Visual summary to show why 3 clusters are inadequate for this data set.
-It is pretty normal in a cluster analysis to check few $k$ before and after the optimal $k$.
-
-. DISCUSS
+Done.
+The caption and description of the figure in the text have been changed to indicate that Fig.
+4.9 depicts a visual summary of why three clusters are insufficient for this data set, however it is recommended by sindex summary plots.
+In a cluster analysis, it is standard practice to test a few cluster sizes before and after the optimal $k$ to see if a different number of clusters is useful.
+In this example, increasing the number of clusters actually unfolds some behavior which was not apparent with $k=3$ as suggested by sindex.
 

### 2022-02-17 Rmd/04-gracsr.Rmd
Ch4 Q14 Catherine

@@
 This clustering almost agrees with the clusters visible in the t-SNE plot.
+This display also serves as a visual summary of why three clusters are insufficient using JS-NQT approach, however it is recommended by *sindex* summary plots in Figure \ref{fig:opt-cluster-tsne-jsd}.
 
@@
 
-(ref:groups-4and5) Summary plots for four and five clusters from JS-NQT showing the distribution of electricity demand combined for all members over *hod*, *moy*, and *wnwd*.
+(ref:groups-4and5) Visual summary of why three clusters are insufficient for JS-NQT approach, however it is recommended by *sindex*.
+The plot shows four and five clusters from JS-NQT showing the distribution of electricity demand combined for all members over *hod*, *moy*, and *wnwd*.
 Groups A-1 (customers 21-24), A-2 (customers 4-9), and A-3 (customers 10-15) profiles correspond to Groups B-1, B-2, and B-3, respectively.

### 2022-02-17 report/response.Rmd
Ch1 Q1 Catherine

@@
 
-DISCUSS
+Done.
+A table outlining the main chapters’ co-authorship arrangements included in Introduction.
 
@@
 
-The conditional distributions are constant. Don't talk about the stationary time series. If there are jumps, then out methods can't model it.
+Done.
+Clarification: "we have assumed the time series to be stationary" has been changed to "the conditional distributions are assumed to remain constant for the observation period" in the Discussion section.
 
@@
 
-> Q17. Again, the computational burden could do with more Discussion. At present, if I were to use this
-method on my data, what kind of sizes are realistic to work with?
+> Q17. Again, the computational burden could do with more Discussion. At present, if I were to use this method on my data, what kind of sizes are realistic to work with?
 
 . DISCUSS
+
 # Prof Juergen Symanzik
@@
 . DISCUSS
+
 > Q3. Finally, it would be helpful to indicate here (and in each chapter) what the names or abbreviations for the three R packages (gravitas, hakear, and gracsr) represent. Even the full R package

### 2022-02-17 Rmd/01-intro.Rmd
Ch1 Q1 Catherine

@@
 Chapter \@ref(ch-conclusion) summarizes the software tools developed for the work, and discusses future plans.
+Table \ref{tab:contribution-outline} details the publications, including my and my fellow co-authors contributions.
+
+\begin{table}
+\centering\footnotesize\tabcolsep=0.12cm
+\caption{Table outlining the main chapters' co-authorship arrangements.}
+\label{tab:contribution-outline}
+\begin{tabular}{|p{1cm}|p{2cm}|p{1.5cm}|p{3.5cm}|p{3.5cm}|p{1.5cm}|}
+\hline
+\RaggedRight\textbf{Thesis Chapter}  & \RaggedRight\textbf{Publication Title}  & \RaggedRight\textbf{Status (published, in press, accepted or returned for revision)}  & \RaggedRight\textbf{Nature and} {\%} \RaggedRight\textbf{of student contribution} & \RaggedRight\textbf{Co-author name(s) Nature and} {\%} \RaggedRight\textbf{of Co-author’s contribution} & \RaggedRight\textbf{Co-author(s), Monash student Y/N} \\ \hline
+2 & Visualizing probability distributions across bivariate cyclic temporal granularities & Published & 70\%.
+Concept, methodology development, writing first draft and software development & (1) Rob J Hyndman, Concept, methodology and software development, writing 12\% (2) Dianne Cook, Concept, methodology and software development, writing 12\% (3) Antony Unwin, Concept and software development 6\% & \multicolumn{1}{c|}{N} \\
+\hline
+3 & Detecting distributional differences between temporal granularities
+for exploratory time series analysis & Working Paper & 80\%.
+Concept, methodology development, writing first draft and software development & (1) Rob J Hyndman, Concept, methodology development and writing 10\% (2) Dianne Cook, Concept, methodology development and writing 10\%  & \multicolumn{1}{c|}{N} \\
+\hline
+4 & Clustering time series based on probability distributions across temporal granularities & To be submitted & 80\%.
+Concept, methodology development, writing first draft and software development & (1) Dianne Cook, Concept, methodology development and writing 15\% (2) Rob J Hyndman, Concept, methodology development and writing 5\% & \multicolumn{1}{c|}{N} \\
+\hline
+\end{tabular}
+\end{table}

### 2022-02-17 scripts/hakear.R
Ch3 Q3 Catherine

@@ -154,8 +154,8 @@ id2_tsibble_hd <- elec %>%
   as_tsibble(index = date_time) %>% 
   create_gran("hour_day") %>% 
   ggplot(aes(x = hour_day, y = kwh)) +
-  geom_boxplot(width = 0.6, outlier.colour = "black", outlier.alpha = 0.1, fill = "#CC79A7", colour =  "#0072B2") + scale_y_log10() +
-  geom_jitter(alpha = 0.06, colour = "#E69F00") +
+  geom_boxplot(width = 0.6, outlier.colour = "black", outlier.alpha = 0.2, fill = "#CC79A7", colour =  "#0072B2") + scale_y_log10() +
+  geom_jitter(alpha = 0.03, colour = "#E69F00", position = position_jitter(height = .2, width = .2)) +
   ggtitle("(a)")+ 
   ylab ("energy usage [kWh]") +
   theme_bw()
@@ -167,8 +167,8 @@ id2_tsibble_dw <- elec %>%
   as_tsibble(index = date_time) %>% 
   create_gran("month_year") %>% 
   ggplot(aes(x = month_year, y = kwh)) +
-  geom_boxplot(width = 0.6, outlier.colour = "black", outlier.alpha = 0.1, fill = "#CC79A7", colour =  "#0072B2") + scale_y_log10()+
-  geom_jitter(alpha = 0.05, colour = "#E69F00") +
+  geom_boxplot(width = 0.6, outlier.colour = "black", outlier.alpha = 0.2, fill = "#CC79A7", colour =  "#0072B2") + scale_y_log10()+
+  geom_jitter(alpha = 0.03, colour = "#E69F00", position = position_jitter(height = .2, width = .2)) +
   ggtitle("(b)")+ 
   ylab ("energy usage [kWh]") +
   theme_bw()

### 2022-02-21 Rmd/03-hakear.Rmd
response update
Juergen Q3 Chapter3 graph tickmarks and labels changed

@@
 ```{r onegran-new, fig.cap="A cyclic granularity can be considered to be a categorical variable, and used to break the data into subsets.
-Here, side-by-side boxplots overlaid on jittered dotplots explore the distribution of of energy use by a household for two different cyclic granularities: (a) hour-of-day and (b) and month-of-year.
+Here, side-by-side boxplots overlaid on jittered dotplots explore the distribution of energy use (on a logarithmic scale) by a household for two different cyclic granularities: (a) hour-of-day and (b) month-of-year.
 Daily peaks occur in morning and evening hours, indicating a working household, where members leave for and return from work.
-More volatility of usage in summer months (Jan, Feb) is probably due to air conditioner use on just some days.", fig.pos="!b"}
+More volatility of usage in summer months in Australia (Jan, Feb) is probably due to air conditioner use on just some days.", fig.pos="!b"}
 ```
@@
 This implies that distributional differences are captured more by this harmony for id 7, which is apparent from this display with more fluctuations across median and 75th percentile for different hours of the day and day of week.
-For id 1, patterns look similar within weekdays and weekends.
+For id 1, patterns look similar within different days of weekdays and weekends.
 Here, the median is represented by the black line, the orange area corresponds to quartile deviation and the green area corresponds to area between $10^{th}$ and $90^{th}$ quantile.
+The display with hod vs wdwnd would have shown higher values of wpd for id:1 (as can be verified from Table 3).
 

### 2022-02-21 scripts/hakear.R

@@ -66,11 +66,15 @@ p1 <- id2_tsibble %>%
             response = "kwh",
             plot_type = "quantile",
             symmetric = TRUE,
-            quantile_prob = c(0.1, 0.25,0.5,0.75, 0.9)) +
+            quantile_prob = c(0.1, 0.25,0.5,0.75, 0.9), size = 1) +
   ggtitle("(a)") + theme(
     strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + 
   scale_colour_brewer(name = "", palette = "PiYG") + 
-  ylab ("energy usage [kWh]")
+  ylab ("energy usage [kWh]") +
+  scale_y_log10(breaks = scales::trans_breaks("log10", 
+                                              function(x) 10^x),
+                 labels = scales::trans_format("log10", scales::math_format(10^.x))) +
+  theme_bw()
 
 
 p2 <- id4_tsibble %>%
@@ -79,11 +83,14 @@ p2 <- id4_tsibble %>%
             response = "kwh",
             plot_type = "quantile",
             symmetric = TRUE,
-            quantile_prob = c(0.1, 0.25,0.5,0.75, 0.9)) + 
+            quantile_prob = c(0.1, 0.25,0.5,0.75, 0.9), size = 1) + 
   ggtitle("(b)") + theme(
     strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))+
-  ylab ("energy usage [kWh]")
-
+  ylab ("energy usage [kWh]")+
+  scale_y_log10(breaks = scales::trans_breaks("log10", 
+                                              function(x) 10^x),
+                labels = scales::trans_format("log10", scales::math_format(10^.x))) +
+  theme_bw()
 
 
 p3 <- id2_tsibble %>%

### 2022-02-22 Rmd/03-hakear.Rmd
Juergen Q5 Chap3

@@
 This suggests that the household does not use as much electricity for heating as it does for air conditioning.
-A lot of households in Victoria use gas heating and hence the heater use might not be reflected in the electricity data.
+A lot of households in Melbourne use gas heating and hence the heater use might not be reflected in the electricity data.
 
@@
 In this paper, the visual discovery of distributional differences is facilitated by choosing a threshold for the proposed numerical distance measure, eventually selecting only those cyclic granularities for which the distributional differences are sufficient to make it an interesting display.
+Two things need to be accomplished here: 1) to see if there are any statistically significant differences between independent groups, and 2) to quantify any differences that do exist.
+One way to address this problem is by using a one-way or two-way ANOVA [@fisher1992statistical].
+Assume we want to examine whether there is a significant difference in electricity demand on various days of the week.
+In this case, each day of the week may be regarded as an independent group, and a one-way ANOVA can be used to assess how the means of the electricity demand varies across different days of the week.
+The approach proposed in this paper looks at the distributional differences in the quantitative variable instead of merely the mean or one measure of central tendency.
+Besides, it also takes into account that the levels in each group have an inherent order in a time series context.
 
@@
 $$
-where $M = \frac{q_1+q_2}{2}$ and $D(q_1||q_2) := \int^\infty_{-\infty} q_1(x)f(\frac{q_1(x)}{q_2(x)})$ is the KL divergence between distributions $q_1$ and $q_2$.
-Other common measures of distance between distributions are Hellinger distance, total variation distance and Fisher information metric.
+where $M = \frac{q_1+q_2}{2}$ and $D(q_1||q_2) := \int^\infty_{-\infty} q_1(x)log(\frac{q_1(x)}{q_2(x)})$ is the KL divergence between distributions $q_1$ and $q_2$.
+There are other ways to obtain distance between distributions like Hellinger distance, total variation distance and Fisher information metric (all of which are special cases of f-divergence), but Jensen-Shannon distance was chosen for ease of computation.
 

### 2022-02-22 report/response.Rmd
Ch4 Q15 Catherine
Ch3 Q3 Catherine
update response document

@@
 
+Not change.
+Figure is already published.
+
 > Q2. Last sentence of Section 2.3.4: fix the table reference
@@
 
-> Q4. The violin plot in Figure 2.5c is barely recognisable, and comparison of the distributions is challenging.
-The plot deserves a bit more space.
+> Q4. The violin plot in Figure 2.5c is barely recognisable, and comparison of the distributions is challenging. The plot deserves a bit more space.
 
@@
 Done.
-Changed to theme_bw() instead of default and outlier.alpha (0.5 ->0.1) and geom_jitter(alpha = 0.04 -> 0.06)
+Changed to theme_bw() instead of default and outlier.alpha (0.5 ->0.2) and geom_jitter(alpha = 0.04 -> 0.03). Also jitter-width decreased to 0.2.
 
@@
 
-No change.
+Done.
 Changed to "distribution of energy usage"
@@
 In Figure 3.1, it would help the reader to mention in the figure caption that a log-scale has been used for the vertical axis.
-Moreover, it might be helpful to add a ticmark label at 10 kWh and add ticmarks representing each fraction of 10 on the log-scale and not only the mid{point between two ticmark labels.
+Moreover, it might be helpful to add a ticmark label at 10 kWh and add ticmarks representing each fraction of 10 on the log-scale and not only the mid-point between two ticmark labels.
 Also, reminding the reader that summer months are in January and February in Australia would be helpful.
-When looking at the figure (prior to reading the main text), my first interpretation for the increased energy usage in January and February was because of (electric) heating | and not cooling.
+When looking at the figure (prior to reading the main text), my first interpretation for the increased energy usage in January and February was because of (electric) heating and not cooling.
 
-Can do.
+Done. (Can't get the tick marks at the right place)
 
@@
 
-Can do.
+Not changed.
+More than comparing the two houses, it is important to compare the patterns of their distributons across months.
+They have very different scales and showing them in the same scale would make it less clear.
 
@@
 
-DISCUSS
-(Existing distance measure to obtain differences between categories like in ANOVA, limitation: only capture one measure of central tendency
-Cite some anova here as lit review?
+CLARIFICATION 1:
+The goal is to determine if there are any statistically significant differences between independent (unrelated) groups similar to a one-way or two-way ANOVA.
+A one-way ANOVA, for example, is analogous to $m=1$ (looking at one cyclic granularity at a time) in our approach.
+Assume we want to examine whether there is a significant difference in electricity demand on various days of the week.
+In this case, each day of the week may be regarded as an independent group, and we are looking at distributional differences rather than mean differences (as in a standard ANOVA).
+Similarly, considering $m=2$ (two cyclic granularities) is equivalent to a two-way ANOVA and is used to assess how distributions of quantitative variables vary according to the levels of two categorical variables rather than merely the mean or one measure of central tendency.
+There are other ways to obtain distributional differences, but we chose Jensen-Shannon distances between quantiles for ease of computation.
+
+CALRIFICATION 2: f in the equation was a typo. Replaced by log now.
+
+CALRIFICATION 3: Distance measures that are mentioned at the end of that paragraph are not explained in details but mentioned that they are all special cases of f-divergence.
+
+CHANGED: Introduction (last but one paragraph) and Distance between distributions section.
 
@@
 > Q8. The term Gestalt theory is mentioned a few times (p. 38 & p. 42), but it is never supported by a reference.
+
+Done.
 

### 2022-02-22 report/response.Rmd
Juergen Q4 Chap3

@@
 
-Not changed.
-More than comparing the two houses, it is important to compare the patterns of their distributons across months.
-They have very different scales and showing them in the same scale would make it less clear.
+The ticmark labels are placed in both the figures but the houses are not brought to a same scale.
+More than comparing the two houses, it is important to compare the patterns of their distributions across months.
+They have different scales and showing them in the same scale is making the individual distributional differences seem less clear.
 

### 2022-02-22 scripts/hakear.R

@@ -209,7 +209,7 @@ v <- hakear::compute_pairwise_norm_scalar(sm, gran_x, gran_facet,
                                           response = kwh, lambda = 0.9, dist_ordered = FALSE)
 
 
-## ---- null4by2
+##----compute-pairwise-norm-null4by2
 
 set.seed(9999)
 
@@ -328,11 +328,6 @@ varall_sd <- compute_pairwise_norm(sim_panel_varall_sd,
                                    nperm = 200
 )
 
-# plot
-p_varall <- sim_panel_varall %>%
-  ggplot(aes(x = as.factor(id_x), y = sim_data)) + facet_wrap(~id_facet) + geom_boxplot() +
-  ggtitle(paste("(d)", round(varall, 2))) + xlab("x level") +
-  ylab ("simulated response")
 
 
 set.seed(9999)
@@ -363,11 +358,6 @@ null_sd <- compute_pairwise_norm(sim_panel_null_sd,
 # 
 # 
 
-p_null <- sim_panel_null %>%
-  ggplot(aes(x = as.factor(id_x), y = sim_data)) + facet_wrap(~id_facet) + geom_boxplot() +
-  ggtitle(paste("(a)", round(null, 2))) + xlab("x level") +
-  ylab ("simulated response")
-
 
 
 
@@ -379,12 +369,6 @@ varf <- compute_pairwise_norm(sim_panel_varf,
                               nperm = 200
 )
 
-p_varf <- sim_panel_varf %>%
-  ggplot(aes(x = as.factor(id_x), y = sim_data)) + facet_wrap(~id_facet) + geom_boxplot() +
-  ggtitle(paste("(b)", round(varf, 2))) + xlab("x level") +
-  ylab ("simulated response")
-
-
 set.seed(9999)
 varx <- compute_pairwise_norm(sim_panel_varx, 
                               gran_x = "id_x",
@@ -393,15 +377,44 @@ varx <- compute_pairwise_norm(sim_panel_varx,
                               nperm = 200
 )
 
+
+
+## ---- null4by2
+# plot
+p_varall <- sim_panel_varall %>%
+  ggplot(aes(x = as.factor(id_x), y = sim_data)) + facet_wrap(~id_facet) + geom_boxplot() +
+  ggtitle(paste("(d)", round(varall, 2))) + xlab("x level") +
[…]

### 2022-02-22 Rmd/03-hakear.Rmd
Juergen Q4 Chap3
Juergen Q12 Ch3

@@
 The value of our proposed distance measure is presented for each panel, the relative differences between which will be explained later in Section \@ref(sec:ranking).
+
+```{r compute-pairwise-norm-null4by2}
+
+```
 
@@
 Specifying $\lambda_x>0.5$ will weight within-facet distances more heavily, while $\lambda_x <0.5$ would weight the between-facet distances more heavily.
-(See the supplements for more details.)
+(See [Section 2.1 supplements](https://github.com/Sayani07/paper-hakear/blob/master/paper/paper-supplementary.pdf) for more details.)
 
@@
 Simulations were carried out to explore the behavior of $\wpd$ as $\nx$ and $\nf$ were varied, in order to compare and evaluate different normalization approaches.
-More detailed simulation results, including an exploration of the size and power of the proposed test, are contained in the supplements.
+More detailed simulation results, including an exploration of the size of the proposed test, are contained in the [supplements](https://github.com/Sayani07/paper-hakear/blob/master/paper/paper-supplementary.pdf).
 
@@
 Table \ref{tab:glm-tab} gives the summary of a Gamma generalized linear model to capture the relationship between $\wpdsub{raw}$ and the number of comparisons.
-The intercepts and slopes are similar, independent of the underlying distributions (see supplementary paper for details) and hence the coefficients are shown for the case when observations are drawn from a N(0,1) distribution.
+The intercepts and slopes are similar, independent of the underlying distributions (see [Table 3 supplementary paper](https://github.com/Sayani07/paper-hakear/blob/master/paper/paper-supplementary.pdf) for details) and hence the coefficients are shown for the case when observations are drawn from a N(0,1) distribution.
 Figure \ref{fig:dist-new-same-scale-link} shows the distribution of $\wpdsub{perm}$ and $\wpdsub{glm-scaled}$ on the same scale to show that a combination approach could be used for higher values of $\nx$ and $\nf$ to alleviate the computational time of the permutation approach.

### 2022-02-22 scripts/hakear.R

@@ -615,7 +615,8 @@ bind_rows(broom::tidy(glm_fit_onegran),
   mutate(estimate = round(estimate, 2),
          std.error = round(std.error, 2),
          statistic = round(statistic, 2),
-         p.value = round(p.value, 2),
+         #p.value = round(p.value, 2),
+         p.value = "<.001",
          term = rep(c("Intercept","$\\log(\\nx \\times \\nf)$"),2)
   ) %>%
   kable(booktabs=TRUE, escape=FALSE,

### 2022-02-22 response.Rmd
Juergen Q13 Ch3

--- 2022-02-22 report/response.Rmd
@@
 
-Discuss.
+Discuss. (Time taking)
 
 > Q7. A few editorial corrections are necessary, e.g.,
-Dang and Wilkinson (2014); Wilkinson, Anand, and Grossman (2005) provide misses an \and" between the two references. The same holds for Buja et al. (2009); Majumder, Hofmann, and Cook (2013) present.
-I would leave it to the author to check for similar omissions. Moreover, these should be past tense: "provided" and "presented".
+Dang and Wilkinson (2014); Wilkinson, Anand, and Grossman (2005) provide misses an \and" between the two references.
+The same holds for Buja et al.
+(2009); Majumder, Hofmann, and Cook (2013) present.
+I would leave it to the author to check for similar omissions.
+Moreover, these should be past tense: "provided" and "presented".
 
@@
 
-Done. (Adding link in supplementary section enough?)
+Done.
+Mentioned which table or subsection needs to be referred and the link of the supplementary material provided.
 
@@
 On p.
-45 and in (3.2), it is not immediately clear whether nx; nf simultaneously have to be less than or equal to 5 so that wpdperm is being used - or whether only one of them has to be less than 5.
+45 and in (3.2), it is not immediately clear whether nx; nf simultaneously have to be less than or equal to 5 so that $wpd_{perm}$ is being used - or whether only one of them has to be less than 5.
 Just a verbal clarification is needed.
@@
 Done.
-We have assumed both nx; nf simultaneously to be less than or equal to 5
+CLARIFICATION: We have assumed both $nx$; $nf$ simultaneously to be less than or equal to 5
 
@@
 
-> Q12. In Table 3.2, listing a p-value as 0 never is a good idea. List it as < 0:01 or < 0:0001 or any other meaningful threshold that matches the number of your simulation runs.
+> Q12. In Table 3.2, listing a p-value as 0 never is a good idea. List it as < 0.01 or < 0.0001 or any other meaningful threshold that matches the number of your simulation runs.
+
+Done.
 
@@
 I suppose this figure relates to m = 2, but this is not stated in the main text.
-Table 3.2
-lists both, m = 1 and m = 2. Moreover, in Figure 3.5, the y{axis ticmark labels overplot and the nx values are partially cut off.
+Table 3.2 lists both, m = 1 and m = 2.
+Moreover, in Figure 3.5, the y-axis ticmark labels overplot and the nx values are partially cut off.
 
-Discuss for changes. Can do.
+Done.
+We decided to remove the y-axis labels in the interest of space. The y-axis is showing the density values but we are only interested in its shape.
+
+CLARIFICATION: Added $m=2$ in the caption of Figures 3.5 and 3.6.
 

### 2022-02-22 Rmd/03-hakear.Rmd
Juergen Q13 Ch3

@@
 
-(ref:raw) Distribution of $\wpdsub{raw}$ is plotted across different $\nx$ and $\nf$ categories under $\Dnull$ through density and rug plots.
-Both location (blue line) and scale (orange marks) of the distribution shifts for different panels.
+(ref:raw) Distribution of $\wpdsub{raw}$ is plotted across different $\nx$ and $\nf$ categories under $\Dnull$ through density and rug plots for $m=2$.
+Both location (blue line) and scale of the distribution shifts for different panels.
 This is not desirable since under the null design, the distribution is not expected to capture any differences.
@@
 
-(ref:dist-new-same-scale-link) The distributions of $\wpdsub{perm}$ and $\wpdsub{glm-scaled}$ are overlaid to compare the location and scale across different $\nx$ and $\nf$.
+(ref:dist-new-same-scale-link) The distributions of $\wpdsub{perm}$ and $\wpdsub{glm-scaled}$ are overlaid to compare the location and scale across different $\nx$ and $\nf$ for $m=2$.
 $\wpdsub{norm}$ takes the value of $\wpdsub{perm}$ for lower levels, and $\wpdsub{glm-scaled}$ for higher levels to to alleviate the problem of computational time in permutation approaches.

### 2022-02-22 Rmd/03-hakear.Rmd
Juergen Q16 and Q18 Ch3

@@
 We consider coarser linear granularities like hour, day, week and month from the commonly used Gregorian calendar.
-From the four linear granularities of hour, day, week, and month, we obtain $N_C = 4\times 3/2 = 6$ cyclic granularities:  "hour_day", "hour_week", "hour_month", "day_ week", "day_month" and "week_month".
+From the four linear granularities of hour, day, week, and month, we obtain $N_C = 4\times 3/2 = 6$ cyclic granularities:  "hour_day", "hour_week", "hour_month", "day_ week", "day_month" and "week_month" abbreviated as *hod*, *how*, *hom*, *dow*, *dom* and *wdwnd* respectively.
 Further, we add cyclic granularity day-type ("wknd wday") to capture weekend and weekday behavior.
@@
 Figure \ref{fig:dotplot-8} helps to compare households through the heatmap (a) across harmony pairs.
-Here *dom*, *dow*, *wdwnd* are abbreviations for day-of-month, day-of-week, weekday/weekend and so on.
-The colors represent the value of $\wpd$.
-Darker cells correspond to higher values of $\wpd$.
+Each household is represented by 25 tiles, each tile representing a pair of cyclic granularities.
+The colors (in shades of red) represent the value of $\wpd$ for each of the harmony pairs (in Table \ref{tab:tab-rank-8}) and the gray tiles correspond to clashes.
+A darker shade of red corresponds to higher values of $\wpd$.
 Those with `*` correspond to $\wpd$ values above $\wpdsub{threshold95}$.
@@
 
-According to Figure \ref{fig:dotplot-8}(c), for the harmony pair (*dow*-*hod*), household id 7 has the greatest value of $\wpd$, while id 1 has the least.
+Figure \ref{fig:dotplot-8}(c) shows a parallel coordinate plot across different harmonies with harmonies arranged from highest to lowest $wpd$ values averaged over all households.
+This display is useful for comparing households across harmonies.
+For example, for the harmony pair (*dow*-*hod*), household id 7 has the greatest value of $\wpd$, while id 1 has the least.
 From Table \ref{tab:tab-rank-8} it can be seen that the harmony pair (*dow*, *hod*) is important for id 7; however, it has been labeled as an insignificant pair for id 1.
@@
 Here, the median is represented by the black line, the orange area corresponds to quartile deviation and the green area corresponds to area between $10^{th}$ and $90^{th}$ quantile.
-The display with hod vs wdwnd would have shown higher values of wpd for id:1 (as can be verified from Table 3).
+The display with *hod* vs *wdwnd* would have shown higher values of wpd for id:1 (as can be verified from Table 3).
 

### 2022-02-22 report/response.Rmd
response document updated

@@
 
-Can do.
+Done.
+Added the abbreviations on Page 50 in "Choosing cyclic granularities of interest and removing clashes" section.
 
@@
 
-Can do.
+Done.
+
+CLARIFICATION 1: Each household is represented by 25 tiles, each tile representing a pair of cyclic granularities.
+The colors (in shades of red) represent the value of $wpd$ for each of the harmony pairs (in Table Table 3.3) and the grey tiles correspond to clashes.
+A darker shade of red corresponds to higher values of $wpd$.
+
+CLARIFICATION 2: The harmony pairs in the parallel coordinate plot are arranged from highest to lowest values of $wpd$ averaged over all households and hence do not correspond to the first or last rows of Table 3.3.
+
+Changed the text description of Figure 3.7.
 

### 2022-02-23 report/response.Rmd
Catherine Q9,14,16 Ch3

@@
 
-. DISCUSS
-No change.
+CLARIFICATION: Null distributions refer to the case when there is no significant distributional differences between categories of one or more granularities.
 Data in Fig 3.3 (a) is consistent with the null distribution.
-Null distributions refer to the case when there is no significant distributional differences between categories of one or more granularities.
 
-> Q10. In table 3.1 Nc is the number of cyclic granularities, whereas in the text (first sentence of 3.3.2 and
+> Q10. In table 3.1 $N_c$ is the number of cyclic granularities, whereas in the text (first sentence of 3.3.2 and
 3.3.3) it is m.
@@
 
-. DISCUSS
-No change.
-We assume we have enough observations for each level or combination of levels to compute distributions. As long as that holds, no adjustment made for varying number of observations.
+CLARIFICATION: As long as that there is enough data for each level or combination of levels, no adjustment for varying number of observations is done.
+If there are few data for one or more levels, subsequent categories could be collapsed to obtain precise estimates of quantiles.
 
-Add it somewhere, need to have reasonable number of observations, otherwise collapse categories (To do)
+Added under the subsection "Characterizing distributions".
 
@@
 
-> Q16. Is there a practical reason why wpdglm is not working as expected for lower nx and nf?
+> Q16. Is there a practical reason why $wpd_{glm}$ is not working as expected for lower nx and nf?
 
-To do. Need to mention it in the paper.
-No practical reason but a more theoretical one.
-For lower $n_x$ and $n_f$, the number of combinations $n_xn_f$ is low such that the max do not have an assymptotic distribution which can be assumed as Normal by CLT.
-With higher combinations, the distribution of the each combination could be assumed to be normal.
+CLARIFICATION: There is no practical explanation for this, but a more empirical one - with smaller $n_x$ and $n_f$, the variance in the values of $wpd_{raw}$ is greater.
+As a result, our modeling technique that attempts to fit $wpd_{raw}$ to the number of comparisons ($n_x\times n_f$) performs poorly.
+The increased volatility in the values of $wpd_{raw}$ with a smaller number of levels is due to the fact that no asymptotic distribution for the statistic "maximum" (which is the foundation of $wpd_{raw}$) across such small number of values could be produced.
 
@@
 
-. DISCUSS
+To do. (Easy)
 first part: How?
@@
 Mention m=1 results are same as m=2, although important not included in the paper but in the supplements. Make subsections for m=1 and 2.
-TO DO
 
@@
 
-To do. DISCUSS.
+To do. (Difficult viz)
 
@@
 
-To do? (Or not)
+To do? (Easy)
 Partially done. Need to add "or"
@@
 
-To do.
+To do.(Medium)
 
@@
 
-. DISCUSS
+To do.(Hard)
 
@@
 
-. DISCUSS
+To do. (Medium)
 
@@
 
-Discuss. (Time taking)
+To do. (Medium)
 
@@
 This may require some experimentation.
-Finally, using a differently colored background (say light yellow) would be helpful to visually demonstrate where wpdperm is being used.
+Finally, using a differently colored background (say light yellow) would be helpful to visually demonstrate where $wpd_{perm}$ is being used.
 For the legend, use the terms as in equation (3.2) and not the full word.
 
-Discuss
+Done. Changed to make it more clear than earlier.
 
@@
 in the text).
-Some explanation and labeling is
-needed here.
+Some explanation and labeling is needed here.
 
-discuss.
+To do. (Difficult viz)
 
@@
 Was this based on a random sample, or manually do obtain rather different energy usage patterns?
-This should be answered earlier on, but this also should be addressed in the Discussion: What can be expected when working with the full data set of Chapter 2, i.e., about 13,000.
-You indicate:
+This should be answered earlier on, but this also should be addressed in the Discussion:
+What can be expected when working with the full data set of Chapter 2, i.e., about 13,000. You indicate:
 A future direction of work is to be able to explore and compare many individuals/subjects together for similar patterns across significant granularities.
 What would be computationally (and visually) feasible as 13,000 is a few orders of magnitude
-higher than the number of eight households used in this example? Clearly, no full answer is expected here, but rather some speculation of what could possibly be done in
-the future (or what is done in the next chapter).
+higher than the number of eight households used in this example?
+Clearly, no full answer is expected here, but rather some speculation of what could possibly be done in the future (or what is done in the next chapter).
 
-Discuss.
+To do. (Hard)
 
@@
 
-Ok. To do in hakear.
+Ok. To do in hakear. (Easy)
 
@@
 
-Discuss.
+To do.(Medium)
 
@@
 
-To do. Make it consistent sentence case.
+To do. Make it consistent sentence case. (Easy)
 

### 2022-02-23 Rmd/03-hakear.Rmd
Catherine Q9,14,16 Ch3

@@
 We use percentiles with $p = {0.01, 0.02, \dots, 0.99}$ to reduce the computational burden in summarizing distributions.
+The assumption is that there is sufficient data for each level or combination of levels, and hence no adjustment is made for the varying number of observations across levels.
+However, if one or more levels has a small number of data points, consecutive categories should be collapsed prior to data transformation and quantile estimation.
 

### 2022-02-23 report/response.Rmd
Catherine Q5 Ch3

@@
 
-To do. (Easy)
-first part: How?
-second part: The results are made available in the supplementary paper (Discuss)
-
-Mention m=1 results are same as m=2, although important not included in the paper but in the supplements. Make subsections for m=1 and 2.
+The simulation design and results corresponding to $m=2$ are presented.
+Similar design and results for $m=1$, although important, are not included in the paper but in the supplements.
+Simulation results involving different underlying distributions are also presented in the supplementary paper.
 

### 2022-02-23 Rmd/03-hakear.Rmd
Catherine Q5 Ch3

@@
 
-Simulations were carried out to explore the behavior of $\wpd$ as $\nx$ and $\nf$ were varied, in order to compare and evaluate different normalization approaches.
-More detailed simulation results, including an exploration of the size of the proposed test, are contained in the [supplements](https://github.com/Sayani07/paper-hakear/blob/master/paper/paper-supplementary.pdf).
+Simulations were carried out to explore how the behavior of $\wpd$ as $\nx$ and $\nf$ were varied, in order to compare and evaluate different normalization approaches for both $m=1$ and $m=2$.
+Here the simulation design and results corresponding to $m=2$ are presented.
+Similar design and results for $m=1$, although important, are not included in the paper but in the [supplements](https://github.com/Sayani07/paper-hakear/blob/master/paper/paper-supplementary.pdf) along with other more detailed simulation results.
 
@@
 
-\noindent $m=1$
-
-\noindent Observations were generated from a N(0,1) distribution for $\nx \in \{2, 3, 5, 7, 9, 14, 17, 20, 24, 31, 42, 50\}$, with $\nsub{times} = 500$ observations drawn for each combination of categories.
+<!-- \noindent $m=1$ -->
+
+<!-- \noindent Observations were generated from a N(0,1) distribution for $\nx \in \{2, 3, 5, 7, 9, 14, 17, 20, 24, 31, 42, 50\}$, with $\nsub{times} = 500$ observations drawn for each combination of categories.
+Let $\wpd_{\ell, s}$ denote the value of $\wpd$ obtained for the $\ell^{th}$ panel and $s^{th}$ simulation.
+-->
+
+<!-- \noindent $m=2$ -->
+
+\noindent Observations were generated from a N(0,1) distribution for each combination of $\nx$ and $\nf$ from $\{2, 3, 5, 7, 14, 20, 31, 50\}$, with $\nsub{times} = 500$ observations drawn for each of the $64$ combinations.
 This design corresponds to $\Dnull$.
 For each of the categories, there were $\nsub{sim}=200$ replications, so that the distribution of $\wpd$ under $\Dnull$ could be observed.
-Let $\wpd_{\ell, s}$ denote the value of $\wpd$ obtained for the $\ell^{th}$ panel and $s^{th}$ simulation.
-
-\noindent $m=2$
-
-\noindent Similarly, observations were generated from a N(0,1) distribution for each combination of $\nx$ and $\nf$ from $\{2, 3, 5, 7, 14, 20, 31, 50\}$, with $\nsub{times} = 500$ observations drawn for each of the $64$ combinations.
 
@@
 We consider coarser linear granularities like hour, day, week and month from the commonly used Gregorian calendar.
-From the four linear granularities of hour, day, week, and month, we obtain $N_C = 4\times 3/2 = 6$ cyclic granularities:  "hour_day", "hour_week", "hour_month", "day_ week", "day_month" and "week_month" abbreviated as *hod*, *how*, *hom*, *dow*, *dom* and *wdwnd* respectively.
+From the four linear granularities of hour, day, week, and month, we obtain $N_C = 4\times 3/2 = 6$ cyclic granularities:  "hour_day", "hour_week", "hour_month", "day_ week", "day_month" and "week_month" abbreviated as *hod*, *how*, *hom*, *dow*, *dom* and *wnwd* respectively.
 Further, we add cyclic granularity day-type ("wknd wday") to capture weekend and weekday behavior.
@@
 Figure \ref{fig:gravitas-plot-8} demonstrates that for id 7, the median (black) and quartile deviation (orange) of energy consumption fluctuates for most hours of the day and days of the week, while for id 1, daily patterns are more consistent within weekdays and weekends.
-As a result, for id 1, it is more appropriate to examine the distributional difference solely across (*dow*, *wdwnd*), which has been rated higher in Table \ref{tab:tab-rank-8}.
+As a result, for id 1, it is more appropriate to examine the distributional difference solely across (*dow*, *wnwd*), which has been rated higher in Table \ref{tab:tab-rank-8}.
 
@@
 However, the difference between morning and evening peaks are higher for weekends.
-Since $\wpd$ is designed to put more weightage on within-facet differences for $\lambda>0.5$, it makes sense that the pair (hod, wdwnd) has been ranked higher than (wdwnd, hod).
+Since $\wpd$ is designed to put more weightage on within-facet differences for $\lambda>0.5$, it makes sense that the pair (hod, wnwd) has been ranked higher than (wnwd, hod).
 Panel c) shows the distribution of energy demand with weekday/weekend as the x-axis and week-of-month as the facets.
@@
 Here, the median is represented by the black line, the orange area corresponds to quartile deviation and the green area corresponds to area between $10^{th}$ and $90^{th}$ quantile.
-The display with *hod* vs *wdwnd* would have shown higher values of wpd for id:1 (as can be verified from Table 3).
+The display with *hod* vs *wnwd* would have shown higher values of wpd for id:1 (as can be verified from Table 3).
 

### 2022-02-23 Rmd/04-gracsr.Rmd
Ch1 Q1 Catherine
Catherine Q5 Ch3

@@
 
-Computationally in R, the data is assumed to be a "tsibble object" [@wang2020tsibble] equipped with an index variable representing inherent ordering from past to present and a key variable that defines observational units over time.
-The measured variable for an observation is a time-indexed sequence of values.
-This sequence, however, could be shown in several ways.
-A shuffle of the raw sequence may represent hourly consumption throughout a day, a week, or a year.
-Cyclic granularities can be expressed in terms of the index set in the tsibble data structure.
+The initial data in R is supposed to be a "tsibble object" [@wang2020tsibble] with an index variable representing inherent ordering from past to present, a key variable that specifies observational units through time, and measured variables.
+As a result, the measured variable for a key is a sequence of values that is time-indexed.
+However, this sequence may be shown in a variety of ways.
+A shuffling of the raw sequence may reflect hourly consumption over the course of a day, a week, or a year.
 
 The data object will change when cyclic granularities are computed, as multiple observations will be categorized into levels of the granularity, thus inducing multiple probability distributions.
-Directly computing Jensen-Shannon distances between the entire probability distributions can be computationally intensive.
-Thus it is recommended that quantiles are used to characterize the probability distributions.
-In the final data object, each category of a cyclic granularity corresponds to a list of numbers which is composed of a few quantiles.
+Directly calculating Jensen-Shannon distances between all probability distributions can be time-consuming.
+As a result, it is suggested that quantiles be employed to characterize probability distributions.
+In the final data object, each category of cyclic granularity corresponds to a list of numbers, which is composed of a few quantiles.
 
@@
 Hence, a way of appropriately scaling it up to many customers such that anomalies are removed before clustering would be useful for bringing forth meaningful, compact and separated clusters.
-Secondly, we have assumed the time series to be stationary, and hence the distributions are assumed to remain constant for the observation period.
+Secondly, the conditional distributions are assumed to remain constant for the observation period.
 In reality, however, it might change.

### 2022-02-23 scripts/hakear.R

@@ -610,18 +610,29 @@ glm_fit_onegran  <- glm(actual ~ log(`nx * nfacet`),
                         family = Gamma(link = "inverse"),
                         data = G21_median_onegran)
 
-bind_rows(broom::tidy(glm_fit_onegran),
-          tidy(glm_fit), .id = "$m$") %>%
+# bind_rows(broom::tidy(glm_fit_onegran),
+#           tidy(glm_fit), .id = "$m$") %>%
+#   mutate(estimate = round(estimate, 2),
+#          std.error = round(std.error, 2),
+#          statistic = round(statistic, 2),
+#          #p.value = round(p.value, 2),
+#          p.value = "<.001",
+#          term = rep(c("Intercept","$\\log(\\nx \\times \\nf)$"),2)
+#   ) %>%
+#   kable(booktabs=TRUE, escape=FALSE,
+#         caption = "Results of generalised linear model to capture the relationship between $\\wpdsub{raw}$ and the number of comparisons.")
+
+
+          bind_rows(tidy(glm_fit)) %>%
   mutate(estimate = round(estimate, 2),
          std.error = round(std.error, 2),
          statistic = round(statistic, 2),
          #p.value = round(p.value, 2),
          p.value = "<.001",
-         term = rep(c("Intercept","$\\log(\\nx \\times \\nf)$"),2)
+         term = rep(c("Intercept","$\\log(\\nx \\times \\nf)$"),1)
   ) %>%
   kable(booktabs=TRUE, escape=FALSE,
         caption = "Results of generalised linear model to capture the relationship between $\\wpdsub{raw}$ and the number of comparisons.")
-
 
 ## ---- wpd-glm-dist
 
@@ -958,6 +969,9 @@ demography %>% add_footnote(label = "***: 99% significant, **: 95% and *: 90%")
 ## ---- dotplot-8
 
 elec_harmony_all <- read_rds(here("data/hakear/elec_harmony_all.rds"))
+
+elec_harmony_all <- elec_harmony_all %>% mutate(facet_variable = if_else(facet_variable=="wdwnd","wnwd",facet_variable))%>% mutate(x_variable = if_else(x_variable=="wdwnd","wnwd",x_variable))
+
 select_split <- str_split(elec_harmony_all$select_harmony, " ", simplify = TRUE)[,2]
 
 elec_sig_split <- elec_harmony_all %>% 

### 2022-02-23 Rmd/03-hakear.Rmd
Catherine Q19 Ch3

@@
 We consider coarser linear granularities like hour, day, week and month from the commonly used Gregorian calendar.
-From the four linear granularities of hour, day, week, and month, we obtain $N_C = 4\times 3/2 = 6$ cyclic granularities:  "hour_day", "hour_week", "hour_month", "day_ week", "day_month" and "week_month" abbreviated as *hod*, *how*, *hom*, *dow*, *dom* and *wnwd* respectively.
+From the four linear granularities of hour, day, week, and month, we obtain $N_C = 4\times 3/2 = 6$ cyclic granularities:  "hour_day", "hour_week", "hour_month", "day_ week", "day_month" and "week_month" abbreviated as *hod*, *how*, *hom*, *dow*, *dom* and *wdwnd* respectively.
 Further, we add cyclic granularity day-type ("wknd wday") to capture weekend and weekday behavior.
@@
 Figure \ref{fig:gravitas-plot-8} demonstrates that for id 7, the median (black) and quartile deviation (orange) of energy consumption fluctuates for most hours of the day and days of the week, while for id 1, daily patterns are more consistent within weekdays and weekends.
-As a result, for id 1, it is more appropriate to examine the distributional difference solely across (*dow*, *wnwd*), which has been rated higher in Table \ref{tab:tab-rank-8}.
+As a result, for id 1, it is more appropriate to examine the distributional difference solely across (*dow*, *wdwnd*), which has been rated higher in Table \ref{tab:tab-rank-8}.
 
@@
 However, the difference between morning and evening peaks are higher for weekends.
-Since $\wpd$ is designed to put more weightage on within-facet differences for $\lambda>0.5$, it makes sense that the pair (hod, wnwd) has been ranked higher than (wnwd, hod).
+Since $\wpd$ is designed to put more weightage on within-facet differences for $\lambda>0.5$, it makes sense that the pair (hod, wdwnd) has been ranked higher than (wdwnd, hod).
 Panel c) shows the distribution of energy demand with weekday/weekend as the x-axis and week-of-month as the facets.
@@
 Here, the median is represented by the black line, the orange area corresponds to quartile deviation and the green area corresponds to area between $10^{th}$ and $90^{th}$ quantile.
-The display with *hod* vs *wnwd* would have shown higher values of wpd for id:1 (as can be verified from Table 3).
+The display with *hod* vs *wdwnd* would have shown higher values of wpd for id:1 (as can be verified from Table 3).
 

### 2022-02-23 scripts/hakear.R

@@ -610,18 +610,29 @@ glm_fit_onegran  <- glm(actual ~ log(`nx * nfacet`),
                         family = Gamma(link = "inverse"),
                         data = G21_median_onegran)
 
-bind_rows(broom::tidy(glm_fit_onegran),
-          tidy(glm_fit), .id = "$m$") %>%
+# bind_rows(broom::tidy(glm_fit_onegran),
+#           tidy(glm_fit), .id = "$m$") %>%
+#   mutate(estimate = round(estimate, 2),
+#          std.error = round(std.error, 2),
+#          statistic = round(statistic, 2),
+#          #p.value = round(p.value, 2),
+#          p.value = "<.001",
+#          term = rep(c("Intercept","$\\log(\\nx \\times \\nf)$"),2)
+#   ) %>%
+#   kable(booktabs=TRUE, escape=FALSE,
+#         caption = "Results of generalised linear model to capture the relationship between $\\wpdsub{raw}$ and the number of comparisons.")
+
+
+          bind_rows(tidy(glm_fit)) %>%
   mutate(estimate = round(estimate, 2),
          std.error = round(std.error, 2),
          statistic = round(statistic, 2),
          #p.value = round(p.value, 2),
          p.value = "<.001",
-         term = rep(c("Intercept","$\\log(\\nx \\times \\nf)$"),2)
+         term = rep(c("Intercept","$\\log(\\nx \\times \\nf)$"),1)
   ) %>%
   kable(booktabs=TRUE, escape=FALSE,
         caption = "Results of generalised linear model to capture the relationship between $\\wpdsub{raw}$ and the number of comparisons.")
-
 
 ## ---- wpd-glm-dist
 

### 2022-02-23 report/response.Rmd
response doc update
Juergen Q11 Ch4

@@
 
-To do? (Easy)
-Partially done. Need to add "or"
+Rephrased to make it fluent.
+Figure 4.1 labels changed to make it consistent with the text.
 
@@
 
-Ok. To do in hakear. (Easy)
+Done.
 
@@
 Changing from red to blue/purple initially hides this information.
-Also arrange the legend from P-1 to P-3 and Q-1 to Q-5 to make it more obvious that P-1 & Q-1 and P-2 & Q-2 are identic
+Also arrange the legend from P-1 to P-3 and Q-1 to Q-5 to make it more obvious that P-1 & Q-1 and P-2 & Q-2 are identical
 
-To do.(Medium)
+Done.
 
@@
 
-To do. Make it consistent sentence case. (Easy)
+Made all bib entries consistent (sentence case)
 

### 2022-02-23 scripts/gracsr/application.R

@@ -836,6 +836,8 @@ knitr::kable(format = "latex",
 
 ## ----summary-plot-wpd------------------------------------------------------------
 
+data_pcp_plot$group5 <- factor(data_pcp_plot$group5, levels = c(1, 2, 3, 4, 5))
+
 parcoord5 <- GGally::ggparcoord(data_pcp_plot %>% select(-group3)%>% left_join(cluster_result_id, by = "customer_id"),
                                 columns = 3:5,
                                 groupColumn = "group5",
@@ -854,9 +856,9 @@ parcoord5 <- GGally::ggparcoord(data_pcp_plot %>% select(-group3)%>% left_join(c
   labs(colour = "group") +
   theme(panel.border = element_blank()) +
   scale_color_discrete(labels = c(c("Q-1","Q-2", "Q-3", "Q-4", "Q-5")))+
-  scale_color_manual(values = c("#3300FF", "#1B9E77", "#3B3178", "#1D7CF2", "#AE6D1C"), labels = c("Q-5","Q-2", "Q-3", "Q-4", "Q-1"))
+  scale_color_manual(values = c("#AE6D1C", "#1B9E77","#3300FF",  "#3B3178", "#1D7CF2"), labels = c("Q-1","Q-2", "Q-3", "Q-4", "Q-5"))
 
-
+data_pcp_plot$group3 <- factor(data_pcp_plot$group3, levels = c(1, 2, 3))
 
 parcoord3 <- GGally::ggparcoord(data_pcp_plot %>% select(-group5)%>% left_join(cluster_result_id, by = "customer_id"),
                                 columns = 3:5,
@@ -875,7 +877,7 @@ parcoord3 <- GGally::ggparcoord(data_pcp_plot %>% select(-group5)%>% left_join(c
   ylab("wpd") +
   labs(colour = "group") +
   theme(panel.border = element_blank()) +
-  scale_color_manual(values = c("#D8367D", "#1B9E77","#AE6D1C"), labels = c("P-3","P-2", "P-1"))
+  scale_color_manual(values = c("#AE6D1C", "#1B9E77","#00008B"), labels = c("P-1","P-2", "P-3"))
 
 
 parcoord3 <- parcoord3 +geom_text_repel(aes(label = sort_group_id), size = 2) &theme_light()&theme(legend.position = "bottom") 

### 2022-02-24 Rmd/01-intro.Rmd
border colors

@@
 This motivates the research presented in this thesis, which aims to provide a platform for systematically exploring probability distributions induced by these multiple observations to support the discovery of regular patterns or anomalies, as well as the exploration of clusters of behaviors or the summarization of the behavior.
+While we were prompted by the example of the smart meters, the challenges and solutions presented in this thesis are essentially applicable to any temporal data collected more than once a year, such as data from traffic sensors, pedestrian movement data in metro stations or public buildings, cab or bike rides, or even more traditional measurements such as temperature or precipitation collected over time.
+In a larger sense, it may be appropriate for data observed over years, decades, or centuries, as in weather or astronomical data.
 

### 2022-02-24 scripts/hakear.R

@@ -991,18 +991,32 @@ elec_sig_split <- elec_harmony_all %>%
 elec_sig_split$significant <- 
   factor(elec_sig_split$significant, levels = c("highest", "high", "medium", "low"))
 
+outline <- data.frame( 
+  id = unique(elec_sig_split$id), 
+  #outline_color = okabeito
+  outline_color = c("#D55E00", "#0072B2","#009E73", "#CC79A7", "#E69F00","#56B4E9", "#F0E442", "#000000")
+) 
+square <- with(elec_sig_split, data.frame( 
+  x = c(-Inf, Inf, Inf, -Inf), 
+  y = c(-Inf, -Inf, Inf, Inf)
+))
+
+
 heatplot <- elec_sig_split %>% 
   mutate(significance_95 = if_else(significant %in% c("high", "highest"), "*", "")) %>% 
   ggplot(aes(x = x_variable,
              y = facet_variable)) +
   geom_tile(aes(fill = wpd)) + 
+  geom_polygon(aes(x = x,y = y, color = id),size=2, fill= NA, data = merge(outline, square))+
+  scale_color_discrete(type = okabeito,
+                       guide = "none")+
   #color = as.factor(significance_95)),
   #size = 0.3) +
   geom_text(aes(label = significance_95), color = "#42275a") +
   scale_fill_gradient(low = "white",high = "#ba5370") +
   #scale_fill_manual(palette = "Dark2") +
   #scale_colour_manual(values = c("white","red")) + 
-  theme(legend.position = "right") +
+  theme(legend.position = "none") +
   coord_fixed() + 
   guides(fill = guide_legend()) +
   theme_void() +
@@ -1011,8 +1025,11 @@ heatplot <- elec_sig_split %>%
         axis.text.x = element_text(angle = 60, hjust=1),
         legend.position = "bottom") + ggtitle("a") +
   theme(
-    #strip.background = element_blank(),
-    strip.text.y  = element_blank(), plot.margin = unit(c(0, -2, 0, 0), "cm")) + ggtitle("(a)") + xlab("x") + ylab("facet")
+    strip.background = element_blank(),
+    strip.text.y  = element_blank(), 
+    strip.text.x  = element_blank(), 
+    plot.margin = unit(c(0, -2, 0, 0), "cm")) + ggtitle("(a)") + xlab("x") + ylab("facet") +
+  facet_wrap(~id, ncol = 1)
[…]

### 2022-02-24 report/response.Rmd
Q17 (Juergen) and Q22(Caatherine) Ch3

@@
 
-I appreciate all the comments and suggestions, and will incorporate them into future work. Specific comments are addressed as follows:
-
 ## Chapter 1
 
-Perhaps though, this would be a good place to detail the co-authorship arrangements for the succeeding chapters, and for the student’s contribution to be stated clearly.
+> This would be a good place to detail the co-authorship arrangements for the succeeding chapters, and for the student’s contribution to be stated clearly.
 
@@
 
-Not change.
+Not changed.
 Figure is already published.
@@
 
-Not change.
+Not changed.
 Figure is already published
@@
 $m$ refers to the number of cyclic granularities we are considering together in the display.
-For example, contextual cyclic granularities could be $hour_day$, $day_week$ and $month_year$ and we want to visualize any one granularity at a time.
+For example, contextual cyclic granularities could be $hour_{day}$, $day_{wee}k$ and $month_{year}$ and we want to visualize any one granularity at a time.
 So $N_c = 3$ and $m=1$.
@@
 
-Not changed.
+$m=1$ removed from the main chapter and included in supplements.
 
@@
 
-To do. (Difficult viz)
+CHANGED.
+- Different box borders (instead of id labels) are added such that (a) and (b) display the household ids in the same order as indicated by the same colour of the line plot in (b) and box border 
+- threshold explained in caption in Table 3.3 
+- grey color explained in the text and caption
+
+CLARIFICATION 1: Using color to convey significance instead of stars (in the Fig 3.3a) is a not a great idea because the varied shades of the tiles represent different values of the $wpd$.
+For example, both tiles may be significant, and their colours may differ to represent distinct $wpd$ values.
+To be consistent, Table 3.3 also has stars to represent significance.
 
@@
 
-> Q9. The description of the data generation on page 67 and Table 4.2 is confusing. Maybe state the distribution of each vt.
+> Q9. The description of the data generation on page 67 and Table 4.2 is confusing. Maybe state the distribution of each $v_t$.
 
-To do.(Medium)
+Stating the distribution of each $v_t$ is difficult because the levels for each granularity and design varies, therefore I chose to divide it into two tables.
 
@@
 
+Suppose there are $n$ observations.
+Our current methodology has a time complexity of $O(n^2)$ since it calculates distance between all pairs.
+The time taken to compute distances between a pair of observations based on $p$ cyclic granularity is
+
 # Prof Juergen Symanzik
@@
 > Q1.
-While the author hints at other sources of similar data, I think it would be worthwhile to give the reader some further specific ideas here in the Introduction where similar complex data may occur, e.g., data from traffic sensors, movement data (turnstiles in metro stations or public buildings), and even more traditional measurements such as temperatures or precipitation over
-time that could benefit from the methods and software described in this thesis.
+While the author hints at other sources of similar data, I think it would be worthwhile to give the reader some further specific ideas here in the Introduction where similar complex data may occur, e.g., data from traffic sensors, movement data (turnstiles in metro stations or public buildings), and even more traditional measurements such as temperatures or precipitation over time that could benefit from the methods and software described in this thesis.
 
-To do. (Medium)
+Done. Added in the Introduction.
 
-> Q2. It may also be helpful to state what is not be covered in this thesis: The spatial component of such data where an additional component could be latitude and longitude or some areal code,
-such as the post code or some administrative units.
+> Q2.
+It may also be helpful to state what is not be covered in this thesis: The spatial component of such data where an additional component could be latitude and longitude or some areal code, such as the post code or some administrative units.
 Extending the work from this thesis with spatial proximity methods for clustering (https://doi.org/10.1016/j.cageo.2011.12.017) [assuming that households in certain residential neighborhoods may have similar energy usage patterns] and trying spatial visualization methods (if possible) such as glyph maps (https:
-//onlinelibrary.wiley.com/doi/abs/10.1002/env.2152) should be mentioned somewhere as a consideration for future work | but that may become a future new thesis by itself.
+//onlinelibrary.wiley.com/doi/abs/10.1002/env.2152) should be mentioned somewhere as a consideration for future work but that may become a future new thesis by itself.
 
-To do. (Medium)
+The name of the thesis implies that only analyzing temporal data is covered.
+Extending it to spatial is too broad a scope.
+The limitations of the study cater to aspects that could be addressed with some more time.
+This is not the case for extending the spatial component, which, as accurately pointed out, would require time comparable to another thesis.
 
@@
 
-To do. (Difficult viz)
+Caption changed to illustrate the plot better.
+Added box borders in (a) such that (a) and (b) display the household ids in the same order as indicated by the same color of the line plot in (b) and box border in (a).
 

### 2022-02-24 Rmd/03-hakear.Rmd
Q17 (Juergen) and Q22(Caatherine) Ch3

--- 2022-02-23 Rmd/03-hakear.Rmd
@@
 (ref:dotplot-8) An ensemble plot with a heatmap (a), line plot (b), parallel coordinate plot (c) to demonstrate energy behavior of the households in different ways.
-Panel (b) shows the raw demand series for September to highlight the repetitive patterns of energy demand.
-Panel (a) shows $\wpd$ values across harmonies where a darker color indicates a higher ranking harmony.
+(a) and (b) display the household ids in the same order as indicated by the same colour of the line plot in (b) and box border in (a).
+(b) shows the raw demand series for September to highlight the repetitive patterns of energy demand.
+(a) shows $\wpd$ values across harmonies where a darker shade of red indicates a higher ranking harmony and gray represents a clash.
 A significant harmony is shown with an asterisk.
-For example, ids 7 and 8 have significant patterns across (*hod*, *dow*) and (*dow*, *hod*).
-Panel (c) is useful for comparing households across harmonies.
+For example, ids 7 and 8 have significant patterns across the pair (*hod*, *dow*) since the tile with *hod* as facet and *dow* as x variable has an asterisk.
+(c) is useful for comparing households across harmonies.
 For example, for the harmony (*dow*-*hod*), ids 1 and 7 have the least and highest $\wpd$ respectively.

### 2022-02-24 scripts/hakear.R

@@ -1007,7 +1007,7 @@ heatplot <- elec_sig_split %>%
   ggplot(aes(x = x_variable,
              y = facet_variable)) +
   geom_tile(aes(fill = wpd)) + 
-  geom_polygon(aes(x = x,y = y, color = id),size=2, fill= NA, data = merge(outline, square))+
+  geom_polygon(aes(x = x,y = y, color = id),size=1, fill= NA, data = merge(outline, square))+
   scale_color_discrete(type = okabeito,
                        guide = "none")+
   #color = as.factor(significance_95)),
@@ -1086,7 +1086,10 @@ elec_zoom <-  elec %>%
   theme(
     strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) +
   scale_color_discrete(type = okabeito) +
-  theme(axis.ticks.y=element_blank())
+  theme(axis.ticks.y=element_blank()) 
+
+# + 
+#   theme(strip.background =element_rect(aes(fill=id))) +scale_fill_discrete(type = okabeito)
 
 
 
@@ -1153,7 +1156,7 @@ elec_rank <- elec_sig_split %>%
 
 elec_rank %>% 
   kable(format = "markdown", 
-        caption = "Ranking of harmonies for the eight households with significance marked for different thresholds. Rankings are different and at most three harmonies are significant for any household. The number of harmonies to explore is reduced from 42 to 3.")
+        caption = "Ranking of harmonies for the eight households with significance marked for different thresholds. ***, ** and * correspond to the 99th, 95th and 90th percentile threshold. Rankings are different and at most three harmonies are significant for any household. The number of harmonies to explore is reduced from 42 to 3.")
 
 
 

### 2022-02-24 report/response.Rmd
last few comments

@@
 
-Added.
-@wertheimer1938gestalt
+Done.
 
@@
 
-To do.(Hard)
-
-Suppose there are $n$ observations.
-Our current methodology has a time complexity of $O(n^2)$ since it calculates distance between all pairs.
-The time taken to compute distances between a pair of observations based on $p$ cyclic granularity is
+A similar question has been answered for the next reviewer.
 
@@
 I am even more lost with hakear.
+
+Added under Thesis structure in Introduction.
+gravitas: GRAnularity VIsualization for Time-series AnalySis
+hakear: HArmonies KEeper And Rater
+gracsr: GRAnularity CluStering in R
 
@@
 
-To do. (Medium)
+We will include this in the supplements while publishing the paper.
 
@@
 
-To do. (Hard)
+Suppose there are $n$ observations.
+Our current methodology has a time complexity of $O(n^2)$ since it calculates distance between all pairs.
+The time taken to compute distances between a pair of observations is also based on the number of cyclic granularities considered.
+For example, $1$, $2$ and $3$ cyclic granularities for JS-NQT are ~ 4.5s, 5.8s and 8.2s respectively.
+The times will be similar for JS-RS.
+For WPD, time taken will be more as the statistic $wpd$ need to be computed for each observation and each cyclic granularity.
+A rough estimate of working with $wpd$ working with two granularities with $(l,m)$ levels: ~(30s- 300s for 6<=l,m<=50) and ~40min-50min for (l,m<=5).
+But this is a one-time job to obtain the significant granularities and can be run for a few individuals to get a sense of the significant granularities for the context.
+for example, for the electricity data, the granularities that were significant for most individuals were "wkdn/wday", "hour-of-day", "day-of-month" and hence clustering was run only on these variables.
 
@@
 We will include it in the journal paper. 
-Split into a "Limitation" and "Discussion" section separately detailing the drawbacks.
+<!-- Split into a "Limitation" and "Discussion" section separately detailing the drawbacks. -->
 

### 2022-02-24 Rmd/01-intro.Rmd
last few comments

@@
 Chapter \@ref(ch-gravitas) provides details of the cyclic granularities, different classes, and computation, and also its usage in exploratory time series analysis through applications.
-This is implemented in the R package `gravitas`.
+This is implemented in the R package `gravitas` (GRAnularity VIsualization for Time-series AnalySis).
 Chapter \@ref(ch-hakear) provides guidance on how to choose significant cyclic granularities, which are likely to have interesting patterns across its categories.
-This is available as the R package `hakear`.
+This is available as the R package `hakear` (HArmonies KEeper And Rater).
 Chapter \@ref(ch-gracsr) provides methods to explore heterogeneity in repetitive behavior for multiple time series over multiple cyclic granularities.
-This is in the developing R package `gracsr`.
+This is in the developing R package `gracsr` (Granularity CluStering in R).
 Chapter \@ref(ch-conclusion) summarizes the software tools developed for the work, and discusses future plans.
@@
 \hline
-4 & Clustering time series based on probability distributions across temporal granularities & To be submitted & 80\%.
+4 & Clustering time series based on probability distributions across temporal granularities & Working Paper & 80\%.
 Concept, methodology development, writing first draft and software development & (1) Dianne Cook, Concept, methodology development and writing 15\% (2) Rob J Hyndman, Concept, methodology development and writing 5\% & \multicolumn{1}{c|}{N} \\

### 2022-02-24 report/response.Rmd
fine tuning responses and incorporate missing comments

@@
 
-> Q3. In Figure 2.4, the outlier glyphs are in my opinion too big and dark and attract too much attention.
+> Q3. On page 23 the references to parts of Figure 2.4 need fixing.
+
+Reference is in an acceptable format. No change as already published.
+
+> Q4. In Figure 2.4, the outlier glyphs are in my opinion too big and dark and attract too much attention.
 The log scale for the y-axis should be mentioned in the caption.
@@
 
-> Q4. The violin plot in Figure 2.5c is barely recognisable, and comparison of the distributions is challenging. The plot deserves a bit more space.
+> Q5. The violin plot in Figure 2.5c is barely recognisable, and comparison of the distributions is challenging. The plot deserves a bit more space.
 
-Not changed.
-Figure is already published
+Not changed as figure is already published. Furthermore, the figure already occupies one page.
 
@@
 Done.
-Added "for a continuous univariate dependent variable" in the last paragraph of Introduction.
+Added "for a continuous univariate dependent variable" in the last paragraph of Introduction of Chapter 3.
 
@@
 
-No change. 
-Refer as months or month_year depending on the context.
+Done. x-axis labels changed to month-of-year for Figure 3.1 and 3.2.
 
@@
 
-CLARIFICATION: Null distributions refer to the case when there is no significant distributional differences between categories of one or more granularities.
+Clarification: Null distributions refer to the case when there is no significant distributional differences between categories of one or more granularities.
 Data in Fig 3.3 (a) is consistent with the null distribution.
@@
 
-Done.
+Clarification: $N_c$ refers to the total number of contextual cyclic granularities.
+$m$ refers to the number of cyclic granularities we are considering together in the display.
+For example, contextual cyclic granularities could be $hour_{day}$, $day_{week}$ and $month_{year}$ and we want to visualize any one granularity at a time.
+So $N_c = 3$ and $m=1$.
 
-Add $m$ to table 3.1
-$m$ = number of cyclic granularities to display together
-
-$N_c$ refers to the total number of contextual cyclic granularities.
-$m$ refers to the number of cyclic granularities we are considering together in the display.
-For example, contextual cyclic granularities could be $hour_{day}$, $day_{wee}k$ and $month_{year}$ and we want to visualize any one granularity at a time.
-So $N_c = 3$ and $m=1$.
+Added $m$ to table 3.1  
+$m$: number of cyclic granularities to display together
 
@@
 
-Done.
 Added v:continuous univariate measured variable
 
-> Q12. In the first sentence on the top of page 42, ordered and unordered are mixed up. Have you considered
-the setting where the facet variable levels are ordered? For the within-facet ordered distances, you
-could consider a distance measure that respects circular order, or choose the start level appropriate to
-the display.
+> Q12.
+In the first sentence on the top of page 42, ordered and unordered are mixed up.
+Have you considered the setting where the facet variable levels are ordered?
+For the within-facet ordered distances, you could consider a distance measure that respects circular order, or choose the start level appropriate to the display.
 
@@
 
-Done.
 Caption changed to incorporate that.
@@
 
-CLARIFICATION: As long as that there is enough data for each level or combination of levels, no adjustment for varying number of observations is done.
+Clarification: As long as that there is enough data for each level or combination of levels, no adjustment for varying number of observations is done.
 If there are few data for one or more levels, subsequent categories could be collapsed to obtain precise estimates of quantiles.
@@
 
-> Q15. The equations on page 44 could be tidied up. There are extra () on the definition of wpdperm and on the residual definition at the bottom of the page. Use $\times$ for the equations.
+> Q15. The equations on page 44 could be tidied up. There are extra () on the definition of $wpd_{perm}$ and on the residual definition at the bottom of the page. Use $\times$ for the equations.
 
@@
 
-CLARIFICATION: There is no practical explanation for this, but a more empirical one - with smaller $n_x$ and $n_f$, the variance in the values of $wpd_{raw}$ is greater.
+Clarification: There is no practical explanation for this, but a more empirical one - with smaller $n_x$ and $n_f$, the variance in the values of $wpd_{raw}$ is greater.
 As a result, our modeling technique that attempts to fit $wpd_{raw}$ to the number of comparisons ($n_x\times n_f$) performs poorly.
@@
 
-> Q17. In the 3.3.1 algorithm, there is m, and then M.
+> Q17. In the 3.3.1 algorithm, there is $m$, and then $M$.
 
@@
 The simulation design and results corresponding to $m=2$ are presented.
-Similar design and results for $m=1$, although important, are not included in the paper but in the supplements.
+Similar design and results for $m=1$, although important, are not included in the main chapter but in the supplements.
 Simulation results involving different underlying distributions are also presented in the supplementary paper.
@@
 
-Done.
+Changed. Axis tick labels made smaller and space between facets minimised to allow for more space.
 
@@
 
-Done.
+Changed. Axis tick labels made smaller and the rug plots are removed to make the display less busy.
 
@@
 
-CLARIFICATION 1: Using color to convey significance instead of stars (in the Fig 3.3a) is a not a great idea because the varied shades of the tiles represent different values of the $wpd$.
+Clarification: Using color to convey significance instead of stars (in the Fig 3.3a) is a not a great idea because the varied shades of the tiles represent different values of the $wpd$.
 For example, both tiles may be significant, and their colours may differ to represent distinct $wpd$ values.
@@
 
-> Q23. The link in the Acknowledgements https://github.com/Sayani07/paper-hakear is not available. Neither
-are the supplementary materials. I understand this is for the paper version, I mention it for completeness.
+> Q23.
+The link in the Acknowledgements https://github.com/Sayani07/paper-hakear is not available.
+Neither are the supplementary materials.
+I understand this is for the paper version, I mention it for completeness.
 
-Changed.
-The link works. The repository has now been made public.
+The Github repository has now been made public. The link works now.
 
@@
 
-Done.
 Rephrased to make it clear.
@@
 
-Done.
 Added the statement in page 60.
@@
 In each box, would be helpful to put in “or”, in places where only one of the steps listed is performed, eq Normal quantile transform or Robust scaling.
-As there are many steps
-in the algorithm, it would be helpful to the reader to label the pipeline steps and to refer back to them in the text. The text in the Data pre-processing box does not make sense to me.
+As there are many steps in the algorithm, it would be helpful to the reader to label the pipeline steps and to refer back to them in the text.
+The text in the Data pre-processing box does not make sense to me.
 
-Rephrased to make it fluent.
+Rephrased the text in the  Data pre-processing section to make it easier to understand.
+
 Figure 4.1 labels changed to make it consistent with the text.
@@
 
-Done.
 Clarification: Each observation (time series) is scaled independently in both JS-NQT and JS-RS.
@@
 
-What is the question? I have mentioned that D stands for Jensen-Shannon distances.
+No change. I had mentioned that $D$ stands for Jensen-Shannon distances.
 
@@
 
-Stating the distribution of each $v_t$ is difficult because the levels for each granularity and design varies, therefore I chose to divide it into two tables.
+No change. Stating the distribution of each $v_t$ is difficult because the levels for each granularity and design varies, therefore I chose to divide it into two tables.
 
@@
 
-Done.
-Clarification: $\mu$ is the difference between means considered for consecutive categories.
+Clarification: $\mu$ is the difference between means considered for consecutive categories. It will vary depending on the chosen design and hence not tabulated.
 
@@
 
-Done.
-Clarification: All
+Clarification: Various range of values of $R$ and $T$ are considered as stated in Table 4.1.
 
@@
 
-MDS plot summary of the data sets from the validation study.
+CLARIFICATION: MDS summary plots of clusters obtained from employing JS-NQT on the validation data sets
+Added in the caption.
 
@@
 
-Done.
+Caption and text changed.  
 The caption and description of the figure in the text have been changed to indicate that Fig.
@@
 
-> Q15. The reference to the stationarity assumption in the Discussion needs clarification.
+> Q15. The reference to the stationarity assumption in the Discussion needs Clarification.
 
-Done.
-Clarification: "we have assumed the time series to be stationary" has been changed to "the conditional distributions are assumed to remain constant for the observation period" in the Discussion section.
+Changed. "we have assumed the time series to be stationary" has been changed to "the conditional distributions are assumed to remain constant for the observation period" in the Discussion section.
 
@@
 
-Done. Added in the Introduction.
+Added in the Introduction.
 
@@
 
-> Q3. Finally, it would be helpful to indicate here (and in each chapter) what the names or abbreviations for the three R packages (gravitas, hakear, and gracsr) represent. Even the full R package
-title of "gravitas: Explore Probability Distributions for Bivariate Temporal Granularities" does not provide a full answer what the "tas" component means (I am assuming that Gravi stands for Granularities Visualization).
+> Q3.
+Finally, it would be helpful to indicate here (and in each chapter) what the names or abbreviations for the three R packages (gravitas, hakear, and gracsr) represent.
+Even the full R package title of "gravitas: Explore Probability Distributions for Bivariate Temporal Granularities" does not provide a full answer what the "tas" component means (I am assuming that Gravi stands for Granularities Visualization).
 I am even more lost with hakear.
@@
 > Q2.
-how to define and handle pay day (or loan day) which often is defined as the 1st (and/or possibly 15th) work day of a month, resulting in some aperiodic granularity as some of these days across the year could initially fall on a weekend, but even more in a few
-cases, may coincide with a public holiday (and thus shifting pay day by 2 or 3 days).
+how to define and handle pay day (or loan day) which often is defined as the 1st (and/or possibly 15th) work day of a month, resulting in some aperiodic granularity as some of these days across the year could initially fall on a weekend, but even more in a few cases, may coincide with a public holiday (and thus shifting pay day by 2 or 3 days).
 Similarly, extended weekends that could stretch from Fridays to Sundays or from Saturdays to Mondays (and possibly even over a 4{day period) might be worth some discussion, in particular as a case of variable{length aperiodic granularity.
@@
 
-Done. (Can't get the tick marks at the right place)
+Done.
 
@@
 > Q5.
-Before introducing a new distance measure in Section 3.2, I would like to see a literature review of existing distance measures | and their limitations.
+Before introducing a new distance measure in Section 3.2, I would like to see a literature review of existing distance measures and their limitations.
 Why is it necessary to introduce a new distance measure here?
@@
 
-CLARIFICATION 1:
+Clarification 1:
 The goal is to determine if there are any statistically significant differences between independent (unrelated) groups similar to a one-way or two-way ANOVA.
@@
 
-Done.
 Made it past tense.
@@
 
-Done.
 Mentioned which table or subsection needs to be referred and the link of the supplementary material provided.
@@
 
-Done.
-CLARIFICATION: We have assumed both $nx$; $nf$ simultaneously to be less than or equal to 5
+Clarification: We have assumed $nx$; $nf$ simultaneously to be less than or equal to 5
 
@@
 
-CLARIFICATION: Added $m=2$ in the caption of Figures 3.5 and 3.6.
+Clarification: Added $m=2$ in the caption of Figures 3.5 and 3.6.
 
@@
 
-Done.
 Added the abbreviations on Page 50 in "Choosing cyclic granularities of interest and removing clashes" section.
@@
 
-Done.
-
-CLARIFICATION 1: Each household is represented by 25 tiles, each tile representing a pair of cyclic granularities.
+Clarification 1: Each household is represented by 25 tiles, each tile representing a pair of cyclic granularities.
 The colors (in shades of red) represent the value of $wpd$ for each of the harmony pairs (in Table Table 3.3) and the grey tiles correspond to clashes.
@@
 
-CLARIFICATION 2: The harmony pairs in the parallel coordinate plot are arranged from highest to lowest values of $wpd$ averaged over all households and hence do not correspond to the first or last rows of Table 3.3.
+Clarification 2: The harmony pairs in the parallel coordinate plot are arranged from highest to lowest values of $wpd$ averaged over all households and hence do not correspond to the first or last rows of Table 3.3.
 
@@
 
-Done.
 Changed to "insignificant"
@@
 
+Clarification 1: The data for eight households for Section 3.4 are from friends and colleagues.
+They were manually chosen from friends with different profession and lifestyle to reflect different energy behavior.
+
 Suppose there are $n$ observations.
-Our current methodology has a time complexity of $O(n^2)$ since it calculates distance between all pairs.
+Our current clustering methodology has a time complexity of $O(n^2)$ since it calculates distance between all pairs.
 The time taken to compute distances between a pair of observations is also based on the number of cyclic granularities considered.
@@
 For WPD, time taken will be more as the statistic $wpd$ need to be computed for each observation and each cyclic granularity.
-A rough estimate of working with $wpd$ working with two granularities with $(l,m)$ levels: ~(30s- 300s for 6<=l,m<=50) and ~40min-50min for (l,m<=5).
+A rough estimate of working with $wpd$ working with two granularities with $(l,m)$ levels: ~(30s - 300s for 6<=l,m<=50) and (~40min - 50min for (l,m<=5).
 But this is a one-time job to obtain the significant granularities and can be run for a few individuals to get a sense of the significant granularities for the context.
@@
 
-Done.
 Rephrased to make it clear that most techniques mentioned in Tureczek and Nielsen (2017) do not treat smart meter data as a data type with temporal component.
@@
 
-Done. 
 Added links in the supplementary materials section at the end of the chapter.
@@
 Done.
-Not showing in the plot, check again.
 
@@
 
-No change.
-Showing the vertical axis would not give us any more information than it is already giving.
+No change. Showing the vertical axis would not give us any more information than it is already giving.
 
@@
 
-Done.
+Done. Used "wnwd" through out Chapters 3 AND 4.
 
@@
 Changing from red to blue/purple initially hides this information.
-Also arrange the legend from P-1 to P-3 and Q-1 to Q-5 to make it more obvious that P-1 & Q-1 and P-2 & Q-2 are identical
+Also arrange the legend from P-1 to P-3 and Q-1 to Q-5 to make it more obvious that P-1 & Q-1 and P-2 & Q-2 are identical.
 

### 2022-02-24 Rmd/04-gracsr.Rmd
fine tuning responses and incorporate missing comments

@@
 
-```{r mds-plot-validation, fig.cap = "MDS summary plots to illustrate the cluster separation for the range of mean differences (rows) under the different scenarios (columns).
+```{r mds-plot-validation, fig.cap = "MDS summary plots of clusters obtained from employing JS-NQT on the validation data sets.
+It illustrates the cluster separation for the range of mean differences (rows) under the different scenarios (columns).
 It can be observed that clusters become more compact and separated for higher mean differences between categories across all scenarios.

### 2022-02-24 scripts/gracsr/validation.R

@@ -1015,7 +1015,7 @@ mds_plot_10 <- ggplot(
     color = group
   )
 ) +
-  geom_point(size = 2, alpha = 0.5, shape = 3) +
+  geom_point(size = 2, alpha = 0.5, aes(shape = group)) +
   facet_grid(diff ~ Scenario, labeller = "label_value") +
   # geom_text(check_overlap = TRUE)  +
   theme_bw() +
@@ -1023,8 +1023,12 @@ mds_plot_10 <- ggplot(
   scale_color_brewer(palette = "Dark2") +
   ylab("mds1") +
   xlab("mds2") +
-  theme(axis.text = element_text(size = 8))+
-  theme(aspect.ratio = 1)
+  theme(axis.text = element_text(size = 7),
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
+  theme(aspect.ratio = 1)+
+  scale_shape_manual(values = c(3, 4, 2, 0, 1)) + 
+  theme(panel.spacing.x=unit(0, "lines"),
+        panel.spacing.y=unit(0.3, "lines"))
 
 mds_plot_10
 

### 2022-02-24 scripts/hakear.R

@@ -71,6 +71,7 @@ p1 <- id2_tsibble %>%
     strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + 
   scale_colour_brewer(name = "", palette = "PiYG") + 
   ylab ("energy usage [kWh]") +
+  xlab("hour-of-day") +
   scale_y_log10(breaks = scales::trans_breaks("log10", 
                                               function(x) 10^x),
                  labels = scales::trans_format("log10", scales::math_format(10^.x))) +
@@ -165,7 +166,7 @@ id2_tsibble_hd <- elec %>%
   geom_jitter(alpha = 0.03, colour = "#E69F00", position = position_jitter(height = .2, width = .2)) +
   ggtitle("(a)")+ 
   ylab ("energy usage [kWh]") +
-  theme_bw()
+  theme_bw() + xlab("hour-of-day")
 
 
 
@@ -178,7 +179,7 @@ id2_tsibble_dw <- elec %>%
   geom_jitter(alpha = 0.03, colour = "#E69F00", position = position_jitter(height = .2, width = .2)) +
   ggtitle("(b)")+ 
   ylab ("energy usage [kWh]") +
-  theme_bw()
+  theme_bw() + xlab("month-of-year")
 
 
 
@@ -1156,7 +1157,7 @@ elec_rank <- elec_sig_split %>%
 
 elec_rank %>% 
   kable(format = "markdown", 
-        caption = "Ranking of harmonies for the eight households with significance marked for different thresholds. ***, ** and * correspond to the 99th, 95th and 90th percentile threshold. Rankings are different and at most three harmonies are significant for any household. The number of harmonies to explore is reduced from 42 to 3.")
+        caption = "Ranking of harmonies for the eight households with significance or different thresholds. ***, ** and * correspond to the 99th, 95th and 90th percentile threshold. Rankings are different and at most three harmonies are significant for any household. The number of harmonies to explore is reduced from 42 to 3.")
 
 
 

### 2022-02-25 Rmd/03-hakear.Rmd
Ch3 Q3 Juergen changed

@@
 Panel (a) is not interesting as the distribution of the variable does not depend on x or facet categories.
+The data set in (a) is consistent with the null distribution.
 Panels (b) and (c) are more interesting than (a) since there is a change in distribution either across facets (b) or x-axis (c).
@@
 For example, a $(2, 3)$ panel will have cyclic granularities with two x-axis levels and three facet levels.
-Let the four elementary designs as described in Figure \ref{fig:null4by2} be $\Dnull$ (referred to as "null distribution") where there is no difference in distribution of $v$ for $A$ or $B$, $\Df$ denotes the set of designs where there is difference in distribution of $v$ for $B$ and not for $A$.
+Let the four elementary designs as described in Figure \ref{fig:null4by2} be $\Dnull$ (referred to as "null distribution") where there is no difference in distribution of $v$ for $A$ or $B$.
+For example, the data set in Figure \ref{fig:null4by2}(a) is consistent with the null distribution.
+$\Df$ denotes the set of designs where there is difference in distribution of $v$ for $B$ and not for $A$.
 Similarly, $\Dx$ denotes the set of designs where difference is observed only across A.

### 2022-02-25 scripts/hakear.R

@@ -73,9 +73,9 @@ p1 <- id2_tsibble %>%
   ylab ("energy usage [kWh]") +
   xlab("hour-of-day") +
   scale_y_log10(breaks = scales::trans_breaks("log10", 
-                                              function(x) 10^x),
-                 labels = scales::trans_format("log10", scales::math_format(10^.x))) +
-  theme_bw()
+                                              function(x) 10^x)) +
+                 #labels = scales::trans_format("log10", scales::math_format(10^.x))) +
+  theme_bw() + ylim(0, 1.5)
 
 
 p2 <- id4_tsibble %>%
@@ -89,9 +89,9 @@ p2 <- id4_tsibble %>%
     strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))+
   ylab ("energy usage [kWh]")+
   scale_y_log10(breaks = scales::trans_breaks("log10", 
-                                              function(x) 10^x),
-                labels = scales::trans_format("log10", scales::math_format(10^.x))) +
-  theme_bw()
+                                              function(x) 10^x))+
+                #labels = scales::trans_format("log10", scales::math_format(10^.x))) +
+  theme_bw() + ylim(0, 1.5)
 
 
 p3 <- id2_tsibble %>%

### 2022-02-25 scripts/hakear.R
Ch3 Q1 Juergen changed

@@ -162,11 +162,18 @@ id2_tsibble_hd <- elec %>%
   as_tsibble(index = date_time) %>% 
   create_gran("hour_day") %>% 
   ggplot(aes(x = hour_day, y = kwh)) +
-  geom_boxplot(width = 0.6, outlier.colour = "black", outlier.alpha = 0.2, fill = "#CC79A7", colour =  "#0072B2") + scale_y_log10() +
+  geom_boxplot(width = 0.6, outlier.colour = "black", outlier.alpha = 0.2, fill = "#CC79A7", colour =  "#0072B2")  +
   geom_jitter(alpha = 0.03, colour = "#E69F00", position = position_jitter(height = .2, width = .2)) +
   ggtitle("(a)")+ 
   ylab ("energy usage [kWh]") +
-  theme_bw() + xlab("hour-of-day")
+  theme_bw() + xlab("hour-of-day")+
+  scale_y_log10(minor_breaks = NULL) +
+  theme_bw()
+
+  # scale_y_log10(breaks = scales::trans_breaks("log10", 
+  #                                             function(x) 10^x)) +
+  # #labels = scales::trans_format("log10", scales::math_format(10^.x))) +
+  # theme_bw()
 
 
 
@@ -175,11 +182,14 @@ id2_tsibble_dw <- elec %>%
   as_tsibble(index = date_time) %>% 
   create_gran("month_year") %>% 
   ggplot(aes(x = month_year, y = kwh)) +
-  geom_boxplot(width = 0.6, outlier.colour = "black", outlier.alpha = 0.2, fill = "#CC79A7", colour =  "#0072B2") + scale_y_log10()+
+  geom_boxplot(width = 0.6, outlier.colour = "black", outlier.alpha = 0.2, fill = "#CC79A7", colour =  "#0072B2") +
   geom_jitter(alpha = 0.03, colour = "#E69F00", position = position_jitter(height = .2, width = .2)) +
   ggtitle("(b)")+ 
   ylab ("energy usage [kWh]") +
-  theme_bw() + xlab("month-of-year")
+  theme_bw() + xlab("month-of-year")+
+  scale_y_log10(minor_breaks = NULL) +
+  #labels = scales::trans_format("log10", scales::math_format(10^.x))) +
+  theme_bw()
 
 
 

### 2022-02-25 report/response.Rmd
Ch3 Q22 Juergen changed

@@
 
-Not changed.
-Figure is already published.
+Done.
 
@@
 
+Added text in the caption of Figure 3.3(a) and "Notation" sub-section to make it clear.
+
 Clarification: Null distributions refer to the case when there is no significant distributional differences between categories of one or more granularities.
@@
 
+Added $m$ to table 3.1  
+$m$: number of cyclic granularities to display together
+
 Clarification: $N_c$ refers to the total number of contextual cyclic granularities.
@@
 So $N_c = 3$ and $m=1$.
-
-Added $m$ to table 3.1  
-$m$: number of cyclic granularities to display together
 
@@
 
+Added under the subsection "Characterizing distributions".
+
 Clarification: As long as that there is enough data for each level or combination of levels, no adjustment for varying number of observations is done.
 If there are few data for one or more levels, subsequent categories could be collapsed to obtain precise estimates of quantiles.
-
-Added under the subsection "Characterizing distributions".
 
@@
 
-The simulation design and results corresponding to $m=2$ are presented.
+Rephrased the simulation design and results corresponding to $m=2$.
 Similar design and results for $m=1$, although important, are not included in the main chapter but in the supplements.
@@
 
-The ticmark labels are placed in both the figures but the houses are not brought to a same scale.
+Done.
+
+<!-- The ticmark labels are placed in both the figures but the houses are not brought to a same scale.
 More than comparing the two houses, it is important to compare the patterns of their distributions across months.
 They have different scales and showing them in the same scale is making the individual distributional differences seem less clear.
+-->
 
@@
 
-Clarification 1:
-The goal is to determine if there are any statistically significant differences between independent (unrelated) groups similar to a one-way or two-way ANOVA.
+Changed the following:
+
+- Introduction (last but one paragraph) to incorporate comments on existing distance measures and their limitations
+
+- "Distance between distributions" section to mention why we chose we chose Jensen-Shannon distances and some details about other distance measures.
+
+- f in the equation was a typo. Replaced by log now.
+
+<!-- Clarification 1: -->
+<!-- The goal is to determine if there are any statistically significant differences between independent (unrelated) groups similar to a one-way or two-way ANOVA.
 A one-way ANOVA, for example, is analogous to $m=1$ (looking at one cyclic granularity at a time) in our approach.
@@
 There are other ways to obtain distributional differences, but we chose Jensen-Shannon distances between quantiles for ease of computation.
+-->
 
-CALRIFICATION 2: f in the equation was a typo. Replaced by log now.
+<!-- CALRIFICATION 2:  -->
 
-CALRIFICATION 3: Distance measures that are mentioned at the end of that paragraph are not explained in details but mentioned that they are all special cases of f-divergence.
-
-CHANGED: Introduction (last but one paragraph) and Distance between distributions section.
+<!-- CALRIFICATION 3: Distance measures that are mentioned at the end of that paragraph are not explained in details but mentioned that they are all special cases of f-divergence. -->
 
@@
 
-Clarification 1: The data for eight households for Section 3.4 are from friends and colleagues.
-They were manually chosen from friends with different profession and lifestyle to reflect different energy behavior.
+- To reflect various energy behaviour, data for eight homes for Section 3.4 is acquired from friends and colleagues with different professions, lifestyles, and household sizes.
+This level of data is private, and we thank our friends and colleagues for agreeing to have their data used for this analysis.
+This is added in the first paragraph of Application.
 
-Suppose there are $n$ observations.
+- A paragraph is added to the Discussion section to incorporate comments on on the computational time.
+
+<!-- Suppose there are $n$ observations.
 Our current clustering methodology has a time complexity of $O(n^2)$ since it calculates distance between all pairs.
@@
 for example, for the electricity data, the granularities that were significant for most individuals were "wkdn/wday", "hour-of-day", "day-of-month" and hence clustering was run only on these variables.
+-->
 

### 2022-02-25 Rmd/03-hakear.Rmd
Ch3 Q22 Juergen changed

@@
 
-The smart meter data set for eight households in Melbourne was procured by downloading the data from the energy supplier/retailer.
+The smart meter data set for eight households in Melbourne was procured by downloading individual level data from energy suppliers/retailers.
+To reflect various energy behaviours, data was obtained from friends and colleagues with different professions, lifestyles, and household sizes.
+This level of data is private, and we thank our friends and colleagues for agreeing to have their data used for this analysis.
 The data has been cleaned to form a `tsibble` [@wang2020tsibble] containing half-hourly electricity consumption from July to December 2019 for each of the households.

### 2022-02-27 Rmd/04-gracsr.Rmd
fine tuning responses and Q13 Ch4 Juergen

@@
 
+### Conclusions
+
+Clustering methods perform effectively when clusters are well-separated and there are no nuisance variables or observations.
+In practice, most problems will contain nuisance observations or variables, and well-separated clusters are uncommon.
+As a result, picking a small group of observations with somewhat diverse and distinctive patterns (prototypes) rather than a brute force method is usually a good way to learn about the clustering and the data set, as demonstrated in this chapter.
 We offer two approaches for calculating pairwise distances between time series based on probability distributions over multiple cyclic granularities at once.
-Depending on the goal of the clustering, these distance metrics, when fed into a hierarchical clustering algorithm using Ward's linkage, yield meaningful clusters.
+Depending on the goal of the clustering, these distance metrics, when fed into a hierarchical clustering algorithm using Ward's linkage, yield meaningful clusters for the prototypes.
 Probability distributions provide an intuitive method to characterize noisy, patchy, long, and unequal-length time series data.
@@
 Furthermore, unlike earlier efforts that group customers based on behavior across only one cyclic granularity (such as hour-of-day), our method is more comprehensive in detecting clusters with repeated patterns at all relevant granularities.
+
 <!--The JS-based approaches are faster than wpd-based approaches.-->
 
+### Limitations and future research
+
 There are a few areas to extend this research.
-First, larger data sets with more uncertainty complicate matters, as is true for any clustering task.
+Firstly, larger data sets with more uncertainty complicate matters, as is true for any clustering task.
 Characterizing clusters with varied or outlying customers can result in a shape that does not represent the group.
@@
 In our application, we have removed those customers and done the analysis but the metrics used should be able to incorporate those customers with such structured missingness.
-Finally, $\wpd$ is computationally heavy even under parallel computation.
-Future work can make the computations more efficient so that they are easily scalable to a large number of customers.
 Moreover, experiments can also be run with non-hierarchy based clustering algorithms to verify if these distances work better with other algorithms.
+
+  For larger data sets, most clustering tasks become increasingly computationally intensive.
+The computational time for determining pairwise distances presented in this paper is determined by the number of cyclic granularities evaluated.
+For JS-NQT, the time required to compute distances between two observations based on $1$, $2$, and $3$ cyclic granularities is 4.5s, 5.8s, and 8.2s, respectively.
+For JS-RS, the time is comparable.
+The time required for WPD will be longer because the statistic $wpd$ must be computed for each observation and cyclic granularity.
+Even with parallel computing, $wpd$ is computationally costly for cyclic granularities with fewer levels.
+Further research may improve the efficiency of the computations to tackle large-scale clustering challenges.
 

### 2022-02-27 report/response.Rmd
fine tuning responses and Q13 Ch4 Juergen
Q22 Ch3 Juergen

@@
 
-For id: 7, there is more distributional difference between consecutive hours of the day compared to id:1.
+<!-- For id: 7, there is more distributional difference between consecutive hours of the day compared to id:1.
 Also all the days of the week are different from each other in id:7.
@@
 The display with hod vs wdwnd would have shown higher values of wpd for id:1 (as can be verified from Table 3).
+-->
 
@@
 Was this based on a random sample, or manually do obtain rather different energy usage patterns?
-This should be answered earlier on, but this also should be addressed in the Discussion:
-What can be expected when working with the full data set of Chapter 2, i.e., about 13,000. You indicate:
+This should be answered earlier on, but this also should be addressed in the Discussion: What can be expected when working with the full data set of Chapter 2, i.e., about 13,000.
+You indicate:
 A future direction of work is to be able to explore and compare many individuals/subjects together for similar patterns across significant granularities.
@@
 
-- To reflect various energy behaviour, data for eight homes for Section 3.4 is acquired from friends and colleagues with different professions, lifestyles, and household sizes.
+- To reflect various energy behavior, data for eight homes for Section 3.4 is acquired from friends and colleagues with different professions, lifestyles, and household sizes.
 This level of data is private, and we thank our friends and colleagues for agreeing to have their data used for this analysis.
@@
 
-- A paragraph is added to the Discussion section to incorporate comments on on the computational time.
+- A paragraph is added to the Discussion section to incorporate comments on the computational time.
 
@@
 
-We will include it in the journal paper. 
+The Discussion section is split into (1) conclusions, and (2) limitations and future research (including a discussion on computational complexity) to address this comment.
+
 <!-- Split into a "Limitation" and "Discussion" section separately detailing the drawbacks. -->

### 2022-02-27 Rmd/03-hakear.Rmd
Q22 Ch3 Juergen

@@
 
+_Computational time _
+
+<!-- A rough estimate of working with $\wpd$ working with two granularities with $(l,m)$ levels: ~(30s - 300s for 6<=l,m<=50) Computing $\wpd$ is computationally more expensive for cyclic granularities with lower levels $\nx,\nf<=5$.
+-->
 <!-- patterns are interesting and others are not.
@@
 to group subjects -->
+The computation time of $\wpd$ for a cyclic granularity or harmony is determined by its levels.
+For example, in a harmony with $6 \leq\nx,\nf \leq 50$, the computation time grows with the number of levels and ranges from $30-300$s.
+$\wpd$ becomes computationally more expensive for cyclic granularities with lower levels $\nx,\nf \leq 5$.
+It is around $40-50$ minutes for $\nx,\nf = \{2,3\}$.
+Even with parallel computation, computing $\wpd$ for many individuals and cyclic granularity/harmonies takes a long time.
+One solution could be to obtain the significant granularities for a smaller number of individuals in order to get a sense of the significant granularities for the context and then computing $\wpd$ only for those granularities for other individuals.
+<!--For example, in the case of electricity data, the important granularities for most people were *wnwd*, *hod*, *dow*, and *dom*.
+After this fact has been established for a few people, $\wpd$ can be computed exclusively for these granularities to save time for others.--> However, information may be lost for individuals with different sets of significant granularities.
+In any case, additional studies may be conducted to speed up the processing of cyclic granularities with low levels.
 

### 2022-02-27 Rmd/04-gracsr.Rmd
Q9 Ch4 Catherine

@@
 To validate our clustering methods, we have created several different data designs containing different granularity features.
-<!--to see where one method works better than the other and where they might give us the same outcome --or the effect of missing data <!---and trends on the proposed methods.--> There are three circular granularities $g1$, $g2$ and $g3$ with categories denoted by $\{g10,g11\}$, $\{g20, g21, g22\}$ and $\{g30, g31, g32, g33, g34\}$ and levels $n_{g_1}=2$, $n_{g_2}=3$ and $n_{g_3}=5$.
+<!--to see where one method works better than the other and where they might give us the same outcome --or the effect of missing data <!---and trends on the proposed methods.--> There are three circular granularities $g1$, $g2$ and $g3$ with categories denoted by $\{g1_{0},g1_{1}\}$, $\{g2_{0}, g2_{1}, g2_{2}\}$ and $\{g3_{0}, g3_{1}, g3_{2}, g3_{3}, g3_{4}\}$ and levels $n_{g_1}=2$, $n_{g_2}=3$ and $n_{g_3}=5$.
 These categories could be integers or some more meaningful labels.

### 2022-02-27 scripts/gracsr.R

@@ -312,9 +312,9 @@ tab_dist <- tibble::tibble(granularity = granularity, `Varying distributions` = 
 
 ##----tab-design
 design = c("design-1", "design-2", "design-3", "design-4", "design-5")
-g1 = c("fixed", "vary", "fixed", "fixed", "vary")
-g2 = c("fixed", "fixed", "vary", "fixed", "vary")
-g3 = c("fixed", "fixed", "fixed", "vary", "vary")
+g1 = c("(0,0)", "(0,2)", "(0,0)", "(0,0)", "(0,2)")
+g2 = c("(0,0,0)", "(0,0,0)", "(2,1,0)", "(0,0,0)", "(2,1,0)")
+g3 = c("(0,0,0,0,0)", "(0,0,0,0,0)", "(0,0,0,0,0)", "(0,1,2,1,0)", "(0,1,2,1,0)")
 table <- tibble(design, g1, g2, g3)
 
 #%>% kable(caption = "5 different designs resulting from considering different distributions across categories.")
@@ -333,7 +333,7 @@ table <- tibble(design, g1, g2, g3)
 # ) %>% kable_styling()
 
 knitr::kable(
-  list(tab_dist,  tibble(design, g1, g2, g3)),
+  table,
   caption = 'For Scenario (a), distributions of different categories when they vary (top). If distributions are fixed, they are set to N(0, 1). 5 designs resulting from different distributions across categories (below)',
   booktabs = TRUE, valign = 't'
 )

### 2022-02-27 scripts/gracsr/validation.R

@@ -18,9 +18,10 @@ tab_dist <- tibble::tibble(
 
 ## ----tab-design-------------------------------------------------------------------------
 design <- c("design-1", "design-2", "design-3", "design-4", "design-5")
-g1 <- c("fixed", "vary", "fixed", "fixed", "vary")
-g2 <- c("fixed", "fixed", "vary", "fixed", "vary")
-g3 <- c("fixed", "fixed", "fixed", "vary", "vary")
+g1 = c("(0,0)", "(0,2)", "(0,0)", "(0,0)", "(0,2)")
+g2 = c("(0,0,0)", "(0,0,0)", "(2,1,0)", "(0,0,0)", "(2,1,0)")
+g3 = c("(0,0,0,0,0)", "(0,0,0,0,0)", "(0,0,0,0,0)", "(0,1,2,1,0)", "(0,1,2,1,0)")
+
 table <- tibble(design, g1, g2, g3)
 
 
@@ -28,9 +29,8 @@ table <- tibble(design, g1, g2, g3)
 ## ----tab-dist-design--------------------------------------------------------------------
 
 knitr::kable(
-  list(tab_dist, tibble(design, g1, g2, g3)),
-  caption = "For S1, distributions of different categories when they vary (displayed on top). If distributions are fixed, they are set to N(0, 1). The various distributions across categories result in five designs (displayed below).",
-  booktabs = TRUE, valign = "t"
+  table,
+  caption = "Summary of the data generating process for S1. The various distributions across levels of different granularities result in five designs. The granularities g1, g2 and g3 have 2,3 and 5 levels, each of which follows a normal distribution. The means of the distribution are presented for each level, and the standard deviation is set to 1. For example, g1 in design-2 is (0,2) implies $g1_{1} \\sim {N(0,1)}$ and $g1_{2} \\sim {N(2,1)}$.",booktabs = TRUE, valign = "t", escape= FALSE
 ) %>%
   kable_styling(latex_options = "hold_position")
 

### 2022-02-27 report/response.Rmd
Q9 Ch4 Catherine
final pages added on response document

@@
 
-Done.
-A table outlining the main chapters’ co-authorship arrangements included in Introduction.
+A table outlining the main chapters’ co-authorship arrangements included in Introduction on page 5.
 
@@
 
-Done. 
 The outliers are made smaller and transparent and caption changed.
@@
 
-Done.
-Added "for a continuous univariate dependent variable" in the last paragraph of Introduction of Chapter 3.
+Added "for a continuous univariate dependent variable" in the last paragraph of Introduction of Chapter 3 on page 39.
 
@@
 
-Added text in the caption of Figure 3.3(a) and "Notation" sub-section to make it clear.
+Added text in the caption of Figure 3.3(a) and Section 3.2.2 (Notation) on page 41 to make it clear.
 
-Clarification: Null distributions refer to the case when there is no significant distributional differences between categories of one or more granularities.
+<!-- Clarification: Null distributions refer to the case when there is no significant distributional differences between categories of one or more granularities.
 Data in Fig 3.3 (a) is consistent with the null distribution.
+-->
 
@@
 
-Added v:continuous univariate measured variable
+Added v:continuous univariate measured variable to the table on page 42.
 
-> Q12.
-In the first sentence on the top of page 42, ordered and unordered are mixed up.
-Have you considered the setting where the facet variable levels are ordered?
-For the within-facet ordered distances, you could consider a distance measure that respects circular order, or choose the start level appropriate to the display.
+> Q12. In the first sentence on the top of page 42, ordered and unordered are mixed up. Have you considered the setting where the facet variable levels are ordered?
 
-first part: Done.
-second part: This is a good suggestion. We will include that into future work.
+Done.
+
+> For the within-facet ordered distances, you could consider a distance measure that respects circular order, or choose the start level appropriate to the display.
+
+This is a good suggestion. We will include that into future work.
 
@@
 
-Caption changed to incorporate that.
-Within-facet distances are illustrated in Panels 3) (when categories are un-ordered, shown only with respect to a1) and Panel 4) (when categories are ordered)
+Caption changed to address this comment.
+
+<!-- Within-facet distances are illustrated in Panels 3) (when categories are un-ordered, shown only with respect to a1) and Panel 4) (when categories are ordered) -->
 
@@
 
-Added under the subsection "Characterizing distributions".
+Added text in the section 3.2.3 under subsection "Characterizing distributions" on page 43 to address this comment.
 
-Clarification: As long as that there is enough data for each level or combination of levels, no adjustment for varying number of observations is done.
+<!-- Clarification: As long as that there is enough data for each level or combination of levels, no adjustment for varying number of observations is done.
 If there are few data for one or more levels, subsequent categories could be collapsed to obtain precise estimates of quantiles.
+-->
 
@@
 
-> Q16. Is there a practical reason why $wpd_{glm}$ is not working as expected for lower nx and nf?
+> Q16. Is there a practical reason why $wpd_{glm}$ is not working as expected for lower $nx$ and $nf$?
 
-Clarification: There is no practical explanation for this, but a more empirical one - with smaller $n_x$ and $n_f$, the variance in the values of $wpd_{raw}$ is greater.
+Added the empirical reason in Section 3.2.4 under sub-section "Combination approach" on page 47.
+
+<!-- Clarification: There is no practical explanation for this, but a more empirical one - with smaller $n_x$ and $n_f$, the variance in the values of $wpd_{raw}$ is greater.
 As a result, our modeling technique that attempts to fit $wpd_{raw}$ to the number of comparisons ($n_x\times n_f$) performs poorly.
 The increased volatility in the values of $wpd_{raw}$ with a smaller number of levels is due to the fact that no asymptotic distribution for the statistic "maximum" (which is the foundation of $wpd_{raw}$) across such small number of values could be produced.
+-->
 
@@
 
-Changed.
-1.b) for $i \in {1, 2, \dots, M}$.
+Changed on page 49.
+<!-- 1.b) for $i \in {1, 2, \dots, M}$. -->
 
@@
 
-Rephrased the simulation design and results corresponding to $m=2$.
+Rephrased the simulation design and results corresponding to $m=2$ on page 50.
 Similar design and results for $m=1$, although important, are not included in the main chapter but in the supplements.
@@
 
-Changed. Axis tick labels made smaller and space between facets minimised to allow for more space.
+Changed (page 51). Axis tick labels made smaller and space between facets minimised to allow for more space.
 
@@
 
-Changed. Axis tick labels made smaller and the rug plots are removed to make the display less busy.
+Changed (page 52). Axis tick labels made smaller and the rug plots are removed to make the display less busy.
 
@@
 
-CHANGED.
+Changed (page 53).
 - Different box borders (instead of id labels) are added such that (a) and (b) display the household ids in the same order as indicated by the same colour of the line plot in (b) and box border    
@@
 
-Rephrased to make it clear.
+Rephrased to make it clear. (page 61)
 
@@
 
-Added the statement in page 60.
+Added the statement in the first paragraph of page 62.
 
@@
 
-Rephrased the text in the  Data pre-processing section to make it easier to understand.
-
-Figure 4.1 labels changed to make it consistent with the text.
+Rephrased the text in the section 4.2.3 (Data pre-processing) to make it easier to understand. Figure 4.1 labels changed to make it consistent with the text.
 
@@
 Is NQT also applied to each observation separately?
-How does “it could be useful to standardize it for
-the selected set of significant granularities prior to computing the distances” relate to the following bullet points?
+How does “it could be useful to standardize it for the selected set of significant granularities prior to computing the distances” relate to the following bullet points?
 
-Clarification: Each observation (time series) is scaled independently in both JS-NQT and JS-RS.
+Text added under Section 4.2.2 on pages 64 and 65 to clarify this.
+
+<!-- Clarification: Each observation (time series) is scaled independently in both JS-NQT and JS-RS.
 For the WPD approach, each observation might be scaled or not scaled, resulting in usable groups in either case.
+-->
 
@@
 
-No change. I had mentioned that $D$ stands for Jensen-Shannon distances.
+No change. I had mentioned that $D$ stands for Jensen-Shannon distances (now page 66).
 
@@
 
-No change. Stating the distribution of each $v_t$ is difficult because the levels for each granularity and design varies, therefore I chose to divide it into two tables.
+The table format and caption changed to address this comment. (page 70)
+
+<!-- No change. Stating the distribution of each $v_t$ is difficult because the levels for each granularity and design varies, therefore I chose to divide it into two tables. -->
 
 > Q10. Where is $\mu$ in Table 4.2?
+
+The changed caption (Table 4.2 page 70) will help clarify this.
 
@@
 
-Clarification: Various range of values of $R$ and $T$ are considered as stated in Table 4.1.
+Added in the first line of Section 4.3.3 (page 72.)
 
@@
 
-CLARIFICATION: MDS summary plots of clusters obtained from employing JS-NQT on the validation data sets
-Added in the caption.
+Added in the caption (page 75) to address this comment.
+
+<!-- Clarification: MDS summary plots of clusters obtained from employing JS-NQT on the validation data sets. -->
 
@@
 
-Done. Added a, b representing 12 customers (rows) in each of them.
+Added a, b representing 12 customers (rows) in each of them (page 78).
 
@@
 
-Caption and text changed.  
-The caption and description of the figure in the text have been changed to indicate that Fig.
+Caption (Figure 4.9 page 80) and description in the text (page 78) changed to address this comment.
+
+<!-- Clarification: Fig.
 4.9 depicts a visual summary of why three clusters are insufficient for this data set, however it is recommended by sindex summary plots.
@@
 In this example, increasing the number of clusters actually unfolds some behavior which was not apparent with $k=3$ as suggested by sindex.
+-->
 
@@
 
-Changed. "we have assumed the time series to be stationary" has been changed to "the conditional distributions are assumed to remain constant for the observation period" in the Discussion section.
+Text changed in the Discussion section (under Section 4.5.2) to address this.
+
+<!-- Clarification: "we have assumed the time series to be stationary" has been changed to "the conditional distributions are assumed to remain constant for the observation period". -->
 
@@
 
-A similar question has been answered for the next reviewer.
+A paragraph is added to the Discussion section to incorporate comments on the computational burden. (second paragraph of section 4.5.2 on pages 83, 84)
 
@@
 
-Added in the Introduction.
+Added in the Introduction (page 2).
 
@@
 //onlinelibrary.wiley.com/doi/abs/10.1002/env.2152) should be mentioned somewhere as a consideration for future work but that may become a future new thesis by itself.
+
+<!-- Added in the conclusion. To do. -->
 
@@
 
-Added under Thesis structure in Introduction.
+Added under Thesis structure in Introduction on page 4.
 gravitas: GRAnularity VIsualization for Time-series AnalySis
@@
 
-- Introduction (last but one paragraph) to incorporate comments on existing distance measures and their limitations
+- Introduction (last but one paragraph on page 39) to incorporate comments on existing distance measures and their limitations
 
-- "Distance between distributions" section to mention why we chose we chose Jensen-Shannon distances and some details about other distance measures.
+- "Distance between distributions" (under section 3.2.3 on page 44) to mention why we chose we chose Jensen-Shannon distances and some details about other distance measures.
 
@@
 
-Added the abbreviations on Page 50 in "Choosing cyclic granularities of interest and removing clashes" section.
+Added the abbreviations on page 54 in "Choosing cyclic granularities of interest and removing clashes" under section 3.4.
 
@@
 
-Caption changed to illustrate the plot better.
+Caption changed on page 53 to illustrate the plot better.
 Added box borders in (a) such that (a) and (b) display the household ids in the same order as indicated by the same color of the line plot in (b) and box border in (a).
@@
 
-Changed the text description of Figure 3.7.
+Changed the text description of Figure 3.7 on page 55.
 
@@
 
-Changed to "insignificant"
+Changed to "insignificant".
 
@@
 
-Caption rephrased to make it more clear.
+Caption rephrased to make it more clear on page 56.
 
@@
 
-- To reflect various energy behavior, data for eight homes for Section 3.4 is acquired from friends and colleagues with different professions, lifestyles, and household sizes.
-This level of data is private, and we thank our friends and colleagues for agreeing to have their data used for this analysis.
-This is added in the first paragraph of Application.
+- Text added in the first paragraph of Application to elaborate how households were selected (page 51).
 
-- A paragraph is added to the Discussion section to incorporate comments on the computational time.
+- A paragraph is added to the Discussion section to incorporate comments on computational time (page 57).
 
@@
 
-Rephrased to make it clear that most techniques mentioned in Tureczek and Nielsen (2017) do not treat smart meter data as a data type with temporal component.
+Rephrased Section 4.1 on page 62 to make it clear that most techniques mentioned in Tureczek and Nielsen (2017) do not treat smart meter data as a data type with temporal component.
 
@@
 
-The Discussion section is split into (1) conclusions, and (2) limitations and future research (including a discussion on computational complexity) to address this comment.
+The Discussion section is split into (1) conclusions, and (2) limitations and future research (including a discussion on computational complexity) on pages 82-84 to address this comment.
 

### 2022-02-27 Rmd/03-hakear.Rmd
final pages added on response document

@@
 The simulation results (in Section \ref{sec:simulations}) show that the distribution of $\wpdsub{glm}$ under the null design is similar for high $\nx$ and $\nf$ (levels higher than $5$) but less so for lower values of $\nx$ and $\nf$.
-Hence, a combination approach is proposed where we use a permutation approach for categories with small numbers of levels, and a modeling approach for categories with higher numbers of levels.
+An empirical explanation for this is that with smaller $n_x$ and $n_f$, the variance in the values of $wpd_{raw}$ is greater [Figure 13 of supplements](https://github.com/Sayani07/paper-hakear/blob/master/paper/paper-supplementary.pdf).
+As a result, our modeling approach that fits $wpd_{raw}$ to the number of comparisons ($n_x\times n_f$) performs poorly.
+In order to address this, a combination approach is proposed where we use a permutation approach for categories with small numbers of levels, and a modeling approach for categories with higher numbers of levels.
 This ensures that the computational load of the permutation approach is alleviated while maintaining a similar null distribution across different categories.

### 2022-02-27 Rmd/04-gracsr.Rmd
final pages added on response document

@@
 Statistical transformations are employed to bring all of them to the same range or normalize each observation.
-This is important because we are not interested in trivial clusters that vary in magnitudes, but rather in uncovering comparable patterns of distributional differences between categories.--> For the JS-based approaches, two data transformation techniques are recommended, normal-quantile transform (NQT) and robust scaling (RS).<!-- NQT is used in @Gupta2021-hakear prior sorting granularities.-- $\wpd$, which is the foundation of wpd-based distances.
-XX wpd was not yet introduced--> While @Gupta2021-hakear already use NQT when computing $\wpd$, it could be useful to standardize it for the selected set of significant granularities prior to computing the distances.
+This is important because we are not interested in trivial clusters that vary in magnitudes, but rather in uncovering comparable patterns of distributional differences between categories.--> For the JS-based approaches, two data transformation techniques are recommended, normal-quantile transform (NQT) and robust scaling (RS) referred to as JS-NQT and JS-RS respectively.
+The approach based on $\wpd$ is referred to as WPD.
+While @Gupta2021-hakear already use NQT when computing $\wpd$, it could be useful to standardize it for the selected set of significant granularities prior to computing the distances.
+In both the JS-NQT and JS-RS approaches, each observation is scaled individually, but in the WPD approach, each observation may or may not be scaled, resulting in usable groups in either scenario.
 
@@
 
-All of the approaches were fitted to each data design and to each combination of the considered parameters.
+All of the approaches were fitted to each data design and to each combination of the considered parameters considered in Table \ref{tab:range-parameter}.
 The formed clusters have to match the design, be well separated, and have minimal intra-cluster variation.
 <!--It is possible to study these desired clustering traits visually in a more comprehensive way than just looking at index values.--> MDS and parallel coordinate graphs are used to demonstrate the findings, as well as an index value plot to provide direction on the number of clusters.
-JS-based approaches corresponding to NQT and RS are referred to as JS-NQT and JS-RS respectively.
-In the following plots, results for JS-NQT are reported, and results with JS-RS or wpd-based distances are in the supplementary material.
+<!--JS-based approaches corresponding to NQT and RS are referred to as JS-NQT and JS-RS respectively.--> In the following plots, results for JS-NQT are reported, and results with JS-RS or wpd-based distances are in the supplementary material.
 

### 2022-02-27 Rmd/05-conclusion.Rmd
Q2 Ch1 Juergen

@@
 
+Another long-term goal would be to develop a comparable framework with a spatial component for analysis.
+Electricity data is commonly found in a spatio-temporal format with some type of area code, such as a postcode or administrative unit.
+Assuming that households in certain residential neighborhoods have similar energy usage patterns, future research could look into clustering and visualization methods for analyzing probability distributions of spatio-temporal data, perhaps through some variation of the glyph map.
+
 <!-- There are a few areas to extend this research.

### 2022-02-28 report/response.Rmd
Q2 Ch1 Juergen
fine tuning all comments and cross checking with thesis all done

@@
 
-I thank my examiners for their thorough and constructive comments. The point by point description of changes are below: the examiners' comments are in red and my response is in black.
+I thank my examiners for such thorough and constructive comments. The point by point description of changes are below: the examiners' comments are in red and my response is in black.
 
@@
 
-No change.
-It appears in the abstract of the paper.
+Reference added as link on page 35. The package reference is also mentioned in the Supplementary section of the paper.
+
+<!-- No change. -->
+<!-- It appears in the abstract of the paper.
 For the thesis, I have added it since the abstract acts as a summary to the chapter.
@@
 Moreover, the plan is to bring all the functions from `gravitas` to `hakear` when we publish it.
+-->
 
@@
 
-Added "for a continuous univariate dependent variable" in the last paragraph of Introduction of Chapter 3 on page 39.
+Added "for a continuous univariate dependent variable" in the last paragraph of section 3.1 (Introduction) on page 39.
 
@@
 
-Done.
-Changed to theme_bw() instead of default and outlier.alpha (0.5 ->0.2) and geom_jitter(alpha = 0.04 -> 0.03). Also jitter-width decreased to 0.2.
+The following are changed in Figure 3.1 to address the comment on page 36:
+
+- changed to theme_bw() instead of default
+- outliers and jitters are made more transparent
+- jitter width reduced to make it narrower
+
+<!-- outlier.alpha (0.5 ->0.2) and geom_jitter(alpha = 0.04 -> 0.03). Also jitter-width decreased to 0.2. -->
 
@@
 
-Done. x-axis labels changed to month-of-year for Figure 3.1 and 3.2.
+The comment is addressed by changing the x-axis labels to month-of-year for Figures 3.1 (page 36) and 3.2 (page 38).
 
@@
 
-Done.
-On p35, $v$ changed to $p$, where $p$ represents the number of variables.
-This is analogous to many time granularities and their pairwise combinations and $v$ represents the univariate measured variable for which distributions are constructed for different time granularities and their pairs.
+On p38, $v$ changed to $p$, where $p$ represents the number of variables.
+This is analogous to many time granularities and their pairwise combinations and $v$ represents the univariate measured variable for which distributions are constructed for different time granularities and their pairs (added to Table 3.1 on page 42).
 
@@
 
-Done.
-Changed to "distribution of energy usage"
+Caption changed on page 38 to address this.
+
+<!-- clarification: If the difference between $90^{th}$ and $75^{th}$ percentile of energy distribution is low implies that energy usage is extremely -->
 
@@
 
-Done.
+Done on page 25.
 
@@
 
-Added $m$ to table 3.1  
+Added $m$ to table 3.1 (page 42).
 $m$: number of cyclic granularities to display together
@@
 
-> Q12. In the first sentence on the top of page 42, ordered and unordered are mixed up. Have you considered the setting where the facet variable levels are ordered?
+> Q12. In the first sentence on the top of page 42, ordered and unordered are mixed up.
 
 Done.
+
+> Have you considered the setting where the facet variable levels are ordered?
+
+This is a good suggestion. We will include that into future work.
 
@@
 
-Caption changed to address this comment.
+Caption changed on page 44 to address this comment.
 
@@
 
-Added text in the section 3.2.3 under subsection "Characterizing distributions" on page 43 to address this comment.
+Added text in the section 3.2.3 under subsection "Characterizing distributions" on page 43 to clarify this.
 
@@
 
-> Q18. In Table 3.2, I would suggest separating the m = 1 and m = 2 tables.
+> Q18. In Table 3.2, I would suggest separating the $m = 1$ and $m = 2$ tables.
 
@@
 
-Rephrased the simulation design and results corresponding to $m=2$ on page 50.
+Presentation of the simulation design and results corresponding to $m=2$ changed on page 50.
 Similar design and results for $m=1$, although important, are not included in the main chapter but in the supplements.
@@
 
-Changed (page 53).
+The following are changed in Figure 3.7 (a) (page 53).
 - Different box borders (instead of id labels) are added such that (a) and (b) display the household ids in the same order as indicated by the same colour of the line plot in (b) and box border
-- threshold explained in caption in Table 3.3   
 - grey color explained in the text and caption
+
+The threshold explained in caption in Table 3.3 (page 54).
 
@@
 
-Added in the first line of Section 4.3.3 (page 72.)
+Added in the first line of Section 4.3.3 (page 72).
 
@@
 
-<!-- Added in the conclusion. To do. -->
+Added in section 5.3.2 (Future ideas) on page 90.
 
-The name of the thesis implies that only analyzing temporal data is covered.
+<!-- The name of the thesis implies that only analyzing temporal data is covered.
 Extending it to spatial is too broad a scope.
@@
 This is not the case for extending the spatial component, which, as accurately pointed out, would require time comparable to another thesis.
+-->
 
@@
 
-- Introduction (last but one paragraph on page 39) to incorporate comments on existing distance measures and their limitations
+- Introduction (last but one paragraph on page 39) to incorporate comments on existing distance measures and their limitations.
 
@@
 - f in the equation was a typo. Replaced by log now.
+
+- dx added at the end of the integral.
 
@@
 
-Done. Changed to make it more clear than earlier.
+Changed Figure 3 on page 52 to address comments.
 
@@
 
-Done.
-$\leqslant$ for both.
+Done. $\leqslant$ for both.
 
@@
 Added box borders in (a) such that (a) and (b) display the household ids in the same order as indicated by the same color of the line plot in (b) and box border in (a).
+Added text on page 55 to state where to look at for item 2 and 3 in the text.
 
@@
 
-Clarification 1: Each household is represented by 25 tiles, each tile representing a pair of cyclic granularities.
+Changed the text description of Figure 3.7 on page 55 to address the comments.
+
+Clarification 1: The harmony pairs are selected using the same procedure described in Chapter 2 (as mentioned on page 54).
+The harmonies in Table 3.3 are sorted according to id 1.
+The cyclic granularities in Figure 3.7 (a) is in a lexicographic order.
+The harmony pairs in the parallel coordinate plot (Figure 3.7 (c)) are arranged from highest to lowest values of $wpd$ averaged over all households and hence do not correspond to the first or last rows of Table 3.3.
+
+Clarification 2: Each household is represented by 25 tiles, each tile representing a pair of cyclic granularities.
 The colors (in shades of red) represent the value of $wpd$ for each of the harmony pairs (in Table Table 3.3) and the grey tiles correspond to clashes.
 A darker shade of red corresponds to higher values of $wpd$.
-
-Clarification 2: The harmony pairs in the parallel coordinate plot are arranged from highest to lowest values of $wpd$ averaged over all households and hence do not correspond to the first or last rows of Table 3.3.
-
-Changed the text description of Figure 3.7 on page 55.
 
@@
 
-- Text added in the first paragraph of Application to elaborate how households were selected (page 51).
+- Text added in the first paragraph of Section 3.4 to elaborate how households were selected (page 51).
 
@@
 
-Added links in the supplementary materials section at the end of the chapter.
+Added links to supplementary material in all the relevant places.
 
@@
 
-Done.
+Thanks for the idea. Done.
 
@@
 
-Done. Used "wnwd" through out Chapters 3 AND 4.
+Done. Used "wnwd" through out Chapters 3 and 4.
 
@@
 
-Made all bib entries consistent (sentence case)
+Made all bib entries consistent (sentence case).
 

### 2022-02-28 Rmd/03-hakear.Rmd
fine tuning all comments and cross checking with thesis all done

@@
 ```{r id2-new2, fig.cap = "Distribution of energy consumption displayed through area quantile plots across two cyclic granularities month-of-year and hour-of-day and two households.
-The black line is the median, whereas the orange band covers the 25th to 75th percentile and the green band covers the 10th to 90th percentile.
-Difference between the 90th and 75th quantiles is less for (Jan, Feb) for the first household (a), suggesting that it is a more frequent user of air conditioners than the second household (b).
+The black line is the median, whereas the orange band covers the $25^{th}$ (lower) to $75^{th}$ (upper) percentile and the green band covers the $10^{th}$ (lower) to $90^{th}$ (upper) percentile.
+Difference between the $90^{th}$ and $75^{th}$ percentile is less for (Jan, Feb) for the first household (a), suggesting that it is a more frequent user of air conditioners than the second household (b).
 Distribution of energy usage for (a) changes across both granularities, whereas for (b) daily pattern stays same irrespective of the months."}
@@
 $$
-where $M = \frac{q_1+q_2}{2}$ and $D(q_1||q_2) := \int^\infty_{-\infty} q_1(x)log(\frac{q_1(x)}{q_2(x)})$ is the KL divergence between distributions $q_1$ and $q_2$.
+where $M = \frac{q_1+q_2}{2}$ and $D(q_1||q_2) := \int^\infty_{-\infty} q_1(x)log(\frac{q_1(x)}{q_2(x)})dx$ is the KL divergence between distributions $q_1$ and $q_2$.
 There are other ways to obtain distance between distributions like Hellinger distance, total variation distance and Fisher information metric (all of which are special cases of f-divergence), but Jensen-Shannon distance was chosen for ease of computation.
@@
 An empirical explanation for this is that with smaller $n_x$ and $n_f$, the variance in the values of $wpd_{raw}$ is greater [Figure 13 of supplements](https://github.com/Sayani07/paper-hakear/blob/master/paper/paper-supplementary.pdf).
-As a result, our modeling approach that fits $wpd_{raw}$ to the number of comparisons ($n_x\times n_f$) performs poorly.
+As a result, our modeling approach that fits $wpd_{raw}$ to the number of comparisons performs poorly.
 In order to address this, a combination approach is proposed where we use a permutation approach for categories with small numbers of levels, and a modeling approach for categories with higher numbers of levels.
@@
 We consider coarser linear granularities like hour, day, week and month from the commonly used Gregorian calendar.
-From the four linear granularities of hour, day, week, and month, we obtain $N_C = 4\times 3/2 = 6$ cyclic granularities:  "hour_day", "hour_week", "hour_month", "day_ week", "day_month" and "week_month" abbreviated as *hod*, *how*, *hom*, *dow*, *dom* and *wnwd* respectively.
-Further, we add cyclic granularity day-type ("wknd wday") to capture weekend and weekday behavior.
+From the four linear granularities of hour, day, week, and month, we obtain $N_C = 4\times 3/2 = 6$ cyclic granularities:  "hour_day", "hour_week", "hour_month", "day_ week", "day_month" and "week_month" abbreviated as *hod*, *how*, *hom*, *dow* and *dom*  respectively.
+Further, we add cyclic granularity day-type ("wknd wday") to capture weekend and weekday behavior (abbreviated as *wnwd*).
 Thus, seven cyclic granularities are considered to be of interest.
@@
 
-We can now see some patterns that were not discernible in Panel (b), including:
+We can now see some patterns in \ref{fig:dotplot-8}(a) that were not discernible in \ref{fig:dotplot-8}(b), including:
 
@@
 The computation time of $\wpd$ for a cyclic granularity or harmony is determined by its levels.
-For example, in a harmony with $6 \leq\nx,\nf \leq 50$, the computation time grows with the number of levels and ranges from $30-300$s.
+For example, in a harmony with $6 \leq\nx,\nf \leq 50$, the computation time grows with the number of levels and ranges from $30$ to $300$ seconds.
 $\wpd$ becomes computationally more expensive for cyclic granularities with lower levels $\nx,\nf \leq 5$.
-It is around $40-50$ minutes for $\nx,\nf = \{2,3\}$.
+It is around $40$ to $50$ minutes for $\nx,\nf = \{2,3\}$.
+To perform these computations, R version 4.0.1 (2020-06-06) with the platform: x86 64-apple-darwin17.0 (64-bit) running under: macOS Mojave 10.14.6 and [MonaRCH](https://docs.monarch.erc.monash.edu/) is utilized.
 Even with parallel computation, computing $\wpd$ for many individuals and cyclic granularity/harmonies takes a long time.

### 2022-02-28 Rmd/04-gracsr.Rmd
fine tuning all comments and cross checking with thesis all done

@@
 As a result, in this scenario, we analyze a combination of categories generated from different distributions.
-Display of design and related results can be found in supplementary material.
+Display of design and related results can be found in [supplementary material](https://github.com/Sayani07/paper-gracsR/blob/master/paper-supplementary.pdf).
 
@@
 <!--It is possible to study these desired clustering traits visually in a more comprehensive way than just looking at index values.--> MDS and parallel coordinate graphs are used to demonstrate the findings, as well as an index value plot to provide direction on the number of clusters.
-<!--JS-based approaches corresponding to NQT and RS are referred to as JS-NQT and JS-RS respectively.--> In the following plots, results for JS-NQT are reported, and results with JS-RS or wpd-based distances are in the supplementary material.
+<!--JS-based approaches corresponding to NQT and RS are referred to as JS-NQT and JS-RS respectively.--> In the following plots, results for JS-NQT are reported, and results with JS-RS or wpd-based distances are in the [supplementary material](https://github.com/Sayani07/paper-gracsR/blob/master/paper-supplementary.pdf).
 
@@
 Additionally, there were missing values in many series.
-(The supplementary material contains details from checking for systematic missingness.) Because our proposed methods evaluate probability distributions rather than raw data, these data issues are not problematic, unless there is any systematic structure related to granularities.
+(The [supplementary material](https://github.com/Sayani07/paper-gracsR/blob/master/paper-supplementary.pdf) contains details from checking for systematic missingness.) Because our proposed methods evaluate probability distributions rather than raw data, these data issues are not problematic, unless there is any systematic structure related to granularities.
 
@@
 This ensured that the final sample of clustered customers were also far apart in the high-dimensional space.
-(See the supplementary materials for further details.)<!--Few of these customers have similar distribution across *moy* and some are similar in their *hod* distribution.
+(See the [supplementary materials](https://github.com/Sayani07/paper-gracsR/blob/master/paper-supplementary.pdf) for further details.)<!--Few of these customers have similar distribution across *moy* and some are similar in their *hod* distribution.
 -->

### 2022-02-28 Rmd/05-conclusion.Rmd
fine tuning all comments and cross checking with thesis all done

--- 2021-11-19 Rmd/05-conclusion.Rmd
@@
 
+Another long-term goal would be to develop a comparable framework with a spatial component for analysis.
+Electricity data is commonly found in a spatio-temporal format with some type of area code, such as a postcode or administrative unit.
+Assuming that households in certain residential neighborhoods have similar energy usage patterns, future research could look into clustering and visualization methods for analyzing probability distributions of spatio-temporal data, perhaps through some variation of the glyph maps.
+
 <!-- There are a few areas to extend this research.

### 2022-02-28 scripts/hakear.R

@@ -72,8 +72,7 @@ p1 <- id2_tsibble %>%
   scale_colour_brewer(name = "", palette = "PiYG") + 
   ylab ("energy usage [kWh]") +
   xlab("hour-of-day") +
-  scale_y_log10(breaks = scales::trans_breaks("log10", 
-                                              function(x) 10^x)) +
+  scale_y_log10(minor_breaks = NULL) +
                  #labels = scales::trans_format("log10", scales::math_format(10^.x))) +
   theme_bw() + ylim(0, 1.5)
 
@@ -88,8 +87,8 @@ p2 <- id4_tsibble %>%
   ggtitle("(b)") + theme(
     strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))+
   ylab ("energy usage [kWh]")+
-  scale_y_log10(breaks = scales::trans_breaks("log10", 
-                                              function(x) 10^x))+
+  xlab("hour-of-day")+
+  scale_y_log10(minor_breaks = NULL)+
                 #labels = scales::trans_format("log10", scales::math_format(10^.x))) +
   theme_bw() + ylim(0, 1.5)
 
@@ -1167,7 +1166,7 @@ elec_rank <- elec_sig_split %>%
 
 elec_rank %>% 
   kable(format = "markdown", 
-        caption = "Ranking of harmonies for the eight households with significance or different thresholds. ***, ** and * correspond to the 99th, 95th and 90th percentile threshold. Rankings are different and at most three harmonies are significant for any household. The number of harmonies to explore is reduced from 42 to 3.")
+        caption = "Ranking of harmonies for the eight households with significance or different thresholds. $^{***}$, $^{**}$ and $^{*}$ corresponds to the $99^{th}$, $95^{th}$ and $90^{th}$ percentile threshold. Rankings are different and at most three harmonies are significant for any household. The number of pair of cyclic granularities to explore is reduced from 42 to 3.", escape = FALSE)
 
 
 

### 2022-02-28 report/response.Rmd
minor changes in responses

@@
 On p38, $v$ changed to $p$, where $p$ represents the number of variables.
-This is analogous to many time granularities and their pairwise combinations and $v$ represents the univariate measured variable for which distributions are constructed for different time granularities and their pairs (added to Table 3.1 on page 42).
+
+Clarification: This is analogous to many time granularities and their pairwise combinations and $v$ represents the univariate measured variable for which distributions are constructed for different time granularities and their pairs (added to Table 3.1 on page 42).
 
@@
 The following are changed in Figure 3.7 (a) (page 53).
+
  - Different box borders (instead of id labels) are added such that (a) and (b) display the household ids in the same order as indicated by the same colour of the line plot in (b) and box border

### 2022-02-28 report/response.Rmd
minor typos

@@
 
-Reference is in an acceptable format. No change as already published.
+Reference is in an acceptable format. Not changed as already published.
 
@@
  - outliers and jitters are made more transparent  
- - jitter width reduced to make it narrower
+ - jitter width reduced to make it clear that whiskers do not extend beyond data
 
@@
 
-The following are changed in Figure 3.7 (a) (page 53).
+The following changes have been made to Figure 3.7 (a) (page 53):
 
- - Different box borders (instead of id labels) are added such that (a) and (b) display the household ids in the same order as indicated by the same colour of the line plot in (b) and box border
+ - different box borders (instead of id labels) are added such that (a) and (b) display the household ids in the same order as indicated by the same colour of the line plot in (b) and box border
  - grey color explained in the text and caption
@@
 
-No change. I had mentioned that $D$ stands for Jensen-Shannon distances (now page 66).
+I have mentioned that $D$ stands for Jensen-Shannon distances (now page 66).
 
@@
 
-The changed caption (Table 4.2 page 70) will help clarify this.
+The changed caption (Table 4.2 page 70) will help to clarify this.
 
@@
 
-# Prof Juergen Symanzik
+# Professor Juergen Symanzik
 
@@
 
-- "Distance between distributions" (under section 3.2.3 on page 44) to mention why we chose we chose Jensen-Shannon distances and some details about other distance measures.
+- "Distance between distributions" (under section 3.2.3 on page 44) to mention why we chose Jensen-Shannon distances and some details about other distance measures.
 
@@
 
-We will include this in the supplements while publishing the paper.
+We will include this in the supplements in the future.
 
@@
 
-Done. $\leqslant$ for both.
+Corrected to be $\leqslant$ for both.
 
@@
 
-No change. Showing the vertical axis would not give us any more information than it is already giving.
+Not changed. Showing the vertical axis would not give us any more information than it is already giving.
 

### 2022-02-28 report/response.Rmd
correcting minor inconsistencies - all done

@@
 $m$ refers to the number of cyclic granularities we are considering together in the display.
-For example, contextual cyclic granularities could be $hour_{day}$, $day_{week}$ and $month_{year}$ and we want to visualize any one granularity at a time.
+For example, contextual cyclic granularities could be hour_day, day_week and month_year and we want to visualize any one granularity at a time.
 So $N_c = 3$ and $m=1$.
@@
 
-- Introduction (last but one paragraph on page 39) to incorporate comments on existing distance measures and their limitations.
+- Introduction (last but one paragraph on page 39) to incorporate comments on existing distance measures and their limitations
 
-- "Distance between distributions" (under section 3.2.3 on page 44) to mention why we chose Jensen-Shannon distances and some details about other distance measures.
+- "Distance between distributions" (under section 3.2.3 on page 44) to mention why we chose Jensen-Shannon distances and some details about other distance measures
 
-- f in the equation was a typo. Replaced by log now.
+- f in the equation was a typo. Replaced by log now
 
-- dx added at the end of the integral.
+- dx added at the end of the integral
 
@@
 
-Changed the text description of Figure 3.7 on page 55 to address the comments.
+Changed the text description of Figure 3.7 (page 55) and Table 3.3 (page 54) to address the comments.
 
@@
 
-- Text added in the first paragraph of Section 3.4 to elaborate how households were selected (page 51).
+Text added in the first paragraph of Section 3.4 to elaborate how households were selected (page 51).
 
-- A paragraph is added to the Discussion section to incorporate comments on computational time (page 57).
+A paragraph is added to the Discussion section to incorporate comments on computational time (page 57).
 

### 2022-02-28 Rmd/03-hakear.Rmd
correcting minor inconsistencies - all done

@@
 We consider coarser linear granularities like hour, day, week and month from the commonly used Gregorian calendar.
-From the four linear granularities of hour, day, week, and month, we obtain $N_C = 4\times 3/2 = 6$ cyclic granularities:  "hour_day", "hour_week", "hour_month", "day_ week", "day_month" and "week_month" abbreviated as *hod*, *how*, *hom*, *dow* and *dom*  respectively.
+From the four linear granularities of hour, day, week, and month, we obtain $N_C = 4\times 3/2 = 6$ cyclic granularities:  "hour_day", "hour_week", "hour_month", "day_ week", "day_month" and "week_month" abbreviated as *hod*, *how*, *hom*, *dow*, *dom* and *wom* respectively.
 Further, we add cyclic granularity day-type ("wknd wday") to capture weekend and weekday behavior (abbreviated as *wnwd*).
@@
 The harmony pairs are then arranged in descending order and highlighted with `***`, `**` and `*` corresponding to the $99^{th}$, $95^{th}$ and $90^{th}$ percentile threshold.
-Table \ref{tab:tab-rank-8} shows the rank of the harmonies for different households.
+Table \ref{tab:tab-rank-8} shows the rank of the harmonies for different households and the harmonies are sorted according to id 1.
 The rankings are different for different households, which is a reflection of their varied behaviors.
@@
 
-Figure \ref{fig:dotplot-8} helps to compare households through the heatmap (a) across harmony pairs.
-Each household is represented by 25 tiles, each tile representing a pair of cyclic granularities.
+Figure \ref{fig:dotplot-8} helps to compare households through the heatmap (a) across harmony pairs (lexicographic order).
+Each household is represented by 25 tiles, each tile representing a pair of cyclic granularities with x-variables on the horizontal axis and facet-variables on the vertical axis.
 The colors (in shades of red) represent the value of $\wpd$ for each of the harmony pairs (in Table \ref{tab:tab-rank-8}) and the gray tiles correspond to clashes.

### 2022-02-28 scripts/hakear.R

@@ -21,6 +21,7 @@ library(GGally)
 #library(viridis)
 #remotes::install_github("Sayani07/hakear", force = TRUE)
 library(hakear)
+library(latex2exp)
 
 # OkabeIto color blind friendly scale
 okabeito <- c("#D55E00", "#0072B2","#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442", "#333333")
@@ -488,7 +489,7 @@ G21 %>%
   geom_density(fill = "#999999", color = "#999999") +
   facet_grid(nx~nfacet,
              labeller = "label_both") + 
-  xlab("raw values of wpd") +
+  xlab(TeX("$wpd_{raw}$")) +
   geom_vline(data = summary_data, aes(xintercept  = mean), color = "#0072B2", size = 0.4)  +
   #geom_rug(sides = "b", colour = "#D55E00") + 
   # scale_x_continuous(breaks = scales::breaks_extended(2)) +
@@ -1228,7 +1229,7 @@ p1 <- id1_tsibble %>%
   scale_y_log10() + theme_bw()  +
   theme(legend.position = "none",
         strip.text = element_text(size = 10, margin = margin())) + 
-  theme(panel.spacing =unit(0, "lines"))
+  theme(panel.spacing =unit(0, "lines")) + xlab("hour-of-day")
 
 p2 <- id7_tsibble %>%
   prob_plot("day_week",
@@ -1243,7 +1244,7 @@ p2 <- id7_tsibble %>%
   scale_y_log10() + theme_bw()  +
   theme(legend.position = "none",
         strip.text = element_text(size = 10, margin = margin()))+ 
-  theme(panel.spacing =unit(0, "lines"))
+  theme(panel.spacing =unit(0, "lines"))+ xlab("hour-of-day")
 
 
 
